### 2024-03-05 study_1/preregistration.Rmd
preregistration study 1

---
title: "Pre-registration Study 1 \n Who does not trust basic science?"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
bibliography: ../references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
```

```{r, include=FALSE, message=FALSE}
# read in data from power simulation

```

# Introduction

Anti-vax people or climate skepticists are often conceived of as simply anti-science. But would these people deny that water is made of molecules containing one oxygen and two hydrogen atoms, and that its chemical formula therefore is H2O? We suspect that everyone (or nearly everyone), including people who hold-anti science conspiracy beliefs, trusts basic science, at least when it doesn’t conflict with their specific beliefs.

# Data collection

No data has been collected yet.

# Procedure

After providing their consent to participate in the study, participants are given an attention check:

> *While watching the television, have you ever had a fatal heart attack?* [1-6; 1 = Never, 6 = Often]

Participants then read the following instructions:

> *We will ask you 11 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.*

Next, participants answer a set of 11 basic science questions, all but two extracted from existing science knowledge questionnaires. After each question, participants will be presented with an answer reflecting the scientific consensus. Participants are asked to choose whether they accept the answer or not, before proceeding to the next question. Figure \@ref(fig:stimulus-example) displays the survey for an example science question. Table \@ref(tab:knowledge) shows all included questions, their scientifically consensual answer, and their source.

(ref:stimulus-example) Example of a science question, the scientific consensus and the corresponding acceptance question.

```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-example)"}
knitr::include_graphics("../figures/study1_question_example.png")
```

After that, participants answer questions on conspiracy thinking and trust in science (see next section).

# Materials

## Knowledge Items

```{r knowledge}
items <- read_csv("materials/knowledge_items.csv") %>% 
  mutate(id = 1:nrow(.)) %>% 
  select(id, Question, `Scientific consensus`, `Reference(s)`)

# Output the table
kbl(items, booktabs = T, longtable = TRUE,
    caption = "Science knowledge items") %>%
  kable_paper(full_width = T) %>% 
  column_spec(1, width = "1em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em") %>%
  column_spec(4, width = "15em") 
```

## Conspiracy scales

We rely on three scales:

1.  The conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 :

I think that . . .

-   ... many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
-   ... government agencies closely monitor all citizens.
-   ... events which superficially seem to lack a connection are often the result of secret activities.
-   ... there are secret organizations that greatly influence political decisions.

[0% - 100%; 0 = certainly not, 100 = certain]

2.  The Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 :

-   I think that the official version of the events given by the authorities very often hides the truth. [1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true]

3.  A selection of science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022, displayed in table \@ref(tab:conspiracy).

```{r conspiracy, echo=FALSE}
# Create the data frame
items <- c(
  "The Apollo moon landings never happened and were staged in a Hollywood film studio.",
  "A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).",
  "The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.",
  "The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.",
  "The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.",
  "There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.",
  "In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.",
  "Proof of alien contact is being concealed from the public.",
  "Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.",
  "Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.")

data <- data.frame(id = 1:length(items), items = items)

# Output the table
kbl(data, booktabs = T, longtable = T, col.names = NULL, 
    caption = "Conspiracy items") %>%
  kable_paper(full_width = F,) %>%
  column_spec(1) %>%
  column_spec(2, width = "30em")
```

## Trust in science

We rely on three items. The first two were selected from the Wellcome Global Monitor survey. The third one is from the Pew research center and has recently been used by a world-wide many labs study [@colognaTrustScientistsTheir2024]. We consider "acting in the best interest of the public" as a sub-aspect of trust, and include this question mainly to be able to compare our sample to a wide ranging global sample.

-   How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

-   In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

-   How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]

# Research questions

-   **RQ1: What is the average science knowledge score (1)?**

We will report the average percentage of questions answered correctly. We will look at the pooled average and the distribution of participant averages.

-   **RQ2: What is the average acceptance of the scientific consensus (2)?**

Similar to RQ1, we will report the average acceptance rate of the scientific consensus. We will look at the pooled average and the distribution of participant averages.

-   **RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**

We will regress trust in science on the average knowledge and acceptance per participant.

If they correlate highly, we will use the average of the two Wellcome Global Monitor trust questions. If not, we will focus on the "How much do you trust science?" question.

We will run a robustness check on the Pew question.

-   **RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**

We will proceed just as for RQ3, but with conspiracy thinking as outcome, instead of trust.

To measure conspiracy thinking, we will use the average score of the Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022.

As robustness checks, we will run the same analysis separately for the two other conspiracy scales (also averaging across all items).

We will also check how well the three conspiracy scales correlate.

# Participants

We will recruit 200 participants from the US.

# Exclusions

We will exclude participants who do not answer “Never” in the attention check.

# References

### 2024-03-13 study_1/analysis.Rmd
analysis experiment 1

---
title: "Analysis Study 1"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
bibliography: ../references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library(stargazer)
```

```{r, include=FALSE, message=FALSE}
# read the cleaned long version data set
data_long <- read_csv("data/cleaned_long.csv")

# read wide version data set
data_wide <- read_csv("data/cleaned_wide.csv")

# make a version with averages per participant
d_avg <- data_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0), 
         knowledge_num = ifelse(knowledge == TRUE, 1, 0)
  ) %>% 
  group_by(id) %>% 
  # calculate by-participant averages
  summarize(avg_acceptance  = sum(acceptance_num)/n(), 
            avg_knowledge = sum(knowledge_num)/n()
  )

# make data frame with average particpant data and wide format
d <- left_join(data_wide %>% 
                 select(-c(ends_with("know"), ends_with("accept"))), 
               d_avg, 
               by = "id")

```

# RQ1

**RQ1: What is the average science knowledge score (1)?**

```{r}
# as table
# d_avg %>% 
#   mutate(avg_knowledge = as.factor(avg_knowledge)) %>% 
#   group_by(avg_knowledge) %>% 
#   count() %>% 
#   ungroup() %>% 
#   mutate(share = n / sum(n)) %>% 
#   mutate_if(is.numeric, round, digits = 3)

# as plot
# Calculate the mean
mean_value <- mean(d_avg$avg_knowledge)

ggplot(d_avg, aes(x = avg_knowledge)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value, y = 40, label = paste("Mean =", round(mean_value, 2))), 
            vjust = -0.5, hjust = 1.2) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average science knowledge", 
       y = "N")
```

# RQ2

**RQ2: What is the average acceptance of the scientific consensus (2)?**

```{r}
# Calculate the mean
mean_value <- mean(d_avg$avg_acceptance)

ggplot(d_avg, aes(x = avg_acceptance)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
            vjust = -0.5, hjust = 1.2) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average consensus acceptance", 
       y = "N")
```

# RQ3

**RQ3: What is the relationship between trust in science and, respectively, knowledge and acceptance?**

```{r}
plot_data <- d %>% 
  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
  summarise(n = n())

# Plot as a tile plot
ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_knowledge, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average science knowledge", 
       y = "Trust in science", 
       fill = "N") +  # Update fill legend title
  theme_minimal() 
```

```{r}
plot_data <- d %>% 
  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
  summarise(n = n())

# Plot as a tile plot
ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_acceptance, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average consensus Acceptance", 
       y = "Trust in science", 
       fill = "N") +  # Update fill legend title
  theme_minimal() 
```

# RQ4

**RQ4: What is the relationship between conspiracy thinking and, respectively, knowledge and acceptance?**

```{r}
# Plot as a tile plot
ggplot(d, aes(x = avg_knowledge, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average science knowledge", 
       y = "Average conspiracy thinking") +  # Update fill legend title
  theme_minimal() 
```

```{r}
ggplot(d, aes(x = avg_acceptance, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average consensus acceptance", 
       y = "Average conspiracy thinking") +  # Update fill legend title
  theme_minimal() 
```

## Difference between false and true answer acceptance

```{r}
# make a version with averages per participant, grouped by false/true ansers 
false_data_avg <- data_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    BCTI_avg = mean(BCTI_avg)
  ) %>% 
  mutate(knowledge = ifelse(knowledge == FALSE, "False answers", "Correct answers"))

# make data frame with average particpant data and wide format
# false_data <- left_join(false_data, false_data_avg, by = "id")

ggplot(false_data_avg, aes(x = avg_acceptance, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average consensus acceptance", 
       y = "Average conspiracy thinking") +  # Update fill legend title
  theme_minimal() +
  facet_wrap(~knowledge)

```

Alternative plot

```{r}
plot_data <- data_long %>% 
  mutate(BCTI_avg = round(BCTI_avg, digits = 0)) %>% 
  group_by(knowledge, acceptance, BCTI_avg) %>% 
  count() %>% 
  group_by(knowledge, BCTI_avg) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3) %>% 
  mutate(knowledge = ifelse(knowledge == FALSE, "False answers", "Correct answers"))

ggplot(plot_data, aes(x = BCTI_avg, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(1, 9, 1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Average conspiracy thinking", fill = "Acceptance \nof consensus") +
  theme_minimal() +
  facet_wrap(~knowledge) +
  theme(legend.position = "top")

ggplot(data_long, aes(x = ))
```

# Additional RQs

## Knowledge and Acceptance by science question

```{r}
# calculate share
plot_data <- data_long %>% 
  group_by(subject) %>% 
  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
```

```{r}
# plot knowledge
ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of correct responses") +
  coord_flip()
```

```{r}
# plot acceptance
ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of acceptance") +
  coord_flip()
```

## Open ended comments
```{r}
levels(as.factor(d$comments))
```

## Is the share of correct responses associated with the share of acceptance ?

```{r}
plot_data <- d %>% 
  group_by(avg_knowledge, avg_acceptance) %>% 
  summarise(n = n())

# Plot as a tile plot
ggplot(plot_data, aes(x = avg_knowledge, y = avg_acceptance, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_knowledge, y = avg_acceptance),
              width = 0.03, height = 0.04) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average science knowledge", 
       y = "Average consensus acceptance", 
       fill = "N") +  # Update fill legend title
  theme_minimal() 
```

# What were the acceptance rates among true and false answers?

```{r}
plot_data <- data_long %>% 
  group_by(knowledge, acceptance) %>% 
  count() %>% 
  group_by(knowledge) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
  theme_minimal()
```

## To which extent did people believe the different conspiracy items ?

```{r}
BCTI_data <- data_long %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  select(id, c(starts_with("BCTI"), score)) %>% 
  group_by(BCTI_item) %>% 
  summarize(avg_score = mean(score)) %>% 
  ungroup()

ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
    geom_pointrange(aes(ymin = 1, ymax = avg_score),
                  # Make the lines a little thicker and the dots a little bigger
                  fatten = 2, size = 1) +
  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
  labs(x = "Conspiracy item", y = "Average score") +
  coord_cartesian(ylim = c(1, 9)) +
  coord_flip()
```

### 2024-03-13 study_1/cleaning.Rmd
analysis experiment 1

---
title: "Data cleaning Study 1"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
```

## Import data
```{r import, message=FALSE}
d <- read_csv("./data/qualtrics.csv")
names(d)
```

```{r}
# inspect
head(d) # you can also use View(d)
# delete first two rows
d <- d %>% 
  slice(3: nrow(.)) 
```

## Attention check
```{r attention}
# attention check
table(d$attention_check)
```

There are 6 failed attention checks.

```{r}
# filter to only valid attention check responses
d <- d %>% filter(attention_check == "Never 1")
```

## Re-shape data

```{r}
# check all names and their order
names(d)
```

```{r}
# Function to extract numbers from a string for BCTI and SICBS
extract_numbers <- function(string_var) {
  as.numeric(gsub("\\D", "", string_var))
}

# test the function
# d %>% 
#   mutate(across(starts_with("BCTI"), ~extract_numbers(.), .names = "{.col}_num")) %>% 
#   select(BCTI_autism, BCTI_autism_num)

# clean and re-shape
d_wide <- d %>% 
  # add an easy to read participant identifier
  mutate(id = 1:nrow(.)) %>%
  # make some character variables numeric
  mutate(across(c(starts_with("BCTI"), starts_with("SICBS"), starts_with("wgm"), 
                  starts_with("pew")), ~extract_numbers(.)
  )
  ) %>% 
  mutate(across(starts_with("CMQ"), as.numeric)
  ) %>% 
  # add average conspiracy measure 
  rowwise() %>% 
  mutate(BCTI_avg = mean(c_across(starts_with("BCTI")
                                  ),
                         na.rm = TRUE
                         )
         ) %>% 
  ungroup()

d_long <- d_wide %>% 
  # bring to long format
  pivot_longer(cols = electrons_know:water_accept, 
               names_to = "subject_question", values_to = "correct_accept") %>% 
  # separate conditions into CONVERGENCE_OUTCOME_STIMULUS
  separate_wider_delim(subject_question, "_", names = c("subject", "question")
                       ) %>%
  # necessary because each participant only saw ten out of 11 questions
  drop_na(correct_accept) %>% 
  pivot_wider(names_from = question, values_from = correct_accept) %>% 
# create better variable names
  rename(knowledge = know, 
         acceptance = accept) 
```

## Export data

```{r}
# wide format
write_csv(d_wide, "data/cleaned_wide.csv")

# long format
write_csv(d_long, "data/cleaned_long.csv")
```

### 2024-03-25 appendix_exp1.Rmd
draft after study 1

# Experiment 1 {#exp1}

## Materials

\FloatBarrier

### Conspiracy scales

Beside Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 which report in the main study, we also assessed two other consipiracy thinking measures:

1. The conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013:
I think that . . .

-   ... many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
-   ... government agencies closely monitor all citizens.
-   ... events which superficially seem to lack a connection are often the result of secret activities.
-   ... there are secret organizations that greatly influence political decisions.

[0% - 100%; 0 = certainly not, 100 = certain]

2.  The Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 :

-   I think that the official version of the events given by the authorities very often hides the truth. [1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true]

### Trust in science

We rely on three items

1. How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

2. In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

3. How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]

## Comparing items

Make correlation table

### Conspiracy theories

Table \@ref(tab:correlation-conspiracy) shows the correlations of the three different scales assessing conspiracy thinking.

```{r correlation-conspiracy}
correlation_matrix <- exp1_wide %>%
  select(BCTI_avg, CMQ_avg, SICBS) %>% 
  rename(BCTI = BCTI_avg, 
         CMQ = CMQ_avg) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different scales assessing conspiracy thinking",
          placement = "h")
```

### Trust in science

Table \@ref(tab:correlation-trust) shows the correlations of the three different items measuring trust in science.

```{r correlation-trust}
correlation_matrix <- exp1_wide %>%
  select(wgm_sciencegeneral, wgm_scientists, pew) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different items measuring trust in science",
          placement = "h")
```

## Correlations with alternative measures

Table \@ref(tab:correlations-outcomes) shows the correlations between knowledge and acceptance, respectively, and outcome variables.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp1_wide %>% 
  select(avg_acceptance, avg_knowledge, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# run models to get correlation coefficient

# Define the function to run regression and tidy results
run_regression <- function(data, dependent_var, independent_var) {
  lm(as.formula(paste(dependent_var, "~", independent_var)), data = data) %>%
    tidy() %>%
    filter(term == independent_var) %>%
    mutate(outcome = {{dependent_var}}) %>% 
    select(outcome, term, estimate)
}

# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list_knowledge <- map(variables, ~ run_regression(standardized_variables, .x, "avg_knowledge"))
results_list_acceptance <- map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance"))
```

```{r correlations-outcomes}
# as data frame/output table
results <- bind_rows(list(results_list_knowledge, results_list_acceptance))
results %>% 
  pivot_wider(names_from = term, 
              values_from = estimate) %>% 
  rename(`Correlation with knowledge` = avg_knowledge, 
         `Correlation with acceptance` = avg_acceptance, 
         ) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 
                             outcome == "CMQ_avg" ~ "CMQ", 
                             outcome == "SICBS" ~ "SICBS", 
                             outcome == "wgm_scientists" ~ "WGM trust scientists",
                             outcome == "wgm_sciencegeneral" ~ "WGM trust general \n(main trust measure)",
                             outcome == "pew" ~ "PEW trust scientists",
                             )) %>% 
  apa_table(caption = "Correlations between knowledge and acceptance, respectively, and outcome variables")
```

## Results conditional on false responses

Table \@ref(tab:false-response-regression) shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp1_false_knowledge %>% 
  select(avg_acceptance, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# Define the function to run regression and tidy results
run_regression <- function(data, dependent_var, independent_var) {
  lm(as.formula(paste(dependent_var, "~", independent_var)), data = data) 
}

# run models to get correlation coefficient

# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance")), variables)
```

```{r false-response-regression}
modelsummary(results_list, 
             stars = TRUE,
             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values") 
```

### 2024-03-25 exp_1/analysis.Rmd
draft after study 1

--- 2024-03-13 study_1/analysis.Rmd
@@
 
-# make a version with averages per participant
-d_avg <- data_long %>% 
-  # make numeric versions
-  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0), 
-         knowledge_num = ifelse(knowledge == TRUE, 1, 0)
-  ) %>% 
-  group_by(id) %>% 
-  # calculate by-participant averages
-  summarize(avg_acceptance  = sum(acceptance_num)/n(), 
-            avg_knowledge = sum(knowledge_num)/n()
-  )
-
-# make data frame with average particpant data and wide format
-d <- left_join(data_wide %>% 
-                 select(-c(ends_with("know"), ends_with("accept"))), 
-               d_avg, 
-               by = "id")
-
+# rename d wide
+d <- data_wide
+d_avg <- data_wide
 ```
@@
               width = 0.03, height = 0.08) +  # Scatter points within each tile
+    geom_smooth(method = "lm") +
   scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
@@
   theme_minimal() 
+```
+
+## Difference between false and true answer acceptance
+
+```{r}
+# make a version with averages per participant, grouped by false/true ansers 
+veracity_data_avg <- data_long %>% 
+  # make numeric versions
+  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
+  ) %>% 
+  group_by(id, knowledge) %>% 
+  # calculate by-participant averages
+  summarize(
+    n = n(),
+    n_accepted = sum(acceptance_num),
+    avg_acceptance  = sum(acceptance_num)/n(), 
+    wgm_sciencegeneral = mean(wgm_sciencegeneral)
+  ) %>% 
+  mutate(knowledge = ifelse(knowledge == FALSE, "False answers", "Correct answers"))
+
+# make data frame with average particpant data and wide format
+# false_data <- left_join(false_data, false_data_avg, by = "id")
+
+ggplot(veracity_data_avg, aes(x = avg_acceptance, y = wgm_sciencegeneral)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  labs(x = "Average consensus acceptance", 
+       y = "Average trust in science") +  # Update fill legend title
+  theme_minimal() +
+  facet_wrap(~knowledge)
+
 ```
@@
 
-# What were the acceptance rates among true and false answers?
+## What were the acceptance rates among true and false answers?
 
@@
 ```
+
+# Redo all analyses without tree question

### 2024-03-25 exp_1/cleaning.Rmd
draft after study 1

--- 2024-03-13 study_1/cleaning.Rmd
@@
 
-## Re-shape data
+## Add demographics
+
+## Recode demographics
+
+```{r, message=FALSE}
+prolific_demographics <- read_csv("./data/prolific_demographics.csv")
+
+d <- left_join(d, prolific_demographics, by = c("PROLIFIC_PID" = "Participant id"))
+```
+
+```{r}
+d <- d %>% 
+  mutate(gender = case_when(Sex == "Male" ~ "male", 
+                            Sex == "Female" ~  "female", 
+                            .default = NA)
+         ) %>% 
+  rename(age = Age)
+```
+
+## Clean and re-shape data
 
@@
 ```
+
+Clean wide format data.
 
@@
   ) %>% 
-  # add average conspiracy measure 
+  # add average conspiracy measures 
   rowwise() %>% 
   mutate(BCTI_avg = mean(c_across(starts_with("BCTI")
+                                  ),
+                         na.rm = TRUE
+                         ),
+         CMQ_avg = mean(c_across(starts_with("CMQ")
                                   ),
@@
   ungroup() 
+```
 
+Make long format data.
+
+```{r}
 d_long <- d_wide %>% 
@@
 
+Add knowledge and acceptance by-participant averages to wide format data.
+
+```{r}
+# make a version with averages per participant
+d_avg <- d_long %>% 
+  # make numeric versions
+  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0), 
+         knowledge_num = ifelse(knowledge == TRUE, 1, 0)
+  ) %>% 
+  group_by(id) %>% 
+  # calculate by-participant averages
+  summarize(avg_acceptance  = sum(acceptance_num)/n(), 
+            avg_knowledge = sum(knowledge_num)/n()
+  )
+
+# make data frame with average particpant data and wide format
+d_wide <- left_join(d_wide %>% 
+                 select(-c(ends_with("know"), ends_with("accept"))), 
+               d_avg, 
+               by = "id")
+```
+
 ## Export data

### 2024-03-25 paper.Rmd
draft after study 1

---
title             : "Who mistrust basic science?"
shorttitle        : "Who mistrust basic science?"

header-includes:  | # to prevent floats from moving past certain points (for the appendix)
  \usepackage{placeins}

author: 
  - name          : "Jan Pfänder"
    affiliation   : "1"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Data collection"
      - "Statistical Analyses"
      - "Writing - Original Draft"
      - "Writing - Review & Editing"
  - name          : "Lou Kerzreho"
    affiliation   : "1"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Models and Simulations"
      - "Writing - Review & Editing"
  - name          : "Hugo Mercier"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "hugo.mercier@gmail.com"
    address: " "
    role:
      - "Conceptualization"
      - "Writing - Original Draft"
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Institut Jean Nicod, Département d’études cognitives, ENS, EHESS, PSL University, CNRS, France"
  - id            : "2"
    institution   : "Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany"

authornote: |
  HM received funding from the ANR (SCALUP, ANR-21-CE28-0016-01), and from the John Tempelton Foundation ("An Evolutionary and Cultural Perspective on Intellectual Humility via Intellectual Curiosity and Epistemic Deference"). The ANR also supported this project through FrontCog (ANR-17-EURE-0017), and PSL (ANR-10-IDEX-0001-02).

abstract: |
  XX

keywords          :

floatsintext      : yes
linenumbers       : no 
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "doc" # "doc" for nice look, "man" for manuscripty
output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors. For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g. set always allow html as true in the yaml heading

always_allow_html: true

appendix:
  - "appendix_exp1.Rmd"

bibliography: references.bib
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  # fig.retina = 3, dpi = conditional_dpi,
  # fig.width = 7, fig.asp = 0.618,
  # fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  },
  options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)
```

```{r packages, include=FALSE}
# load required packages
library("papaja")      # For APA style manuscript   
library("lme4")        # model specification / estimation
library("lmerTest")    # provides p-values in the output
library("tidyverse")   # data wrangling and visualisation
library("afex")        # anova and deriving p-values from lmer
library("broom")       # extracting data from model fits 
library("broom.mixed") # extracting data from mixed models
library("metafor")     # doing mata analysis
library("patchwork")   # put several plots together
library("ggridges")    # for plots
library("gghalves")    # for plots
library("ggbeeswarm")  # Special distribution-shaped point jittering
library("knitr")       # for tables
library("kableExtra")  # also for tables
library("ggpubr")      # for combining plots with ggarrange() 
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library("modelsummary") # for regression tables
```

```{r functions}
# load plot theme
source("functions/plot_theme.R")

# load other functions
source("functions/own_functions.R")
```

# Introduction

There is increasingly more talk of defiance towards science and towards experts more generally (for science, see the intro to the stability paper, see also Paresman et al on populism?). In every population, many people only trust science ‘some of the time’, and a sizable minority doesn’t trust it.

This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.

But how deep is distrust of science? Do people reject science wholesale, or is it more driven by attitudes towards specific contents, maybe in particular those seen as being politicized, or as going against deeply held beliefs, or behaviors they want to engage in? [previous work suggests that issue-specific claims provoke more negative attitudes]

Here, we look at whether people, across different levels of trust in science (and of other relevant attitudes? Conspi? Populists? Apparently on Prolific we can specifically target antivax people for instance, which could be quite useful), accept basic scientific facts.

[We can draw some inspiration from the “Westwood et al 2021 Current research overstates American support for political violence” paper: we want to put in perspective the current rise (? discourse around?) in mistrust towards science]

## Levels of trust in science

## Why do people distrust science?

[here? We know from the many studies on the gateway belief model that people don’t always accept the scientific consensus (indeed, the effect sizes are usually not massive)]

## How much science do people know?

This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven… so the first thing to do would be to look at data on knowledge of science, and see what the highest rates of correct answer is, to see how much room for improvement there is.

“The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D. Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989). Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007). The results of the U.S. portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch. Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators’ questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).”

# First series of experiments

The goal is to see how many people really don’t accept basic science (in practice, get as close to ceiling in acceptance as we can, first with a generic sample, then see if that still works with an anti-vax sample).

# Experiment 1

```{r exp1}
# Analyze data of experiments and store results

# Experiment 1

# read the cleaned long version data set
exp1_long <- read_csv("exp_1/data/cleaned_long.csv")

# read wide version data set
exp1_wide <- read_csv("exp_1/data/cleaned_wide.csv")

# Research questions

# RQ3
cor_trust_knowledge <- cor(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_knowledge) %>% round(digits = 2)

cor_trust_acceptance <- cor(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_acceptance) %>% round(digits = 2)

# RQ4
cor_conspiracy_knowledge <- cor(exp1_wide$BCTI_avg, exp1_wide$avg_knowledge) %>% round(digits = 2)

cor_conspiracy_acceptance <- cor(exp1_wide$BCTI_avg, exp1_wide$avg_acceptance) %>% round(digits = 2)

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp1_false_knowledge <- exp1_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

false_answers_cor_trust_acceptance <- cor(exp1_false_knowledge$wgm_sciencegeneral, exp1_false_knowledge$avg_acceptance) %>% round(digits = 2)

false_answers_cor_conspiracy_acceptance <- cor(exp1_false_knowledge$BCTI_avg, exp1_false_knowledge$avg_acceptance) %>% round(digits = 2)

# extract descriptives for inline reporting
exp1_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp1_wide$id),
  gender = exp1_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp1_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp1_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  # RQ3
  cor_trust_knowledge = cor_trust_knowledge,
  cor_trust_acceptance = cor_trust_acceptance,
  # RQ4
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp1_long %>% 
    group_by(knowledge, acceptance) %>% 
    count() %>% 
    group_by(knowledge) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge, acceptance),
  # acceptance and trust in science/conspiracy thinking
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance
)
```

The main goal of experiment one was to test whether people would accept the scientific consensus on basic knowledge questions. Additionally, we wanted to know if both science knowledge and acceptance of the scientific consensus are associated with trust in science and conspiracy thinking. We had the following research questions:

-   **RQ1: What is the average science knowledge score (1)?**

-   **RQ2: What is the average acceptance of the scientific consensus (2)?**

-   **RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**

-   **RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**

## Methods

### Participants

We recruited 200 participants from the US via prolific. 6 participants failed our attention check, resulting in a final sample of `r exp1_descriptives$n_subj` participants (`r exp1_descriptives$gender$female$n` female, `r exp1_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp1_descriptives$age$mean`, $age_\text{sd}$: `r exp1_descriptives$age$sd`, $age_\text{median}$: `r exp1_descriptives$age$median`). Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.

### Procedure

After providing their consent to participate in the study, participants were given an attention check "While watching the television, have you ever had a fatal heart attack?" [1-6; 1 = Never, 6 = Often]. All participants who did not answer "1 = Never" were excluded. Participants then read the following instructions:"We will ask you 11 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions, which were randomly selected from a pool of 11 questions, in random order. After each question, participants were presented with an answer reflecting the scientific consensus. Participants were asked to choose whether they accept the answer or not, before proceeding to the next question. Figure \@ref(fig:stimulus-example) displays the survey for an example science question. Finally, participants answered questions on conspiracy thinking and trust in science.

(ref:stimulus-example) Example of a science question, the scientific consensus and the corresponding acceptance question.

```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-example)"}
knitr::include_graphics("./figures/study1_question_example.png")
```

### Materials

#### Science knowledge and acceptance

Table \@ref(tab:knowledge) shows all questions, their scientifically consensual answer, and their source. All but two questions were selected from existing science knowledge questionnaires. We tried to select non-political questions.

```{r knowledge}
items <- read_csv("exp_1/materials/knowledge_items.csv") %>% 
  mutate(id = 1:nrow(.)) %>% 
  select(id, Question, `Scientific consensus`, `Reference(s)`)

# Output the table
kbl(items, booktabs = T, longtable = TRUE,
    caption = "Science knowledge items", 
    full_width = T) %>%
  column_spec(1, width = "1em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em") %>%
  column_spec(4, width = "15em") 
```

#### Conspiracy scales

To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 (Table \@ref(tab:conspiracy)).

```{r conspiracy, echo=FALSE}
# Create the data frame
items <- c(
  "The Apollo moon landings never happened and were staged in a Hollywood film studio.",
  "A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).",
  "The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.",
  "The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.",
  "The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.",
  "There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.",
  "In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.",
  "Proof of alien contact is being concealed from the public.",
  "Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.",
  "Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.")

data <- data.frame(id = 1:length(items), items = items)

# Output the table
kbl(data, booktabs = T, longtable = T, col.names = NULL, 
    caption = "Conspiracy items", 
    full_width = F) %>%
  column_spec(1) %>%
  column_spec(2, width = "30em")
```

To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 and the Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 (see Appendix \@ref:exp1).

#### Trust in science

Our main item for measuring trust in science is selected from the Wellcome Global Monitor survey: "In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"

We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey ("How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), the other from the Pew research center ("How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]"). We selected these items so that we could compare the ratings in our sample to global survey results. The WGM survey has been administered in over 140 countries and included over 140000 respondents. The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents [@colognaTrustScientistsTheir2024].

## Results

Regarding RQ1 and RQ2, participants answered on average `r exp1_descriptives$means$knowledge_mean` (sd = `r exp1_descriptives$means$knowledge_sd`) of the questions correctly, and accepted the scientific consensus on average for `r exp1_descriptives$means$acceptance_mean` (sd = `r exp1_descriptives$means$acceptance_sd`) of the questions.

Fig. \@ref(fig:exp1-conditional-acceptance) illustrates the relationship between knowledge and acceptance. In most cases (`r exp1_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having given the wrong answer to a question. In very few cases (`r exp1_descriptives$conditional_acceptance$true$No$share`), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response. We believe this might have been due to inattention.

(ref:exp1-conditional-acceptance) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus  **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus

```{r exp1-conditional-acceptance, fig.cap="(ref:exp1-conditional-acceptance)"}
plot_data <- exp1_long %>% 
  group_by(knowledge, acceptance) %>% 
  count() %>% 
  group_by(knowledge) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

absolute_numbers <- plot_data %>% 
  group_by(knowledge) %>% 
  summarise(n = sum(n))

ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
  theme_minimal() +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
                   xmax = as.numeric(factor(knowledge)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n)),
               inherit.aes = FALSE,
               step.increase = 0)
```

For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = `r exp1_descriptives$cor_trust_knowledge`), and acceptance of scientific consensus and trust in science ( r = `r exp1_descriptives$cor_trust_acceptance`). The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science. These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig. \@ref(fig:exp1-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.

For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = `r exp1_descriptives$cor_conspiracy_knowledge`), and conspiracy thinking and acceptance of scientific consensus ( r = `r exp1_descriptives$cor_conspiracy_acceptance`).

In appendix XX we show that these results hold for our alternative measures of trust and conspiracy thinking. We also include more descriptive statistics, such as knowledge and acceptance by science questions.

Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus? In our main analyses, we looked at correlations of acceptance across all observations. One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus. To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question. We than calculated the correlation between trust in science and average acceptance rate by participant. We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance`), and only a very small correlation for trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance`). These associations are not statistically significant (see Appendix \@ref:exp1).

The correlations of acceptance with

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

## Make a combined plot with all results

(ref:exp1-plot) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus  **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus

```{r exp1-plot, fig.cap="(ref:exp1-plot)", fig.height= 10, fig.width=10}
# Plot distributions of knowledge and acceptance and their mean

mean_value_knowledge <- mean(exp1_wide$avg_knowledge)
max_n <- max(table(exp1_wide$avg_acceptance))

knowledge_mean <- ggplot(exp1_wide, aes(x = avg_knowledge)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, max_n, 10)) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, max_n)) +
  labs(x = "Average science knowledge", 
       y = "N")

mean_value <- mean(exp1_wide$avg_acceptance)

acceptance_mean <- ggplot(exp1_wide, aes(x = avg_acceptance)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, max_n, 10)) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, max_n)) +
  labs(x = "Average consensus acceptance", 
       y = "N")

# Trust in science plots
plot_data <- exp1_wide %>% 
  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
  summarise(n = n())

trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(xlim = c(0, 1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average science knowledge", 
       y = "Trust in science", 
       fill = "N")

plot_data <- exp1_wide %>% 
  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
  summarise(n = n())

trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  
  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) + 
  scale_fill_viridis_c(option = "plasma") + 
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(xlim = c(0, 1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average consensus Acceptance", 
       y = "Trust in science", 
       fill = "N") +
  theme(legend.position = "top")

# Conspiracy Thinking plots

conspiracy_knowledge <- ggplot(exp1_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:9) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, 9)) +
  labs(x = "Average science knowledge", 
       y = "Average conspiracy thinking")

conspiracy_acceptance <- ggplot(exp1_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:9) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, 9)) +
  labs(x = "Average consensus acceptance", 
       y = "Average conspiracy thinking") +
  theme(legend.position = "top")

# Combine plots

# distributions
distributions <- ggarrange(knowledge_mean + 
                             rremove("xlab") , 
                           acceptance_mean + 
                             rremove("xlab") +
                             rremove("ylab"))

# trust
trust <- ggarrange(trust_knowledge + 
                             rremove("xlab") +
                             guides(fill = "none") +
                             labs(caption = NULL) , 
                           trust_acceptance + 
                             rremove("xlab") +
                             rremove("ylab"), 
                   common.legend = TRUE)

# conspiracy
conspiracy <- ggarrange(conspiracy_knowledge, 
                           conspiracy_acceptance + 
                             rremove("ylab"), 
                   common.legend = TRUE)

# common plot
patchwork <- (distributions / trust / conspiracy) + 
  plot_annotation(tag_levels = 'A')

patchwork

```

## Discussion

These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus. Yet, in `r exp1_descriptives$conditional_acceptance$false$No$share` of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes. In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.

However, conditioning our analyses to initial false knowledge responses, i.e. looking at change of opinion towards the scientific consensus, we find that trust in science is onl

(Experiment 1 Lou write up)

If we ask basic science questions, and we give feedback:
How many people accept the correct answer at the end?
Is that a function of TIS?
Is that a function of CT?

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

So we look at who changes their minds. We still see an influence of CT (? do stats) (and of TIS?).

Takeaway: 
Relatively high K in this population
After feedback even higher
But not at ceiling
Relation with TIS/CT, both in K, and in integrating feedback (?)

However, for those who don’t accept the feedback, we don’t know if it’s because they don’t trust us, or they don’t trust science.
We can also try slightly more stringent attention checks (the fact that some people go from good to bad suggests some inattention)

\FloatBarrier

# References

::: {#refs}
:::

\newpage

# (APPENDIX) Appendix {-}

```{r child = "appendix_exp1.Rmd"}
```

### 2024-03-25 paper.tex
draft after study 1

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  doc,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\usepackage{csquotes}
\usepackage{placeins}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Who mistrust basic science?},
  pdfauthor={Jan Pfänder1, Lou Kerzreho1, \& Hugo Mercier1},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Who mistrust basic science?}
\author{Jan Pfänder\textsuperscript{1}, Lou Kerzreho\textsuperscript{1}, \& Hugo Mercier\textsuperscript{1}}
\date{}

\shorttitle{Who mistrust basic science?}

\authornote{

HM received funding from the ANR (SCALUP, ANR-21-CE28-0016-01), and from the John Tempelton Foundation (``An Evolutionary and Cultural Perspective on Intellectual Humility via Intellectual Curiosity and Epistemic Deference''). The ANR also supported this project through FrontCog (ANR-17-EURE-0017), and PSL (ANR-10-IDEX-0001-02).

The authors made the following contributions. Jan Pfänder: Conceptualization, Data collection, Statistical Analyses, Writing - Original Draft, Writing - Review \& Editing; Lou Kerzreho: Models and Simulations, Writing - Review \& Editing; Hugo Mercier: Conceptualization, Writing - Original Draft, Writing - Review \& Editing, Supervision.

Correspondence concerning this article should be addressed to Hugo Mercier, . E-mail: \href{mailto:hugo.mercier@gmail.com}{\nolinkurl{hugo.mercier@gmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Institut Jean Nicod, Département d'études cognitives, ENS, EHESS, PSL University, CNRS, France\\\textsuperscript{2} Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany}

\abstract{%
XX
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

There is increasingly more talk of defiance towards science and towards experts more generally (for science, see the intro to the stability paper, see also Paresman et al on populism?). In every population, many people only trust science `some of the time', and a sizable minority doesn't trust it.

This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.

But how deep is distrust of science? Do people reject science wholesale, or is it more driven by attitudes towards specific contents, maybe in particular those seen as being politicized, or as going against deeply held beliefs, or behaviors they want to engage in? {[}previous work suggests that issue-specific claims provoke more negative attitudes{]}

Here, we look at whether people, across different levels of trust in science (and of other relevant attitudes? Conspi? Populists? Apparently on Prolific we can specifically target antivax people for instance, which could be quite useful), accept basic scientific facts.

{[}We can draw some inspiration from the ``Westwood et al 2021 Current research overstates American support for political violence'' paper: we want to put in perspective the current rise (? discourse around?) in mistrust towards science{]}

\hypertarget{levels-of-trust-in-science}{%
\subsection{Levels of trust in science}\label{levels-of-trust-in-science}}

\hypertarget{why-do-people-distrust-science}{%
\subsection{Why do people distrust science?}\label{why-do-people-distrust-science}}

{[}here? We know from the many studies on the gateway belief model that people don't always accept the scientific consensus (indeed, the effect sizes are usually not massive){]}

\hypertarget{how-much-science-do-people-know}{%
\subsection{How much science do people know?}\label{how-much-science-do-people-know}}

This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven\ldots{} so the first thing to do would be to look at data on knowledge of science, and see what the highest rates of correct answer is, to see how much room for improvement there is.

``The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D. Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989). Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007). The results of the U.S. portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch. Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators' questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).''

\hypertarget{first-series-of-experiments}{%
\section{First series of experiments}\label{first-series-of-experiments}}

The goal is to see how many people really don't accept basic science (in practice, get as close to ceiling in acceptance as we can, first with a generic sample, then see if that still works with an anti-vax sample).

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

The main goal of experiment one was to test whether people would accept the scientific consensus on basic knowledge questions. Additionally, we wanted to know if both science knowledge and acceptance of the scientific consensus are associated with trust in science and conspiracy thinking. We had the following research questions:

\begin{itemize}
\item
  \textbf{RQ1: What is the average science knowledge score (1)?}
\item
  \textbf{RQ2: What is the average acceptance of the scientific consensus (2)?}
\item
  \textbf{RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?}
\item
  \textbf{RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?}
\end{itemize}

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

We recruited 200 participants from the US via prolific. 6 participants failed our attention check, resulting in a final sample of 194 participants (98 female, 96 male; \(age_\text{mean}\): 41.71, \(age_\text{sd}\): 13.07, \(age_\text{median}\): 39). Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

After providing their consent to participate in the study, participants were given an attention check ``While watching the television, have you ever had a fatal heart attack?'' {[}1-6; 1 = Never, 6 = Often{]}. All participants who did not answer ``1 = Never'' were excluded. Participants then read the following instructions:``We will ask you 11 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.'' Next, participants answered a set of 10 basic science questions, which were randomly selected from a pool of 11 questions, in random order. After each question, participants were presented with an answer reflecting the scientific consensus. Participants were asked to choose whether they accept the answer or not, before proceeding to the next question. Figure \ref{fig:stimulus-example} displays the survey for an example science question. Finally, participants answered questions on conspiracy thinking and trust in science.

\begin{figure}

\includegraphics[width=1\linewidth]{./figures/study1_question_example} \hfill{}

\caption{Example of a science question, the scientific consensus and the corresponding acceptance question.}\label{fig:stimulus-example}
\end{figure}

\hypertarget{materials}{%
\subsubsection{Materials}\label{materials}}

\hypertarget{science-knowledge-and-acceptance}{%
\paragraph{Science knowledge and acceptance}\label{science-knowledge-and-acceptance}}

Table \ref{tab:knowledge} shows all questions, their scientifically consensual answer, and their source. All but two questions were selected from existing science knowledge questionnaires. We tried to select non-political questions.

\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{1em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{15em}}
\caption{\label{tab:knowledge}Science knowledge items}\\
\toprule
id & Question & Scientific consensus & Reference(s)\\
\midrule
1 & Do antibiotics kill viruses as well as bacteria? & There is a consensus among scientists that antibiotics kill bacteria, but not viruses. & (Sturgis \& Allum, 2004), (Miller, 2004), (Miller, 1998) (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat, 2011), (Pardo \& Calvo, 2004), (Hayes \& Tariq, 2000), (Committee et al., 2016)\\
2 & Are electrons smaller, larger, or the same size as atoms ? & There is a consensus among scientists that electrons are smaller than atoms. & (Sturgis \& Allum, 2004), (Miller, 2004), (Miller, 1998), (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat 2011), (Pardo \& Calvo, 2004), Committee et al., 2016\\
3 & Have the continents on Earth been moving for millions of years or have they always been where they are now? & There is a consensus among scientists that the continents on Earth have been moving for millions of years due to plate tectonics. & (Miller, 2004) and (Miller, 1998), (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989)
(Gauchat, 2011), (Pardo \& Calvo, 2004)\\
4 & What decides whether a baby is a boy or a girl ? Is it the father’s genes, the mother’s genes, or both? & There is a consensus among scientists that it is the genes in the father's sperm which are decisive on whether a baby is a boy or a girl. & (Sturgis \& Allum, 2004), (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat, 2011), (Pardo \& Calvo, 2004), (Committee et al., 2016)\\
5 & Do lasers work by focusing sound waves? & There is a consensus among scientists that lasers do not work by focusing sound waves. & (Sturgis \& Allum, 2004), (Miller, 2004) and (Miller, 1998), (Evans \& Durant, 1995) and (Durant et al., 1989), (Tourangeau et al., 2016), (Gauchat , 2011), (Pardo \& Calvo, 2004), (Committee et al., 2016)\\
\addlinespace
6 & How long does it take for Earth to go around the sun : one day, one month, or one year ? & There is a consensus among scientists that it takes one year for Earth to go around the sun. & (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat, 2011)\\
7 & Are diamonds made of carbon ? & There is a consensus among scientists that diamonds are made of carbon. & (Evans \& Durant, 1995) and (Durant et al., 1989)\\
8 & Which travels faster : light or sound? & There is a consensus among scientists that light travels faster than sound. & (Miller, 2004) and (Miller, 1998), (Evans \& Durant, 1995) and (Durant et al., 1989)\\
9 & Is common table salt made of calcium carbonate? & There is a consensus among scientists that common table salt is not made of calcium carbonate; it is made of sodium chloride. & (Evans \& Durant, 1995) and (Durant et al., 1989)\\
10 & Where do trees mainly draw the materials with which they create their mass? & There is a consensus among scientists that carbon drawn from the air during photosynthesis makes up most of the materials that trees use to build new leaves, stems, and roots. & https://www.canr.msu.edu/news/where\_do\_trees\_get\_their\_mass\_from; 
https://www.tenereteam.com/blogs/where-do-trees-get-their-mass/;
https://askabiologist.asu.edu/recipe-plant-growth\\
\addlinespace
11 & Is water made of molecules containing one oxygen and two hydrogen atoms? & There is a consensus among scientists that water is made of molecules containing one oxygen and two hydrogen atoms, and that its chemical formula is therefore H2O. & https://en.wikipedia.org/wiki/Water\\
\bottomrule
\end{longtable}

\hypertarget{conspiracy-scales}{%
\paragraph{Conspiracy scales}\label{conspiracy-scales}}

To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by Pennycook, Binnendyk, and Rand (2022) (Table \ref{tab:conspiracy}).

\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
\caption{\label{tab:conspiracy}Conspiracy items}\\
\toprule
1 & The Apollo moon landings never happened and were staged in a Hollywood film studio.\\
2 & A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).\\
3 & The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.\\
4 & The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.\\
5 & The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.\\
\addlinespace
6 & There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.\\
7 & In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.\\
8 & Proof of alien contact is being concealed from the public.\\
9 & Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.\\
10 & Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.\\
\bottomrule
\end{longtable}

To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix @ref:exp1).

\hypertarget{trust-in-science}{%
\paragraph{Trust in science}\label{trust-in-science}}

Our main item for measuring trust in science is selected from the Wellcome Global Monitor survey: ``In general, would you say that you trust science a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''

We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey (``How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''), the other from the Pew research center (``How much confidence do you have in scientists to act in the best interests of the public? {[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}''). We selected these items so that we could compare the ratings in our sample to global survey results. The WGM survey has been administered in over 140 countries and included over 140000 respondents. The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents (Cologna et al., 2024).

\hypertarget{results}{%
\subsection{Results}\label{results}}

Regarding RQ1 and RQ2, participants answered on average 74 \% (sd = 0.16) of the questions correctly, and accepted the scientific consensus on average for 93 \% (sd = 0.12) of the questions.

Fig. \ref{fig:exp1-conditional-acceptance} illustrates the relationship between knowledge and acceptance. In most cases (76.3 \%), participants readily accepted the scientific consensus after having given the wrong answer to a question. In very few cases (1.6 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response. We believe this might have been due to inattention.

\begin{figure}
\centering
\includegraphics{output/figures/exp1-conditional-acceptance.pdf}
\caption{\label{fig:exp1-conditional-acceptance}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
\end{figure}

For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29), and acceptance of scientific consensus and trust in science ( r = 0.27). The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science. These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig. \ref{fig:exp1-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.

For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = -0.39), and conspiracy thinking and acceptance of scientific consensus ( r = -0.33).

In appendix XX we show that these results hold for our alternative measures of trust and conspiracy thinking. We also include more descriptive statistics, such as knowledge and acceptance by science questions.

Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus? In our main analyses, we looked at correlations of acceptance across all observations. One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus. To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question. We than calculated the correlation between trust in science and average acceptance rate by participant. We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = -0.14), and only a very small correlation for trust in science (r = 0.06). These associations are not statistically significant (see Appendix @ref:exp1).

The correlations of acceptance with

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

\hypertarget{make-a-combined-plot-with-all-results}{%
\subsection{Make a combined plot with all results}\label{make-a-combined-plot-with-all-results}}

\begin{figure}
\centering
\includegraphics{output/figures/exp1-plot.pdf}
\caption{\label{fig:exp1-plot}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
\end{figure}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus. Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes. In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.

However, conditioning our analyses to initial false knowledge responses, i.e.~looking at change of opinion towards the scientific consensus, we find that trust in science is onl

(Experiment 1 Lou write up)

If we ask basic science questions, and we give feedback:
How many people accept the correct answer at the end?
Is that a function of TIS?
Is that a function of CT?

However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.

So we look at who changes their minds. We still see an influence of CT (? do stats) (and of TIS?).

Takeaway:
Relatively high K in this population
After feedback even higher
But not at ceiling
Relation with TIS/CT, both in K, and in integrating feedback (?)

However, for those who don't accept the feedback, we don't know if it's because they don't trust us, or they don't trust science.
We can also try slightly more stringent attention checks (the fact that some people go from good to bad suggests some inattention)

\FloatBarrier

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-bruderMeasuringIndividualDifferences2013}{}}%
Bruder, M., Haffke, P., Neave, N., Nouripanah, N., \& Imhoff, R. (2013). Measuring individual differences in generic beliefs in conspiracy theories across cultures: Conspiracy mentality questionnaire. \emph{Frontiers in Psychology}, \emph{4}. Retrieved from \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2013.00225}

\leavevmode\vadjust pre{\hypertarget{ref-colognaTrustScientistsTheir2024}{}}%
Cologna, V., Mede, N. G., Berger, S., Besley, J., Brick, C., Joubert, M., \ldots{} Linden, D. S. van der. (2024). \emph{Trust in scientists and their role in society across 67 countries}. \url{https://doi.org/10.31219/osf.io/6ay7s}

\leavevmode\vadjust pre{\hypertarget{ref-lantianMeasuringBeliefConspiracy2016}{}}%
Lantian, A., Muller, D., Nurra, C., \& Douglas, K. M. (2016). Measuring Belief in Conspiracy Theories: Validation of a French and English Single-Item Scale. \emph{International Review of Social Psychology}, \emph{29}(1), 1. \url{https://doi.org/10.5334/irsp.8}

\leavevmode\vadjust pre{\hypertarget{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}{}}%
Pennycook, G., Binnendyk, J., \& Rand, D. (2022). \emph{Overconfidently conspiratorial: Conspiracy believers are dispositionally overconfident and massively overestimate how much others agree with them}. \url{https://doi.org/10.31234/osf.io/d5fz2}

\end{CSLReferences}

\newpage

\hypertarget{appendix-appendix}{%
\appendix}

\hypertarget{exp1}{%
\section{Experiment 1}\label{exp1}}

\hypertarget{materials-1}{%
\subsection{Materials}\label{materials-1}}

\FloatBarrier

\hypertarget{conspiracy-scales-1}{%
\subsubsection{Conspiracy scales}\label{conspiracy-scales-1}}

Beside Belief in Conspiracy Theory Inventory (BCTI) by Pennycook et al. (2022) which report in the main study, we also assessed two other consipiracy thinking measures:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The conspiracy mentality questionnaire (CMQ) by Bruder et al. (2013):
  I think that . . .
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \ldots{} many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
\item
  \ldots{} government agencies closely monitor all citizens.
\item
  \ldots{} events which superficially seem to lack a connection are often the result of secret activities.
\item
  \ldots{} there are secret organizations that greatly influence political decisions.
\end{itemize}

{[}0\% - 100\%; 0 = certainly not, 100 = certain{]}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  The Single Item Conspiracy Beliefs Scale (SICBS) by Lantian et al. (2016) :
\end{enumerate}

\begin{itemize}
\tightlist
\item
  I think that the official version of the events given by the authorities very often hides the truth. {[}1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true{]}
\end{itemize}

\hypertarget{trust-in-science-1}{%
\subsubsection{Trust in science}\label{trust-in-science-1}}

We rely on three items

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
\item
  In general, would you say that you trust science a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
\item
  How much confidence do you have in scientists to act in the best interests of the public? {[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}
\end{enumerate}

\hypertarget{comparing-items}{%
\subsection{Comparing items}\label{comparing-items}}

Make correlation table

\hypertarget{conspiracy-theories}{%
\subsubsection{Conspiracy theories}\label{conspiracy-theories}}

Table \ref{tab:correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}

\begin{tabular}{llll}
\toprule
 & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
\midrule
BCTI & 1.00 & 0.58 & 0.56\\
CMQ & 0.58 & 1.00 & 0.77\\
SICBS & 0.56 & 0.77 & 1.00\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{trust-in-science-2}{%
\subsubsection{Trust in science}\label{trust-in-science-2}}

Table \ref{tab:correlation-trust} shows the correlations of the three different items measuring trust in science.

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:correlation-trust}Correlations of the three different items measuring trust in science}

\begin{tabular}{llll}
\toprule
 & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
\midrule
wgm\_sciencegeneral & 1.00 & 0.85 & 0.75\\
wgm\_scientists & 0.85 & 1.00 & 0.82\\
pew & 0.75 & 0.82 & 1.00\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{correlations-with-alternative-measures}{%
\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures}}

Table \ref{tab:correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}

\begin{tabular}{lll}
\toprule
outcome & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
\midrule
BCTI 
(main conspiracy measure) & -0.39 & -0.33\\
CMQ & -0.11 & -0.07\\
SICBS & -0.10 & -0.09\\
WGM trust scientists & 0.23 & 0.26\\
WGM trust general 
(main trust measure) & 0.29 & 0.27\\
PEW trust scientists & 0.16 & 0.21\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{results-conditional-on-false-responses}{%
\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses}}

Table \ref{tab:false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.

\begin{table}

\caption{\label{tab:false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
\midrule
(Intercept) & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
 & (0.073) & (0.074) & (0.074) & (0.073) & (0.074) & (0.074)\\
avg\_acceptance & -0.139+ & 0.008 & -0.028 & 0.108 & 0.064 & 0.056\\
 & (0.073) & (0.074) & (0.074) & (0.074) & (0.074) & (0.074)\\
\midrule
Num.Obs. & 184 & 184 & 184 & 184 & 184 & 184\\
R2 & 0.019 & 0.000 & 0.001 & 0.012 & 0.004 & 0.003\\
R2 Adj. & 0.014 & -0.005 & -0.005 & 0.006 & -0.001 & -0.002\\
AIC & 523.6 & 527.2 & 527.0 & 525.0 & 526.4 & 526.6\\
BIC & 533.2 & 536.8 & 536.7 & 534.7 & 536.1 & 536.2\\
Log.Lik. & -258.801 & -260.578 & -260.512 & -259.506 & -260.204 & -260.295\\
RMSE & 0.99 & 1.00 & 1.00 & 0.99 & 1.00 & 1.00\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\end{tabular}
\end{table}

\end{document}

### 2024-03-25 paper.Rmd
update author info after first study

@@
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - "Conceptualization"
-      - "Data collection"
-      - "Statistical Analyses"
-      - "Writing - Original Draft"
-      - "Writing - Review & Editing"
+
   - name          : "Lou Kerzreho"
     affiliation   : "1"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - "Models and Simulations"
-      - "Writing - Review & Editing"
+    role:
+
   - name          : "Hugo Mercier"
@@
     role:
-      - "Conceptualization"
-      - "Writing - Original Draft"
-      - "Writing - Review & Editing"
-      - "Supervision"
 
@@
     institution   : "Institut Jean Nicod, Département d’études cognitives, ENS, EHESS, PSL University, CNRS, France"
-  - id            : "2"
-    institution   : "Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany"
 
@@
 classoption       : "doc" # "doc" for nice look, "man" for manuscripty
-output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
+output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
 For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g.
@@
 
-(ref:exp1-conditional-acceptance) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus  **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus
+(ref:exp1-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
 

### 2024-03-25 paper.tex
update author info after first study

@@
 
-The authors made the following contributions.
-Jan Pfänder: Conceptualization, Data collection, Statistical Analyses, Writing - Original Draft, Writing - Review \& Editing; Lou Kerzreho: Models and Simulations, Writing - Review \& Editing; Hugo Mercier: Conceptualization, Writing - Original Draft, Writing - Review \& Editing, Supervision.
+The authors made the following contributions. Jan Pfänder: ; Lou Kerzreho: ; Hugo Mercier: .
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Institut Jean Nicod, Département d'études cognitives, ENS, EHESS, PSL University, CNRS, France\\\textsuperscript{2} Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Institut Jean Nicod, Département d'études cognitives, ENS, EHESS, PSL University, CNRS, France}
 
@@
 \includegraphics{output/figures/exp1-conditional-acceptance.pdf}
-\caption{\label{fig:exp1-conditional-acceptance}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
+\caption{\label{fig:exp1-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
 \end{figure}

### 2024-03-25 paper.Rmd
update draft study 1

@@
 
--   **RQ1: What is the average science knowledge score (1)?**
+**RQ1: What is the average science knowledge score (1)?**
 
--   **RQ2: What is the average acceptance of the scientific consensus (2)?**
+**RQ2: What is the average acceptance of the scientific consensus (2)?**
 
--   **RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**
+**RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**
 
--   **RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**
+**RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**
 
@@
 
-To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 and the Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 (see Appendix \@ref:exp1).
+To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 and the Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 (see Appendix \@ref(exp1)).
 
@@
 
-In appendix XX we show that these results hold for our alternative measures of trust and conspiracy thinking.
+In Appendix \@ref(exp1) we show that these results hold for our alternative measures of trust and conspiracy thinking.
 We also include more descriptive statistics, such as knowledge and acceptance by science questions.
@@
 We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance`), and only a very small correlation for trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance`).
-These associations are not statistically significant (see Appendix \@ref:exp1).
-
-The correlations of acceptance with
-
-However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.
-
-## Make a combined plot with all results
+These associations are not statistically significant (see Appendix \@ref(exp1)).
 
@@
 In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
-However, conditioning our analyses to initial false knowledge responses, i.e. looking at change of opinion towards the scientific consensus, we find that trust in science is onl
-
-(Experiment 1 Lou write up)
-
-If we ask basic science questions, and we give feedback:
-How many people accept the correct answer at the end?
-Is that a function of TIS?
-Is that a function of CT?
-
-However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.
-
-So we look at who changes their minds. We still see an influence of CT (? do stats) (and of TIS?).
-
-Takeaway: 
-Relatively high K in this population
-After feedback even higher
-But not at ceiling
-Relation with TIS/CT, both in K, and in integrating feedback (?)
-
-However, for those who don’t accept the feedback, we don’t know if it’s because they don’t trust us, or they don’t trust science.
-We can also try slightly more stringent attention checks (the fact that some people go from good to bad suggests some inattention)
+However, conditioning our analyses to initial false knowledge responses, i.e.
+looking at change of opinion towards the scientific consensus, we find that trust in science is onl
 

### 2024-03-25 paper.tex
update draft study 1

@@
 
-\begin{itemize}
-\item
 \textbf{RQ1: What is the average science knowledge score (1)?}
-\item
+
 \textbf{RQ2: What is the average acceptance of the scientific consensus (2)?}
-\item
+
 \textbf{RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?}
-\item
+
 \textbf{RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?}
-\end{itemize}
 
@@
 
-To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix @ref:exp1).
+To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix \ref{exp1}).
 
@@
 
-In appendix XX we show that these results hold for our alternative measures of trust and conspiracy thinking.
+In Appendix \ref{exp1} we show that these results hold for our alternative measures of trust and conspiracy thinking.
 We also include more descriptive statistics, such as knowledge and acceptance by science questions.
@@
 We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = -0.14), and only a very small correlation for trust in science (r = 0.06).
-These associations are not statistically significant (see Appendix @ref:exp1).
-
-The correlations of acceptance with
-
-However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.
-
-\hypertarget{make-a-combined-plot-with-all-results}{%
-\subsection{Make a combined plot with all results}\label{make-a-combined-plot-with-all-results}}
+These associations are not statistically significant (see Appendix \ref{exp1}).
 
@@
 In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
 However, conditioning our analyses to initial false knowledge responses, i.e.~looking at change of opinion towards the scientific consensus, we find that trust in science is onl
-
-(Experiment 1 Lou write up)
-
-If we ask basic science questions, and we give feedback:
-How many people accept the correct answer at the end?
-Is that a function of TIS?
-Is that a function of CT?
-
-However, trust/CT is related to basic knowledge, so the fact that it is also related to the post feedback answer could just be a function of that.
-
-So we look at who changes their minds. We still see an influence of CT (? do stats) (and of TIS?).
-
-Takeaway:
-Relatively high K in this population
-After feedback even higher
-But not at ceiling
-Relation with TIS/CT, both in K, and in integrating feedback (?)
-
-However, for those who don't accept the feedback, we don't know if it's because they don't trust us, or they don't trust science.
-We can also try slightly more stringent attention checks (the fact that some people go from good to bad suggests some inattention)
 

### 2024-04-01 exp_2/preregistration.Rmd
preregistration study 2

--- 2024-03-05 study_1/preregistration.Rmd
@@
 ---
-title: "Pre-registration Study 1 \n Who does not trust basic science?"
+title: "Pre-registration Study 2 \n Who does not trust basic science?"
 output:
@@
 
-```{r, include=FALSE, message=FALSE}
-# read in data from power simulation
-
-```
-
 # Introduction
 
-Anti-vax people or climate skepticists are often conceived of as simply anti-science.
-But would these people deny that water is made of molecules containing one oxygen and two hydrogen atoms, and that its chemical formula therefore is H2O?
-We suspect that everyone (or nearly everyone), including people who hold-anti science conspiracy beliefs, trusts basic science, at least when it doesn’t conflict with their specific beliefs.
+In a first study (https://osf.io/8utsj/) we tested whether participants would accept the scientific consensus on basic science facts.
+Most participants did, but not for all the questions.
+In this survey, we make two main adjustments to further investigate rejection of consensus: (i) we add a longer explanation and sources, instead of only stating the consensus.
+If it was the case that participants did not mistrust the science, but us the experimenters, this should increase acceptance; (ii) every time a participant rejects the scientific consensus, we ask them an open-ended question to explain why they did so.
+We also remove one science question that we had added ourselves (on where trees mainly draw there materials from to grow; answer: air), because the results were very different compared to other questions, suggesting the question might have been confusing.
 
@@
 
-> *We will ask you 11 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.*
+> *We will ask you 10 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.*
 
-Next, participants answer a set of 11 basic science questions, all but two extracted from existing science knowledge questionnaires.
-After each question, participants will be presented with an answer reflecting the scientific consensus.
+Next, participants answer a set of 10 basic science questions, all but one extracted from existing science knowledge questionnaires.
+After each question, participants will be presented with an answer reflecting the scientific consensus, along with an explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
 Participants are asked to choose whether they accept the answer or not, before proceeding to the next question.
-Figure \@ref(fig:stimulus-example) displays the survey for an example science question.
-Table \@ref(tab:knowledge) shows all included questions, their scientifically consensual answer, and their source.
+Figure 1 displays the survey for an example science question.
+Table 1 shows all included questions, their scientifically consensual answer, the explanations, and their source.
 
-(ref:stimulus-example) Example of a science question, the scientific consensus and the corresponding acceptance question.
+(ref:stimulus-example) Example of a science question, the explanation with references and the corresponding acceptance question.
 
-```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-example)"}
-knitr::include_graphics("../figures/study1_question_example.png")
+```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="center", fig.show="hold", fig.cap="(ref:stimulus-example)"}
+knitr::include_graphics("materials/study2_question_example.png")
 ```
@@
 ```{r knowledge}
-items <- read_csv("materials/knowledge_items.csv") %>% 
+items <- read_csv("materials/knowledge_items_2.csv") %>% 
   mutate(id = 1:nrow(.)) %>% 
-  select(id, Question, `Scientific consensus`, `Reference(s)`)
+  select(id, Question, `Scientific consensus`, `Explanation`)
 
@@
 kbl(items, booktabs = T, longtable = TRUE,
-    caption = "Science knowledge items") %>%
+    caption = "Science knowledge items", linesep = "\\addlinespace") %>%
   kable_paper(full_width = T) %>% 
   column_spec(1, width = "1em") %>%
-  column_spec(2, width = "10em") %>%
+  column_spec(2, width = "9em") %>%
   column_spec(3, width = "10em") %>%
-  column_spec(4, width = "15em") 
+  column_spec(4, width = "20em") 
 ```
-
 ## Conspiracy scales
@@
 
--   **RQ2: What is the average acceptance of the scientific consensus (2)?**
+In light of the previous study, we change RQ2 to:
 
-Similar to RQ1, we will report the average acceptance rate of the scientific consensus. We will look at the pooled average and the distribution of participant averages.
+-  **RQ2: When a participant’s answer does not match the consensus, how often do they change their mind and accept the consensus (2)?**
 
--   **RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**
+Similar to RQ1, we will report the average acceptance rate of the scientific consensus, when it was not already known. We will look at the pooled average and the distribution of participant averages.
 
-We will regress trust in science on the average knowledge and acceptance per participant.
+-  **RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
 
-If they correlate highly, we will use the average of the two Wellcome Global Monitor trust questions. If not, we will focus on the "How much do you trust science?" question.
+We will conduct a qualitative analysis and, if there are enough answers, attempt to categorize them.
 
-We will run a robustness check on the Pew question.
+In light of the results to the first study, we formulate the following hypotheses:
 
--   **RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**
+-  **H1a: Higher trust in science is associated with more science knowledge.**
+-  **H1b: Higher trust in science is associated with more acceptance of the scientific consensus, given the consensus is not already known.**
 
-We will proceed just as for RQ3, but with conspiracy thinking as outcome, instead of trust.
+We will regress trust in science on the average knowledge and acceptance (for those who initially disagreed) per participant.
+We will standardize all measures such that the regression coefficient yields a the correlation.
 
-To measure conspiracy thinking, we will use the average score of the Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022.
+We will focus on the “How much do you trust science?” question, but conduct robustness checks on the two other trust questions.
 
-As robustness checks, we will run the same analysis separately for the two other conspiracy scales (also averaging across all items).
+We will also check how well the three trust items correlate.
+
+-  **H2a: Higher conspiracy thinking is associated with more science knowledge.**
+-  **H2b: Higher conspiracy thinking is associated with more acceptance of the scientific consensus, given the consensus is not already known.**
+
+We will proceed just as for H1, but with conspiracy thinking as predictor, instead of trust.
+
+To measure conspiracy thinking, we will use the average score of the Conspiracy Theory Inventory (BCTI) by Pennycook, Binnendyk, and Rand (2022).
+As robustness checks, we will run the same analysis separately for the two other conspiracy scales.
 

### 2024-04-17 exp_3/cleaning.Rmd
preregistration study 3

--- 2024-03-25 exp_1/cleaning.Rmd
@@
 ---
-title: "Data cleaning Study 1"
+title: "Data cleaning Study 3"
 output: 
@@
 ```{r import, message=FALSE}
-d <- read_csv("./data/qualtrics.csv")
+d <- read_csv("data/test_study_3.csv")
 names(d)
@@
 
-There are 6 failed attention checks.
+There are % failed attention checks.
 
@@
 ```{r, message=FALSE}
-prolific_demographics <- read_csv("./data/prolific_demographics.csv")
+prolific_demographics <- read_csv("data/prolific_demographics.csv")
 
@@
 
+transform_text <- function(x) {
+  ifelse(str_detect(x, "^I agree"), "Yes", "No")
+}
+
 # clean and re-shape
@@
   ) %>% 
+  # clean the text for correction 
+  mutate_at(vars(ends_with("_confirm")), ~ transform_text(.))%>%
   # add average conspiracy measures 
@@
 
+We use the initial acceptance question and the confirmation question (only participants who did not accept the consensus for the intial question answered the confirmation question) to make a final acceptance question.
+This final acceptance will take on the initial answer in case of initial acceptance, and the answer of the confirmation question in case of initial non-acceptance.
+The final question will simply be named `acceptance`.
+
 ```{r}
@@
   # bring to long format
-  pivot_longer(cols = electrons_know:water_accept, 
-               names_to = "subject_question", values_to = "correct_accept") %>% 
+  pivot_longer(cols = c(ends_with("know"), ends_with("accept"), ends_with("confirm")), 
+               names_to = "subject_question", values_to = "answer") %>% 
   # separate conditions into CONVERGENCE_OUTCOME_STIMULUS
@@
                        ) %>%
-  # necessary because each participant only saw ten out of 11 questions
-  drop_na(correct_accept) %>% 
-  pivot_wider(names_from = question, values_from = correct_accept) %>% 
+  pivot_wider(names_from = question, values_from = answer) %>% 
 # create better variable names
   rename(knowledge = know, 
-         acceptance = accept) 
+         acceptance_initial = accept, 
+         acceptance_confirmation = confirm) %>% 
+  # make a final acceptance variable
+  mutate(acceptance = ifelse(is.na(acceptance_confirmation), acceptance_initial, acceptance_confirmation))
 ```
@@
 # make data frame with average particpant data and wide format
-d_wide <- left_join(d_wide %>% 
-                 select(-c(ends_with("know"), ends_with("accept"))), 
-               d_avg, 
+d_wide <- left_join(d_wide, d_avg, 
                by = "id")

### 2024-04-17 exp_3/preregistration.Rmd
preregistration study 3

--- 2024-03-05 study_1/preregistration.Rmd
@@
 ---
-title: "Pre-registration Study 1 \n Who does not trust basic science?"
+title: "Pre-registration Study 3 \n Who does not trust basic science?"
 output:
-  bookdown::pdf_document2: default
+  bookdown::pdf_document2:
+   extra_dependencies: ["float"]
   bookdown::word_document2: default
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE)
+knitr::opts_chunk$set(fig.pos="H", echo = FALSE, message = FALSE)
 ```
@@
 
-```{r, include=FALSE, message=FALSE}
-# read in data from power simulation
-
-```
-
 # Introduction
 
-Anti-vax people or climate skepticists are often conceived of as simply anti-science.
-But would these people deny that water is made of molecules containing one oxygen and two hydrogen atoms, and that its chemical formula therefore is H2O?
-We suspect that everyone (or nearly everyone), including people who hold-anti science conspiracy beliefs, trusts basic science, at least when it doesn’t conflict with their specific beliefs.
+In two previous studies (https://osf.io/8utsj/) we tested whether participants would accept the scientific consensus on basic science facts.
+We first asked participants to answer basic science questions in a multiple choice format.
+After they had provided their answer, we presented participants with the scientific consensus on the question.
+Most participants did accept this consensus, even after having provided the wrong answers - but not for all the questions.
+In the second study, we provided participants with more extensive explanations and sources for the scientific consensus, as we anticipated that this might increase acceptance.
+For cases in which participants still did not accept the consensus, we asked them to explain why.
+Both study 1 and 2 were run on convenience samples.
+Here, we will run study 2-–with some minor modifications-–on a sample of people holding anti-vaccination beliefs (see below).
+By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in case they didn’t), we provide them with an explicit opportunity to change their answer.
+Based on the answers to the open questions we got in study 2, we also propose a categorization scheme of reasons why people rejected the consensus.
 
@@
 
-> *We will ask you 11 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.*
+> *We will ask you 10 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.*
 
-Next, participants answer a set of 11 basic science questions, all but two extracted from existing science knowledge questionnaires.
-After each question, participants will be presented with an answer reflecting the scientific consensus.
+Next, participants answer a set of 10 basic science questions, all but one extracted from existing science knowledge questionnaires.
+After each question, participants will be presented with an answer reflecting the scientific consensus, along with an explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
 Participants are asked to choose whether they accept the answer or not, before proceeding to the next question.
-Figure \@ref(fig:stimulus-example) displays the survey for an example science question.
-Table \@ref(tab:knowledge) shows all included questions, their scientifically consensual answer, and their source.
+Figure \@ref(fig:stimulus-example) provides an example for one of the science questions.
+Table \@ref(tab:knowledge) shows all included questions, their scientifically consensual answer, the explanations, and their source.
 
-(ref:stimulus-example) Example of a science question, the scientific consensus and the corresponding acceptance question.
+(ref:stimulus-example) Example of a science question, the explanation with references and the corresponding acceptance question.
 
-```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-example)"}
-knitr::include_graphics("../figures/study1_question_example.png")
+```{r stimulus-example, echo=FALSE, out.width= "50%", fig.align="center", fig.show="hold", fig.cap="(ref:stimulus-example)"}
+knitr::include_graphics("materials/study2_question_example.png")
+```
+
+After answering all questions, if they had rejected the scientific consensus in some cases, we ask participants to explain why for each of these cases.
+We also offer them the possibility to change their answer (Fig.
+\@ref(fig:explanation-example)).
+
+(ref:explanation-example) Example of an explanation question and the opportunity to change the previous answer.
+
+```{r explanation-example, echo=FALSE, out.width= "50%", fig.align="center", fig.show="hold", fig.cap="(ref:explanation-example)"}
+knitr::include_graphics("materials/example_explanation_study3.png")
 ```
@@
 ```{r knowledge}
-items <- read_csv("materials/knowledge_items.csv") %>% 
+items <- read_csv("materials/knowledge_items_3.csv") %>% 
   mutate(id = 1:nrow(.)) %>% 
-  select(id, Question, `Scientific consensus`, `Reference(s)`)
+  select(id, Question, `Scientific consensus`, `Explanation`)
 
@@
 kbl(items, booktabs = T, longtable = TRUE,
-    caption = "Science knowledge items") %>%
+    caption = "Science knowledge items", linesep = "\\addlinespace") %>%
   kable_paper(full_width = T) %>% 
   column_spec(1, width = "1em") %>%
-  column_spec(2, width = "10em") %>%
+  column_spec(2, width = "9em") %>%
   column_spec(3, width = "10em") %>%
-  column_spec(4, width = "15em") 
+  column_spec(4, width = "20em") 
 ```
@@
 
--   **RQ1: What is the average science knowledge score (1)?**
+-   **RQ1: What is the average science knowledge score?**
 
@@
 
--   **RQ2: What is the average acceptance of the scientific consensus (2)?**
+-  **RQ2: What is the average acceptance of the scientific consensus**
 
-Similar to RQ1, we will report the average acceptance rate of the scientific consensus. We will look at the pooled average and the distribution of participant averages.
+Similar to RQ1, we will report the average acceptance rate of the scientific consensus.
+We will look at the pooled average and the distribution of participant averages.
+Note that we will proceed as in study 1, i.e.
+we will not condition on initially false answers as we did in study 2.
 
--   **RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**
+-  **RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
 
-We will regress trust in science on the average knowledge and acceptance per participant.
+Based on the results of our last study, we propose 5 categories of justification for the rejection of the scientific consensus:
 
-If they correlate highly, we will use the average of the two Wellcome Global Monitor trust questions. If not, we will focus on the "How much do you trust science?" question.
+1. No justification provided/no clear justification (e.g. “I read the article.“)
+2. Mistake, de facto acceptance of the consensus (e.g. "I made an error in my selection.")
+3. Not convinced by the explanation provided (e.g. "I feel like the excerpt that was provided did not definitively state the answer.")
 
-We will run a robustness check on the Pew question.
+*Motivated rejection*:
 
--   **RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**
+4. Motivated rejection due to personal convictions (e.g. "I do not believe in the heliocentric model of an earth spinning around a centralized sun.")
 
-We will proceed just as for RQ3, but with conspiracy thinking as outcome, instead of trust.
+5. Motivated rejection for religious reasons  (e.g. "I believe God make it they way it is and if He wanted it changed, He would have done it")
 
+When answers fall in category 2, we expect people to change their answers and accept the scientific consensus when asked a second time.
+Those answers will be counted in the analysis as acceptance of the consensus.
+
+# Hypotheses
+
+In light of the results to the previous studies, we formulate the following hypotheses:
+
+-  **H1a: Higher trust in science is associated with more science knowledge.**
+-  **H1b: Higher trust in science is associated with more acceptance of the scientific consensus.**
+
+We will regress trust in science on the average knowledge and acceptance per participant. We will focus on the “How much do you trust science?” question.
+We will run a robustness check on the second Wellcome Global Monitor trust question, on the Pew question, and a combined measure of all three question items.
+We will also check how well the three trust items correlate.
+
+-  **H2a: Higher conspiracy thinking is associated with less science knowledge.**
+-  **H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.**
+
+We will proceed just as for RQ3, but with conspiracy thinking as predictor, instead of trust.
 To measure conspiracy thinking, we will use the average score of the Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022.
-
 As robustness checks, we will run the same analysis separately for the two other conspiracy scales (also averaging across all items).
-
 We will also check how well the three conspiracy scales correlate.
@@
 
-We will recruit 200 participants from the US.
+We will recruit 200 participants from the US. To select a sample of participants distrusting vaccines, we combined three specific sample filters on  Prolific:
+
+1. **COVID-19 Vaccine Opinions**: "Participants were asked the following question: Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines:
+
+  - For (I feel positively about the vaccines)
+  - Against (I feel negatively about the vaccines)
+  - Neutral (I don't have strong opinions either way)
+  - Prefer not to say"
+
+    We select only people who answered "Against".
+
+2. **COVID-19 Vaccination**: "Participants were asked the following question: Have you received a coronavirus (COVID-19) vaccination?"
+
+  - Yes (at least one dose)
+  - No
+  - Prefer not to answer"
+
+    We select only people who answered "No".
+
+3.
+**Vaccine Opinions 2**: "Participants were asked the following question: On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children.
+
+ - 1 (TOTALLY DISAGREEE)
+ - 2 (DISAGREE)
+ - 3 (SOMEWHAT DISAGREE)
+ - 4 (NEITHER AGREE NOR DISAGREE)
+ - 5 (SOMEWHAT AGREE)
+ - 6 (AGREE) 
+ - 7 (TOTALLY AGREE)
+ - Rather not say"
+
+    We select only people who answered "1", "2", or "3".
+
+The total sample available on Prolific with those three criteria combined is 667.
+If this pool of participants ist too small and we don’t reach the targeted sample size of 200 within a week of data collection, we’ll first remove the criterion “1.
+COVID-19 Vaccine Opinions”.
+If within the next three days we still don’t reach the targeted sample size we will also remove the criterion “2.
+COVID-19 Vaccination”.
 

### 2024-05-31 appendix_exp2.Rmd
first write up studies 1 to 3

# Experiment 2 {#exp2}

## Materials

\FloatBarrier

See Appendix \@ref(exp1).

## Comparing items

### Conspiracy theories

Table \@ref(tab:exp2-correlation-conspiracy) shows the correlations of the three different scales assessing conspiracy thinking.

```{r exp2-correlation-conspiracy}
correlation_matrix <- exp2_wide %>%
  select(BCTI_avg, CMQ_avg, SICBS) %>% 
  rename(BCTI = BCTI_avg, 
         CMQ = CMQ_avg) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different scales assessing conspiracy thinking",
          placement = "h")
```

### Trust in science

Table \@ref(tab:exp2-correlation-trust) shows the correlations of the three different items measuring trust in science.

```{r exp2-correlation-trust}
correlation_matrix <- exp2_wide %>%
  select(wgm_sciencegeneral, wgm_scientists, pew) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different items measuring trust in science",
          placement = "h")
```

## Correlations with alternative measures

Table \@ref(tab:exp2-correlations-outcomes) shows the correlations between knowledge and acceptance, respectively, and outcome variables.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp2_wide %>% 
  select(avg_acceptance, avg_knowledge, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list_knowledge <- map(variables, ~ run_regression(standardized_variables, .x, "avg_knowledge"))
results_list_acceptance <- map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance"))
```

```{r exp2-correlations-outcomes}
# as data frame/output table
results <- bind_rows(list(results_list_knowledge, results_list_acceptance))
results %>% 
  filter(term != "intercept") %>%
  pivot_wider(names_from = term, 
              values_from = estimate) %>% 
  rename(`Correlation with knowledge` = avg_knowledge, 
         `Correlation with acceptance` = avg_acceptance, 
         ) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 
                             outcome == "CMQ_avg" ~ "CMQ", 
                             outcome == "SICBS" ~ "SICBS", 
                             outcome == "wgm_scientists" ~ "WGM trust scientists",
                             outcome == "wgm_sciencegeneral" ~ "WGM trust general \n(main trust measure)",
                             outcome == "pew" ~ "PEW trust scientists",
                             )) %>% 
  apa_table(caption = "Correlations between knowledge and acceptance, respectively, and outcome variables")
```

## Results conditional on false responses

Table \@ref(tab:exp2-false-response-regression) shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp2_false_knowledge %>% 
  select(avg_acceptance, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# Run regression for each variable and store results in a list
results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance", return = "model")), variables)
```

```{r exp2-false-response-regression}
modelsummary(results_list, 
             stars = TRUE,
             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values") 
```

## By-question variation

```{r}
# calculate share
plot_data <- exp2_long %>% 
  group_by(subject) %>% 
  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
```

### Knowledge

(ref:exp2-questions-knowledge) Distribution of correct answers by question.

```{r exp2-questions-knowledge, fig.cap="(ref:exp2-questions-knowledge)"}
# plot knowledge
ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of correct responses") +
  coord_flip()
```

### Acceptance

(ref:exp2-questions-acceptance) Distribution of consensus acceptance by question.

```{r exp2-questions-acceptance, fig.cap="(ref:exp2-questions-acceptance)"}
# plot acceptance
ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of acceptance") +
  coord_flip()
```

## Trust in science

(ref:exp2-trust-scientists) Distribution of trust in scientists.

```{r exp2-trust-scientists, fig.cap="(ref:exp2-trust-scientists)"}
# calculate share
plot_data <- exp2_wide %>% 
  group_by(wgm_scientists) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(share = n/sum(n)) %>% 
  rounded_numbers()

# plot knowledge
ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Trust in science", y = "Share of participants") +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) 
```

## Conspiracy thinking

### Distribution

(ref:exp2-conspiracy-distribution) Distribution of conspiracy thinking.

```{r exp2-conspiracy-distribution, fig.cap="(ref:exp2-conspiracy-distribution)"}
# plot knowledge
ggplot(exp2_wide, aes(x = BCTI_avg)) +
  geom_histogram() +
  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
```

### By-item variation

(ref:exp2-conspiracy-items) Distribution of conspiracy thinking by item.

```{r exp2-conspiracy-items, fig.cap="(ref:exp2-conspiracy-items)"}
BCTI_data <- exp2_long %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  select(id, c(starts_with("BCTI"), score)) %>% 
  group_by(BCTI_item) %>% 
  summarize(avg_score = mean(score, na.rm=TRUE)) %>% 
  ungroup()

ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
    geom_pointrange(aes(ymin = 1, ymax = avg_score),
                  # Make the lines a little thicker and the dots a little bigger
                  fatten = 2, size = 1) +
  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
  labs(x = "Conspiracy item", y = "Average score") +
  coord_cartesian(ylim = c(1, 9)) +
  coord_flip()
```

## Reasons for consensus rejection

### By category

[To do]

### All raw answers

```{r exp2-reasons-rejection}
# store all variables
reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
               "orbit", "diamonds", "speed", "salt", "water")

table <- exp2_wide %>% 
  select(id, all_of(reasons_rejection)) %>% 
  pivot_longer(all_of(reasons_rejection), 
               names_to = "question",
               values_to = "answer") %>% 
  drop_na(answer) %>% 
  arrange(id)

kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2) %>%
  column_spec(3, width = "30em")

```

### 2024-05-31 appendix_exp3.Rmd
first write up studies 1 to 3

# Experiment 3 {#exp3}

## Materials

\FloatBarrier

See Appendix \@ref(exp1).

## Comparing items

### Conspiracy theories

Table \@ref(tab:exp3-correlation-conspiracy) shows the correlations of the three different scales assessing conspiracy thinking.

```{r exp3-correlation-conspiracy}
correlation_matrix <- exp3_wide %>%
  select(BCTI_avg, CMQ_avg, SICBS) %>% 
  rename(BCTI = BCTI_avg, 
         CMQ = CMQ_avg) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different scales assessing conspiracy thinking",
          placement = "h")
```

### Trust in science

Table \@ref(tab:exp3-correlation-trust) shows the correlations of the three different items measuring trust in science.

```{r exp3-correlation-trust}
correlation_matrix <- exp3_wide %>%
  select(wgm_sciencegeneral, wgm_scientists, pew) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different items measuring trust in science",
          placement = "h")
```

## Correlations with alternative measures

Table \@ref(tab:exp3-correlations-outcomes) shows the correlations between knowledge and acceptance, respectively, and outcome variables.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp3_wide %>% 
  select(avg_acceptance, avg_knowledge, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list_knowledge <- map(variables, ~ run_regression(standardized_variables, .x, "avg_knowledge"))
results_list_acceptance <- map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance"))
```

```{r exp3-correlations-outcomes}
# as data frame/output table
results <- bind_rows(list(results_list_knowledge, results_list_acceptance))
results %>% 
  filter(term != "intercept") %>%
  pivot_wider(names_from = term, 
              values_from = estimate) %>% 
  rename(`Correlation with knowledge` = avg_knowledge, 
         `Correlation with acceptance` = avg_acceptance, 
  ) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 
                             outcome == "CMQ_avg" ~ "CMQ", 
                             outcome == "SICBS" ~ "SICBS", 
                             outcome == "wgm_scientists" ~ "WGM trust scientists",
                             outcome == "wgm_sciencegeneral" ~ "WGM trust general \n(main trust measure)",
                             outcome == "pew" ~ "PEW trust scientists",
                             )) %>% 
  apa_table(caption = "Correlations between knowledge and acceptance, respectively, and outcome variables")
```

## Results conditional on false responses

Table \@ref(tab:exp3-false-response-regression) shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp3_false_knowledge %>% 
  select(avg_acceptance, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# Run regression for each variable and store results in a list
results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance", return = "model")), variables)
```

```{r exp3-false-response-regression}
modelsummary(results_list, 
             stars = TRUE,
             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values") 
```

## By-question variation

```{r}
# calculate share
plot_data <- exp3_long %>% 
  group_by(subject) %>% 
  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
```

### Knowledge

(ref:exp3-questions-knowledge) Distribution of correct answers by question.

```{r exp3-questions-knowledge, fig.cap="(ref:exp3-questions-knowledge)"}
# plot knowledge
ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of correct responses") +
  coord_flip()
```

(ref:exp3-questions-acceptance) Distribution of consensus acceptance by question.

```{r exp3-questions-acceptance, fig.cap="(ref:exp3-questions-acceptance)"}
# plot acceptance
ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of acceptance") +
  coord_flip()
```

## Trust in science

(ref:exp3-trust-scientists) Distribution of trust in scientists.

```{r exp3-trust-scientists, fig.cap="(ref:exp3-trust-scientists)"}
# calculate share
plot_data <- exp3_wide %>% 
  group_by(wgm_scientists) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(share = n/sum(n)) %>% 
  rounded_numbers()

# plot knowledge
ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Trust in science", y = "Share of participants") +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) 
```

## Conspiracy thinking

(ref:exp3-conspiracy-distribution) Distribution of conspiracy thinking.

```{r exp3-conspiracy-distribution, fig.cap="(ref:exp3-conspiracy-distribution)"}
# plot knowledge
ggplot(exp3_wide, aes(x = BCTI_avg)) +
  geom_histogram() +
  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
```

### By-item variation

(ref:exp3-conspiracy-items) Distribution of conspiracy thinking by item.

```{r exp3-conspiracy-items, fig.cap="(ref:exp3-conspiracy-items)"}
BCTI_data <- exp3_long %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  select(id, c(starts_with("BCTI"), score)) %>% 
  group_by(BCTI_item) %>% 
  summarize(avg_score = mean(score, na.rm=TRUE)) %>% 
  ungroup()

ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
    geom_pointrange(aes(ymin = 1, ymax = avg_score),
                  # Make the lines a little thicker and the dots a little bigger
                  fatten = 2, size = 1) +
  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
  labs(x = "Conspiracy item", y = "Average score") +
  coord_cartesian(ylim = c(1, 9)) +
  coord_flip()
```

## Reasons for consensus rejection

### All raw answers

```{r exp3-reasons-rejection}
# store all variables
reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
               "orbit", "diamonds", "speed", "salt", "water")

table <- exp3_wide %>% 
  select(id, all_of(reasons_rejection)) %>% 
  pivot_longer(all_of(reasons_rejection), 
               names_to = "question",
               values_to = "answer") %>% 
  drop_na(answer) %>% 
  arrange(id)

# apa_table(table,
#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
#           longtable = TRUE
#           )
kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2) %>%
  column_spec(3, width = "30em")

```

## Why do people accept the scientific consensus:

### Participants who answered "other"

```{r exp3-other-reasons-acceptance}
# store all variables
table <- exp3_wide %>% 
  select(id, other_reason_agreement) %>% 
  drop_na(other_reason_agreement) %>% 
  arrange(id)

kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2, width = "30em")
```

## How did participants independently verifiy?

### Participants who answered "other"

```{r exp3-independent-verification}
# store all variables
table <- exp3_wide %>% 
  select(id, reason_followup) %>% 
  drop_na(reason_followup) %>% 
  arrange(id)

kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2, width = "30em")
```

### 2024-05-31 exp_2/cleaning.Rmd
preregistration study 3
first write up studies 1 to 3

--- 2024-03-25 exp_1/cleaning.Rmd
@@
 ---
-title: "Data cleaning Study 1"
+title: "Data cleaning Study 2"
 output: 
@@
 
-There are 6 failed attention checks.
+There are 11 failed attention checks.
 
@@
 
+## Justifications
+
+### Extract justifications
+
+```{r}
+# questions <- c("electrons", "antibiotics", "continents", "sex", "lasers", "orbit", "diamonds", "speed", "salt", "water")
+# 
+# justifications <- exp2_wide %>% 
+#   select(id, all_of(questions)) %>% 
+#   pivot_longer(all_of(questions), 
+#                names_to = "question", 
+#                values_to = "answer") %>% 
+#   drop_na(answer) %>% 
+#   arrange(id)
+# 
+# write_csv(justifications, "data/justifications.csv")
+```
+
+### Clean justifications
+
+Note that we had used the following coding scheme (where 4 would be the top-category for 5 and 6, i.e. 4 has never been coded):
+
+1	- no justification provided / no clear justification	
+2	- mistake, they actually accept the consensus	
+3	- not convinced by the explanation provided	
+4	- motivated rejection of the belief	
+5	-- do to personal convictions	
+6	-- for religious reasons
+
+```{r}
+justifications <- read_csv("data/justifications_coded.csv")
+
+# Research questions
+justifications <- justifications %>%
+  mutate(category = case_when(category == 1 ~ "No justification",
+                              category  == 2 ~  "Mistake",
+                              category  == 3 ~ "Not convinced",
+                              category  == 5 ~ "Personal convictions",
+                              category  == 6 ~ "Religious Beliefs",
+                              TRUE ~ NA_character_
+  )
+  )
+
+write_csv(justifications, "data/justifications_clean.csv")
+```
+
 ## Export data

### 2024-05-31 exp_3/analysis.Rmd
first write up studies 1 to 3

--- 2024-03-13 study_1/analysis.Rmd
@@
 ---
-title: "Analysis Study 1"
+title: "Analysis Study 3"
 output:
@@
 data_wide <- read_csv("data/cleaned_wide.csv") 
-
-# make a version with averages per participant
-d_avg <- data_long %>% 
-  # make numeric versions
-  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0), 
-         knowledge_num = ifelse(knowledge == TRUE, 1, 0)
-  ) %>% 
-  group_by(id) %>% 
-  # calculate by-participant averages
-  summarize(avg_acceptance  = sum(acceptance_num)/n(), 
-            avg_knowledge = sum(knowledge_num)/n()
-  )
-
-# make data frame with average particpant data and wide format
-d <- left_join(data_wide %>% 
-                 select(-c(ends_with("know"), ends_with("accept"))), 
-               d_avg, 
-               by = "id")
-
 ```
-
-# RQ1
-
-**RQ1: What is the average science knowledge score (1)?**
-
-```{r}
-# as table
-# d_avg %>% 
-#   mutate(avg_knowledge = as.factor(avg_knowledge)) %>% 
-#   group_by(avg_knowledge) %>% 
-#   count() %>% 
-#   ungroup() %>% 
-#   mutate(share = n / sum(n)) %>% 
-#   mutate_if(is.numeric, round, digits = 3)
-
-# as plot
-# Calculate the mean
-mean_value <- mean(d_avg$avg_knowledge)
-
-ggplot(d_avg, aes(x = avg_knowledge)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value, y = 40, label = paste("Mean =", round(mean_value, 2))), 
-            vjust = -0.5, hjust = 1.2) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  labs(x = "Average science knowledge", 
-       y = "N")
-```
-
-# RQ2
-
-**RQ2: What is the average acceptance of the scientific consensus (2)?**
-
-```{r}
-# Calculate the mean
-mean_value <- mean(d_avg$avg_acceptance)
-
-ggplot(d_avg, aes(x = avg_acceptance)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
-            vjust = -0.5, hjust = 1.2) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  labs(x = "Average consensus acceptance", 
-       y = "N")
-```
-
-# RQ3
-
-**RQ3: What is the relationship between trust in science and, respectively, knowledge and acceptance?**
-
-```{r}
-plot_data <- d %>% 
-  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
-  summarise(n = n())
-
-# Plot as a tile plot
-ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  # Add black border around tiles for better visibility
-  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_knowledge, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) +  # Scatter points within each tile
-  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average science knowledge", 
-       y = "Trust in science", 
-       fill = "N") +  # Update fill legend title
-  theme_minimal() 
-```
-
-```{r}
-plot_data <- d %>% 
-  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
-  summarise(n = n())
-
-# Plot as a tile plot
-ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  # Add black border around tiles for better visibility
-  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_acceptance, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) +  # Scatter points within each tile
-  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average consensus Acceptance", 
-       y = "Trust in science", 
-       fill = "N") +  # Update fill legend title
-  theme_minimal() 
-```
-
-# RQ4
-
-**RQ4: What is the relationship between conspiracy thinking and, respectively, knowledge and acceptance?**
-
-```{r}
-# Plot as a tile plot
-ggplot(d, aes(x = avg_knowledge, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  labs(x = "Average science knowledge", 
-       y = "Average conspiracy thinking") +  # Update fill legend title
-  theme_minimal() 
-```
-
-```{r}
-ggplot(d, aes(x = avg_acceptance, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  labs(x = "Average consensus acceptance", 
-       y = "Average conspiracy thinking") +  # Update fill legend title
-  theme_minimal() 
-```
-
-## Difference between false and true answer acceptance
-
-```{r}
-# make a version with averages per participant, grouped by false/true ansers 
-false_data_avg <- data_long %>% 
-  # make numeric versions
-  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
-  ) %>% 
-  group_by(id, knowledge) %>% 
-  # calculate by-participant averages
-  summarize(
-    n = n(),
-    n_accepted = sum(acceptance_num),
-    avg_acceptance  = sum(acceptance_num)/n(), 
-    BCTI_avg = mean(BCTI_avg)
-  ) %>% 
-  mutate(knowledge = ifelse(knowledge == FALSE, "False answers", "Correct answers"))
-
-# make data frame with average particpant data and wide format
-# false_data <- left_join(false_data, false_data_avg, by = "id")
-
-ggplot(false_data_avg, aes(x = avg_acceptance, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  labs(x = "Average consensus acceptance", 
-       y = "Average conspiracy thinking") +  # Update fill legend title
-  theme_minimal() +
-  facet_wrap(~knowledge)
-
-```
-
-Alternative plot
-
-```{r}
-plot_data <- data_long %>% 
-  mutate(BCTI_avg = round(BCTI_avg, digits = 0)) %>% 
-  group_by(knowledge, acceptance, BCTI_avg) %>% 
-  count() %>% 
-  group_by(knowledge, BCTI_avg) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3) %>% 
-  mutate(knowledge = ifelse(knowledge == FALSE, "False answers", "Correct answers"))
-
-ggplot(plot_data, aes(x = BCTI_avg, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  scale_x_continuous(breaks = seq(1, 9, 1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Average conspiracy thinking", fill = "Acceptance \nof consensus") +
-  theme_minimal() +
-  facet_wrap(~knowledge) +
-  theme(legend.position = "top")
-
-ggplot(data_long, aes(x = ))
-```
-
-# Additional RQs
-
-## Knowledge and Acceptance by science question
-
-```{r}
-# calculate share
-plot_data <- data_long %>% 
-  group_by(subject) %>% 
-  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
-            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
-```
-
-```{r}
-# plot knowledge
-ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of correct responses") +
-  coord_flip()
-```
-
-```{r}
-# plot acceptance
-ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of acceptance") +
-  coord_flip()
-```
-
-## Open ended comments
-```{r}
-levels(as.factor(d$comments))
-```
-
-## Is the share of correct responses associated with the share of acceptance ?
-
-```{r}
-plot_data <- d %>% 
-  group_by(avg_knowledge, avg_acceptance) %>% 
-  summarise(n = n())
-
-# Plot as a tile plot
-ggplot(plot_data, aes(x = avg_knowledge, y = avg_acceptance, fill = n)) +
-  geom_tile() +  # Add black border around tiles for better visibility
-  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_knowledge, y = avg_acceptance),
-              width = 0.03, height = 0.04) +  # Scatter points within each tile
-  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
-  labs(x = "Average science knowledge", 
-       y = "Average consensus acceptance", 
-       fill = "N") +  # Update fill legend title
-  theme_minimal() 
-```
-
-# What were the acceptance rates among true and false answers?
 
@@
 
+absolute_numbers <- plot_data %>% 
+  group_by(knowledge) %>% 
+  summarise(n = sum(n))
+
 ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
@@
   labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  theme_minimal()
+  theme_minimal() +
+  geom_bracket(data = absolute_numbers,
+               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
+                   xmax = as.numeric(factor(knowledge)) + 0.4, 
+                   y.position = 1.05, label =  paste0("n = ", n)),
+               inherit.aes = FALSE,
+               step.increase = 0)
 ```
-
-## To which extent did people believe the different conspiracy items ?
-
-```{r}
-BCTI_data <- data_long %>% 
-  # bring to long format
-  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
-               names_to = "BCTI_item", values_to = "score") %>% 
-  select(id, c(starts_with("BCTI"), score)) %>% 
-  group_by(BCTI_item) %>% 
-  summarize(avg_score = mean(score)) %>% 
-  ungroup()
-
-ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
-    geom_pointrange(aes(ymin = 1, ymax = avg_score),
-                  # Make the lines a little thicker and the dots a little bigger
-                  fatten = 2, size = 1) +
-  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
-  labs(x = "Conspiracy item", y = "Average score") +
-  coord_cartesian(ylim = c(1, 9)) +
-  coord_flip()
-```

### 2024-05-31 exp_3/cleaning.Rmd
first write up studies 1 to 3

@@
 ```{r import, message=FALSE}
-d <- read_csv("data/test_study_3.csv")
+d <- read_csv("data/qualtrics.csv")
 names(d)
@@
 
+Code the reasons for why people accept the scientific consensus:
+
+```{r}
+d_wide <- d_wide %>%
+  mutate(reason_agreement = case_when(
+    grepl("Other:", reason_agreement) ~ "other",
+    grepl("independently", reason_agreement) ~ "independent verification",
+    grepl("trust scientists", reason_agreement) ~ "trust in scientists",
+    TRUE ~ NA_character_  # This handles any NAs
+  )) %>% 
+  rename(other_reason_agreement = reason_agreement_3_TEXT)
+```
+
 Make long format data.
@@
                names_to = "subject_question", values_to = "answer") %>% 
-  # separate conditions into CONVERGENCE_OUTCOME_STIMULUS
+  # separate subject (e.g. water) and question (e.g. acceptance)
   separate_wider_delim(subject_question, "_", names = c("subject", "question")
@@
   mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0), 
-         knowledge_num = ifelse(knowledge == TRUE, 1, 0)
+         knowledge_num = ifelse(knowledge == TRUE, 1, 0), 
+         acceptance_initial_num = ifelse(acceptance_initial == "Yes", 1, 0)
   ) %>% 
@@
   summarize(avg_acceptance  = sum(acceptance_num)/n(), 
-            avg_knowledge = sum(knowledge_num)/n()
+            avg_knowledge = sum(knowledge_num)/n(),
+            avg_acceptance_initial  = sum(acceptance_initial_num)/n()
   )
 
-# make data frame with average particpant data and wide format
+# make data frame with average participant data and wide format
 d_wide <- left_join(d_wide, d_avg, 
                by = "id")
+```
+
+## Check NAs
+
+```{r}
+d_wide %>% 
+  summarize(across(c("wgm_scientists", "wgm_sciencegeneral", "pew", "reason_agreement", "BCTI_avg", "CMQ_avg", "SICBS", "avg_acceptance", "avg_knowledge"), 
+                   ~sum(is.na(.x))
+                   )
+            )
+```
+
+It's technically not possible to have NA's for trust in science and conspiracy questions, because we used forced responses on them.
+Checking the qualtrics, we know that we accidentally randomized people into two of three outcome measures: trust in science, conspiracy or reasons for accepting consensus.
+This leaves us with reduced sample sizes for all analyses concerning these outcomes (XX for trust in science question, XX for conspiracy measures, XX for accepting consensus).
+
+```{r}
+d_wide %>% 
+  summarize(across(c("wgm_scientists", "wgm_sciencegeneral", "pew", "reason_agreement", "BCTI_avg", "CMQ_avg", "SICBS"), 
+                   ~sum(!is.na(.x))
+                   )
+            ) %>% 
+  pivot_longer(everything(), 
+               names_to = "outcome", 
+               values_to = "n_valid")
+```
+
+## Clean justifications
+
+```{r}
+justifications <- read_csv("data/justifications_coded.csv")
+
+# Research questions
+justifications <- justifications %>%
+  mutate(category = case_when(final_coding == 1 ~ "No justification",
+                              final_coding == 2 ~  "Mistake",
+                              final_coding == 3 ~ "Not convinced",
+                              final_coding == 4 ~ "Personal convictions",
+                              final_coding == 5 ~ "Religious Beliefs",
+                              TRUE ~ NA_character_
+                              )
+         )
+
+write_csv(justifications, "data/justifications_clean.csv")
 ```

### 2024-05-31 paper.Rmd
first write up studies 1 to 3

@@
 author: 
-  - name          : "Jan Pfänder"
-    affiliation   : "1"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-
-  - name          : "Lou Kerzreho"
-    affiliation   : "1"
-    role:
-
-  - name          : "Hugo Mercier"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    email         : "hugo.mercier@gmail.com"
-    address: " "
-    role:
+  - name          : ""
+    affiliation   : ""
 
 affiliation:
-  - id            : "1"
-    institution   : "Institut Jean Nicod, Département d’études cognitives, ENS, EHESS, PSL University, CNRS, France"
-
-authornote: |
-  HM received funding from the ANR (SCALUP, ANR-21-CE28-0016-01), and from the John Tempelton Foundation ("An Evolutionary and Cultural Perspective on Intellectual Humility via Intellectual Curiosity and Epistemic Deference").
-The ANR also supported this project through FrontCog (ANR-17-EURE-0017), and PSL (ANR-10-IDEX-0001-02).
+  - id            : ""
+    institution   : ""
 
@@
 classoption       : "doc" # "doc" for nice look, "man" for manuscripty
-output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
+output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
 For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g.
@@
   - "appendix_exp1.Rmd"
+  - "appendix_exp2.Rmd"
+  - "appendix_exp3.Rmd"
 
@@
 
-There is increasingly more talk of defiance towards science and towards experts more generally (for science, see the intro to the stability paper, see also Paresman et al on populism?).
+There is increasingly more talk of defiance towards science and towards experts more generally.
 In every population, many people only trust science ‘some of the time’, and a sizable minority doesn’t trust it.
@@
 
-Here, we look at whether people, across different levels of trust in science (and of other relevant attitudes?
-Conspi?
-Populists?
-Apparently on Prolific we can specifically target antivax people for instance, which could be quite useful), accept basic scientific facts.
+Here, we look at whether people in the US, across different levels of trust in science (including a sample with anti-vaccine attitudes), accept basic scientific facts.
 
@@
 
-# First series of experiments
+# Overview of experiments
 
-The goal is to see how many people really don’t accept basic science (in practice, get as close to ceiling in acceptance as we can, first with a generic sample, then see if that still works with an anti-vax sample).
+The goal is to see to which extent people do not accept basic science.
 
@@
 # RQ3
-cor_trust_knowledge <- cor(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_knowledge) %>% round(digits = 2)
+cor_trust_knowledge <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_knowledge) %>% text_ready()
 
-cor_trust_acceptance <- cor(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_acceptance) %>% round(digits = 2)
+cor_trust_acceptance <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_acceptance) %>% text_ready()
 
 # RQ4
-cor_conspiracy_knowledge <- cor(exp1_wide$BCTI_avg, exp1_wide$avg_knowledge) %>% round(digits = 2)
+cor_conspiracy_knowledge <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_knowledge) %>% text_ready()
 
-cor_conspiracy_acceptance <- cor(exp1_wide$BCTI_avg, exp1_wide$avg_acceptance) %>% round(digits = 2)
+cor_conspiracy_acceptance <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_acceptance)%>% text_ready()
 
@@
 
-false_answers_cor_trust_acceptance <- cor(exp1_false_knowledge$wgm_sciencegeneral, exp1_false_knowledge$avg_acceptance) %>% round(digits = 2)
+false_answers_cor_trust_acceptance <- cor.test(exp1_false_knowledge$wgm_sciencegeneral, exp1_false_knowledge$avg_acceptance) %>% text_ready()
 
-false_answers_cor_conspiracy_acceptance <- cor(exp1_false_knowledge$BCTI_avg, exp1_false_knowledge$avg_acceptance) %>% round(digits = 2)
+false_answers_cor_conspiracy_acceptance <- cor.test(exp1_false_knowledge$BCTI_avg, exp1_false_knowledge$avg_acceptance) %>% text_ready()
 
@@
 ```{r knowledge}
-items <- read_csv("exp_1/materials/knowledge_items.csv") %>% 
+items <- read_csv("exp_3/materials/overview_questions.csv") %>% 
   mutate(id = 1:nrow(.)) %>% 
-  select(id, Question, `Scientific consensus`, `Reference(s)`)
+  select(id, "Question", "Scientific consensus (Study 1)", "Explanation (Study 2 & 3)")
 
@@
     full_width = T) %>%
+  kable_styling(font_size = 8) %>% 
   column_spec(1, width = "1em") %>%
@@
   column_spec(3, width = "10em") %>%
-  column_spec(4, width = "15em") 
+  column_spec(4, width = "23em") 
 ```
@@
     full_width = F) %>%
+  kable_styling(font_size = 8) %>% 
   column_spec(1) %>%
-  column_spec(2, width = "30em")
+  column_spec(2, width = "40em")
 ```
@@
   labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  theme_minimal() +
+  plot_theme +
   geom_bracket(data = absolute_numbers,
@@
 
-For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = `r exp1_descriptives$cor_trust_knowledge`), and acceptance of scientific consensus and trust in science ( r = `r exp1_descriptives$cor_trust_acceptance`).
+For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus and trust in science (r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`).
 The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
@@
 
-For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = `r exp1_descriptives$cor_conspiracy_knowledge`), and conspiracy thinking and acceptance of scientific consensus ( r = `r exp1_descriptives$cor_conspiracy_acceptance`).
+For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`), and conspiracy thinking and acceptance of scientific consensus (r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`).
 
-In Appendix \@ref(exp1) we show that these results hold for our alternative measures of trust and conspiracy thinking.
-We also include more descriptive statistics, such as knowledge and acceptance by science questions.
+In Appendix \@ref(exp1), we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \@ref(exp1) also includes more descriptive statistics, such as knowledge and acceptance by science questions.
 
@@
 We than calculated the correlation between trust in science and average acceptance rate by participant.
-We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance`), and only a very small correlation for trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance`).
-These associations are not statistically significant (see Appendix \@ref(exp1)).
+We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`), nor with trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`).
 
@@
 
-These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus.
+These results suggest that most people accept the scientific consensus most of the time.
+Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
 Yet, in `r exp1_descriptives$conditional_acceptance$false$No$share` of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
 In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-However, conditioning our analyses to initial false knowledge responses, i.e.
-looking at change of opinion towards the scientific consensus, we find that trust in science is onl
+
+# Experiment 2
+
+```{r exp2}
+# Analyze data of experiments and store results
+
+# Experiment 2
+
+# read the cleaned long version data set
+exp2_long <- read_csv("exp_2/data/cleaned_long.csv")
+
+# read wide version data set
+exp2_wide <- read_csv("exp_2/data/cleaned_wide.csv")
+
+# read coded justifications
+exp2_justifications <- read_csv("exp_2/data/justifications_clean.csv")
+
+# Research questions
+
+# H1a
+cor_trust_knowledge <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_knowledge) %>% text_ready()
+
+# H2a
+cor_conspiracy_knowledge <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_knowledge) %>% text_ready()
+
+# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
+exp2_false_knowledge <- exp2_long %>% 
+  # make numeric versions
+  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
+  ) %>% 
+  group_by(id, knowledge) %>% 
+  # calculate by-participant averages
+  summarize(
+    n = n(),
+    n_accepted = sum(acceptance_num),
+    avg_acceptance  = sum(acceptance_num)/n(), 
+    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
+    BCTI_avg = mean(BCTI_avg), 
+    # add additonal measures for robustness checks
+    CMQ_avg = mean(CMQ_avg), 
+    SICBS = mean(SICBS),
+    wgm_scientists = mean(wgm_scientists), 
+    pew = mean(pew)
+  ) %>% 
+  # filter to only false responses
+  filter(knowledge == FALSE) %>% 
+  ungroup()
+
+# H1b
+# conditional on false answers
+false_answers_cor_trust_acceptance <- cor.test(exp2_false_knowledge$wgm_sciencegeneral, exp2_false_knowledge$avg_acceptance) %>% text_ready()
+
+# pooled across all cases
+cor_trust_acceptance <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_acceptance) %>% text_ready()
+
+# H2b
+# conditional on false answers
+false_answers_cor_conspiracy_acceptance <- cor.test(exp2_false_knowledge$BCTI_avg, exp2_false_knowledge$avg_acceptance) %>% text_ready()
+
+# pooled across all cases
+cor_conspiracy_acceptance <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_acceptance) %>% text_ready()
+
+# extract descriptives for inline reporting
+exp2_descriptives <- list(
+  # Demographics
+  n_subj = n_distinct(exp2_wide$id),
+  gender = exp2_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
+  age = exp2_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
+                                       .names = "{.fn}")) %>% rounded_numbers(),
+  # RQ1 & RQ2
+  means = exp2_wide %>% 
+    summarize(across(c(avg_knowledge, avg_acceptance), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()) %>% 
+    mutate_if(is.numeric, round, digits = 2) %>% 
+  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
+  # H1
+  # a
+  cor_trust_knowledge = cor_trust_knowledge,
+  # b (conditional on false responses)
+  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
+  # b (all cases)
+  cor_trust_acceptance = cor_trust_acceptance,
+  # H2
+  # a
+  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
+  # b (conditional on false responses)
+  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
+  # b(all cases)
+  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
+  # Exploratory
+  # acceptance and knowledge
+  conditional_acceptance = exp2_long %>% 
+    group_by(knowledge, acceptance) %>% 
+    count() %>% 
+    group_by(knowledge) %>% 
+    mutate (share = n/sum(n), 
+            # rename knowledge values
+            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate (share = paste0(share*100, " %")) %>% 
+    super_split(knowledge, acceptance),
+  # Justifications
+  justifications_n = nrow(exp2_justifications), 
+  justifications_n_participants = n_distinct(exp2_justifications$id),
+  justification_by_category = exp2_justifications %>%
+    group_by(category)%>%
+    summarize(n = n(), 
+              n_subjects = n_distinct(id)) %>%
+    mutate(share = n/sum(n)) %>%
+    mutate_if(is.numeric, round, digits=3) %>% 
+    mutate(share = paste0(share*100, "%")) %>% 
+    split(.$category)
+)
+```
+
+In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
+In most instances they did, but not always.
+In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
+To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
+To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
+We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \@ref(exp1)).
+
+Based on experiment 1, we formulated the following hypotheses:
+
+**H1a: Higher trust in science is associated with more science knowledge?**
+
+**H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?**
+
+**H2a: Higher conspiracy thinking is associated with less science knowledge?**
+
+**H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?**
+
+We had the following research questions:
+
+**RQ1: What is the average science knowledge score?**
+
+**RQ2: When a participant’s answer does not match the consensus, how often do they change their mind and accept the consensus?**
+
+**RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
+
+## Methods
+
+### Participants
+
+We recruited 201 participants from the US via prolific.
+11 participants failed our attention check, resulting in a final sample of `r exp2_descriptives$n_subj` participants (`r exp2_descriptives$gender$female$n` female, `r exp2_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp2_descriptives$age$mean`, $age_\text{sd}$: `r exp2_descriptives$age$sd`, $age_\text{median}$: `r exp2_descriptives$age$median`).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+### Procedure
+
+The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
+\@ref(fig:exp2-stimulus-example).
+
+(ref:exp2-stimulus-example) Example of a science question, the scientific consensus and corresponding explanation and sources.
+
+```{r exp2-stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:exp2-stimulus-example)"}
+knitr::include_graphics("./figures/study2_question_example.png")
+```
+
+### Materials
+
+We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
+
+## Results
+
+As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`).
+By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
+For trust in science, we find no statistically significant correlation (r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`).
+For conspiracy thinking we find a small negative one (r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
+
+(ref:exp2-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
+
+```{r exp2-conditional-acceptance, fig.cap="(ref:exp2-conditional-acceptance)"}
+plot_data <- exp2_long %>% 
+  group_by(knowledge, acceptance) %>% 
+  count() %>% 
+  group_by(knowledge) %>% 
+  mutate (share = n/sum(n)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
+
+absolute_numbers <- plot_data %>% 
+  group_by(knowledge) %>% 
+  summarise(n = sum(n))
+
+ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
+  geom_col(position = "dodge") +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5, size = 3) +
+  scale_fill_viridis_d(option = "cividis") +
+  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
+  plot_theme  +
+  geom_bracket(data = absolute_numbers,
+               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
+                   xmax = as.numeric(factor(knowledge)) + 0.4, 
+                   y.position = 1.05, label =  paste0("n = ", n)),
+               inherit.aes = FALSE,
+               step.increase = 0)
+```
+
+Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\@ref(fig:exp2-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+In Appendix \@ref(exp2) we show that these results hold for our alternative measures of trust and conspiracy thinking.
+We also include more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Regarding RQ1, participants answered on average `r exp2_descriptives$means$knowledge_mean` (sd = `r exp2_descriptives$means$knowledge_sd`) of the questions correctly, and accepted the scientific consensus on average for `r exp2_descriptives$means$acceptance_mean` (sd = `r exp2_descriptives$means$acceptance_sd`) of the questions.
+Fig.
+\@ref(fig:exp2-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
+In response to RQ2, in most cases (`r exp2_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In very few cases (`r exp2_descriptives$conditional_acceptance$true$No$share`), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+
+(ref:exp2-plot) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
+
+```{r exp2-plot, fig.cap="(ref:exp2-plot)", fig.height= 12, fig.width=12}
+# Plot distributions of knowledge and acceptance and their mean
+
+mean_value_knowledge <- mean(exp2_wide$avg_knowledge)
+max_n <- max(table(exp2_wide$avg_acceptance))
+
+knowledge_mean <- ggplot(exp2_wide, aes(x = avg_knowledge)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average science knowledge", 
+       y = "N")
+
+mean_value <- mean(exp2_wide$avg_acceptance)
+
+acceptance_mean <- ggplot(exp2_wide, aes(x = avg_acceptance)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average consensus acceptance", 
+       y = "N")
+
+# Trust in science plots
+plot_data <- exp2_wide %>% 
+  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
+  summarise(n = n())
+
+trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  # Add black border around tiles for better visibility
+  geom_jitter(inherit.aes = FALSE, data = exp2_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) +  # Scatter points within each tile
+  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average science knowledge", 
+       y = "Trust in science", 
+       fill = "N")
+
+plot_data <- exp2_wide %>% 
+  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
+  summarise(n = n())
+
+trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  
+  geom_jitter(inherit.aes = FALSE, data = exp2_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) + 
+  scale_fill_viridis_c(option = "plasma") + 
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average consensus Acceptance", 
+       y = "Trust in science", 
+       fill = "N") +
+  theme(legend.position = "top")
+
+# Conspiracy Thinking plots
+
+conspiracy_knowledge <- ggplot(exp2_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average science knowledge", 
+       y = "Average conspiracy thinking")
+
+conspiracy_acceptance <- ggplot(exp2_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average consensus acceptance", 
+       y = "Average conspiracy thinking") +
+  theme(legend.position = "top")
+
+# Combine plots
+
+# distributions
+distributions <- ggarrange(knowledge_mean + 
+                             rremove("xlab") , 
+                           acceptance_mean + 
+                             rremove("xlab") +
+                             rremove("ylab"))
+
+# trust
+trust <- ggarrange(trust_knowledge + 
+                             rremove("xlab") +
+                             guides(fill = "none") +
+                             labs(caption = NULL) , 
+                           trust_acceptance + 
+                             rremove("xlab") +
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# conspiracy
+conspiracy <- ggarrange(conspiracy_knowledge, 
+                           conspiracy_acceptance + 
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# common plot
+patchwork <- (distributions / trust / conspiracy) + 
+  plot_annotation(tag_levels = 'A')
+
+patchwork
+```
+
+For RQ3, we got `r exp2_descriptives$justifications_n` answers from `r exp2_descriptives$justifications_n_participants` different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \@ref(tab:exp2-justifications) summarizes these answers by five categories.
+All answers can be read in Appendix \@ref(exp2).
+
+```{r exp2-justifications}
+exp2_justifications %>%
+  group_by(category)%>%
+  summarize(n = n(), 
+            n_subjects = n_distinct(id)) %>%
+  mutate(Share = n/sum(n)) %>%
+  mutate_if(is.numeric, round, digits=3) %>% 
+  mutate(Share = paste0(Share*100, "%")) %>% 
+  arrange(desc(n)) %>% 
+  select(category, n, Share, n_subjects) %>% 
+  rename(Category = category,
+         `N (instances)` = n, 
+         `Share (instances)` = Share, 
+         `N (unique participants)` = n_subjects
+         ) %>% 
+  apa_table(caption = "Justifications by category")
+```
+
+## Discussion
+
+Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.
+answered the knowledge question wrongly).
+The share of these latter is considerably larger in experiment 2 (`r exp2_descriptives$conditional_acceptance$false$Yes$share`) than in experiment 1 (`r exp1_descriptives$conditional_acceptance$false$Yes$share`).
+While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
+We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
+
+# Experiment 3
+
+```{r exp3}
+# Analyze data of experiments and store results
+
+# Experiment 3
+
+# read the cleaned long version data set
+exp3_long <- read_csv("exp_3/data/cleaned_long.csv")
+
+# read wide version data set
+exp3_wide <- read_csv("exp_3/data/cleaned_wide.csv")
+
+# read coded justifications
+exp3_justifications <- read_csv("exp_3/data/justifications_clean.csv")
+
+# Research questions
+
+# H1
+cor_trust_knowledge <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_knowledge,
+  use = "complete.obs") %>% text_ready()
+cor_trust_acceptance <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_acceptance,
+  use = "complete.obs") %>% text_ready()
+
+# H2
+cor_conspiracy_knowledge <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_knowledge,
+  use = "complete.obs") %>% text_ready()
+cor_conspiracy_acceptance <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_acceptance,
+  use = "complete.obs") %>% text_ready()
+
+# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
+exp3_false_knowledge <- exp3_long %>% 
+  # make numeric versions
+  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
+  ) %>% 
+  group_by(id, knowledge) %>% 
+  # calculate by-participant averages
+  summarize(
+    n = n(),
+    n_accepted = sum(acceptance_num),
+    avg_acceptance  = sum(acceptance_num)/n(), 
+    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
+    BCTI_avg = mean(BCTI_avg), 
+    # add additonal measures for robustness checks
+    CMQ_avg = mean(CMQ_avg), 
+    SICBS = mean(SICBS),
+    wgm_scientists = mean(wgm_scientists), 
+    pew = mean(pew)
+  ) %>% 
+  # filter to only false responses
+  filter(knowledge == FALSE) %>% 
+  ungroup()
+
+# H1b conditional on false answers
+false_answers_cor_trust_acceptance <- cor.test(exp3_false_knowledge$wgm_sciencegeneral, exp3_false_knowledge$avg_acceptance,
+  use = "complete.obs") %>% 
+  text_ready()
+
+# H2b conditional on false answers
+false_answers_cor_conspiracy_acceptance <- cor.test(exp3_false_knowledge$BCTI_avg, exp3_false_knowledge$avg_acceptance,
+  use = "complete.obs") %>% 
+  text_ready()
+
+# extract descriptives for inline reporting
+exp3_descriptives <- list(
+  # Demographics
+  n_subj = n_distinct(exp3_wide$id),
+  n_subject_by_outcome = exp3_wide %>% 
+    summarize(across(c("wgm_scientists", "reason_agreement", "BCTI_avg"), 
+                     ~sum(!is.na(.x)
+                     )
+    )
+    ),
+  gender = exp3_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
+  age = exp3_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
+                                       .names = "{.fn}")) %>% rounded_numbers(),
+  # RQ1 & RQ2
+  means = exp3_wide %>% 
+    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()) %>% 
+    mutate_if(is.numeric, round, digits = 2) %>% 
+  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
+  # H1
+  # a
+  cor_trust_knowledge = cor_trust_knowledge,
+  # b (conditional on false responses)
+  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
+  # b (all cases)
+  cor_trust_acceptance = cor_trust_acceptance,
+  # H2
+  # a
+  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
+  # b (conditional on false responses)
+  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
+  # b(all cases)
+  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
+  # Exploratory
+  # acceptance and knowledge
+  conditional_acceptance = exp3_long %>% 
+    pivot_longer(c(acceptance, acceptance_initial), 
+                 names_to = "measure", 
+                 values_to = "acceptance") %>% 
+    group_by(knowledge, measure, acceptance) %>% 
+    count() %>% 
+    group_by(knowledge, measure) %>% 
+    mutate (share = n/sum(n), 
+            # rename knowledge values
+            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate (share = paste0(share*100, " %")) %>% 
+    super_split(knowledge,  measure, acceptance),
+  # Justifications
+  justifications_n = nrow(exp3_justifications), 
+  justifications_n_participants = n_distinct(exp3_justifications$id),
+  justification_by_category = exp3_justifications %>%
+    group_by(category)%>%
+    summarize(n = n(), 
+              n_subjects = n_distinct(id)) %>%
+    mutate(share = n/sum(n)) %>%
+    mutate_if(is.numeric, round, digits=3) %>% 
+    mutate(share = paste0(share*100, "%")) %>% 
+    split(.$category),
+  # Consensus acceptance
+  acceptance_n = exp3_wide %>% 
+    filter(!is.na(reason_agreement)) %>% 
+    nrow(.),
+  acceptance_by_reason = exp3_wide %>% 
+    group_by(reason_agreement) %>% 
+    summarize(n = n()) %>%
+    drop_na(reason_agreement) %>% 
+    mutate(share = n/sum(n)) %>%
+    mutate_if(is.numeric, round, digits=3) %>% 
+    mutate(share = paste0(share*100, "%")) %>% 
+    split(.$reason_agreement),
+  reason_followup_n = exp3_wide %>% 
+    filter(!is.na(reason_followup)) %>% 
+    nrow(.)
+)
+
+```
+
+Study 3 is essentially a replication-–with some minor modifications-–of study 2, but on a different type of sample.
+Both study 1 and 2 were run on convenience samples.
+For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
+By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
+\@ref(fig:exp3-explanation-example)).
+Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
+Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
+We want to know if participants perceive that this is because of trust, or other factors.
+
+As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
+
+**H1a: Higher trust in science is associated with more science knowledge?**
+
+**H1b: Higher trust in science is associated with more acceptance of the scientific consensus.**
+
+**H2a: Higher conspiracy thinking is associated with less science knowledge?**
+
+**H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.**
+
+We had the following research questions:
+
+**RQ1: What is the average science knowledge score?**
+
+**RQ2: What is the average acceptance of the scientific consensus**
+
+**RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
+
+**RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?**
+
+## Methods
+
+### Participants
+
+We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of `r exp3_descriptives$n_subj` participants (`r exp3_descriptives$gender$female$n` female, `r exp3_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp3_descriptives$age$mean`, $age_\text{sd}$: `r exp3_descriptives$age$sd`, $age_\text{median}$: `r exp3_descriptives$age$median`).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
+This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = `r exp3_descriptives$n_subject_by_outcome$wgm_scientists` for trust in science, N = `r exp3_descriptives$n_subject_by_outcome$BCTI_avg` for conspiracy measures, N = `r exp3_descriptives$n_subject_by_outcome$reason_agreement` for reason for accepting consensus).
+
+### Procedure
+
+The procedure was mostly the same as in experiment 2.
+In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
+Finally, at the end of the survey, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain.
+Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".
+
+(ref:exp3-explanation-example) Example of an explanation question and the opportunity to change the previous answer.
+
+```{r exp3-explanation-example, echo=FALSE, out.width= "50%", fig.align="center", fig.show="hold", fig.cap="(ref:exp3-explanation-example)"}
+knitr::include_graphics("figures/study3_example_explanation.png")
+```
+
+### Materials
+
+We relied on the same items as in experiment 2.
+We used a pre-registered categorization scheme for the open ended answers for why people reject the consensus.
+The categories were XX and XX.
+[Describe coding process, with independent reviewer etc.]
+
+## Results
+
+As in study 1 and 2, we find that participants answered on average `r exp3_descriptives$means$knowledge_mean` (sd = `r exp3_descriptives$means$knowledge_sd`) of the questions correctly, and initially accepted the scientific consensus on average for `r exp3_descriptives$means$acceptance_initial_mean` (sd = `r exp3_descriptives$means$acceptance_initial_sd`) of the questions (RQ2).
+The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (`r exp3_descriptives$means$acceptance_mean`, sd = `r exp3_descriptives$means$acceptance_sd`).
+
+Fig.
+\@ref(fig:exp3-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
+In most cases (`r exp3_descriptives$conditional_acceptance$false$acceptance_initial$Yes$share`), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
+After providing the chance to revise the initial consensus rejection, this share is even larger (`r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`).
+In very few cases (`r exp3_descriptives$conditional_acceptance$true$acceptance_initial$No$share`), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
+This share drops slightly after providing the chance to revise the initial consensus rejection (`r exp3_descriptives$conditional_acceptance$true$acceptance$No$share`).
+
+In the appendix XX we provide an analysis that this drop is statistically significant [TO DO].
+
+For all correlations, we include opinion revisions for measuring consensus acceptance.
+We find no statistically significant correlation between science knowledge and trust in science (r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`).
+Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
+\@ref(fig:exp3-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+We find no statistically significant correlation between conspiracy thinking and science knowledge (r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`), and between conspiracy thinking and acceptance of scientific consensus (r = `r exp3_descriptives$cor_conspiracy_acceptance`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`).
+
+In Appendix \@ref(exp3), we show that these associations are statistically significant (except the last one?).
+We show that they are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \@ref(exp3) also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+
+(ref:exp3-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
+
+```{r exp3-conditional-acceptance, fig.cap="(ref:exp3-conditional-acceptance)"}
+plot_data <- exp3_long %>% 
+  pivot_longer(c(acceptance, acceptance_initial), 
+               names_to = "measure", 
+               values_to = "acceptance") %>% 
+  # give nicer names
+  mutate(measure = case_when(measure == "acceptance" ~ "Acceptance including revisions", 
+                             measure == "acceptance_initial" ~ "Initial acceptance"), 
+         # make a factor
+         measure = factor(measure, levels = unique(measure)[c(2, 1)])
+         ) %>% 
+  group_by(measure, knowledge, acceptance) %>% 
+  count() %>% 
+  group_by(measure, knowledge) %>% 
+  mutate (share = n/sum(n)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
+
+absolute_numbers <- plot_data %>% 
+  group_by(measure, knowledge) %>% 
+  summarise(n = sum(n))
+
+ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
+  geom_col(position = "dodge") +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5, size = 3) +
+  scale_fill_viridis_d(option = "cividis") +
+  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
+  plot_theme +
+  geom_bracket(data = absolute_numbers,
+               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
+                   xmax = as.numeric(factor(knowledge)) + 0.4, 
+                   y.position = 1.05, label =  paste0("n = ", n)),
+               inherit.aes = FALSE,
+               step.increase = 0) +
+  facet_wrap(~measure)
+```
+
+Regarding RQ1, participants answered on average `r exp3_descriptives$means$knowledge_mean` (sd = `r exp3_descriptives$means$knowledge_sd`) of the questions correctly.
+For RQ3, we got `r exp3_descriptives$justifications_n` answers from `r exp3_descriptives$justifications_n_participants` different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \@ref(tab:exp3-justifications) summarizes these answers by five categories.
+
+All answers can be read in the appendix.
+
+```{r exp3-justifications}
+exp3_justifications %>%
+  group_by(category)%>%
+  summarize(n = n(), 
+            n_subjects = n_distinct(id)) %>%
+  mutate(Share = n/sum(n)) %>%
+  mutate_if(is.numeric, round, digits=3) %>% 
+  mutate(Share = paste0(Share*100, "%")) %>% 
+  arrange(desc(n)) %>% 
+  select(category, n, Share, n_subjects) %>% 
+  rename(Category = category,
+         `N (instances)` = n, 
+         `Share (instances)` = Share, 
+         `N (unique participants)` = n_subjects
+         ) %>% 
+  apa_table(caption = "Justifications by category")
+```
+
+For RQ4, we had `r exp3_descriptives$acceptance_n` participants answering the question.
+Of these `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share` said they accepted the scientific consensus because they trust scientists on this question, while `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share` said they independently verified the fact.
+`r exp3_descriptives$acceptance_by_reason[["other"]]$share` answered with other "other" and gave an open-ended explanation (see Appendix).
+
+We also asked all `r exp3_descriptives$reason_followup_n` participants who answered that they had independently verified the answer to explain how they did so.
+The open-ended answers are listed in Appendix XX.
+
+(ref:exp3-plot) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
+
+```{r exp3-plot, fig.cap="(ref:exp3-plot)", fig.height= 12, fig.width=12}
+# Plot distributions of knowledge and acceptance and their mean
+
+mean_value_knowledge <- mean(exp3_wide$avg_knowledge)
+max_n <- max(table(exp3_wide$avg_acceptance))
+
+knowledge_mean <- ggplot(exp3_wide, aes(x = avg_knowledge)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average science knowledge", 
+       y = "N")
+
+mean_value <- mean(exp3_wide$avg_acceptance)
+
+acceptance_mean <- ggplot(exp3_wide, aes(x = avg_acceptance)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average consensus acceptance", 
+       y = "N")
+
+# Trust in science plots
+plot_data <- exp3_wide %>% 
+  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
+  summarise(n = n())
+
+trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  # Add black border around tiles for better visibility
+  geom_jitter(inherit.aes = FALSE, data = exp3_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) +  # Scatter points within each tile
+  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average science knowledge", 
+       y = "Trust in science", 
+       fill = "N")
+
+plot_data <- exp3_wide %>% 
+  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
+  summarise(n = n())
+
+trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  
+  geom_jitter(inherit.aes = FALSE, data = exp3_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) + 
+  scale_fill_viridis_c(option = "plasma") + 
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average consensus Acceptance", 
+       y = "Trust in science", 
+       fill = "N") +
+  theme(legend.position = "top")
+
+# Conspiracy Thinking plots
+
+conspiracy_knowledge <- ggplot(exp3_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average science knowledge", 
+       y = "Average conspiracy thinking")
+
+conspiracy_acceptance <- ggplot(exp3_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average consensus acceptance", 
+       y = "Average conspiracy thinking") +
+  theme(legend.position = "top")
+
+# Combine plots
+
+# distributions
+distributions <- ggarrange(knowledge_mean + 
+                             rremove("xlab") , 
+                           acceptance_mean + 
+                             rremove("xlab") +
+                             rremove("ylab"))
+
+# trust
+trust <- ggarrange(trust_knowledge + 
+                             rremove("xlab") +
+                             guides(fill = "none") +
+                             labs(caption = NULL) , 
+                           trust_acceptance + 
+                             rremove("xlab") +
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# conspiracy
+conspiracy <- ggarrange(conspiracy_knowledge, 
+                           conspiracy_acceptance + 
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# common plot
+patchwork <- (distributions / trust / conspiracy) + 
+  plot_annotation(tag_levels = 'A')
+
+patchwork
+```
 
@@
 ```
+
+\clearpage
+
+```{r child = "appendix_exp2.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_exp3.Rmd"}
+```

### 2024-05-31 paper.tex
first write up studies 1 to 3

@@
 \fi
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
   \fi
   % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
   pdftitle={Who mistrust basic science?},
-  pdfauthor={Jan Pfänder1, Lou Kerzreho1, \& Hugo Mercier1},
   pdflang={en-EN},
@@
 \title{Who mistrust basic science?}
-\author{Jan Pfänder\textsuperscript{1}, Lou Kerzreho\textsuperscript{1}, \& Hugo Mercier\textsuperscript{1}}
+\author{\phantom{0}}
 \date{}
@@
 
-\authornote{
-
-HM received funding from the ANR (SCALUP, ANR-21-CE28-0016-01), and from the John Tempelton Foundation (``An Evolutionary and Cultural Perspective on Intellectual Humility via Intellectual Curiosity and Epistemic Deference'').
-The ANR also supported this project through FrontCog (ANR-17-EURE-0017), and PSL (ANR-10-IDEX-0001-02).
-
-The authors made the following contributions. Jan Pfänder: ; Lou Kerzreho: ; Hugo Mercier: .
-
-Correspondence concerning this article should be addressed to Hugo Mercier, . E-mail: \href{mailto:hugo.mercier@gmail.com}{\nolinkurl{hugo.mercier@gmail.com}}
-
-}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Institut Jean Nicod, Département d'études cognitives, ENS, EHESS, PSL University, CNRS, France}
+\affiliation{\vspace{0.5cm}\textsuperscript{} }
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-There is increasingly more talk of defiance towards science and towards experts more generally (for science, see the intro to the stability paper, see also Paresman et al on populism?).
+\section{Introduction}\label{introduction}
+
+There is increasingly more talk of defiance towards science and towards experts more generally.
 In every population, many people only trust science `some of the time', and a sizable minority doesn't trust it.
@@
 
-\hypertarget{levels-of-trust-in-science}{%
-\subsection{Levels of trust in science}\label{levels-of-trust-in-science}}
-
-\hypertarget{why-do-people-distrust-science}{%
-\subsection{Why do people distrust science?}\label{why-do-people-distrust-science}}
+\subsection{Levels of trust in science}\label{levels-of-trust-in-science}
+
+\subsection{Why do people distrust science?}\label{why-do-people-distrust-science}
 
@@
 
-\hypertarget{how-much-science-do-people-know}{%
-\subsection{How much science do people know?}\label{how-much-science-do-people-know}}
+\subsection{How much science do people know?}\label{how-much-science-do-people-know}
 
@@
 
-\hypertarget{first-series-of-experiments}{%
-\section{First series of experiments}\label{first-series-of-experiments}}
-
-The goal is to see how many people really don't accept basic science (in practice, get as close to ceiling in acceptance as we can, first with a generic sample, then see if that still works with an anti-vax sample).
-
-\hypertarget{experiment-1}{%
-\section{Experiment 1}\label{experiment-1}}
+\section{Overview of experiments}\label{overview-of-experiments}
+
+The goal is to see to which extent people do not accept basic science.
+
+\section{Experiment 1}\label{experiment-1}
 
@@
 
-\hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
-
-\hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
+\subsection{Methods}\label{methods}
+
+\subsubsection{Participants}\label{participants}
 
@@
 
-\hypertarget{procedure}{%
-\subsubsection{Procedure}\label{procedure}}
+\subsubsection{Procedure}\label{procedure}
 
@@
 
-\hypertarget{materials}{%
-\subsubsection{Materials}\label{materials}}
-
-\hypertarget{science-knowledge-and-acceptance}{%
-\paragraph{Science knowledge and acceptance}\label{science-knowledge-and-acceptance}}
+\subsubsection{Materials}\label{materials}
+
+\paragraph{Science knowledge and acceptance}\label{science-knowledge-and-acceptance}
 
@@
 
-\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{1em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{15em}}
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{1em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{23em}}
 \caption{\label{tab:knowledge}Science knowledge items}\\
 \toprule
-id & Question & Scientific consensus & Reference(s)\\
+id & Question & Scientific consensus (Study 1) & Explanation (Study 2 \& 3)\\
 \midrule
@@
 & There is a consensus among scientists that antibiotics kill bacteria, but not viruses.
-& (Sturgis \& Allum, 2004), (Miller, 2004), (Miller, 1998) (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat, 2011), (Pardo \& Calvo, 2004), (Hayes \& Tariq, 2000), (Committee et al., 2016)\\
+& Antibiotics specifically target and kill bacteria, not viruses.
+Scientists know this thanks to extensive laboratory experiments and clinical trials where antibiotics have been observed to be effective against bacterial infections but not viral ones.
+Scientists have also studied how antibiotics work, which typically involve disrupting bacterial cell processes like cell wall synthesis or protein production, which are absent in viruses.
+
+Here are some resources confirming that antibiotics only kill bacteria: 
+American Chemical Society
+Queensland Government 
+Wikipedia\\
 2 & Are electrons smaller, larger, or the same size as atoms ?
 & There is a consensus among scientists that electrons are smaller than atoms.
-& (Sturgis \& Allum, 2004), (Miller, 2004), (Miller, 1998), (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat 2011), (Pardo \& Calvo, 2004), Committee et al., 2016\\
+& Electrons are much smaller than atoms.
+Scientists know this thanks to experiments in particle physics, such as electron scattering experiments.
+They are also able to measure the electron charge and mass.
+Additionally, atomic theory and quantum mechanics provide a theoretical framework explaining the structure of atoms, where electrons orbit the nucleus in distinct energy levels.
+
+Here are some resources confirming that electrons are smaller than atoms:
+Science Focus 
+Britannica 
+Wikipedia\\
 3 & Have the continents on Earth been moving for millions of years or have they always been where they are now?
 & There is a consensus among scientists that the continents on Earth have been moving for millions of years due to plate tectonics.
-& (Miller, 2004) and (Miller, 1998), (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989)
-(Gauchat, 2011), (Pardo \& Calvo, 2004)\\
+& The continents on Earth have been moving for millions of years due to plate tectonics.
+Scientists have gathered evidence from various sources, including fossil records, geological formations, and the magnetic properties of rocks.
+The theory of plate tectonics explains how the Earth's lithosphere is divided into large, rigid plates that move over the semi-fluid asthenosphere, leading to phenomena like continental drift, earthquakes, and volcanic activity.
+
+Here are some resources confirming that continents on Earth have been moving for millions years:
+Britannica 
+Live Science 
+National Geographic\\
 4 & What decides whether a baby is a boy or a girl ?
@@
 & There is a consensus among scientists that it is the genes in the father's sperm which are decisive on whether a baby is a boy or a girl.
-& (Sturgis \& Allum, 2004), (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat, 2011), (Pardo \& Calvo, 2004), (Committee et al., 2016)\\
+& Chromosomes are structures found in the nucleus of cells that carry long pieces of DNA.
+Two of the chromosomes (the X and the Y chromosome) are called sex chromosomes.
+In most cases, females have two X chromosomes, while males have one X and one Y chromosome.
+At conception, the mother transmits an X chromosome, and the father may contribute an X or a Y.
+The chromosome from the father therefore determines if the baby is female (if the father transmits an X) or male (if the father transmits a Y).
+In most cases, children are born a male or female according to their chromosomes.
+However, some children may be born with genitalia that do not match their chromosomes.
+
+Here are some resources confirming that it is the father's genes that decide the sex of a baby: 
+National Library of Medicine 
+Nature 
+Wikipedia\\
 5 & Do lasers work by focusing sound waves?
 & There is a consensus among scientists that lasers do not work by focusing sound waves.
-& (Sturgis \& Allum, 2004), (Miller, 2004) and (Miller, 1998), (Evans \& Durant, 1995) and (Durant et al., 1989), (Tourangeau et al., 2016), (Gauchat , 2011), (Pardo \& Calvo, 2004), (Committee et al., 2016)\\
-\addlinespace
-6 & How long does it take for Earth to go around the sun : one day, one month, or one year ?
+& Lasers produce a narrow beam of light in which all of the particles of light have very similar wavelengths.
+The laser’s lightwaves travel together with their peaks all lined up, in phase.
+This is why laser beams are very narrow, very bright, and can be focused into a very tiny spot.
+Because laser light stays focused and does not spread out much (like a flashlight would), laser beams can travel very long distances.
+
+Here are some resources confirming that lasers do not work by focusing sound waves: 
+NASA 
+Lawrence Livermore National Library
+Wikipedia\\
+\addlinespace
+6 & How long does it take for Earth to go around the sun: one day, one month, or one year ?
 & There is a consensus among scientists that it takes one year for Earth to go around the sun.
-& (Tourangeau et al., 2016), (Evans \& Durant, 1995) and (Durant et al., 1989), (Gauchat, 2011)\\
-7 & Are diamonds made of carbon ? & There is a consensus among scientists that diamonds are made of carbon. & (Evans \& Durant, 1995) and (Durant et al., 1989)\\
+& Earth takes approximately one year to orbit the sun.
+This knowledge is based on astronomical observations and measurements of celestial motion.
+Early astronomers tracked the movement of the Earth relative to the stars and planets, leading to the development of models that accurately predict the Earth's orbital period around the sun.
+
+Here are some resources confirming that Earth goes around the sun in one year: 
+NASA
+National Geographic 
+Wikipedia\\
+7 & Are diamonds made of carbon ?
+& There is a consensus among scientists that diamonds are made of carbon.
+& Diamonds are made of only one element, carbon.
+Carbon is the same element that makes coal or graphite used for pencils.
+Why are diamonds transparent and hard while coal and graphite are opaque and soft?
+It all comes down to the placement of their atoms.
+In diamonds, each carbon atom is bonded to 4 other carbon atoms, while in graphite, each atom is only bonded to 3 other carbon atoms.
+The bonds in diamonds are held in such a tight structure that all light passes around them, which is why diamonds look transparent.
+In coal and graphite, light gets trapped between the atoms, which is why they look dark and opaque.
+Why do carbon atoms bond differently in diamonds?
+At very high pressures and temperatures, the carbon atoms are squeezed so much that they start touching more atoms.
+When the pressure is about 50,000 times the pressure at the surface of the Earth and the temperature is about 1600°C, the carbon atoms bond with 4 other atoms and result in diamonds.
+
+Here are some resources confirming that diamonds are made of carbon: 
+Arizona State University 
+University of Bristol 
+Wikipedia\\
 8 & Which travels faster : light or sound?
 & There is a consensus among scientists that light travels faster than sound.
-& (Miller, 2004) and (Miller, 1998), (Evans \& Durant, 1995) and (Durant et al., 1989)\\
+& Light travels faster than sound.
+This conclusion arises from numerous experiments and observations in physics.
+You can also experience this for yourself during a thunderstorm.
+Unless the lighting is right above you, you will first see the lighting, then hear the thunder some time after (you can even tell how far the lighting struck by counting how many seconds it takes for the thunder to reach you).
+More precisely, the speed of light has been measured accurately using techniques such as the Michelson–Morley experiment.
+Similarly, the speed of sound has been measured through experiments involving the propagation of sound waves through various materials.
+
+Here are some resources confirming that light travers faster than sound: 
+NASA 
+Wikipedia : Speed of light 
+Wikipedia: Speed of sound\\
 9 & Is common table salt made of calcium carbonate?
 & There is a consensus among scientists that common table salt is not made of calcium carbonate; it is made of sodium chloride.
-& (Evans \& Durant, 1995) and (Durant et al., 1989)\\
-10 & Where do trees mainly draw the materials with which they create their mass?
+& Common table salt, or sodium chloride (NaCl), is not made of calcium carbonate.
+Scientists can use a variety of tests to understand what matter is made of.
+The most used test for sodium chloride is the chemical reaction with silver nitrate.
+The presence of sodium chloride yields a white precipitate upon drops of a silver nitrate solution.
+Calcium carbonate, by contrast, is the substance that makes up, for instance, chalk.
+
+Here are some resources confirming that common table salt is not made of calcium carbonate: 
+National Library of Medicine 
+Chem Europe 
+Wikipedia\\
+10 & Is water made of molecules containing one oxygen and two hydrogen atoms?
+& There is a consensus among scientists that water is made of molecules containing one oxygen and two hydrogen atoms, and that its chemical formula is therefore H2O.
+& Water is made of molecules containing one oxygen atom and two hydrogen atoms, chemically represented as H2O.
+We know this for instance because if you set hydrogen on fire in a container with oxygen, water will form on the sides of the container.
+More recent techniques such as spectroscopy and X-ray crystallography have allowed scientists to more directly see the composition and structure of water molecules, confirming the presence of two hydrogen atoms bonded to one oxygen atom.
+
+Here are some resources confirming that water is made of molecules containing one oxygen and two hydrogen atoms:
+National Library of Medicine 
+Wikipedia: Chemical structure 
+Wikipedia: Water\\
+\addlinespace
+11 & Where do trees mainly draw the materials with which they create their mass?
 & There is a consensus among scientists that carbon drawn from the air during photosynthesis makes up most of the materials that trees use to build new leaves, stems, and roots.
-& https://www.canr.msu.edu/news/where\_do\_trees\_get\_their\_mass\_from;
-https://www.tenereteam.com/blogs/where-do-trees-get-their-mass/;
-https://askabiologist.asu.edu/recipe-plant-growth\\
-\addlinespace
-11 & Is water made of molecules containing one oxygen and two hydrogen atoms?
-& There is a consensus among scientists that water is made of molecules containing one oxygen and two hydrogen atoms, and that its chemical formula is therefore H2O.
-& https://en.wikipedia.org/wiki/Water\\
+& \\
 \bottomrule
 \end{longtable}
-
-\hypertarget{conspiracy-scales}{%
-\paragraph{Conspiracy scales}\label{conspiracy-scales}}
+\endgroup{}
+
+\paragraph{Conspiracy scales}\label{conspiracy-scales}
 
@@
 
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
 \caption{\label{tab:conspiracy}Conspiracy items}\\
@@
 \end{longtable}
+\endgroup{}
 
@@
 
-\hypertarget{trust-in-science}{%
-\paragraph{Trust in science}\label{trust-in-science}}
+\paragraph{Trust in science}\label{trust-in-science}
 
@@
 
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
+\subsection{Results}\label{results}
 
@@
 
-For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29), and acceptance of scientific consensus and trust in science ( r = 0.27).
+For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29, p \textless{} .001), and acceptance of scientific consensus and trust in science (r = 0.27, p \textless{} .001).
 The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
@@
 
-For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = -0.39), and conspiracy thinking and acceptance of scientific consensus ( r = -0.33).
-
-In Appendix \ref{exp1} we show that these results hold for our alternative measures of trust and conspiracy thinking.
-We also include more descriptive statistics, such as knowledge and acceptance by science questions.
+For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = -0.38, p \textless{} .001), and conspiracy thinking and acceptance of scientific consensus (r = -0.33, p \textless{} .001).
+
+In Appendix \ref{exp1}, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \ref{exp1} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
 
@@
 We than calculated the correlation between trust in science and average acceptance rate by participant.
-We find a slightly smaller negative correlation of acceptance with for conspiracy thinking (r = -0.14), and only a very small correlation for trust in science (r = 0.06).
-These associations are not statistically significant (see Appendix \ref{exp1}).
+We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = -0.14, p = 0.061), nor with trust in science (r = 0.06, p = 0.387).
 
@@
 
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
-
-These results suggest that even when people do not know the answer to science questions, they tend to mostly accept the scientific consensus.
+\subsection{Discussion}\label{discussion}
+
+These results suggest that most people accept the scientific consensus most of the time.
+Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
 Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
 In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-However, conditioning our analyses to initial false knowledge responses, i.e.~looking at change of opinion towards the scientific consensus, we find that trust in science is onl
+
+\section{Experiment 2}\label{experiment-2}
+
+In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
+In most instances they did, but not always.
+In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
+To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
+To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
+We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \ref{exp1}).
+
+Based on experiment 1, we formulated the following hypotheses:
+
+\textbf{H1a: Higher trust in science is associated with more science knowledge?}
+
+\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?}
+
+\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
+
+\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?}
+
+We had the following research questions:
+
+\textbf{RQ1: What is the average science knowledge score?}
+
+\textbf{RQ2: When a participant's answer does not match the consensus, how often do they change their mind and accept the consensus?}
+
+\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
+
+\subsection{Methods}\label{methods-1}
+
+\subsubsection{Participants}\label{participants-1}
+
+We recruited 201 participants from the US via prolific.
+11 participants failed our attention check, resulting in a final sample of 190 participants (96 female, 94 male; \(age_\text{mean}\): 43.48, \(age_\text{sd}\): 12.25, \(age_\text{median}\): 42).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+\subsubsection{Procedure}\label{procedure-1}
+
+The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
+\ref{fig:exp2-stimulus-example}.
+
+\begin{figure}
+
+\includegraphics[width=1\linewidth]{./figures/study2_question_example} \hfill{}
+
+\caption{Example of a science question, the scientific consensus and corresponding explanation and sources.}\label{fig:exp2-stimulus-example}
+\end{figure}
+
+\subsubsection{Materials}\label{materials-1}
+
+We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
+
+\subsection{Results}\label{results-1}
+
+As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = 0.28, p \textless{} .001) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = -0.40, p \textless{} .001).
+By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
+For trust in science, we find no statistically significant correlation (r = 0.16, p = 0.051).
+For conspiracy thinking we find a small negative one (r = -0.22, p = 0.006).
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-conditional-acceptance.pdf}
+\caption{\label{fig:exp2-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp2-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+In Appendix \ref{exp2} we show that these results hold for our alternative measures of trust and conspiracy thinking.
+We also include more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Regarding RQ1, participants answered on average 79 \% (sd = 0.19) of the questions correctly, and accepted the scientific consensus on average for 98 \% (sd = 0.05) of the questions.
+Fig.
+\ref{fig:exp2-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In response to RQ2, in most cases (92.9 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In very few cases (0.5 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-plot.pdf}
+\caption{\label{fig:exp2-plot}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
+\end{figure}
+
+For RQ3, we got 35 answers from 25 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \ref{tab:exp2-justifications} summarizes these answers by five categories.
+All answers can be read in Appendix \ref{exp2}.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-justifications}Justifications by category}
+
+\begin{tabular}{llll}
+\toprule
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+\midrule
+Personal convictions & 12.00 & 34.3\% & 8.00\\
+Mistake & 9.00 & 25.7\% & 8.00\\
+No justification & 7.00 & 20\% & 5.00\\
+Not convinced & 5.00 & 14.3\% & 5.00\\
+Religious Beliefs & 2.00 & 5.7\% & 2.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Discussion}\label{discussion-1}
+
+Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.~answered the knowledge question wrongly).
+The share of these latter is considerably larger in experiment 2 (92.9 \%) than in experiment 1 (76.3 \%).
+While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
+We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
+
+\section{Experiment 3}\label{experiment-3}
+
+Study 3 is essentially a replication-\/--with some minor modifications-\/--of study 2, but on a different type of sample.
+Both study 1 and 2 were run on convenience samples.
+For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
+By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
+\ref{fig:exp3-explanation-example}).
+Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
+Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
+We want to know if participants perceive that this is because of trust, or other factors.
+
+As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
+
+\textbf{H1a: Higher trust in science is associated with more science knowledge?}
+
+\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus.}
+
+\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
+
+\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.}
+
+We had the following research questions:
+
+\textbf{RQ1: What is the average science knowledge score?}
+
+\textbf{RQ2: What is the average acceptance of the scientific consensus}
+
+\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
+
+\textbf{RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?}
+
+\subsection{Methods}\label{methods-2}
+
+\subsubsection{Participants}\label{participants-2}
+
+We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of 200 participants (125 female, 73 male; \(age_\text{mean}\): 42.90, \(age_\text{sd}\): 12.00, \(age_\text{median}\): 41).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
+This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = 140 for trust in science, N = 138 for conspiracy measures, N = 122 for reason for accepting consensus).
+
+\subsubsection{Procedure}\label{procedure-2}
+
+The procedure was mostly the same as in experiment 2.
+In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
+Finally, at the end of the survey, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
+Participants who selected ``You mostly agree with the consensus because you have been able to independently verify it'', were asked the open-ended follow-up question: ``Could you please tell us how you independently verified the information?''.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{figures/study3_example_explanation}
+
+}
+
+\caption{Example of an explanation question and the opportunity to change the previous answer.}\label{fig:exp3-explanation-example}
+\end{figure}
+
+\subsubsection{Materials}\label{materials-2}
+
+We relied on the same items as in experiment 2.
+We used a pre-registered categorization scheme for the open ended answers for why people reject the consensus.
+The categories were XX and XX.
+{[}Describe coding process, with independent reviewer etc.{]}
+
+\subsection{Results}\label{results-2}
+
+As in study 1 and 2, we find that participants answered on average 75 \% (sd = 0.18) of the questions correctly, and initially accepted the scientific consensus on average for 96 \% (sd = 0.08) of the questions (RQ2).
+The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (98 \%, sd = 0.06).
+
+Fig.
+\ref{fig:exp3-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In most cases (87 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
+After providing the chance to revise the initial consensus rejection, this share is even larger (95.5 \%).
+In very few cases (0.9 \%), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
+This share drops slightly after providing the chance to revise the initial consensus rejection (0.7 \%).
+
+In the appendix XX we provide an analysis that this drop is statistically significant {[}TO DO{]}.
+
+For all correlations, we include opinion revisions for measuring consensus acceptance.
+We find no statistically significant correlation between science knowledge and trust in science (r = 0.14, p = 0.100), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = 0.20, p = 0.019).
+Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp3-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+We find no statistically significant correlation between conspiracy thinking and science knowledge (r = -0.16, p = 0.055), and between conspiracy thinking and acceptance of scientific consensus (r = c(cor = -0.023), c(t = -0.27), = 0.788, c(df = 136), -0.189, 0.145, Pearson's product-moment correlation, two.sided, {[}-0.189, 0.145{]}, p = 0.788).
+
+In Appendix \ref{exp3}, we show that these associations are statistically significant (except the last one?).
+We show that they are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \ref{exp3} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-conditional-acceptance.pdf}
+\caption{\label{fig:exp3-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+Regarding RQ1, participants answered on average 75 \% (sd = 0.18) of the questions correctly.
+For RQ3, we got 74 answers from 47 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \ref{tab:exp3-justifications} summarizes these answers by five categories.
+
+All answers can be read in the appendix.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp3-justifications}Justifications by category}
+
+\begin{tabular}{llll}
+\toprule
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+\midrule
+Personal convictions & 25.00 & 33.8\% & 19.00\\
+Mistake & 21.00 & 28.4\% & 18.00\\
+No justification & 18.00 & 24.3\% & 13.00\\
+Not convinced & 7.00 & 9.5\% & 3.00\\
+Religious Beliefs & 3.00 & 4.1\% & 3.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+For RQ4, we had 122 participants answering the question.
+Of these 41.8\% said they accepted the scientific consensus because they trust scientists on this question, while 47.5\% said they independently verified the fact.
+10.7\% answered with other ``other'' and gave an open-ended explanation (see Appendix).
+
+We also asked all 58 participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed in Appendix XX.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-plot.pdf}
+\caption{\label{fig:exp3-plot}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
+\end{figure}
 
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
+\section{References}\label{references}
+
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-bruderMeasuringIndividualDifferences2013}{}}%
+\bibitem[\citeproctext]{ref-bruderMeasuringIndividualDifferences2013}
 Bruder, M., Haffke, P., Neave, N., Nouripanah, N., \& Imhoff, R.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-colognaTrustScientistsTheir2024}{}}%
+\bibitem[\citeproctext]{ref-colognaTrustScientistsTheir2024}
 Cologna, V., Mede, N.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lantianMeasuringBeliefConspiracy2016}{}}%
+\bibitem[\citeproctext]{ref-lantianMeasuringBeliefConspiracy2016}
 Lantian, A., Muller, D., Nurra, C., \& Douglas, K.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}{}}%
+\bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
 Pennycook, G., Binnendyk, J., \& Rand, D.
@@
 
-\hypertarget{appendix-appendix}{%
-\appendix}
-
-\hypertarget{exp1}{%
-\section{Experiment 1}\label{exp1}}
-
-\hypertarget{materials-1}{%
-\subsection{Materials}\label{materials-1}}
+\appendix
+
+\section{Experiment 1}\label{exp1}
+
+\subsection{Materials}\label{materials-3}
 
@@
 
-\hypertarget{conspiracy-scales-1}{%
-\subsubsection{Conspiracy scales}\label{conspiracy-scales-1}}
+\subsubsection{Conspiracy scales}\label{conspiracy-scales-1}
 
@@
 
-\hypertarget{trust-in-science-1}{%
-\subsubsection{Trust in science}\label{trust-in-science-1}}
+\subsubsection{Trust in science}\label{trust-in-science-1}
 
@@
 
-\hypertarget{comparing-items}{%
-\subsection{Comparing items}\label{comparing-items}}
-
-Make correlation table
-
-\hypertarget{conspiracy-theories}{%
-\subsubsection{Conspiracy theories}\label{conspiracy-theories}}
+\subsection{Comparing items}\label{comparing-items}
+
+\subsubsection{Conspiracy theories}\label{conspiracy-theories}
 
@@
 
-\hypertarget{trust-in-science-2}{%
-\subsubsection{Trust in science}\label{trust-in-science-2}}
+\subsubsection{Trust in science}\label{trust-in-science-2}
 
@@
 
-\hypertarget{correlations-with-alternative-measures}{%
-\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures}}
+\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures}
 
@@
 BCTI 
-(main conspiracy measure) & -0.39 & -0.33\\
-CMQ & -0.11 & -0.07\\
-SICBS & -0.10 & -0.09\\
-WGM trust scientists & 0.23 & 0.26\\
+(main conspiracy measure) & -0.38 (p< .001) & -0.33 (p< .001)\\
+CMQ & -0.11 (p= 0.121) & -0.07 (p= 0.317)\\
+SICBS & -0.1 (p= 0.158) & -0.09 (p= 0.213)\\
+WGM trust scientists & 0.23 (p= 0.001) & 0.26 (p< .001)\\
 WGM trust general 
-(main trust measure) & 0.29 & 0.27\\
-PEW trust scientists & 0.16 & 0.21\\
+(main trust measure) & 0.29 (p< .001) & 0.27 (p< .001)\\
+PEW trust scientists & 0.16 (p= 0.027) & 0.21 (p= 0.003)\\
 \bottomrule
@@
 
-\hypertarget{results-conditional-on-false-responses}{%
-\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses}}
+\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses}
 
@@
 
+\subsection{By-question variation}\label{by-question-variation}
+
+\subsubsection{Knowledge}\label{knowledge}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-questions-knowledge.pdf}
+\caption{\label{fig:exp1-questions-knowledge}Distribution of correct answers by question.}
+\end{figure}
+
+\subsubsection{Acceptance}\label{acceptance}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-questions-acceptance.pdf}
+\caption{\label{fig:exp1-questions-acceptance}Distribution of consensus acceptance by question.}
+\end{figure}
+
+\subsection{Trust in science}\label{trust-in-science-3}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-trust-scientists.pdf}
+\caption{\label{fig:exp1-trust-scientists}Distribution of trust in scientists.}
+\end{figure}
+
+\subsection{Conspiracy thinking}\label{conspiracy-thinking}
+
+\subsubsection{Distribution}\label{distribution}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-conspiracy-distribution.pdf}
+\caption{\label{fig:exp1-conspiracy-distribution}Distribution of conspiracy thinking.}
+\end{figure}
+
+\subsubsection{By-item variation}\label{by-item-variation}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-conspiracy-items.pdf}
+\caption{\label{fig:exp1-conspiracy-items}Distribution of conspiracy thinking by item.}
+\end{figure}
+
+\subsection{Open ended comments}\label{open-ended-comments}
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
+\caption{\label{tab:unnamed-chunk-29}Reasons for consensus rejection}\\
+\toprule
+id & comments\\
+\midrule
+1 & no\\
+20 & None\\
+22 & I answered thoughtfully, good luck!\\
+27 & interesting, thank you.\\
+30 & n/a\\
+\addlinespace
+32 & None\\
+39 & no\\
+41 & none\\
+42 & n/a\\
+67 & thanks\\
+\addlinespace
+69 & No comments\\
+74 & no\\
+76 & none\\
+78 & None - thank you!\\
+79 & Should've clicked agree about plants getting carbon from the air, but I wanted to look it up first.
+Whenever I encounter something I didn't know online, I like to cross-reference.
+In retrospect though, everything else you asked about was simple and on the level, and I'm sure you are correct.
+
+Thanks for your hard work.\\
+\addlinespace
+84 & No, thank you.\\
+86 & n/a\\
+89 & N/A\\
+95 & no\\
+98 & Great topics! Thank you for this study.\\
+\addlinespace
+109 & To provide context for my responses, I currently study physiology (STEM) and perform academic research as an undergrad at my university.\\
+110 & n/a\\
+114 & Everything in the survey went fine with no issues.
+We seem to have a lot of conspiracy theory nuts in the world today and I think it's a sign of the larger mental health crisis that's currently going on.\\
+115 & I like the subject content of the survey\\
+116 & No\\
+\addlinespace
+132 & None.\\
+133 & very interesting\\
+135 & Nope\\
+139 & none\\
+140 & Air AND water make up the bulk of trees mass and the water is drawn from the earth.
+Water has much more mass than air and I stand by my answer of Earth being that is where the water is drawn from.
+The question was very vague and as you see...
+open to interpretation\\
+\addlinespace
+141 & n/a\\
+148 & no\\
+150 & no\\
+152 & None.\\
+154 & none, tysm!\\
+\addlinespace
+155 & Not virus, just metals/DDT. And I think it would be impossible for dinosaurs to have sex with the way they are shaped.\\
+157 & none\\
+159 & I think scientists are acting in the best interests of science rather than the community directly. An unethical scientist could advance scientific knowledge while hurting humanity.\\
+161 & Mo Thanks\\
+164 & I appreciate science but understand that if scientists don't go along with what investors/govt want they won't receive grants for funding.
+I also very recently lived through several years of gaslighting about covid.
+While I realize I'm not the sharpest tool in the shed, I'm not a total idiot.
+Things that were censored at the beginning of the pandemic are now allowed to be discussed freely.
+Science should be publicly debated otherwise it isn't science.\\
+\addlinespace
+165 & it was interesting\\
+166 & Everything in the study was good I enjoyed taking it\\
+167 & none\\
+169 & good study\\
+170 & Thanks for the opportunity!\\
+\addlinespace
+173 & No\\
+180 & The study was great.\\
+183 & Very interesting study.\\
+185 & None\\
+186 & Thanks, have  good day\\
+\addlinespace
+189 & no\\
+\bottomrule
+\end{longtable}
+
+\clearpage
+
+\section{Experiment 2}\label{exp2}
+
+\subsection{Materials}\label{materials-4}
+
+\FloatBarrier
+
+See Appendix \ref{exp1}.
+
+\subsection{Comparing items}\label{comparing-items-1}
+
+\subsubsection{Conspiracy theories}\label{conspiracy-theories-1}
+
+Table \ref{tab:exp2-correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
+\midrule
+BCTI & 1.00 & 0.66 & 0.63\\
+CMQ & 0.66 & 1.00 & 0.86\\
+SICBS & 0.63 & 0.86 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsubsection{Trust in science}\label{trust-in-science-4}
+
+Table \ref{tab:exp2-correlation-trust} shows the correlations of the three different items measuring trust in science.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-correlation-trust}Correlations of the three different items measuring trust in science}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
+\midrule
+wgm\_sciencegeneral & 1.00 & 0.89 & 0.79\\
+wgm\_scientists & 0.89 & 1.00 & 0.85\\
+pew & 0.79 & 0.85 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures-1}
+
+Table \ref{tab:exp2-correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}
+
+\begin{tabular}{lllllllll}
+\toprule
+std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+0.07 & -6.01 & < .001 & -0.53 & -0.27 & BCTI 
+(main conspiracy measure) & {}[-0.533, -0.27] & -0.40 & NA\\
+0.07 & -2.36 & = 0.019 & -0.31 & -0.03 & CMQ & {}[-0.311, -0.028] & -0.17 & NA\\
+0.07 & -1.91 & = 0.058 & -0.28 & 0.00 & SICBS & {}[-0.28, 0.005] & -0.14 & NA\\
+0.07 & 3.20 & = 0.002 & 0.09 & 0.37 & WGM trust scientists & {}[0.087, 0.368] & 0.23 & NA\\
+0.07 & 3.98 & < .001 & 0.14 & 0.42 & WGM trust general 
+(main trust measure) & {}[0.14, 0.417] & 0.28 & NA\\
+0.07 & 2.23 & = 0.027 & 0.02 & 0.30 & PEW trust scientists & {}[0.019, 0.303] & 0.16 & NA\\
+0.07 & -5.47 & < .001 & -0.50 & -0.24 & BCTI 
+(main conspiracy measure) & {}[-0.504, -0.237] & NA & -0.37\\
+0.07 & -3.03 & = 0.003 & -0.36 & -0.08 & CMQ & {}[-0.356, -0.076] & NA & -0.22\\
+0.07 & -2.76 & = 0.006 & -0.34 & -0.06 & SICBS & {}[-0.338, -0.056] & NA & -0.20\\
+0.07 & 3.51 & < .001 & 0.11 & 0.39 & WGM trust scientists & {}[0.108, 0.387] & NA & 0.25\\
+0.07 & 4.29 & < .001 & 0.16 & 0.44 & WGM trust general 
+(main trust measure) & {}[0.161, 0.436] & NA & 0.30\\
+0.07 & 3.04 & = 0.003 & 0.08 & 0.36 & PEW trust scientists & {}[0.076, 0.357] & NA & 0.22\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses-1}
+
+Table \ref{tab:exp2-false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.
+
+\begin{table}
+
+\caption{\label{tab:exp2-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
+\centering
+\begin{tabular}[t]{lcccccc}
+\toprule
+  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
+\midrule
+(Intercept) & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
+ & (0.081) & (0.080) & (0.081) & (0.082) & (0.082) & \vphantom{1} (0.082)\\
+avg\_acceptance & -0.225** & -0.247** & -0.201* & 0.131 & 0.161+ & 0.130\\
+ & (0.081) & (0.080) & (0.081) & (0.082) & (0.082) & (0.082)\\
+\midrule
+Num.Obs. & 147 & 147 & 147 & 147 & 147 & 147\\
+R2 & 0.051 & 0.061 & 0.041 & 0.017 & 0.026 & 0.017\\
+R2 Adj. & 0.044 & 0.054 & 0.034 & 0.010 & 0.019 & 0.010\\
+AIC & 414.5 & 412.9 & 416.1 & 419.6 & 418.3 & 419.7\\
+BIC & 423.5 & 421.9 & 425.0 & 428.6 & 427.3 & 428.6\\
+Log.Lik. & -204.248 & -203.465 & -205.036 & -206.802 & -206.141 & -206.826\\
+RMSE & 0.97 & 0.97 & 0.98 & 0.99 & 0.98 & 0.99\\
+\bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
+\end{tabular}
+\end{table}
+
+\subsection{By-question variation}\label{by-question-variation-1}
+
+\subsubsection{Knowledge}\label{knowledge-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-questions-knowledge.pdf}
+\caption{\label{fig:exp2-questions-knowledge}Distribution of correct answers by question.}
+\end{figure}
+
+\subsubsection{Acceptance}\label{acceptance-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-questions-acceptance.pdf}
+\caption{\label{fig:exp2-questions-acceptance}Distribution of consensus acceptance by question.}
+\end{figure}
+
+\subsection{Trust in science}\label{trust-in-science-5}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-trust-scientists.pdf}
+\caption{\label{fig:exp2-trust-scientists}Distribution of trust in scientists.}
+\end{figure}
+
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-1}
+
+\subsubsection{Distribution}\label{distribution-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-conspiracy-distribution.pdf}
+\caption{\label{fig:exp2-conspiracy-distribution}Distribution of conspiracy thinking.}
+\end{figure}
+
+\subsubsection{By-item variation}\label{by-item-variation-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-conspiracy-items.pdf}
+\caption{\label{fig:exp2-conspiracy-items}Distribution of conspiracy thinking by item.}
+\end{figure}
+
+\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection}
+
+\subsubsection{By category}\label{by-category}
+
+{[}To do{]}
+
+\subsubsection{All raw answers}\label{all-raw-answers}
+
+\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp2-reasons-rejection}Reasons for consensus rejection}\\
+\toprule
+id & question & answer\\
+\midrule
+33 & sex & Because I didn't think the information provided here was accurate, more or less.\\
+45 & lasers & I read the article.\\
+45 & salt & Read the article provided.\\
+47 & lasers & The statement "lasers do not work by focusing sound waves" is correct according to scientific consensus.
+Lasers operate on the principles of stimulated emission of radiation, wherein photons are emitted in a coherent, focused beam through the process of stimulated emission.\\
+47 & diamonds & Are diamonds made of carbon? Scientific consensus: Diamonds are made of carbon.  Do you think you could explain why you answered differently?\\
+\addlinespace
+47 & water & it likely arose from a mistake or miscommunication.\\
+50 & orbit & I honestly couldn't remember.\\
+50 & speed & When I looked it up I got two different answers.\\
+71 & water & After I put "No", I realized that I was thinking about "H2o" incorrectly. So yes, I actually do agree.\\
+73 & electrons & Atoms are known to be the smallest unit of matter.\\
+\addlinespace
+77 & continents & I believe God make it they way it is and if He wanted it changed, He would have done it\\
+78 & sex & no\\
+78 & salt & no\\
+79 & orbit & It’s early and I’m tired. I misread it.\\
+105 & salt & I think I simply misunderstood the paragraph. I'm actually trying to read it thoroughly and answering truthfully, but I have misread it.\\
+\addlinespace
+109 & electrons & Maybe\\
+127 & continents & I'm a young earth Christian. I believe the Christian Bible (and outside sources) offers ample evidence that the earth is thousands of years old, not millions.\\
+127 & orbit & I do not believe in the heliocentric model of an earth spinning around a centralized sun.\\
+130 & continents & I don't think the continent move rather it is the atmosphere\\
+130 & sex & I think that the gene both parent release should determine the sex of the baby\\
+\addlinespace
+131 & antibiotics & There are some antibiotics that also kill viruses\\
+132 & salt & Because it says that table salt is not made of calcium carbonate, so I clicked no.\\
+133 & lasers & I feel like the excerpt that was provided did not definitively state the answer.\\
+140 & lasers & It was confusing to understand the explanation of why lasers don’t focus on sound\\
+152 & continents & I think that they were that way after the ice age.
+It split apart the whole surface and then water filled in and they have remained the same for ever.
+They are not moving around or have they for millions of years.\\
+\addlinespace
+152 & orbit & The  Earth circles the sun every day.  The dark side is when the moon comes out.  So everyday, you have daylight and dark (moon).  Dark comes from the dark side of the sun.\\
+170 & lasers & I may have misread, but I did not see confirmation of this.\\
+172 & salt & no\\
+178 & electrons & This answer is different because atoms is everywhere and you can't see it but electrons aren't out that as much at all but you could possibly know it's there\\
+178 & antibiotics & Because antibiotics can kill bacteria in which kills the virus as well in the process so you get healed with this product\\
+\addlinespace
+178 & sex & There is no 100\% evidence that this is the case and I still fully believe that it is from both genes to make this decision.\\
+179 & antibiotics & Antibiotics seem that if they are capable of killing tiny things such as bacteria then they are most likely to attach to other things in out bodies and cause it harm although unintentional.\\
+183 & salt & I thought the scientific consensus was right\\
+186 & sex & The scientific consensus is clear: both parents' genes determine the sex of a baby, specifically the sex chromosomes they contribute.\\
+189 & salt & I made an error in my selection.\\
+\bottomrule
+\end{longtable}
+
+\clearpage
+
+\section{Experiment 3}\label{exp3}
+
+\subsection{Materials}\label{materials-5}
+
+\FloatBarrier
+
+See Appendix \ref{exp1}.
+
+\subsection{Comparing items}\label{comparing-items-2}
+
+\subsubsection{Conspiracy theories}\label{conspiracy-theories-2}
+
+Table \ref{tab:exp3-correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp3-correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
+\midrule
+BCTI & 1.00 & NA & NA\\
+CMQ & NA & 1.00 & NA\\
+SICBS & NA & NA & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsubsection{Trust in science}\label{trust-in-science-6}
+
+Table \ref{tab:exp3-correlation-trust} shows the correlations of the three different items measuring trust in science.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp3-correlation-trust}Correlations of the three different items measuring trust in science}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
+\midrule
+wgm\_sciencegeneral & 1.00 & NA & NA\\
+wgm\_scientists & NA & 1.00 & NA\\
+pew & NA & NA & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures-2}
+
+Table \ref{tab:exp3-correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp3-correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}
+
+\begin{tabular}{lllllllll}
+\toprule
+std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+0.08 & -1.93 & = 0.055 & -0.32 & 0.00 & BCTI 
+(main conspiracy measure) & {}[-0.317, 0.004] & -0.16 & NA\\
+0.08 & 0.10 & = 0.925 & -0.15 & 0.17 & CMQ & {}[-0.154, 0.17] & 0.01 & NA\\
+0.08 & 0.30 & = 0.767 & -0.14 & 0.19 & SICBS & {}[-0.138, 0.186] & 0.02 & NA\\
+0.08 & 1.31 & = 0.193 & -0.06 & 0.28 & WGM trust scientists & {}[-0.056, 0.276] & 0.11 & NA\\
+0.08 & 1.66 & = 0.100 & -0.03 & 0.30 & WGM trust general 
+(main trust measure) & {}[-0.027, 0.304] & 0.14 & NA\\
+0.08 & -0.91 & = 0.367 & -0.24 & 0.09 & PEW trust scientists & {}[-0.243, 0.09] & -0.08 & NA\\
+0.11 & -0.27 & = 0.788 & -0.25 & 0.19 & BCTI 
+(main conspiracy measure) & {}[-0.246, 0.187] & NA & -0.03\\
+0.11 & -0.08 & = 0.934 & -0.23 & 0.21 & CMQ & {}[-0.226, 0.208] & NA & -0.01\\
+0.11 & 0.02 & = 0.987 & -0.22 & 0.22 & SICBS & {}[-0.215, 0.218] & NA & 0.00\\
+0.07 & 2.72 & = 0.007 & 0.05 & 0.34 & WGM trust scientists & {}[0.054, 0.343] & NA & 0.20\\
+0.07 & 2.37 & = 0.019 & 0.03 & 0.32 & WGM trust general 
+(main trust measure) & {}[0.029, 0.32] & NA & 0.17\\
+0.07 & 1.64 & = 0.104 & -0.03 & 0.27 & PEW trust scientists & {}[-0.025, 0.268] & NA & 0.12\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses-2}
+
+Table \ref{tab:exp3-false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.
+
+\begin{table}
+
+\caption{\label{tab:exp3-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
+\centering
+\begin{tabular}[t]{lcccccc}
+\toprule
+  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
+\midrule
+(Intercept) & -0.001 & 0.000 & 0.000 & 0.005 & 0.002 & 0.008\\
+ & (0.091) & (0.091) & (0.091) & (0.088) & (0.089) & (0.088)\\
+avg\_acceptance & -0.043 & -0.025 & -0.009 & 0.083 & 0.039 & 0.125\\
+ & (0.090) & (0.090) & (0.090) & (0.079) & (0.079) & (0.078)\\
+\midrule
+Num.Obs. & 122 & 122 & 122 & 128 & 128 & 128\\
+R2 & 0.002 & 0.001 & 0.000 & 0.009 & 0.002 & 0.020\\
+R2 Adj. & -0.006 & -0.008 & -0.008 & 0.001 & -0.006 & 0.012\\
+AIC & 351.0 & 351.1 & 351.2 & 367.1 & 368.0 & 365.7\\
+BIC & 359.4 & 359.6 & 359.6 & 375.7 & 376.5 & 374.2\\
+Log.Lik. & -172.491 & -172.571 & -172.604 & -180.560 & -180.996 & -179.843\\
+RMSE & 0.99 & 1.00 & 1.00 & 0.99 & 1.00 & 0.99\\
+\bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
+\end{tabular}
+\end{table}
+
+\subsection{By-question variation}\label{by-question-variation-2}
+
+\subsubsection{Knowledge}\label{knowledge-2}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-questions-knowledge.pdf}
+\caption{\label{fig:exp3-questions-knowledge}Distribution of correct answers by question.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-questions-acceptance.pdf}
+\caption{\label{fig:exp3-questions-acceptance}Distribution of consensus acceptance by question.}
+\end{figure}
+
+\subsection{Trust in science}\label{trust-in-science-7}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-trust-scientists.pdf}
+\caption{\label{fig:exp3-trust-scientists}Distribution of trust in scientists.}
+\end{figure}
+
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-2}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-conspiracy-distribution.pdf}
+\caption{\label{fig:exp3-conspiracy-distribution}Distribution of conspiracy thinking.}
+\end{figure}
+
+\subsubsection{By-item variation}\label{by-item-variation-2}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-conspiracy-items.pdf}
+\caption{\label{fig:exp3-conspiracy-items}Distribution of conspiracy thinking by item.}
+\end{figure}
+
+\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection-1}
+
+\subsubsection{All raw answers}\label{all-raw-answers-1}
+
+\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp3-reasons-rejection}Reasons for consensus rejection}\\
+\toprule
+id & question & answer\\
+\midrule
+5 & lasers & It is what I believed I was taught\\
+12 & antibiotics & I do not know\\
+20 & water & I think I got it confused in my head. I was thinking 2 oxygen and 1 hydrogen.\\
+25 & electrons & I made a mistake.\\
+26 & antibiotics & I have always experienced receiving antibiotic for viral infections. So i do not agree that antibiotic does not kill viruses.\\
+\addlinespace
+32 & antibiotics & Because I realized that my thinking was not correct.\\
+32 & orbit & After thinking about it, I realized  that it does take a year for the sun to go around the earth.\\
+38 & continents & The earth is not millions of years old.\\
+38 & orbit & The earth is flat and stationary.\\
+43 & water & I got it confused. Its true. One oxygen and  2 hydrogen is what i learned in school.\\
+\addlinespace
+46 & sex & Yes.  Its normal to think that, however, not always.\\
+53 & continents & The Earth was created aged by God and has not been around for millions of years. I do not believe in evolution but rather creation based on the Word of God.\\
+61 & continents & I do not believe that the universe is millions of years old\\
+62 & continents & I do not believe the theory that the earth is that old.\\
+65 & electrons & Because I would like to see more research on this matter.\\
+\addlinespace
+65 & antibiotics & Because I would like to see more research on this matter.\\
+65 & continents & Because I would like to see more research on this matter.\\
+65 & diamonds & Because I would like to see more research on this matter.\\
+65 & water & Because I would like to see more research on this matter.\\
+72 & lasers & That’s the truth\\
+\addlinespace
+72 & salt & I wasn’t sure\\
+83 & continents & The Earth has only been around for 6-7 thousand years. As far as I know however, the Continents have been moving.\\
+83 & water & thought it was "2 parts oxide, one hydrogen." I was of course wrong. I also miss clicked the wrong answer for how long it takes the Earth the revolve around the sun... I think.\\
+87 & electrons & I believe atoms are the smallest particles.\\
+87 & continents & On second thought, I agree.\\
+\addlinespace
+87 & diamonds & On second thought, I believe I agree.\\
+87 & water & This does not fit according to my recollection.\\
+92 & salt & I meant to agree, I did not know the answer.\\
+96 & electrons & I don't think electrons are smaller, just because that is what we were taught in science\\
+96 & continents & This is something we are taught academically in science, it may not be true\\
+\addlinespace
+98 & antibiotics & I think there are two different kinds of antibiotics. One type kills bacteria infections and the other type kills viral infections\\
+99 & continents & I don't have a reason to believe Earth has been here for millions of years, so I don't agree with the consensus. Of course, I know I could be wrong.\\
+102 & continents & This information is based in evolution which is a theory.\\
+109 & continents & I partially agree that continents have moved due to plate tectonics.
+But I don't believe the Earth is as old as it mentions in the explanation, and is much younger.
+I also disagree with scientific consensus and there is really no such thing.
+Because with science, when you base things on evidence and then find new evidence, you have to change your theory.
+So there is really no such thing as consensus, there is data, facts and theories.\\
+112 & antibiotics & No\\
+\addlinespace
+112 & speed & No\\
+112 & salt & No\\
+117 & antibiotics & Not really. I thought they killed virus' but maybe I am wrong. heh.\\
+117 & sex & I thought both male and female decided the fate of the babies gender.\\
+122 & antibiotics & antibotics only kill bacteria\\
+\addlinespace
+122 & lasers & I thought they light was a lot faster than sound waves\\
+122 & salt & not sure\\
+129 & continents & YES, I believe in a young earth.  I do not accept the tax supported theory of evolution.\\
+130 & antibiotics & I was confused.\\
+132 & continents & Continental drift is an unproven theory. I have seen no actual evidence for it. Flooding and receding of waters are more likely the cause of coastline changes throughout history.\\
+\addlinespace
+132 & orbit & There is no scientific evidence that the earth moves at all except for earthquakes.
+The sun is clearly moving around the earth in a daily circular pattern.
+The sun is much smaller and closer than we have been taught.\\
+142 & salt & I thought it was made up of that.\\
+143 & lasers & Lasers operate based on the principles of stimulated emission of electromagnetic radiation, typically light.\\
+146 & antibiotics & Different knowledge and experience.\\
+146 & sex & A mothers choice not the fathers choice because of religious beliefs.\\
+\addlinespace
+146 & diamonds & Gained through knowledge and skills.\\
+150 & speed & To be honest, I did actually hear the concept that like travels fashion and sound so I’m not sure why I picked that answer.\\
+151 & lasers & I accidentally clicked the other answer I'm not going to lie. As soon as I clicked it on accident I was like "Sh*t". I meant to choose the other answer on this one, I'm sorry.\\
+151 & salt & I misunderstood this question. This one confused me the most even though it's right in front of me. Sorry.\\
+155 & salt & error\\
+\addlinespace
+158 & lasers & No I did not know this question.\\
+158 & salt & I knew this one and just blanked on the answer.\\
+159 & sex & I thought I read it said in most cases women transmit an X chromosome, but now I'm not sure.\\
+161 & antibiotics & I thought antibiotics kills every bit of poison.\\
+166 & electrons & I'm not sure, I dont' understand it\\
+\addlinespace
+166 & water & I'm not sure, I dont' understand it\\
+171 & diamonds & It just doesn't sit well with me.\\
+172 & sex & I was wrong. I agree with scientific consensus.\\
+173 & orbit & Yes the scientist now more than I do\\
+178 & electrons & I was incorrect on this.\\
+\addlinespace
+191 & antibiotics & I could be wrong, but I thought they killed viruses as well.\\
+191 & continents & I do not believe without a doubt the earth has been around for millions of years.\\
+196 & sex & Answered incorrectly in error.  I agree with the scientific consensus.\\
+197 & antibiotics & I’ve taken antibiotics numerous of times and although my sickness is usually cured, I always get a yeast infection as result.
+This always happens since the antibiotics kill the good and bad bacteria in the body.\\
+197 & sex & I believe that God decides whether parents will be blessed with a boy or girl. It had nothing to do with genetics.\\
+\addlinespace
+197 & diamonds & This just doesn’t seem correct even after viewing the article from the university. I can’t explain why but I can’t justify it.\\
+198 & diamonds & I just never heard of diamonds being made of carbon and it makes zero sense to me while thinking about it.\\
+200 & continents & The earth is not old but young. Not billions or millions but thanks of years.\\
+200 & orbit & One day because the sun is not taking a year to light up other parts of the earth.\\
+\bottomrule
+\end{longtable}
+
+\subsection{Why do people accept the scientific consensus:}\label{why-do-people-accept-the-scientific-consensus}
+
+\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other}
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp3-other-reasons-acceptance}Reasons for consensus rejection}\\
+\toprule
+id & other\_reason\_agreement\\
+\midrule
+24 & some stuff I already knew, other stuff I don't care about like atoms and molecules\\
+37 & I mostly agreed based on my personal knowledge.\\
+62 & I have never seen compelling evidence to disprove the consensus.\\
+89 & Because it is true.\\
+101 & I can't say that I trust scientists on many things, but on most in this survey.
+However, while I believe the continents have moved around, I don't believe the earth is millions of years old, as I'm a Christian.
+Though God never notes exactly how old the earth is, so it could be millions of years old.\\
+\addlinespace
+102 & Agreed due to vague knowledge about the subject.\\
+166 & I'm not sure, you cant trust any party\\
+172 & Because i recall previously being taught that\\
+182 & I agree because the explaination makes sense\\
+188 & I don't trust scientists, but those are the current established 'facts'. Which, on the mundane side of things, are likely true.\\
+\addlinespace
+189 & I take nothing that anyone tells me at face value. Lots of studying and verifying accuracy of the things that I was taught my entire life.\\
+195 & I mostly agree with the consensus because I don't have the equipment to independently test it myself and I also know scientists have proven those who don't have the so called equipment to know what they know.
+We can only observe what we observe at the moment, but there's more to what the scientists are observing and theorizing.\\
+\bottomrule
+\end{longtable}
+
+\subsection{How did participants independently verifiy?}\label{how-did-participants-independently-verifiy}
+
+\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other-1}
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp3-independent-verification}Reasons for consensus rejection}\\
+\toprule
+id & reason\_followup\\
+\midrule
+5 & By trying to remember\\
+10 & Curiosity and the internet\\
+12 & Guessed\\
+14 & I double-checked the information on Google by looking up various websites, I did go with my initial gut instincts in most cases however if it was wrong upon verifying it then I would change my answer\\
+18 & Prior knowledge from a formal education\\
+\addlinespace
+19 & I can check several different sources on the internet and see if they have the same verifiable information.\\
+21 & I use intuition.  I use common sense.\\
+22 & I have read it before, and did some experiments in chemistry class.
+I look for scientific consensus on everything.
+On these long established principles, I could have selected both answers.
+I believe sometimes scientists are bribed or exhibit bias depending on who is funding them and what answers the funding entity wants.
+COVID treatments were a big example of that.
+Hospitals were paid to execute certain protocols and disallow others.\\
+27 & Most of the questions, the consensus answer was the same as my answer.
+For the one or two questions where it was not, I looked at who their sources were for the answers, and decided whether I believed those sources and how unsure I was about my answer and made my decision from there.\\
+31 & Through exposure to scientific experiments and research that prove the claims valid and can be repeated with the same results occurring.\\
+\addlinespace
+36 & There is evidence and a good amount of understand about these issues. There isn't as much understanding when it comes to things like vaccines.\\
+38 & Through many sources, including my own common sense and a lifetime of research into these matters.\\
+39 & I used the information that I have learned in school and in the past.\\
+40 & Most things I knew already as some were common knowledge\\
+41 & My gut instinct.\\
+\addlinespace
+42 & Most was knowledge I already knew, however if I were to find out or look up info, I would simply consult the internet\\
+43 & I thought about what i was told in school.\\
+44 & I checked reliable sources, and I knew most of them from learning this information in science class.\\
+48 & some things are obvious and don't need verification\\
+49 & Research on my own as opposed to jist blindly following what the media is told to tell me.\\
+\addlinespace
+56 & extensive research\\
+58 & I verified it with what then knowledge I know\\
+67 & I remember some of the stuff from school.\\
+68 & I clicked on two out of three of the links for most statements that I was unsure about.\\
+71 & I have a lot of people who do deep research that verified the info.\\
+\addlinespace
+77 & Just various amounts of research online. There have been times where simple questions were simply ignored or answered in a perplexing manner. Those types of answers always raise a flag.\\
+81 & By  the knowledge I already had.\\
+82 & Some of these questions I remember from high school. And I learned this in science. Therefore I do remember alot of these questions.\\
+88 & I don't know\\
+97 & Either I already knew the information or after reading the information I can to my conclusion\\
+\addlinespace
+98 & My wife has taken two different kinds of antibiotics, one for viral infections and one for bacterial infections\\
+100 & I independently verified the information by doing a lot of research and comparing studies.\\
+103 & I do my own research.\\
+114 & I guess I haven't independently verified the information.\\
+116 & wikipedia articles and google searches\\
+\addlinespace
+120 & The questions I got wrong were facts that I had simply forgotten. Once made aware of the correct answer, I remembered that it was in fact the correct answer.\\
+122 & heard of it before\\
+123 & I used the links provided and then googled the question if I was unsure\\
+126 & It is an obvious pattern whenever these events happen that there is a coverup, you just have to research the track record\\
+129 & I'm unsure of which question, you asked several.
+it's a bit confusing.
+the vaccine question has been verified by personal treatment and independent studies that were banned on social platforms\\
+\addlinespace
+131 & I used the information and links given to verify it.\\
+132 & I have been studying this information for many years.
+Real science is not consensus alone.
+Peer review is rigged and fraudulent.
+We are not being given the truth about many things and most "scientists" don't even know that they are wrong and refuse to reinvestigate the topics.\\
+136 & Have an educational background in anthropology and took a lot of science courses. I trust scientists generally but not always the way the findings are relayed.\\
+137 & My own research, we were lied to in school\\
+144 & By looking through multiple sources on a given topic.\\
+\addlinespace
+152 & A lot of this I have already learned, but I can't personally verify it bc I can't run the tests.\\
+154 & I have done several science experiments to prove a lot of these ideas.\\
+156 & looking it up online from at least 2-3 sources\\
+159 & Instead of taking what the mainstream talking point is and immediately accepting it, I like to do further research myself.\\
+161 & I already knew of the information presented to me for years.\\
+\addlinespace
+163 & I googled it and read information from various sources.\\
+164 & Many of these tidbits of information are taught as introduction to different subjects. I have either studied this or have used Google  Academic to verify.\\
+178 & Some of the questions I've done my own research on, others I can see with my own eyes.\\
+185 & most of the questions i already knew the answer to because i researched it a long time ago\\
+186 & The information presented is basic knowledge we learn throughout ours lives.
+Light travels faster than sound.
+The fathers genes decide whether ba baby will be male or female.
+Diamonds are formed from carbon.
+Calcium carbonate is not sodium chloride.
+It takes the earth 365.2ish to orbit our sun.
+Yes, water is 2 hydrogen, 1 oxygen.
+And and electron is much smaller than an atom.
+All things we learn in school.
+Oh and lasers use light waves not sound waves.\\
+\addlinespace
+187 & For light traveling faster then sound it's obviously verifiable because you see lightning before you hear thunder, lasers are visible so obviously they are made of light not sound, the definition of water is H2O so if something isn't H2O then it isn't water.
+The vast majority of the questions were extremely common sense if you think about them for any amount of time.\\
+191 & I either realized I was wrong or it sounded logical in my mind.\\
+200 & The Bible.\\
+\bottomrule
+\end{longtable}
+
 \end{document}

### 2024-05-31 functions/own_functions.R

@@ -4,23 +4,34 @@ rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)
 # get model outputs ready for inline text reporting
 text_ready <- function(model_output) {
   
-  result <- tidy(model_output, conf.int = TRUE) %>% 
-    filter(effect == "fixed") %>% 
+  result <- tidy(model_output, conf.int = TRUE)
+  
+  if ("effect" %in% colnames(result)) {
+    # Filter for effect == "fixed" if the variable exists
+    result <- result %>% 
+      filter(effect == "fixed")
+  } 
+  
+  result <- result %>% 
     # report p.value according to apa standards
     mutate(p.value = case_when(p.value < 0.001 ~ "< .001",
-                               TRUE ~ sprintf("%.3f", p.value)
+                               TRUE ~ paste0("= ", sprintf("%.3f", p.value))
     )
     ) %>% 
     # all other terms
     rounded_numbers() %>% 
-    mutate(
-      term = ifelse(term == "(Intercept)", "intercept", term),
-      ci = glue::glue("[{conf.low}, {conf.high}]"), 
-      # if there is an interaction (recognized by ":"), name it just interaction
-      term = str_replace(term, ".*:.*", "interaction")
-    ) %>% 
-    select(term, estimate, std.error, ci, p.value) %>% 
-    split(.$term)
+    mutate(ci = glue::glue("[{conf.low}, {conf.high}]")) 
+  
+  if ("term" %in% colnames(result)) {
+    # Filter for effect == "fixed" if the variable exists
+    result <- result %>% 
+      mutate(
+        term = ifelse(term == "(Intercept)", "intercept", term),
+        # if there is an interaction (recognized by ":"), name it just interaction
+        term = str_replace(term, ".*:.*", "interaction")
+      ) %>% 
+      split(.$term)
+  } 
   
   return(result)
 }
@@ -42,48 +53,6 @@ clean_t_test <- function(t.test) {
   
 }
 
-# Plot results from simulation
-
-# Plot competence distributions
-plot_competence_distributions <- function(data) {
-  # Your data frame
-  d <- data %>% 
-    group_by(competence) %>% 
-    reframe(alpha = unique(alpha), beta = unique(beta))
-  
-  # Function to plot the different competence distributions 
-  plot_competence_distribution <- function(alpha, beta, competence) {
-    ggplot() +
-      geom_function(
-        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
-      ) +
[…]

### 2024-05-31 functions/plot_theme.R

@@ -9,5 +9,7 @@ plot_theme <- theme_minimal(base_size = 12) +
     # Bold axis titles
     axis.title = element_text(face = "bold"), 
     # Bold legend labels
-    legend.text = element_text(face = "bold")
+    legend.text = element_text(face = "bold"),
+    # Increase facet wrap titles
+    strip.text = element_text(face = "bold")
     )

### 2024-06-24 appendix_exp1.Rmd
making summary plot and adding some first introduction bits

@@
 \FloatBarrier
+
+(ref:stimulus-example) Example of a science question, the scientific consensus and the corresponding acceptance question.
+
+```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-example)"}
+knitr::include_graphics("./figures/study1_question_example.png")
+```
+
+```{r questions-with-sources}
+# items <- read_csv("exp_1/materials/knowledge_items.csv") %>% 
+#   mutate(id = 1:nrow(.)) %>% 
+#   select(id, Question, `Scientific consensus`, `Reference(s)`)
+# 
+# # Output the table
+# kbl(items, booktabs = T, longtable = TRUE,
+#     caption = "Science knowledge items", 
+#     full_width = T) %>%
+#   column_spec(1, width = "1em") %>%
+#   column_spec(2, width = "10em") %>%
+#   column_spec(3, width = "10em") %>%
+#   column_spec(4, width = "15em") 
+```
 
@@
 ## Comparing items
-
-Make correlation table
 
@@
 ```{r}
-# run models to get correlation coefficient
-
-# Define the function to run regression and tidy results
-run_regression <- function(data, dependent_var, independent_var) {
-  lm(as.formula(paste(dependent_var, "~", independent_var)), data = data) %>%
-    tidy() %>%
-    filter(term == independent_var) %>%
-    mutate(outcome = {{dependent_var}}) %>% 
-    select(outcome, term, estimate)
-}
-
 # List of variables
@@
 results %>% 
+  filter(term != "intercept") %>%
+    mutate_if(is.numeric, round, digits = 2) %>% 
+  # make a combined variable of estimate and p.value
+  mutate(estimate = paste0(estimate, " (p", p.value, ")")) %>% 
+  select(outcome, term, estimate) %>% 
   pivot_wider(names_from = term, 
@@
          ) %>% 
-  mutate_if(is.numeric, round, digits = 2) %>% 
   mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 
@@
 ```{r}
-# Define the function to run regression and tidy results
-run_regression <- function(data, dependent_var, independent_var) {
-  lm(as.formula(paste(dependent_var, "~", independent_var)), data = data) 
-}
-
-# run models to get correlation coefficient
-
-# List of variables
-variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")
-
 # Run regression for each variable and store results in a list
-results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance")), variables)
+results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance", return = "model")), variables)
 ```
@@
 ```
+
+## By-question variation
+
+```{r}
+# calculate share
+plot_data <- exp1_long %>% 
+  group_by(subject) %>% 
+  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
+            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
+```
+
+### Knowledge
+
+(ref:exp1-questions-knowledge) Distribution of correct answers by question.
+
+```{r exp1-questions-knowledge, fig.cap="(ref:exp1-questions-knowledge)"}
+# plot knowledge
+ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  labs(x = "Question", y = "Share of correct responses") +
+  coord_flip()
+```
+
+### Acceptance
+
+(ref:exp1-questions-acceptance) Distribution of consensus acceptance by question.
+
+```{r exp1-questions-acceptance, fig.cap="(ref:exp1-questions-acceptance)"}
+# plot acceptance
+ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  labs(x = "Question", y = "Share of acceptance") +
+  coord_flip()
+```
+
+## Trust in science
+
+(ref:exp1-trust-scientists) Distribution of trust in scientists.
+
+```{r exp1-trust-scientists, fig.cap="(ref:exp1-trust-scientists)"}
+# calculate share
+plot_data <- exp1_wide %>% 
+  group_by(wgm_scientists) %>% 
+  summarize(n = n()) %>% 
+  ungroup() %>% 
+  mutate(share = n/sum(n)) %>% 
+  rounded_numbers()
+
+# plot knowledge
+ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
+  geom_col() +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  labs(x = "Trust in science", y = "Share of participants") +
+  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5, size = 3) 
+```
+
+## Conspiracy thinking
+
+### Distribution
+
+(ref:exp1-conspiracy-distribution) Distribution of conspiracy thinking.
+
+```{r exp1-conspiracy-distribution, fig.cap="(ref:exp1-conspiracy-distribution)"}
+# plot knowledge
+ggplot(exp1_wide, aes(x = BCTI_avg)) +
+  geom_histogram() +
+  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
+```
+
+### By-item variation
+
+(ref:exp1-conspiracy-items) Distribution of conspiracy thinking by item.
+
+```{r exp1-conspiracy-items, fig.cap="(ref:exp1-conspiracy-items)"}
+BCTI_data <- exp1_long %>% 
+  # bring to long format
+  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
+               names_to = "BCTI_item", values_to = "score") %>% 
+  select(id, c(starts_with("BCTI"), score)) %>% 
+  group_by(BCTI_item) %>% 
+  summarize(avg_score = mean(score)) %>% 
+  ungroup()
+
+ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
+    geom_pointrange(aes(ymin = 1, ymax = avg_score),
+                  # Make the lines a little thicker and the dots a little bigger
+                  fatten = 2, size = 1) +
+  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
+  labs(x = "Conspiracy item", y = "Average score") +
+  coord_cartesian(ylim = c(1, 9)) +
+  coord_flip()
+```
+
+## Open ended comments
+```{r}
+# levels(as.factor(exp1_wide$comments))
+table <- exp1_wide %>% 
+  drop_na(comments) %>% 
+  select(id, comments)
+
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Reasons for consensus rejection") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2, width = "40em") 
+```
+
+## Additional plots
+
+(ref:exp1-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
+
+```{r exp1-conditional-acceptance, fig.cap="(ref:exp1-conditional-acceptance)"}
+plot_data <- exp1_long %>% 
+  group_by(knowledge, acceptance) %>% 
+  count() %>% 
+  group_by(knowledge) %>% 
+  mutate (share = n/sum(n)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
+
+absolute_numbers <- plot_data %>% 
+  group_by(knowledge) %>% 
+  summarise(n = sum(n))
+
+ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
+  geom_col(position = "dodge") +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5, size = 3) +
+  scale_fill_viridis_d(option = "cividis") +
+  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
+  plot_theme +
+  geom_bracket(data = absolute_numbers,
+               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
+                   xmax = as.numeric(factor(knowledge)) + 0.4, 
+                   y.position = 1.05, label =  paste0("n = ", n)),
+               inherit.aes = FALSE,
+               step.increase = 0)
+```
+
+(ref:exp1-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus
+
+```{r exp1-plot-overview, fig.cap="(ref:exp1-plot-overview)", fig.height= 10, fig.width=10}
+# Plot distributions of knowledge and acceptance and their mean
+
+mean_value_knowledge <- mean(exp1_wide$avg_knowledge)
+max_n <- max(table(exp1_wide$avg_acceptance))
+
+knowledge_mean <- ggplot(exp1_wide, aes(x = avg_knowledge)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average science knowledge", 
+       y = "N")
+
+mean_value <- mean(exp1_wide$avg_acceptance)
+
+acceptance_mean <- ggplot(exp1_wide, aes(x = avg_acceptance)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average consensus acceptance", 
+       y = "N")
+
+# Trust in science plots
+plot_data <- exp1_wide %>% 
+  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
+  summarise(n = n())
+
+trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  # Add black border around tiles for better visibility
+  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) +  # Scatter points within each tile
+  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average science knowledge", 
+       y = "Trust in science", 
+       fill = "N")
+
+plot_data <- exp1_wide %>% 
+  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
+  summarise(n = n())
+
+trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  
+  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) + 
+  scale_fill_viridis_c(option = "plasma") + 
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average consensus Acceptance", 
+       y = "Trust in science", 
+       fill = "N") +
+  theme(legend.position = "top")
+
+# Conspiracy Thinking plots
+
+conspiracy_knowledge <- ggplot(exp1_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average science knowledge", 
+       y = "Average conspiracy thinking")
+
+conspiracy_acceptance <- ggplot(exp1_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average consensus acceptance", 
+       y = "Average conspiracy thinking") +
+  theme(legend.position = "top")
+
+# Combine plots
+
+# distributions
+distributions <- ggarrange(knowledge_mean + 
+                             rremove("xlab") , 
+                           acceptance_mean + 
+                             rremove("xlab") +
+                             rremove("ylab"))
+
+# trust
+trust <- ggarrange(trust_knowledge + 
+                             rremove("xlab") +
+                             guides(fill = "none") +
+                             labs(caption = NULL) , 
+                           trust_acceptance + 
+                             rremove("xlab") +
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# conspiracy
+conspiracy <- ggarrange(conspiracy_knowledge, 
+                           conspiracy_acceptance + 
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# common plot
+patchwork <- (distributions / trust / conspiracy) + 
+  plot_annotation(tag_levels = 'A')
+
+patchwork
+
+```

### 2024-06-24 appendix_exp2.Rmd
making summary plot and adding some first introduction bits

@@
 
-See Appendix \@ref(exp1).
+(ref:exp2-stimulus-example) Example of a science question, the scientific consensus and corresponding explanation and sources.
+
+```{r exp2-stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:exp2-stimulus-example)"}
+knitr::include_graphics("./figures/study2_question_example.png")
+```
 
@@
 ```
+
+## Additional plots
+
+(ref:exp2-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
+
+```{r exp2-conditional-acceptance, fig.cap="(ref:exp2-conditional-acceptance)"}
+plot_data <- exp2_long %>% 
+  group_by(knowledge, acceptance) %>% 
+  count() %>% 
+  group_by(knowledge) %>% 
+  mutate (share = n/sum(n)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
+
+absolute_numbers <- plot_data %>% 
+  group_by(knowledge) %>% 
+  summarise(n = sum(n))
+
+ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
+  geom_col(position = "dodge") +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5, size = 3) +
+  scale_fill_viridis_d(option = "cividis") +
+  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
+  plot_theme  +
+  geom_bracket(data = absolute_numbers,
+               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
+                   xmax = as.numeric(factor(knowledge)) + 0.4, 
+                   y.position = 1.05, label =  paste0("n = ", n)),
+               inherit.aes = FALSE,
+               step.increase = 0)
+```
+
+(ref:exp2-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
+```{r exp2-plot-overview, fig.cap="(ref:exp2-plot-overview)", fig.height= 12, fig.width=12}
+
+# Plot distributions of knowledge and acceptance and their mean
+mean_value_knowledge <- mean(exp2_wide$avg_knowledge)
+max_n <- max(table(exp2_wide$avg_acceptance))
+
+knowledge_mean <- ggplot(exp2_wide, aes(x = avg_knowledge)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average science knowledge", 
+       y = "N")
+
+mean_value <- mean(exp2_wide$avg_acceptance)
+
+acceptance_mean <- ggplot(exp2_wide, aes(x = avg_acceptance)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average consensus acceptance", 
+       y = "N")
+
+# Trust in science plots
+plot_data <- exp2_wide %>% 
+  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
+  summarise(n = n())
+
+trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  # Add black border around tiles for better visibility
+  geom_jitter(inherit.aes = FALSE, data = exp2_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) +  # Scatter points within each tile
+  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average science knowledge", 
+       y = "Trust in science", 
+       fill = "N")
+
+plot_data <- exp2_wide %>% 
+  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
+  summarise(n = n())
+
+trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  
+  geom_jitter(inherit.aes = FALSE, data = exp2_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) + 
+  scale_fill_viridis_c(option = "plasma") + 
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average consensus Acceptance", 
+       y = "Trust in science", 
+       fill = "N") +
+  theme(legend.position = "top")
+
+# Conspiracy Thinking plots
+
+conspiracy_knowledge <- ggplot(exp2_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average science knowledge", 
+       y = "Average conspiracy thinking")
+
+conspiracy_acceptance <- ggplot(exp2_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average consensus acceptance", 
+       y = "Average conspiracy thinking") +
+  theme(legend.position = "top")
+
+# Combine plots
+
+# distributions
+distributions <- ggarrange(knowledge_mean + 
+                             rremove("xlab") , 
+                           acceptance_mean + 
+                             rremove("xlab") +
+                             rremove("ylab"))
+
+# trust
+trust <- ggarrange(trust_knowledge + 
+                             rremove("xlab") +
+                             guides(fill = "none") +
+                             labs(caption = NULL) , 
+                           trust_acceptance + 
+                             rremove("xlab") +
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# conspiracy
+conspiracy <- ggarrange(conspiracy_knowledge, 
+                           conspiracy_acceptance + 
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# common plot
+patchwork <- (distributions / trust / conspiracy) + 
+  plot_annotation(tag_levels = 'A')
+
+patchwork
+
+```

### 2024-06-24 appendix_exp3.Rmd
making summary plot and adding some first introduction bits

@@
 
-See Appendix \@ref(exp1).
+(ref:exp3-explanation-example) Example of an explanation question and the opportunity to change the previous answer.
+
+```{r exp3-explanation-example, echo=FALSE, out.width= "50%", fig.align="center", fig.show="hold", fig.cap="(ref:exp3-explanation-example)"}
+knitr::include_graphics("figures/study3_example_explanation.png")
+```
 
@@
 ```
+
+## Additional plots
+
+(ref:exp3-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
+
+```{r exp3-conditional-acceptance, fig.cap="(ref:exp3-conditional-acceptance)"}
+plot_data <- exp3_long %>% 
+  pivot_longer(c(acceptance, acceptance_initial), 
+               names_to = "measure", 
+               values_to = "acceptance") %>% 
+  # give nicer names
+  mutate(measure = case_when(measure == "acceptance" ~ "Acceptance including revisions", 
+                             measure == "acceptance_initial" ~ "Initial acceptance"), 
+         # make a factor
+         measure = factor(measure, levels = unique(measure)[c(2, 1)])
+         ) %>% 
+  group_by(measure, knowledge, acceptance) %>% 
+  count() %>% 
+  group_by(measure, knowledge) %>% 
+  mutate (share = n/sum(n)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
+
+absolute_numbers <- plot_data %>% 
+  group_by(measure, knowledge) %>% 
+  summarise(n = sum(n))
+
+ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
+  geom_col(position = "dodge") +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5, size = 3) +
+  scale_fill_viridis_d(option = "cividis") +
+  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
+  plot_theme +
+  geom_bracket(data = absolute_numbers,
+               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
+                   xmax = as.numeric(factor(knowledge)) + 0.4, 
+                   y.position = 1.05, label =  paste0("n = ", n)),
+               inherit.aes = FALSE,
+               step.increase = 0) +
+  facet_wrap(~measure)
+```
+
+(ref:exp3-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
+
+```{r exp3-plot-overview, fig.cap="(ref:exp3-plot-overview)", fig.height= 12, fig.width=12}
+# Plot distributions of knowledge and acceptance and their mean
+
+mean_value_knowledge <- mean(exp3_wide$avg_knowledge)
+max_n <- max(table(exp3_wide$avg_acceptance))
+
+knowledge_mean <- ggplot(exp3_wide, aes(x = avg_knowledge)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average science knowledge", 
+       y = "N")
+
+mean_value <- mean(exp3_wide$avg_acceptance)
+
+acceptance_mean <- ggplot(exp3_wide, aes(x = avg_acceptance)) +
+  geom_bar() + 
+  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
+  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
+            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = seq(0, max_n, 10)) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, max_n)) +
+  labs(x = "Average consensus acceptance", 
+       y = "N")
+
+# Trust in science plots
+plot_data <- exp3_wide %>% 
+  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
+  summarise(n = n())
+
+trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  # Add black border around tiles for better visibility
+  geom_jitter(inherit.aes = FALSE, data = exp3_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) +  # Scatter points within each tile
+  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average science knowledge", 
+       y = "Trust in science", 
+       fill = "N")
+
+plot_data <- exp3_wide %>% 
+  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
+  summarise(n = n())
+
+trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
+  geom_tile() +  
+  geom_jitter(inherit.aes = FALSE, data = exp3_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
+              width = 0.03, height = 0.08) + 
+  scale_fill_viridis_c(option = "plasma") + 
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  coord_cartesian(xlim = c(0, 1)) +
+  scale_y_continuous(breaks = 1:5) +
+  labs(x = "Average consensus Acceptance", 
+       y = "Trust in science", 
+       fill = "N") +
+  theme(legend.position = "top")
+
+# Conspiracy Belief plot
+
+conspiracy_knowledge <- ggplot(exp3_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average science knowledge", 
+       y = "Average conspiracy thinking")
+
+conspiracy_acceptance <- ggplot(exp3_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(breaks = 1:9) +
+  coord_cartesian(xlim = c(0, 1), 
+                  ylim = c(0, 9)) +
+  labs(x = "Average consensus acceptance", 
+       y = "Average conspiracy thinking") +
+  theme(legend.position = "top")
+
+# Combine plots
+
+# distributions
+distributions <- ggarrange(knowledge_mean + 
+                             rremove("xlab") , 
+                           acceptance_mean + 
+                             rremove("xlab") +
+                             rremove("ylab"))
+
+# trust
+trust <- ggarrange(trust_knowledge + 
+                             rremove("xlab") +
+                             guides(fill = "none") +
+                             labs(caption = NULL) , 
+                           trust_acceptance + 
+                             rremove("xlab") +
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# conspiracy
+conspiracy <- ggarrange(conspiracy_knowledge, 
+                           conspiracy_acceptance + 
+                             rremove("ylab"), 
+                   common.legend = TRUE)
+
+# common plot
+patchwork <- (distributions / trust / conspiracy) + 
+  plot_annotation(tag_levels = 'A')
+
+patchwork
+```

### 2024-06-24 paper.Rmd
making summary plot and adding some first introduction bits

@@
 ---
-title             : "Who mistrust basic science?"
-shorttitle        : "Who mistrust basic science?"
+title             : "Universal acceptance of (nearly all of) basic science in the US"
+shorttitle        : "Acceptance of basic science"
 
@@
 library("modelsummary") # for regression tables
+library("ggrepel") # for better random patterns in plots
 ```
@@
 
-There is increasingly more talk of defiance towards science and towards experts more generally.
-In every population, many people only trust science ‘some of the time’, and a sizable minority doesn’t trust it.
-
-This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
-
-But how deep is distrust of science?
-Do people reject science wholesale, or is it more driven by attitudes towards specific contents, maybe in particular those seen as being politicized, or as going against deeply held beliefs, or behaviors they want to engage in?
-[previous work suggests that issue-specific claims provoke more negative attitudes]
-
-Here, we look at whether people in the US, across different levels of trust in science (including a sample with anti-vaccine attitudes), accept basic scientific facts.
-
-[We can draw some inspiration from the “Westwood et al 2021 Current research overstates American support for political violence” paper: we want to put in perspective the current rise (?
-discourse around?) in mistrust towards science]
-
-## Levels of trust in science
-
-## Why do people distrust science?
-
-[here? We know from the many studies on the gateway belief model that people don’t always accept the scientific consensus (indeed, the effect sizes are usually not massive)]
-
-## How much science do people know?
-
-This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven… so the first thing to do would be to look at data on knowledge of science, and see what the highest rates of correct answer is, to see how much room for improvement there is.
-
-“The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
-Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
-Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
-The results of the U.S.
-portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
-Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators’ questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).”
+How much do people trust and accept science?
+The recent COVID pandemic has highlighted the importance of this question, with trust in science being the best predictor of people’s proclivity to follow health guidelines [@alganTrustScientistsTimes2021].
+
+Unfortunately, trust in science is far from being at ceiling [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020], and there are suggestions that it has recently been dropping in some countries [@alganTrustScientistsTimes2021] and that it is increasingly politicized in the US [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
+
+This relative lack of trust in science could mean different things.
+It could mean that some people reject science wholesale, or at least to a significant extent.
+Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (creationism, etc.), or beliefs that are associated with disliked behaviors (vaccination, acting for climate change).
+
+We answer this question by looking at how much Americans accept basic scientific knowledge.
+We’ll pay special attention to two groups of participants: people who say they don’t trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
+
+We presently briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
+
+## Trust in science
+
+Across the globe, most people say they trust science at least to some extent.
+In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018] .
+In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants [@wellcomeglobalmonitorWellcomeGlobalMonitor2020].
+Across all countries, in 2018, 32% of participants said they trust science "a lot" (41% in 2020), 45% trust "some" science (39% in 2020), and only 13% percent trust science "not much/ not at all" (13% in 2020), while 10% indicated "don't know" (7% in 2020) (<https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020#gid=6acd&pid=0>).
+
+Trust in science also appears to be relatively stable in the US [@funkPublicConfidenceScientists2020].
+However, there are some potentially worrying trends: The US might be increasingly polarized on trust in science [@gauchatPoliticizationSciencePublic2012].
+For example, the partisan divide on the question of whether climate change is a major threat to societey [has been growing in the last decade](https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/) [@kennedyWhatDataSays2023].
+And while by some measures only about half of the US population believes in anthropogenic climate change, people do mostly not seem to change their mind on on the issue [@mottaChangingMindsChanging2021].
+
+Trust in science is typically measured in three ways: either asking explictly about general trust in science or scientists, by asking about science attitudes (e.g.
+XX), or asking questions about specific, typically contentious, science topics (e.g.
+vaccines, climate change, GMOs).
+It is unclear to which extent these measures capture trust in basic science.
+To assess this, we can look at the literature on science knowledge.
+
+## Science knowledge
+
+A lot of research has been conducted on the public understanding of science.
+Part of this research are surveys measuring to which extent people have basic science knowledge (e.g.
+"Does the Earth go around the Sun, or does the Sun go around the Earth?").
+For us, these surveys provide a lower bound to how much people trust basic science: people who give the right answer (and not by chance) supposedly trust science on that question.
+Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge [@millerPublicUnderstandingAttitudes2004; @millerMeasurementCivicScientific1998a].
+For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria but not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016].
+A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.
+
+[To do: Make a descriptive overview of science knowledge across time/countries.
+This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven.
+But there are enough questions where people's knowledge appears quite poor anyways]
+
+## CT and TiS
+
+Recently, the relationship of one group to trust in science (TiS) has been particularly scrutinized: conspiracy theorists (CTists), either defined as people who endorse specific conspiracy theories, or as people who have a conspiracy mindset [@rutjensConspiracyBeliefsScience2022; @vranicDidMyOwn2022].
+CTists reject specific science knowledge (related to their CT), and they tend to trust science less on the whole.
+What remains unclear is to which extent CTists (dis)trust basic science.
+
+## The present studies
+
+In a series of studies, we ask basic science knowledge questions, inform people of the scientific consensus, and see if they accept it.
+Of particular interest are people who are low in TiS, and CTists.
+
+In study 1, we plainly inform participants of the scientific consensus and ask if they accept it.
+We use a subset of questions on basic science knowledge that have been used in numerous public opinion surveys [@allumScienceKnowledgeAttitudes2008; @durantPublicUnderstandingScience1989; @millerMeasurementCivicScientific1998a] sometimes referred to as the "Oxford scale" [@gauchatCulturalAuthorityScience2011].
+In study 2, we remove a problematic science question and present people with a more elaborated explanation of the scientific consensus, as well as additional sources.
+This allows us to remove the issue that participants might simply not have trusted us to correctly report the scientific consensus in study 1.
+Study 3 follows essentially the same design as study 2, but is run on a vaccine-skeptic sample.
+Further, we provided participants with an explicit option to revise their answer, in case they rejected the consensus.
+We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently.
+Study 4 uses the same design as study 3, but asking science questions that people are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves [on an anti-vaxx sample again?].
+
+Then: replications in other countries?
 
@@
 
-(ref:stimulus-example) Example of a science question, the scientific consensus and the corresponding acceptance question.
-
-```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-example)"}
-knitr::include_graphics("./figures/study1_question_example.png")
-```
-
 ### Materials
@@
 To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 (Table \@ref(tab:conspiracy)).
+Participants were asked: "Below is a list of events for which the official version has been disputed.
+For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
+[1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true]".
 
@@
 
-data <- data.frame(id = 1:length(items), items = items)
+conspiracy_items <- data.frame(id = 1:length(items), items = items)
 
 # Output the table
-kbl(data, booktabs = T, longtable = T, col.names = NULL, 
+kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL, 
     caption = "Conspiracy items", 
@@
 
-(ref:exp1-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
-
-```{r exp1-conditional-acceptance, fig.cap="(ref:exp1-conditional-acceptance)"}
-plot_data <- exp1_long %>% 
-  group_by(knowledge, acceptance) %>% 
-  count() %>% 
-  group_by(knowledge) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-absolute_numbers <- plot_data %>% 
-  group_by(knowledge) %>% 
-  summarise(n = sum(n))
-
-ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  plot_theme +
-  geom_bracket(data = absolute_numbers,
-               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
-                   xmax = as.numeric(factor(knowledge)) + 0.4, 
-                   y.position = 1.05, label =  paste0("n = ", n)),
-               inherit.aes = FALSE,
-               step.increase = 0)
-```
-
 For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus and trust in science (r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`).
@@
 We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`), nor with trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`).
-
-(ref:exp1-plot) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus  **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus
-
-```{r exp1-plot, fig.cap="(ref:exp1-plot)", fig.height= 10, fig.width=10}
-# Plot distributions of knowledge and acceptance and their mean
-
-mean_value_knowledge <- mean(exp1_wide$avg_knowledge)
-max_n <- max(table(exp1_wide$avg_acceptance))
-
-knowledge_mean <- ggplot(exp1_wide, aes(x = avg_knowledge)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
-            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = seq(0, max_n, 10)) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, max_n)) +
-  labs(x = "Average science knowledge", 
-       y = "N")
-
-mean_value <- mean(exp1_wide$avg_acceptance)
-
-acceptance_mean <- ggplot(exp1_wide, aes(x = avg_acceptance)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
-            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = seq(0, max_n, 10)) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, max_n)) +
-  labs(x = "Average consensus acceptance", 
-       y = "N")
-
-# Trust in science plots
-plot_data <- exp1_wide %>% 
-  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
-  summarise(n = n())
-
-trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  # Add black border around tiles for better visibility
-  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) +  # Scatter points within each tile
-  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  coord_cartesian(xlim = c(0, 1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average science knowledge", 
-       y = "Trust in science", 
-       fill = "N")
-
-plot_data <- exp1_wide %>% 
-  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
-  summarise(n = n())
-
-trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  
-  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) + 
-  scale_fill_viridis_c(option = "plasma") + 
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  coord_cartesian(xlim = c(0, 1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average consensus Acceptance", 
-       y = "Trust in science", 
-       fill = "N") +
-  theme(legend.position = "top")
-
-# Conspiracy Thinking plots
-
-conspiracy_knowledge <- ggplot(exp1_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:9) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, 9)) +
-  labs(x = "Average science knowledge", 
-       y = "Average conspiracy thinking")
-
-conspiracy_acceptance <- ggplot(exp1_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:9) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, 9)) +
-  labs(x = "Average consensus acceptance", 
-       y = "Average conspiracy thinking") +
-  theme(legend.position = "top")
-
-# Combine plots
-
-# distributions
-distributions <- ggarrange(knowledge_mean + 
-                             rremove("xlab") , 
-                           acceptance_mean + 
-                             rremove("xlab") +
-                             rremove("ylab"))
-
-# trust
-trust <- ggarrange(trust_knowledge + 
-                             rremove("xlab") +
-                             guides(fill = "none") +
-                             labs(caption = NULL) , 
-                           trust_acceptance + 
-                             rremove("xlab") +
-                             rremove("ylab"), 
-                   common.legend = TRUE)
-
-# conspiracy
-conspiracy <- ggarrange(conspiracy_knowledge, 
-                           conspiracy_acceptance + 
-                             rremove("ylab"), 
-                   common.legend = TRUE)
-
-# common plot
-patchwork <- (distributions / trust / conspiracy) + 
-  plot_annotation(tag_levels = 'A')
-
-patchwork
-
-```
 
@@
 
-(ref:exp2-stimulus-example) Example of a science question, the scientific consensus and corresponding explanation and sources.
-
-```{r exp2-stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:exp2-stimulus-example)"}
-knitr::include_graphics("./figures/study2_question_example.png")
-```
-
 ### Materials
@@
 For conspiracy thinking we find a small negative one (r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
-
-(ref:exp2-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
-
-```{r exp2-conditional-acceptance, fig.cap="(ref:exp2-conditional-acceptance)"}
-plot_data <- exp2_long %>% 
-  group_by(knowledge, acceptance) %>% 
-  count() %>% 
-  group_by(knowledge) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-absolute_numbers <- plot_data %>% 
-  group_by(knowledge) %>% 
-  summarise(n = sum(n))
-
-ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  plot_theme  +
-  geom_bracket(data = absolute_numbers,
-               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
-                   xmax = as.numeric(factor(knowledge)) + 0.4, 
-                   y.position = 1.05, label =  paste0("n = ", n)),
-               inherit.aes = FALSE,
-               step.increase = 0)
-```
 
@@
 In very few cases (`r exp2_descriptives$conditional_acceptance$true$No$share`), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
-
-(ref:exp2-plot) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
-
-```{r exp2-plot, fig.cap="(ref:exp2-plot)", fig.height= 12, fig.width=12}
-# Plot distributions of knowledge and acceptance and their mean
-
-mean_value_knowledge <- mean(exp2_wide$avg_knowledge)
-max_n <- max(table(exp2_wide$avg_acceptance))
-
-knowledge_mean <- ggplot(exp2_wide, aes(x = avg_knowledge)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
-            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = seq(0, max_n, 10)) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, max_n)) +
-  labs(x = "Average science knowledge", 
-       y = "N")
-
-mean_value <- mean(exp2_wide$avg_acceptance)
-
-acceptance_mean <- ggplot(exp2_wide, aes(x = avg_acceptance)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
-            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = seq(0, max_n, 10)) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, max_n)) +
-  labs(x = "Average consensus acceptance", 
-       y = "N")
-
-# Trust in science plots
-plot_data <- exp2_wide %>% 
-  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
-  summarise(n = n())
-
-trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  # Add black border around tiles for better visibility
-  geom_jitter(inherit.aes = FALSE, data = exp2_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) +  # Scatter points within each tile
-  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  coord_cartesian(xlim = c(0, 1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average science knowledge", 
-       y = "Trust in science", 
-       fill = "N")
-
-plot_data <- exp2_wide %>% 
-  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
-  summarise(n = n())
-
-trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  
-  geom_jitter(inherit.aes = FALSE, data = exp2_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) + 
-  scale_fill_viridis_c(option = "plasma") + 
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  coord_cartesian(xlim = c(0, 1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average consensus Acceptance", 
-       y = "Trust in science", 
-       fill = "N") +
-  theme(legend.position = "top")
-
-# Conspiracy Thinking plots
-
-conspiracy_knowledge <- ggplot(exp2_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:9) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, 9)) +
-  labs(x = "Average science knowledge", 
-       y = "Average conspiracy thinking")
-
-conspiracy_acceptance <- ggplot(exp2_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:9) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, 9)) +
-  labs(x = "Average consensus acceptance", 
-       y = "Average conspiracy thinking") +
-  theme(legend.position = "top")
-
-# Combine plots
-
-# distributions
-distributions <- ggarrange(knowledge_mean + 
-                             rremove("xlab") , 
-                           acceptance_mean + 
-                             rremove("xlab") +
-                             rremove("ylab"))
-
-# trust
-trust <- ggarrange(trust_knowledge + 
-                             rremove("xlab") +
-                             guides(fill = "none") +
-                             labs(caption = NULL) , 
-                           trust_acceptance + 
-                             rremove("xlab") +
-                             rremove("ylab"), 
-                   common.legend = TRUE)
-
-# conspiracy
-conspiracy <- ggarrange(conspiracy_knowledge, 
-                           conspiracy_acceptance + 
-                             rremove("ylab"), 
-                   common.legend = TRUE)
-
-# common plot
-patchwork <- (distributions / trust / conspiracy) + 
-  plot_annotation(tag_levels = 'A')
-
-patchwork
-```
 
@@
   n_subject_by_outcome = exp3_wide %>% 
-    summarize(across(c("wgm_scientists", "reason_agreement", "BCTI_avg"), 
+    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
                      ~sum(!is.na(.x)
@@
 
+(ref:summary-plot) Points represent the average share of acceptance as a function of the **A** level of trust in science ("In general, would you say that you trust science a lot, some, not much, or not at all?
+[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"); **B** average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); **C** belief in a specific conspiracy theory.
+Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: "1 - completly false, 5 - unsure, 9 - completely true".
+Averages are only based on participants who believed at least to some degree in the conspiracy theory.
+We consider a believer as anyone scoring higher than the scale midpoint, i.e.
+\>5.
+Number labels in plots represent the number of participants for the corresponding data point.
+
+```{r summary-plot, fig.cap="(ref:summary-plot)", fig.height= 10, fig.width=10}
+# combine data of all three studies
+combined_data <- bind_rows(exp1_long %>% 
+                             mutate(study = "study 1"), 
+                           exp2_long %>% 
+                             mutate(study = "study 2"), 
+                           exp3_long %>% 
+                             mutate(study = "study 3")
+                           )
+
+# make plot data for trust
+plot_trust <- combined_data %>% 
+  drop_na(wgm_sciencegeneral, acceptance) %>% 
+  # make a labels version of trust in science
+  mutate(wgm_sciencegeneral = case_when(
+    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
+    wgm_sciencegeneral == 2 ~ "2 (Not much)",
+    wgm_sciencegeneral == 3 ~ "3 (Some)",
+    wgm_sciencegeneral == 4 ~ "4 (A lot)",
+    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
+  )) %>% 
+  group_by(study, wgm_sciencegeneral, acceptance) %>% 
+  summarize(n = n(), 
+            n_participants = n_distinct(id)) %>% 
+  group_by(study, wgm_sciencegeneral) %>% 
+  mutate (share = n/sum(n), 
+          n_participants = max(n_participants)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
+
+# make trust plot
+trust <- ggplot(plot_trust %>% 
+         filter(acceptance == "Yes"), 
+       aes(x = wgm_sciencegeneral, y = share, color = study, shape = study)) +
+  geom_point(position = "dodge", alpha = 0.5) +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
+  geom_text_repel(aes(label = paste0(n_participants)),
+            vjust = -0.5, size = 3,
+            show.legend = FALSE) +
+  scale_color_viridis_d(option = "plasma", end = 0.8) +
+  labs(x = "Trust in science", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
+  plot_theme + 
+  theme(legend.position = "top")
+
+# make plot data for conspiracy belief
+plot_conspiracy <- combined_data %>% 
+  # bring to long format
+  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
+               names_to = "BCTI_item", values_to = "score") %>% 
+  # make a belief variable if participant indicates a numeric answer above the
+  # scale midpoint
+  mutate(belief = ifelse(score > 5, TRUE, 
+                         ifelse(is.na(score), NA, FALSE)
+                         ),
+         BCTI_item = sub("^BCTI_", "", BCTI_item)
+         ) %>% 
+  group_by(study, BCTI_item, belief, acceptance) %>% 
+  summarize(n = n(),
+            n_participants = n_distinct(id)) %>% 
+  group_by(study, BCTI_item, belief) %>% 
+  mutate (share = n/sum(n), 
+          n_participants = max(n_participants)) %>% 
+  mutate_if(is.numeric, round, digits = 3) %>% 
+  ungroup()
+
+# make conspiracy belief plot
+conspiracy <- ggplot(plot_conspiracy %>% 
+         filter(acceptance == "Yes" & belief == TRUE), 
+       aes(x = BCTI_item, y = share, color = study, shape = study)) +
+  geom_point(position = "dodge", alpha = 0.5) +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
+  geom_text_repel(aes(label = paste0(n_participants)),
+            vjust = -0.5, size = 3,
+            show.legend = FALSE) +
+  scale_color_viridis_d(option = "plasma", end = 0.8) +
+  labs(x = "Conspiracy theory belief", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
+  plot_theme +
+  theme(axis.text.x = element_text(angle = 30, hjust = 1))
+
+# Conspiracy Thinking
+
+# Define breaks for each integer from 1 to 9
+breaks <- seq(0, 100, by = 20)
+
+# Define labels for each bin in the desired format
+labels <- paste0(c(0, 20, 40, 60, 80), "-", c(20, 40, 60, 80, 100))
+
+plot_data <- combined_data %>% 
+  drop_na(CMQ_avg) %>% 
+  # Create the binned version of CMQ_avg
+  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
+  group_by(study, CMQ_avg_binned, acceptance) %>% 
+  summarize(n = n()) %>% 
+  group_by(study, CMQ_avg_binned) %>% 
+  mutate (share = n/sum(n)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
+
+n_participants <- combined_data %>% 
+  drop_na(CMQ_avg) %>% 
+  # Create the binned version of CMQ_avg
+  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
+  group_by(study, CMQ_avg_binned) %>% 
+  summarize(n_participants = n_distinct(id))
+
+plot_conspiracy_thinking <- left_join(plot_data, n_participants)
+
+# make plot
+conspiracy_thinking  <- ggplot(plot_conspiracy_thinking  %>% 
+         filter(acceptance == "Yes"), 
+       aes(x = CMQ_avg_binned, y = share, color = study, shape = study)) +
+  geom_point(position = "dodge", alpha = 0.5) +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
+  geom_text_repel(aes(label = paste0(n_participants)),
+            vjust = -0.5, size = 3,
+            show.legend = FALSE) +
+  scale_color_viridis_d(option = "plasma", end = 0.8) +
+  labs(x = "Conspiracy Thinking", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
+  plot_theme + 
+  theme(legend.position = "top")
+
+# remove elements of plots for better overall plot
+conspiracy_thinking <- conspiracy_thinking + 
+  #rremove("ylab") +
+  theme(axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank()) 
+  #guides(color = "none", shape = "none") +
+
+conspiracy <- conspiracy +
+  rremove("ylab") +
+  guides(color = "none", shape = "none")
+
+trust <- trust +
+  rremove("ylab")
+
+# common plot
+patchwork <- (trust + conspiracy_thinking)/conspiracy + 
+  plot_layout(guides = 'collect', widths = c(1, 1)) +
+  plot_annotation(tag_levels = 'A')  & theme(legend.position = "top")
+
+wrap_elements(patchwork) +
+  labs(tag = "Acceptance rate of consensus") +
+    theme(
+    plot.tag = element_text(angle = 90, , face = "bold", size = 12),
+    plot.tag.position = "left"
+  )
+```
+
 Study 3 is essentially a replication-–with some minor modifications-–of study 2, but on a different type of sample.
@@
 However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
-This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = `r exp3_descriptives$n_subject_by_outcome$wgm_scientists` for trust in science, N = `r exp3_descriptives$n_subject_by_outcome$BCTI_avg` for conspiracy measures, N = `r exp3_descriptives$n_subject_by_outcome$reason_agreement` for reason for accepting consensus).
+This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = `r exp3_descriptives$n_subject_by_outcome$wgm_sciencegeneral` for trust in science, N = `r exp3_descriptives$n_subject_by_outcome$BCTI_avg` for conspiracy measures, N = `r exp3_descriptives$n_subject_by_outcome$reason_agreement` for reason for accepting consensus).
 
@@
 Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".
-
-(ref:exp3-explanation-example) Example of an explanation question and the opportunity to change the previous answer.
-
-```{r exp3-explanation-example, echo=FALSE, out.width= "50%", fig.align="center", fig.show="hold", fig.cap="(ref:exp3-explanation-example)"}
-knitr::include_graphics("figures/study3_example_explanation.png")
-```
 
@@
 Appendix \@ref(exp3) also includes more descriptive statistics, such as knowledge and acceptance by science questions.
-
-(ref:exp3-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
-
-```{r exp3-conditional-acceptance, fig.cap="(ref:exp3-conditional-acceptance)"}
-plot_data <- exp3_long %>% 
-  pivot_longer(c(acceptance, acceptance_initial), 
-               names_to = "measure", 
-               values_to = "acceptance") %>% 
-  # give nicer names
-  mutate(measure = case_when(measure == "acceptance" ~ "Acceptance including revisions", 
-                             measure == "acceptance_initial" ~ "Initial acceptance"), 
-         # make a factor
-         measure = factor(measure, levels = unique(measure)[c(2, 1)])
-         ) %>% 
-  group_by(measure, knowledge, acceptance) %>% 
-  count() %>% 
-  group_by(measure, knowledge) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-absolute_numbers <- plot_data %>% 
-  group_by(measure, knowledge) %>% 
-  summarise(n = sum(n))
-
-ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  plot_theme +
-  geom_bracket(data = absolute_numbers,
-               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
-                   xmax = as.numeric(factor(knowledge)) + 0.4, 
-                   y.position = 1.05, label =  paste0("n = ", n)),
-               inherit.aes = FALSE,
-               step.increase = 0) +
-  facet_wrap(~measure)
-```
 
@@
 
-(ref:exp3-plot) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
-
-```{r exp3-plot, fig.cap="(ref:exp3-plot)", fig.height= 12, fig.width=12}
-# Plot distributions of knowledge and acceptance and their mean
-
-mean_value_knowledge <- mean(exp3_wide$avg_knowledge)
-max_n <- max(table(exp3_wide$avg_acceptance))
-
-knowledge_mean <- ggplot(exp3_wide, aes(x = avg_knowledge)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
-            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = seq(0, max_n, 10)) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, max_n)) +
-  labs(x = "Average science knowledge", 
-       y = "N")
-
-mean_value <- mean(exp3_wide$avg_acceptance)
-
-acceptance_mean <- ggplot(exp3_wide, aes(x = avg_acceptance)) +
-  geom_bar() + 
-  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
-  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
-            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = seq(0, max_n, 10)) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, max_n)) +
-  labs(x = "Average consensus acceptance", 
-       y = "N")
-
-# Trust in science plots
-plot_data <- exp3_wide %>% 
-  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
-  summarise(n = n())
-
-trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  # Add black border around tiles for better visibility
-  geom_jitter(inherit.aes = FALSE, data = exp3_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) +  # Scatter points within each tile
-  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  coord_cartesian(xlim = c(0, 1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average science knowledge", 
-       y = "Trust in science", 
-       fill = "N")
-
-plot_data <- exp3_wide %>% 
-  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
-  summarise(n = n())
-
-trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
-  geom_tile() +  
-  geom_jitter(inherit.aes = FALSE, data = exp3_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
-              width = 0.03, height = 0.08) + 
-  scale_fill_viridis_c(option = "plasma") + 
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  coord_cartesian(xlim = c(0, 1)) +
-  scale_y_continuous(breaks = 1:5) +
-  labs(x = "Average consensus Acceptance", 
-       y = "Trust in science", 
-       fill = "N") +
-  theme(legend.position = "top")
-
-# Conspiracy Thinking plots
-
-conspiracy_knowledge <- ggplot(exp3_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:9) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, 9)) +
-  labs(x = "Average science knowledge", 
-       y = "Average conspiracy thinking")
-
-conspiracy_acceptance <- ggplot(exp3_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
-  geom_point() +
-  geom_smooth(method = "lm") +
-  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-  scale_y_continuous(breaks = 1:9) +
-  coord_cartesian(xlim = c(0, 1), 
-                  ylim = c(0, 9)) +
-  labs(x = "Average consensus acceptance", 
-       y = "Average conspiracy thinking") +
-  theme(legend.position = "top")
-
-# Combine plots
-
-# distributions
-distributions <- ggarrange(knowledge_mean + 
-                             rremove("xlab") , 
-                           acceptance_mean + 
-                             rremove("xlab") +
-                             rremove("ylab"))
-
-# trust
-trust <- ggarrange(trust_knowledge + 
-                             rremove("xlab") +
-                             guides(fill = "none") +
-                             labs(caption = NULL) , 
-                           trust_acceptance + 
-                             rremove("xlab") +
-                             rremove("ylab"), 
-                   common.legend = TRUE)
-
-# conspiracy
-conspiracy <- ggarrange(conspiracy_knowledge, 
-                           conspiracy_acceptance + 
-                             rremove("ylab"), 
-                   common.legend = TRUE)
-
-# common plot
-patchwork <- (distributions / trust / conspiracy) + 
-  plot_annotation(tag_levels = 'A')
-
-patchwork
-```
+# Discussion
+
+[We can draw some inspiration from the “Westwood et al 2021 Current research overstates American support for political violence” paper: we want to put in perspective the current rise (?
+discourse around?) in mistrust towards science]
+
+There is increasingly more talk of defiance towards science and towards experts more generally.
+In every population, many people only trust science ‘some of the time’, and a sizable minority doesn’t trust it.
+This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
+
+But how deep is distrust of science? In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
+
+Traditionally, people have interpreted people's lack of basic knowledge in light of the deficit-model. Here, we show that people readily accept the scientifically consensual answer to these questions.
+
+“The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
+Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
+Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
+The results of the U.S.
+portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
+Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators’ questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).”
 
@@
 
-# (APPENDIX) Appendix {-}
+# (APPENDIX) Appendix {.unnumbered}
 

### 2024-06-24 paper.tex
making summary plot and adding some first introduction bits

@@
 \hypersetup{
-  pdftitle={Who mistrust basic science?},
+  pdftitle={Universal acceptance of (nearly all of) basic science in the US},
   pdflang={en-EN},
@@
 
-\title{Who mistrust basic science?}
-\author{\phantom{0}}
+\title{Universal acceptance of (nearly all of) basic science in the US}
+\author{\textsuperscript{}}
 \date{}
 
-\shorttitle{Who mistrust basic science?}
+\shorttitle{Acceptance of basic science}
 
@@
 
-There is increasingly more talk of defiance towards science and towards experts more generally.
-In every population, many people only trust science `some of the time', and a sizable minority doesn't trust it.
-
-This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
-
-But how deep is distrust of science?
-Do people reject science wholesale, or is it more driven by attitudes towards specific contents, maybe in particular those seen as being politicized, or as going against deeply held beliefs, or behaviors they want to engage in?
-{[}previous work suggests that issue-specific claims provoke more negative attitudes{]}
-
-Here, we look at whether people in the US, across different levels of trust in science (including a sample with anti-vaccine attitudes), accept basic scientific facts.
-
-{[}We can draw some inspiration from the ``Westwood et al 2021 Current research overstates American support for political violence'' paper: we want to put in perspective the current rise (?
-discourse around?) in mistrust towards science{]}
-
-\subsection{Levels of trust in science}\label{levels-of-trust-in-science}
-
-\subsection{Why do people distrust science?}\label{why-do-people-distrust-science}
-
-{[}here? We know from the many studies on the gateway belief model that people don't always accept the scientific consensus (indeed, the effect sizes are usually not massive){]}
-
-\subsection{How much science do people know?}\label{how-much-science-do-people-know}
-
-This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven\ldots{} so the first thing to do would be to look at data on knowledge of science, and see what the highest rates of correct answer is, to see how much room for improvement there is.
-
-``The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
-Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
-Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
-The results of the U.S.
-portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
-Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators' questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).''
+How much do people trust and accept science?
+The recent COVID pandemic has highlighted the importance of this question, with trust in science being the best predictor of people's proclivity to follow health guidelines (Algan, Cohen, Davoine, Foucault, \& Stantcheva, 2021).
+
+Unfortunately, trust in science is far from being at ceiling (Wellcome Global Monitor, 2018, 2020), and there are suggestions that it has recently been dropping in some countries (Algan et al., 2021) and that it is increasingly politicized in the US (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
+
+This relative lack of trust in science could mean different things.
+It could mean that some people reject science wholesale, or at least to a significant extent.
+Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (creationism, etc.), or beliefs that are associated with disliked behaviors (vaccination, acting for climate change).
+
+We answer this question by looking at how much Americans accept basic scientific knowledge.
+We'll pay special attention to two groups of participants: people who say they don't trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
+
+We presently briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
+
+\subsection{Trust in science}\label{trust-in-science}
+
+Across the globe, most people say they trust science at least to some extent.
+In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science (Wellcome Global Monitor, 2018) .
+In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants (Wellcome Global Monitor, 2020).
+Across all countries, in 2018, 32\% of participants said they trust science ``a lot'' (41\% in 2020), 45\% trust ``some'' science (39\% in 2020), and only 13\% percent trust science ``not much/ not at all'' (13\% in 2020), while 10\% indicated ``don't know'' (7\% in 2020) (\url{https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020\#gid=6acd&pid=0}).
+
+Trust in science also appears to be relatively stable in the US (Funk \& Kennedy, 2020).
+However, there are some potentially worrying trends: The US might be increasingly polarized on trust in science (Gauchat, 2012).
+For example, the partisan divide on the question of whether climate change is a major threat to societey \href{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}{has been growing in the last decade} (Kennedy, 2023).
+And while by some measures only about half of the US population believes in anthropogenic climate change, people do mostly not seem to change their mind on on the issue (Motta, 2021).
+
+Trust in science is typically measured in three ways: either asking explictly about general trust in science or scientists, by asking about science attitudes (e.g.~XX), or asking questions about specific, typically contentious, science topics (e.g.~vaccines, climate change, GMOs).
+It is unclear to which extent these measures capture trust in basic science.
+To assess this, we can look at the literature on science knowledge.
+
+\subsection{Science knowledge}\label{science-knowledge}
+
+A lot of research has been conducted on the public understanding of science.
+Part of this research are surveys measuring to which extent people have basic science knowledge (e.g.~``Does the Earth go around the Sun, or does the Sun go around the Earth?'').
+For us, these surveys provide a lower bound to how much people trust basic science: people who give the right answer (and not by chance) supposedly trust science on that question.
+Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge (Miller, 1998, 2004).
+For example, only 55\% in the US (in 2014) and only 46\% in the EU (in 2005) knew that antibiotics kill only bacteria but not viruses (Science Literacy et al., 2016).
+A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.
+
+{[}To do: Make a descriptive overview of science knowledge across time/countries.
+This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven.
+But there are enough questions where people's knowledge appears quite poor anyways{]}
+
+\subsection{CT and TiS}\label{ct-and-tis}
+
+Recently, the relationship of one group to trust in science (TiS) has been particularly scrutinized: conspiracy theorists (CTists), either defined as people who endorse specific conspiracy theories, or as people who have a conspiracy mindset (Rutjens \& Većkalov, 2022; Vranic, Hromatko, \& Tonković, 2022).
+CTists reject specific science knowledge (related to their CT), and they tend to trust science less on the whole.
+What remains unclear is to which extent CTists (dis)trust basic science.
+
+\subsection{The present studies}\label{the-present-studies}
+
+In a series of studies, we ask basic science knowledge questions, inform people of the scientific consensus, and see if they accept it.
+Of particular interest are people who are low in TiS, and CTists.
+
+In study 1, we plainly inform participants of the scientific consensus and ask if they accept it.
+We use a subset of questions on basic science knowledge that have been used in numerous public opinion surveys (Allum, Sturgis, Tabourazi, \& Brunton-Smith, 2008; Durant, Evans, \& Thomas, 1989; Miller, 1998) sometimes referred to as the ``Oxford scale'' (Gauchat, 2011).
+In study 2, we remove a problematic science question and present people with a more elaborated explanation of the scientific consensus, as well as additional sources.
+This allows us to remove the issue that participants might simply not have trusted us to correctly report the scientific consensus in study 1.
+Study 3 follows essentially the same design as study 2, but is run on a vaccine-skeptic sample.
+Further, we provided participants with an explicit option to revise their answer, in case they rejected the consensus.
+We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently.
+Study 4 uses the same design as study 3, but asking science questions that people are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves {[}on an anti-vaxx sample again?{]}.
+
+Then: replications in other countries?
 
@@
 Finally, participants answered questions on conspiracy thinking and trust in science.
-
-\begin{figure}
-
-\includegraphics[width=1\linewidth]{./figures/study1_question_example} \hfill{}
-
-\caption{Example of a science question, the scientific consensus and the corresponding acceptance question.}\label{fig:stimulus-example}
-\end{figure}
 
@@
 To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by Pennycook, Binnendyk, and Rand (2022) (Table \ref{tab:conspiracy}).
+Participants were asked: ``Below is a list of events for which the official version has been disputed.
+For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
+{[}1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true{]}''.
 
@@
 
-\paragraph{Trust in science}\label{trust-in-science}
+\paragraph{Trust in science}\label{trust-in-science-1}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-conditional-acceptance.pdf}
-\caption{\label{fig:exp1-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
-
 For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29, p \textless{} .001), and acceptance of scientific consensus and trust in science (r = 0.27, p \textless{} .001).
@@
 
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-plot.pdf}
-\caption{\label{fig:exp1-plot}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
-\end{figure}
-
 \subsection{Discussion}\label{discussion}
@@
 
-\begin{figure}
-
-\includegraphics[width=1\linewidth]{./figures/study2_question_example} \hfill{}
-
-\caption{Example of a science question, the scientific consensus and corresponding explanation and sources.}\label{fig:exp2-stimulus-example}
-\end{figure}
-
 \subsubsection{Materials}\label{materials-1}
@@
 For conspiracy thinking we find a small negative one (r = -0.22, p = 0.006).
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-conditional-acceptance.pdf}
-\caption{\label{fig:exp2-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-plot.pdf}
-\caption{\label{fig:exp2-plot}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
-\end{figure}
-
 For RQ3, we got 35 answers from 25 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
@@
 \section{Experiment 3}\label{experiment-3}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/summary-plot.pdf}
+\caption{\label{fig:summary-plot}Points represent the average share of acceptance as a function of the \textbf{A} level of trust in science (``In general, would you say that you trust science a lot, some, not much, or not at all?
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''); \textbf{B} average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); \textbf{C} belief in a specific conspiracy theory.
+Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: ``1 - completly false, 5 - unsure, 9 - completely true''.
+Averages are only based on participants who believed at least to some degree in the conspiracy theory.
+We consider a believer as anyone scoring higher than the scale midpoint, i.e.~\textgreater5.
+Number labels in plots represent the number of participants for the corresponding data point.}
+\end{figure}
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.5\linewidth]{figures/study3_example_explanation}
-
-}
-
-\caption{Example of an explanation question and the opportunity to change the previous answer.}\label{fig:exp3-explanation-example}
-\end{figure}
-
 \subsubsection{Materials}\label{materials-2}
@@
 
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-conditional-acceptance.pdf}
-\caption{\label{fig:exp3-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
-
 Regarding RQ1, participants answered on average 75 \% (sd = 0.18) of the questions correctly.
@@
 
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-plot.pdf}
-\caption{\label{fig:exp3-plot}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
-\end{figure}
+\section{Discussion}\label{discussion-2}
+
+{[}We can draw some inspiration from the ``Westwood et al 2021 Current research overstates American support for political violence'' paper: we want to put in perspective the current rise (?
+discourse around?) in mistrust towards science{]}
+
+There is increasingly more talk of defiance towards science and towards experts more generally.
+In every population, many people only trust science `some of the time', and a sizable minority doesn't trust it.
+This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
+
+But how deep is distrust of science? In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
+
+Traditionally, people have interpreted people's lack of basic knowledge in light of the deficit-model. Here, we show that people readily accept the scientifically consensual answer to these questions.
+
+``The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
+Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
+Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
+The results of the U.S.
+portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
+Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators' questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).''
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-alganTrustScientistsTimes2021}
+Algan, Y., Cohen, D., Davoine, E., Foucault, M., \& Stantcheva, S.
+(2021).
+Trust in scientists in times of pandemic: Panel evidence from 12 countries.
+\emph{Proceedings of the National Academy of Sciences}, \emph{118}(40), e2108576118.
+\url{https://doi.org/10.1073/pnas.2108576118}
+
+\bibitem[\citeproctext]{ref-allumScienceKnowledgeAttitudes2008}
+Allum, N., Sturgis, P., Tabourazi, D., \& Brunton-Smith, I.
+(2008).
+Science knowledge and attitudes across cultures: a meta-analysis.
+\emph{Public Understanding of Science}, \emph{17}(1), 35--54.
+\url{https://doi.org/10.1177/0963662506070159}
+
 \bibitem[\citeproctext]{ref-bruderMeasuringIndividualDifferences2013}
@@
 
+\bibitem[\citeproctext]{ref-durantPublicUnderstandingScience1989}
+Durant, J. R., Evans, G. A., \& Thomas, G. P. (1989). The public understanding of science. \emph{Nature}, \emph{340}(6228), 11--14. \url{https://doi.org/10.1038/340011a0}
+
+\bibitem[\citeproctext]{ref-funkPublicConfidenceScientists2020}
+Funk, C., \& Kennedy, B. (2020). \emph{Public confidence in scientists has remained stable for decades}.
+
+\bibitem[\citeproctext]{ref-gauchatCulturalAuthorityScience2011}
+Gauchat, G.
+(2011).
+The cultural authority of science: Public trust and acceptance of organized science.
+\emph{Public Understanding of Science}, \emph{20}(6), 751--770.
+\url{https://doi.org/10.1177/0963662510365246}
+
+\bibitem[\citeproctext]{ref-gauchatPoliticizationSciencePublic2012}
+Gauchat, G.
+(2012).
+Politicization of Science in the Public Sphere: A Study of Public Trust in the United States, 1974 to 2010.
+\emph{American Sociological Review}, \emph{77}(2), 167--187.
+\url{https://doi.org/10.1177/0003122412438225}
+
+\bibitem[\citeproctext]{ref-kennedyWhatDataSays2023}
+Kennedy, A.
+T.
+C.
+F.
+and B.
+(2023).
+\emph{What the data says about americans{'} views of climate change}.
+Retrieved from \url{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}
+
+\bibitem[\citeproctext]{ref-krauseTrendsAmericansTrust2019a}
+Krause, N.
+M., Brossard, D., Scheufele, D.
+A., Xenos, M.
+A., \& Franke, K.
+(2019).
+Trends{\textemdash}americans{'} trust in science and scientists.
+\emph{Public Opinion Quarterly}, \emph{83}(4), 817--836.
+\url{https://doi.org/10.1093/poq/nfz041}
+
 \bibitem[\citeproctext]{ref-lantianMeasuringBeliefConspiracy2016}
@@
 
+\bibitem[\citeproctext]{ref-liPolarizationPublicTrust2022}
+Li, N., \& Qian, Y.
+(2022).
+Polarization of public trust in scientists between 1978 and 2018: Insights from a cross-decade comparison using interpretable machine learning.
+\emph{Politics and the Life Sciences}, \emph{41}(1), 45--54.
+\url{https://doi.org/10.1017/pls.2021.18}
+
+\bibitem[\citeproctext]{ref-millerMeasurementCivicScientific1998a}
+Miller, J. D. (1998). The measurement of civic scientific literacy. \emph{Public Understanding of Science}, \emph{7}(3), 203--223. \url{https://doi.org/10.1088/0963-6625/7/3/001}
+
+\bibitem[\citeproctext]{ref-millerPublicUnderstandingAttitudes2004}
+Miller, J.
+D.
+(2004).
+Public Understanding of, and Attitudes toward, Scientific Research: What We Know and What We Need to Know.
+\emph{Public Understanding of Science}, \emph{13}(3), 273--294.
+\url{https://doi.org/10.1177/0963662504044908}
+
+\bibitem[\citeproctext]{ref-mottaChangingMindsChanging2021}
+Motta, M.
+(2021).
+Changing Minds or Changing Samples?
+Disentangling Microlevel Stability and Macrolevel Growth in Anthropogenic Climate Change Beliefs.
+\emph{International Journal of Public Opinion Research}, \emph{33}(2), 477--489.
+\url{https://doi.org/10.1093/ijpor/edaa020}
+
 \bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
@@
 
+\bibitem[\citeproctext]{ref-rutjensConspiracyBeliefsScience2022}
+Rutjens, B. T., \& Većkalov, B. (2022). Conspiracy beliefs and science rejection. \emph{Current Opinion in Psychology}, \emph{46}, 101392. \url{https://doi.org/10.1016/j.copsyc.2022.101392}
+
+\bibitem[\citeproctext]{ref-committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016}
+Science Literacy, C.
+on, Public Perception of Science, Board on Science Education, Behavioral, D.
+of, Sciences, S., Education, \& National Academies of Sciences, Engineering, and Medicine.
+(2016).
+\emph{Science Literacy: Concepts, Contexts, and Consequences} (C.
+E.
+Snow \& K.
+A.
+Dibner, Eds.).
+Washington, D.C.: National Academies Press.
+\url{https://doi.org/10.17226/23595}
+
+\bibitem[\citeproctext]{ref-vranicDidMyOwn2022}
+Vranic, A., Hromatko, I., \& Tonković, M.
+(2022).
+{``}I did my own research{''}: Overconfidence, (dis)trust in science, and endorsement of conspiracy theories.
+\emph{Frontiers in Psychology}, \emph{13}.
+\url{https://doi.org/10.3389/fpsyg.2022.931865}
+
+\bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2018}
+Wellcome Global Monitor. (2018). \emph{Wellcome Global Monitor 2018}. Retrieved from \url{https://wellcome.org/reports/wellcome-global-monitor/2018}
+
+\bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2020}
+Wellcome Global Monitor. (2020). \emph{Wellcome Global Monitor 2020: Covid-19}. Retrieved from \url{https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020}
+
 \end{CSLReferences}
@@
 \FloatBarrier
+
+\begin{figure}
+
+\includegraphics[width=1\linewidth]{./figures/study1_question_example} \hfill{}
+
+\caption{Example of a science question, the scientific consensus and the corresponding acceptance question.}\label{fig:stimulus-example}
+\end{figure}
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-1}
+\subsubsection{Trust in science}\label{trust-in-science-2}
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-2}
+\subsubsection{Trust in science}\label{trust-in-science-3}
 
@@
 
-\subsection{Trust in science}\label{trust-in-science-3}
+\subsection{Trust in science}\label{trust-in-science-4}
 
@@
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:unnamed-chunk-29}Reasons for consensus rejection}\\
+\caption{\label{tab:unnamed-chunk-21}Reasons for consensus rejection}\\
 \toprule
@@
 
+\subsection{Additional plots}\label{additional-plots}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-conditional-acceptance.pdf}
+\caption{\label{fig:exp1-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-plot-overview.pdf}
+\caption{\label{fig:exp1-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
+\end{figure}
+
 \clearpage
@@
 
-See Appendix \ref{exp1}.
+\begin{figure}
+
+\includegraphics[width=1\linewidth]{./figures/study2_question_example} \hfill{}
+
+\caption{Example of a science question, the scientific consensus and corresponding explanation and sources.}\label{fig:exp2-stimulus-example}
+\end{figure}
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-4}
+\subsubsection{Trust in science}\label{trust-in-science-5}
 
@@
 
-\subsection{Trust in science}\label{trust-in-science-5}
+\subsection{Trust in science}\label{trust-in-science-6}
 
@@
 
+\subsection{Additional plots}\label{additional-plots-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-conditional-acceptance.pdf}
+\caption{\label{fig:exp2-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
 \clearpage
@@
 
-See Appendix \ref{exp1}.
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{figures/study3_example_explanation}
+
+}
+
+\caption{Example of an explanation question and the opportunity to change the previous answer.}\label{fig:exp3-explanation-example}
+\end{figure}
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-6}
+\subsubsection{Trust in science}\label{trust-in-science-7}
 
@@
 
-\subsection{Trust in science}\label{trust-in-science-7}
+\subsection{Trust in science}\label{trust-in-science-8}
 
@@
 
+\subsection{Additional plots}\label{additional-plots-2}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-conditional-acceptance.pdf}
+\caption{\label{fig:exp3-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp3-plot-overview.pdf}
+\caption{\label{fig:exp3-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
+\end{figure}
+
 \end{document}

### 2024-07-23 appendix_exp3.Rmd
preliminary write up of experiment 1-3 before merging distinct experimental sections

@@
 knitr::include_graphics("figures/study3_example_explanation.png")
+```
+
+Table \@ref(tab:knowledge-overview) shows all questions, their scientifically consensual answer, and their source.
+All but two questions were selected from existing science knowledge questionnaires.
+We tried to select non-political questions.
+
+```{r knowledge-overview}
+items <- read_csv("exp_3/materials/overview_questions.csv") %>% 
+  mutate(id = 1:nrow(.)) %>% 
+  select(id, "Question", "Scientific consensus (Study 1)", "Explanation (Study 2 & 3)")
+
+# Output the table
+kbl(items, booktabs = T, longtable = TRUE,
+    caption = "Science knowledge items", 
+    full_width = T) %>%
+  kable_styling(font_size = 8) %>% 
+  column_spec(1, width = "1em") %>%
+  column_spec(2, width = "10em") %>%
+  column_spec(3, width = "10em") %>%
+  column_spec(4, width = "23em") 
 ```

### 2024-07-23 paper.Rmd
preliminary write up of experiment 1-3 before merging distinct experimental sections

@@
 ```{r knowledge}
-items <- read_csv("exp_3/materials/overview_questions.csv") %>% 
+# Function to replace special characters in all cells of a data frame
+replace_special_characters <- function(df) {
+  df %>%
+    mutate(across(everything(), ~ gsub("\u2013|\u2014|\u2212", "-", .))) %>% # Replace en-dash, em-dash, and Unicode minus with ASCII hyphen-minus
+    mutate(across(everything(), ~ iconv(., from = "UTF-8", to = "ASCII//TRANSLIT"))) # Ensure all text is ASCII
+}
+
+# Read the CSV file
+items <- read_csv("exp_4/materials/overview_questions.csv")
+
+# Replace special characters
+items_cleaned <- replace_special_characters(items) %>% 
   mutate(id = 1:nrow(.)) %>%
-  select(id, "Question", "Scientific consensus (Study 1)", "Explanation (Study 2 & 3)")
+  select(id, everything())  # Ensure 'id' is the first column
+
+# add footnote for the tree items
+items_cleaned[11, 2] <- paste0("*", 
+                               items_cleaned[11, 2])
 
 # Output the table
-kbl(items, booktabs = T, longtable = TRUE,
+knitr::kable(items_cleaned, booktabs = T, longtable = TRUE,
     caption = "Science knowledge items", 
-    full_width = T) %>%
+    full_width = T, 
+    col.names = c("", "Study 1-3", "Study 4")) %>%
   kable_styling(font_size = 8) %>%
-  column_spec(1, width = "1em") %>%
-  column_spec(2, width = "10em") %>%
-  column_spec(3, width = "10em") %>%
-  column_spec(4, width = "23em") 
+  column_spec(1, width = "2em") %>%
+  column_spec(2, width = "22em")%>%
+  column_spec(3, width = "22em") %>% 
+  footnote(symbol = "Only used in Study 1")
 ```

### 2024-07-23 paper.tex
preliminary write up of experiment 1-3 before merging distinct experimental sections

@@
 
+\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{2em}>{\raggedright\arraybackslash}p{22em}>{\raggedright\arraybackslash}p{22em}}
+\caption{\label{tab:knowledge}Science knowledge items}\\
+\toprule
+ & Study 1-3 & Study 4\\
+\midrule
+1 & Do antibiotics kill viruses as well as bacteria?
+[Yes, both; No, only viruses; No, only bacteria] & For which disease is the drug bedaquiline, developed in 2007, a treatment?
+[Tetanus; Tuberculosis; Malaria]\\
+2 & Are electrons smaller, larger, or the same size as atoms?
+[Smaller; Same size; Larger] & What is the maximum speed a proton can attain in the largest particle collider as to 2015?
+[90\% of the speed of light; 99\% of the speed of light; the speed of light]\\
+3 & Have the continents on Earth been moving for millions of years or have they always been where they are now?
+[They have been moving; They have always been where they are now] & Kepler-452b is an exoplanet revolving around the star Kepler-452.
+How far away from the star is it, as established by astronomers in 2015?
+[97 million mi; 1,2 million mi; 1254 million mi]\\
+4 & What decides whether a baby is a boy or a girl ?
+Is it the father's genes, the mother's genes, or both?
+[The mother's genes; the father's genes; both] & Using bomb-pusle dating with carbon 14, what is the age of the oldest known vertebrate, as established in 2016?
+[138 years; 205 years;  392 years]\\
+5 & Do lasers work by focusing sound waves?
+[Yes; No] & How many more glials cells are there in the brain in comparison with neurons, as established in 2016?
+[The same amount; Twice as many; Tenth as many]\\
+\addlinespace
+6 & How long does it take for Earth to go around the sun: one day, one month, or one year?
+[One day; One month; One year] & As predicted by the general theory of relativity, how many times would the Earth keep orbiting if the Sun disappeared, as established in 2012?
+[47 seconds; 8 minutes;  2 hours]\\
+7 & Are diamonds made of carbon? [Yes; No] & What is the electric charge of the Higgs Boson, as established in 2012? [1.602176634 x 10-19; 0; 3.2x10-19C]\\
+8 & Which travels faster : light or sound? [Light; Sound] & What is the age of the oldest materials formed on Earth, as established in 2020? [Less than 4.6 Ga; Around 4.6 Ga; More than 4.6 Ga]\\
+9 & Is common table salt made of calcium carbonate? [Yes; No] & With the best current cloning techniques, what is the average success rate when operated on mice, as of 2010? [2,7\%; 9,4\%; 17,2\%]\\
+10 & Is water made of molecules containing one oxygen and two hydrogen atoms?
+[Yes; No] & What was the strength of the Earth magnetic field 3.7 billion years ago, as discovered this year?
+[15 microtesla; 30 microtesla; 45 microtesla]\\
+\addlinespace
+11 & *Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air] & \\
+\bottomrule
+\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{*} Only used in Study 1}\\
+\end{longtable}
+\endgroup{}
+
+\paragraph{Conspiracy scales}\label{conspiracy-scales}
+
+To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by Pennycook, Binnendyk, and Rand (2022) (Table \ref{tab:conspiracy}).
+Participants were asked: ``Below is a list of events for which the official version has been disputed.
+For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
+{[}1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true{]}''.
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
+\caption{\label{tab:conspiracy}Conspiracy items}\\
+\toprule
+1 & The Apollo moon landings never happened and were staged in a Hollywood film studio.\\
+2 & A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).\\
+3 & The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.\\
+4 & The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.\\
+5 & The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.\\
+\addlinespace
+6 & There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.\\
+7 & In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.\\
+8 & Proof of alien contact is being concealed from the public.\\
+9 & Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.\\
+10 & Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.\\
+\bottomrule
+\end{longtable}
+\endgroup{}
+
+To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix \ref{exp1}).
+
+\paragraph{Trust in science}\label{trust-in-science-1}
+
+Our main item for measuring trust in science is selected from the Wellcome Global Monitor survey: ``In general, would you say that you trust science a lot, some, not much, or not at all?
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''
+
+We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey (``How much do you trust scientists in this country?
+Do you trust them a lot, some, not much, or not at all?
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''), the other from the Pew research center (``How much confidence do you have in scientists to act in the best interests of the public?
+{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}'').
+We selected these items so that we could compare the ratings in our sample to global survey results.
+The WGM survey has been administered in over 140 countries and included over 140000 respondents.
+The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents (Cologna et al., 2024).
+
+\subsection{Results}\label{results}
+
+Regarding RQ1 and RQ2, participants answered on average 74 \% (sd = 0.16) of the questions correctly, and accepted the scientific consensus on average for 93 \% (sd = 0.12) of the questions.
+
+Fig.
+\ref{fig:exp1-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In most cases (76.3 \%), participants readily accepted the scientific consensus after having given the wrong answer to a question.
+In very few cases (1.6 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+We believe this might have been due to inattention.
+
+For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29, p \textless{} .001), and acceptance of scientific consensus and trust in science (r = 0.27, p \textless{} .001).
+The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp1-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+
+For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = -0.38, p \textless{} .001), and conspiracy thinking and acceptance of scientific consensus (r = -0.33, p \textless{} .001).
+
+In Appendix \ref{exp1}, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \ref{exp1} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus?
+In our main analyses, we looked at correlations of acceptance across all observations.
+One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus.
+To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question.
+We than calculated the correlation between trust in science and average acceptance rate by participant.
+We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = -0.14, p = 0.061), nor with trust in science (r = 0.06, p = 0.387).
+
+\subsection{Discussion}\label{discussion}
+
+These results suggest that most people accept the scientific consensus most of the time.
+Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
+Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
+In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
+
+\section{Experiment 2}\label{experiment-2}
+
+In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
+In most instances they did, but not always.
+In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
+To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
+To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
+We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \ref{exp1}).
+
+Based on experiment 1, we formulated the following hypotheses:
+
+\textbf{H1a: Higher trust in science is associated with more science knowledge?}
+
+\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?}
+
+\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
+
+\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?}
+
+We had the following research questions:
+
+\textbf{RQ1: What is the average science knowledge score?}
+
+\textbf{RQ2: When a participant's answer does not match the consensus, how often do they change their mind and accept the consensus?}
+
+\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
+
+\subsection{Methods}\label{methods-1}
+
+\subsubsection{Participants}\label{participants-1}
+
+We recruited 201 participants from the US via prolific.
+11 participants failed our attention check, resulting in a final sample of 190 participants (96 female, 94 male; \(age_\text{mean}\): 43.48, \(age_\text{sd}\): 12.25, \(age_\text{median}\): 42).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+\subsubsection{Procedure}\label{procedure-1}
+
+The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
+\ref{fig:exp2-stimulus-example}.
+
+\subsubsection{Materials}\label{materials-1}
+
+We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
+
+\subsection{Results}\label{results-1}
+
+As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = 0.28, p \textless{} .001) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = -0.40, p \textless{} .001).
+By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
+For trust in science, we find no statistically significant correlation (r = 0.16, p = 0.051).
+For conspiracy thinking we find a small negative one (r = -0.22, p = 0.006).
+
+Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp2-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+In Appendix \ref{exp2} we show that these results hold for our alternative measures of trust and conspiracy thinking.
+We also include more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Regarding RQ1, participants answered on average 79 \% (sd = 0.19) of the questions correctly, and accepted the scientific consensus on average for 98 \% (sd = 0.05) of the questions.
+Fig.
+\ref{fig:exp2-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In response to RQ2, in most cases (92.9 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In very few cases (0.5 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+
+For RQ3, we got 35 answers from 25 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \ref{tab:exp2-justifications} summarizes these answers by five categories.
+All answers can be read in Appendix \ref{exp2}.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-justifications}Justifications by category}
+
+\begin{tabular}{llll}
+\toprule
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+\midrule
+Personal convictions & 12.00 & 34.3\% & 8.00\\
+Mistake & 9.00 & 25.7\% & 8.00\\
+No justification & 7.00 & 20\% & 5.00\\
+Not convinced & 5.00 & 14.3\% & 5.00\\
+Religious Beliefs & 2.00 & 5.7\% & 2.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Discussion}\label{discussion-1}
+
+Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.~answered the knowledge question wrongly).
+The share of these latter is considerably larger in experiment 2 (92.9 \%) than in experiment 1 (76.3 \%).
+While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
+We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
+
+\section{Experiment 3}\label{experiment-3}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/summary-plot.pdf}
+\caption{\label{fig:summary-plot}Points represent the average share of acceptance as a function of the \textbf{A} level of trust in science (``In general, would you say that you trust science a lot, some, not much, or not at all?
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''); \textbf{B} average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); \textbf{C} belief in a specific conspiracy theory.
+Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: ``1 - completly false, 5 - unsure, 9 - completely true''.
+Averages are only based on participants who believed at least to some degree in the conspiracy theory.
+We consider a believer as anyone scoring higher than the scale midpoint, i.e.~\textgreater5.
+Number labels in plots represent the number of participants for the corresponding data point.}
+\end{figure}
+
+Study 3 is essentially a replication-\/--with some minor modifications-\/--of study 2, but on a different type of sample.
+Both study 1 and 2 were run on convenience samples.
+For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
+By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
+\ref{fig:exp3-explanation-example}).
+Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
+Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
+We want to know if participants perceive that this is because of trust, or other factors.
+
+As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
+
+\textbf{H1a: Higher trust in science is associated with more science knowledge?}
+
+\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus.}
+
+\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
+
+\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.}
+
+We had the following research questions:
+
+\textbf{RQ1: What is the average science knowledge score?}
+
+\textbf{RQ2: What is the average acceptance of the scientific consensus}
+
+\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
+
+\textbf{RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?}
+
+\subsection{Methods}\label{methods-2}
+
+\subsubsection{Participants}\label{participants-2}
+
+We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of 200 participants (125 female, 73 male; \(age_\text{mean}\): 42.90, \(age_\text{sd}\): 12.00, \(age_\text{median}\): 41).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
+This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = 140 for trust in science, N = 138 for conspiracy measures, N = 122 for reason for accepting consensus).
+
+\subsubsection{Procedure}\label{procedure-2}
+
+The procedure was mostly the same as in experiment 2.
+In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
+Finally, at the end of the survey, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
+Participants who selected ``You mostly agree with the consensus because you have been able to independently verify it'', were asked the open-ended follow-up question: ``Could you please tell us how you independently verified the information?''.
+
+\subsubsection{Materials}\label{materials-2}
+
+We relied on the same items as in experiment 2.
+We used a pre-registered categorization scheme for the open ended answers for why people reject the consensus.
+The categories were XX and XX.
+{[}Describe coding process, with independent reviewer etc.{]}
+
+\subsection{Results}\label{results-2}
+
+As in study 1 and 2, we find that participants answered on average 75 \% (sd = 0.18) of the questions correctly, and initially accepted the scientific consensus on average for 96 \% (sd = 0.08) of the questions (RQ2).
+The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (98 \%, sd = 0.06).
+
+Fig.
+\ref{fig:exp3-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In most cases (87 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
+After providing the chance to revise the initial consensus rejection, this share is even larger (95.5 \%).
+In very few cases (0.9 \%), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
+This share drops slightly after providing the chance to revise the initial consensus rejection (0.7 \%).
+
+In the appendix XX we provide an analysis that this drop is statistically significant {[}TO DO{]}.
+
+For all correlations, we include opinion revisions for measuring consensus acceptance.
+We find no statistically significant correlation between science knowledge and trust in science (r = 0.14, p = 0.100), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = 0.20, p = 0.019).
+Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp3-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+We find no statistically significant correlation between conspiracy thinking and science knowledge (r = -0.16, p = 0.055), and between conspiracy thinking and acceptance of scientific consensus (r = c(cor = -0.023), c(t = -0.27), = 0.788, c(df = 136), -0.189, 0.145, Pearson's product-moment correlation, two.sided, {[}-0.189, 0.145{]}, p = 0.788).
+
+In Appendix \ref{exp3}, we show that these associations are statistically significant (except the last one?).
+We show that they are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \ref{exp3} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Regarding RQ1, participants answered on average 75 \% (sd = 0.18) of the questions correctly.
+For RQ3, we got 74 answers from 47 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \ref{tab:exp3-justifications} summarizes these answers by five categories.
+
+All answers can be read in the appendix.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp3-justifications}Justifications by category}
+
+\begin{tabular}{llll}
+\toprule
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+\midrule
+Personal convictions & 25.00 & 33.8\% & 19.00\\
+Mistake & 21.00 & 28.4\% & 18.00\\
+No justification & 18.00 & 24.3\% & 13.00\\
+Not convinced & 7.00 & 9.5\% & 3.00\\
+Religious Beliefs & 3.00 & 4.1\% & 3.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+For RQ4, we had 122 participants answering the question.
+Of these 41.8\% said they accepted the scientific consensus because they trust scientists on this question, while 47.5\% said they independently verified the fact.
+10.7\% answered with other ``other'' and gave an open-ended explanation (see Appendix).
+
+We also asked all 58 participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed in Appendix XX.
+
+\section{Discussion}\label{discussion-2}
+
+{[}We can draw some inspiration from the ``Westwood et al 2021 Current research overstates American support for political violence'' paper: we want to put in perspective the current rise (?
+discourse around?) in mistrust towards science{]}
+
+There is increasingly more talk of defiance towards science and towards experts more generally.
+In every population, many people only trust science `some of the time', and a sizable minority doesn't trust it.
+This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
+
+But how deep is distrust of science? In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
+
+Traditionally, people have interpreted people's lack of basic knowledge in light of the deficit-model. Here, we show that people readily accept the scientifically consensual answer to these questions.
+
+``The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
+Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
+Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
+The results of the U.S.
+portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
+Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators' questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).''
+
+\FloatBarrier
+
+\section{References}\label{references}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-alganTrustScientistsTimes2021}
+Algan, Y., Cohen, D., Davoine, E., Foucault, M., \& Stantcheva, S.
+(2021).
+Trust in scientists in times of pandemic: Panel evidence from 12 countries.
+\emph{Proceedings of the National Academy of Sciences}, \emph{118}(40), e2108576118.
+\url{https://doi.org/10.1073/pnas.2108576118}
+
+\bibitem[\citeproctext]{ref-allumScienceKnowledgeAttitudes2008}
+Allum, N., Sturgis, P., Tabourazi, D., \& Brunton-Smith, I.
+(2008).
+Science knowledge and attitudes across cultures: a meta-analysis.
+\emph{Public Understanding of Science}, \emph{17}(1), 35--54.
+\url{https://doi.org/10.1177/0963662506070159}
+
+\bibitem[\citeproctext]{ref-bruderMeasuringIndividualDifferences2013}
+Bruder, M., Haffke, P., Neave, N., Nouripanah, N., \& Imhoff, R.
+(2013).
+Measuring individual differences in generic beliefs in conspiracy theories across cultures: Conspiracy mentality questionnaire.
+\emph{Frontiers in Psychology}, \emph{4}.
+Retrieved from \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2013.00225}
+
+\bibitem[\citeproctext]{ref-colognaTrustScientistsTheir2024}
+Cologna, V., Mede, N.
+G., Berger, S., Besley, J., Brick, C., Joubert, M., \ldots{} Linden, D.
+S.
+van der.
+(2024).
+\emph{Trust in scientists and their role in society across 67 countries}.
+\url{https://doi.org/10.31219/osf.io/6ay7s}
+
+\bibitem[\citeproctext]{ref-durantPublicUnderstandingScience1989}
+Durant, J. R., Evans, G. A., \& Thomas, G. P. (1989). The public understanding of science. \emph{Nature}, \emph{340}(6228), 11--14. \url{https://doi.org/10.1038/340011a0}
+
+\bibitem[\citeproctext]{ref-funkPublicConfidenceScientists2020}
+Funk, C., \& Kennedy, B. (2020). \emph{Public confidence in scientists has remained stable for decades}.
+
+\bibitem[\citeproctext]{ref-gauchatCulturalAuthorityScience2011}
+Gauchat, G.
+(2011).
+The cultural authority of science: Public trust and acceptance of organized science.
+\emph{Public Understanding of Science}, \emph{20}(6), 751--770.
+\url{https://doi.org/10.1177/0963662510365246}
+
+\bibitem[\citeproctext]{ref-gauchatPoliticizationSciencePublic2012}
+Gauchat, G.
+(2012).
+Politicization of Science in the Public Sphere: A Study of Public Trust in the United States, 1974 to 2010.
+\emph{American Sociological Review}, \emph{77}(2), 167--187.
+\url{https://doi.org/10.1177/0003122412438225}
+
+\bibitem[\citeproctext]{ref-kennedyWhatDataSays2023}
+Kennedy, A.
+T.
+C.
+F.
+and B.
+(2023).
+\emph{What the data says about americans{'} views of climate change}.
+Retrieved from \url{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}
+
+\bibitem[\citeproctext]{ref-krauseTrendsAmericansTrust2019a}
+Krause, N.
+M., Brossard, D., Scheufele, D.
+A., Xenos, M.
+A., \& Franke, K.
+(2019).
+Trends{\textemdash}americans{'} trust in science and scientists.
+\emph{Public Opinion Quarterly}, \emph{83}(4), 817--836.
+\url{https://doi.org/10.1093/poq/nfz041}
+
+\bibitem[\citeproctext]{ref-lantianMeasuringBeliefConspiracy2016}
+Lantian, A., Muller, D., Nurra, C., \& Douglas, K.
+M.
+(2016).
+Measuring Belief in Conspiracy Theories: Validation of a French and English Single-Item Scale.
+\emph{International Review of Social Psychology}, \emph{29}(1), 1.
+\url{https://doi.org/10.5334/irsp.8}
+
+\bibitem[\citeproctext]{ref-liPolarizationPublicTrust2022}
+Li, N., \& Qian, Y.
+(2022).
+Polarization of public trust in scientists between 1978 and 2018: Insights from a cross-decade comparison using interpretable machine learning.
+\emph{Politics and the Life Sciences}, \emph{41}(1), 45--54.
+\url{https://doi.org/10.1017/pls.2021.18}
+
+\bibitem[\citeproctext]{ref-millerMeasurementCivicScientific1998a}
+Miller, J. D. (1998). The measurement of civic scientific literacy. \emph{Public Understanding of Science}, \emph{7}(3), 203--223. \url{https://doi.org/10.1088/0963-6625/7/3/001}
+
+\bibitem[\citeproctext]{ref-millerPublicUnderstandingAttitudes2004}
+Miller, J.
+D.
+(2004).
+Public Understanding of, and Attitudes toward, Scientific Research: What We Know and What We Need to Know.
+\emph{Public Understanding of Science}, \emph{13}(3), 273--294.
+\url{https://doi.org/10.1177/0963662504044908}
+
+\bibitem[\citeproctext]{ref-mottaChangingMindsChanging2021}
+Motta, M.
+(2021).
+Changing Minds or Changing Samples?
+Disentangling Microlevel Stability and Macrolevel Growth in Anthropogenic Climate Change Beliefs.
+\emph{International Journal of Public Opinion Research}, \emph{33}(2), 477--489.
+\url{https://doi.org/10.1093/ijpor/edaa020}
+
+\bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
+Pennycook, G., Binnendyk, J., \& Rand, D.
+(2022).
+\emph{Overconfidently conspiratorial: Conspiracy believers are dispositionally overconfident and massively overestimate how much others agree with them}.
+\url{https://doi.org/10.31234/osf.io/d5fz2}
+
+\bibitem[\citeproctext]{ref-rutjensConspiracyBeliefsScience2022}
+Rutjens, B. T., \& Većkalov, B. (2022). Conspiracy beliefs and science rejection. \emph{Current Opinion in Psychology}, \emph{46}, 101392. \url{https://doi.org/10.1016/j.copsyc.2022.101392}
+
+\bibitem[\citeproctext]{ref-committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016}
+Science Literacy, C.
+on, Public Perception of Science, Board on Science Education, Behavioral, D.
+of, Sciences, S., Education, \& National Academies of Sciences, Engineering, and Medicine.
+(2016).
+\emph{Science Literacy: Concepts, Contexts, and Consequences} (C.
+E.
+Snow \& K.
+A.
+Dibner, Eds.).
+Washington, D.C.: National Academies Press.
+\url{https://doi.org/10.17226/23595}
+
+\bibitem[\citeproctext]{ref-vranicDidMyOwn2022}
+Vranic, A., Hromatko, I., \& Tonković, M.
+(2022).
+{``}I did my own research{''}: Overconfidence, (dis)trust in science, and endorsement of conspiracy theories.
+\emph{Frontiers in Psychology}, \emph{13}.
+\url{https://doi.org/10.3389/fpsyg.2022.931865}
+
+\bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2018}
+Wellcome Global Monitor. (2018). \emph{Wellcome Global Monitor 2018}. Retrieved from \url{https://wellcome.org/reports/wellcome-global-monitor/2018}
+
+\bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2020}
+Wellcome Global Monitor. (2020). \emph{Wellcome Global Monitor 2020: Covid-19}. Retrieved from \url{https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020}
+
+\end{CSLReferences}
+
+\newpage
+
+\appendix
+
+\section{Experiment 1}\label{exp1}
+
+\subsection{Materials}\label{materials-3}
+
+\FloatBarrier
+
+\begin{figure}
+
+\includegraphics[width=1\linewidth]{./figures/study1_question_example} \hfill{}
+
+\caption{Example of a science question, the scientific consensus and the corresponding acceptance question.}\label{fig:stimulus-example}
+\end{figure}
+
+\subsubsection{Conspiracy scales}\label{conspiracy-scales-1}
+
+Beside Belief in Conspiracy Theory Inventory (BCTI) by Pennycook et al. (2022) which report in the main study, we also assessed two other consipiracy thinking measures:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  The conspiracy mentality questionnaire (CMQ) by Bruder et al. (2013):
+  I think that . . .
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  \ldots{} many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
+\item
+  \ldots{} government agencies closely monitor all citizens.
+\item
+  \ldots{} events which superficially seem to lack a connection are often the result of secret activities.
+\item
+  \ldots{} there are secret organizations that greatly influence political decisions.
+\end{itemize}
+
+{[}0\% - 100\%; 0 = certainly not, 100 = certain{]}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  The Single Item Conspiracy Beliefs Scale (SICBS) by Lantian et al. (2016) :
+\end{enumerate}
+
+\begin{itemize}
+\tightlist
+\item
+  I think that the official version of the events given by the authorities very often hides the truth. {[}1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true{]}
+\end{itemize}
+
+\subsubsection{Trust in science}\label{trust-in-science-2}
+
+We rely on three items
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
+\item
+  In general, would you say that you trust science a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
+\item
+  How much confidence do you have in scientists to act in the best interests of the public? {[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}
+\end{enumerate}
+
+\subsection{Comparing items}\label{comparing-items}
+
+\subsubsection{Conspiracy theories}\label{conspiracy-theories}
+
+Table \ref{tab:correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
+\midrule
+BCTI & 1.00 & 0.58 & 0.56\\
+CMQ & 0.58 & 1.00 & 0.77\\
+SICBS & 0.56 & 0.77 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsubsection{Trust in science}\label{trust-in-science-3}
+
+Table \ref{tab:correlation-trust} shows the correlations of the three different items measuring trust in science.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:correlation-trust}Correlations of the three different items measuring trust in science}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
+\midrule
+wgm\_sciencegeneral & 1.00 & 0.85 & 0.75\\
+wgm\_scientists & 0.85 & 1.00 & 0.82\\
+pew & 0.75 & 0.82 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures}
+
+Table \ref{tab:correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}
+
+\begin{tabular}{lll}
+\toprule
+outcome & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+BCTI 
+(main conspiracy measure) & -0.38 (p< .001) & -0.33 (p< .001)\\
+CMQ & -0.11 (p= 0.121) & -0.07 (p= 0.317)\\
+SICBS & -0.1 (p= 0.158) & -0.09 (p= 0.213)\\
+WGM trust scientists & 0.23 (p= 0.001) & 0.26 (p< .001)\\
+WGM trust general 
+(main trust measure) & 0.29 (p< .001) & 0.27 (p< .001)\\
+PEW trust scientists & 0.16 (p= 0.027) & 0.21 (p= 0.003)\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses}
+
+Table \ref{tab:false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.
+
+\begin{table}
+
+\caption{\label{tab:false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
+\centering
+\begin{tabular}[t]{lcccccc}
+\toprule
+  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
+\midrule
+(Intercept) & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
+ & (0.073) & (0.074) & (0.074) & (0.073) & (0.074) & (0.074)\\
+avg\_acceptance & -0.139+ & 0.008 & -0.028 & 0.108 & 0.064 & 0.056\\
+ & (0.073) & (0.074) & (0.074) & (0.074) & (0.074) & (0.074)\\
+\midrule
+Num.Obs. & 184 & 184 & 184 & 184 & 184 & 184\\
+R2 & 0.019 & 0.000 & 0.001 & 0.012 & 0.004 & 0.003\\
+R2 Adj. & 0.014 & -0.005 & -0.005 & 0.006 & -0.001 & -0.002\\
+AIC & 523.6 & 527.2 & 527.0 & 525.0 & 526.4 & 526.6\\
+BIC & 533.2 & 536.8 & 536.7 & 534.7 & 536.1 & 536.2\\
+Log.Lik. & -258.801 & -260.578 & -260.512 & -259.506 & -260.204 & -260.295\\
+RMSE & 0.99 & 1.00 & 1.00 & 0.99 & 1.00 & 1.00\\
+\bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
+\end{tabular}
+\end{table}
+
+\subsection{By-question variation}\label{by-question-variation}
+
+\subsubsection{Knowledge}\label{knowledge}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-questions-knowledge.pdf}
+\caption{\label{fig:exp1-questions-knowledge}Distribution of correct answers by question.}
+\end{figure}
+
+\subsubsection{Acceptance}\label{acceptance}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-questions-acceptance.pdf}
+\caption{\label{fig:exp1-questions-acceptance}Distribution of consensus acceptance by question.}
+\end{figure}
+
+\subsection{Trust in science}\label{trust-in-science-4}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-trust-scientists.pdf}
+\caption{\label{fig:exp1-trust-scientists}Distribution of trust in scientists.}
+\end{figure}
+
+\subsection{Conspiracy thinking}\label{conspiracy-thinking}
+
+\subsubsection{Distribution}\label{distribution}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-conspiracy-distribution.pdf}
+\caption{\label{fig:exp1-conspiracy-distribution}Distribution of conspiracy thinking.}
+\end{figure}
+
+\subsubsection{By-item variation}\label{by-item-variation}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-conspiracy-items.pdf}
+\caption{\label{fig:exp1-conspiracy-items}Distribution of conspiracy thinking by item.}
+\end{figure}
+
+\subsection{Open ended comments}\label{open-ended-comments}
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
+\caption{\label{tab:unnamed-chunk-21}Reasons for consensus rejection}\\
+\toprule
+id & comments\\
+\midrule
+1 & no\\
+20 & None\\
+22 & I answered thoughtfully, good luck!\\
+27 & interesting, thank you.\\
+30 & n/a\\
+\addlinespace
+32 & None\\
+39 & no\\
+41 & none\\
+42 & n/a\\
+67 & thanks\\
+\addlinespace
+69 & No comments\\
+74 & no\\
+76 & none\\
+78 & None - thank you!\\
+79 & Should've clicked agree about plants getting carbon from the air, but I wanted to look it up first.
+Whenever I encounter something I didn't know online, I like to cross-reference.
+In retrospect though, everything else you asked about was simple and on the level, and I'm sure you are correct.
+
+Thanks for your hard work.\\
+\addlinespace
+84 & No, thank you.\\
+86 & n/a\\
+89 & N/A\\
+95 & no\\
+98 & Great topics! Thank you for this study.\\
+\addlinespace
+109 & To provide context for my responses, I currently study physiology (STEM) and perform academic research as an undergrad at my university.\\
+110 & n/a\\
+114 & Everything in the survey went fine with no issues.
+We seem to have a lot of conspiracy theory nuts in the world today and I think it's a sign of the larger mental health crisis that's currently going on.\\
+115 & I like the subject content of the survey\\
+116 & No\\
+\addlinespace
+132 & None.\\
+133 & very interesting\\
+135 & Nope\\
+139 & none\\
+140 & Air AND water make up the bulk of trees mass and the water is drawn from the earth.
+Water has much more mass than air and I stand by my answer of Earth being that is where the water is drawn from.
+The question was very vague and as you see...
+open to interpretation\\
+\addlinespace
+141 & n/a\\
+148 & no\\
+150 & no\\
+152 & None.\\
+154 & none, tysm!\\
+\addlinespace
+155 & Not virus, just metals/DDT. And I think it would be impossible for dinosaurs to have sex with the way they are shaped.\\
+157 & none\\
+159 & I think scientists are acting in the best interests of science rather than the community directly. An unethical scientist could advance scientific knowledge while hurting humanity.\\
+161 & Mo Thanks\\
+164 & I appreciate science but understand that if scientists don't go along with what investors/govt want they won't receive grants for funding.
+I also very recently lived through several years of gaslighting about covid.
+While I realize I'm not the sharpest tool in the shed, I'm not a total idiot.
+Things that were censored at the beginning of the pandemic are now allowed to be discussed freely.
+Science should be publicly debated otherwise it isn't science.\\
+\addlinespace
+165 & it was interesting\\
+166 & Everything in the study was good I enjoyed taking it\\
+167 & none\\
+169 & good study\\
+170 & Thanks for the opportunity!\\
+\addlinespace
+173 & No\\
+180 & The study was great.\\
+183 & Very interesting study.\\
+185 & None\\
+186 & Thanks, have  good day\\
+\addlinespace
+189 & no\\
+\bottomrule
+\end{longtable}
+
+\subsection{Additional plots}\label{additional-plots}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-conditional-acceptance.pdf}
+\caption{\label{fig:exp1-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp1-plot-overview.pdf}
+\caption{\label{fig:exp1-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
+\end{figure}
+
+\clearpage
+
+\section{Experiment 2}\label{exp2}
+
+\subsection{Materials}\label{materials-4}
+
+\FloatBarrier
+
+\begin{figure}
+
+\includegraphics[width=1\linewidth]{./figures/study2_question_example} \hfill{}
+
+\caption{Example of a science question, the scientific consensus and corresponding explanation and sources.}\label{fig:exp2-stimulus-example}
+\end{figure}
+
+\subsection{Comparing items}\label{comparing-items-1}
+
+\subsubsection{Conspiracy theories}\label{conspiracy-theories-1}
+
+Table \ref{tab:exp2-correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
+\midrule
+BCTI & 1.00 & 0.66 & 0.63\\
+CMQ & 0.66 & 1.00 & 0.86\\
+SICBS & 0.63 & 0.86 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsubsection{Trust in science}\label{trust-in-science-5}
+
+Table \ref{tab:exp2-correlation-trust} shows the correlations of the three different items measuring trust in science.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-correlation-trust}Correlations of the three different items measuring trust in science}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
+\midrule
+wgm\_sciencegeneral & 1.00 & 0.89 & 0.79\\
+wgm\_scientists & 0.89 & 1.00 & 0.85\\
+pew & 0.79 & 0.85 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures-1}
+
+Table \ref{tab:exp2-correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}
+
+\begin{tabular}{lllllllll}
+\toprule
+std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+0.07 & -6.01 & < .001 & -0.53 & -0.27 & BCTI 
+(main conspiracy measure) & {}[-0.533, -0.27] & -0.40 & NA\\
+0.07 & -2.36 & = 0.019 & -0.31 & -0.03 & CMQ & {}[-0.311, -0.028] & -0.17 & NA\\
+0.07 & -1.91 & = 0.058 & -0.28 & 0.00 & SICBS & {}[-0.28, 0.005] & -0.14 & NA\\
+0.07 & 3.20 & = 0.002 & 0.09 & 0.37 & WGM trust scientists & {}[0.087, 0.368] & 0.23 & NA\\
+0.07 & 3.98 & < .001 & 0.14 & 0.42 & WGM trust general 
+(main trust measure) & {}[0.14, 0.417] & 0.28 & NA\\
+0.07 & 2.23 & = 0.027 & 0.02 & 0.30 & PEW trust scientists & {}[0.019, 0.303] & 0.16 & NA\\
+0.07 & -5.47 & < .001 & -0.50 & -0.24 & BCTI 
+(main conspiracy measure) & {}[-0.504, -0.237] & NA & -0.37\\
+0.07 & -3.03 & = 0.003 & -0.36 & -0.08 & CMQ & {}[-0.356, -0.076] & NA & -0.22\\
+0.07 & -2.76 & = 0.006 & -0.34 & -0.06 & SICBS & {}[-0.338, -0.056] & NA & -0.20\\
+0.07 & 3.51 & < .001 & 0.11 & 0.39 & WGM trust scientists & {}[0.108, 0.387] & NA & 0.25\\
+0.07 & 4.29 & < .001 & 0.16 & 0.44 & WGM trust general 
+(main trust measure) & {}[0.161, 0.436] & NA & 0.30\\
+0.07 & 3.04 & = 0.003 & 0.08 & 0.36 & PEW trust scientists & {}[0.076, 0.357] & NA & 0.22\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses-1}
+
+Table \ref{tab:exp2-false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.
+
+\begin{table}
+
+\caption{\label{tab:exp2-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
+\centering
+\begin{tabular}[t]{lcccccc}
+\toprule
+  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
+\midrule
+(Intercept) & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
+ & (0.081) & (0.080) & (0.081) & (0.082) & (0.082) & \vphantom{1} (0.082)\\
+avg\_acceptance & -0.225** & -0.247** & -0.201* & 0.131 & 0.161+ & 0.130\\
+ & (0.081) & (0.080) & (0.081) & (0.082) & (0.082) & (0.082)\\
+\midrule
+Num.Obs. & 147 & 147 & 147 & 147 & 147 & 147\\
+R2 & 0.051 & 0.061 & 0.041 & 0.017 & 0.026 & 0.017\\
+R2 Adj. & 0.044 & 0.054 & 0.034 & 0.010 & 0.019 & 0.010\\
+AIC & 414.5 & 412.9 & 416.1 & 419.6 & 418.3 & 419.7\\
+BIC & 423.5 & 421.9 & 425.0 & 428.6 & 427.3 & 428.6\\
+Log.Lik. & -204.248 & -203.465 & -205.036 & -206.802 & -206.141 & -206.826\\
+RMSE & 0.97 & 0.97 & 0.98 & 0.99 & 0.98 & 0.99\\
+\bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
+\end{tabular}
+\end{table}
+
+\subsection{By-question variation}\label{by-question-variation-1}
+
+\subsubsection{Knowledge}\label{knowledge-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-questions-knowledge.pdf}
+\caption{\label{fig:exp2-questions-knowledge}Distribution of correct answers by question.}
+\end{figure}
+
+\subsubsection{Acceptance}\label{acceptance-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-questions-acceptance.pdf}
+\caption{\label{fig:exp2-questions-acceptance}Distribution of consensus acceptance by question.}
+\end{figure}
+
+\subsection{Trust in science}\label{trust-in-science-6}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-trust-scientists.pdf}
+\caption{\label{fig:exp2-trust-scientists}Distribution of trust in scientists.}
+\end{figure}
+
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-1}
+
+\subsubsection{Distribution}\label{distribution-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-conspiracy-distribution.pdf}
+\caption{\label{fig:exp2-conspiracy-distribution}Distribution of conspiracy thinking.}
+\end{figure}
+
+\subsubsection{By-item variation}\label{by-item-variation-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-conspiracy-items.pdf}
+\caption{\label{fig:exp2-conspiracy-items}Distribution of conspiracy thinking by item.}
+\end{figure}
+
+\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection}
+
+\subsubsection{By category}\label{by-category}
+
+{[}To do{]}
+
+\subsubsection{All raw answers}\label{all-raw-answers}
+
+\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp2-reasons-rejection}Reasons for consensus rejection}\\
+\toprule
+id & question & answer\\
+\midrule
+33 & sex & Because I didn't think the information provided here was accurate, more or less.\\
+45 & lasers & I read the article.\\
+45 & salt & Read the article provided.\\
+47 & lasers & The statement "lasers do not work by focusing sound waves" is correct according to scientific consensus.
+Lasers operate on the principles of stimulated emission of radiation, wherein photons are emitted in a coherent, focused beam through the process of stimulated emission.\\
+47 & diamonds & Are diamonds made of carbon? Scientific consensus: Diamonds are made of carbon.  Do you think you could explain why you answered differently?\\
+\addlinespace
+47 & water & it likely arose from a mistake or miscommunication.\\
+50 & orbit & I honestly couldn't remember.\\
+50 & speed & When I looked it up I got two different answers.\\
+71 & water & After I put "No", I realized that I was thinking about "H2o" incorrectly. So yes, I actually do agree.\\
+73 & electrons & Atoms are known to be the smallest unit of matter.\\
+\addlinespace
+77 & continents & I believe God make it they way it is and if He wanted it changed, He would have done it\\
+78 & sex & no\\
+78 & salt & no\\
+79 & orbit & It’s early and I’m tired. I misread it.\\
+105 & salt & I think I simply misunderstood the paragraph. I'm actually trying to read it thoroughly and answering truthfully, but I have misread it.\\
+\addlinespace
+109 & electrons & Maybe\\
+127 & continents & I'm a young earth Christian. I believe the Christian Bible (and outside sources) offers ample evidence that the earth is thousands of years old, not millions.\\
+127 & orbit & I do not believe in the heliocentric model of an earth spinning around a centralized sun.\\
+130 & continents & I don't think the continent move rather it is the atmosphere\\
+130 & sex & I think that the gene both parent release should determine the sex of the baby\\
+\addlinespace
+131 & antibiotics & There are some antibiotics that also kill viruses\\
+132 & salt & Because it says that table salt is not made of calcium carbonate, so I clicked no.\\
+133 & lasers & I feel like the excerpt that was provided did not definitively state the answer.\\
+140 & lasers & It was confusing to understand the explanation of why lasers don’t focus on sound\\
+152 & continents & I think that they were that way after the ice age.
+It split apart the whole surface and then water filled in and they have remained the same for ever.
+They are not moving around or have they for millions of years.\\
+\addlinespace
+152 & orbit & The  Earth circles the sun every day.  The dark side is when the moon comes out.  So everyday, you have daylight and dark (moon).  Dark comes from the dark side of the sun.\\
+170 & lasers & I may have misread, but I did not see confirmation of this.\\
+172 & salt & no\\
+178 & electrons & This answer is different because atoms is everywhere and you can't see it but electrons aren't out that as much at all but you could possibly know it's there\\
+178 & antibiotics & Because antibiotics can kill bacteria in which kills the virus as well in the process so you get healed with this product\\
+\addlinespace
+178 & sex & There is no 100\% evidence that this is the case and I still fully believe that it is from both genes to make this decision.\\
+179 & antibiotics & Antibiotics seem that if they are capable of killing tiny things such as bacteria then they are most likely to attach to other things in out bodies and cause it harm although unintentional.\\
+183 & salt & I thought the scientific consensus was right\\
+186 & sex & The scientific consensus is clear: both parents' genes determine the sex of a baby, specifically the sex chromosomes they contribute.\\
+189 & salt & I made an error in my selection.\\
+\bottomrule
+\end{longtable}
+
+\subsection{Additional plots}\label{additional-plots-1}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp2-conditional-acceptance.pdf}
+\caption{\label{fig:exp2-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+\clearpage
+
+\section{Experiment 3}\label{exp3}
+
+\subsection{Materials}\label{materials-5}
+
+\FloatBarrier
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{figures/study3_example_explanation}
+
+}
+
+\caption{Example of an explanation question and the opportunity to change the previous answer.}\label{fig:exp3-explanation-example}
+\end{figure}
+
+Table \ref{tab:knowledge-overview} shows all questions, their scientifically consensual answer, and their source.
+All but two questions were selected from existing science knowledge questionnaires.
+We tried to select non-political questions.
+
+\begingroup\fontsize{8}{10}\selectfont
+
 \begin{longtable}[t]{>{\raggedleft\arraybackslash}p{1em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{23em}}
-\caption{\label{tab:knowledge}Science knowledge items}\\
+\caption{\label{tab:knowledge-overview}Science knowledge items}\\
 \toprule
@@
 
-\paragraph{Conspiracy scales}\label{conspiracy-scales}
-
-To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by Pennycook, Binnendyk, and Rand (2022) (Table \ref{tab:conspiracy}).
-Participants were asked: ``Below is a list of events for which the official version has been disputed.
-For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
-{[}1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true{]}''.
-
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:conspiracy}Conspiracy items}\\
-\toprule
-1 & The Apollo moon landings never happened and were staged in a Hollywood film studio.\\
-2 & A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).\\
-3 & The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.\\
-4 & The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.\\
-5 & The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.\\
-\addlinespace
-6 & There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.\\
-7 & In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.\\
-8 & Proof of alien contact is being concealed from the public.\\
-9 & Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.\\
-10 & Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.\\
-\bottomrule
-\end{longtable}
-\endgroup{}
-
-To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix \ref{exp1}).
-
-\paragraph{Trust in science}\label{trust-in-science-1}
-
-Our main item for measuring trust in science is selected from the Wellcome Global Monitor survey: ``In general, would you say that you trust science a lot, some, not much, or not at all?
-{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''
-
-We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey (``How much do you trust scientists in this country?
-Do you trust them a lot, some, not much, or not at all?
-{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''), the other from the Pew research center (``How much confidence do you have in scientists to act in the best interests of the public?
-{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}'').
-We selected these items so that we could compare the ratings in our sample to global survey results.
-The WGM survey has been administered in over 140 countries and included over 140000 respondents.
-The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents (Cologna et al., 2024).
-
-\subsection{Results}\label{results}
-
-Regarding RQ1 and RQ2, participants answered on average 74 \% (sd = 0.16) of the questions correctly, and accepted the scientific consensus on average for 93 \% (sd = 0.12) of the questions.
-
-Fig.
-\ref{fig:exp1-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
-In most cases (76.3 \%), participants readily accepted the scientific consensus after having given the wrong answer to a question.
-In very few cases (1.6 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
-We believe this might have been due to inattention.
-
-For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29, p \textless{} .001), and acceptance of scientific consensus and trust in science (r = 0.27, p \textless{} .001).
-The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp1-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-
-For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = -0.38, p \textless{} .001), and conspiracy thinking and acceptance of scientific consensus (r = -0.33, p \textless{} .001).
-
-In Appendix \ref{exp1}, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \ref{exp1} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus?
-In our main analyses, we looked at correlations of acceptance across all observations.
-One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus.
-To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question.
-We than calculated the correlation between trust in science and average acceptance rate by participant.
-We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = -0.14, p = 0.061), nor with trust in science (r = 0.06, p = 0.387).
-
-\subsection{Discussion}\label{discussion}
-
-These results suggest that most people accept the scientific consensus most of the time.
-Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
-Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
-In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
-\section{Experiment 2}\label{experiment-2}
-
-In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
-In most instances they did, but not always.
-In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
-To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
-To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
-We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \ref{exp1}).
-
-Based on experiment 1, we formulated the following hypotheses:
-
-\textbf{H1a: Higher trust in science is associated with more science knowledge?}
-
-\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?}
-
-\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
-
-\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?}
-
-We had the following research questions:
-
-\textbf{RQ1: What is the average science knowledge score?}
-
-\textbf{RQ2: When a participant's answer does not match the consensus, how often do they change their mind and accept the consensus?}
-
-\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
-
-\subsection{Methods}\label{methods-1}
-
-\subsubsection{Participants}\label{participants-1}
-
-We recruited 201 participants from the US via prolific.
-11 participants failed our attention check, resulting in a final sample of 190 participants (96 female, 94 male; \(age_\text{mean}\): 43.48, \(age_\text{sd}\): 12.25, \(age_\text{median}\): 42).
-Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
-
-\subsubsection{Procedure}\label{procedure-1}
-
-The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
-\ref{fig:exp2-stimulus-example}.
-
-\subsubsection{Materials}\label{materials-1}
-
-We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
-
-\subsection{Results}\label{results-1}
-
-As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = 0.28, p \textless{} .001) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = -0.40, p \textless{} .001).
-By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
-For trust in science, we find no statistically significant correlation (r = 0.16, p = 0.051).
-For conspiracy thinking we find a small negative one (r = -0.22, p = 0.006).
-
-Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp2-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-In Appendix \ref{exp2} we show that these results hold for our alternative measures of trust and conspiracy thinking.
-We also include more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Regarding RQ1, participants answered on average 79 \% (sd = 0.19) of the questions correctly, and accepted the scientific consensus on average for 98 \% (sd = 0.05) of the questions.
-Fig.
-\ref{fig:exp2-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
-In response to RQ2, in most cases (92.9 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-In very few cases (0.5 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
-
-For RQ3, we got 35 answers from 25 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Table \ref{tab:exp2-justifications} summarizes these answers by five categories.
-All answers can be read in Appendix \ref{exp2}.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp2-justifications}Justifications by category}
-
-\begin{tabular}{llll}
-\toprule
-Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
-\midrule
-Personal convictions & 12.00 & 34.3\% & 8.00\\
-Mistake & 9.00 & 25.7\% & 8.00\\
-No justification & 7.00 & 20\% & 5.00\\
-Not convinced & 5.00 & 14.3\% & 5.00\\
-Religious Beliefs & 2.00 & 5.7\% & 2.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsection{Discussion}\label{discussion-1}
-
-Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.~answered the knowledge question wrongly).
-The share of these latter is considerably larger in experiment 2 (92.9 \%) than in experiment 1 (76.3 \%).
-While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
-We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
-\section{Experiment 3}\label{experiment-3}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/summary-plot.pdf}
-\caption{\label{fig:summary-plot}Points represent the average share of acceptance as a function of the \textbf{A} level of trust in science (``In general, would you say that you trust science a lot, some, not much, or not at all?
-{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''); \textbf{B} average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); \textbf{C} belief in a specific conspiracy theory.
-Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: ``1 - completly false, 5 - unsure, 9 - completely true''.
-Averages are only based on participants who believed at least to some degree in the conspiracy theory.
-We consider a believer as anyone scoring higher than the scale midpoint, i.e.~\textgreater5.
-Number labels in plots represent the number of participants for the corresponding data point.}
-\end{figure}
-
-Study 3 is essentially a replication-\/--with some minor modifications-\/--of study 2, but on a different type of sample.
-Both study 1 and 2 were run on convenience samples.
-For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
-By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
-\ref{fig:exp3-explanation-example}).
-Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
-Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
-We want to know if participants perceive that this is because of trust, or other factors.
-
-As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
-
-\textbf{H1a: Higher trust in science is associated with more science knowledge?}
-
-\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus.}
-
-\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
-
-\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.}
-
-We had the following research questions:
-
-\textbf{RQ1: What is the average science knowledge score?}
-
-\textbf{RQ2: What is the average acceptance of the scientific consensus}
-
-\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
-
-\textbf{RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?}
-
-\subsection{Methods}\label{methods-2}
-
-\subsubsection{Participants}\label{participants-2}
-
-We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of 200 participants (125 female, 73 male; \(age_\text{mean}\): 42.90, \(age_\text{sd}\): 12.00, \(age_\text{median}\): 41).
-Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
-
-However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
-This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = 140 for trust in science, N = 138 for conspiracy measures, N = 122 for reason for accepting consensus).
-
-\subsubsection{Procedure}\label{procedure-2}
-
-The procedure was mostly the same as in experiment 2.
-In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
-Finally, at the end of the survey, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
-Participants who selected ``You mostly agree with the consensus because you have been able to independently verify it'', were asked the open-ended follow-up question: ``Could you please tell us how you independently verified the information?''.
-
-\subsubsection{Materials}\label{materials-2}
-
-We relied on the same items as in experiment 2.
-We used a pre-registered categorization scheme for the open ended answers for why people reject the consensus.
-The categories were XX and XX.
-{[}Describe coding process, with independent reviewer etc.{]}
-
-\subsection{Results}\label{results-2}
-
-As in study 1 and 2, we find that participants answered on average 75 \% (sd = 0.18) of the questions correctly, and initially accepted the scientific consensus on average for 96 \% (sd = 0.08) of the questions (RQ2).
-The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (98 \%, sd = 0.06).
-
-Fig.
-\ref{fig:exp3-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
-In most cases (87 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
-After providing the chance to revise the initial consensus rejection, this share is even larger (95.5 \%).
-In very few cases (0.9 \%), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
-This share drops slightly after providing the chance to revise the initial consensus rejection (0.7 \%).
-
-In the appendix XX we provide an analysis that this drop is statistically significant {[}TO DO{]}.
-
-For all correlations, we include opinion revisions for measuring consensus acceptance.
-We find no statistically significant correlation between science knowledge and trust in science (r = 0.14, p = 0.100), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = 0.20, p = 0.019).
-Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp3-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-We find no statistically significant correlation between conspiracy thinking and science knowledge (r = -0.16, p = 0.055), and between conspiracy thinking and acceptance of scientific consensus (r = c(cor = -0.023), c(t = -0.27), = 0.788, c(df = 136), -0.189, 0.145, Pearson's product-moment correlation, two.sided, {[}-0.189, 0.145{]}, p = 0.788).
-
-In Appendix \ref{exp3}, we show that these associations are statistically significant (except the last one?).
-We show that they are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \ref{exp3} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Regarding RQ1, participants answered on average 75 \% (sd = 0.18) of the questions correctly.
-For RQ3, we got 74 answers from 47 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Table \ref{tab:exp3-justifications} summarizes these answers by five categories.
-
-All answers can be read in the appendix.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp3-justifications}Justifications by category}
-
-\begin{tabular}{llll}
-\toprule
-Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
-\midrule
-Personal convictions & 25.00 & 33.8\% & 19.00\\
-Mistake & 21.00 & 28.4\% & 18.00\\
-No justification & 18.00 & 24.3\% & 13.00\\
-Not convinced & 7.00 & 9.5\% & 3.00\\
-Religious Beliefs & 3.00 & 4.1\% & 3.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-For RQ4, we had 122 participants answering the question.
-Of these 41.8\% said they accepted the scientific consensus because they trust scientists on this question, while 47.5\% said they independently verified the fact.
-10.7\% answered with other ``other'' and gave an open-ended explanation (see Appendix).
-
-We also asked all 58 participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed in Appendix XX.
-
-\section{Discussion}\label{discussion-2}
-
-{[}We can draw some inspiration from the ``Westwood et al 2021 Current research overstates American support for political violence'' paper: we want to put in perspective the current rise (?
-discourse around?) in mistrust towards science{]}
-
-There is increasingly more talk of defiance towards science and towards experts more generally.
-In every population, many people only trust science `some of the time', and a sizable minority doesn't trust it.
-This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
-
-But how deep is distrust of science? In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
-
-Traditionally, people have interpreted people's lack of basic knowledge in light of the deficit-model. Here, we show that people readily accept the scientifically consensual answer to these questions.
-
-``The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
-Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
-Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
-The results of the U.S.
-portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
-Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators' questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).''
-
-\FloatBarrier
-
-\section{References}\label{references}
-
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-alganTrustScientistsTimes2021}
-Algan, Y., Cohen, D., Davoine, E., Foucault, M., \& Stantcheva, S.
-(2021).
-Trust in scientists in times of pandemic: Panel evidence from 12 countries.
-\emph{Proceedings of the National Academy of Sciences}, \emph{118}(40), e2108576118.
-\url{https://doi.org/10.1073/pnas.2108576118}
-
-\bibitem[\citeproctext]{ref-allumScienceKnowledgeAttitudes2008}
-Allum, N., Sturgis, P., Tabourazi, D., \& Brunton-Smith, I.
-(2008).
-Science knowledge and attitudes across cultures: a meta-analysis.
-\emph{Public Understanding of Science}, \emph{17}(1), 35--54.
-\url{https://doi.org/10.1177/0963662506070159}
-
-\bibitem[\citeproctext]{ref-bruderMeasuringIndividualDifferences2013}
-Bruder, M., Haffke, P., Neave, N., Nouripanah, N., \& Imhoff, R.
-(2013).
-Measuring individual differences in generic beliefs in conspiracy theories across cultures: Conspiracy mentality questionnaire.
-\emph{Frontiers in Psychology}, \emph{4}.
-Retrieved from \url{https://www.frontiersin.org/articles/10.3389/fpsyg.2013.00225}
-
-\bibitem[\citeproctext]{ref-colognaTrustScientistsTheir2024}
-Cologna, V., Mede, N.
-G., Berger, S., Besley, J., Brick, C., Joubert, M., \ldots{} Linden, D.
-S.
-van der.
-(2024).
-\emph{Trust in scientists and their role in society across 67 countries}.
-\url{https://doi.org/10.31219/osf.io/6ay7s}
-
-\bibitem[\citeproctext]{ref-durantPublicUnderstandingScience1989}
-Durant, J. R., Evans, G. A., \& Thomas, G. P. (1989). The public understanding of science. \emph{Nature}, \emph{340}(6228), 11--14. \url{https://doi.org/10.1038/340011a0}
-
-\bibitem[\citeproctext]{ref-funkPublicConfidenceScientists2020}
-Funk, C., \& Kennedy, B. (2020). \emph{Public confidence in scientists has remained stable for decades}.
-
-\bibitem[\citeproctext]{ref-gauchatCulturalAuthorityScience2011}
-Gauchat, G.
-(2011).
-The cultural authority of science: Public trust and acceptance of organized science.
-\emph{Public Understanding of Science}, \emph{20}(6), 751--770.
-\url{https://doi.org/10.1177/0963662510365246}
-
-\bibitem[\citeproctext]{ref-gauchatPoliticizationSciencePublic2012}
-Gauchat, G.
-(2012).
-Politicization of Science in the Public Sphere: A Study of Public Trust in the United States, 1974 to 2010.
-\emph{American Sociological Review}, \emph{77}(2), 167--187.
-\url{https://doi.org/10.1177/0003122412438225}
-
-\bibitem[\citeproctext]{ref-kennedyWhatDataSays2023}
-Kennedy, A.
-T.
-C.
-F.
-and B.
-(2023).
-\emph{What the data says about americans{'} views of climate change}.
-Retrieved from \url{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}
-
-\bibitem[\citeproctext]{ref-krauseTrendsAmericansTrust2019a}
-Krause, N.
-M., Brossard, D., Scheufele, D.
-A., Xenos, M.
-A., \& Franke, K.
-(2019).
-Trends{\textemdash}americans{'} trust in science and scientists.
-\emph{Public Opinion Quarterly}, \emph{83}(4), 817--836.
-\url{https://doi.org/10.1093/poq/nfz041}
-
-\bibitem[\citeproctext]{ref-lantianMeasuringBeliefConspiracy2016}
-Lantian, A., Muller, D., Nurra, C., \& Douglas, K.
-M.
-(2016).
-Measuring Belief in Conspiracy Theories: Validation of a French and English Single-Item Scale.
-\emph{International Review of Social Psychology}, \emph{29}(1), 1.
-\url{https://doi.org/10.5334/irsp.8}
-
-\bibitem[\citeproctext]{ref-liPolarizationPublicTrust2022}
-Li, N., \& Qian, Y.
-(2022).
-Polarization of public trust in scientists between 1978 and 2018: Insights from a cross-decade comparison using interpretable machine learning.
-\emph{Politics and the Life Sciences}, \emph{41}(1), 45--54.
-\url{https://doi.org/10.1017/pls.2021.18}
-
-\bibitem[\citeproctext]{ref-millerMeasurementCivicScientific1998a}
-Miller, J. D. (1998). The measurement of civic scientific literacy. \emph{Public Understanding of Science}, \emph{7}(3), 203--223. \url{https://doi.org/10.1088/0963-6625/7/3/001}
-
-\bibitem[\citeproctext]{ref-millerPublicUnderstandingAttitudes2004}
-Miller, J.
-D.
-(2004).
-Public Understanding of, and Attitudes toward, Scientific Research: What We Know and What We Need to Know.
-\emph{Public Understanding of Science}, \emph{13}(3), 273--294.
-\url{https://doi.org/10.1177/0963662504044908}
-
-\bibitem[\citeproctext]{ref-mottaChangingMindsChanging2021}
-Motta, M.
-(2021).
-Changing Minds or Changing Samples?
-Disentangling Microlevel Stability and Macrolevel Growth in Anthropogenic Climate Change Beliefs.
-\emph{International Journal of Public Opinion Research}, \emph{33}(2), 477--489.
-\url{https://doi.org/10.1093/ijpor/edaa020}
-
-\bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
-Pennycook, G., Binnendyk, J., \& Rand, D.
-(2022).
-\emph{Overconfidently conspiratorial: Conspiracy believers are dispositionally overconfident and massively overestimate how much others agree with them}.
-\url{https://doi.org/10.31234/osf.io/d5fz2}
-
-\bibitem[\citeproctext]{ref-rutjensConspiracyBeliefsScience2022}
-Rutjens, B. T., \& Većkalov, B. (2022). Conspiracy beliefs and science rejection. \emph{Current Opinion in Psychology}, \emph{46}, 101392. \url{https://doi.org/10.1016/j.copsyc.2022.101392}
-
-\bibitem[\citeproctext]{ref-committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016}
-Science Literacy, C.
-on, Public Perception of Science, Board on Science Education, Behavioral, D.
-of, Sciences, S., Education, \& National Academies of Sciences, Engineering, and Medicine.
-(2016).
-\emph{Science Literacy: Concepts, Contexts, and Consequences} (C.
-E.
-Snow \& K.
-A.
-Dibner, Eds.).
-Washington, D.C.: National Academies Press.
-\url{https://doi.org/10.17226/23595}
-
-\bibitem[\citeproctext]{ref-vranicDidMyOwn2022}
-Vranic, A., Hromatko, I., \& Tonković, M.
-(2022).
-{``}I did my own research{''}: Overconfidence, (dis)trust in science, and endorsement of conspiracy theories.
-\emph{Frontiers in Psychology}, \emph{13}.
-\url{https://doi.org/10.3389/fpsyg.2022.931865}
-
-\bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2018}
-Wellcome Global Monitor. (2018). \emph{Wellcome Global Monitor 2018}. Retrieved from \url{https://wellcome.org/reports/wellcome-global-monitor/2018}
-
-\bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2020}
-Wellcome Global Monitor. (2020). \emph{Wellcome Global Monitor 2020: Covid-19}. Retrieved from \url{https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020}
-
-\end{CSLReferences}
-
-\newpage
-
-\appendix
-
-\section{Experiment 1}\label{exp1}
-
-\subsection{Materials}\label{materials-3}
-
-\FloatBarrier
-
-\begin{figure}
-
-\includegraphics[width=1\linewidth]{./figures/study1_question_example} \hfill{}
-
-\caption{Example of a science question, the scientific consensus and the corresponding acceptance question.}\label{fig:stimulus-example}
-\end{figure}
-
-\subsubsection{Conspiracy scales}\label{conspiracy-scales-1}
-
-Beside Belief in Conspiracy Theory Inventory (BCTI) by Pennycook et al. (2022) which report in the main study, we also assessed two other consipiracy thinking measures:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  The conspiracy mentality questionnaire (CMQ) by Bruder et al. (2013):
-  I think that . . .
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  \ldots{} many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
-\item
-  \ldots{} government agencies closely monitor all citizens.
-\item
-  \ldots{} events which superficially seem to lack a connection are often the result of secret activities.
-\item
-  \ldots{} there are secret organizations that greatly influence political decisions.
-\end{itemize}
-
-{[}0\% - 100\%; 0 = certainly not, 100 = certain{]}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  The Single Item Conspiracy Beliefs Scale (SICBS) by Lantian et al. (2016) :
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  I think that the official version of the events given by the authorities very often hides the truth. {[}1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true{]}
-\end{itemize}
-
-\subsubsection{Trust in science}\label{trust-in-science-2}
-
-We rely on three items
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
-\item
-  In general, would you say that you trust science a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
-\item
-  How much confidence do you have in scientists to act in the best interests of the public? {[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}
-\end{enumerate}
-
-\subsection{Comparing items}\label{comparing-items}
-
-\subsubsection{Conspiracy theories}\label{conspiracy-theories}
-
-Table \ref{tab:correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.
-
-\begin{table}[h]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}
-
-\begin{tabular}{llll}
-\toprule
- & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
-\midrule
-BCTI & 1.00 & 0.58 & 0.56\\
-CMQ & 0.58 & 1.00 & 0.77\\
-SICBS & 0.56 & 0.77 & 1.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsubsection{Trust in science}\label{trust-in-science-3}
-
-Table \ref{tab:correlation-trust} shows the correlations of the three different items measuring trust in science.
-
-\begin{table}[h]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:correlation-trust}Correlations of the three different items measuring trust in science}
-
-\begin{tabular}{llll}
-\toprule
- & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
-\midrule
-wgm\_sciencegeneral & 1.00 & 0.85 & 0.75\\
-wgm\_scientists & 0.85 & 1.00 & 0.82\\
-pew & 0.75 & 0.82 & 1.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures}
-
-Table \ref{tab:correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}
-
-\begin{tabular}{lll}
-\toprule
-outcome & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
-\midrule
-BCTI 
-(main conspiracy measure) & -0.38 (p< .001) & -0.33 (p< .001)\\
-CMQ & -0.11 (p= 0.121) & -0.07 (p= 0.317)\\
-SICBS & -0.1 (p= 0.158) & -0.09 (p= 0.213)\\
-WGM trust scientists & 0.23 (p= 0.001) & 0.26 (p< .001)\\
-WGM trust general 
-(main trust measure) & 0.29 (p< .001) & 0.27 (p< .001)\\
-PEW trust scientists & 0.16 (p= 0.027) & 0.21 (p= 0.003)\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses}
-
-Table \ref{tab:false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.
-
-\begin{table}
-
-\caption{\label{tab:false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
-\centering
-\begin{tabular}[t]{lcccccc}
-\toprule
-  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
-\midrule
-(Intercept) & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
- & (0.073) & (0.074) & (0.074) & (0.073) & (0.074) & (0.074)\\
-avg\_acceptance & -0.139+ & 0.008 & -0.028 & 0.108 & 0.064 & 0.056\\
- & (0.073) & (0.074) & (0.074) & (0.074) & (0.074) & (0.074)\\
-\midrule
-Num.Obs. & 184 & 184 & 184 & 184 & 184 & 184\\
-R2 & 0.019 & 0.000 & 0.001 & 0.012 & 0.004 & 0.003\\
-R2 Adj. & 0.014 & -0.005 & -0.005 & 0.006 & -0.001 & -0.002\\
-AIC & 523.6 & 527.2 & 527.0 & 525.0 & 526.4 & 526.6\\
-BIC & 533.2 & 536.8 & 536.7 & 534.7 & 536.1 & 536.2\\
-Log.Lik. & -258.801 & -260.578 & -260.512 & -259.506 & -260.204 & -260.295\\
-RMSE & 0.99 & 1.00 & 1.00 & 0.99 & 1.00 & 1.00\\
-\bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
-\end{tabular}
-\end{table}
-
-\subsection{By-question variation}\label{by-question-variation}
-
-\subsubsection{Knowledge}\label{knowledge}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-questions-knowledge.pdf}
-\caption{\label{fig:exp1-questions-knowledge}Distribution of correct answers by question.}
-\end{figure}
-
-\subsubsection{Acceptance}\label{acceptance}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-questions-acceptance.pdf}
-\caption{\label{fig:exp1-questions-acceptance}Distribution of consensus acceptance by question.}
-\end{figure}
-
-\subsection{Trust in science}\label{trust-in-science-4}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-trust-scientists.pdf}
-\caption{\label{fig:exp1-trust-scientists}Distribution of trust in scientists.}
-\end{figure}
-
-\subsection{Conspiracy thinking}\label{conspiracy-thinking}
-
-\subsubsection{Distribution}\label{distribution}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-conspiracy-distribution.pdf}
-\caption{\label{fig:exp1-conspiracy-distribution}Distribution of conspiracy thinking.}
-\end{figure}
-
-\subsubsection{By-item variation}\label{by-item-variation}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-conspiracy-items.pdf}
-\caption{\label{fig:exp1-conspiracy-items}Distribution of conspiracy thinking by item.}
-\end{figure}
-
-\subsection{Open ended comments}\label{open-ended-comments}
-
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:unnamed-chunk-21}Reasons for consensus rejection}\\
-\toprule
-id & comments\\
-\midrule
-1 & no\\
-20 & None\\
-22 & I answered thoughtfully, good luck!\\
-27 & interesting, thank you.\\
-30 & n/a\\
-\addlinespace
-32 & None\\
-39 & no\\
-41 & none\\
-42 & n/a\\
-67 & thanks\\
-\addlinespace
-69 & No comments\\
-74 & no\\
-76 & none\\
-78 & None - thank you!\\
-79 & Should've clicked agree about plants getting carbon from the air, but I wanted to look it up first.
-Whenever I encounter something I didn't know online, I like to cross-reference.
-In retrospect though, everything else you asked about was simple and on the level, and I'm sure you are correct.
-
-Thanks for your hard work.\\
-\addlinespace
-84 & No, thank you.\\
-86 & n/a\\
-89 & N/A\\
-95 & no\\
-98 & Great topics! Thank you for this study.\\
-\addlinespace
-109 & To provide context for my responses, I currently study physiology (STEM) and perform academic research as an undergrad at my university.\\
-110 & n/a\\
-114 & Everything in the survey went fine with no issues.
-We seem to have a lot of conspiracy theory nuts in the world today and I think it's a sign of the larger mental health crisis that's currently going on.\\
-115 & I like the subject content of the survey\\
-116 & No\\
-\addlinespace
-132 & None.\\
-133 & very interesting\\
-135 & Nope\\
-139 & none\\
-140 & Air AND water make up the bulk of trees mass and the water is drawn from the earth.
-Water has much more mass than air and I stand by my answer of Earth being that is where the water is drawn from.
-The question was very vague and as you see...
-open to interpretation\\
-\addlinespace
-141 & n/a\\
-148 & no\\
-150 & no\\
-152 & None.\\
-154 & none, tysm!\\
-\addlinespace
-155 & Not virus, just metals/DDT. And I think it would be impossible for dinosaurs to have sex with the way they are shaped.\\
-157 & none\\
-159 & I think scientists are acting in the best interests of science rather than the community directly. An unethical scientist could advance scientific knowledge while hurting humanity.\\
-161 & Mo Thanks\\
-164 & I appreciate science but understand that if scientists don't go along with what investors/govt want they won't receive grants for funding.
-I also very recently lived through several years of gaslighting about covid.
-While I realize I'm not the sharpest tool in the shed, I'm not a total idiot.
-Things that were censored at the beginning of the pandemic are now allowed to be discussed freely.
-Science should be publicly debated otherwise it isn't science.\\
-\addlinespace
-165 & it was interesting\\
-166 & Everything in the study was good I enjoyed taking it\\
-167 & none\\
-169 & good study\\
-170 & Thanks for the opportunity!\\
-\addlinespace
-173 & No\\
-180 & The study was great.\\
-183 & Very interesting study.\\
-185 & None\\
-186 & Thanks, have  good day\\
-\addlinespace
-189 & no\\
-\bottomrule
-\end{longtable}
-
-\subsection{Additional plots}\label{additional-plots}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-conditional-acceptance.pdf}
-\caption{\label{fig:exp1-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-plot-overview.pdf}
-\caption{\label{fig:exp1-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
-\end{figure}
-
-\clearpage
-
-\section{Experiment 2}\label{exp2}
-
-\subsection{Materials}\label{materials-4}
-
-\FloatBarrier
-
-\begin{figure}
-
-\includegraphics[width=1\linewidth]{./figures/study2_question_example} \hfill{}
-
-\caption{Example of a science question, the scientific consensus and corresponding explanation and sources.}\label{fig:exp2-stimulus-example}
-\end{figure}
-
-\subsection{Comparing items}\label{comparing-items-1}
-
-\subsubsection{Conspiracy theories}\label{conspiracy-theories-1}
-
-Table \ref{tab:exp2-correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.
-
-\begin{table}[h]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp2-correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}
-
-\begin{tabular}{llll}
-\toprule
- & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
-\midrule
-BCTI & 1.00 & 0.66 & 0.63\\
-CMQ & 0.66 & 1.00 & 0.86\\
-SICBS & 0.63 & 0.86 & 1.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsubsection{Trust in science}\label{trust-in-science-5}
-
-Table \ref{tab:exp2-correlation-trust} shows the correlations of the three different items measuring trust in science.
-
-\begin{table}[h]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp2-correlation-trust}Correlations of the three different items measuring trust in science}
-
-\begin{tabular}{llll}
-\toprule
- & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
-\midrule
-wgm\_sciencegeneral & 1.00 & 0.89 & 0.79\\
-wgm\_scientists & 0.89 & 1.00 & 0.85\\
-pew & 0.79 & 0.85 & 1.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures-1}
-
-Table \ref{tab:exp2-correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp2-correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}
-
-\begin{tabular}{lllllllll}
-\toprule
-std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
-\midrule
-0.07 & -6.01 & < .001 & -0.53 & -0.27 & BCTI 
-(main conspiracy measure) & {}[-0.533, -0.27] & -0.40 & NA\\
-0.07 & -2.36 & = 0.019 & -0.31 & -0.03 & CMQ & {}[-0.311, -0.028] & -0.17 & NA\\
-0.07 & -1.91 & = 0.058 & -0.28 & 0.00 & SICBS & {}[-0.28, 0.005] & -0.14 & NA\\
-0.07 & 3.20 & = 0.002 & 0.09 & 0.37 & WGM trust scientists & {}[0.087, 0.368] & 0.23 & NA\\
-0.07 & 3.98 & < .001 & 0.14 & 0.42 & WGM trust general 
-(main trust measure) & {}[0.14, 0.417] & 0.28 & NA\\
-0.07 & 2.23 & = 0.027 & 0.02 & 0.30 & PEW trust scientists & {}[0.019, 0.303] & 0.16 & NA\\
-0.07 & -5.47 & < .001 & -0.50 & -0.24 & BCTI 
-(main conspiracy measure) & {}[-0.504, -0.237] & NA & -0.37\\
-0.07 & -3.03 & = 0.003 & -0.36 & -0.08 & CMQ & {}[-0.356, -0.076] & NA & -0.22\\
-0.07 & -2.76 & = 0.006 & -0.34 & -0.06 & SICBS & {}[-0.338, -0.056] & NA & -0.20\\
-0.07 & 3.51 & < .001 & 0.11 & 0.39 & WGM trust scientists & {}[0.108, 0.387] & NA & 0.25\\
-0.07 & 4.29 & < .001 & 0.16 & 0.44 & WGM trust general 
-(main trust measure) & {}[0.161, 0.436] & NA & 0.30\\
-0.07 & 3.04 & = 0.003 & 0.08 & 0.36 & PEW trust scientists & {}[0.076, 0.357] & NA & 0.22\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses-1}
-
-Table \ref{tab:exp2-false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.
-
-\begin{table}
-
-\caption{\label{tab:exp2-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
-\centering
-\begin{tabular}[t]{lcccccc}
-\toprule
-  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
-\midrule
-(Intercept) & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
- & (0.081) & (0.080) & (0.081) & (0.082) & (0.082) & \vphantom{1} (0.082)\\
-avg\_acceptance & -0.225** & -0.247** & -0.201* & 0.131 & 0.161+ & 0.130\\
- & (0.081) & (0.080) & (0.081) & (0.082) & (0.082) & (0.082)\\
-\midrule
-Num.Obs. & 147 & 147 & 147 & 147 & 147 & 147\\
-R2 & 0.051 & 0.061 & 0.041 & 0.017 & 0.026 & 0.017\\
-R2 Adj. & 0.044 & 0.054 & 0.034 & 0.010 & 0.019 & 0.010\\
-AIC & 414.5 & 412.9 & 416.1 & 419.6 & 418.3 & 419.7\\
-BIC & 423.5 & 421.9 & 425.0 & 428.6 & 427.3 & 428.6\\
-Log.Lik. & -204.248 & -203.465 & -205.036 & -206.802 & -206.141 & -206.826\\
-RMSE & 0.97 & 0.97 & 0.98 & 0.99 & 0.98 & 0.99\\
-\bottomrule
-\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
-\end{tabular}
-\end{table}
-
-\subsection{By-question variation}\label{by-question-variation-1}
-
-\subsubsection{Knowledge}\label{knowledge-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-questions-knowledge.pdf}
-\caption{\label{fig:exp2-questions-knowledge}Distribution of correct answers by question.}
-\end{figure}
-
-\subsubsection{Acceptance}\label{acceptance-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-questions-acceptance.pdf}
-\caption{\label{fig:exp2-questions-acceptance}Distribution of consensus acceptance by question.}
-\end{figure}
-
-\subsection{Trust in science}\label{trust-in-science-6}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-trust-scientists.pdf}
-\caption{\label{fig:exp2-trust-scientists}Distribution of trust in scientists.}
-\end{figure}
-
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-1}
-
-\subsubsection{Distribution}\label{distribution-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-conspiracy-distribution.pdf}
-\caption{\label{fig:exp2-conspiracy-distribution}Distribution of conspiracy thinking.}
-\end{figure}
-
-\subsubsection{By-item variation}\label{by-item-variation-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-conspiracy-items.pdf}
-\caption{\label{fig:exp2-conspiracy-items}Distribution of conspiracy thinking by item.}
-\end{figure}
-
-\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection}
-
-\subsubsection{By category}\label{by-category}
-
-{[}To do{]}
-
-\subsubsection{All raw answers}\label{all-raw-answers}
-
-\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp2-reasons-rejection}Reasons for consensus rejection}\\
-\toprule
-id & question & answer\\
-\midrule
-33 & sex & Because I didn't think the information provided here was accurate, more or less.\\
-45 & lasers & I read the article.\\
-45 & salt & Read the article provided.\\
-47 & lasers & The statement "lasers do not work by focusing sound waves" is correct according to scientific consensus.
-Lasers operate on the principles of stimulated emission of radiation, wherein photons are emitted in a coherent, focused beam through the process of stimulated emission.\\
-47 & diamonds & Are diamonds made of carbon? Scientific consensus: Diamonds are made of carbon.  Do you think you could explain why you answered differently?\\
-\addlinespace
-47 & water & it likely arose from a mistake or miscommunication.\\
-50 & orbit & I honestly couldn't remember.\\
-50 & speed & When I looked it up I got two different answers.\\
-71 & water & After I put "No", I realized that I was thinking about "H2o" incorrectly. So yes, I actually do agree.\\
-73 & electrons & Atoms are known to be the smallest unit of matter.\\
-\addlinespace
-77 & continents & I believe God make it they way it is and if He wanted it changed, He would have done it\\
-78 & sex & no\\
-78 & salt & no\\
-79 & orbit & It’s early and I’m tired. I misread it.\\
-105 & salt & I think I simply misunderstood the paragraph. I'm actually trying to read it thoroughly and answering truthfully, but I have misread it.\\
-\addlinespace
-109 & electrons & Maybe\\
-127 & continents & I'm a young earth Christian. I believe the Christian Bible (and outside sources) offers ample evidence that the earth is thousands of years old, not millions.\\
-127 & orbit & I do not believe in the heliocentric model of an earth spinning around a centralized sun.\\
-130 & continents & I don't think the continent move rather it is the atmosphere\\
-130 & sex & I think that the gene both parent release should determine the sex of the baby\\
-\addlinespace
-131 & antibiotics & There are some antibiotics that also kill viruses\\
-132 & salt & Because it says that table salt is not made of calcium carbonate, so I clicked no.\\
-133 & lasers & I feel like the excerpt that was provided did not definitively state the answer.\\
-140 & lasers & It was confusing to understand the explanation of why lasers don’t focus on sound\\
-152 & continents & I think that they were that way after the ice age.
-It split apart the whole surface and then water filled in and they have remained the same for ever.
-They are not moving around or have they for millions of years.\\
-\addlinespace
-152 & orbit & The  Earth circles the sun every day.  The dark side is when the moon comes out.  So everyday, you have daylight and dark (moon).  Dark comes from the dark side of the sun.\\
-170 & lasers & I may have misread, but I did not see confirmation of this.\\
-172 & salt & no\\
-178 & electrons & This answer is different because atoms is everywhere and you can't see it but electrons aren't out that as much at all but you could possibly know it's there\\
-178 & antibiotics & Because antibiotics can kill bacteria in which kills the virus as well in the process so you get healed with this product\\
-\addlinespace
-178 & sex & There is no 100\% evidence that this is the case and I still fully believe that it is from both genes to make this decision.\\
-179 & antibiotics & Antibiotics seem that if they are capable of killing tiny things such as bacteria then they are most likely to attach to other things in out bodies and cause it harm although unintentional.\\
-183 & salt & I thought the scientific consensus was right\\
-186 & sex & The scientific consensus is clear: both parents' genes determine the sex of a baby, specifically the sex chromosomes they contribute.\\
-189 & salt & I made an error in my selection.\\
-\bottomrule
-\end{longtable}
-
-\subsection{Additional plots}\label{additional-plots-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-conditional-acceptance.pdf}
-\caption{\label{fig:exp2-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
-
-\clearpage
-
-\section{Experiment 3}\label{exp3}
-
-\subsection{Materials}\label{materials-5}
-
-\FloatBarrier
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.5\linewidth]{figures/study3_example_explanation}
-
-}
-
-\caption{Example of an explanation question and the opportunity to change the previous answer.}\label{fig:exp3-explanation-example}
-\end{figure}
-
 \subsection{Comparing items}\label{comparing-items-2}

### 2024-08-13 exp_4/analysis.Rmd
pre-registration study 4

---
title: "Analysis Study 4"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
bibliography: ../references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library(stargazer)
```

```{r, include=FALSE, message=FALSE}
# read the cleaned long version data set
data_long <- read_csv("data/cleaned_long.csv")

# read wide version data set
data_wide <- read_csv("data/cleaned_wide.csv") 
```

```{r}
plot_data <- data_long %>% 
  group_by(knowledge, acceptance) %>% 
  count() %>% 
  group_by(knowledge) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

absolute_numbers <- plot_data %>% 
  group_by(knowledge) %>% 
  summarise(n = sum(n))

ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
  theme_minimal() +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
                   xmax = as.numeric(factor(knowledge)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n)),
               inherit.aes = FALSE,
               step.increase = 0)
```

## New plots

```{r}
plot_data <- data_long %>% 
  drop_na(wgm_sciencegeneral) %>% 
  group_by(wgm_sciencegeneral, acceptance) %>% 
  count() %>% 
  group_by(wgm_sciencegeneral) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

absolute_numbers <- plot_data %>% 
  group_by(wgm_sciencegeneral) %>% 
  summarise(n = sum(n))

ggplot(plot_data, aes(x = wgm_sciencegeneral, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Trust in science", fill = "Acceptance \nof consensus") +
  theme_minimal() +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(wgm_sciencegeneral)) - 0.4, 
                   xmax = as.numeric(factor(wgm_sciencegeneral)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n)),
               inherit.aes = FALSE,
               step.increase = 0)
```

```{r}
# Define breaks for each integer from 1 to 9
breaks <- seq(1, 9, by = 1)

# Define labels for each bin in the desired format
labels <- paste0(1:8, "-", 2:9)

plot_data <- data_long %>% 
  drop_na(BCTI_avg) %>% 
  # # Create the binned version of BCTI_avg
  mutate(BCTI_avg_binned = cut(BCTI_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
  group_by(BCTI_avg_binned, acceptance) %>% 
  summarize(n = n()) %>% 
  group_by(BCTI_avg_binned) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

n_participants <-  data_long %>% 
  drop_na(BCTI_avg) %>% 
  # # Create the binned version of BCTI_avg
  mutate(BCTI_avg_binned = cut(BCTI_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
  group_by(BCTI_avg_binned) %>% 
  summarize(n_participants = n_distinct(id))

plot_data <- left_join(plot_data, n_participants)

absolute_numbers <- plot_data %>% 
  group_by(BCTI_avg_binned) %>% 
  summarise(n_participants = sum(n_participants), 
            n = sum(n))

ggplot(plot_data, aes(x = BCTI_avg_binned, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Average conspiracy thinking (binned)", fill = "Acceptance \nof consensus") +
  theme_minimal() +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(BCTI_avg_binned)) - 0.4, 
                   xmax = as.numeric(factor(BCTI_avg_binned)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n)),
               inherit.aes = FALSE,
               step.increase = 0)
```

```{r}
plot_data
```

### 2024-08-13 exp_4/cleaning.Rmd
pre-registration study 4

--- 2024-05-31 exp_3/cleaning.Rmd
@@
 ---
-title: "Data cleaning Study 3"
+title: "Data cleaning Study 4"
 output: 
@@
 ```{r import, message=FALSE}
-d <- read_csv("data/qualtrics.csv")
+d <- read_csv("data/test.csv")
 names(d)
@@
 ```{r}
+
 d_long <- d_wide %>% 
   # bring to long format
-  pivot_longer(cols = c(ends_with("know"), ends_with("accept"), ends_with("confirm")), 
+  pivot_longer(cols = c(
+    ends_with("know"),
+    ends_with("accept"),
+    ends_with("confirm"),
+    ends_with("link1"),
+    ends_with("link2")
+  ), 
   names_to = "subject_question", values_to = "answer") %>% 
@@
   pivot_wider(names_from = question, values_from = answer) %>% 
+  # make a click count variable of the sum of both link clicks
+  mutate(
+    # make link variables numeric
+    across(contains("link"), ~as.numeric(.x)),
+    # make sum variable
+    link_clicks = link1 + link2) %>% 
 # create better variable names
@@
 
-It's technically not possible to have NA's for trust in science and conspiracy questions, because we used forced responses on them.
-Checking the qualtrics, we know that we accidentally randomized people into two of three outcome measures: trust in science, conspiracy or reasons for accepting consensus.
-This leaves us with reduced sample sizes for all analyses concerning these outcomes (XX for trust in science question, XX for conspiracy measures, XX for accepting consensus).
-
 ```{r}

### 2024-09-02 appendix_exp1.Rmd
draft before edits Hugo

@@
 # Experiment 1 {#exp1}
+
+## Participants
+
+We recruited 200 participants from the US via prolific.
+6 participants failed our attention check, resulting in a final sample of `r exp1_descriptives$n_subj` participants (`r exp1_descriptives$gender$female$n` female, `r exp1_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp1_descriptives$age$mean`, $age_\text{sd}$: `r exp1_descriptives$age$sd`, $age_\text{median}$: `r exp1_descriptives$age$median`).
+Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.
 
@@
 3. How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]
+
+## Results
+
+Regarding RQ1 and RQ2, participants answered on average `r exp1_descriptives$means$knowledge_mean` (sd = `r exp1_descriptives$means$knowledge_sd`) of the questions correctly, and accepted the scientific consensus on average for `r exp1_descriptives$means$acceptance_mean` (sd = `r exp1_descriptives$means$acceptance_sd`) of the questions.
+
+Fig.
+\@ref(fig:exp1-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
+In most cases (`r exp1_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having given the wrong answer to a question.
+In very few cases (`r exp1_descriptives$conditional_acceptance$true$No$share`), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+We believe this might have been due to inattention.
+
+For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus and trust in science (r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`).
+The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\@ref(fig:exp1-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+
+For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`), and conspiracy thinking and acceptance of scientific consensus (r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`).
+
+In Appendix \@ref(exp1), we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \@ref(exp1) also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus?
+In our main analyses, we looked at correlations of acceptance across all observations.
+One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus.
+To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question.
+We then calculated the correlation between trust in science and average acceptance rate by participant.
+We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`), nor with trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`).
+
+## Discussion
+
+These results suggest that most people accept the scientific consensus most of the time.
+Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
+Yet, in `r exp1_descriptives$conditional_acceptance$false$No$share` of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
+In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
 

### 2024-09-02 appendix_exp2.Rmd
draft before edits Hugo

@@
 
+In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
+In most instances they did, but not always.
+In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
+To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
+To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
+We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \@ref(exp1)).
+
+Based on experiment 1, we formulated the following hypotheses:
+
+**H1a: Higher trust in science is associated with more science knowledge?**
+
+**H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?**
+
+**H2a: Higher conspiracy thinking is associated with less science knowledge?**
+
+**H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?**
+
+We had the following research questions:
+
+**RQ1: What is the average science knowledge score?**
+
+**RQ2: When a participant’s answer does not match the consensus, how often do they change their mind and accept the consensus?**
+
+**RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
+
+## Participants
+
+We recruited 201 participants from the US via prolific.
+11 participants failed our attention check, resulting in a final sample of `r exp2_descriptives$n_subj` participants (`r exp2_descriptives$gender$female$n` female, `r exp2_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp2_descriptives$age$mean`, $age_\text{sd}$: `r exp2_descriptives$age$sd`, $age_\text{median}$: `r exp2_descriptives$age$median`).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+## Procedure
+
+The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
+\@ref(fig:exp2-stimulus-example).
+
 ## Materials
+
+We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
 
@@
 ```
+
+## Results
+
+As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`).
+By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
+For trust in science, we find no statistically significant correlation (r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`).
+For conspiracy thinking we find a small negative one (r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
+
+Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\@ref(fig:exp2-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+In Appendix \@ref(exp2) we show that these results hold for our alternative measures of trust and conspiracy thinking.
+We also include more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Regarding RQ1, participants answered on average `r exp2_descriptives$means$knowledge_mean` (sd = `r exp2_descriptives$means$knowledge_sd`) of the questions correctly, and accepted the scientific consensus on average for `r exp2_descriptives$means$acceptance_mean` (sd = `r exp2_descriptives$means$acceptance_sd`) of the questions.
+Fig.
+\@ref(fig:exp2-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
+In response to RQ2, in most cases (`r exp2_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In very few cases (`r exp2_descriptives$conditional_acceptance$true$No$share`), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+
+```{r exp2-justifications}
+exp2_justifications %>%
+  group_by(category)%>%
+  summarize(n = n(), 
+            n_subjects = n_distinct(id)) %>%
+  mutate(Share = n/sum(n)) %>%
+  mutate_if(is.numeric, round, digits=3) %>% 
+  mutate(Share = paste0(Share*100, "%")) %>% 
+  arrange(desc(n)) %>% 
+  select(category, n, Share, n_subjects) %>% 
+  rename(Category = category,
+         `N (instances)` = n, 
+         `Share (instances)` = Share, 
+         `N (unique participants)` = n_subjects
+         ) %>% 
+  apa_table(caption = "Justifications by category")
+```
+
+For RQ3, we got `r exp2_descriptives$justifications_n` answers from `r exp2_descriptives$justifications_n_participants` different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \@ref(tab:exp2-justifications) summarizes these answers by five categories.
+All answers can be read in Appendix \@ref(exp2).
+
+## Discussion
+
+Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.
+answered the knowledge question wrongly).
+The share of these latter is considerably larger in experiment 2 (`r exp2_descriptives$conditional_acceptance$false$Yes$share`) than in experiment 1 (`r exp1_descriptives$conditional_acceptance$false$Yes$share`).
+While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
+We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
 

### 2024-09-02 appendix_exp4.Rmd
draft before edits Hugo

# Experiment 4 {#exp4}

Study 4 relies on broadly the same design as study 3, and is again run on a sample of participants with vaccine-skeptic beliefs. The main difference is that study 4 uses a set of science questions which, instead of being on well-established facts, we are about more recent discoveries (Table \@ref(tab:exp4-knowledge)). Our main question is whether participants will nevertheless readily accept the scientific consensus, as they did for the basic facts in previous studies.

As for the previous studies, we had the following hypotheses:

**H1a: Higher trust in science is associated with more science knowledge?**

**H1b: Higher trust in science is associated with more acceptance of the scientific consensus.**

**H2a: Higher conspiracy thinking is associated with less science knowledge?**

**H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.**

We had the following research questions:

**RQ1: What is the average science knowledge score?**

**RQ2: What is the average acceptance of the scientific consensus**

**RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**

**RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?**

## Participants

We recruited 200 participants from the US via prolific, of which two failed our attention check, resulting in a final sample of `r exp4_descriptives$n_subj` participants (`r exp4_descriptives$gender$female$n` female, `r exp4_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp4_descriptives$age$mean`, $age_\text{sd}$: `r exp4_descriptives$age$sd`, $age_\text{median}$: `r exp4_descriptives$age$median`). Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.

## Procedure

The procedure was identical to the one in experiment 3.

## Materials

\FloatBarrier

Table \@ref(tab:exp4-knowledge) shows all questions, their scientifically consensual answer, and their source. All but two questions were selected from existing science knowledge questionnaires. We tried to select non-political questions.

```{r exp4-knowledge}
# Function to replace special characters in all cells of a data frame
replace_special_characters <- function(df) {
  df %>%
    mutate(across(everything(), ~ gsub("\u2013|\u2014|\u2212", "-", .))) %>% # Replace en-dash, em-dash, and Unicode minus with ASCII hyphen-minus
    mutate(across(everything(), ~ iconv(., from = "UTF-8", to = "ASCII//TRANSLIT"))) # Ensure all text is ASCII
}

items <- read_csv("exp_4/materials/questions_study4.csv")

# Replace special characters
items_cleaned <- replace_special_characters(items) %>% 
  mutate(id = 1:nrow(.)) %>%
  select(id, everything())  # Ensure 'id' is the first column

# Output the table
kbl(items_cleaned, booktabs = T, longtable = TRUE,
    caption = "Science knowledge items", full_width = T) %>%
  kable_styling(font_size = 8) %>%
  column_spec(1, width = "1em") %>%
  column_spec(2, width = "12em") %>%
  column_spec(3, width = "10em") %>%
  column_spec(4, width = "12em")
```

## Results

By contrast to previous studies on basic science knowledge, participants on average answered only  `r exp4_descriptives$means$knowledge_mean` (sd = `r exp4_descriptives$means$knowledge_sd`) of the questions correctly (RQ1). Similar with previous studies, however, they initially accepted the scientific consensus on average for `r exp4_descriptives$means$acceptance_initial_mean` (sd = `r exp4_descriptives$means$acceptance_initial_sd`) of the questions (RQ2). The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (`r exp4_descriptives$means$acceptance_mean`, sd = `r exp4_descriptives$means$acceptance_sd`).

Fig. \@ref(fig:exp4-conditional-acceptance) illustrates the relationship between knowledge and acceptance. In most cases (`r exp4_descriptives$conditional_acceptance$false$acceptance_initial$Yes$share`), participants readily accepted the scientific consensus right after having given the wrong answer to a question. After providing the chance to revise the initial consensus rejection, this share is even larger (`r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`). In few (but still perhaps surprisingly many) cases (`r exp4_descriptives$conditional_acceptance$true$acceptance_initial$No$share`), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response. This share drops slightly after providing the chance to revise the initial consensus rejection (`r exp4_descriptives$conditional_acceptance$true$acceptance$No$share`).

For all correlations, we include opinion revisions for measuring consensus acceptance. We find no statistically significant correlation between science knowledge and trust in science (r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`), but a small positive correlation between acceptance of scientific consensus and trust in science (r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`). Again, these findings might be partly due to ceiling effects: As illustrated in Fig. \@ref(fig:exp4-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates. We do not find a statistically significant correlation between neither conspiracy thinking and science knowledge (r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`), nor between conspiracy thinking and acceptance of scientific consensus (r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`).

For RQ3, we got `r exp4_descriptives$justifications_n` answers from `r exp4_descriptives$justifications_n_participants` different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question. Table \@ref(tab:exp4-justifications) summarizes these answers by five categories.

All answers can be read in the appendix.

```{r exp4-justifications}
exp4_justifications %>%
  group_by(category)%>%
  summarize(n = n(), 
            n_subjects = n_distinct(id)) %>%
  mutate(Share = n/sum(n)) %>%
  mutate_if(is.numeric, round, digits=3) %>% 
  mutate(Share = paste0(Share*100, "%")) %>% 
  arrange(desc(n)) %>% 
  select(category, n, Share, n_subjects) %>% 
  rename(Category = category,
         `N (instances)` = n, 
         `Share (instances)` = Share, 
         `N (unique participants)` = n_subjects
         ) %>% 
  apa_table(caption = "Justifications by category")
```

For RQ4, only `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share` of participants said they accepted the scientific consensus because they trust scientists on this question, while `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` said they independently verified the fact. `r exp4_descriptives$acceptance_by_reason[["other"]]$share` answered with other "other" and gave an open-ended explanation (see below for all open-ended answers).

We also asked all `r exp4_descriptives$reason_followup_n` participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed below.

## Comparing items

### Conspiracy theories

Table \@ref(tab:exp4-correlation-conspiracy) shows the correlations of the three different scales assessing conspiracy thinking.

```{r exp4-correlation-conspiracy}
correlation_matrix <- exp4_wide %>%
  select(BCTI_avg, CMQ_avg, SICBS) %>% 
  rename(BCTI = BCTI_avg, 
         CMQ = CMQ_avg) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different scales assessing conspiracy thinking",
          placement = "h")
```

### Trust in science

Table \@ref(tab:exp4-correlation-trust) shows the correlations of the three different items measuring trust in science.

```{r exp4-correlation-trust}
correlation_matrix <- exp4_wide %>%
  select(wgm_sciencegeneral, wgm_scientists, pew) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different items measuring trust in science",
          placement = "h")
```

## Correlations with alternative measures

Table \@ref(tab:exp4-correlations-outcomes) shows the correlations between knowledge and acceptance, respectively, and outcome variables.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp4_wide %>% 
  select(avg_acceptance, avg_knowledge, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list_knowledge <- map(variables, ~ run_regression(standardized_variables, .x, "avg_knowledge"))
results_list_acceptance <- map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance"))
```

```{r exp4-correlations-outcomes}
# as data frame/output table
results <- bind_rows(list(results_list_knowledge, results_list_acceptance))
results %>% 
  filter(term != "intercept") %>%
  pivot_wider(names_from = term, 
              values_from = estimate) %>% 
  rename(`Correlation with knowledge` = avg_knowledge, 
         `Correlation with acceptance` = avg_acceptance, 
  ) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 
                             outcome == "CMQ_avg" ~ "CMQ", 
                             outcome == "SICBS" ~ "SICBS", 
                             outcome == "wgm_scientists" ~ "WGM trust scientists",
                             outcome == "wgm_sciencegeneral" ~ "WGM trust general \n(main trust measure)",
                             outcome == "pew" ~ "PEW trust scientists",
                             )) %>% 
  apa_table(caption = "Correlations between knowledge and acceptance, respectively, and outcome variables")
```

## Results conditional on false responses

Table \@ref(tab:exp4-false-response-regression) shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp4_false_knowledge %>% 
  select(avg_acceptance, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# Run regression for each variable and store results in a list
results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance", return = "model")), variables)
```

```{r exp4-false-response-regression}
modelsummary(results_list, 
             stars = TRUE,
             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values") 
```

## By-question variation

```{r}
# calculate share
plot_data <- exp4_long %>% 
  group_by(subject) %>% 
  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
```

### Knowledge

(ref:exp4-questions-knowledge) Distribution of correct answers by question.

```{r exp4-questions-knowledge, fig.cap="(ref:exp4-questions-knowledge)"}
# plot knowledge
ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of correct responses") +
  coord_flip()
```

(ref:exp4-questions-acceptance) Distribution of consensus acceptance by question.

```{r exp4-questions-acceptance, fig.cap="(ref:exp4-questions-acceptance)"}
# plot acceptance
ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of acceptance") +
  coord_flip()
```

## Trust in science

(ref:exp4-trust-scientists) Distribution of trust in scientists.

```{r exp4-trust-scientists, fig.cap="(ref:exp4-trust-scientists)"}
# calculate share
plot_data <- exp4_wide %>% 
  group_by(wgm_scientists) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(share = n/sum(n)) %>% 
  rounded_numbers()

# plot knowledge
ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Trust in science", y = "Share of participants") +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) 
```

## Conspiracy thinking

(ref:exp4-conspiracy-distribution) Distribution of conspiracy thinking.

```{r exp4-conspiracy-distribution, fig.cap="(ref:exp4-conspiracy-distribution)"}
# plot knowledge
ggplot(exp4_wide, aes(x = BCTI_avg)) +
  geom_histogram() +
  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
```

### By-item variation

(ref:exp4-conspiracy-items) Distribution of conspiracy thinking by item.

```{r exp4-conspiracy-items, fig.cap="(ref:exp4-conspiracy-items)"}
BCTI_data <- exp4_long %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  select(id, c(starts_with("BCTI"), score)) %>% 
  group_by(BCTI_item) %>% 
  summarize(avg_score = mean(score, na.rm=TRUE)) %>% 
  ungroup()

ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
    geom_pointrange(aes(ymin = 1, ymax = avg_score),
                  # Make the lines a little thicker and the dots a little bigger
                  fatten = 2, size = 1) +
  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
  labs(x = "Conspiracy item", y = "Average score") +
  coord_cartesian(ylim = c(1, 9)) +
  coord_flip()
```

## Reasons for consensus rejection

### All raw answers

```{r exp4-reasons-rejection}
# store all variables
reasons_rejection <- unique(exp4_long$subject)

table <- exp4_wide %>% 
  select(id, all_of(reasons_rejection)) %>% 
  pivot_longer(all_of(reasons_rejection), 
               names_to = "question",
               values_to = "answer") %>% 
  drop_na(answer) %>% 
  arrange(id)

# apa_table(table,
#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
#           longtable = TRUE
#           )
kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2) %>%
  column_spec(3, width = "30em")

```

## Why do people accept the scientific consensus:

### Participants who answered "other"

```{r exp4-other-reasons-acceptance}
# store all variables
table <- exp4_wide %>% 
  select(id, other_reason_agreement) %>% 
  drop_na(other_reason_agreement) %>% 
  arrange(id)

kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2, width = "30em")
```

## How did participants independently verifiy?

### Participants who answered "other"

```{r exp4-independent-verification}
# store all variables
table <- exp4_wide %>% 
  select(id, reason_followup) %>% 
  drop_na(reason_followup) %>% 
  arrange(id)

kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2, width = "30em")
```

## Additional plots

(ref:exp4-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.

```{r exp4-conditional-acceptance, fig.cap="(ref:exp4-conditional-acceptance)"}
plot_data <- exp4_long %>% 
  pivot_longer(c(acceptance, acceptance_initial), 
               names_to = "measure", 
               values_to = "acceptance") %>% 
  # give nicer names
  mutate(measure = case_when(measure == "acceptance" ~ "Acceptance including revisions", 
                             measure == "acceptance_initial" ~ "Initial acceptance"), 
         # make a factor
         measure = factor(measure, levels = unique(measure)[c(2, 1)])
         ) %>% 
  group_by(measure, knowledge, acceptance) %>% 
  count() %>% 
  group_by(measure, knowledge) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

absolute_numbers <- plot_data %>% 
  group_by(measure, knowledge) %>% 
  summarise(n = sum(n))

ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
  plot_theme +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
                   xmax = as.numeric(factor(knowledge)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n)),
               inherit.aes = FALSE,
               step.increase = 0) +
  facet_wrap(~measure)
```

(ref:exp4-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus

```{r exp4-plot-overview, fig.cap="(ref:exp4-plot-overview)", fig.height= 12, fig.width=12}
# Plot distributions of knowledge and acceptance and their mean

mean_value_knowledge <- mean(exp4_wide$avg_knowledge)
max_n <- max(table(exp4_wide$avg_acceptance))

knowledge_mean <- ggplot(exp4_wide, aes(x = avg_knowledge)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, max_n, 10)) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, max_n)) +
  labs(x = "Average science knowledge", 
       y = "N")

mean_value <- mean(exp4_wide$avg_acceptance)

acceptance_mean <- ggplot(exp4_wide, aes(x = avg_acceptance)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, max_n, 10)) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, max_n)) +
  labs(x = "Average consensus acceptance", 
       y = "N")

# Trust in science plots
plot_data <- exp4_wide %>% 
  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
  summarise(n = n())

trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = exp4_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(xlim = c(0, 1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average science knowledge", 
       y = "Trust in science", 
       fill = "N")

plot_data <- exp4_wide %>% 
  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
  summarise(n = n())

trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  
  geom_jitter(inherit.aes = FALSE, data = exp4_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) + 
  scale_fill_viridis_c(option = "plasma") + 
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(xlim = c(0, 1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average consensus Acceptance", 
       y = "Trust in science", 
       fill = "N") +
  theme(legend.position = "top")

# Conspiracy Belief plot

conspiracy_knowledge <- ggplot(exp4_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:9) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, 9)) +
  labs(x = "Average science knowledge", 
       y = "Average conspiracy thinking")

conspiracy_acceptance <- ggplot(exp4_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:9) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, 9)) +
  labs(x = "Average consensus acceptance", 
       y = "Average conspiracy thinking") +
  theme(legend.position = "top")

# Combine plots

# distributions
distributions <- ggarrange(knowledge_mean + 
                             rremove("xlab") , 
                           acceptance_mean + 
                             rremove("xlab") +
                             rremove("ylab"))

# trust
trust <- ggarrange(trust_knowledge + 
                             rremove("xlab") +
                             guides(fill = "none") +
                             labs(caption = NULL) , 
                           trust_acceptance + 
                             rremove("xlab") +
                             rremove("ylab"), 
                   common.legend = TRUE)

# conspiracy
conspiracy <- ggarrange(conspiracy_knowledge, 
                           conspiracy_acceptance + 
                             rremove("ylab"), 
                   common.legend = TRUE)

# common plot
patchwork <- (distributions / trust / conspiracy) + 
  plot_annotation(tag_levels = 'A')

patchwork
```

### 2024-09-02 exp_4/cleaning.Rmd
draft before edits Hugo

@@
 ```{r import, message=FALSE}
-d <- read_csv("data/test.csv")
+d <- read_csv("data/qualtrics.csv")
 names(d)
@@
 
-There are % failed attention checks.
+There are 2 failed attention checks.
 
@@
 ```{r}
-
 d_long <- d_wide %>% 
@@
 ```{r}
+# Extract justifications
+# questions <- unique(data_long$subject)
+# 
+# justifications <- d_wide %>%
+#   select(id, all_of(questions)) %>%
+#   pivot_longer(all_of(questions),
+#                names_to = "question",
+#                values_to = "answer") %>%
+#   drop_na(answer) %>%
+#   arrange(id)
+# 
+# write_csv(justifications, "data/justifications.csv")
+```
+
+```{r}
 justifications <- read_csv("data/justifications_coded.csv")

### 2024-09-02 paper.Rmd
draft before edits Hugo

@@
   - "appendix_exp3.Rmd"
+  - "appendix_exp4.Rmd"
 
@@
 
-Unfortunately, trust in science is far from being at ceiling [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020], and there are suggestions that it has recently been dropping in some countries [@alganTrustScientistsTimes2021] and that it is increasingly politicized in the US [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
+Unfortunately, trust in science is far from being at ceiling [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020], and there are suggestions that it has recently been dropping in some countries [@alganTrustScientistsTimes2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly politicized [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
 
@@
 It could mean that some people reject science wholesale, or at least to a significant extent.
-Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (creationism, etc.), or beliefs that are associated with disliked behaviors (vaccination, acting for climate change).
+Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (e.g.
+creationism), or beliefs that are associated with disliked behaviors (e.g.
+vaccination, acting for climate change).
 
 We answer this question by looking at how much Americans accept basic scientific knowledge.
-We’ll pay special attention to two groups of participants: people who say they don’t trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
-
-We presently briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
+We pay special attention to two groups of participants: people who say they don’t trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
+
+We briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
 
@@
 Across the globe, most people say they trust science at least to some extent.
-In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018] .
+In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018].
 In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants [@wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-Across all countries, in 2018, 32% of participants said they trust science "a lot" (41% in 2020), 45% trust "some" science (39% in 2020), and only 13% percent trust science "not much/ not at all" (13% in 2020), while 10% indicated "don't know" (7% in 2020) (<https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020#gid=6acd&pid=0>).
+Across all countries, in 2018, 32% of participants said they trust science "a lot" (41% in 2020), 45% trust "some" science (39% in 2020), and only 13% percent trust science "not much/ not at all" (13% in 2020), while 10% indicated "don't know" (7% in 2020)[^1].
+
+[^1]: <https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020#gid=6acd&pid=0>\>
 
 Trust in science also appears to be relatively stable in the US [@funkPublicConfidenceScientists2020].
-However, there are some potentially worrying trends: The US might be increasingly polarized on trust in science [@gauchatPoliticizationSciencePublic2012].
-For example, the partisan divide on the question of whether climate change is a major threat to societey [has been growing in the last decade](https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/) [@kennedyWhatDataSays2023].
-And while by some measures only about half of the US population believes in anthropogenic climate change, people do mostly not seem to change their mind on on the issue [@mottaChangingMindsChanging2021].
-
-Trust in science is typically measured in three ways: either asking explictly about general trust in science or scientists, by asking about science attitudes (e.g.
-XX), or asking questions about specific, typically contentious, science topics (e.g.
+However, there are some potentially worrying trends: Recently, trust in US has dropped [@brianAmericansTrustScientists2023], and some evidence suggests increasing polarization of trust in science along political lines [@gauchatPoliticizationSciencePublic2012].
+For example, the partisan divide on the question of whether climate change is a major threat to society has been growing in the last decade[^2] [@kennedyWhatDataSays2023].
+A recent study suggests that by some measure, only about half of the US population believes in anthropogenic climate change, and that individuals do mostly not seem to change their mind on on the issue [@mottaChangingMindsChanging2021].
+
+[^2]: <https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/>
+
+Trust in science is typically measured in three ways: either asking explicitly about general trust in science or scientists, by asking about science attitudes (e.g.
+support for public funding of science), or asking questions about specific, typically contentious, science topics (e.g.
 vaccines, climate change, GMOs).
@@
 
-A lot of research has been conducted on the public understanding of science.
-Part of this research are surveys measuring to which extent people have basic science knowledge (e.g.
+Much of the research on public understanding of science measures to which extent people have basic science knowledge (e.g.
 "Does the Earth go around the Sun, or does the Sun go around the Earth?").
@@
 Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge [@millerPublicUnderstandingAttitudes2004; @millerMeasurementCivicScientific1998a].
-For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria but not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016].
+For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016].
 A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.
 
-[To do: Make a descriptive overview of science knowledge across time/countries.
-This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven.
-But there are enough questions where people's knowledge appears quite poor anyways]
+[To do: Make a descriptive overview of science knowledge across time/countries.]
 
@@
 
-In a series of studies, we ask basic science knowledge questions, inform people of the scientific consensus, and see if they accept it.
-Of particular interest are people who are low in TiS, and CTists.
+In a series of four pre-registered studies (n = 782), we asked participants about basic (studies 1, 2 and 3) and recent/advanced (study 4) science knowledge questions, informed them of the scientific consensus, and aked if they accept it.
+Of particular interest were people who are low in TiS, and CTists.
+Our main hypotheses were:
+
+**H1: Higher trust in science is associated with more science knowledge and more acceptance of the scientific consensus**
+
+**H2: Higher conspiracy thinking/belief is associated with less science knowledge and less acceptance of the scientific consensus**
 
@@
 We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently.
-Study 4 uses the same design as study 3, but asking science questions that people are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves [on an anti-vaxx sample again?].
-
-Then: replications in other countries?
-
-# Overview of experiments
-
-The goal is to see to which extent people do not accept basic science.
-
-# Experiment 1
+Study 4 uses the same design as study 3, but using a set of more recent, advance science questions that participants are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves.
+Just as study 3, study 4 is run on a sample of vaccine-skeptic participants.
+
+# Methods
 
@@
 
-The main goal of experiment one was to test whether people would accept the scientific consensus on basic knowledge questions.
-Additionally, we wanted to know if both science knowledge and acceptance of the scientific consensus are associated with trust in science and conspiracy thinking.
-We had the following research questions:
-
-**RQ1: What is the average science knowledge score (1)?**
-
-**RQ2: What is the average acceptance of the scientific consensus (2)?**
-
-**RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?**
-
-**RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?**
-
-## Methods
-
-### Participants
-
-We recruited 200 participants from the US via prolific.
-6 participants failed our attention check, resulting in a final sample of `r exp1_descriptives$n_subj` participants (`r exp1_descriptives$gender$female$n` female, `r exp1_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp1_descriptives$age$mean`, $age_\text{sd}$: `r exp1_descriptives$age$sd`, $age_\text{median}$: `r exp1_descriptives$age$median`).
-Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.
-
-### Procedure
-
-After providing their consent to participate in the study, participants were given an attention check "While watching the television, have you ever had a fatal heart attack?" [1-6; 1 = Never, 6 = Often].
-All participants who did not answer "1 = Never" were excluded.
-Participants then read the following instructions:"We will ask you 11 questions about science.
-After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions, which were randomly selected from a pool of 11 questions, in random order.
-After each question, participants were presented with an answer reflecting the scientific consensus.
-Participants were asked to choose whether they accept the answer or not, before proceeding to the next question.
-Figure \@ref(fig:stimulus-example) displays the survey for an example science question.
-Finally, participants answered questions on conspiracy thinking and trust in science.
-
-### Materials
-
-#### Science knowledge and acceptance
-
-Table \@ref(tab:knowledge) shows all questions, their scientifically consensual answer, and their source.
-All but two questions were selected from existing science knowledge questionnaires.
-We tried to select non-political questions.
-
-```{r knowledge}
-# Function to replace special characters in all cells of a data frame
-replace_special_characters <- function(df) {
-  df %>%
-    mutate(across(everything(), ~ gsub("\u2013|\u2014|\u2212", "-", .))) %>% # Replace en-dash, em-dash, and Unicode minus with ASCII hyphen-minus
-    mutate(across(everything(), ~ iconv(., from = "UTF-8", to = "ASCII//TRANSLIT"))) # Ensure all text is ASCII
-}
-
-# Read the CSV file
-items <- read_csv("exp_4/materials/overview_questions.csv")
-
-# Replace special characters
-items_cleaned <- replace_special_characters(items) %>% 
-  mutate(id = 1:nrow(.)) %>%
-  select(id, everything())  # Ensure 'id' is the first column
-
-# add footnote for the tree items
-items_cleaned[11, 2] <- paste0("*", 
-                               items_cleaned[11, 2])
-
-# Output the table
-knitr::kable(items_cleaned, booktabs = T, longtable = TRUE,
-    caption = "Science knowledge items", 
-    full_width = T, 
-    col.names = c("", "Study 1-3", "Study 4")) %>%
-  kable_styling(font_size = 8) %>%
-  column_spec(1, width = "2em") %>%
-  column_spec(2, width = "22em")%>%
-  column_spec(3, width = "22em") %>% 
-  footnote(symbol = "Only used in Study 1")
-```
-
-#### Conspiracy scales
-
-To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 (Table \@ref(tab:conspiracy)).
-Participants were asked: "Below is a list of events for which the official version has been disputed.
-For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
-[1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true]".
-
-```{r conspiracy, echo=FALSE}
-# Create the data frame
-items <- c(
-  "The Apollo moon landings never happened and were staged in a Hollywood film studio.",
-  "A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).",
-  "The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.",
-  "The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.",
-  "The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.",
-  "There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.",
-  "In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.",
-  "Proof of alien contact is being concealed from the public.",
-  "Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.",
-  "Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.")
-
-conspiracy_items <- data.frame(id = 1:length(items), items = items)
-
-# Output the table
-kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL, 
-    caption = "Conspiracy items", 
-    full_width = F) %>%
-  kable_styling(font_size = 8) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "40em")
-```
-
-To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 and the Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 (see Appendix \@ref(exp1)).
-
-#### Trust in science
-
-Our main item for measuring trust in science is selected from the Wellcome Global Monitor survey: "In general, would you say that you trust science a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"
-
-We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey ("How much do you trust scientists in this country?
-Do you trust them a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), the other from the Pew research center ("How much confidence do you have in scientists to act in the best interests of the public?
-[1-5; 1 = No confidence at all, 5 = A great deal of confidence]").
-We selected these items so that we could compare the ratings in our sample to global survey results.
-The WGM survey has been administered in over 140 countries and included over 140000 respondents.
-The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents [@colognaTrustScientistsTheir2024].
-
-## Results
-
-Regarding RQ1 and RQ2, participants answered on average `r exp1_descriptives$means$knowledge_mean` (sd = `r exp1_descriptives$means$knowledge_sd`) of the questions correctly, and accepted the scientific consensus on average for `r exp1_descriptives$means$acceptance_mean` (sd = `r exp1_descriptives$means$acceptance_sd`) of the questions.
-
-Fig.
-\@ref(fig:exp1-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
-In most cases (`r exp1_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having given the wrong answer to a question.
-In very few cases (`r exp1_descriptives$conditional_acceptance$true$No$share`), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
-We believe this might have been due to inattention.
-
-For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus and trust in science (r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`).
-The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\@ref(fig:exp1-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-
-For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`), and conspiracy thinking and acceptance of scientific consensus (r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`).
-
-In Appendix \@ref(exp1), we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \@ref(exp1) also includes more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus?
-In our main analyses, we looked at correlations of acceptance across all observations.
-One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus.
-To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question.
-We than calculated the correlation between trust in science and average acceptance rate by participant.
-We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`), nor with trust in science (r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`).
-
-## Discussion
-
-These results suggest that most people accept the scientific consensus most of the time.
-Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
-Yet, in `r exp1_descriptives$conditional_acceptance$false$No$share` of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
-In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
-# Experiment 2
-
 ```{r exp2}
@@
 false_answers_cor_trust_acceptance <- cor.test(exp2_false_knowledge$wgm_sciencegeneral, exp2_false_knowledge$avg_acceptance) %>% text_ready()
-
-# pooled across all cases
+# pooled across all cases (in line with all other studies' analysis)
 cor_trust_acceptance <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_acceptance) %>% text_ready()
@@
 false_answers_cor_conspiracy_acceptance <- cor.test(exp2_false_knowledge$BCTI_avg, exp2_false_knowledge$avg_acceptance) %>% text_ready()
-
-# pooled across all cases
+# pooled across all cases (in line with all other studies' analysis)
 cor_conspiracy_acceptance <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_acceptance) %>% text_ready()
@@
 ```
-
-In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
-In most instances they did, but not always.
-In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
-To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
-To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
-We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \@ref(exp1)).
-
-Based on experiment 1, we formulated the following hypotheses:
-
-**H1a: Higher trust in science is associated with more science knowledge?**
-
-**H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?**
-
-**H2a: Higher conspiracy thinking is associated with less science knowledge?**
-
-**H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?**
-
-We had the following research questions:
-
-**RQ1: What is the average science knowledge score?**
-
-**RQ2: When a participant’s answer does not match the consensus, how often do they change their mind and accept the consensus?**
-
-**RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
-
-## Methods
-
-### Participants
-
-We recruited 201 participants from the US via prolific.
-11 participants failed our attention check, resulting in a final sample of `r exp2_descriptives$n_subj` participants (`r exp2_descriptives$gender$female$n` female, `r exp2_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp2_descriptives$age$mean`, $age_\text{sd}$: `r exp2_descriptives$age$sd`, $age_\text{median}$: `r exp2_descriptives$age$median`).
-Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
-
-### Procedure
-
-The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
-\@ref(fig:exp2-stimulus-example).
-
-### Materials
-
-We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
-
-## Results
-
-As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`).
-By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
-For trust in science, we find no statistically significant correlation (r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`).
-For conspiracy thinking we find a small negative one (r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
-
-Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\@ref(fig:exp2-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-In Appendix \@ref(exp2) we show that these results hold for our alternative measures of trust and conspiracy thinking.
-We also include more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Regarding RQ1, participants answered on average `r exp2_descriptives$means$knowledge_mean` (sd = `r exp2_descriptives$means$knowledge_sd`) of the questions correctly, and accepted the scientific consensus on average for `r exp2_descriptives$means$acceptance_mean` (sd = `r exp2_descriptives$means$acceptance_sd`) of the questions.
-Fig.
-\@ref(fig:exp2-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
-In response to RQ2, in most cases (`r exp2_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-In very few cases (`r exp2_descriptives$conditional_acceptance$true$No$share`), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
-
-For RQ3, we got `r exp2_descriptives$justifications_n` answers from `r exp2_descriptives$justifications_n_participants` different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Table \@ref(tab:exp2-justifications) summarizes these answers by five categories.
-All answers can be read in Appendix \@ref(exp2).
-
-```{r exp2-justifications}
-exp2_justifications %>%
-  group_by(category)%>%
-  summarize(n = n(), 
-            n_subjects = n_distinct(id)) %>%
-  mutate(Share = n/sum(n)) %>%
-  mutate_if(is.numeric, round, digits=3) %>% 
-  mutate(Share = paste0(Share*100, "%")) %>% 
-  arrange(desc(n)) %>% 
-  select(category, n, Share, n_subjects) %>% 
-  rename(Category = category,
-         `N (instances)` = n, 
-         `Share (instances)` = Share, 
-         `N (unique participants)` = n_subjects
-         ) %>% 
-  apa_table(caption = "Justifications by category")
-```
-
-## Discussion
-
-Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.
-answered the knowledge question wrongly).
-The share of these latter is considerably larger in experiment 2 (`r exp2_descriptives$conditional_acceptance$false$Yes$share`) than in experiment 1 (`r exp1_descriptives$conditional_acceptance$false$Yes$share`).
-While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
-We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
-# Experiment 3
 
@@
 
+```{r exp4}
+# Analyze data of experiments and store results
+
+# Experiment 4
+
+# read the cleaned long version data set
+exp4_long <- read_csv("exp_4/data/cleaned_long.csv")
+
+# read wide version data set
+exp4_wide <- read_csv("exp_4/data/cleaned_wide.csv")
+
+# read coded justifications
+exp4_justifications <- read_csv("exp_4/data/justifications_clean.csv")
+
+# Research questions
+
+# H1
+cor_trust_knowledge <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_knowledge,
+  use = "complete.obs") %>% text_ready()
+cor_trust_acceptance <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_acceptance,
+  use = "complete.obs") %>% text_ready()
+
+# H2
+cor_conspiracy_knowledge <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_knowledge,
+  use = "complete.obs") %>% text_ready()
+cor_conspiracy_acceptance <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_acceptance,
+  use = "complete.obs") %>% text_ready()
+
+# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
+exp4_false_knowledge <- exp4_long %>% 
+  # make numeric versions
+  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
+  ) %>% 
+  group_by(id, knowledge) %>% 
+  # calculate by-participant averages
+  summarize(
+    n = n(),
+    n_accepted = sum(acceptance_num),
+    avg_acceptance  = sum(acceptance_num)/n(), 
+    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
+    BCTI_avg = mean(BCTI_avg), 
+    # add additonal measures for robustness checks
+    CMQ_avg = mean(CMQ_avg), 
+    SICBS = mean(SICBS),
+    wgm_scientists = mean(wgm_scientists), 
+    pew = mean(pew)
+  ) %>% 
+  # filter to only false responses
+  filter(knowledge == FALSE) %>% 
+  ungroup()
+
+# H1b conditional on false answers
+false_answers_cor_trust_acceptance <- cor.test(exp4_false_knowledge$wgm_sciencegeneral, exp4_false_knowledge$avg_acceptance,
+  use = "complete.obs") %>% 
+  text_ready()
+
+# H2b conditional on false answers
+false_answers_cor_conspiracy_acceptance <- cor.test(exp4_false_knowledge$BCTI_avg, exp4_false_knowledge$avg_acceptance,
+  use = "complete.obs") %>% 
+  text_ready()
+
+# extract descriptives for inline reporting
+exp4_descriptives <- list(
+  # Demographics
+  n_subj = n_distinct(exp4_wide$id),
+  n_subject_by_outcome = exp4_wide %>% 
+    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
+                     ~sum(!is.na(.x)
+                     )
+    )
+    ),
+  gender = exp4_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
+  age = exp4_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
+                                       .names = "{.fn}")) %>% rounded_numbers(),
+  # RQ1 & RQ2
+  means = exp4_wide %>% 
+    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()) %>% 
+    mutate_if(is.numeric, round, digits = 2) %>% 
+  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
+  # H1
+  # a
+  cor_trust_knowledge = cor_trust_knowledge,
+  # b (conditional on false responses)
+  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
+  # b (all cases)
+  cor_trust_acceptance = cor_trust_acceptance,
+  # H2
+  # a
+  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
+  # b (conditional on false responses)
+  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
+  # b(all cases)
+  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
+  # Exploratory
+  # acceptance and knowledge
+  conditional_acceptance = exp4_long %>% 
+    pivot_longer(c(acceptance, acceptance_initial), 
+                 names_to = "measure", 
+                 values_to = "acceptance") %>% 
+    group_by(knowledge, measure, acceptance) %>% 
+    count() %>% 
+    group_by(knowledge, measure) %>% 
+    mutate (share = n/sum(n), 
+            # rename knowledge values
+            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate (share = paste0(share*100, " %")) %>% 
+    super_split(knowledge,  measure, acceptance),
+  # Justifications
+  justifications_n = nrow(exp4_justifications), 
+  justifications_n_participants = n_distinct(exp4_justifications$id),
+  justification_by_category = exp4_justifications %>%
+    group_by(category)%>%
+    summarize(n = n(), 
+              n_subjects = n_distinct(id)) %>%
+    mutate(share = n/sum(n)) %>%
+    mutate_if(is.numeric, round, digits=3) %>% 
+    mutate(share = paste0(share*100, "%")) %>% 
+    split(.$category),
+  # Consensus acceptance
+  acceptance_n = exp4_wide %>% 
+    filter(!is.na(reason_agreement)) %>% 
+    nrow(.),
+  acceptance_by_reason = exp4_wide %>% 
+    group_by(reason_agreement) %>% 
+    summarize(n = n()) %>%
+    drop_na(reason_agreement) %>% 
+    mutate(share = n/sum(n)) %>%
+    mutate_if(is.numeric, round, digits=3) %>% 
+    mutate(share = paste0(share*100, "%")) %>% 
+    split(.$reason_agreement),
+  reason_followup_n = exp4_wide %>% 
+    filter(!is.na(reason_followup)) %>% 
+    nrow(.)
+)
+
+```
+
+## Procedure
+
+After providing their consent to participate in the study, participants were given an attention check "While watching the television, have you ever had a fatal heart attack?" [1-6; 1 = Never, 6 = Often].
+All participants who did not answer "1 = Never" were excluded.
+Participants then read the following instructions:"We will ask you 10 questions about science.
+After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions (in study 1 they were randomly selected from a pool of 11 questions) in random order.
+After each question, participants were presented with an answer reflecting the scientific consensus.
+In study 2 and 3, participants additionally saw a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
+In study 4, we provided only two links and no explanation.
+Participants were asked to choose whether they accept the answer or not, before proceeding to the next question.
+Participants then answered questions on conspiracy thinking, conspiracy beliefs and trust in science.
+
+In study 2, 3 and 4, we presented participants with open-ended questions to explain cases where they rejected the scientific consensus.
+In study 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus.
+Finally, at the end of study 3 and 4, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain.
+Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".
+
+## Participants
+
+We recruited a total of `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in study 1, `r exp2_descriptives$n_subj` in study 2, `r exp3_descriptives$n_subj` in study 3, `r exp4_descriptives$n_subj` in study 4) participants in the US via prolific.
+Details and demographics can be found in the online supplemental material.
+While samples for study 1 and 2 were convenience samples, study 3 and 4 were run on a sample holding vaccine-skeptic beliefs (see OSM for details).
+
+## Materials
+
+Table \@ref(tab:knowledge) shows all questions and their answer options used in the four studies.
+
+```{r knowledge}
+# Function to replace special characters in all cells of a data frame
+replace_special_characters <- function(df) {
+  df %>%
+    mutate(across(everything(), ~ gsub("\u2013|\u2014|\u2212", "-", .))) %>% # Replace en-dash, em-dash, and Unicode minus with ASCII hyphen-minus
+    mutate(across(everything(), ~ iconv(., from = "UTF-8", to = "ASCII//TRANSLIT"))) # Ensure all text is ASCII
+}
+
+# Read the CSV file
+items <- read_csv("exp_4/materials/overview_questions.csv")
+
+# Replace special characters
+items_cleaned <- replace_special_characters(items) %>% 
+  mutate(id = 1:nrow(.)) %>%
+  select(id, everything())  # Ensure 'id' is the first column
+
+# add footnote for the tree items
+items_cleaned[11, 2] <- paste0("*", 
+                               items_cleaned[11, 2])
+
+# Output the table
+knitr::kable(items_cleaned, booktabs = T, longtable = TRUE,
+    caption = "Science knowledge items", 
+    full_width = T, 
+    col.names = c("", "Study 1-3", "Study 4")) %>%
+  kable_styling(font_size = 8) %>%
+  column_spec(1, width = "2em") %>%
+  column_spec(2, width = "22em")%>%
+  column_spec(3, width = "22em") %>% 
+  footnote(symbol = "Only used in Study 1")
+```
+
+Our main measure of conspiracy belief was asking participants about specific conspiracy theories.
+We selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) [@pennycookOverconfidentlyConspiratorialConspiracy2022] (Table \@ref(tab:conspiracy)).
+Participants were asked: "Below is a list of events for which the official version has been disputed.
+For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
+[1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true]".
+
+As a robustness check, we also used established scales on conspiracy thinking, namely the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013] and the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016].
+Details and comparisons between the scales can be found in the OSM.
+
+```{r conspiracy, echo=FALSE}
+# Create the data frame
+items <- c(
+  "The Apollo moon landings never happened and were staged in a Hollywood film studio.",
+  "A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).",
+  "The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.",
+  "The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.",
+  "The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.",
+  "There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.",
+  "In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.",
+  "Proof of alien contact is being concealed from the public.",
+  "Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.",
+  "Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.")
+
+conspiracy_items <- data.frame(id = 1:length(items), items = items)
+
+# Output the table
+kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL, 
+    caption = "Conspiracy items", 
+    full_width = F) %>%
+  kable_styling(font_size = 8) %>% 
+  column_spec(1) %>%
+  column_spec(2, width = "40em")
+```
+
+Our main item for measuring trust in science was selected from the Wellcome Global Monitor survey: "In general, would you say that you trust science a lot, some, not much, or not at all?
+[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]".
+We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey ("How much do you trust scientists in this country?
+Do you trust them a lot, some, not much, or not at all?
+[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), the other from the Pew research center ("How much confidence do you have in scientists to act in the best interests of the public?
+[1-5; 1 = No confidence at all, 5 = A great deal of confidence]").
+We selected these items so that we could compare the ratings in our sample to global survey results.
+The WGM survey has been administered in over 140 countries and included over 140000 respondents.
+The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents [@colognaTrustScientistsTheir2024].
+
+# Results
+
 (ref:summary-plot) Points represent the average share of acceptance as a function of the **A** level of trust in science ("In general, would you say that you trust science a lot, some, not much, or not at all?
@@
                            exp3_long %>% 
-                             mutate(study = "study 3")
+                             mutate(study = "study 3"),
+                           exp4_long %>% 
+                             mutate(study = "study 4")
                            )
@@
         axis.ticks.y = element_blank()) 
-  #guides(color = "none", shape = "none") +
+  #guides(color = "none", shape = "none")
 
@@
 
-Study 3 is essentially a replication-–with some minor modifications-–of study 2, but on a different type of sample.
-Both study 1 and 2 were run on convenience samples.
-For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
-By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
-\@ref(fig:exp3-explanation-example)).
-Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
-Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
-We want to know if participants perceive that this is because of trust, or other factors.
-
-As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
-
-**H1a: Higher trust in science is associated with more science knowledge?**
-
-**H1b: Higher trust in science is associated with more acceptance of the scientific consensus.**
-
-**H2a: Higher conspiracy thinking is associated with less science knowledge?**
-
-**H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.**
-
-We had the following research questions:
-
-**RQ1: What is the average science knowledge score?**
-
-**RQ2: What is the average acceptance of the scientific consensus**
-
-**RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
-
-**RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?**
-
-## Methods
-
-### Participants
-
-We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of `r exp3_descriptives$n_subj` participants (`r exp3_descriptives$gender$female$n` female, `r exp3_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp3_descriptives$age$mean`, $age_\text{sd}$: `r exp3_descriptives$age$sd`, $age_\text{median}$: `r exp3_descriptives$age$median`).
-Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
-
-However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
-This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = `r exp3_descriptives$n_subject_by_outcome$wgm_sciencegeneral` for trust in science, N = `r exp3_descriptives$n_subject_by_outcome$BCTI_avg` for conspiracy measures, N = `r exp3_descriptives$n_subject_by_outcome$reason_agreement` for reason for accepting consensus).
-
-### Procedure
-
-The procedure was mostly the same as in experiment 2.
-In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
-Finally, at the end of the survey, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain.
-Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".
-
-### Materials
-
-We relied on the same items as in experiment 2.
-We used a pre-registered categorization scheme for the open ended answers for why people reject the consensus.
-The categories were XX and XX.
-[Describe coding process, with independent reviewer etc.]
-
-## Results
-
-As in study 1 and 2, we find that participants answered on average `r exp3_descriptives$means$knowledge_mean` (sd = `r exp3_descriptives$means$knowledge_sd`) of the questions correctly, and initially accepted the scientific consensus on average for `r exp3_descriptives$means$acceptance_initial_mean` (sd = `r exp3_descriptives$means$acceptance_initial_sd`) of the questions (RQ2).
-The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (`r exp3_descriptives$means$acceptance_mean`, sd = `r exp3_descriptives$means$acceptance_sd`).
-
-Fig.
-\@ref(fig:exp3-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
-In most cases (`r exp3_descriptives$conditional_acceptance$false$acceptance_initial$Yes$share`), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
-After providing the chance to revise the initial consensus rejection, this share is even larger (`r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`).
-In very few cases (`r exp3_descriptives$conditional_acceptance$true$acceptance_initial$No$share`), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
-This share drops slightly after providing the chance to revise the initial consensus rejection (`r exp3_descriptives$conditional_acceptance$true$acceptance$No$share`).
-
-In the appendix XX we provide an analysis that this drop is statistically significant [TO DO].
-
-For all correlations, we include opinion revisions for measuring consensus acceptance.
-We find no statistically significant correlation between science knowledge and trust in science (r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`).
-Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
-\@ref(fig:exp3-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-We find no statistically significant correlation between conspiracy thinking and science knowledge (r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`), and between conspiracy thinking and acceptance of scientific consensus (r = `r exp3_descriptives$cor_conspiracy_acceptance`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`).
-
-In Appendix \@ref(exp3), we show that these associations are statistically significant (except the last one?).
-We show that they are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \@ref(exp3) also includes more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Regarding RQ1, participants answered on average `r exp3_descriptives$means$knowledge_mean` (sd = `r exp3_descriptives$means$knowledge_sd`) of the questions correctly.
-For RQ3, we got `r exp3_descriptives$justifications_n` answers from `r exp3_descriptives$justifications_n_participants` different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Table \@ref(tab:exp3-justifications) summarizes these answers by five categories.
-
-All answers can be read in the appendix.
-
-```{r exp3-justifications}
+Across all studies we find that almost all participants trust basic and recent/advance science facts almost all the time.
+That is true even for participants who expressed low general trust in science, and those who have a tendency to conspiracy think or hold science-related conspiracy beliefs (Fig.
+\@ref(fig:summary-plot)).
+
+Our results suggest that the more people tend to trust science, the more they tend to know about science and accept the scientific consensus.
+In study 2 and 3, we find a positive but small correlation between trust in science and both science knowledge (study 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`).
+In study 3 and 4--both run on vaccine-skeptic samples--we find no statistically significant correlation between science knowledge and trust in science (study 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`), but a small positive correlation between acceptance of scientific consensus and trust in science (study 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`).
+
+As for conspiracy beliefs, results are mixed.
+In studies 1 and 2, we find a small negative correlation between conspiracy thinking and both science knowledge (study 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`).
+However, in study 3 and 4, we do not find a statistically significant correlation with neither science knowledge (study 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`), and between nor acceptance of scientific consensus (study 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`).
+
+Overall, the correlations we find are relatively weak, which might be partly due to ceiling effects.
+This was the case in particular for acceptance: In all studies, in the vast majority of cases, participants accepted accepted science facts.
+The absence of correlations between conspiracy beliefs and science knowledge or acceptance in studies 3 and 4 might be explained by our sample selection.
+For these studies, we recruited samples scoring considerably higher in conspiracy belief than the convenience samples of study 1 and 2 (see OSM).
+The presence of effects in study 1 and 2, and the absence in study 3 and 4 might be because differences between moderate and high conspiracy thinking might not matter, while, differences between low and moderate/high conspiracy thinking do matter.
+
+Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
+Yes.
+In most cases (study 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+Only in study 4 do we find evidence[^3] that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in study 2 evidence that it is associated with (less) conspiracy thinking (study 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
+In very few cases, participants changed their mind away from the scientific consensus.
+These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: `r exp1_descriptives$conditional_acceptance$true$No$share`; study 2: `r exp2_descriptives$conditional_acceptance$true$No$share`; study 3: `r exp3_descriptives$conditional_acceptance$true$acceptance$No$share`; study 4: `r exp4_descriptives$conditional_acceptance$true$acceptance$No$share`).
+
+[^3]: For this analysis, we restricted our data to those cases in which participants initially gave the wrong answer.
+
+Why did participants reject the scientific consensus?
+We got a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (study 2: `r exp2_descriptives$justifications_n`; study 3: `r exp3_descriptives$justifications_n`; study 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (study 2: `r exp2_descriptives$justifications_n_participants`; study 3: `r exp3_descriptives$justifications_n_participants`; study 4: `r exp4_descriptives$justifications_n_participants`) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Based on the answers, we created five categories.
+Table \@ref(tab:justifications) summarizes these answers by five categories.
+All answers can be read in the OSM.
+
+```{r justifications}
+
+## make a common table of exp 2 and exp 3
+justifications <- bind_rows(exp2_justifications %>% 
+                              mutate(study = "study 2"), 
                             exp3_justifications %>% 
+                              mutate(study = "study 3", 
+                                     final_coding = as.character(final_coding)),
+                            exp4_justifications %>% 
+                              mutate(study = "study 4"))
+
+justifications %>%
   group_by(category)%>%
@@
   select(category, n, Share, n_subjects) %>% 
-  rename(Category = category,
+  ungroup() %>% 
+  rename(
+    Category = category,
     `N (instances)` = n, 
@@
   ) %>% 
-  apa_table(caption = "Justifications by category")
-```
-
-For RQ4, we had `r exp3_descriptives$acceptance_n` participants answering the question.
-Of these `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share` said they accepted the scientific consensus because they trust scientists on this question, while `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share` said they independently verified the fact.
-`r exp3_descriptives$acceptance_by_reason[["other"]]$share` answered with other "other" and gave an open-ended explanation (see Appendix).
-
-We also asked all `r exp3_descriptives$reason_followup_n` participants who answered that they had independently verified the answer to explain how they did so.
-The open-ended answers are listed in Appendix XX.
+  apa_table(caption = "Justifications by category for data from studies 2, 3 and 4 combined.")
+```
+
+Why did participants accept the scientific consensus?
+In studies 3 and 4, we had asked participants about cases where they agreed with the scientific consensus.
+A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (study 3: `r exp3_descriptives$acceptance_n`; study 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
+Of these, `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 4 said they accepted the scientific consensus because they trust scientists, while `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 4 said they independently verified the fact.
+We asked all participants who gave this answer to explain how they did it (answers can be found in the OSM).
+`r exp3_descriptives$acceptance_by_reason[["other"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in study 4 answered with other "other" and gave an open-ended explanation (see OSM).
+
+More detailed results addressing all our pre-registered research questions can be found in the OSM.
 
@@
 
-[We can draw some inspiration from the “Westwood et al 2021 Current research overstates American support for political violence” paper: we want to put in perspective the current rise (?
-discourse around?) in mistrust towards science]
-
-There is increasingly more talk of defiance towards science and towards experts more generally.
-In every population, many people only trust science ‘some of the time’, and a sizable minority doesn’t trust it.
-This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
-
-But how deep is distrust of science? In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
-
-Traditionally, people have interpreted people's lack of basic knowledge in light of the deficit-model. Here, we show that people readily accept the scientifically consensual answer to these questions.
-
-“The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
-Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
-Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
-The results of the U.S.
-portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
-Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators’ questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).”
+Survey data from the US suggests a decline of trust in science in recent years [@brianAmericansTrustScientists2023].
+Across the globe, many people only trust science ‘some of the time’ , and a sizable minority doesn’t trust it [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
+This appears difficult to reconcile with the fact that, in many countries, most people nowadays receive at least a basic education [@ritchieGlobalEducation2023], and that science education has been found to be the most strongly associated variable with trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018].
+
+But how deep is distrust of science?
+In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
+Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
+That is true not only for basic science questions (studies 1-3), but also for more recent scientific findings that most people are unfamiliar with (study 4).
+In study 1 the acceptance rate after having given a wrong answer was considerably lower (`r exp1_descriptives$conditional_acceptance$false$No$share`) than in subsequent studies (study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`).
+While this could be just sampling variation, it might be that adding explanations and/or sources convinced participants more than merely stating the consensus.
+In general, participants with lower trust in science and who believe more in conspiracy theories tended to both know less about science and accept the scientific consensus less.
+
+Scholars have long interpreted the lack of basic science knowledge as the root cause of distrust in and more generally unfavorable attitudes towards science, an explanation known under the term "deficit-model" [@millerMeasurementCivicScientific1998a; @millerPublicUnderstandingAttitudes2004; @durantPublicUnderstandingScience1989].
+In line with proponents of the deficit-model, we find that people indeed err even on very basic science questions.
+However, we also find that almost everyone instantly accepted the scientific consensus once informed about it.
+Moreover, the same was true for non-basic science questions.
+This calls into question the deficit model: If participants trust science on basic (and non-basic) questions—as suggests immediate acceptance of the scientific consensus—then unfavorable attitudes towards science must have other explanations than a mere lack of knowledge.
+This argument is supported by past research.
+In a seminal meta-analysis, \@allumScienceKnowledgeAttitudes2008 found correlations between science knowledge and general attitudes towards science to be at best small and largely context dependent [\@allumScienceKnowledgeAttitudes2008].
+A more recent report by the National Academies of Sciences, Engineering, and Medicine states that science knowledge is "unlikely to substantially affect attitudes on scientific issues" [@nationalacademiesofsciencesCommunicatingScienceEffectively2017].
+
+How to make sense of the detachment of basic science knowledge from attitudes towards science?
+One explanation might be that most scientific findings are very abstract.
+For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second.
+It is therefor easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
+In other words, people might hold on to most scientific findings as reflective beliefs—beliefs that have little effect on our behaviors [@mercierHowGoodAre2021].
+It is probably when it becomes tangible that science shape's people's attitudes.
+This can happen, for example, when science appears to support conclusions in favor of certain policies or moral positions, as for example in the case of anthropogenic climate change or vaccines.
+
+It is tempting, especially for scientists, to explain mistrust in science with ignorance.
+After all, careful proof of objective truth is what makes science trustworthy—that is, for scientists.
+For most people, what shapes trust in science might not be so different from what shapes trust in other institutions, such as the government, banks or church.
 
@@
 ```
+
+\clearpage
+
+```{r child = "appendix_exp4.Rmd"}
+```

### 2024-09-02 paper.tex
draft before edits Hugo

@@
 
-Unfortunately, trust in science is far from being at ceiling (Wellcome Global Monitor, 2018, 2020), and there are suggestions that it has recently been dropping in some countries (Algan et al., 2021) and that it is increasingly politicized in the US (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
+Unfortunately, trust in science is far from being at ceiling (Wellcome Global Monitor, 2018, 2020), and there are suggestions that it has recently been dropping in some countries (Algan et al., 2021), including the US (Brian \& Tyson, 2023), where it is also increasingly politicized (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
 
@@
 It could mean that some people reject science wholesale, or at least to a significant extent.
-Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (creationism, etc.), or beliefs that are associated with disliked behaviors (vaccination, acting for climate change).
+Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (e.g.~creationism), or beliefs that are associated with disliked behaviors (e.g.~vaccination, acting for climate change).
 
 We answer this question by looking at how much Americans accept basic scientific knowledge.
-We'll pay special attention to two groups of participants: people who say they don't trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
-
-We presently briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
+We pay special attention to two groups of participants: people who say they don't trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
+
+We briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
 
@@
 Across the globe, most people say they trust science at least to some extent.
-In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science (Wellcome Global Monitor, 2018) .
+In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science (Wellcome Global Monitor, 2018).
 In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants (Wellcome Global Monitor, 2020).
-Across all countries, in 2018, 32\% of participants said they trust science ``a lot'' (41\% in 2020), 45\% trust ``some'' science (39\% in 2020), and only 13\% percent trust science ``not much/ not at all'' (13\% in 2020), while 10\% indicated ``don't know'' (7\% in 2020) (\url{https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020\#gid=6acd&pid=0}).
+Across all countries, in 2018, 32\% of participants said they trust science ``a lot'' (41\% in 2020), 45\% trust ``some'' science (39\% in 2020), and only 13\% percent trust science ``not much/ not at all'' (13\% in 2020), while 10\% indicated ``don't know'' (7\% in 2020)\footnote{\url{https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020\#gid=6acd&pid=0}\textgreater{}}.
 
 Trust in science also appears to be relatively stable in the US (Funk \& Kennedy, 2020).
-However, there are some potentially worrying trends: The US might be increasingly polarized on trust in science (Gauchat, 2012).
-For example, the partisan divide on the question of whether climate change is a major threat to societey \href{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}{has been growing in the last decade} (Kennedy, 2023).
-And while by some measures only about half of the US population believes in anthropogenic climate change, people do mostly not seem to change their mind on on the issue (Motta, 2021).
-
-Trust in science is typically measured in three ways: either asking explictly about general trust in science or scientists, by asking about science attitudes (e.g.~XX), or asking questions about specific, typically contentious, science topics (e.g.~vaccines, climate change, GMOs).
+However, there are some potentially worrying trends: Recently, trust in US has dropped (Brian \& Tyson, 2023), and some evidence suggests increasing polarization of trust in science along political lines (Gauchat, 2012).
+For example, the partisan divide on the question of whether climate change is a major threat to society has been growing in the last decade\footnote{\url{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}} (Kennedy, 2023).
+A recent study suggests that by some measure, only about half of the US population believes in anthropogenic climate change, and that individuals do mostly not seem to change their mind on on the issue (Motta, 2021).
+
+Trust in science is typically measured in three ways: either asking explicitly about general trust in science or scientists, by asking about science attitudes (e.g.~support for public funding of science), or asking questions about specific, typically contentious, science topics (e.g.~vaccines, climate change, GMOs).
 It is unclear to which extent these measures capture trust in basic science.
@@
 
-A lot of research has been conducted on the public understanding of science.
-Part of this research are surveys measuring to which extent people have basic science knowledge (e.g.~``Does the Earth go around the Sun, or does the Sun go around the Earth?'').
+Much of the research on public understanding of science measures to which extent people have basic science knowledge (e.g.~``Does the Earth go around the Sun, or does the Sun go around the Earth?'').
 For us, these surveys provide a lower bound to how much people trust basic science: people who give the right answer (and not by chance) supposedly trust science on that question.
 Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge (Miller, 1998, 2004).
-For example, only 55\% in the US (in 2014) and only 46\% in the EU (in 2005) knew that antibiotics kill only bacteria but not viruses (Science Literacy et al., 2016).
+For example, only 55\% in the US (in 2014) and only 46\% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses (Science Literacy et al., 2016).
 A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.
 
-{[}To do: Make a descriptive overview of science knowledge across time/countries.
-This is a crucial point: if there are questions with specific scientific answers that just about everyone already knows, then in a sense our point is already proven.
-But there are enough questions where people's knowledge appears quite poor anyways{]}
+{[}To do: Make a descriptive overview of science knowledge across time/countries.{]}
 
@@
 
-In a series of studies, we ask basic science knowledge questions, inform people of the scientific consensus, and see if they accept it.
-Of particular interest are people who are low in TiS, and CTists.
+In a series of four pre-registered studies (n = 782), we asked participants about basic (studies 1, 2 and 3) and recent/advanced (study 4) science knowledge questions, informed them of the scientific consensus, and aked if they accept it.
+Of particular interest were people who are low in TiS, and CTists.
+Our main hypotheses were:
+
+\textbf{H1: Higher trust in science is associated with more science knowledge and more acceptance of the scientific consensus}
+
+\textbf{H2: Higher conspiracy thinking/belief is associated with less science knowledge and less acceptance of the scientific consensus}
 
@@
 We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently.
-Study 4 uses the same design as study 3, but asking science questions that people are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves {[}on an anti-vaxx sample again?{]}.
-
-Then: replications in other countries?
-
-\section{Overview of experiments}\label{overview-of-experiments}
-
-The goal is to see to which extent people do not accept basic science.
-
-\section{Experiment 1}\label{experiment-1}
-
-The main goal of experiment one was to test whether people would accept the scientific consensus on basic knowledge questions.
-Additionally, we wanted to know if both science knowledge and acceptance of the scientific consensus are associated with trust in science and conspiracy thinking.
-We had the following research questions:
-
-\textbf{RQ1: What is the average science knowledge score (1)?}
-
-\textbf{RQ2: What is the average acceptance of the scientific consensus (2)?}
-
-\textbf{RQ3: What is the relationship between trust in science and, respectively, (1) and (2)?}
-
-\textbf{RQ4: What is the relationship between conspiracy thinking and, respectively, (1) and (2)?}
-
-\subsection{Methods}\label{methods}
-
-\subsubsection{Participants}\label{participants}
-
-We recruited 200 participants from the US via prolific.
-6 participants failed our attention check, resulting in a final sample of 194 participants (98 female, 96 male; \(age_\text{mean}\): 41.71, \(age_\text{sd}\): 13.07, \(age_\text{median}\): 39).
-Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.
-
-\subsubsection{Procedure}\label{procedure}
+Study 4 uses the same design as study 3, but using a set of more recent, advance science questions that participants are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves.
+Just as study 3, study 4 is run on a sample of vaccine-skeptic participants.
+
+\section{Methods}\label{methods}
+
+\subsection{Procedure}\label{procedure}
 
@@
 All participants who did not answer ``1 = Never'' were excluded.
-Participants then read the following instructions:``We will ask you 11 questions about science.
-After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.'' Next, participants answered a set of 10 basic science questions, which were randomly selected from a pool of 11 questions, in random order.
+Participants then read the following instructions:``We will ask you 10 questions about science.
+After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.'' Next, participants answered a set of 10 basic science questions (in study 1 they were randomly selected from a pool of 11 questions) in random order.
 After each question, participants were presented with an answer reflecting the scientific consensus.
+In study 2 and 3, participants additionally saw a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
+In study 4, we provided only two links and no explanation.
 Participants were asked to choose whether they accept the answer or not, before proceeding to the next question.
-Figure \ref{fig:stimulus-example} displays the survey for an example science question.
-Finally, participants answered questions on conspiracy thinking and trust in science.
-
-\subsubsection{Materials}\label{materials}
-
-\paragraph{Science knowledge and acceptance}\label{science-knowledge-and-acceptance}
-
-Table \ref{tab:knowledge} shows all questions, their scientifically consensual answer, and their source.
-All but two questions were selected from existing science knowledge questionnaires.
-We tried to select non-political questions.
+Participants then answered questions on conspiracy thinking, conspiracy beliefs and trust in science.
+
+In study 2, 3 and 4, we presented participants with open-ended questions to explain cases where they rejected the scientific consensus.
+In study 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus.
+Finally, at the end of study 3 and 4, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
+Participants who selected ``You mostly agree with the consensus because you have been able to independently verify it'', were asked the open-ended follow-up question: ``Could you please tell us how you independently verified the information?''.
+
+\subsection{Participants}\label{participants}
+
+We recruited a total of 782 (194 in study 1, 190 in study 2, 200 in study 3, 198 in study 4) participants in the US via prolific.
+Details and demographics can be found in the online supplemental material.
+While samples for study 1 and 2 were convenience samples, study 3 and 4 were run on a sample holding vaccine-skeptic beliefs (see OSM for details).
+
+\subsection{Materials}\label{materials}
+
+Table \ref{tab:knowledge} shows all questions and their answer options used in the four studies.
 
@@
 
-\paragraph{Conspiracy scales}\label{conspiracy-scales}
-
-To measure conspiracy thinking, we selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) by Pennycook, Binnendyk, and Rand (2022) (Table \ref{tab:conspiracy}).
+Our main measure of conspiracy belief was asking participants about specific conspiracy theories.
+We selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) (Pennycook, Binnendyk, \& Rand, 2022) (Table \ref{tab:conspiracy}).
 Participants were asked: ``Below is a list of events for which the official version has been disputed.
@@
 {[}1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true{]}''.
+
+As a robustness check, we also used established scales on conspiracy thinking, namely the four-item conspiracy mentality questionnaire (CMQ) (Bruder, Haffke, Neave, Nouripanah, \& Imhoff, 2013) and the single item conspiracy beliefs scale (SICBS) (Lantian, Muller, Nurra, \& Douglas, 2016).
+Details and comparisons between the scales can be found in the OSM.
 
@@
 
-To cross-check our results with alternative measures, we also assessed the conspiracy mentality questionnaire (CMQ) by Bruder, Haffke, Neave, Nouripanah, and Imhoff (2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian, Muller, Nurra, and Douglas (2016) (see Appendix \ref{exp1}).
-
-\paragraph{Trust in science}\label{trust-in-science-1}
-
-Our main item for measuring trust in science is selected from the Wellcome Global Monitor survey: ``In general, would you say that you trust science a lot, some, not much, or not at all?
-{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''
-
+Our main item for measuring trust in science was selected from the Wellcome Global Monitor survey: ``In general, would you say that you trust science a lot, some, not much, or not at all?
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''.
 We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey (``How much do you trust scientists in this country?
@@
 
-\subsection{Results}\label{results}
-
-Regarding RQ1 and RQ2, participants answered on average 74 \% (sd = 0.16) of the questions correctly, and accepted the scientific consensus on average for 93 \% (sd = 0.12) of the questions.
-
-Fig.
-\ref{fig:exp1-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
-In most cases (76.3 \%), participants readily accepted the scientific consensus after having given the wrong answer to a question.
-In very few cases (1.6 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
-We believe this might have been due to inattention.
-
-For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29, p \textless{} .001), and acceptance of scientific consensus and trust in science (r = 0.27, p \textless{} .001).
-The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp1-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-
-For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = -0.38, p \textless{} .001), and conspiracy thinking and acceptance of scientific consensus (r = -0.33, p \textless{} .001).
-
-In Appendix \ref{exp1}, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \ref{exp1} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus?
-In our main analyses, we looked at correlations of acceptance across all observations.
-One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus.
-To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question.
-We than calculated the correlation between trust in science and average acceptance rate by participant.
-We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = -0.14, p = 0.061), nor with trust in science (r = 0.06, p = 0.387).
-
-\subsection{Discussion}\label{discussion}
-
-These results suggest that most people accept the scientific consensus most of the time.
-Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
-Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
-In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
-\section{Experiment 2}\label{experiment-2}
-
-In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
-In most instances they did, but not always.
-In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
-To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
-To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
-We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \ref{exp1}).
-
-Based on experiment 1, we formulated the following hypotheses:
-
-\textbf{H1a: Higher trust in science is associated with more science knowledge?}
-
-\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?}
-
-\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
-
-\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?}
-
-We had the following research questions:
-
-\textbf{RQ1: What is the average science knowledge score?}
-
-\textbf{RQ2: When a participant's answer does not match the consensus, how often do they change their mind and accept the consensus?}
-
-\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
-
-\subsection{Methods}\label{methods-1}
-
-\subsubsection{Participants}\label{participants-1}
-
-We recruited 201 participants from the US via prolific.
-11 participants failed our attention check, resulting in a final sample of 190 participants (96 female, 94 male; \(age_\text{mean}\): 43.48, \(age_\text{sd}\): 12.25, \(age_\text{median}\): 42).
-Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
-
-\subsubsection{Procedure}\label{procedure-1}
-
-The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
-\ref{fig:exp2-stimulus-example}.
-
-\subsubsection{Materials}\label{materials-1}
-
-We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
-
-\subsection{Results}\label{results-1}
-
-As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = 0.28, p \textless{} .001) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = -0.40, p \textless{} .001).
-By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
-For trust in science, we find no statistically significant correlation (r = 0.16, p = 0.051).
-For conspiracy thinking we find a small negative one (r = -0.22, p = 0.006).
-
-Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp2-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-In Appendix \ref{exp2} we show that these results hold for our alternative measures of trust and conspiracy thinking.
-We also include more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Regarding RQ1, participants answered on average 79 \% (sd = 0.19) of the questions correctly, and accepted the scientific consensus on average for 98 \% (sd = 0.05) of the questions.
-Fig.
-\ref{fig:exp2-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
-In response to RQ2, in most cases (92.9 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-In very few cases (0.5 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
-
-For RQ3, we got 35 answers from 25 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Table \ref{tab:exp2-justifications} summarizes these answers by five categories.
-All answers can be read in Appendix \ref{exp2}.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp2-justifications}Justifications by category}
-
-\begin{tabular}{llll}
-\toprule
-Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
-\midrule
-Personal convictions & 12.00 & 34.3\% & 8.00\\
-Mistake & 9.00 & 25.7\% & 8.00\\
-No justification & 7.00 & 20\% & 5.00\\
-Not convinced & 5.00 & 14.3\% & 5.00\\
-Religious Beliefs & 2.00 & 5.7\% & 2.00\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
-
-\subsection{Discussion}\label{discussion-1}
-
-Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.~answered the knowledge question wrongly).
-The share of these latter is considerably larger in experiment 2 (92.9 \%) than in experiment 1 (76.3 \%).
-While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
-We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
-
-\section{Experiment 3}\label{experiment-3}
+\section{Results}\label{results}
 
@@
 
-Study 3 is essentially a replication-\/--with some minor modifications-\/--of study 2, but on a different type of sample.
-Both study 1 and 2 were run on convenience samples.
-For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
-By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
-\ref{fig:exp3-explanation-example}).
-Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
-Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
-We want to know if participants perceive that this is because of trust, or other factors.
-
-As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
-
-\textbf{H1a: Higher trust in science is associated with more science knowledge?}
-
-\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus.}
-
-\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
-
-\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.}
-
-We had the following research questions:
-
-\textbf{RQ1: What is the average science knowledge score?}
-
-\textbf{RQ2: What is the average acceptance of the scientific consensus}
-
-\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
-
-\textbf{RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?}
-
-\subsection{Methods}\label{methods-2}
-
-\subsubsection{Participants}\label{participants-2}
-
-We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of 200 participants (125 female, 73 male; \(age_\text{mean}\): 42.90, \(age_\text{sd}\): 12.00, \(age_\text{median}\): 41).
-Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
-
-However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
-This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = 140 for trust in science, N = 138 for conspiracy measures, N = 122 for reason for accepting consensus).
-
-\subsubsection{Procedure}\label{procedure-2}
-
-The procedure was mostly the same as in experiment 2.
-In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
-Finally, at the end of the survey, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
-Participants who selected ``You mostly agree with the consensus because you have been able to independently verify it'', were asked the open-ended follow-up question: ``Could you please tell us how you independently verified the information?''.
-
-\subsubsection{Materials}\label{materials-2}
-
-We relied on the same items as in experiment 2.
-We used a pre-registered categorization scheme for the open ended answers for why people reject the consensus.
-The categories were XX and XX.
-{[}Describe coding process, with independent reviewer etc.{]}
-
-\subsection{Results}\label{results-2}
-
-As in study 1 and 2, we find that participants answered on average 75 \% (sd = 0.18) of the questions correctly, and initially accepted the scientific consensus on average for 96 \% (sd = 0.08) of the questions (RQ2).
-The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (98 \%, sd = 0.06).
-
-Fig.
-\ref{fig:exp3-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
-In most cases (87 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
-After providing the chance to revise the initial consensus rejection, this share is even larger (95.5 \%).
-In very few cases (0.9 \%), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
-This share drops slightly after providing the chance to revise the initial consensus rejection (0.7 \%).
-
-In the appendix XX we provide an analysis that this drop is statistically significant {[}TO DO{]}.
-
-For all correlations, we include opinion revisions for measuring consensus acceptance.
-We find no statistically significant correlation between science knowledge and trust in science (r = 0.14, p = 0.100), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = 0.20, p = 0.019).
-Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp3-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-We find no statistically significant correlation between conspiracy thinking and science knowledge (r = -0.16, p = 0.055), and between conspiracy thinking and acceptance of scientific consensus (r = c(cor = -0.023), c(t = -0.27), = 0.788, c(df = 136), -0.189, 0.145, Pearson's product-moment correlation, two.sided, {[}-0.189, 0.145{]}, p = 0.788).
-
-In Appendix \ref{exp3}, we show that these associations are statistically significant (except the last one?).
-We show that they are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \ref{exp3} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
-
-Regarding RQ1, participants answered on average 75 \% (sd = 0.18) of the questions correctly.
-For RQ3, we got 74 answers from 47 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Table \ref{tab:exp3-justifications} summarizes these answers by five categories.
-
-All answers can be read in the appendix.
+Across all studies we find that almost all participants trust basic and recent/advance science facts almost all the time.
+That is true even for participants who expressed low general trust in science, and those who have a tendency to conspiracy think or hold science-related conspiracy beliefs (Fig.
+\ref{fig:summary-plot}).
+
+Our results suggest that the more people tend to trust science, the more they tend to know about science and accept the scientific consensus.
+In study 2 and 3, we find a positive but small correlation between trust in science and both science knowledge (study 1: r = 0.29, p \textless{} .001; study 2: r = 0.28, p \textless{} .001), and acceptance of scientific consensus (study 1: r = 0.27, p \textless{} .001; study 2: r = 0.30, p \textless{} .001).
+In study 3 and 4--both run on vaccine-skeptic samples--we find no statistically significant correlation between science knowledge and trust in science (study 3: r = 0.14, p = 0.100; study 4: r = 0.10, p = 0.144), but a small positive correlation between acceptance of scientific consensus and trust in science (study 3: r = 0.20, p = 0.019; study 4: r = 0.16, p = 0.029).
+
+As for conspiracy beliefs, results are mixed.
+In studies 1 and 2, we find a small negative correlation between conspiracy thinking and both science knowledge (study 1: r = -0.38, p \textless{} .001; study 2: r = -0.40, p \textless{} .001), and acceptance of scientific consensus (study 1: r = -0.33, p \textless{} .001; study 2: r = -0.37, p \textless{} .001).
+However, in study 3 and 4, we do not find a statistically significant correlation with neither science knowledge (study 3: r = -0.16, p = 0.055; study 4: r = -0.02, p = 0.742), and between nor acceptance of scientific consensus (study 3: r = -0.02, p = 0.788; study 4: r = -0.05, p = 0.498).
+
+Overall, the correlations we find are relatively weak, which might be partly due to ceiling effects.
+This was the case in particular for acceptance: In all studies, in the vast majority of cases, participants accepted accepted science facts.
+The absence of correlations between conspiracy beliefs and science knowledge or acceptance in studies 3 and 4 might be explained by our sample selection.
+For these studies, we recruited samples scoring considerably higher in conspiracy belief than the convenience samples of study 1 and 2 (see OSM).
+The presence of effects in study 1 and 2, and the absence in study 3 and 4 might be because differences between moderate and high conspiracy thinking might not matter, while, differences between low and moderate/high conspiracy thinking do matter.
+
+Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
+Yes.
+In most cases (study 1: 76.3 \%; study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+Only in study 4 do we find evidence\footnote{For this analysis, we restricted our data to those cases in which participants initially gave the wrong answer.} that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = 0.06, p = 0.387; study 2: r = 0.16, p = 0.051; study 3: r = 0.04, p = 0.619; study 4: r = 0.15, p = 0.037) and only in study 2 evidence that it is associated with (less) conspiracy thinking (study 1: r = -0.14, p = 0.061; study 2: r = -0.22, p = 0.006; study 3: r = -0.04, p = 0.631; study 4: r = -0.05, p = 0.455).
+In very few cases, participants changed their mind away from the scientific consensus.
+These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: 1.6 \%; study 2: 0.5 \%; study 3: 0.7 \%; study 4: 5.7 \%).
+
+Why did participants reject the scientific consensus?
+We got a total of 364 answers (study 2: 35; study 3: 74; study 4: 255) from 167 (study 2: 25; study 3: 47; study 4: 95) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Based on the answers, we created five categories.
+Table \ref{tab:justifications} summarizes these answers by five categories.
+All answers can be read in the OSM.
 
@@
 
-\caption{\label{tab:exp3-justifications}Justifications by category}
+\caption{\label{tab:justifications}Justifications by category for data from studies 2, 3 and 4 combined.}
 
@@
 \midrule
-Personal convictions & 25.00 & 33.8\% & 19.00\\
-Mistake & 21.00 & 28.4\% & 18.00\\
-No justification & 18.00 & 24.3\% & 13.00\\
-Not convinced & 7.00 & 9.5\% & 3.00\\
-Religious Beliefs & 3.00 & 4.1\% & 3.00\\
+NA & 255.00 & 70.1\% & 95.00\\
+Personal convictions & 37.00 & 10.2\% & 27.00\\
+Mistake & 30.00 & 8.2\% & 26.00\\
+No justification & 25.00 & 6.9\% & 18.00\\
+Not convinced & 12.00 & 3.3\% & 8.00\\
+Religious Beliefs & 5.00 & 1.4\% & 5.00\\
 \bottomrule
@@
 
-For RQ4, we had 122 participants answering the question.
-Of these 41.8\% said they accepted the scientific consensus because they trust scientists on this question, while 47.5\% said they independently verified the fact.
-10.7\% answered with other ``other'' and gave an open-ended explanation (see Appendix).
-
-We also asked all 58 participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed in Appendix XX.
-
-\section{Discussion}\label{discussion-2}
-
-{[}We can draw some inspiration from the ``Westwood et al 2021 Current research overstates American support for political violence'' paper: we want to put in perspective the current rise (?
-discourse around?) in mistrust towards science{]}
-
-There is increasingly more talk of defiance towards science and towards experts more generally.
-In every population, many people only trust science `some of the time', and a sizable minority doesn't trust it.
-This appears difficult to reconcile with the fact that, in many countries, most people receive at least a basic science education, that science education (along with education more generally) is increasing, and that science education seems to be the main driver of trust in science.
-
-But how deep is distrust of science? In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
-
-Traditionally, people have interpreted people's lack of basic knowledge in light of the deficit-model. Here, we show that people readily accept the scientifically consensual answer to these questions.
-
-``The dominant approach to conceptualizing and measuring science literacy in population surveys has arisen out of work by Jon D.
-Miller and Kenneth Prewitt in the United States (see Miller, 1983, 1998, 2004) alongside collabora- tors in Great Britain (see Durant et al., 1989).
-Underlying these efforts appears to have been widespread concern among policy makers and the scientific com- munity that nonscientists were becoming skeptical about the benefits of sci- ence and that such skepticism might result in cuts to science funding that would harm the scientific progress that many argue underpins both American and European economic development (Bauer et al., 2007).
-The results of the U.S.
-portion of this work have formed the core of a chapter of a biennial report called Science and Engineering Indicators (hereafter, Indicators) that the National Science Board provides to Congress and the Executive Branch.
-Scholars have also used the raw data collected for Indicators (which is made publicly available) for peer-reviewed research (e.g., Gauchat, 2012; Losh, 2010), and other countries have used many of the Indicators' questions for their own national surveys (e.g., Bauer et al., 2012a; National Science Board, 2016).''
+Why did participants accept the scientific consensus?
+In studies 3 and 4, we had asked participants about cases where they agreed with the scientific consensus.
+A total of 320 (study 3: 122; study 4: 198) participants answered this question.
+Of these, 41.8\% in study 3 and 36.4\% in study 4 said they accepted the scientific consensus because they trust scientists, while 47.5\% in study 3 and 47\% in study 4 said they independently verified the fact.
+We asked all participants who gave this answer to explain how they did it (answers can be found in the OSM).
+10.7\% in study 3 and 16.7\% in study 4 answered with other ``other'' and gave an open-ended explanation (see OSM).
+
+More detailed results addressing all our pre-registered research questions can be found in the OSM.
+
+\section{Discussion}\label{discussion}
+
+Survey data from the US suggests a decline of trust in science in recent years (Brian \& Tyson, 2023).
+Across the globe, many people only trust science `some of the time' , and a sizable minority doesn't trust it (Wellcome Global Monitor, 2018, 2020).
+This appears difficult to reconcile with the fact that, in many countries, most people nowadays receive at least a basic education (Ritchie, Samborska, Ahuja, Ortiz-Ospina, \& Roser, 2023), and that science education has been found to be the most strongly associated variable with trust in science (Wellcome Global Monitor, 2018).
+
+But how deep is distrust of science?
+In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
+Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
+That is true not only for basic science questions (studies 1-3), but also for more recent scientific findings that most people are unfamiliar with (study 4).
+In study 1 the acceptance rate after having given a wrong answer was considerably lower (23.7 \%) than in subsequent studies (study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%).
+While this could be just sampling variation, it might be that adding explanations and/or sources convinced participants more than merely stating the consensus.
+In general, participants with lower trust in science and who believe more in conspiracy theories tended to both know less about science and accept the scientific consensus less.
+
+Scholars have long interpreted the lack of basic science knowledge as the root cause of distrust in and more generally unfavorable attitudes towards science, an explanation known under the term ``deficit-model'' (Durant et al., 1989; Miller, 1998, 2004).
+In line with proponents of the deficit-model, we find that people indeed err even on very basic science questions.
+However, we also find that almost everyone instantly accepted the scientific consensus once informed about it.
+Moreover, the same was true for non-basic science questions.
+This calls into question the deficit model: If participants trust science on basic (and non-basic) questions---as suggests immediate acceptance of the scientific consensus---then unfavorable attitudes towards science must have other explanations than a mere lack of knowledge.
+This argument is supported by past research.
+In a seminal meta-analysis, @allumScienceKnowledgeAttitudes2008 found correlations between science knowledge and general attitudes towards science to be at best small and largely context dependent {[}@allumScienceKnowledgeAttitudes2008{]}.
+A more recent report by the National Academies of Sciences, Engineering, and Medicine states that science knowledge is ``unlikely to substantially affect attitudes on scientific issues'' (National Academies of Sciences, 2017).
+
+How to make sense of the detachment of basic science knowledge from attitudes towards science?
+One explanation might be that most scientific findings are very abstract.
+For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second.
+It is therefor easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
+In other words, people might hold on to most scientific findings as reflective beliefs---beliefs that have little effect on our behaviors (Mercier, 2021).
+It is probably when it becomes tangible that science shape's people's attitudes.
+This can happen, for example, when science appears to support conclusions in favor of certain policies or moral positions, as for example in the case of anthropogenic climate change or vaccines.
+
+It is tempting, especially for scientists, to explain mistrust in science with ignorance.
+After all, careful proof of objective truth is what makes science trustworthy---that is, for scientists.
+For most people, what shapes trust in science might not be so different from what shapes trust in other institutions, such as the government, banks or church.
 
@@
 \url{https://doi.org/10.1177/0963662506070159}
+
+\bibitem[\citeproctext]{ref-brianAmericansTrustScientists2023}
+Brian, K., \& Tyson, A.
+(2023).
+\emph{Americans{'} trust in scientists, positive views of science continue to decline}.
+Retrieved from \url{https://www.pewresearch.org/science/2023/11/14/americans-trust-in-scientists-positive-views-of-science-continue-to-decline/}
 
@@
 
+\bibitem[\citeproctext]{ref-mercierHowGoodAre2021}
+Mercier, H. (2021). How Good Are We At Evaluating Communicated Information? \emph{Royal Institute of Philosophy Supplement}, \emph{89}, 257--272. \url{https://doi.org/10.1017/S1358246121000096}
+
 \bibitem[\citeproctext]{ref-millerMeasurementCivicScientific1998a}
@@
 
+\bibitem[\citeproctext]{ref-nationalacademiesofsciencesCommunicatingScienceEffectively2017}
+National Academies of Sciences, E. (2017). \emph{Communicating science effectively: A research agenda}. National Academies Press.
+
 \bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
@@
 \url{https://doi.org/10.31234/osf.io/d5fz2}
+
+\bibitem[\citeproctext]{ref-ritchieGlobalEducation2023}
+Ritchie, H., Samborska, V., Ahuja, N., Ortiz-Ospina, E., \& Roser, M. (2023). Global education. \emph{Our World in Data}. Retrieved from \url{https://ourworldindata.org/global-education}
 
@@
 
-\subsection{Materials}\label{materials-3}
+\subsection{Participants}\label{participants-1}
+
+We recruited 200 participants from the US via prolific.
+6 participants failed our attention check, resulting in a final sample of 194 participants (98 female, 96 male; \(age_\text{mean}\): 41.71, \(age_\text{sd}\): 13.07, \(age_\text{median}\): 39).
+Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.
+
+\subsection{Materials}\label{materials-1}
 
@@
 
-\subsubsection{Conspiracy scales}\label{conspiracy-scales-1}
+\subsubsection{Conspiracy scales}\label{conspiracy-scales}
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-2}
+\subsubsection{Trust in science}\label{trust-in-science-1}
 
@@
 
+\subsection{Results}\label{results-1}
+
+Regarding RQ1 and RQ2, participants answered on average 74 \% (sd = 0.16) of the questions correctly, and accepted the scientific consensus on average for 93 \% (sd = 0.12) of the questions.
+
+Fig.
+\ref{fig:exp1-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In most cases (76.3 \%), participants readily accepted the scientific consensus after having given the wrong answer to a question.
+In very few cases (1.6 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+We believe this might have been due to inattention.
+
+For RQ3, we find a positive but small correlation between both science knowledge and trust in science (r = 0.29, p \textless{} .001), and acceptance of scientific consensus and trust in science (r = 0.27, p \textless{} .001).
+The more people are knowledgeable about science and the more they tend to accept the scientific consensus, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp1-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+
+For RQ4, we find a negative correlation of similar magnitude between conspiracy thinking and science knowledge (r = -0.38, p \textless{} .001), and conspiracy thinking and acceptance of scientific consensus (r = -0.33, p \textless{} .001).
+
+In Appendix \ref{exp1}, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
+Appendix \ref{exp1} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Are trust in science and conspiracy thinking, respectively, associated with being more easily convinced of the scientific consensus?
+In our main analyses, we looked at correlations of acceptance across all observations.
+One possibility is that the associations between trust in science/conspiracy thinking and acceptance of scientific consensus are explained by science knowledge: People who give the right answer in the first place are more ready to accept the consensus, and trust in science/conspiracy thinking are mostly associated with this knowledge, but not with willingness to accept the consensus.
+To addressed this potential confound, in a non-preregistered analysis, we restricted our sample to cases where participants gave the wrong answer to the knowledge question.
+We then calculated the correlation between trust in science and average acceptance rate by participant.
+We find no statistically significant correlation of acceptance with neither conspiracy thinking (r = -0.14, p = 0.061), nor with trust in science (r = 0.06, p = 0.387).
+
+\subsection{Discussion}\label{discussion-1}
+
+These results suggest that most people accept the scientific consensus most of the time.
+Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
+Yet, in 23.7 \% of the cases, participants rejected the scientific consensus after having given the wrong answer, suggesting that simply stating the consensus is not sufficient to convince participants sometimes.
+In general, people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
+
 \subsection{Comparing items}\label{comparing-items}
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-3}
+\subsubsection{Trust in science}\label{trust-in-science-2}
 
@@
 
-\subsection{Trust in science}\label{trust-in-science-4}
+\subsection{Trust in science}\label{trust-in-science-3}
 
@@
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:unnamed-chunk-21}Reasons for consensus rejection}\\
+\caption{\label{tab:unnamed-chunk-22}Reasons for consensus rejection}\\
 \toprule
@@
 
-\subsection{Materials}\label{materials-4}
+In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
+In most instances they did, but not always.
+In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
+To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
+To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
+We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \ref{exp1}).
+
+Based on experiment 1, we formulated the following hypotheses:
+
+\textbf{H1a: Higher trust in science is associated with more science knowledge?}
+
+\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus, for participants who did not already know it?}
+
+\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
+
+\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus, for participants who did not already know it?}
+
+We had the following research questions:
+
+\textbf{RQ1: What is the average science knowledge score?}
+
+\textbf{RQ2: When a participant's answer does not match the consensus, how often do they change their mind and accept the consensus?}
+
+\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
+
+\subsection{Participants}\label{participants-2}
+
+We recruited 201 participants from the US via prolific.
+11 participants failed our attention check, resulting in a final sample of 190 participants (96 female, 94 male; \(age_\text{mean}\): 43.48, \(age_\text{sd}\): 12.25, \(age_\text{median}\): 42).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+\subsection{Procedure}\label{procedure-1}
+
+The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
+\ref{fig:exp2-stimulus-example}.
+
+\subsection{Materials}\label{materials-2}
+
+We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
 
@@
 \end{figure}
+
+\subsection{Results}\label{results-2}
+
+As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = 0.28, p \textless{} .001) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = -0.40, p \textless{} .001).
+By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
+For trust in science, we find no statistically significant correlation (r = 0.16, p = 0.051).
+For conspiracy thinking we find a small negative one (r = -0.22, p = 0.006).
+
+Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
+These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp2-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+In Appendix \ref{exp2} we show that these results hold for our alternative measures of trust and conspiracy thinking.
+We also include more descriptive statistics, such as knowledge and acceptance by science questions.
+
+Regarding RQ1, participants answered on average 79 \% (sd = 0.19) of the questions correctly, and accepted the scientific consensus on average for 98 \% (sd = 0.05) of the questions.
+Fig.
+\ref{fig:exp2-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In response to RQ2, in most cases (92.9 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In very few cases (0.5 \%), participants who gave the correct response afterwards rejected the scientific consensus, thereby contradicting their own initial response.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp2-justifications}Justifications by category}
+
+\begin{tabular}{llll}
+\toprule
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+\midrule
+Personal convictions & 12.00 & 34.3\% & 8.00\\
+Mistake & 9.00 & 25.7\% & 8.00\\
+No justification & 7.00 & 20\% & 5.00\\
+Not convinced & 5.00 & 14.3\% & 5.00\\
+Religious Beliefs & 2.00 & 5.7\% & 2.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+For RQ3, we got 35 answers from 25 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \ref{tab:exp2-justifications} summarizes these answers by five categories.
+All answers can be read in Appendix \ref{exp2}.
+
+\subsection{Discussion}\label{discussion-2}
+
+Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.~answered the knowledge question wrongly).
+The share of these latter is considerably larger in experiment 2 (92.9 \%) than in experiment 1 (76.3 \%).
+While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
+We also show, again, that people with lower trust in science and who believe more in conspiracy theories tend to both know less about science and accept the scientific consensus less.
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-5}
+\subsubsection{Trust in science}\label{trust-in-science-4}
 
@@
 
-\subsection{Trust in science}\label{trust-in-science-6}
+\subsection{Trust in science}\label{trust-in-science-5}
 
@@
 
-\subsection{Materials}\label{materials-5}
+Study 3 is essentially a replication-\/--with some minor modifications-\/--of study 2, but on a different type of sample.
+Both study 1 and 2 were run on convenience samples.
+For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
+By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
+\ref{fig:exp3-explanation-example}).
+Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
+Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
+We want to know if participants perceive that this is because of trust, or other factors.
+
+As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
+
+\textbf{H1a: Higher trust in science is associated with more science knowledge?}
+
+\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus.}
+
+\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
+
+\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.}
+
+We had the following research questions:
+
+\textbf{RQ1: What is the average science knowledge score?}
+
+\textbf{RQ2: What is the average acceptance of the scientific consensus}
+
+\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
+
+\textbf{RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?}
+
+\subsection{Participants}\label{participants-3}
+
+We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of 200 participants (125 female, 73 male; \(age_\text{mean}\): 42.90, \(age_\text{sd}\): 12.00, \(age_\text{median}\): 41).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
+This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = 140 for trust in science, N = 138 for conspiracy measures, N = 122 for reason for accepting consensus).
+
+\subsection{Procedure}\label{procedure-2}
+
+The procedure was mostly the same as in experiment 2.
+In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
+Finally, at the end of the survey, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
+Participants who selected ``You mostly agree with the consensus because you have been able to independently verify it'', were asked the open-ended follow-up question: ``Could you please tell us how you independently verified the information?''.
+
+\subsection{Materials}\label{materials-3}
 
@@
 
+\subsection{Results}\label{results-3}
+
+As in study 1 and 2, we find that participants answered on average 75 \% (sd = 0.18) of the questions correctly (RQ1), and initially accepted the scientific consensus on average for 96 \% (sd = 0.08) of the questions (RQ2).
+The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (98 \%, sd = 0.06).
+
+Fig.
+\ref{fig:exp3-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In most cases (87 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
+After providing the chance to revise the initial consensus rejection, this share is even larger (95.5 \%).
+In very few cases (0.9 \%), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
+This share drops slightly after providing the chance to revise the initial consensus rejection (0.7 \%).
+
+For all correlations, we include opinion revisions for measuring consensus acceptance.
+We find no statistically significant correlation between science knowledge and trust in science (r = 0.14, p = 0.100), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = 0.20, p = 0.019).
+Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp3-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+We find no statistically significant correlation between conspiracy thinking and science knowledge (r = -0.16, p = 0.055), and between conspiracy thinking and acceptance of scientific consensus (r = -0.02, p = 0.788).
+
+For RQ3, we got 74 answers from 47 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \ref{tab:exp3-justifications} summarizes these answers by five categories.
+
+All answers can be read in the appendix.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp3-justifications}Justifications by category}
+
+\begin{tabular}{llll}
+\toprule
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+\midrule
+Personal convictions & 25.00 & 33.8\% & 19.00\\
+Mistake & 21.00 & 28.4\% & 18.00\\
+No justification & 18.00 & 24.3\% & 13.00\\
+Not convinced & 7.00 & 9.5\% & 3.00\\
+Religious Beliefs & 3.00 & 4.1\% & 3.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+For RQ4, we had 122 participants answering the question.
+Of these 41.8\% said they accepted the scientific consensus because they trust scientists on this question, while 47.5\% said they independently verified the fact.
+10.7\% answered with other ``other'' and gave an open-ended explanation (see below for all open-ended answers).
+
+We also asked all 58 participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed below.
+
 \subsection{Comparing items}\label{comparing-items-2}
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-7}
+\subsubsection{Trust in science}\label{trust-in-science-6}
 
@@
 
-\subsection{Trust in science}\label{trust-in-science-8}
+\subsection{Trust in science}\label{trust-in-science-7}
 
@@
 
+\clearpage
+
+\section{Experiment 4}\label{exp4}
+
+Study 4 relies on broadly the same design as study 3, and is again run on a sample of participants with vaccine-skeptic beliefs.
+The main difference is that study 4 uses a set of science questions which, instead of being on well-established facts, we are about more recent discoveries (Table \ref{tab:exp4-knowledge}).
+Our main question is whether participants will nevertheless readily accept the scientific consensus, as they did for the basic facts in previous studies.
+
+As for the previous studies, we had the following hypotheses:
+
+\textbf{H1a: Higher trust in science is associated with more science knowledge?}
+
+\textbf{H1b: Higher trust in science is associated with more acceptance of the scientific consensus.}
+
+\textbf{H2a: Higher conspiracy thinking is associated with less science knowledge?}
+
+\textbf{H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.}
+
+We had the following research questions:
+
+\textbf{RQ1: What is the average science knowledge score?}
+
+\textbf{RQ2: What is the average acceptance of the scientific consensus}
+
+\textbf{RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?}
+
+\textbf{RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?}
+
+\subsection{Participants}\label{participants-4}
+
+We recruited 200 participants from the US via prolific, of which two failed our attention check, resulting in a final sample of 198 participants (98 female, 100 male; \(age_\text{mean}\): 42.67, \(age_\text{sd}\): 12.32, \(age_\text{median}\): 41).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+\subsection{Procedure}\label{procedure-3}
+
+The procedure was identical to the one in experiment 3.
+
+\subsection{Materials}\label{materials-4}
+
+\FloatBarrier
+
+Table \ref{tab:exp4-knowledge} shows all questions, their scientifically consensual answer, and their source.
+All but two questions were selected from existing science knowledge questionnaires.
+We tried to select non-political questions.
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{1em}>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{12em}}
+\caption{\label{tab:exp4-knowledge}Science knowledge items}\\
+\toprule
+id & Question & Scientific consensus & Sources\\
+\midrule
+1 & For which disease is the drug bedaquiline, developed in 2007, a treatment?
+[Tetanus; Tuberculosis; Malaria] & There is a consensus among scientists that the drug bedaquiline is used against tubercolisis.
+& Sources: Wikipedia: https://en.wikipedia.org/wiki/Bedaquiline ;  Journal of applied microbiology: https://enviromicro-journals.onlinelibrary.wiley.com/doi/full/10.1111/jam.14478\#:\textasciitilde{}:text=The\%20drug\%20bedaquiline\%20was\%20approved,\%2DTB\%20and\%20XDR\%2DTB\\
+2 & What is the maximum speed a proton can attain in the largest particle collider as to 2015?
+[90\% of the speed of light; 99\% of the speed of light; the speed of light] & There is a consensus among scientists that the maximum speed a proton can attain in the largest particle collider (the Large Hadron Collider) is 99\% of the speed of light.
+& Sources: Wikipedia: https://en.wikipedia.org/wiki/Large\_Hadron\_Collider\#Findings\_and\_discoveries ; CERN: https://home.cern/news/news/accelerators/first-successful-beam-record-energy-65-tev\\
+3 & Kepler-452b is an exoplanet revolving around the star Kepler-452.
+How far away from the star is it, as established by astronomers in 2015?
+[97 million mi; 1,2 million mi; 1254 million mi] & There is a consensus among scientists that Kepler-452b is 97 million mi away from its star Kepler-452.
+& Wikipedia: https://en.wikipedia.org/wiki/Kepler-452b; NASA: https://science.nasa.gov/exoplanet-catalog/kepler-452-b/\\
+4 & Using bomb-pusle dating with carbon 14, what is the age of the oldest known vertebrate, as established in 2016?
+[138 years; 205 years;  392 years] & There is a consensus among scientists that the oldest known vertebrate in 2016 was 392 years old, it is the Greenland shark.
+& Sources:  Wikipedia: https://en.wikipedia.org/wiki/Bomb\_pulse\#Applications Oxford University Research Archive: https://ora.ox.ac.uk/objects/uuid:6c040460-9519-4720-9669-9911bdd03b09\\
+5 & How many more glials cells are there in the brain in comparison with neurons, as established in 2016?
+[The same amount; Twice as many; Tenth as many] & There is a consensus among scientists that there is the same amount of glial cells and neurons in the brain.
+& Sources: The Journal of Comparative neurology: https://onlinelibrary.wiley.com/doi/10.1002/cne.24040; Wikipedia: https://en.wikipedia.org/wiki/Glia\#Total\_number\\
+\addlinespace
+6 & As predicted by the general theory of relativity, how many times would the Earth keep orbiting if the Sun disappeared, as established in 2012?
+[47 seconds; 8 minutes;  2 hours] & There is a consensus among scientists that if the Sun disappeared, the Earth would keep orbiting for 8 minutes.
+& Sources: Wikipedia: https://en.wikipedia.org/wiki/Gravity\#Tests\_of\_general\_relativity ; Science Bulletin: https://link.springer.com/article/10.1007/s11434-012-5603-3\\
+7 & What is the electric charge of the Higgs Boson, as established in 2012?
+[1.602176634 x 10-19; 0; 3.2x10-19C] & There is a consensus among scientists that the Higgs Boson does not have an electric charge, it is electrically neutral.
+& Sources: Wikipedia: https://en.wikipedia.org/wiki/Higgs\_boson\#cite\_ref-npr-interview\_189-0; Cern: https://home.cern/science/physics/higgs-boson\\
+8 & What is the age of the oldest materials formed on Earth, as established in 2020?
+[Less than 4.6 Ga; Around 4.6 Ga; More than 4.6 Ga] & There is a consensus among scientists that the oldest materials formed on Earth are more than 4.6 Ga years old.
+& Sources: Wikipedia: https://en.wikipedia.org/wiki/Oldest\_dated\_rocks ; Proceedings of the National Academy of Sciences: https://www.pnas.org/doi/10.1073/pnas.1904573117\\
+9 & With the best current cloning techniques, what is the average success rate when operated on mice, as of 2010?
+[2,7\%; 9,4\%; 17,2\%] & There is a consensus among scientists that the average success rate of the best current cloning techniques on mice is 9.4\%.
+& Sources: Wikipedia: https://en.wikipedia.org/wiki/Cloning\#cite\_note-112 ; Oxford Academic: https://academic.oup.com/biolreprod/article/83/6/929/2530195?login=false\\
+10 & What was the strength of the Earth magnetic field 3.7 billion years ago, as discovered this year?
+[15 microtesla; 30 microtesla; 45 microtesla] & There is a consensus among scientists that the strength of Earth magnetic field 3.7 billion years ago was 15 microtesla.
+& Sources: Phys.org : https://phys.org/news/2024-04-oldest-undisputed-evidence-earth-magnetic.html; Science News: https://www.sci.news/othersciences/geoscience/earliest-evidence-earths-magnetic-field-greenland-12884.html\\
+\bottomrule
+\end{longtable}
+\endgroup{}
+
+\subsection{Results}\label{results-4}
+
+By contrast to previous studies on basic science knowledge, participants on average answered only 36 \% (sd = 0.17) of the questions correctly (RQ1).
+Similar with previous studies, however, they initially accepted the scientific consensus on average for 86 \% (sd = 0.18) of the questions (RQ2).
+The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (91 \%, sd = 0.16).
+
+Fig.
+\ref{fig:exp4-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
+In most cases (83.7 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
+After providing the chance to revise the initial consensus rejection, this share is even larger (89.5 \%).
+In few (but still perhaps surprisingly many) cases (8.6 \%), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
+This share drops slightly after providing the chance to revise the initial consensus rejection (5.7 \%).
+
+For all correlations, we include opinion revisions for measuring consensus acceptance.
+We find no statistically significant correlation between science knowledge and trust in science (r = 0.10, p = 0.144), but a small positive correlation between acceptance of scientific consensus and trust in science (r = 0.16, p = 0.029).
+Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
+\ref{fig:exp4-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+We do not find a statistically significant correlation between neither conspiracy thinking and science knowledge (r = -0.02, p = 0.742), nor between conspiracy thinking and acceptance of scientific consensus (r = -0.05, p = 0.498).
+
+For RQ3, we got 255 answers from 95 different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \ref{tab:exp4-justifications} summarizes these answers by five categories.
+
+All answers can be read in the appendix.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp4-justifications}Justifications by category}
+
+\begin{tabular}{llll}
+\toprule
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+\midrule
+NA & 255.00 & 100\% & 95.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+For RQ4, only 36.4\% of participants said they accepted the scientific consensus because they trust scientists on this question, while 47\% said they independently verified the fact.
+16.7\% answered with other ``other'' and gave an open-ended explanation (see below for all open-ended answers).
+
+We also asked all 93 participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed below.
+
+\subsection{Comparing items}\label{comparing-items-3}
+
+\subsubsection{Conspiracy theories}\label{conspiracy-theories-3}
+
+Table \ref{tab:exp4-correlation-conspiracy} shows the correlations of the three different scales assessing conspiracy thinking.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp4-correlation-conspiracy}Correlations of the three different scales assessing conspiracy thinking}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{BCTI} & \multicolumn{1}{c}{CMQ} & \multicolumn{1}{c}{SICBS}\\
+\midrule
+BCTI & 1.00 & 0.48 & 0.27\\
+CMQ & 0.48 & 1.00 & 0.42\\
+SICBS & 0.27 & 0.42 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsubsection{Trust in science}\label{trust-in-science-8}
+
+Table \ref{tab:exp4-correlation-trust} shows the correlations of the three different items measuring trust in science.
+
+\begin{table}[h]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp4-correlation-trust}Correlations of the three different items measuring trust in science}
+
+\begin{tabular}{llll}
+\toprule
+ & \multicolumn{1}{c}{wgm\_sciencegeneral} & \multicolumn{1}{c}{wgm\_scientists} & \multicolumn{1}{c}{pew}\\
+\midrule
+wgm\_sciencegeneral & 1.00 & 0.71 & 0.57\\
+wgm\_scientists & 0.71 & 1.00 & 0.69\\
+pew & 0.57 & 0.69 & 1.00\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Correlations with alternative measures}\label{correlations-with-alternative-measures-3}
+
+Table \ref{tab:exp4-correlations-outcomes} shows the correlations between knowledge and acceptance, respectively, and outcome variables.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:exp4-correlations-outcomes}Correlations between knowledge and acceptance, respectively, and outcome variables}
+
+\begin{tabular}{lllllllll}
+\toprule
+std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+0.07 & -0.33 & = 0.742 & -0.16 & 0.12 & BCTI 
+(main conspiracy measure) & {}[-0.164, 0.117] & -0.02 & NA\\
+0.07 & 1.60 & = 0.112 & -0.03 & 0.25 & CMQ & {}[-0.027, 0.253] & 0.11 & NA\\
+0.07 & 0.55 & = 0.586 & -0.10 & 0.18 & SICBS & {}[-0.102, 0.18] & 0.04 & NA\\
+0.07 & 2.01 & = 0.045 & 0.00 & 0.28 & WGM trust scientists & {}[0.003, 0.282] & 0.14 & NA\\
+0.07 & 1.47 & = 0.144 & -0.04 & 0.24 & WGM trust general 
+(main trust measure) & {}[-0.036, 0.244] & 0.10 & NA\\
+0.07 & 2.04 & = 0.043 & 0.00 & 0.28 & PEW trust scientists & {}[0.005, 0.284] & 0.14 & NA\\
+0.07 & -0.68 & = 0.498 & -0.19 & 0.09 & BCTI 
+(main conspiracy measure) & {}[-0.189, 0.092] & NA & -0.05\\
+0.07 & -1.42 & = 0.156 & -0.24 & 0.04 & CMQ & {}[-0.241, 0.039] & NA & -0.10\\
+0.07 & -1.92 & = 0.056 & -0.28 & 0.00 & SICBS & {}[-0.276, 0.004] & NA & -0.14\\
+0.07 & 3.74 & < .001 & 0.12 & 0.39 & WGM trust scientists & {}[0.122, 0.394] & NA & 0.26\\
+0.07 & 2.20 & = 0.029 & 0.02 & 0.30 & WGM trust general 
+(main trust measure) & {}[0.016, 0.295] & NA & 0.16\\
+0.07 & 3.00 & = 0.003 & 0.07 & 0.35 & PEW trust scientists & {}[0.072, 0.348] & NA & 0.21\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\subsection{Results conditional on false responses}\label{results-conditional-on-false-responses-3}
+
+Table \ref{tab:exp4-false-response-regression} shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.
+
+\begin{table}
+
+\caption{\label{tab:exp4-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
+\centering
+\begin{tabular}[t]{lcccccc}
+\toprule
+  & BCTI\_avg & CMQ\_avg & SICBS & wgm\_scientists & wgm\_sciencegeneral & pew\\
+\midrule
+(Intercept) & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000\\
+ & (0.071) & (0.071) & (0.070) & (0.069) & (0.070) & (0.070)\\
+avg\_acceptance & -0.053 & -0.111 & -0.157* & 0.239*** & 0.148* & 0.186**\\
+ & (0.071) & (0.071) & (0.071) & (0.069) & (0.071) & (0.070)\\
+\midrule
+Num.Obs. & 198 & 198 & 198 & 198 & 198 & 198\\
+R2 & 0.003 & 0.012 & 0.025 & 0.057 & 0.022 & 0.035\\
+R2 Adj. & -0.002 & 0.007 & 0.020 & 0.052 & 0.017 & 0.030\\
+AIC & 566.3 & 564.4 & 561.9 & 555.2 & 562.5 & 559.9\\
+BIC & 576.2 & 574.3 & 571.8 & 565.1 & 572.4 & 569.8\\
+Log.Lik. & -280.165 & -279.212 & -277.970 & -274.621 & -278.246 & -276.951\\
+RMSE & 1.00 & 0.99 & 0.99 & 0.97 & 0.99 & 0.98\\
+\bottomrule
+\multicolumn{7}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
+\end{tabular}
+\end{table}
+
+\subsection{By-question variation}\label{by-question-variation-3}
+
+\subsubsection{Knowledge}\label{knowledge-3}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp4-questions-knowledge.pdf}
+\caption{\label{fig:exp4-questions-knowledge}Distribution of correct answers by question.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp4-questions-acceptance.pdf}
+\caption{\label{fig:exp4-questions-acceptance}Distribution of consensus acceptance by question.}
+\end{figure}
+
+\subsection{Trust in science}\label{trust-in-science-9}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp4-trust-scientists.pdf}
+\caption{\label{fig:exp4-trust-scientists}Distribution of trust in scientists.}
+\end{figure}
+
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-3}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp4-conspiracy-distribution.pdf}
+\caption{\label{fig:exp4-conspiracy-distribution}Distribution of conspiracy thinking.}
+\end{figure}
+
+\subsubsection{By-item variation}\label{by-item-variation-3}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp4-conspiracy-items.pdf}
+\caption{\label{fig:exp4-conspiracy-items}Distribution of conspiracy thinking by item.}
+\end{figure}
+
+\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection-2}
+
+\subsubsection{All raw answers}\label{all-raw-answers-2}
+
+\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp4-reasons-rejection}Reasons for consensus rejection}\\
+\toprule
+id & question & answer\\
+\midrule
+5 & planet & I thought it was fake\\
+6 & magnetic & I dont believe the earth has been here 3.7 billion years\\
+10 & orbit & I honestly do not know anything about this stuff so i took a guess\\
+10 & cloning & I took a guess\\
+12 & material & seemed like too long of a time\\
+\addlinespace
+14 & cloning & I didn't think that many mice are actually cloned.\\
+16 & orbit & I don't believe the Earth orbits the Sun.\\
+16 & material & Nothing but their hypotheses without real evidence.\\
+16 & magnetic & I'm not convinced the Earth existed 3.7 billion years ago.\\
+23 & vertebrate & I don't believe in this kind of carbon dating. We can only slightly estimate the age, therefore I highly doubt the exact number of 392 years is accurate.\\
+\addlinespace
+23 & material & Earth is not billions of years old. At most, maybe 10,000 years. There is absolutely no way anything on Earth is billions of years old.\\
+23 & magnetic & There was no magnetic field 3.7 billion years ago because the Earth was formed only around 6,000-10,000 years ago.\\
+24 & cloning & I think it's less than that percentage\\
+26 & vertebrate & no, just a guess\\
+26 & brain & I have no idea.\\
+\addlinespace
+26 & material & No Idea\\
+28 & material & I really could not explain this\\
+28 & cloning & No, I could not explain this either\\
+29 & orbit & Not enough science to prove.\\
+29 & material & How can anyone put an  that age on something? Humans were not around the track.\\
+\addlinespace
+29 & cloning & Do not believe that anything can be cloned 100\%\\
+32 & drug & Science is not always accurate.\\
+32 & planet & N/A\\
+32 & vertebrate & That doesn't seem accurate.\\
+32 & brain & No.\\
+\addlinespace
+32 & boson & No.\\
+32 & orbit & No.\\
+32 & material & No.\\
+32 & cloning & No.\\
+32 & magnetic & No.\\
+\addlinespace
+35 & vertebrate & No, because I just guessed.\\
+35 & orbit & No, I just guessed.\\
+35 & material & No\\
+41 & planet & i guessed\\
+41 & vertebrate & i guessed\\
+\addlinespace
+41 & boson & i guessed\\
+41 & magnetic & i guessed\\
+42 & brain & There was a difference between what wiki said and the other article.\\
+43 & vertebrate & I did not see anything to support this answer.\\
+43 & brain & I did not feel like this was the correct answer.\\
+\addlinespace
+43 & material & Based on the article, I did not see this stated.\\
+43 & cloning & There was nothing stated to prove this.\\
+44 & vertebrate & I don’t see how it is possible to know the oldest vertebrate when there are so many species out there undiscovered\\
+44 & brain & I don’t think I can explain it differently because I’m not an expert in science\\
+44 & orbit & I am unsure if I can explain it differently, but how would we know if it never happened?\\
+\addlinespace
+44 & cloning & Again, this is a very detailed scientific question.
+I feel it so I will have to have extensive knowledge to answer this or I would have to have done an experiment on my own to to know this.\\
+45 & drug & I am sure I heard it was used for malaria.\\
+45 & cloning & I am not sure cloning is a definite technique.\\
+49 & material & Religious reasons.\\
+49 & magnetic & Religious reasons.\\
+\addlinespace
+54 & vertebrate & did not know the answer\\
+54 & brain & did not know the answer\\
+58 & cloning & The success rate just seems too high.\\
+59 & vertebrate & The websites I read, seemed to have disagreed, unless I may have misread what I was fact checking.\\
+60 & vertebrate & No.\\
+\addlinespace
+60 & orbit & No.\\
+63 & planet & I don't know, but in my opinion when you are talking distance from star that is unknowable.\\
+63 & vertebrate & I was guessing.  But knowing age via carbon dating is just guessing as well\\
+63 & brain & I was guessing.  Same answer as others.\\
+63 & boson & I was guessing.  But same answer.  They pretend to know things that are unknowable.\\
+\addlinespace
+63 & orbit & I was guessing.  Same answer as others.  Unknowable\\
+63 & material & I was guessing.  This again is something unknowable.\\
+63 & speed & I was guessing.  This is unknowable.\\
+63 & magnetic & I was guessing.  This is unknowable.\\
+65 & vertebrate & I couldn’t locate the proof\\
+\addlinespace
+65 & speed & The proof wasn’t clear to me\\
+65 & cloning & Didn’t see the info to back this up\\
+67 & material & Just a guess, I didn't think it made sense.\\
+67 & speed & Figured it was the speed of light, turns out I was close.\\
+68 & orbit & I'm not really able to give an explanation\\
+\addlinespace
+68 & speed & not really\\
+68 & cloning & Not really\\
+69 & orbit & It didn't make sense to me at the time.\\
+69 & cloning & I thought it was easier to clone mice\\
+70 & orbit & these were hypothesis on the ones i got wrong.\\
+\addlinespace
+70 & speed & it was a hypothesis, i had no idea\\
+70 & magnetic & it was a hypothesis i had no clue\\
+72 & drug & I do not believe in this consensus since it usually is fabricated\\
+72 & vertebrate & This is usually fabricated\\
+72 & brain & This is normally a fabricated fact\\
+\addlinespace
+72 & boson & This is probably fabricated\\
+72 & orbit & I do not agree with nearly scientific fact since it is made up\\
+72 & material & This is probably made up\\
+72 & speed & I do think this consensus is faked\\
+72 & magnetic & these are all made up terms\\
+\addlinespace
+74 & vertebrate & Because there was not evidence that I understood to support this theory.\\
+74 & brain & Because the research that was provided does not fully support this theory.\\
+74 & boson & Because the research that I read did not support our object this statement.\\
+74 & orbit & The research that was credited does not give a direct amount of time just a formula to get the answer.\\
+74 & material & I disagree with this statement because the research that was given says 4.6 GA years old not more than.\\
+\addlinespace
+74 & speed & I disagree because there are not enough facts to support this claim.\\
+75 & orbit & Looking at the website given, I could not find the exact time of how long the Earth would keep orbiting, so I didn't agree.\\
+75 & speed & I could not find the answer to that in the given websites\\
+75 & cloning & I could not find data that supported that percentage.\\
+78 & material & No, Just a thought.\\
+\addlinespace
+78 & cloning & Just a random thought\\
+79 & orbit & N/A\\
+80 & vertebrate & I actually meant to press agree after reading the wikipedia , it had mentioned how the oldest 'recorded' vertebrate was 392 years old\\
+81 & drug & Malaria sounds more like the answer to this question, since it was developed more recently.\\
+81 & brain & This question I have no idea. But I would be willing to answer differently if needed.\\
+\addlinespace
+81 & orbit & The Earth would still keep its orbit if the sun disappeared. I believe it would still take 365 days to do one full orbit on its axis.\\
+83 & material & The Bible says the Earth is 6000 years old\\
+84 & material & From a biblical perspective, this is not the case. Other than that, no.\\
+87 & vertebrate & No, I was just guessing.\\
+87 & speed & No. I was just guessing.\\
+\addlinespace
+88 & planet & God hasn't created stars that are millions of miles away from Earth or each other.\\
+88 & orbit & God created the Earth as flat and stationary!\\
+88 & material & God created the Earth 6,000 years ago.\\
+88 & magnetic & God created the Earth 6,000 years ago.\\
+92 & cloning & I feel that they've been cloning people with pretty great success for some time now.\\
+\addlinespace
+92 & magnetic & Because I don't agree that the earth is billions years old, since I'm a Christian.\\
+93 & material & There is no way for man to know if it was actually formed on earth or if it was formed somewhere else and then ended up on earth.\\
+94 & boson & I agree\\
+94 & cloning & The articles I read were quite vague but all stated the success of cloning mice has been incredibly low, which made me disagree with the consensus\\
+96 & orbit & I felt if the Sun disappeared than the Earth orbiting would be fast than 8 minutes.\\
+\addlinespace
+96 & speed & I would like to change my answer to agree.\\
+96 & cloning & I would like to change my answer to agree.\\
+97 & material & How do they know how old anything is?.  Everything we have been told by the powers that be for our whole lives has been a lie.  It is hard to believe anything anymore.\\
+98 & speed & I was not able to find that information in the reference material.\\
+98 & cloning & I was not able to find that answer in the reference material.
+And I guess I have a deep distrust of scientists, as well.
+Even if I read something that a scientist said on Wikipedia or Google I would not believe it.
+Remember it was the scientists who said safe and effective.
+And the Nazis have been cloning people since the 30s, so why believe anything they say about cloning mice.\\
+\addlinespace
+104 & vertebrate & It was just a guess.\\
+104 & brain & It was just a guess.\\
+104 & boson & It was just a guess.\\
+104 & speed & It was just a guess.\\
+105 & drug & I am unfamiliar.\\
+\addlinespace
+105 & orbit & Again I have no knowledge. It was a guess\\
+106 & planet & Not sure I took a guess\\
+106 & vertebrate & Not sure I took a guess\\
+106 & brain & Not sure I took a guess\\
+106 & material & Not sure I took a guess\\
+\addlinespace
+108 & material & I don’t trust the sources as anyone can make changes\\
+113 & vertebrate & Because I dont think they really know themselves\\
+113 & material & The earth has only been around a short time and it thousands not millions or billions of years\\
+113 & magnetic & No dont believe the bible is the true map of mans existance and its not billions of years old\\
+114 & planet & I tried to research it and didnt find the answer so i just chose anything\\
+\addlinespace
+114 & vertebrate & I just chose an answer because I didn't know\\
+114 & material & I just chose anything because I didn't know the answer\\
+114 & cloning & I just chose anything because I didn't know the answer\\
+115 & vertebrate & It said the oldest known vertebrate was 252 years old\\
+115 & speed & I didn’t see anywhere in the articles that backs this claim.\\
+\addlinespace
+118 & vertebrate & Aren't dinosaur bones older than 392 years old?\\
+119 & material & If I understood the resources correctly I thought the time period was less. I'm questioning my answer now though!\\
+121 & vertebrate & one of the articles i looked at said it was a smaller number I can't remember the number.\\
+121 & material & I don't believe the earth is that old.\\
+121 & cloning & I am not completely sure.  the number just seemed very large and I think that many of them die during the process.\\
+\addlinespace
+121 & magnetic & I'm not sure.  I still do not believe that the Earth is that old.\\
+122 & speed & When reading the material that were given, I could not find the answer to confirm it. Therefore, since I could not confirm it, I did not accept the scientific consensus.\\
+124 & planet & I didn't really understand the information on this one.\\
+124 & boson & While reading the information provided I didn't see exact confirmation on if it has an electric charge or not.\\
+124 & material & The answer was vague and not exact it just said more than 4 billion but not more than 4.6 billion.\\
+\addlinespace
+128 & drug & Because I thought it was the answer I had out.\\
+128 & planet & It’s that far away.\\
+128 & vertebrate & It shows that it was correct\\
+128 & brain & Because of what I read.\\
+128 & orbit & I had read it wrong.\\
+\addlinespace
+130 & brain & Neurons are everywhere.  I assumed that glials cells were not as prevalent as neurons.\\
+130 & cloning & It was hard to believe science had progressed that much regarding cloning.  Cloning brings to mind science in the far future.\\
+132 & orbit & I don't have anything scientifically to offer other than I just don't believe that's true.\\
+133 & brain & I remember briefly coming across information on this subject and at the time they were thing they out numbered them 10.1. however that could of changed the more research that was done.\\
+133 & cloning & i just assumed that cloning would not have a high success rate from reading about years ago, but I assume things have changed.\\
+\addlinespace
+134 & planet & I do not believe we can measure the distance of stars.\\
+134 & orbit & I took a guess because I do not believe we can predict such a thing.\\
+134 & material & I do not believe the earth is that old.\\
+136 & cloning & I’m not sure that I believe that they are that successful.\\
+137 & material & Not enough scientific proof\\
+\addlinespace
+137 & cloning & I don't believe in cloning techniques\\
+137 & magnetic & It's possible that the earth magnetic field 3.7 billion years ago was 15 micro Tesla\\
+139 & planet & I misread and thought they were talking about the distant from earth to Kepler.\\
+139 & brain & It depends on the area of the brain and there could be some differences depending on which brains you are looking at.\\
+139 & magnetic & I just wasn't sure.\\
+\addlinespace
+140 & vertebrate & No.\\
+140 & brain & No\\
+140 & material & No\\
+140 & cloning & No\\
+143 & planet & I don't trust that astronomers can accurately measure this.\\
+\addlinespace
+143 & vertebrate & I don't trust carbon 14 dating. I don't know what bomb-pusle is.\\
+143 & orbit & I don't trust that scientists can estimate this accurately.\\
+143 & material & I don't trust that scientists have a way to measure this accurately.\\
+143 & magnetic & I don't trust that scientists have a way to measure this accurately.\\
+145 & magnetic & i cant explain why i answered differently i probably forgot\\
+\addlinespace
+151 & boson & In some theoretical models or speculative physics, there might be hypotheses about different properties of the Higgs boson.\\
+151 & speed & The exact percentage of the speed of light can vary slightly depending on the specific conditions and the state of the accelerator at any given time.
+Precision in reporting such values is crucial.\\
+151 & cloning & The success rate might also depend on the specific strain of mice used or the conditions under which the cloning was performed.
+Variations in experimental conditions can affect the reported success rates.\\
+152 & boson & Neutrality is a toxic line of thinking and only establishes constraints\\
+152 & material & Universe is very old yes\\
+\addlinespace
+152 & speed & LHC is not what they say is it\\
+152 & cloning & "sucess" for one thing but also this is vague and doesn't really say much\\
+152 & magnetic & No way to really know this for sure and it also doesnt matter much\\
+153 & orbit & I think if the sun disappeared the gravitational field would disappear nearly immediately; and the Earth would continue on in a straight line.\\
+154 & orbit & I accept the scientific consensus but I do not accept Wikipedia as a confirming source of scientific data.\\
+\addlinespace
+154 & material & I don't have an explanation as to why I answered differently.\\
+154 & speed & I thought from memory of my previous reading that my answer was correct.\\
+155 & planet & I believe that there is not a way to accurately measure such a thing\\
+155 & vertebrate & I believe that a lot of scientists make up numbers and the numbers aren't even accurate\\
+155 & brain & This one might be accurate but I just wanted to disagree because I feel like it might not be accurate\\
+\addlinespace
+155 & orbit & I believe that a lot of those things about the Earth orbiting and stuff like that are misunderstood and not true\\
+155 & material & I believe that the Earth is much younger than we're told and that allow those numbers about the Earth being trillions and trillions of years old is just crazy talk\\
+155 & magnetic & I believe that the Earth does not have a magnetic field that's just a theory that is wrong in my opinion\\
+157 & orbit & Sure, it's a prediction and one that doesn't sound very plausible to me.\\
+157 & material & I don't believe the earth is that old.\\
+\addlinespace
+157 & magnetic & Once again, I don't believe the earth is that old.\\
+159 & planet & Couldn't havesais it better myself\\
+159 & material & No\\
+163 & cloning & Because one of the links said 8 out of 10 success rate on cloning mice.\\
+163 & magnetic & Because i didn't see anything showing that the researcher pulled samples from the northern most region to do the dating.
+There had been 2 other heating events that could have corrupted the data if they took it from the southern area.\\
+\addlinespace
+164 & orbit & I don't fully trust the sources and the Sun's never disappeared so how would they know.\\
+167 & material & They are just guessing. There is no way to verify their estimates.\\
+167 & magnetic & Earth age estimates are just guesses there is no way to verify them.\\
+168 & planet & Scientific "Consensus" has low credibility in my perception.\\
+168 & brain & Scientific "Consensus" has low credibility in my perception.\\
+\addlinespace
+168 & boson & Scientific "Consensus" has low credibility in my perception.\\
+168 & orbit & Scientific "Consensus" has low credibility in my perception.\\
+168 & material & Scientific "Consensus" has low credibility in my perception.\\
+168 & speed & Scientific "Consensus" has low credibility in my perception.\\
+168 & cloning & Scientific "Consensus" has low credibility in my perception.\\
+\addlinespace
+168 & magnetic & Scientific "Consensus" has low credibility in my perception.\\
+169 & material & I don't think has to be that old.\\
+171 & vertebrate & It did not sound right to me, and that science can be wrong too.\\
+171 & cloning & I will agree with science on this one because they  create weird things.\\
+172 & speed & I thought the article showed a different answer, but I looked back and read it again and confirmed the original statement is in fact true\\
+\addlinespace
+172 & cloning & I thought the article showed a different answer, but I looked back and read it again and confirmed the original statement is in fact true\\
+173 & vertebrate & I think an older one has been found since 2016.\\
+173 & brain & I think it is more likely there is a differing amount which may be very small but I really have no idea.
+I just know we humans are highly fallible and fallen and are wrong about things all the time.\\
+173 & material & The Earth is no where near this old, it is around 6000-7000 years old.
+I cannot wait until we come to realize how utterly foolish we are in believing in the millions-billions or years old myth.\\
+173 & magnetic & Same reason as above.\\
+\addlinespace
+174 & boson & The website was not convincing\\
+174 & orbit & I thought it would be longer\\
+175 & material & My final authority is the Bible.  It only goes back 6500 years to the creation.\\
+175 & cloning & I just dont believe cloning techniques are that good my researched showed its around 1-2\%.\\
+175 & magnetic & Earth was created around 6500 years ago.\\
+\addlinespace
+177 & material & I am a creationist and believe the earth is much younger than that and am not convinced the testing methods are correct\\
+177 & cloning & number just seemed too high\\
+181 & brain & There is a similar amount, but not the same amount.\\
+181 & orbit & It's impossible to state as fact.\\
+181 & material & It's impossible to state as fact.\\
+\addlinespace
+181 & magnetic & It's impossible to state as fact.\\
+183 & planet & i read the distance in kilometers instead of miles.\\
+183 & magnetic & i read that during this time the microteslas was 30-50\\
+184 & vertebrate & Knowing how long vertebrates live today it seems unlikely that they lived significantly longer in the past\\
+184 & material & There is so much variation when it comes to measuring the early days of earth that I would assume given enough time this estimate will change dramatically in the future\\
+\addlinespace
+188 & vertebrate & No I am not able to explain my answer.\\
+188 & orbit & No.\\
+192 & vertebrate & Based on what I read it’s not true.\\
+192 & material & Based on what I read it’s not true.\\
+192 & speed & I’m not sure if it’s true because I couldn’t find the proof for the claim in the material.\\
+\addlinespace
+192 & cloning & I don’t know if it is true because I couldn’t find the proof in the material.\\
+194 & planet & It how I was taught.\\
+194 & boson & I don't know\\
+195 & planet & im a flat earther and i believe all the stars are right in our own atmosphere\\
+195 & orbit & i dont believe that the earth orbits around the sun at all to be honest i think the sun and moon travel over the earth\\
+\addlinespace
+195 & material & i dont believe that the earth is that old\\
+195 & magnetic & dont believe the earth is that old\\
+197 & orbit & I could not find the reference to corroborate the information.\\
+197 & material & The oldest material I found was just over 4.4 Ga\\
+197 & speed & I could not find the reference to corroborate the information.\\
+\bottomrule
+\end{longtable}
+
+\subsection{Why do people accept the scientific consensus:}\label{why-do-people-accept-the-scientific-consensus-1}
+
+\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other-2}
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp4-other-reasons-acceptance}Reasons for consensus rejection}\\
+\toprule
+id & other\_reason\_agreement\\
+\midrule
+8 & I didn’t know the answers at all so they could have been either one.\\
+12 & questions were out of my realm and I had no interest in researching it\\
+15 & I am not privy to information in he scientific field. I believe it to mostly be factual, but I'm unsure about discrepancies so I have to assume it may be true.\\
+20 & I agreed with certain things because it's just how I felt.\\
+23 & I didn't verify anything, nor do I really trust scientists, I guess I just believe that it's generally possible and seems like a reasonable calculation.\\
+\addlinespace
+25 & Not sure\\
+26 & I wasn't sure either way.\\
+35 & Neither, because I would have to do my own research beyond the untrustworthy sources you provided.\\
+39 & I mostly agree because I obtained information from opposing sources and formed an opinion.\\
+46 & I guessed\\
+\addlinespace
+63 & I haven't been able to independently verify it.  But I could if I wanted to and that information was something we could accurately figure out and prove it.\\
+85 & I mostly agree with the consensus because they weren't politically-charged topics, and also, I've been able to independently verify the statements.\\
+90 & There is no way for normal people to verify claims made by scientists with access to vast resources\\
+92 & Seemed to be accurate\\
+97 & I didn't really care what the answer was to those questions, so I just went with the flow.\\
+\addlinespace
+99 & I do not agree\\
+104 & It was just a guess.\\
+109 & i agreed with the consensus because i have not done my own research to refute it so what choice did i have.
+there’s not time enough on a survey for me to genuinely refute or agree so i mildly agree\\
+110 & I chose that I didn’t\\
+113 & I agree that things have been made and are scientifically real but not a billion or million years the bible does not confirm this and its smarter then science\\
+\addlinespace
+130 & I trust honest and ethical scientists.  The challenge is to find out who they are.  Science is an extraordinary discipline.  It should never be used to deceive and control us.\\
+132 & A combination of both as I didn't check all of them but the ones that I did in deed checked out.\\
+143 & I didn't have a reason to disagree with it.\\
+152 & ALmost everything is subjective and "science" is wrong again and again although some things are objectively true with empirical evidence like math and 2 + 2 = 4\\
+157 & I believe some of the questions are easier to verify than others.\\
+\addlinespace
+158 & I don't trust them that much because they are bought and paid for.\\
+162 & I agreed because of my trust in scientists.\\
+168 & I would agree if the subject involved could be independently verified, independent as in a non-aligned individual with any interests involving it.\\
+173 & A mix of both, it largely depends on the subject matter.
+I will say I do not have a whole lot of trust in scientists as they are by and large secular in their views which distorts them a tremendous amount.\\
+175 & The points I can agree with make sense and dont go against the Word of God.\\
+\addlinespace
+187 & I agreed because of my own research and knowledge coincides with the consensus.\\
+189 & Agreed because I didn’t want to verify the information.\\
+\bottomrule
+\end{longtable}
+
+\subsection{How did participants independently verifiy?}\label{how-did-participants-independently-verifiy-1}
+
+\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other-3}
+
+\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
+\caption{\label{tab:exp4-independent-verification}Reasons for consensus rejection}\\
+\toprule
+id & reason\_followup\\
+\midrule
+2 & Online ,Google, youtube\\
+3 & From the source and knowledge\\
+5 & I compared the articles and their sources for accuracy\\
+6 & Through research or facts I obtained previously\\
+11 & Through the links provided or a google search. Some of these things I've heard in passing on TV or the radio\\
+\addlinespace
+17 & its just information i have researched over the years.
+One example would be aliens.
+The US government denied aliens being real for years but a couple years ago they told the public they are in fact real.\\
+19 & I have not yet but will through factcheck.org.\\
+21 & Reading links and googling more articles\\
+24 & I have a friend and also Google.  I'm interested in scientific facts\\
+27 & After I was showed the answer, I copied and pasted in into search bar and looked at other sources.\\
+\addlinespace
+31 & I researched it on a fact checked website\\
+33 & I looked up each question. I have to just trust it at some point to answer the survey, but there are some things I'm not going to trust the internet for.\\
+34 & I dif a lot of research on some of the questions..\\
+36 & just an educated guess\\
+37 & Through painstakingly tiring, wikipedia combing and years of just generally being inquisitive\\
+\addlinespace
+38 & Do my own research times ten\\
+40 & I went to a couple of the provided links\\
+41 & i did not\\
+42 & By clicking on the links provided and reading the information.\\
+43 & I searched multiple sites.\\
+\addlinespace
+44 & I would have to conduct experiments sent independently, verify it or read trustworthy sources\\
+45 & Some of the information I have heard before.\\
+48 & By finding additional sources\\
+49 & My own research.\\
+50 & I go and read multiple articles or studies\\
+\addlinespace
+51 & Through study and resources.\\
+52 & I have done extensive research, read books from the library, watched DVDs and verified the information.
+After my mom died from a brain tumor I've researched the cures for cancer for 20 years now and know that doctors who come up with a cure are jailed or murdered to keep the status quo of the cancer industry.
+It's upsetting what our own government did in the 911 attacks on our own country as well.\\
+56 & There were 2 links shown to confirm the answers.  The references were from scientists as well.\\
+58 & There's lots of evidence.\\
+59 & I only quickly checked. I have a hard time following Science, as it is not something I find interesting. However, I do like to fact check things when possible, and when provoked to do so.\\
+\addlinespace
+60 & I verified the information by clicking the links provided and using critical reasoning.\\
+65 & Using the websites probided\\
+67 & I always like to fact check stuff because I don't necessarily believe things at face value.\\
+71 & I independently verified some of it a long time ago\\
+72 & By lots of reading, research, and logic common sense\\
+\addlinespace
+74 & I read the documents that were given myself\\
+75 & Read the provide websites to find agreeable information, also quick searches provided some verifications.\\
+76 & Wikepedia\\
+77 & I didn't\\
+79 & I used google. Which I would say is an elementary resource for information, but it was good for answering such questions\\
+\addlinespace
+82 & Non- mainstream media sources, government documents, intuition.\\
+83 & Research on my own\\
+86 & I researched on google.\\
+87 & By the websites you listed.\\
+95 & I knew some things and from personal opinions\\
+\addlinespace
+98 & I use the Brave browser. Less censorship.\\
+101 & I checked online with various sources to make sure they were true\\
+102 & Checking the websites provided\\
+107 & you need to add questions about where scientists get there funding from, and whether or not we would trust a statement like, all scientists agree\\
+108 & I just went with my gut\\
+\addlinespace
+111 & I like to read a lot from all sorts of different sources, then read arguments that disagree with them.
+Eventually I'll find a grain of truth in all of that and use pattern recognition to fill in gaps based on what I know.\\
+114 & Just did a google search\\
+115 & I read the articles.\\
+116 & Though previous knowledge and research\\
+117 & I clicked the links to read what was stated.\\
+\addlinespace
+118 & I did a quick look on Google and searched on my own\\
+119 & Google and the resources shown during the study\\
+120 & I do my own research.\\
+121 & i used the links in the study and looked briefly, at the articles.  Didn't understand most of it but made best decision I could.\\
+122 & I read the material that was given within each of the questions and found the information.\\
+\addlinespace
+123 & I checked various sources and informational circles\\
+124 & I searched up the information and researched myself.\\
+126 & I looked at the websites cited in the questions.\\
+127 & did web searches\\
+133 & Well not exactly verified , but Just things I have researched or came across while online. there is a lot of information out there and a lot of misinformation. You just have to take a side\\
+\addlinespace
+135 & doing research and reading articles and scientific studies.\\
+136 & By clicking the links provided\\
+137 & I somewhat trust scientist but overall, I independently do my own research to verify such existing's\\
+138 & Did much research and corroborated each finding or information and lined up against each other to see similar outcomes.\\
+139 & By looking up the links to reputable sources and seeing if they match up with what other reputable sources are saying.\\
+\addlinespace
+140 & Verified persons in specialized authority.\\
+141 & I just do my own research and look into things and how they could or couldn’t work\\
+145 & i was informed  by trusted sources\\
+147 & I checked the links given\\
+154 & Online research of more trusted sources\\
+\addlinespace
+155 & Basically it aligns with other things I believe to be true so I will agree with it I'm not saying that I went out of my way and did a bunch of different studies but the thing about a lot of scientific research and studies if the foundation of the study is unstable or unreliable then I can't trust that a lot of the results are accurate\\
+156 & I clicked on the articles and read for myself. Instead of trusting the information listed in the survey on its own I clicked the links to verify the information.\\
+163 & i either knew the information or i followed the links provided and believed the information.\\
+166 & I search multiple doctors and scientists and compare there work and theories\\
+167 & I read research papers and look for inconsistencies or bad methodology.\\
+\addlinespace
+170 & Google search it\\
+171 & I heard of these questions before, and I looked at the sources and my sources.\\
+172 & I read the attached files to see if the info could be backed up on there as well.
+The 2 I thought I read as false I ended up doing a google search to confirm that the statements were in fact true\\
+176 & I didn’t but I will\\
+179 & I research various sources that lean in either direction (left or right) as well as sources that appear, at least on the surface, to be impartial.
+After gathering all of this information, I make my informed decision.\\
+\addlinespace
+183 & asking ai such as chatgpt\\
+184 & Through my own research and comparison to other studies and others research.\\
+188 & I looked it up using the links provided to me.\\
+191 & By clicking and reading the link\\
+192 & I read the articles and found the actual answer to the question.\\
+\addlinespace
+194 & Online\\
+195 & google and research\\
+197 & With the references provided in the study.\\
+\bottomrule
+\end{longtable}
+
+\subsection{Additional plots}\label{additional-plots-3}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp4-conditional-acceptance.pdf}
+\caption{\label{fig:exp4-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/exp4-plot-overview.pdf}
+\caption{\label{fig:exp4-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
+\end{figure}
+
 \end{document}

### 2024-09-10 appendix_exp2.Rmd
before final edits Hugo, pre-submission

@@
 filter(term != "intercept") %>%
+    mutate_if(is.numeric, round, digits = 2) %>% 
+  # make a combined variable of estimate and p.value
+  mutate(estimate = paste0(estimate, " (p", p.value, ")")) %>% 
+  select(outcome, term, estimate) %>% 
   pivot_wider(names_from = term, 

### 2024-09-10 appendix_exp3.Rmd
before final edits Hugo, pre-submission

@@
 # Experiment 3 {#exp3}
+
+Study 3 is essentially a replication-–with some minor modifications-–of study 2, but on a different type of sample.
+Both study 1 and 2 were run on convenience samples.
+For study 3, we recruited a sample of people holding anti-vaccination beliefs (see below).
+By contrast to study 2, after asking participants an open question about why they did not accept the consensus (in cases where they didn't), we provide them with an explicit opportunity to change their answer (Fig.
+\@ref(fig:exp3-explanation-example)).
+Based on the answers to the open-ended questions, we also pre-registered a categorization scheme of reasons why people rejected the consensus.
+Finally, we also ask participants about why they agree with the scientific consensus on certain questions, in case they do.
+We want to know if participants perceive that this is because of trust, or other factors.
+
+As for study 2 (but without conditioning on wrong answers, as we did in study 1), we had the following hypotheses:
+
+**H1a: Higher trust in science is associated with more science knowledge?**
+
+**H1b: Higher trust in science is associated with more acceptance of the scientific consensus.**
+
+**H2a: Higher conspiracy thinking is associated with less science knowledge?**
+
+**H2b: Higher conspiracy thinking is associated with less acceptance of the scientific consensus.**
+
+We had the following research questions:
+
+**RQ1: What is the average science knowledge score?**
+
+**RQ2: What is the average acceptance of the scientific consensus**
+
+**RQ3: What reasons do participants provide to justify their rejection of the scientific consensus?**
+
+**RQ4: In case they agree with the scientific consensus, do people feel that this is because of trust?**
+
+## Participants
+
+We recruited a sample of participants holing vaccine-skeptic beliefs.
+Prolific Academic allows to filter out participants who have given specific answers to certain questions.
+We selected three available questions: 1.
+"Participants were asked the following question: Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: [For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say"].
+We selected particpants who answered "Against".
+2.
+"Participants were asked the following question: Have you received a coronavirus (COVID-19) vaccination?
+[Yes (at least one dose); No; Prefer not to answer]".
+We select only people who answered "No".
+3.
+"Participants were asked the following question: On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children.
+[1 (TOTALLY DISAGREEE); 2 (DISAGREE); 3 (SOMEWHAT DISAGREE); 4 (NEITHER AGREE NOR DISAGREE); 5 (SOMEWHAT AGREE); 6 (AGREE); 7 (TOTALLY AGREE); Rather not say"].
+We select only people who answered "1", "2", or "3".
+
+Based on these criteria, we recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of `r exp3_descriptives$n_subj` participants (`r exp3_descriptives$gender$female$n` female, `r exp3_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp3_descriptives$age$mean`, $age_\text{sd}$: `r exp3_descriptives$age$sd`, $age_\text{median}$: `r exp3_descriptives$age$median`).
+Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
+
+However, due to a randomization mistake for our outcomes, participant answered only two of the three outcome measure blocs (trust in science, conspiracy thinking, and reason for accepting consensus).
+This leaves us with reduced sample sizes for all analyses concerning these outcomes (N = `r exp3_descriptives$n_subject_by_outcome$wgm_sciencegeneral` for trust in science, N = `r exp3_descriptives$n_subject_by_outcome$BCTI_avg` for conspiracy measures, N = `r exp3_descriptives$n_subject_by_outcome$reason_agreement` for reason for accepting consensus).
+
+## Procedure
+
+The procedure was mostly the same as in experiment 2.
+In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
+Finally, at the end of the survey, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain.
+Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".
 
@@
 
+## Results
+
+As in study 1 and 2, we find that participants answered on average `r exp3_descriptives$means$knowledge_mean` (sd = `r exp3_descriptives$means$knowledge_sd`) of the questions correctly (RQ1), and initially accepted the scientific consensus on average for `r exp3_descriptives$means$acceptance_initial_mean` (sd = `r exp3_descriptives$means$acceptance_initial_sd`) of the questions (RQ2).
+The acceptance rate is even higher when accounting for opinion revisions towards acceptance of the consensus, after initial rejection (`r exp3_descriptives$means$acceptance_mean`, sd = `r exp3_descriptives$means$acceptance_sd`).
+
+Fig.
+\@ref(fig:exp3-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
+In most cases (`r exp3_descriptives$conditional_acceptance$false$acceptance_initial$Yes$share`), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
+After providing the chance to revise the initial consensus rejection, this share is even larger (`r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`).
+In very few cases (`r exp3_descriptives$conditional_acceptance$true$acceptance_initial$No$share`), participants who initially gave the correct response afterwards rejected the scientific consensus right after, thereby contradicting their own initial response.
+This share drops slightly after providing the chance to revise the initial consensus rejection (`r exp3_descriptives$conditional_acceptance$true$acceptance$No$share`).
+
+For all correlations, we include opinion revisions for measuring consensus acceptance.
+We find no statistically significant correlation between science knowledge and trust in science (r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`).
+Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
+\@ref(fig:exp3-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+We find no statistically significant correlation between conspiracy thinking and science knowledge (r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`), and between conspiracy thinking and acceptance of scientific consensus (r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`).
+
+For RQ3, we got `r exp3_descriptives$justifications_n` answers from `r exp3_descriptives$justifications_n_participants` different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Table \@ref(tab:exp3-justifications) summarizes these answers by five categories.
+
+All answers can be read in the appendix.
+
+```{r exp3-justifications}
+exp3_justifications %>%
+  group_by(category)%>%
+  summarize(n = n(), 
+            n_subjects = n_distinct(id)) %>%
+  mutate(Share = n/sum(n)) %>%
+  mutate_if(is.numeric, round, digits=3) %>% 
+  mutate(Share = paste0(Share*100, "%")) %>% 
+  arrange(desc(n)) %>% 
+  select(category, n, Share, n_subjects) %>% 
+  rename(Category = category,
+         `N (instances)` = n, 
+         `Share (instances)` = Share, 
+         `N (unique participants)` = n_subjects
+         ) %>% 
+  apa_table(caption = "Justifications by category")
+```
+
+For RQ4, we had `r exp3_descriptives$acceptance_n` participants answering the question.
+Of these `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share` said they accepted the scientific consensus because they trust scientists on this question, while `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share` said they independently verified the fact.
+`r exp3_descriptives$acceptance_by_reason[["other"]]$share` answered with other "other" and gave an open-ended explanation (see below for all open-ended answers).
+
+We also asked all `r exp3_descriptives$reason_followup_n` participants who answered that they had independently verified the answer to explain how they did so. The open-ended answers are listed below.
+
 ## Comparing items
@@
   select(wgm_sciencegeneral, wgm_scientists, pew) %>% 
-  cor()
+  cor(use = "complete.obs")
 
@@
   ) %>% 
-  mutate_if(is.numeric, round, digits = 2) %>% 
   mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 

### 2024-09-10 appendix_exp4.Rmd
before final edits Hugo, pre-submission

@@
 
-We recruited 200 participants from the US via prolific, of which two failed our attention check, resulting in a final sample of `r exp4_descriptives$n_subj` participants (`r exp4_descriptives$gender$female$n` female, `r exp4_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp4_descriptives$age$mean`, $age_\text{sd}$: `r exp4_descriptives$age$sd`, $age_\text{median}$: `r exp4_descriptives$age$median`).
+We recruited a sample of participants holing vaccine-skeptic beliefs.
+Prolific Academic allows to filter out participants who have given specific answers to certain questions.
+We selected three available questions: 1.
+"Participants were asked the following question: Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: [For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say"].
+We selected particpants who answered "Against".
+2.
+"Participants were asked the following question: Have you received a coronavirus (COVID-19) vaccination?
+[Yes (at least one dose); No; Prefer not to answer]".
+We select only people who answered "No".
+3.
+"Participants were asked the following question: On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children.
+[1 (TOTALLY DISAGREEE); 2 (DISAGREE); 3 (SOMEWHAT DISAGREE); 4 (NEITHER AGREE NOR DISAGREE); 5 (SOMEWHAT AGREE); 6 (AGREE); 7 (TOTALLY AGREE); Rather not say"].
+We select only people who answered "1", "2", or "3".
+
+Based on these criteria, we recruited 200 participants from the US via prolific, of which two failed our attention check, resulting in a final sample of `r exp4_descriptives$n_subj` participants (`r exp4_descriptives$gender$female$n` female, `r exp4_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp4_descriptives$age$mean`, $age_\text{sd}$: `r exp4_descriptives$age$sd`, $age_\text{median}$: `r exp4_descriptives$age$median`).
 Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
@@
   filter(term != "intercept") %>%
+  mutate_if(is.numeric, round, digits = 2) %>% 
+  # make a combined variable of estimate and p.value
+  mutate(estimate = paste0(estimate, " (p", p.value, ")")) %>% 
+  select(outcome, term, estimate) %>% 
   pivot_wider(names_from = term, 

### 2024-09-10 exp_4/cleaning.Rmd
before final edits Hugo, pre-submission

@@
 
-Add knowledge and acceptance by-participant averages to wide format data.
+Add knowledge and acceptance by-participant averages to wide format data. Also add sum of link clicks.
 
 ```{r}
-# make a version with averages per participant
+# make a version with averages per participant (and sum of link clicks)
 d_avg <- d_long %>% 
@@
             avg_knowledge = sum(knowledge_num)/n(),
-            avg_acceptance_initial  = sum(acceptance_initial_num)/n()
+            avg_acceptance_initial  = sum(acceptance_initial_num)/n(), 
+            sum_link_clicks = sum(link_clicks)
   )
@@
 justifications <- justifications %>%
-  mutate(category = case_when(final_coding == 1 ~ "No justification",
-                              final_coding == 2 ~  "Mistake",
-                              final_coding == 3 ~ "Not convinced",
-                              final_coding == 4 ~ "Personal convictions",
-                              final_coding == 5 ~ "Religious Beliefs",
-                              TRUE ~ NA_character_
-                              )
-         )
+  mutate(category = final_coding)
 

### 2024-09-10 paper.Rmd
before final edits Hugo, pre-submission

@@
 ---
-title             : "Universal acceptance of (nearly all of) basic science in the US"
-shorttitle        : "Acceptance of basic science"
+title             : "Quasi-universal acceptance of basic science in the US"
+shorttitle        : "Does anyone mistrust basic science?"
 
@@
 
-How much do people trust and accept science?
-The recent COVID pandemic has highlighted the importance of this question, with trust in science being the best predictor of people’s proclivity to follow health guidelines [@alganTrustScientistsTimes2021].
-
-Unfortunately, trust in science is far from being at ceiling [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020], and there are suggestions that it has recently been dropping in some countries [@alganTrustScientistsTimes2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly politicized [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
+Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change (ref) or vaccination (ref) to following recommendations during COVID [@alganTrustScientistsTimes2021], which suggests that trust in science was the most important predictor of these behaviors).
+
+Unfortunately, people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
+Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; although see @wellcomeglobalmonitorPublicTrustScientists2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
+Rejection of science can take the more extreme form of conspiracy theories, many of which question the scientific consensus on vaccination, climate change, and even the shape of the Earth.
+
+What does this apparent rejection of science actually entail?
+Do people who say they do not trust science, or who believe in conspiracy theories at odds with well-established science, reject most of science?
+Or, on the contrary, do they object to a few specific facets of science, while still accepting the overwhelming majority of basic science?
+Answering this question has theoretical implications: if some people reject science wholesale, then it’s possible that their distrust of science causes their rejection of specific scientific theories, or facilitates their acceptance of conspiracy theories; by contrast, if everyone trusts most of science, then the mistrust of specific facets of science is more likely to be post-hoc, a rationalization of other beliefs or behaviors.
+The present question also has practical implications: many communication attempts leverage the scientific consensus [e.g.
+on vaccination, climate change, etc., for review, see @vanstekelenburgScientificConsensusCommunicationContested2022; see also @veckalov27countryTestCommunicating2024].
+These attempts are more likely to be successful if everyone trusts basic science than if some people reject science wholesale.
+
+Studies of science knowledge provide relevant evidence, as people who answer science knowledge questions correctly presumably trust scientists in this respect.
+The average level of science knowledge is quite low, even in countries with extensive science education.
+For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016].
+However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.
 
@@
 To assess this, we can look at the literature on science knowledge.
+
+[introduce SK as: that gives us a lower bound of how much people trust basic science]
 
@@
 
-In a series of four pre-registered studies (n = 782), we asked participants about basic (studies 1, 2 and 3) and recent/advanced (study 4) science knowledge questions, informed them of the scientific consensus, and aked if they accept it.
-Of particular interest were people who are low in TiS, and CTists.
-Our main hypotheses were:
+In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts.
+For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted correct answer, and asked them whether they accepted it (measuring their trust in basic science).
+We also measured participants’ trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF)).
+The differences between the four studies are summarized presently, and the methods are detailed below.
+
+### Materials
+
+In Study 1, we used questions drawn from questionnaires of scientific knowledge (e.g.
+"Are electrons smaller, larger, or the same size as atoms?
+[Smaller; Same size; Larger]"), supplemented by a ‘trick’ question ("Where do trees mainly draw the materials with which they create their mass?
+[Earth; Water; Air]"; correct answer: Air).
+In Studies 2 and 3, this last question was removed.
+The scientific facts used in Studies 1 to 3 represent long-established and basic knowledge.
+In Study 4 we used more recent much less basic scientific discoveries (e.g.
+"How many more glial cells are there in the brain in comparison with neurons, as established in 2016?
+[The same amount; Twice as many; Tenth as many]"; correct answer: The same amount).
+
+### Presentation of the scientific consensus
+
+In Study 1, we simply told participants that they would be provided with the scientifically consensual answer.
+However, for participants to accept this answer, they must not only trust science, but also trust us (that we are presenting them with the actual scientifically consensual answer).
+To remove this issue, in Studies 2 to 3, we presented participants with a short explanation of the correct answer, as well as links to three sources per answer (e.g.
+Wikipedia, National Geographic or NASA).
+In Study 4, as the topics were more complex, we did not provide an explanation, but we provided two sources per answer.
+
+### Measure of acceptance of the scientific consensus
+
+In Study 1, we simply look at whether participants say that they accept the scientifically consensual answer.
+In the subsequent studies, we asked participants to explain why they disagreed with the scientific consensus.
+This revealed that a number of participants had made a mistake (misunderstanding, selecting the wrong answer).
+As a result, in Studies 3 and 4, participants who have indicated that they rejected the scientifically consensual answer were offered the option to revise their answer, or to keep rejecting it.
+
+### Additional questions
+
+In Studies 3 and 4, we attempted to understand how some people who say they do not trust science still accept scientifically consensual answers, by asking them whether they accepted the answers on the basis of trust in science or because they had independently verified them.
+
+### Samples
+
+Studies 1 and 2 were conducted on the standard sample of US participants recruited on the platform Prolific Academic.
+In order to increase the share of participants with low trust in science, and who endorse conspiracy theories, Studies 3 and 4 used the same platform, but only recruited participants who had declared previously being skeptical of vaccination.
+
+### Hypotheses
+
+The main goal of the present studies is descriptive: to find out whether participants who report not trusting science, or who believe in conspiracy theories, still accept most well-established scientific facts.
+However, we also tested two directional hypotheses (pre-registered as research questions in the Study 1):
 
@@
 **H2: Higher conspiracy thinking/belief is associated with less science knowledge and less acceptance of the scientific consensus**
-
-In study 1, we plainly inform participants of the scientific consensus and ask if they accept it.
-We use a subset of questions on basic science knowledge that have been used in numerous public opinion surveys [@allumScienceKnowledgeAttitudes2008; @durantPublicUnderstandingScience1989; @millerMeasurementCivicScientific1998a] sometimes referred to as the "Oxford scale" [@gauchatCulturalAuthorityScience2011].
-In study 2, we remove a problematic science question and present people with a more elaborated explanation of the scientific consensus, as well as additional sources.
-This allows us to remove the issue that participants might simply not have trusted us to correctly report the scientific consensus in study 1.
-Study 3 follows essentially the same design as study 2, but is run on a vaccine-skeptic sample.
-Further, we provided participants with an explicit option to revise their answer, in case they rejected the consensus.
-We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently.
-Study 4 uses the same design as study 3, but using a set of more recent, advance science questions that participants are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves.
-Just as study 3, study 4 is run on a sample of vaccine-skeptic participants.
 
@@
     mutate_if(is.numeric, round, digits = 2) %>% 
-  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
+    mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))),
+  means_numeric = exp1_wide %>% 
+    summarize(across(c(avg_knowledge, avg_acceptance,), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()),
   # RQ3
@@
     mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
+  means_numeric = exp2_wide %>% 
+    summarize(across(c(avg_knowledge, avg_acceptance,), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()),
   # H1
@@
 
+# Exploratory: Differences by reason for trust
+trust_by_reason_model <- run_regression(exp3_wide, dependent_var = "wgm_sciencegeneral", independent_var = "reason_agreement")
+
+acceptance_by_reason_model <-run_regression(exp3_wide, dependent_var = "avg_acceptance", independent_var = "reason_agreement")
+
 # extract descriptives for inline reporting
@@
   mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
+  means_numeric = exp3_wide %>% 
+    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()),
   # H1
@@
     filter(!is.na(reason_followup)) %>% 
-    nrow(.)
+    nrow(.),
+  # Exploratory: Differences by reason for trust
+  reason_means = exp3_wide %>% 
+    group_by(reason_agreement) %>% 
+    summarize(across(c(wgm_sciencegeneral, avg_acceptance), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()) %>% 
+    mutate_if(is.numeric, round, digits = 2) %>% 
+    mutate(across(acceptance_mean, ~paste0(.*100, " %"))) %>% 
+    split(.$reason_agreement), 
+  reason_trust_model = trust_by_reason_model %>% split(.$term), 
+  reason_acceptance_model = acceptance_by_reason_model %>% split(.$term)
 )
-
 ```
@@
 
+# Exploratory: Differences by reason for trust
+trust_by_reason_model <- run_regression(exp4_wide, dependent_var = "wgm_sciencegeneral", independent_var = "reason_agreement")
+
+acceptance_by_reason_model <- run_regression(exp4_wide, dependent_var = "avg_acceptance", independent_var = "reason_agreement")
+
+clicks_by_reason_model <- run_regression(exp4_wide, dependent_var = "sum_link_clicks", independent_var = "reason_agreement")
+
 # extract descriptives for inline reporting
@@
     mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
+  means_numeric = exp4_wide %>% 
+    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()),
   # H1
@@
     filter(!is.na(reason_followup)) %>% 
-    nrow(.)
+    nrow(.),
+  # Exploratory: Differences by reason for trust
+  reason_means = exp4_wide %>% 
+    group_by(reason_agreement) %>% 
+    summarize(across(c(wgm_sciencegeneral, avg_acceptance, sum_link_clicks), 
+                     list(mean = mean, sd = sd), 
+                     .names = "{.col}_{.fn}")) %>%
+    rename_with(~sub("^avg_", "", .), everything()) %>% 
+    mutate_if(is.numeric, round, digits = 2) %>% 
+    mutate(across(acceptance_mean, ~paste0(.*100, " %"))) %>% 
+    split(.$reason_agreement), 
+  reason_trust_model = trust_by_reason_model %>% split(.$term), 
+  reason_acceptance_model = acceptance_by_reason_model %>% split(.$term), 
+  reason_clicks_model = clicks_by_reason_model %>% split(.$term)
 )
-
-```
+```
+
+```{r combined-data}
+# combine data of all three studies
+combined_data <- bind_rows(exp1_long %>% 
+                             mutate(study = "1"), 
+                           exp2_long %>% 
+                             mutate(study = "2"), 
+                           exp3_long %>% 
+                             mutate(study = "3"),
+                           exp4_long %>% 
+                             mutate(study = "4")
+                           )
+
+# add a binned version for Conspiracy Thinking
+
+# Define breaks 
+breaks <- seq(0, 100, by = 20)
+# Define labels for each bin in the desired format
+labels <- paste0(c(0, 20, 40, 60, 80), "-", c(20, 40, 60, 80, 100))
+
+combined_data <- combined_data %>% 
+    # Create the binned version of CMQ_avg
+  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, 
+                              include.lowest = TRUE, right = FALSE))
+
+# extract descriptives for inline reporting
+combined_descriptives <- list(
+  # acceptance
+  mean_acceptance = combined_data %>% 
+  group_by(acceptance) %>% 
+  count() %>% 
+    ungroup() %>% 
+    mutate(share = n/sum(n)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    split(.$acceptance), 
+  # acceptance by trust
+  acceptance_by_trust = combined_data %>% 
+    group_by(acceptance, wgm_sciencegeneral) %>% 
+    count() %>% 
+    group_by(wgm_sciencegeneral) %>% 
+    mutate(share = n/sum(n)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    super_split(wgm_sciencegeneral, acceptance), 
+  participants_by_trust = combined_data %>% 
+  group_by(wgm_sciencegeneral) %>% 
+  summarize(n_participants = n_distinct(id)) %>% 
+    mutate(share = n_participants/sum(n_participants)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    split(.$wgm_sciencegeneral),
+  # acceptance by conspiracy thinking
+  acceptance_by_conspiracy_thinking = combined_data %>% 
+    group_by(acceptance, CMQ_avg_binned) %>% 
+    count() %>% 
+    group_by(CMQ_avg_binned) %>% 
+    mutate(share = n/sum(n)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    super_split(CMQ_avg_binned, acceptance),
+  participants_by_conspiracy_thinking  = combined_data %>% 
+    group_by(CMQ_avg_binned) %>% 
+    summarize(n_participants = n_distinct(id)) %>% 
+    mutate(share = n_participants/sum(n_participants)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    split(.$CMQ_avg_binned),
+  # acceptance by conspiracy belief
+  acceptance_by_conspiracy_belief = combined_data %>% 
+    # bring to long format
+    pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
+                 names_to = "BCTI_item", values_to = "score") %>% 
+    mutate(BCTI_item = sub("^BCTI_", "", BCTI_item)) %>% 
+    group_by(acceptance, BCTI_item, score) %>% 
+    count() %>% 
+    group_by(BCTI_item, score) %>% 
+    mutate(share = n/sum(n)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    super_split(acceptance, BCTI_item, score),
+  participants_by_conspiracy_belief  = combined_data %>% 
+    # bring to long format
+    pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
+                 names_to = "BCTI_item", values_to = "score") %>% 
+    mutate(BCTI_item = sub("^BCTI_", "", BCTI_item)) %>% 
+    group_by(BCTI_item, score) %>% 
+    summarize(n_participants = n_distinct(id)) %>% 
+    mutate(share = n_participants/sum(n_participants)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    super_split(BCTI_item, score),
+  # knowledge
+  mean_knowledge = combined_data %>% 
+    group_by(knowledge) %>% 
+    count() %>% 
+    ungroup() %>% 
+    mutate (share = n/sum(n), 
+            # rename knowledge values
+            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    split(.$knowledge)
+)
+```
+
+## Deviations from preregistration
+
+For study 2, we restricted our main hypotheses about acceptance to cases in which participants initially provided a wrong answer.
+The rationale was to see if trust in science or conspiracy belief were associated with a change of mind towards the consensus.
+We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM[^3].
+However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting simple, unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
+
+[^3]: Only in study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (study 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
 
@@
 Participants then read the following instructions:"We will ask you 10 questions about science.
-After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions (in study 1 they were randomly selected from a pool of 11 questions) in random order.
-After each question, participants were presented with an answer reflecting the scientific consensus.
-In study 2 and 3, participants additionally saw a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
-In study 4, we provided only two links and no explanation.
-Participants were asked to choose whether they accept the answer or not, before proceeding to the next question.
-Participants then answered questions on conspiracy thinking, conspiracy beliefs and trust in science.
-
-In study 2, 3 and 4, we presented participants with open-ended questions to explain cases where they rejected the scientific consensus.
-In study 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus.
-Finally, at the end of study 3 and 4, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain.
+After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions (in Study 1 they were randomly selected from a pool of 11 questions, as one ‘trick’ question had been added) in random order.
+After each question, participants were presented with an answer reflecting the scientific consensus, and asked whether they accepted it.
+In Studies 2 and 3, participants additionally saw a short explanation, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
+In Study 4, we provided only two links and no explanation.
+Participants then answered questions on conspiracy thinking, conspiracy beliefs, and trust in science.
+
+In Studies 2, 3, and 4, we presented participants with open-ended questions so they could explain their rejection of the scientific consensus.
+In Studies 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus.
+Finally, at the end of Studies 3 and 4, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain.
 Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".
@@
 
-We recruited a total of `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in study 1, `r exp2_descriptives$n_subj` in study 2, `r exp3_descriptives$n_subj` in study 3, `r exp4_descriptives$n_subj` in study 4) participants in the US via prolific.
+After removing failed attention checks, we were left with a total sample size of `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in Study 1, 6 failed attention checks; `r exp2_descriptives$n_subj` in Study 2, 11 failed attention checks; `r exp3_descriptives$n_subj` in Study 3, no failed attention checks; `r exp4_descriptives$n_subj` in Study 4, 2 failed attention checks) participants in the US via prolific.
 Details and demographics can be found in the online supplemental material.
-While samples for study 1 and 2 were convenience samples, study 3 and 4 were run on a sample holding vaccine-skeptic beliefs (see OSM for details).
+While samples for Studies 1 and 2 were convenience samples, Studies 3 and 4 were conducted on a sample holding vaccine-skeptic beliefs.
+Prolific Academic allows to select participants based on their answers to a range of questions.
+We picked three of these questions and only recruited participants who met our criteria for each of them:
+
+1.
+"Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: [For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say"].
+We selected particpants who answered "Against".
+2.  "Have you received a coronavirus (COVID-19) vaccination? [Yes (at least one dose); No; Prefer not to answer]". We select only people who answered "No".
+3.
+"On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children.
+[1 (totally disagree); 2 (disagree); 3 (somewhat disagree); 4 (neither agree nor disagree); 5 (somewhat agree); 6 (agree); 7 (totally agree); rather not say]".
+We select only people who answered '1', '2', or '3'.
 
@@
 
+### Scientific facts
+
+Studies 1 to 3 used 10 facts drawn from widely used questionnaires about science knowledge [@allumScienceKnowledgeAttitudes2008; @durantPublicUnderstandingScience1989; @millerMeasurementCivicScientific1998a] sometimes referred to as the "Oxford scale" [@gauchatCulturalAuthorityScience2011].
+A ‘trick’ question was added in Study 1 and removed as its wording proved unclear.
+Study 4 used 10 more recent scientific discoveries.
 Table \@ref(tab:knowledge) shows all questions and their answer options used in the four studies.
@@
 
-Our main measure of conspiracy belief was asking participants about specific conspiracy theories.
+### Conspiracy beliefs
+
 We selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) [@pennycookOverconfidentlyConspiratorialConspiracy2022] (Table \@ref(tab:conspiracy)).
@@
 
-As a robustness check, we also used established scales on conspiracy thinking, namely the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013] and the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016].
+### Conspiracy thinking
+
+We used the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013] and the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016].
 Details and comparisons between the scales can be found in the OSM.
@@
 kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL,
-    caption = "Conspiracy items", 
-    full_width = F) %>%
+    caption = "Conspiracy items", full_width = T) %>%
   kable_styling(font_size = 8) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "40em")
-```
-
-Our main item for measuring trust in science was selected from the Wellcome Global Monitor survey: "In general, would you say that you trust science a lot, some, not much, or not at all?
+  column_spec(1, width = "3em") %>%
+  column_spec(2, width = "40em") %>% 
+  footnote(general = "Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.",
+           escape = FALSE,
+           footnote_as_chunk = TRUE,
+           threeparttable = TRUE)
+```
+
+### Trust in science
+
+In all analyses reported in the main paper, we measure trust in science via a question selected from the Wellcome Global Monitor surveys [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020]: "In general, would you say that you trust science a lot, some, not much, or not at all?
 [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]".
-We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey ("How much do you trust scientists in this country?
+We chose this question as it seemed to be the most general one.
+In the appendix, we additionally report results for two alternative measures of trust included in our studies: Another from the WGM surveys ("How much do you trust scientists in this country?
 Do you trust them a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), the other from the Pew research center ("How much confidence do you have in scientists to act in the best interests of the public?
-[1-5; 1 = No confidence at all, 5 = A great deal of confidence]").
-We selected these items so that we could compare the ratings in our sample to global survey results.
-The WGM survey has been administered in over 140 countries and included over 140000 respondents.
-The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents [@colognaTrustScientistsTheir2024].
+[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), and one from the Pew Research Center ("How much confidence do you have in scientists to act in the best interests of the public?
+[1-5; 1 = No confidence at all, 5 = A great deal of confidence]"; see e.g.
+@funkKeyFindingsPublic2019 and used by e.g.
+@colognaTrustScientistsTheir2024).
+We selected these items so that we could compare the answers in our sample to global survey results.
+We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM).
 
@@
 
-(ref:summary-plot) Points represent the average share of acceptance as a function of the **A** level of trust in science ("In general, would you say that you trust science a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"); **B** average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); **C** belief in a specific conspiracy theory.
-Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: "1 - completly false, 5 - unsure, 9 - completely true".
-Averages are only based on participants who believed at least to some degree in the conspiracy theory.
-We consider a believer as anyone scoring higher than the scale midpoint, i.e.
-\>5.
-Number labels in plots represent the number of participants for the corresponding data point.
+(ref:summary-plot) Points represent the average share of acceptance and numbers the absolute count of participants as a function of: **A** the level of trust in science ("In general, would you say that you trust science a lot, some, not much, or not at all?
+[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"); **B** the average conspiracy thinking (CMQ, five items on a scale from 0 to 100); **C** the belief in specific conspiracy theories (i.e.
+participants who answered 9, "completely true", for a given theory, see Table \@ref(tab:conspiracy) for the list of the theories).
 
 ```{r summary-plot, fig.cap="(ref:summary-plot)", fig.height= 10, fig.width=10}
-# combine data of all three studies
-combined_data <- bind_rows(exp1_long %>% 
-                             mutate(study = "study 1"), 
-                           exp2_long %>% 
-                             mutate(study = "study 2"), 
-                           exp3_long %>% 
-                             mutate(study = "study 3"),
-                           exp4_long %>% 
-                             mutate(study = "study 4")
-                           )
-
 # make plot data for trust
@@
                 aes(x = wgm_sciencegeneral, y = share, color = study, shape = study)) +
-  geom_point(position = "dodge", alpha = 0.5) +
+  geom_pointrange(aes(ymin = 0, ymax = share), 
+                  position = position_dodge(width = 0.8), 
+                  alpha = 0.5, 
+                  show.legend = FALSE) +
+  geom_text(aes(label = n_participants, y = share + 0.01),  # Align text labels with dodged points
+            position = position_dodge(width = 0.8), 
+            vjust = -0.5, 
+            size = 3, 
+            show.legend = FALSE) +
+  #geom_point(position = "dodge", alpha = 0.5) +
+  # geom_text_repel(aes(label = paste0(n_participants)),
+  #                 vjust = -0.5, size = 3,
+  #                 show.legend = FALSE) +
   scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
-  geom_text_repel(aes(label = paste0(n_participants)),
-            vjust = -0.5, size = 3,
-            show.legend = FALSE) +
   scale_color_viridis_d(option = "plasma", end = 0.8) +
@@
   # make a belief variable if participant indicates a numeric answer above the
-  # scale midpoint
-  mutate(belief = ifelse(score > 5, TRUE, 
+  # a belief threshold 
+  mutate(belief = ifelse(score == 9, TRUE, 
                          ifelse(is.na(score), NA, FALSE)
                          ),
-         BCTI_item = sub("^BCTI_", "", BCTI_item)
+         BCTI_item = sub("^BCTI_", "", BCTI_item),
+         # rename BCTI items
+         BCTI_item = case_when(
+           BCTI_item == "apollo" ~ "Moon landing", 
+           BCTI_item == "cancer" ~ "Cancer cure", 
+           BCTI_item == "viruses" ~ "Viruses",
+           BCTI_item == "climatechange" ~ "Climate change",
+           BCTI_item == "flatearth" ~ "Flat earth",
+           BCTI_item == "autism" ~ "Vaccine autism",
+           BCTI_item == "polio" ~ "Vaccine cancer",
+           BCTI_item == "alien" ~ "Aliens",
+           BCTI_item == "hydorxychlor" ~ "Covid hydroxychloroquine",
+           BCTI_item == "evolution" ~ "Evolution"
+         )
          ) %>% 
@@
          aes(x = BCTI_item, y = share, color = study, shape = study)) +
-  geom_point(position = "dodge", alpha = 0.5) +
+  geom_pointrange(aes(ymin = 0, ymax = share), 
+                  position = position_dodge(width = 0.8), 
+                  alpha = 0.5, 
+                  show.legend = FALSE) +
+  geom_text(aes(label = n_participants, y = share + 0.01),  # Align text labels with dodged points
+            position = position_dodge(width = 0.8), 
+            vjust = -0.5, 
+            size = 3, 
+            show.legend = FALSE) +
+  # geom_point(position = "dodge", alpha = 0.5) +
+  # geom_text_repel(aes(label = paste0(n_participants)),
+  #                 vjust = -0.5, size = 3,
+  #                 show.legend = FALSE) +
   scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
-  geom_text_repel(aes(label = paste0(n_participants)),
-            vjust = -0.5, size = 3,
-            show.legend = FALSE) +
   scale_color_viridis_d(option = "plasma", end = 0.8) +
@@
                                aes(x = CMQ_avg_binned, y = share, color = study, shape = study)) +
-  geom_point(position = "dodge", alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
-  geom_text_repel(aes(label = paste0(n_participants)),
-            vjust = -0.5, size = 3,
+  geom_pointrange(aes(ymin = 0, ymax = share), 
+                  position = position_dodge(width = 0.8), 
+                  alpha = 0.5) +
+  geom_text(aes(label = n_participants, y = share + 0.01),  # Align text labels with dodged points
+            position = position_dodge(width = 0.8), 
+            vjust = -0.5, 
+            size = 3, 
             show.legend = FALSE) +
+  # geom_point(position = "dodge", alpha = 0.5) +
+  # geom_text_repel(aes(label = paste0(n_participants)),
+  #                 vjust = -0.5, size = 3,
+  #                 show.legend = FALSE) +
   scale_color_viridis_d(option = "plasma", end = 0.8) +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.01)) +
   labs(x = "Conspiracy Thinking", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
@@
 wrap_elements(patchwork) +
-  labs(tag = "Acceptance rate of consensus") +
+  labs(tag = "Rate of acceptance of consensual scientific facts") +
     theme(
@@
 
-Across all studies we find that almost all participants trust basic and recent/advance science facts almost all the time.
-That is true even for participants who expressed low general trust in science, and those who have a tendency to conspiracy think or hold science-related conspiracy beliefs (Fig.
-\@ref(fig:summary-plot)).
-
-Our results suggest that the more people tend to trust science, the more they tend to know about science and accept the scientific consensus.
-In study 2 and 3, we find a positive but small correlation between trust in science and both science knowledge (study 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`).
-In study 3 and 4--both run on vaccine-skeptic samples--we find no statistically significant correlation between science knowledge and trust in science (study 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`), but a small positive correlation between acceptance of scientific consensus and trust in science (study 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`).
-
-As for conspiracy beliefs, results are mixed.
-In studies 1 and 2, we find a small negative correlation between conspiracy thinking and both science knowledge (study 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`).
-However, in study 3 and 4, we do not find a statistically significant correlation with neither science knowledge (study 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`), and between nor acceptance of scientific consensus (study 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`).
-
-Overall, the correlations we find are relatively weak, which might be partly due to ceiling effects.
-This was the case in particular for acceptance: In all studies, in the vast majority of cases, participants accepted accepted science facts.
-The absence of correlations between conspiracy beliefs and science knowledge or acceptance in studies 3 and 4 might be explained by our sample selection.
-For these studies, we recruited samples scoring considerably higher in conspiracy belief than the convenience samples of study 1 and 2 (see OSM).
-The presence of effects in study 1 and 2, and the absence in study 3 and 4 might be because differences between moderate and high conspiracy thinking might not matter, while, differences between low and moderate/high conspiracy thinking do matter.
-
+The main outcome of interest is acceptance of the scientifically consensual facts presented.
+Overall, acceptance was very high (aggregating across all studies: `r combined_descriptives$mean_acceptance$Yes$share`; Study 1: `r exp1_descriptives$means$acceptance_mean`; Study 2: `r exp2_descriptives$means$acceptance_mean`; Study 3: `r exp3_descriptives$means$acceptance_mean`; Study 4: `r exp4_descriptives$means$acceptance_mean`).
+Note that this includes both participants who had previously correctly answered the knowledge question, and participants who changed their mind when presented with the scientific consensus.
+In Studies 3 & 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: `r exp3_descriptives$means$acceptance_initial_mean`; in Study 4: `r exp4_descriptives$means$acceptance_initial_mean`).
+
+As shown in Figure \@ref(fig:summary-plot), these very high rates of acceptance hold for: participants who they do not trust science at all (`r combined_descriptives$participants_by_trust$'1'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_trust$'1'$Yes$share`), participants who rank in the top two deciles of the conspiracy thinking scale (`r combined_descriptives$participants_by_conspiracy_thinking$'80-100'$share` of participants, average acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_thinking$'80-100'$Yes$share`), participants who consider as "completely true" (the maximum of the 9-point scale) conspiracy theories stating that the earth is flat (`r combined_descriptives$participants_by_conspiracy_belief$flatearth$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share`), or that climate change due to fossil emissions is a hoax (`r combined_descriptives$participants_by_conspiracy_belief$climatechange$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$climatechange$'9'$share`).
+This doesn't merely reflect science knowledge: aggregated across all studies--participants only correctly answered `r combined_descriptives$mean_knowledge$true$share` of the questions (Study 1: `r exp1_descriptives$means$knowledge_mean`; Study 2: `r exp2_descriptives$means$knowledge_mean`; Study 3: `r exp3_descriptives$means$knowledge_mean`; Study 4: `r exp4_descriptives$means$knowledge_mean`) before they were provided with the scientifically consensual answer.
 Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
@@
 In most cases (study 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-Only in study 4 do we find evidence[^3] that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in study 2 evidence that it is associated with (less) conspiracy thinking (study 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
 In very few cases, participants changed their mind away from the scientific consensus.
@@
 
-[^3]: For this analysis, we restricted our data to those cases in which participants initially gave the wrong answer.
+Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`), but a less consistent relation between trust in science and science knowledge (Studies 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`).
+
+Regarding H2, the results are mixed both for the relation between conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) and acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`) as well as science knowledge (Studies 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`).
 
@@
 ```{r justifications}
-
 ## make a common table of exp 2 and exp 3
@@
 Why did participants accept the scientific consensus?
-In studies 3 and 4, we had asked participants about cases where they agreed with the scientific consensus.
-A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (study 3: `r exp3_descriptives$acceptance_n`; study 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
-Of these, `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 4 said they accepted the scientific consensus because they trust scientists, while `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 4 said they independently verified the fact.
-We asked all participants who gave this answer to explain how they did it (answers can be found in the OSM).
-`r exp3_descriptives$acceptance_by_reason[["other"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in study 4 answered with other "other" and gave an open-ended explanation (see OSM).
+In Studies 3 and 4--the vaccine skeptic samples--we had asked participants about cases where they agreed with the scientific consensus.
+A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (Study 3: `r exp3_descriptives$acceptance_n`; Study 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than there were participants saying it was because they trust scientists (Study 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`).
+We asked participants who said they verified independently to explain how they did it (answers can be found in the ESM).
+`r exp3_descriptives$acceptance_by_reason[["other"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in study 4 answered with other "other" and gave an open-ended explanation (see ESM).
+
+In an exploratory analysis, we ran linear regressions to test whether there are differences between these two groups.
+Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = `r exp3_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 4; Study 4: mean = `r exp4_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value`) than those who said they verified independently (Study 3 mean = `r exp3_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`; Study 4 mean = `r exp4_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`).
+We did not find a difference regarding acceptance (Study 3: $\hat{\beta}_{\text{Acceptance}}$ = `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value` on a scale from 0 to 1; Study 4: $\hat{\beta}_{\text{Acceptance}}$ = `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value`).
+In study 4, where we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+On average, participants clicked only clicked on `r exp4_descriptives$reason_means$'trust in scientists'$sum_link_clicks_mean` links (out of 20 possible clicks) and there was no difference between the two groups ($\hat{\beta}_{\text{Clicks}}$ = `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$estimate`, `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$p.value`).
 
@@
 In line with proponents of the deficit-model, we find that people indeed err even on very basic science questions.
-However, we also find that almost everyone instantly accepted the scientific consensus once informed about it.
+However, we also find that almost everyone accepted the scientific consensus once informed about it.
 Moreover, the same was true for non-basic science questions.
@@
 For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second.
-It is therefor easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
+It is therefore easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
 In other words, people might hold on to most scientific findings as reflective beliefs—beliefs that have little effect on our behaviors [@mercierHowGoodAre2021].

### 2024-09-10 paper.tex
before final edits Hugo, pre-submission

@@
 \hypersetup{
-  pdftitle={Universal acceptance of (nearly all of) basic science in the US},
+  pdftitle={Quasi-universal acceptance of basic science in the US},
   pdflang={en-EN},
@@
 
-\title{Universal acceptance of (nearly all of) basic science in the US}
+\title{Quasi-universal acceptance of basic science in the US}
 \author{\textsuperscript{}}
@@
 
-\shorttitle{Acceptance of basic science}
+\shorttitle{Does anyone mistrust basic science?}
 
@@
 
-How much do people trust and accept science?
-The recent COVID pandemic has highlighted the importance of this question, with trust in science being the best predictor of people's proclivity to follow health guidelines (Algan, Cohen, Davoine, Foucault, \& Stantcheva, 2021).
-
-Unfortunately, trust in science is far from being at ceiling (Wellcome Global Monitor, 2018, 2020), and there are suggestions that it has recently been dropping in some countries (Algan et al., 2021), including the US (Brian \& Tyson, 2023), where it is also increasingly politicized (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
+Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change (ref) or vaccination (ref) to following recommendations during COVID (Algan, Cohen, Davoine, Foucault, \& Stantcheva, 2021), which suggests that trust in science was the most important predictor of these behaviors).
+
+Unfortunately, people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia (Wellcome Global Monitor, 2018, 2020).
+Moreover, trust in science has recently been declining in some countries (Algan et al., 2021; although see Wellcome Global Monitor, 2021), including the US (Brian \& Tyson, 2023), where it is also increasingly polarizing (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
+Rejection of science can take the more extreme form of conspiracy theories, many of which question the scientific consensus on vaccination, climate change, and even the shape of the Earth.
+
+What does this apparent rejection of science actually entail?
+Do people who say they do not trust science much, or who believe in conspiracy theories at odds with well-established science, reject most of science?
+Or, on the contrary, do they object to a few specific facets of science, while still accepting the overwhelming majority of basic science?
+Answering this question has theoretical implications: if some people reject science wholesale, then it's possible that their distrust of science causes their rejection of specific scientific theories, or facilitates their acceptance of conspiracy theories; by contrast, if everyone trusts most of science, then the mistrust of specific facets of science is more likely to be post-hoc, a rationalization of other beliefs or behaviors.
+The present question also has practical implications: many communication attempts leverage the scientific consensus (e.g.~on vaccination, climate change, etc., for review, see Van Stekelenburg, Schaap, Veling, Van 'T Riet, \& Buijzen, 2022; see also Većkalov et al., 2024).
+These attempts are more likely to be successful if everyone trusts basic science than if some people reject science wholesale.
 
@@
 To assess this, we can look at the literature on science knowledge.
+
+{[}introduce SK as: that gives us a lower bound of how much people trust basic science{]}
 
@@
 
-In a series of four pre-registered studies (n = 782), we asked participants about basic (studies 1, 2 and 3) and recent/advanced (study 4) science knowledge questions, informed them of the scientific consensus, and aked if they accept it.
-Of particular interest were people who are low in TiS, and CTists.
-Our main hypotheses were:
+In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts.
+For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted correct answer, and asked them whether they accepted it (measuring their trust in basic science).
+We also measured participants' trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF)).
+The differences between the four studies are summarized presently, and the methods are detailed below.
+
+\subsubsection{Materials}\label{materials}
+
+In Study 1, we used questions drawn from questionnaires of scientific knowledge (e.g.~``Are electrons smaller, larger, or the same size as atoms?
+{[}Smaller; Same size; Larger{]}''), supplemented by a `trick' question (``Where do trees mainly draw the materials with which they create their mass?
+{[}Earth; Water; Air{]}''; correct answer: Air).
+In Studies 2 and 3, this last question was removed.
+The scientific facts used in Studies 1 to 3 represent long-established and basic knowledge.
+In Study 4 we used more recent much less basic scientific discoveries (e.g.~``How many more glial cells are there in the brain in comparison with neurons, as established in 2016?
+{[}The same amount; Twice as many; Tenth as many{]}''; correct answer: The same amount).
+
+\subsubsection{Presentation of the scientific consensus}\label{presentation-of-the-scientific-consensus}
+
+In Study 1, we simply told participants that they would be provided with the scientifically consensual answer.
+However, for participants to accept this answer, they must not only trust science, but also trust us (that we are presenting them with the actual scientifically consensual answer).
+To remove this issue, in Studies 2 to 3, we presented participants with a short explanation of the correct answer, as well as links to three sources per answer (e.g.~Wikipedia, National Geographic or NASA).
+In Study 4, as the topics were more complex, we did not provide an explanation, but we provided two sources per answer.
+
+\subsubsection{Measure of acceptance of the scientific consensus}\label{measure-of-acceptance-of-the-scientific-consensus}
+
+In Study 1, we simply look at whether participants say that they accept the scientifically consensual answer.
+In the subsequent studies, we asked participants to explain why they disagreed with the scientific consensus.
+This revealed that a number of participants had made a mistake (misunderstanding, selecting the wrong answer).
+As a result, in Studies 3 and 4, participants who have indicated that they rejected the scientifically consensual answer were offered the option to revise their answer, or to keep rejecting it.
+
+\subsubsection{Additional questions}\label{additional-questions}
+
+In Studies 3 and 4, we attempted to understand how some people who say they do not trust science still accept scientifically consensual answers, by asking them whether they accepted the answers on the basis of trust in science or because they had independently verified them.
+
+\subsubsection{Samples}\label{samples}
+
+Studies 1 and 2 were conducted on the standard sample of US participants recruited on the platform Prolific Academic.
+In order to increase the share of participants with low trust in science, and who endorse conspiracy theories, Studies 3 and 4 used the same platform, but only recruited participants who had declared previously being skeptical of vaccination.
+
+\subsubsection{Hypotheses}\label{hypotheses}
+
+The main goal of the present studies is descriptive: to find out whether participants who report not trusting science, or who believe in conspiracy theories, still accept most well-established scientific facts.
+However, we also tested two directional hypotheses (pre-registered as research questions in the Study 1):
 
@@
 
-In study 1, we plainly inform participants of the scientific consensus and ask if they accept it.
-We use a subset of questions on basic science knowledge that have been used in numerous public opinion surveys (Allum, Sturgis, Tabourazi, \& Brunton-Smith, 2008; Durant, Evans, \& Thomas, 1989; Miller, 1998) sometimes referred to as the ``Oxford scale'' (Gauchat, 2011).
-In study 2, we remove a problematic science question and present people with a more elaborated explanation of the scientific consensus, as well as additional sources.
-This allows us to remove the issue that participants might simply not have trusted us to correctly report the scientific consensus in study 1.
-Study 3 follows essentially the same design as study 2, but is run on a vaccine-skeptic sample.
-Further, we provided participants with an explicit option to revise their answer, in case they rejected the consensus.
-We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently.
-Study 4 uses the same design as study 3, but using a set of more recent, advance science questions that participants are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves.
-Just as study 3, study 4 is run on a sample of vaccine-skeptic participants.
-
 \section{Methods}\label{methods}
+
+\subsection{Deviations from preregistration}\label{deviations-from-preregistration}
+
+For study 2, we restricted our main hypotheses about acceptance to cases in which participants initially provided a wrong answer.
+The rationale was to see if trust in science or conspiracy belief were associated with a change of mind towards the consensus.
+We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM\footnote{Only in study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = 0.06, p = 0.387; study 2: r = 0.16, p = 0.051; study 3: r = 0.04, p = 0.619; study 4: r = 0.15, p = 0.037) and only in study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (study 1: r = -0.14, p = 0.061; study 2: r = -0.22, p = 0.006; study 3: r = -0.04, p = 0.631; study 4: r = -0.05, p = 0.455).}.
+However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting simple, unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
 
@@
 Participants then read the following instructions:``We will ask you 10 questions about science.
-After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.'' Next, participants answered a set of 10 basic science questions (in study 1 they were randomly selected from a pool of 11 questions) in random order.
-After each question, participants were presented with an answer reflecting the scientific consensus.
-In study 2 and 3, participants additionally saw a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
-In study 4, we provided only two links and no explanation.
-Participants were asked to choose whether they accept the answer or not, before proceeding to the next question.
-Participants then answered questions on conspiracy thinking, conspiracy beliefs and trust in science.
-
-In study 2, 3 and 4, we presented participants with open-ended questions to explain cases where they rejected the scientific consensus.
-In study 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus.
-Finally, at the end of study 3 and 4, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
+After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.'' Next, participants answered a set of 10 basic science questions (in Study 1 they were randomly selected from a pool of 11 questions, as one `trick' question had been added) in random order.
+After each question, participants were presented with an answer reflecting the scientific consensus, and asked whether they accepted it.
+In Studies 2 and 3, participants additionally saw a short explanation, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
+In Study 4, we provided only two links and no explanation.
+Participants then answered questions on conspiracy thinking, conspiracy beliefs, and trust in science.
+
+In Studies 2, 3, and 4, we presented participants with open-ended questions so they could explain their rejection of the scientific consensus.
+In Studies 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus.
+Finally, at the end of Studies 3 and 4, we asked participants: ``For the questions in which you agreed with the scientific consensus, would you say that\ldots?'' The answer options were: (i) ``You mostly agree with the consensus because, on that question, you trust scientists'', (ii) ``You mostly agree with the consensus because you have been able to independently verify it'', and (iii) ``Other'', with a text box for participants to explain.
 Participants who selected ``You mostly agree with the consensus because you have been able to independently verify it'', were asked the open-ended follow-up question: ``Could you please tell us how you independently verified the information?''.
@@
 
-We recruited a total of 782 (194 in study 1, 190 in study 2, 200 in study 3, 198 in study 4) participants in the US via prolific.
+After removing failed attention checks, we were left with a total sample size of 782 (194 in Study 1, 6 failed attention checks; 190 in Study 2, 11 failed attention checks; 200 in Study 3, no failed attention checks; 198 in Study 4, 2 failed attention checks) participants in the US via prolific.
 Details and demographics can be found in the online supplemental material.
-While samples for study 1 and 2 were convenience samples, study 3 and 4 were run on a sample holding vaccine-skeptic beliefs (see OSM for details).
-
-\subsection{Materials}\label{materials}
-
+While samples for Studies 1 and 2 were convenience samples, Studies 3 and 4 were conducted on a sample holding vaccine-skeptic beliefs.
+Prolific Academic allows to select participants based on their answers to a range of questions.
+We picked three of these questions and only recruited participants who met our criteria for each of them:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  ``Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: {[}For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say''{]}.
+We selected particpants who answered''Against''.
+\item
+  ``Have you received a coronavirus (COVID-19) vaccination? {[}Yes (at least one dose); No; Prefer not to answer{]}''. We select only people who answered ``No''.
+\item
+  ``On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children.
+{[}1 (totally disagree); 2 (disagree); 3 (somewhat disagree); 4 (neither agree nor disagree); 5 (somewhat agree); 6 (agree); 7 (totally agree); rather not say{]}''.
+We select only people who answered `1', `2', or `3'.
+\end{enumerate}
+
+\subsection{Materials}\label{materials-1}
+
+\subsubsection{Scientific facts}\label{scientific-facts}
+
+Studies 1 to 3 used 10 facts drawn from widely used questionnaires about science knowledge (Allum, Sturgis, Tabourazi, \& Brunton-Smith, 2008; Durant, Evans, \& Thomas, 1989; Miller, 1998) sometimes referred to as the ``Oxford scale'' (Gauchat, 2011).
+A `trick' question was added in Study 1 and removed as its wording proved unclear.
+Study 4 used 10 more recent scientific discoveries.
 Table \ref{tab:knowledge} shows all questions and their answer options used in the four studies.
@@
 
-Our main measure of conspiracy belief was asking participants about specific conspiracy theories.
+\subsubsection{Conspiracy beliefs}\label{conspiracy-beliefs}
+
 We selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) (Pennycook, Binnendyk, \& Rand, 2022) (Table \ref{tab:conspiracy}).
@@
 
-As a robustness check, we also used established scales on conspiracy thinking, namely the four-item conspiracy mentality questionnaire (CMQ) (Bruder, Haffke, Neave, Nouripanah, \& Imhoff, 2013) and the single item conspiracy beliefs scale (SICBS) (Lantian, Muller, Nurra, \& Douglas, 2016).
+\subsubsection{Conspiracy thinking}\label{conspiracy-thinking}
+
+We used the four-item conspiracy mentality questionnaire (CMQ) (Bruder, Haffke, Neave, Nouripanah, \& Imhoff, 2013) and the single item conspiracy beliefs scale (SICBS) (Lantian, Muller, Nurra, \& Douglas, 2016).
 Details and comparisons between the scales can be found in the OSM.
@@
 
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\item \textit{Note: } 
+\item Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.
+\end{TableNotes}
+\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{3em}>{\raggedright\arraybackslash}p{40em}}
 \caption{\label{tab:conspiracy}Conspiracy items}\\
@@
 \bottomrule
+\insertTableNotes
 \end{longtable}
+\end{ThreePartTable}
 \endgroup{}
 
-Our main item for measuring trust in science was selected from the Wellcome Global Monitor survey: ``In general, would you say that you trust science a lot, some, not much, or not at all?
+\subsubsection{Trust in science}\label{trust-in-science-1}
+
+In all analyses reported in the main paper, we measure trust in science via a question selected from the Wellcome Global Monitor surveys (Wellcome Global Monitor, 2018, 2020): ``In general, would you say that you trust science a lot, some, not much, or not at all?
 {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''.
-We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey (``How much do you trust scientists in this country?
+We chose this question as it seemed to be the most general one.
+In the appendix, we additionally report results for two alternative measures of trust included in our studies: Another from the WGM surveys (``How much do you trust scientists in this country?
 Do you trust them a lot, some, not much, or not at all?
-{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''), the other from the Pew research center (``How much confidence do you have in scientists to act in the best interests of the public?
-{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}'').
-We selected these items so that we could compare the ratings in our sample to global survey results.
-The WGM survey has been administered in over 140 countries and included over 140000 respondents.
-The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents (Cologna et al., 2024).
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''), and one from the Pew Research Center (``How much confidence do you have in scientists to act in the best interests of the public?
+{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}''; see e.g.
+Funk, Johnson, and Hefferon (2019) and used by e.g.
+Cologna et al.
+(2024)).
+We selected these items so that we could compare the answers in our sample to global survey results.
+We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM).
 
@@
 \includegraphics{output/figures/summary-plot.pdf}
-\caption{\label{fig:summary-plot}Points represent the average share of acceptance as a function of the \textbf{A} level of trust in science (``In general, would you say that you trust science a lot, some, not much, or not at all?
-{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''); \textbf{B} average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); \textbf{C} belief in a specific conspiracy theory.
-Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: ``1 - completly false, 5 - unsure, 9 - completely true''.
-Averages are only based on participants who believed at least to some degree in the conspiracy theory.
-We consider a believer as anyone scoring higher than the scale midpoint, i.e.~\textgreater5.
-Number labels in plots represent the number of participants for the corresponding data point.}
-\end{figure}
-
-Across all studies we find that almost all participants trust basic and recent/advance science facts almost all the time.
-That is true even for participants who expressed low general trust in science, and those who have a tendency to conspiracy think or hold science-related conspiracy beliefs (Fig.
-\ref{fig:summary-plot}).
-
-Our results suggest that the more people tend to trust science, the more they tend to know about science and accept the scientific consensus.
-In study 2 and 3, we find a positive but small correlation between trust in science and both science knowledge (study 1: r = 0.29, p \textless{} .001; study 2: r = 0.28, p \textless{} .001), and acceptance of scientific consensus (study 1: r = 0.27, p \textless{} .001; study 2: r = 0.30, p \textless{} .001).
-In study 3 and 4--both run on vaccine-skeptic samples--we find no statistically significant correlation between science knowledge and trust in science (study 3: r = 0.14, p = 0.100; study 4: r = 0.10, p = 0.144), but a small positive correlation between acceptance of scientific consensus and trust in science (study 3: r = 0.20, p = 0.019; study 4: r = 0.16, p = 0.029).
-
-As for conspiracy beliefs, results are mixed.
-In studies 1 and 2, we find a small negative correlation between conspiracy thinking and both science knowledge (study 1: r = -0.38, p \textless{} .001; study 2: r = -0.40, p \textless{} .001), and acceptance of scientific consensus (study 1: r = -0.33, p \textless{} .001; study 2: r = -0.37, p \textless{} .001).
-However, in study 3 and 4, we do not find a statistically significant correlation with neither science knowledge (study 3: r = -0.16, p = 0.055; study 4: r = -0.02, p = 0.742), and between nor acceptance of scientific consensus (study 3: r = -0.02, p = 0.788; study 4: r = -0.05, p = 0.498).
-
-Overall, the correlations we find are relatively weak, which might be partly due to ceiling effects.
-This was the case in particular for acceptance: In all studies, in the vast majority of cases, participants accepted accepted science facts.
-The absence of correlations between conspiracy beliefs and science knowledge or acceptance in studies 3 and 4 might be explained by our sample selection.
-For these studies, we recruited samples scoring considerably higher in conspiracy belief than the convenience samples of study 1 and 2 (see OSM).
-The presence of effects in study 1 and 2, and the absence in study 3 and 4 might be because differences between moderate and high conspiracy thinking might not matter, while, differences between low and moderate/high conspiracy thinking do matter.
-
+\caption{\label{fig:summary-plot}Points represent the average share of acceptance and numbers the absolute count of participants as a function of: \textbf{A} the level of trust in science (``In general, would you say that you trust science a lot, some, not much, or not at all?
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''); \textbf{B} the average conspiracy thinking (CMQ, five items on a scale from 0 to 100); \textbf{C} the belief in specific conspiracy theories (i.e.~participants who answered 9, ``completely true'', for a given theory, see Table \ref{tab:conspiracy} for the list of the theories).}
+\end{figure}
+
+The main outcome of interest is acceptance of the scientifically consensual facts presented.
+Overall, acceptance was very high (aggregating across all studies: 95.1 \%; Study 1: 93 \%; Study 2: 98 \%; Study 3: 98 \%; Study 4: 91 \%).
+Note that this includes both participants who had previously correctly answered the knowledge question, and participants who changed their mind when presented with the scientific consensus.
+In Studies 3 \& 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: 96 \%; in Study 4: 86 \%).
+
+As shown in Figure \ref{fig:summary-plot}, these very high rates of acceptance hold for: participants who they do not trust science at all (5.7 \% of participants, acceptance rate of 87.3 \%), participants who rank in the top two deciles of the conspiracy thinking scale (28.8 \% of participants, average acceptance rate of 93.2 \%), participants who consider as ``completely true'' (the maximum of the 9-point scale) conspiracy theories stating that the earth is flat (5.8 \% of participants, acceptance rate of 87.2 \%), or that climate change due to fossil emissions is a hoax (12 \% of participants, acceptance rate of 91.7 \%).
+This doesn't merely reflect science knowledge: aggregated across all studies--participants only correctly answered 65.8 \% of the questions (Study 1: 74 \%; Study 2: 79 \%; Study 3: 75 \%; Study 4: 36 \%) before they were provided with the scientifically consensual answer.
 Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
@@
 In most cases (study 1: 76.3 \%; study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-Only in study 4 do we find evidence\footnote{For this analysis, we restricted our data to those cases in which participants initially gave the wrong answer.} that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = 0.06, p = 0.387; study 2: r = 0.16, p = 0.051; study 3: r = 0.04, p = 0.619; study 4: r = 0.15, p = 0.037) and only in study 2 evidence that it is associated with (less) conspiracy thinking (study 1: r = -0.14, p = 0.061; study 2: r = -0.22, p = 0.006; study 3: r = -0.04, p = 0.631; study 4: r = -0.05, p = 0.455).
 In very few cases, participants changed their mind away from the scientific consensus.
 These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: 1.6 \%; study 2: 0.5 \%; study 3: 0.7 \%; study 4: 5.7 \%).
+
+\begin{verbatim}
+## # A tibble: 1 x 2
+##   mean_acceptance mean_knowledge
+##             <dbl>          <dbl>
+## 1           0.805          0.416
+\end{verbatim}
+
+Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = 0.27, p \textless{} .001; 2: r = 0.30, p \textless{} .001; 3: r = 0.20, p = 0.019; 4: r = 0.16, p = 0.029), but a less consistent relation between trust in science and science knowledge (Studies 1: r = 0.29, p \textless{} .001; 2: r = 0.28, p \textless{} .001; 3: r = 0.14, p = 0.100; study 4: r = 0.10, p = 0.144).
+
+Regarding H2, the results are mixed both for the relation between conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) and acceptance of the scientific consensus (Studies 1: r = -0.33, p \textless{} .001; 2: r = -0.37, p \textless{} .001; 3: r = -0.02, p = 0.788; 4: r = -0.05, p = 0.498) as well as science knowledge (Studies 1: r = -0.38, p \textless{} .001; study 2: r = -0.40, p \textless{} .001; 3: r = -0.16, p = 0.055; 4: r = -0.02, p = 0.742).
 
@@
 \midrule
-NA & 255.00 & 70.1\% & 95.00\\
-Personal convictions & 37.00 & 10.2\% & 27.00\\
-Mistake & 30.00 & 8.2\% & 26.00\\
-No justification & 25.00 & 6.9\% & 18.00\\
-Not convinced & 12.00 & 3.3\% & 8.00\\
-Religious Beliefs & 5.00 & 1.4\% & 5.00\\
+Not convinced & 158.00 & 43.4\% & 70.00\\
+No justification & 103.00 & 28.3\% & 51.00\\
+Personal convictions & 43.00 & 11.8\% & 31.00\\
+Mistake & 42.00 & 11.5\% & 34.00\\
+Religious Beliefs & 18.00 & 4.9\% & 13.00\\
 \bottomrule
@@
 Why did participants accept the scientific consensus?
-In studies 3 and 4, we had asked participants about cases where they agreed with the scientific consensus.
-A total of 320 (study 3: 122; study 4: 198) participants answered this question.
-Of these, 41.8\% in study 3 and 36.4\% in study 4 said they accepted the scientific consensus because they trust scientists, while 47.5\% in study 3 and 47\% in study 4 said they independently verified the fact.
-We asked all participants who gave this answer to explain how they did it (answers can be found in the OSM).
-10.7\% in study 3 and 16.7\% in study 4 answered with other ``other'' and gave an open-ended explanation (see OSM).
+In Studies 3 and 4--the vaccine skeptic samples--we had asked participants about cases where they agreed with the scientific consensus.
+A total of 320 (Study 3: 122; Study 4: 198) participants answered this question.
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: \texttt{exp3\_descriptives\$acceptance\_by\_reason{[}{[}"independent\ verification"{]}{]}\$share}; Study 4: 47\% ), than there were participants saying it was because they trust scientists (Study 3: 41.8\%; Study 4: 36.4\%).
+We asked participants who said they verified independently to explain how they did it (answers can be found in the ESM).
+10.7\% in study 3 and 16.7\% in study 4 answered with other ``other'' and gave an open-ended explanation (see ESM).
+
+In an exploratory analysis, we ran linear regressions to test whether there are differences between these two groups.
+Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = NA; \(\hat{\beta}_{\text{Trust}}\) = 0.19, = 0.330 on a scale from 1 to 4; Study 4: mean = 2.92; \(\hat{\beta}_{\text{Trust}}\) = 0.46, \textless{} .001) than those who said they verified independently (Study 3 mean = NA; Study 4 mean = 2.45).
+We did not find a difference regarding acceptance (Study 3: \(\hat{\beta}_{\text{Acceptance}}\) = 0.01, = 0.523 on a scale from 0 to 1; Study 4: \(\hat{\beta}_{\text{Acceptance}}\) = 0.04, = 0.116).
+In study 4, where we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+On average, participants clicked only clicked on 1.36 links (out of 20 possible clicks) and there was no difference between the two groups (\(\hat{\beta}_{\text{Clicks}}\) = -1.09, = 0.062).
 
@@
 For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second.
-It is therefor easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
+It is therefore easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
 In other words, people might hold on to most scientific findings as reflective beliefs---beliefs that have little effect on our behaviors (Mercier, 2021).
@@
 Durant, J. R., Evans, G. A., \& Thomas, G. P. (1989). The public understanding of science. \emph{Nature}, \emph{340}(6228), 11--14. \url{https://doi.org/10.1038/340011a0}
+
+\bibitem[\citeproctext]{ref-funkKeyFindingsPublic2019}
+Funk, C., Johnson, C., \& Hefferon, M.
+(2019).
+\emph{5 key findings about public trust in scientists in the u.s.} Retrieved from \url{https://www.pewresearch.org/fact-tank/2019/08/05/5-key-findings-about-public-trust-in-scientists-in-the-u-s/}
 
@@
 
+\bibitem[\citeproctext]{ref-vanstekelenburgScientificConsensusCommunicationContested2022}
+Van Stekelenburg, A., Schaap, G., Veling, H., Van 'T Riet, J., \& Buijzen, M.
+(2022).
+Scientific-Consensus Communication About Contested Science: A Preregistered Meta-Analysis.
+\emph{Psychological Science}, \emph{33}(12), 1989--2008.
+\url{https://doi.org/10.1177/09567976221083219}
+
+\bibitem[\citeproctext]{ref-veckalov27countryTestCommunicating2024}
+Većkalov, B., Geiger, S.
+J., Bartoš, F., White, M.
+P., Rutjens, B.
+T., Harreveld, F.
+van, \ldots{} Linden, S.
+van der.
+(2024).
+A 27-country test of communicating the scientific consensus on climate change.
+\emph{Nature Human Behaviour}, 1--14.
+\url{https://doi.org/10.1038/s41562-024-01928-2}
+
 \bibitem[\citeproctext]{ref-vranicDidMyOwn2022}
@@
 
+\bibitem[\citeproctext]{ref-wellcomeglobalmonitorPublicTrustScientists2021}
+Wellcome Global Monitor.
+(2021).
+\emph{Public trust in scientists rose during the Covid-19 pandemic \textbar{} News}.
+Retrieved from \url{https://wellcome.org/news/public-trust-scientists-rose-during-covid-19-pandemic}
+
 \end{CSLReferences}
@@
 
-\subsection{Materials}\label{materials-1}
+\subsection{Materials}\label{materials-2}
 
@@
 
-\subsection{Conspiracy thinking}\label{conspiracy-thinking}
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-1}
 
@@
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:unnamed-chunk-22}Reasons for consensus rejection}\\
+\caption{\label{tab:unnamed-chunk-25}Reasons for consensus rejection}\\
 \toprule
@@
 
-\subsection{Materials}\label{materials-2}
+\subsection{Materials}\label{materials-3}
 
@@
 
-\begin{tabular}{lllllllll}
-\toprule
-std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
-\midrule
-0.07 & -6.01 & < .001 & -0.53 & -0.27 & BCTI 
-(main conspiracy measure) & {}[-0.533, -0.27] & -0.40 & NA\\
-0.07 & -2.36 & = 0.019 & -0.31 & -0.03 & CMQ & {}[-0.311, -0.028] & -0.17 & NA\\
-0.07 & -1.91 & = 0.058 & -0.28 & 0.00 & SICBS & {}[-0.28, 0.005] & -0.14 & NA\\
-0.07 & 3.20 & = 0.002 & 0.09 & 0.37 & WGM trust scientists & {}[0.087, 0.368] & 0.23 & NA\\
-0.07 & 3.98 & < .001 & 0.14 & 0.42 & WGM trust general 
-(main trust measure) & {}[0.14, 0.417] & 0.28 & NA\\
-0.07 & 2.23 & = 0.027 & 0.02 & 0.30 & PEW trust scientists & {}[0.019, 0.303] & 0.16 & NA\\
-0.07 & -5.47 & < .001 & -0.50 & -0.24 & BCTI 
-(main conspiracy measure) & {}[-0.504, -0.237] & NA & -0.37\\
-0.07 & -3.03 & = 0.003 & -0.36 & -0.08 & CMQ & {}[-0.356, -0.076] & NA & -0.22\\
-0.07 & -2.76 & = 0.006 & -0.34 & -0.06 & SICBS & {}[-0.338, -0.056] & NA & -0.20\\
-0.07 & 3.51 & < .001 & 0.11 & 0.39 & WGM trust scientists & {}[0.108, 0.387] & NA & 0.25\\
-0.07 & 4.29 & < .001 & 0.16 & 0.44 & WGM trust general 
-(main trust measure) & {}[0.161, 0.436] & NA & 0.30\\
-0.07 & 3.04 & = 0.003 & 0.08 & 0.36 & PEW trust scientists & {}[0.076, 0.357] & NA & 0.22\\
+\begin{tabular}{lll}
+\toprule
+outcome & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+BCTI 
+(main conspiracy measure) & -0.4 (p< .001) & -0.37 (p< .001)\\
+CMQ & -0.17 (p= 0.019) & -0.22 (p= 0.003)\\
+SICBS & -0.14 (p= 0.058) & -0.2 (p= 0.006)\\
+WGM trust scientists & 0.23 (p= 0.002) & 0.25 (p< .001)\\
+WGM trust general 
+(main trust measure) & 0.28 (p< .001) & 0.3 (p< .001)\\
+PEW trust scientists & 0.16 (p= 0.027) & 0.22 (p= 0.003)\\
 \bottomrule
@@
 
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-1}
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-2}
 
@@
 
-We recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of 200 participants (125 female, 73 male; \(age_\text{mean}\): 42.90, \(age_\text{sd}\): 12.00, \(age_\text{median}\): 41).
+We recruited a sample of participants holing vaccine-skeptic beliefs.
+Prolific Academic allows to filter out participants who have given specific answers to certain questions.
+We selected three available questions: 1.
+``Participants were asked the following question: Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: {[}For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say''{]}.
+We selected particpants who answered''Against''.
+2.
+``Participants were asked the following question: Have you received a coronavirus (COVID-19) vaccination?
+{[}Yes (at least one dose); No; Prefer not to answer{]}''.
+We select only people who answered ``No''.
+3.
+``Participants were asked the following question: On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children.
+{[}1 (TOTALLY DISAGREEE); 2 (DISAGREE); 3 (SOMEWHAT DISAGREE); 4 (NEITHER AGREE NOR DISAGREE); 5 (SOMEWHAT AGREE); 6 (AGREE); 7 (TOTALLY AGREE); Rather not say''{]}.
+We select only people who answered''1'', ``2'', or ``3''.
+
+Based on these criteria, we recruited 200 participants from the US via prolific, of which none failed our attention check, resulting in a final sample of 200 participants (125 female, 73 male; \(age_\text{mean}\): 42.90, \(age_\text{sd}\): 12.00, \(age_\text{median}\): 41).
 Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
@@
 
-\subsection{Materials}\label{materials-3}
+\subsection{Materials}\label{materials-4}
 
@@
 \midrule
-wgm\_sciencegeneral & 1.00 & NA & NA\\
-wgm\_scientists & NA & 1.00 & NA\\
-pew & NA & NA & 1.00\\
+wgm\_sciencegeneral & 1.00 & 0.77 & 0.68\\
+wgm\_scientists & 0.77 & 1.00 & 0.68\\
+pew & 0.68 & 0.68 & 1.00\\
 \bottomrule
@@
 
-\begin{tabular}{lllllllll}
-\toprule
-std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
-\midrule
-0.08 & -1.93 & = 0.055 & -0.32 & 0.00 & BCTI 
-(main conspiracy measure) & {}[-0.317, 0.004] & -0.16 & NA\\
-0.08 & 0.10 & = 0.925 & -0.15 & 0.17 & CMQ & {}[-0.154, 0.17] & 0.01 & NA\\
-0.08 & 0.30 & = 0.767 & -0.14 & 0.19 & SICBS & {}[-0.138, 0.186] & 0.02 & NA\\
-0.08 & 1.31 & = 0.193 & -0.06 & 0.28 & WGM trust scientists & {}[-0.056, 0.276] & 0.11 & NA\\
-0.08 & 1.66 & = 0.100 & -0.03 & 0.30 & WGM trust general 
-(main trust measure) & {}[-0.027, 0.304] & 0.14 & NA\\
-0.08 & -0.91 & = 0.367 & -0.24 & 0.09 & PEW trust scientists & {}[-0.243, 0.09] & -0.08 & NA\\
-0.11 & -0.27 & = 0.788 & -0.25 & 0.19 & BCTI 
-(main conspiracy measure) & {}[-0.246, 0.187] & NA & -0.03\\
-0.11 & -0.08 & = 0.934 & -0.23 & 0.21 & CMQ & {}[-0.226, 0.208] & NA & -0.01\\
-0.11 & 0.02 & = 0.987 & -0.22 & 0.22 & SICBS & {}[-0.215, 0.218] & NA & 0.00\\
-0.07 & 2.72 & = 0.007 & 0.05 & 0.34 & WGM trust scientists & {}[0.054, 0.343] & NA & 0.20\\
-0.07 & 2.37 & = 0.019 & 0.03 & 0.32 & WGM trust general 
-(main trust measure) & {}[0.029, 0.32] & NA & 0.17\\
-0.07 & 1.64 & = 0.104 & -0.03 & 0.27 & PEW trust scientists & {}[-0.025, 0.268] & NA & 0.12\\
+\begin{tabular}{lll}
+\toprule
+outcome & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+BCTI 
+(main conspiracy measure) & -0.16 (p= 0.055) & -0.03 (p= 0.788)\\
+CMQ & 0.01 (p= 0.925) & -0.01 (p= 0.934)\\
+SICBS & 0.02 (p= 0.767) & 0 (p= 0.987)\\
+WGM trust scientists & 0.11 (p= 0.193) & 0.2 (p= 0.007)\\
+WGM trust general 
+(main trust measure) & 0.14 (p= 0.100) & 0.17 (p= 0.019)\\
+PEW trust scientists & -0.08 (p= 0.367) & 0.12 (p= 0.104)\\
 \bottomrule
@@
 
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-2}
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-3}
 
@@
 
-We recruited 200 participants from the US via prolific, of which two failed our attention check, resulting in a final sample of 198 participants (98 female, 100 male; \(age_\text{mean}\): 42.67, \(age_\text{sd}\): 12.32, \(age_\text{median}\): 41).
+We recruited a sample of participants holing vaccine-skeptic beliefs.
+Prolific Academic allows to filter out participants who have given specific answers to certain questions.
+We selected three available questions: 1.
+``Participants were asked the following question: Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: {[}For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say''{]}.
+We selected particpants who answered''Against''.
+2.
+``Participants were asked the following question: Have you received a coronavirus (COVID-19) vaccination?
+{[}Yes (at least one dose); No; Prefer not to answer{]}''.
+We select only people who answered ``No''.
+3.
+``Participants were asked the following question: On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children.
+{[}1 (TOTALLY DISAGREEE); 2 (DISAGREE); 3 (SOMEWHAT DISAGREE); 4 (NEITHER AGREE NOR DISAGREE); 5 (SOMEWHAT AGREE); 6 (AGREE); 7 (TOTALLY AGREE); Rather not say''{]}.
+We select only people who answered''1'', ``2'', or ``3''.
+
+Based on these criteria, we recruited 200 participants from the US via prolific, of which two failed our attention check, resulting in a final sample of 198 participants (98 female, 100 male; \(age_\text{mean}\): 42.67, \(age_\text{sd}\): 12.32, \(age_\text{median}\): 41).
 Since we did not have any prior assumptions on effect sizes and our analyses were descriptive, we did not do a power analysis.
@@
 
-\subsection{Materials}\label{materials-4}
+\subsection{Materials}\label{materials-5}
 
@@
 \midrule
-NA & 255.00 & 100\% & 95.00\\
+Not convinced & 146.00 & 57.3\% & 65.00\\
+No justification & 78.00 & 30.6\% & 35.00\\
+Religious Beliefs & 13.00 & 5.1\% & 8.00\\
+Mistake & 12.00 & 4.7\% & 10.00\\
+Personal convictions & 6.00 & 2.4\% & 4.00\\
 \bottomrule
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-8}
+\subsubsection{Trust in science}\label{trust-in-science-9}
 
@@
 
-\begin{tabular}{lllllllll}
-\toprule
-std.error & \multicolumn{1}{c}{statistic} & \multicolumn{1}{c}{p.value} & \multicolumn{1}{c}{conf.low} & \multicolumn{1}{c}{conf.high} & \multicolumn{1}{c}{outcome} & \multicolumn{1}{c}{ci} & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
-\midrule
-0.07 & -0.33 & = 0.742 & -0.16 & 0.12 & BCTI 
-(main conspiracy measure) & {}[-0.164, 0.117] & -0.02 & NA\\
-0.07 & 1.60 & = 0.112 & -0.03 & 0.25 & CMQ & {}[-0.027, 0.253] & 0.11 & NA\\
-0.07 & 0.55 & = 0.586 & -0.10 & 0.18 & SICBS & {}[-0.102, 0.18] & 0.04 & NA\\
-0.07 & 2.01 & = 0.045 & 0.00 & 0.28 & WGM trust scientists & {}[0.003, 0.282] & 0.14 & NA\\
-0.07 & 1.47 & = 0.144 & -0.04 & 0.24 & WGM trust general 
-(main trust measure) & {}[-0.036, 0.244] & 0.10 & NA\\
-0.07 & 2.04 & = 0.043 & 0.00 & 0.28 & PEW trust scientists & {}[0.005, 0.284] & 0.14 & NA\\
-0.07 & -0.68 & = 0.498 & -0.19 & 0.09 & BCTI 
-(main conspiracy measure) & {}[-0.189, 0.092] & NA & -0.05\\
-0.07 & -1.42 & = 0.156 & -0.24 & 0.04 & CMQ & {}[-0.241, 0.039] & NA & -0.10\\
-0.07 & -1.92 & = 0.056 & -0.28 & 0.00 & SICBS & {}[-0.276, 0.004] & NA & -0.14\\
-0.07 & 3.74 & < .001 & 0.12 & 0.39 & WGM trust scientists & {}[0.122, 0.394] & NA & 0.26\\
-0.07 & 2.20 & = 0.029 & 0.02 & 0.30 & WGM trust general 
-(main trust measure) & {}[0.016, 0.295] & NA & 0.16\\
-0.07 & 3.00 & = 0.003 & 0.07 & 0.35 & PEW trust scientists & {}[0.072, 0.348] & NA & 0.21\\
+\begin{tabular}{lll}
+\toprule
+outcome & \multicolumn{1}{c}{Correlation with knowledge} & \multicolumn{1}{c}{Correlation with acceptance}\\
+\midrule
+BCTI 
+(main conspiracy measure) & -0.02 (p= 0.742) & -0.05 (p= 0.498)\\
+CMQ & 0.11 (p= 0.112) & -0.1 (p= 0.156)\\
+SICBS & 0.04 (p= 0.586) & -0.14 (p= 0.056)\\
+WGM trust scientists & 0.14 (p= 0.045) & 0.26 (p< .001)\\
+WGM trust general 
+(main trust measure) & 0.1 (p= 0.144) & 0.16 (p= 0.029)\\
+PEW trust scientists & 0.14 (p= 0.043) & 0.21 (p= 0.003)\\
 \bottomrule
@@
 
-\subsection{Trust in science}\label{trust-in-science-9}
+\subsection{Trust in science}\label{trust-in-science-10}
 
@@
 
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-3}
+\subsection{Conspiracy thinking}\label{conspiracy-thinking-4}
 

### 2024-09-12 appendix_acceptance_knowledge.Rmd
first version within word limit of 4k

--- 2024-09-02 paper.Rmd
@@
----
-title             : "Universal acceptance of (nearly all of) basic science in the US"
-shorttitle        : "Acceptance of basic science"
-
-header-includes:  | # to prevent floats from moving past certain points (for the appendix)
-  \usepackage{placeins}
-
-author: 
-  - name          : ""
-    affiliation   : ""
-
-affiliation:
-  - id            : ""
-    institution   : ""
-
-abstract: |
-  XX
-
-keywords          :
-
-floatsintext      : yes
-linenumbers       : no 
-draft             : no
-mask              : no
-
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-
-classoption       : "doc" # "doc" for nice look, "man" for manuscripty
-output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
-For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g.
-set always allow html as true in the yaml heading
-
-always_allow_html: true
-
-appendix:
-  - "appendix_exp1.Rmd"
-  - "appendix_exp2.Rmd"
-  - "appendix_exp3.Rmd"
-  - "appendix_exp4.Rmd"
-
-bibliography: references.bib
----
-
-```{r setup, include=FALSE}
-# Figure out output format
-is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
-is_latex <- knitr::pandoc_to("latex")
-is_html <- knitr::pandoc_to("html")
-
-# Word-specific things
-table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
-conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
-conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align
-
-# Knitr options
-knitr::opts_chunk$set(
-  echo = FALSE, warning = FALSE, message = FALSE,
-  # tidy.opts = list(width.cutoff = 120),  # Code width
-  # fig.retina = 3, dpi = conditional_dpi,
-  # fig.width = 7, fig.asp = 0.618,
-  # fig.align = conditional_align, out.width = "100%",
-  fig.path = "output/figures/",
-  cache.path = "output/_cache/",
-  fig.process = function(x) {  # Remove "-1" from figure names
-    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
-    if (file.rename(x, x2)) x2 else x
-  },
-  options(scipen = 99999999)  # Prevent scientific notation
-)
-
-# R options
-options(
-  width = 90,  # Output width
-  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
-  knitr.kable.NA = "",  # Make NAs blank in kables
-  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
-  modelsummary_factory_default = table_format,  # Set modelsummary backend
-  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
-)
-```
-
-```{r packages, include=FALSE}
-# load required packages
-library("papaja")      # For APA style manuscript   
-library("lme4")        # model specification / estimation
-library("lmerTest")    # provides p-values in the output
-library("tidyverse")   # data wrangling and visualisation
-library("afex")        # anova and deriving p-values from lmer
-library("broom")       # extracting data from model fits 
-library("broom.mixed") # extracting data from mixed models
-library("metafor")     # doing mata analysis
-library("patchwork")   # put several plots together
-library("ggridges")    # for plots
-library("gghalves")    # for plots
-library("ggbeeswarm")  # Special distribution-shaped point jittering
-library("knitr")       # for tables
-library("kableExtra")  # also for tables
-library("ggpubr")      # for combining plots with ggarrange() 
-library("grid")        # for image plots   
-library("gridExtra")   # for image plots
-library("png")         # for image plots
-library("modelsummary") # for regression tables
-library("ggrepel") # for better random patterns in plots
-```
-
-```{r functions}
-# load plot theme
-source("functions/plot_theme.R")
-
-# load other functions
-source("functions/own_functions.R")
-```
-
-# Introduction
-
-How much do people trust and accept science?
-The recent COVID pandemic has highlighted the importance of this question, with trust in science being the best predictor of people’s proclivity to follow health guidelines [@alganTrustScientistsTimes2021].
-
-Unfortunately, trust in science is far from being at ceiling [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020], and there are suggestions that it has recently been dropping in some countries [@alganTrustScientistsTimes2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly politicized [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
-
-This relative lack of trust in science could mean different things.
-It could mean that some people reject science wholesale, or at least to a significant extent.
-Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (e.g.
-creationism), or beliefs that are associated with disliked behaviors (e.g.
-vaccination, acting for climate change).
-
-We answer this question by looking at how much Americans accept basic scientific knowledge.
-We pay special attention to two groups of participants: people who say they don’t trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
-
-We briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
-
-## Trust in science
-
-Across the globe, most people say they trust science at least to some extent.
-In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018].
-In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants [@wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-Across all countries, in 2018, 32% of participants said they trust science "a lot" (41% in 2020), 45% trust "some" science (39% in 2020), and only 13% percent trust science "not much/ not at all" (13% in 2020), while 10% indicated "don't know" (7% in 2020)[^1].
-
-[^1]: <https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020#gid=6acd&pid=0>\>
-
-Trust in science also appears to be relatively stable in the US [@funkPublicConfidenceScientists2020].
-However, there are some potentially worrying trends: Recently, trust in US has dropped [@brianAmericansTrustScientists2023], and some evidence suggests increasing polarization of trust in science along political lines [@gauchatPoliticizationSciencePublic2012].
-For example, the partisan divide on the question of whether climate change is a major threat to society has been growing in the last decade[^2] [@kennedyWhatDataSays2023].
-A recent study suggests that by some measure, only about half of the US population believes in anthropogenic climate change, and that individuals do mostly not seem to change their mind on on the issue [@mottaChangingMindsChanging2021].
-
-[^2]: <https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/>
-
-Trust in science is typically measured in three ways: either asking explicitly about general trust in science or scientists, by asking about science attitudes (e.g.
-support for public funding of science), or asking questions about specific, typically contentious, science topics (e.g.
-vaccines, climate change, GMOs).
-It is unclear to which extent these measures capture trust in basic science.
-To assess this, we can look at the literature on science knowledge.
-
-## Science knowledge
-
-Much of the research on public understanding of science measures to which extent people have basic science knowledge (e.g.
-"Does the Earth go around the Sun, or does the Sun go around the Earth?").
-For us, these surveys provide a lower bound to how much people trust basic science: people who give the right answer (and not by chance) supposedly trust science on that question.
-Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge [@millerPublicUnderstandingAttitudes2004; @millerMeasurementCivicScientific1998a].
-For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016].
-A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.
-
-[To do: Make a descriptive overview of science knowledge across time/countries.]
-
-## CT and TiS
-
-Recently, the relationship of one group to trust in science (TiS) has been particularly scrutinized: conspiracy theorists (CTists), either defined as people who endorse specific conspiracy theories, or as people who have a conspiracy mindset [@rutjensConspiracyBeliefsScience2022; @vranicDidMyOwn2022].
-CTists reject specific science knowledge (related to their CT), and they tend to trust science less on the whole.
-What remains unclear is to which extent CTists (dis)trust basic science.
-
-## The present studies
-
-In a series of four pre-registered studies (n = 782), we asked participants about basic (studies 1, 2 and 3) and recent/advanced (study 4) science knowledge questions, informed them of the scientific consensus, and aked if they accept it.
-Of particular interest were people who are low in TiS, and CTists.
-Our main hypotheses were:
-
-**H1: Higher trust in science is associated with more science knowledge and more acceptance of the scientific consensus**
-
-**H2: Higher conspiracy thinking/belief is associated with less science knowledge and less acceptance of the scientific consensus**
-
-In study 1, we plainly inform participants of the scientific consensus and ask if they accept it.
-We use a subset of questions on basic science knowledge that have been used in numerous public opinion surveys [@allumScienceKnowledgeAttitudes2008; @durantPublicUnderstandingScience1989; @millerMeasurementCivicScientific1998a] sometimes referred to as the "Oxford scale" [@gauchatCulturalAuthorityScience2011].
-In study 2, we remove a problematic science question and present people with a more elaborated explanation of the scientific consensus, as well as additional sources.
-This allows us to remove the issue that participants might simply not have trusted us to correctly report the scientific consensus in study 1.
-Study 3 follows essentially the same design as study 2, but is run on a vaccine-skeptic sample.
-Further, we provided participants with an explicit option to revise their answer, in case they rejected the consensus.
-We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently.
-Study 4 uses the same design as study 3, but using a set of more recent, advance science questions that participants are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves.
-Just as study 3, study 4 is run on a sample of vaccine-skeptic participants.
-
-# Methods
-
-```{r exp1}
-# Analyze data of experiments and store results
-
-# Experiment 1
-
-# read the cleaned long version data set
-exp1_long <- read_csv("exp_1/data/cleaned_long.csv")
-
-# read wide version data set
-exp1_wide <- read_csv("exp_1/data/cleaned_wide.csv")
-
-# Research questions
-
-# RQ3
-cor_trust_knowledge <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_knowledge) %>% text_ready()
-
-cor_trust_acceptance <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_acceptance) %>% text_ready()
-
-# RQ4
-cor_conspiracy_knowledge <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_knowledge) %>% text_ready()
-
-cor_conspiracy_acceptance <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_acceptance)%>% text_ready()
-
-# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
-exp1_false_knowledge <- exp1_long %>% 
-  # make numeric versions
-  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
-  ) %>% 
-  group_by(id, knowledge) %>% 
-  # calculate by-participant averages
-  summarize(
-    n = n(),
-    n_accepted = sum(acceptance_num),
-    avg_acceptance  = sum(acceptance_num)/n(), 
-    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
-    BCTI_avg = mean(BCTI_avg), 
-    # add additonal measures for robustness checks
-    CMQ_avg = mean(CMQ_avg), 
-    SICBS = mean(SICBS),
-    wgm_scientists = mean(wgm_scientists), 
-    pew = mean(pew)
-  ) %>% 
-  # filter to only false responses
-  filter(knowledge == FALSE) %>% 
-  ungroup()
-
-false_answers_cor_trust_acceptance <- cor.test(exp1_false_knowledge$wgm_sciencegeneral, exp1_false_knowledge$avg_acceptance) %>% text_ready()
-
-false_answers_cor_conspiracy_acceptance <- cor.test(exp1_false_knowledge$BCTI_avg, exp1_false_knowledge$avg_acceptance) %>% text_ready()
-
-# extract descriptives for inline reporting
-exp1_descriptives <- list(
-  # Demographics
-  n_subj = n_distinct(exp1_wide$id),
-  gender = exp1_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
-  age = exp1_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
-                                       .names = "{.fn}")) %>% rounded_numbers(),
-  # RQ1 & RQ2
-  means = exp1_wide %>% 
-    summarize(across(c(avg_knowledge, avg_acceptance), 
-                     list(mean = mean, sd = sd), 
-                     .names = "{.col}_{.fn}")) %>%
-    rename_with(~sub("^avg_", "", .), everything()) %>% 
-    mutate_if(is.numeric, round, digits = 2) %>% 
-  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
-  # RQ3
-  cor_trust_knowledge = cor_trust_knowledge,
-  cor_trust_acceptance = cor_trust_acceptance,
-  # RQ4
-  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
-  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
-  # Exploratory
-  # acceptance and knowledge
-  conditional_acceptance = exp1_long %>% 
-    group_by(knowledge, acceptance) %>% 
-    count() %>% 
-    group_by(knowledge) %>% 
-    mutate (share = n/sum(n), 
-            # rename knowledge values
-            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
-    mutate_if(is.numeric, round, digits = 3) %>% 
-    mutate (share = paste0(share*100, " %")) %>% 
-    super_split(knowledge, acceptance),
-  # acceptance and trust in science/conspiracy thinking
-  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
-  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance
-)
-```
-
-```{r exp2}
-# Analyze data of experiments and store results
-
-# Experiment 2
-
-# read the cleaned long version data set
-exp2_long <- read_csv("exp_2/data/cleaned_long.csv")
-
-# read wide version data set
-exp2_wide <- read_csv("exp_2/data/cleaned_wide.csv")
-
-# read coded justifications
-exp2_justifications <- read_csv("exp_2/data/justifications_clean.csv")
-
-# Research questions
-
-# H1a
-cor_trust_knowledge <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_knowledge) %>% text_ready()
-
-# H2a
-cor_conspiracy_knowledge <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_knowledge) %>% text_ready()
-
-# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
-exp2_false_knowledge <- exp2_long %>% 
-  # make numeric versions
-  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
-  ) %>% 
-  group_by(id, knowledge) %>% 
-  # calculate by-participant averages
-  summarize(
-    n = n(),
-    n_accepted = sum(acceptance_num),
-    avg_acceptance  = sum(acceptance_num)/n(), 
-    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
-    BCTI_avg = mean(BCTI_avg), 
-    # add additonal measures for robustness checks
-    CMQ_avg = mean(CMQ_avg), 
-    SICBS = mean(SICBS),
-    wgm_scientists = mean(wgm_scientists), 
-    pew = mean(pew)
-  ) %>% 
-  # filter to only false responses
-  filter(knowledge == FALSE) %>% 
-  ungroup()
-
-# H1b
-# conditional on false answers
-false_answers_cor_trust_acceptance <- cor.test(exp2_false_knowledge$wgm_sciencegeneral, exp2_false_knowledge$avg_acceptance) %>% text_ready()
-# pooled across all cases (in line with all other studies' analysis)
-cor_trust_acceptance <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_acceptance) %>% text_ready()
-
-# H2b
-# conditional on false answers
-false_answers_cor_conspiracy_acceptance <- cor.test(exp2_false_knowledge$BCTI_avg, exp2_false_knowledge$avg_acceptance) %>% text_ready()
-# pooled across all cases (in line with all other studies' analysis)
-cor_conspiracy_acceptance <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_acceptance) %>% text_ready()
-
-# extract descriptives for inline reporting
-exp2_descriptives <- list(
-  # Demographics
-  n_subj = n_distinct(exp2_wide$id),
-  gender = exp2_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
-  age = exp2_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
-                                       .names = "{.fn}")) %>% rounded_numbers(),
-  # RQ1 & RQ2
-  means = exp2_wide %>% 
-    summarize(across(c(avg_knowledge, avg_acceptance), 
-                     list(mean = mean, sd = sd), 
-                     .names = "{.col}_{.fn}")) %>%
-    rename_with(~sub("^avg_", "", .), everything()) %>% 
-    mutate_if(is.numeric, round, digits = 2) %>% 
-  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
-  # H1
-  # a
-  cor_trust_knowledge = cor_trust_knowledge,
-  # b (conditional on false responses)
-  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
-  # b (all cases)
-  cor_trust_acceptance = cor_trust_acceptance,
-  # H2
-  # a
-  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
-  # b (conditional on false responses)
-  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
-  # b(all cases)
-  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
-  # Exploratory
-  # acceptance and knowledge
-  conditional_acceptance = exp2_long %>% 
-    group_by(knowledge, acceptance) %>% 
-    count() %>% 
-    group_by(knowledge) %>% 
-    mutate (share = n/sum(n), 
-            # rename knowledge values
-            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
-    mutate_if(is.numeric, round, digits = 3) %>% 
-    mutate (share = paste0(share*100, " %")) %>% 
-    super_split(knowledge, acceptance),
-  # Justifications
-  justifications_n = nrow(exp2_justifications), 
-  justifications_n_participants = n_distinct(exp2_justifications$id),
-  justification_by_category = exp2_justifications %>%
-    group_by(category)%>%
-    summarize(n = n(), 
-              n_subjects = n_distinct(id)) %>%
-    mutate(share = n/sum(n)) %>%
-    mutate_if(is.numeric, round, digits=3) %>% 
-    mutate(share = paste0(share*100, "%")) %>% 
-    split(.$category)
-)
-```
-
-```{r exp3}
-# Analyze data of experiments and store results
-
-# Experiment 3
-
-# read the cleaned long version data set
-exp3_long <- read_csv("exp_3/data/cleaned_long.csv")
-
-# read wide version data set
-exp3_wide <- read_csv("exp_3/data/cleaned_wide.csv")
-
-# read coded justifications
-exp3_justifications <- read_csv("exp_3/data/justifications_clean.csv")
-
-# Research questions
-
-# H1
-cor_trust_knowledge <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_knowledge,
-  use = "complete.obs") %>% text_ready()
-cor_trust_acceptance <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_acceptance,
-  use = "complete.obs") %>% text_ready()
-
-# H2
-cor_conspiracy_knowledge <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_knowledge,
-  use = "complete.obs") %>% text_ready()
-cor_conspiracy_acceptance <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_acceptance,
-  use = "complete.obs") %>% text_ready()
-
-# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
-exp3_false_knowledge <- exp3_long %>% 
-  # make numeric versions
-  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
-  ) %>% 
-  group_by(id, knowledge) %>% 
-  # calculate by-participant averages
-  summarize(
-    n = n(),
-    n_accepted = sum(acceptance_num),
-    avg_acceptance  = sum(acceptance_num)/n(), 
-    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
-    BCTI_avg = mean(BCTI_avg), 
-    # add additonal measures for robustness checks
-    CMQ_avg = mean(CMQ_avg), 
-    SICBS = mean(SICBS),
-    wgm_scientists = mean(wgm_scientists), 
-    pew = mean(pew)
-  ) %>% 
-  # filter to only false responses
-  filter(knowledge == FALSE) %>% 
-  ungroup()
-
-# H1b conditional on false answers
-false_answers_cor_trust_acceptance <- cor.test(exp3_false_knowledge$wgm_sciencegeneral, exp3_false_knowledge$avg_acceptance,
-  use = "complete.obs") %>% 
-  text_ready()
-
-# H2b conditional on false answers
-false_answers_cor_conspiracy_acceptance <- cor.test(exp3_false_knowledge$BCTI_avg, exp3_false_knowledge$avg_acceptance,
-  use = "complete.obs") %>% 
-  text_ready()
-
-# extract descriptives for inline reporting
-exp3_descriptives <- list(
-  # Demographics
-  n_subj = n_distinct(exp3_wide$id),
-  n_subject_by_outcome = exp3_wide %>% 
-    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
-                     ~sum(!is.na(.x)
-                     )
-    )
-    ),
-  gender = exp3_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
-  age = exp3_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
-                                       .names = "{.fn}")) %>% rounded_numbers(),
-  # RQ1 & RQ2
-  means = exp3_wide %>% 
-    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
-                     list(mean = mean, sd = sd), 
-                     .names = "{.col}_{.fn}")) %>%
-    rename_with(~sub("^avg_", "", .), everything()) %>% 
-    mutate_if(is.numeric, round, digits = 2) %>% 
-  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
-  # H1
-  # a
-  cor_trust_knowledge = cor_trust_knowledge,
-  # b (conditional on false responses)
-  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
-  # b (all cases)
-  cor_trust_acceptance = cor_trust_acceptance,
-  # H2
-  # a
-  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
-  # b (conditional on false responses)
-  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
-  # b(all cases)
-  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
-  # Exploratory
-  # acceptance and knowledge
-  conditional_acceptance = exp3_long %>% 
-    pivot_longer(c(acceptance, acceptance_initial), 
-                 names_to = "measure", 
-                 values_to = "acceptance") %>% 
-    group_by(knowledge, measure, acceptance) %>% 
-    count() %>% 
-    group_by(knowledge, measure) %>% 
-    mutate (share = n/sum(n), 
-            # rename knowledge values
-            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
-    mutate_if(is.numeric, round, digits = 3) %>% 
-    mutate (share = paste0(share*100, " %")) %>% 
-    super_split(knowledge,  measure, acceptance),
-  # Justifications
-  justifications_n = nrow(exp3_justifications), 
-  justifications_n_participants = n_distinct(exp3_justifications$id),
-  justification_by_category = exp3_justifications %>%
-    group_by(category)%>%
-    summarize(n = n(), 
-              n_subjects = n_distinct(id)) %>%
-    mutate(share = n/sum(n)) %>%
-    mutate_if(is.numeric, round, digits=3) %>% 
-    mutate(share = paste0(share*100, "%")) %>% 
-    split(.$category),
-  # Consensus acceptance
-  acceptance_n = exp3_wide %>% 
-    filter(!is.na(reason_agreement)) %>% 
-    nrow(.),
-  acceptance_by_reason = exp3_wide %>% 
-    group_by(reason_agreement) %>% 
-    summarize(n = n()) %>%
-    drop_na(reason_agreement) %>% 
-    mutate(share = n/sum(n)) %>%
-    mutate_if(is.numeric, round, digits=3) %>% 
-    mutate(share = paste0(share*100, "%")) %>% 
-    split(.$reason_agreement),
-  reason_followup_n = exp3_wide %>% 
-    filter(!is.na(reason_followup)) %>% 
-    nrow(.)
-)
-
-```
-
-```{r exp4}
-# Analyze data of experiments and store results
-
-# Experiment 4
-
-# read the cleaned long version data set
-exp4_long <- read_csv("exp_4/data/cleaned_long.csv")
-
-# read wide version data set
-exp4_wide <- read_csv("exp_4/data/cleaned_wide.csv")
-
-# read coded justifications
-exp4_justifications <- read_csv("exp_4/data/justifications_clean.csv")
-
-# Research questions
-
-# H1
-cor_trust_knowledge <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_knowledge,
-  use = "complete.obs") %>% text_ready()
-cor_trust_acceptance <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_acceptance,
-  use = "complete.obs") %>% text_ready()
-
-# H2
-cor_conspiracy_knowledge <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_knowledge,
-  use = "complete.obs") %>% text_ready()
-cor_conspiracy_acceptance <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_acceptance,
-  use = "complete.obs") %>% text_ready()
-
-# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
-exp4_false_knowledge <- exp4_long %>% 
-  # make numeric versions
-  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
-  ) %>% 
-  group_by(id, knowledge) %>% 
-  # calculate by-participant averages
-  summarize(
-    n = n(),
-    n_accepted = sum(acceptance_num),
-    avg_acceptance  = sum(acceptance_num)/n(), 
-    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
-    BCTI_avg = mean(BCTI_avg), 
-    # add additonal measures for robustness checks
-    CMQ_avg = mean(CMQ_avg), 
-    SICBS = mean(SICBS),
-    wgm_scientists = mean(wgm_scientists), 
-    pew = mean(pew)
-  ) %>% 
-  # filter to only false responses
-  filter(knowledge == FALSE) %>% 
-  ungroup()
-
-# H1b conditional on false answers
-false_answers_cor_trust_acceptance <- cor.test(exp4_false_knowledge$wgm_sciencegeneral, exp4_false_knowledge$avg_acceptance,
-  use = "complete.obs") %>% 
-  text_ready()
-
-# H2b conditional on false answers
-false_answers_cor_conspiracy_acceptance <- cor.test(exp4_false_knowledge$BCTI_avg, exp4_false_knowledge$avg_acceptance,
-  use = "complete.obs") %>% 
-  text_ready()
-
-# extract descriptives for inline reporting
-exp4_descriptives <- list(
-  # Demographics
-  n_subj = n_distinct(exp4_wide$id),
-  n_subject_by_outcome = exp4_wide %>% 
-    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
-                     ~sum(!is.na(.x)
-                     )
-    )
-    ),
-  gender = exp4_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
-  age = exp4_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
-                                       .names = "{.fn}")) %>% rounded_numbers(),
-  # RQ1 & RQ2
-  means = exp4_wide %>% 
-    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
-                     list(mean = mean, sd = sd), 
-                     .names = "{.col}_{.fn}")) %>%
-    rename_with(~sub("^avg_", "", .), everything()) %>% 
-    mutate_if(is.numeric, round, digits = 2) %>% 
-  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
-  # H1
-  # a
-  cor_trust_knowledge = cor_trust_knowledge,
-  # b (conditional on false responses)
-  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
-  # b (all cases)
-  cor_trust_acceptance = cor_trust_acceptance,
-  # H2
-  # a
-  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
-  # b (conditional on false responses)
-  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
-  # b(all cases)
-  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
-  # Exploratory
-  # acceptance and knowledge
-  conditional_acceptance = exp4_long %>% 
-    pivot_longer(c(acceptance, acceptance_initial), 
-                 names_to = "measure", 
-                 values_to = "acceptance") %>% 
-    group_by(knowledge, measure, acceptance) %>% 
-    count() %>% 
-    group_by(knowledge, measure) %>% 
-    mutate (share = n/sum(n), 
-            # rename knowledge values
-            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
-    mutate_if(is.numeric, round, digits = 3) %>% 
-    mutate (share = paste0(share*100, " %")) %>% 
-    super_split(knowledge,  measure, acceptance),
-  # Justifications
-  justifications_n = nrow(exp4_justifications), 
-  justifications_n_participants = n_distinct(exp4_justifications$id),
-  justification_by_category = exp4_justifications %>%
-    group_by(category)%>%
-    summarize(n = n(), 
-              n_subjects = n_distinct(id)) %>%
-    mutate(share = n/sum(n)) %>%
-    mutate_if(is.numeric, round, digits=3) %>% 
-    mutate(share = paste0(share*100, "%")) %>% 
-    split(.$category),
-  # Consensus acceptance
-  acceptance_n = exp4_wide %>% 
-    filter(!is.na(reason_agreement)) %>% 
-    nrow(.),
-  acceptance_by_reason = exp4_wide %>% 
-    group_by(reason_agreement) %>% 
-    summarize(n = n()) %>%
-    drop_na(reason_agreement) %>% 
-    mutate(share = n/sum(n)) %>%
-    mutate_if(is.numeric, round, digits=3) %>% 
-    mutate(share = paste0(share*100, "%")) %>% 
-    split(.$reason_agreement),
-  reason_followup_n = exp4_wide %>% 
-    filter(!is.na(reason_followup)) %>% 
-    nrow(.)
-)
-
-```
-
-## Procedure
-
-After providing their consent to participate in the study, participants were given an attention check "While watching the television, have you ever had a fatal heart attack?" [1-6; 1 = Never, 6 = Often].
-All participants who did not answer "1 = Never" were excluded.
-Participants then read the following instructions:"We will ask you 10 questions about science.
-After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions (in study 1 they were randomly selected from a pool of 11 questions) in random order.
-After each question, participants were presented with an answer reflecting the scientific consensus.
-In study 2 and 3, participants additionally saw a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer.
-In study 4, we provided only two links and no explanation.
-Participants were asked to choose whether they accept the answer or not, before proceeding to the next question.
-Participants then answered questions on conspiracy thinking, conspiracy beliefs and trust in science.
-
-In study 2, 3 and 4, we presented participants with open-ended questions to explain cases where they rejected the scientific consensus.
-In study 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus.
-Finally, at the end of study 3 and 4, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain.
-Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".
-
-## Participants
-
-We recruited a total of `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in study 1, `r exp2_descriptives$n_subj` in study 2, `r exp3_descriptives$n_subj` in study 3, `r exp4_descriptives$n_subj` in study 4) participants in the US via prolific.
-Details and demographics can be found in the online supplemental material.
-While samples for study 1 and 2 were convenience samples, study 3 and 4 were run on a sample holding vaccine-skeptic beliefs (see OSM for details).
-
-## Materials
-
-Table \@ref(tab:knowledge) shows all questions and their answer options used in the four studies.
-
-```{r knowledge}
-# Function to replace special characters in all cells of a data frame
-replace_special_characters <- function(df) {
-  df %>%
-    mutate(across(everything(), ~ gsub("\u2013|\u2014|\u2212", "-", .))) %>% # Replace en-dash, em-dash, and Unicode minus with ASCII hyphen-minus
-    mutate(across(everything(), ~ iconv(., from = "UTF-8", to = "ASCII//TRANSLIT"))) # Ensure all text is ASCII
-}
-
-# Read the CSV file
-items <- read_csv("exp_4/materials/overview_questions.csv")
-
-# Replace special characters
-items_cleaned <- replace_special_characters(items) %>% 
-  mutate(id = 1:nrow(.)) %>%
-  select(id, everything())  # Ensure 'id' is the first column
-
-# add footnote for the tree items
-items_cleaned[11, 2] <- paste0("*", 
-                               items_cleaned[11, 2])
-
-# Output the table
-knitr::kable(items_cleaned, booktabs = T, longtable = TRUE,
-    caption = "Science knowledge items", 
-    full_width = T, 
-    col.names = c("", "Study 1-3", "Study 4")) %>%
-  kable_styling(font_size = 8) %>%
-  column_spec(1, width = "2em") %>%
-  column_spec(2, width = "22em")%>%
-  column_spec(3, width = "22em") %>% 
-  footnote(symbol = "Only used in Study 1")
-```
-
-Our main measure of conspiracy belief was asking participants about specific conspiracy theories.
-We selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) [@pennycookOverconfidentlyConspiratorialConspiracy2022] (Table \@ref(tab:conspiracy)).
-Participants were asked: "Below is a list of events for which the official version has been disputed.
-For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
-[1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true]".
-
-As a robustness check, we also used established scales on conspiracy thinking, namely the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013] and the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016].
-Details and comparisons between the scales can be found in the OSM.
-
-```{r conspiracy, echo=FALSE}
-# Create the data frame
-items <- c(
-  "The Apollo moon landings never happened and were staged in a Hollywood film studio.",
-  "A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).",
-  "The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.",
-  "The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.",
-  "The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.",
-  "There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.",
-  "In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.",
-  "Proof of alien contact is being concealed from the public.",
-  "Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.",
-  "Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.")
-
-conspiracy_items <- data.frame(id = 1:length(items), items = items)
-
-# Output the table
-kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL, 
-    caption = "Conspiracy items", 
-    full_width = F) %>%
-  kable_styling(font_size = 8) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "40em")
-```
-
-Our main item for measuring trust in science was selected from the Wellcome Global Monitor survey: "In general, would you say that you trust science a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]".
-We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey ("How much do you trust scientists in this country?
-Do you trust them a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), the other from the Pew research center ("How much confidence do you have in scientists to act in the best interests of the public?
-[1-5; 1 = No confidence at all, 5 = A great deal of confidence]").
-We selected these items so that we could compare the ratings in our sample to global survey results.
-The WGM survey has been administered in over 140 countries and included over 140000 respondents.
-The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents [@colognaTrustScientistsTheir2024].
-
-# Results
-
-(ref:summary-plot) Points represent the average share of acceptance as a function of the **A** level of trust in science ("In general, would you say that you trust science a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"); **B** average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); **C** belief in a specific conspiracy theory.
-Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: "1 - completly false, 5 - unsure, 9 - completely true".
-Averages are only based on participants who believed at least to some degree in the conspiracy theory.
-We consider a believer as anyone scoring higher than the scale midpoint, i.e.
-\>5.
-Number labels in plots represent the number of participants for the corresponding data point.
-
-```{r summary-plot, fig.cap="(ref:summary-plot)", fig.height= 10, fig.width=10}
-# combine data of all three studies
-combined_data <- bind_rows(exp1_long %>% 
-                             mutate(study = "study 1"), 
-                           exp2_long %>% 
-                             mutate(study = "study 2"), 
-                           exp3_long %>% 
-                             mutate(study = "study 3"),
-                           exp4_long %>% 
-                             mutate(study = "study 4")
-                           )
-
-# make plot data for trust
-plot_trust <- combined_data %>% 
-  drop_na(wgm_sciencegeneral, acceptance) %>% 
-  # make a labels version of trust in science
-  mutate(wgm_sciencegeneral = case_when(
-    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
-    wgm_sciencegeneral == 2 ~ "2 (Not much)",
-    wgm_sciencegeneral == 3 ~ "3 (Some)",
-    wgm_sciencegeneral == 4 ~ "4 (A lot)",
-    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
-  )) %>% 
-  group_by(study, wgm_sciencegeneral, acceptance) %>% 
-  summarize(n = n(), 
-            n_participants = n_distinct(id)) %>% 
-  group_by(study, wgm_sciencegeneral) %>% 
-  mutate (share = n/sum(n), 
-          n_participants = max(n_participants)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-# make trust plot
-trust <- ggplot(plot_trust %>% 
-         filter(acceptance == "Yes"), 
-       aes(x = wgm_sciencegeneral, y = share, color = study, shape = study)) +
-  geom_point(position = "dodge", alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
-  geom_text_repel(aes(label = paste0(n_participants)),
-            vjust = -0.5, size = 3,
-            show.legend = FALSE) +
-  scale_color_viridis_d(option = "plasma", end = 0.8) +
-  labs(x = "Trust in science", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
-  plot_theme + 
-  theme(legend.position = "top")
-
-# make plot data for conspiracy belief
-plot_conspiracy <- combined_data %>% 
-  # bring to long format
-  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
-               names_to = "BCTI_item", values_to = "score") %>% 
-  # make a belief variable if participant indicates a numeric answer above the
-  # scale midpoint
-  mutate(belief = ifelse(score > 5, TRUE, 
-                         ifelse(is.na(score), NA, FALSE)
-                         ),
-         BCTI_item = sub("^BCTI_", "", BCTI_item)
-         ) %>% 
-  group_by(study, BCTI_item, belief, acceptance) %>% 
-  summarize(n = n(),
-            n_participants = n_distinct(id)) %>% 
-  group_by(study, BCTI_item, belief) %>% 
-  mutate (share = n/sum(n), 
-          n_participants = max(n_participants)) %>% 
-  mutate_if(is.numeric, round, digits = 3) %>% 
-  ungroup()
-
-# make conspiracy belief plot
-conspiracy <- ggplot(plot_conspiracy %>% 
-         filter(acceptance == "Yes" & belief == TRUE), 
-       aes(x = BCTI_item, y = share, color = study, shape = study)) +
-  geom_point(position = "dodge", alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
-  geom_text_repel(aes(label = paste0(n_participants)),
-            vjust = -0.5, size = 3,
-            show.legend = FALSE) +
-  scale_color_viridis_d(option = "plasma", end = 0.8) +
-  labs(x = "Conspiracy theory belief", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
-  plot_theme +
-  theme(axis.text.x = element_text(angle = 30, hjust = 1))
-
-# Conspiracy Thinking
-
-# Define breaks for each integer from 1 to 9
-breaks <- seq(0, 100, by = 20)
-
-# Define labels for each bin in the desired format
-labels <- paste0(c(0, 20, 40, 60, 80), "-", c(20, 40, 60, 80, 100))
-
-plot_data <- combined_data %>% 
-  drop_na(CMQ_avg) %>% 
-  # Create the binned version of CMQ_avg
-  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
-  group_by(study, CMQ_avg_binned, acceptance) %>% 
-  summarize(n = n()) %>% 
-  group_by(study, CMQ_avg_binned) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-n_participants <- combined_data %>% 
-  drop_na(CMQ_avg) %>% 
-  # Create the binned version of CMQ_avg
-  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
-  group_by(study, CMQ_avg_binned) %>% 
-  summarize(n_participants = n_distinct(id))
-
-plot_conspiracy_thinking <- left_join(plot_data, n_participants)
-
-# make plot
-conspiracy_thinking  <- ggplot(plot_conspiracy_thinking  %>% 
-         filter(acceptance == "Yes"), 
-       aes(x = CMQ_avg_binned, y = share, color = study, shape = study)) +
-  geom_point(position = "dodge", alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
-  geom_text_repel(aes(label = paste0(n_participants)),
-            vjust = -0.5, size = 3,
-            show.legend = FALSE) +
-  scale_color_viridis_d(option = "plasma", end = 0.8) +
-  labs(x = "Conspiracy Thinking", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
-  plot_theme + 
-  theme(legend.position = "top")
-
-# remove elements of plots for better overall plot
-conspiracy_thinking <- conspiracy_thinking + 
-  #rremove("ylab") +
-  theme(axis.title.y = element_blank(),
-        axis.text.y = element_blank(),
-        axis.ticks.y = element_blank()) 
-  #guides(color = "none", shape = "none")
-
-conspiracy <- conspiracy +
-  rremove("ylab") +
-  guides(color = "none", shape = "none")
-
-trust <- trust +
-  rremove("ylab")
-
-# common plot
-patchwork <- (trust + conspiracy_thinking)/conspiracy + 
-  plot_layout(guides = 'collect', widths = c(1, 1)) +
-  plot_annotation(tag_levels = 'A')  & theme(legend.position = "top")
-
-wrap_elements(patchwork) +
-  labs(tag = "Acceptance rate of consensus") +
-    theme(
-    plot.tag = element_text(angle = 90, , face = "bold", size = 12),
-    plot.tag.position = "left"
-  )
-```
-
-Across all studies we find that almost all participants trust basic and recent/advance science facts almost all the time.
-That is true even for participants who expressed low general trust in science, and those who have a tendency to conspiracy think or hold science-related conspiracy beliefs (Fig.
-\@ref(fig:summary-plot)).
-
-Our results suggest that the more people tend to trust science, the more they tend to know about science and accept the scientific consensus.
-In study 2 and 3, we find a positive but small correlation between trust in science and both science knowledge (study 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`).
-In study 3 and 4--both run on vaccine-skeptic samples--we find no statistically significant correlation between science knowledge and trust in science (study 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`), but a small positive correlation between acceptance of scientific consensus and trust in science (study 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`).
-
-As for conspiracy beliefs, results are mixed.
-In studies 1 and 2, we find a small negative correlation between conspiracy thinking and both science knowledge (study 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`).
-However, in study 3 and 4, we do not find a statistically significant correlation with neither science knowledge (study 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`), and between nor acceptance of scientific consensus (study 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`).
-
-Overall, the correlations we find are relatively weak, which might be partly due to ceiling effects.
-This was the case in particular for acceptance: In all studies, in the vast majority of cases, participants accepted accepted science facts.
-The absence of correlations between conspiracy beliefs and science knowledge or acceptance in studies 3 and 4 might be explained by our sample selection.
-For these studies, we recruited samples scoring considerably higher in conspiracy belief than the convenience samples of study 1 and 2 (see OSM).
-The presence of effects in study 1 and 2, and the absence in study 3 and 4 might be because differences between moderate and high conspiracy thinking might not matter, while, differences between low and moderate/high conspiracy thinking do matter.
+# Acceptance conditional on knowledge {#acceptance-knowledge}
 
@@
 In most cases (study 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-Only in study 4 do we find evidence[^3] that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in study 2 evidence that it is associated with (less) conspiracy thinking (study 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
 In very few cases, participants changed their mind away from the scientific consensus.
 These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: `r exp1_descriptives$conditional_acceptance$true$No$share`; study 2: `r exp2_descriptives$conditional_acceptance$true$No$share`; study 3: `r exp3_descriptives$conditional_acceptance$true$acceptance$No$share`; study 4: `r exp4_descriptives$conditional_acceptance$true$acceptance$No$share`).
+Figure \@ref(fig:acceptance-knowledge) shows acceptance rates conditional on whether the initial answer to the questions was correct or not.
 
-[^3]: For this analysis, we restricted our data to those cases in which participants initially gave the wrong answer.
+(ref:acceptance-knowledge) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
 
-Why did participants reject the scientific consensus?
-We got a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (study 2: `r exp2_descriptives$justifications_n`; study 3: `r exp3_descriptives$justifications_n`; study 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (study 2: `r exp2_descriptives$justifications_n_participants`; study 3: `r exp3_descriptives$justifications_n_participants`; study 4: `r exp4_descriptives$justifications_n_participants`) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Based on the answers, we created five categories.
-Table \@ref(tab:justifications) summarizes these answers by five categories.
-All answers can be read in the OSM.
+```{r acceptance-knowledge, fig.cap="(ref:acceptance-knowledge)", fig.width=8, fig.height=8}
+plot_data <- combined_data %>% 
+  mutate(study = paste0("Study ", study)) %>% 
+  group_by(study, knowledge, acceptance) %>% 
+  count() %>% 
+  group_by(study, knowledge) %>% 
+  mutate (share = n/sum(n)) %>% 
+  mutate_if(is.numeric, round, digits = 3)
 
-```{r justifications}
+absolute_numbers <- plot_data %>% 
+  group_by(study, knowledge) %>% 
+  summarise(n = sum(n))
 
-## make a common table of exp 2 and exp 3
-justifications <- bind_rows(exp2_justifications %>% 
-                              mutate(study = "study 2"), 
-                            exp3_justifications %>% 
-                              mutate(study = "study 3", 
-                                     final_coding = as.character(final_coding)),
-                            exp4_justifications %>% 
-                              mutate(study = "study 4"))
-
-justifications %>%
-  group_by(category)%>%
-  summarize(n = n(), 
-            n_subjects = n_distinct(id)) %>%
-  mutate(Share = n/sum(n)) %>%
-  mutate_if(is.numeric, round, digits=3) %>% 
-  mutate(Share = paste0(Share*100, "%")) %>% 
-  arrange(desc(n)) %>% 
-  select(category, n, Share, n_subjects) %>% 
-  ungroup() %>% 
-  rename(
-    Category = category,
-    `N (instances)` = n, 
-    `Share (instances)` = Share, 
-    `N (unique participants)` = n_subjects
-  ) %>% 
-  apa_table(caption = "Justifications by category for data from studies 2, 3 and 4 combined.")
+ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
+  geom_col(position = "dodge") +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
+  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
+            position = position_dodge(width = 0.9),
+            vjust = -0.5, size = 3) +
+  scale_fill_viridis_d(option = "cividis") +
+  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
+  plot_theme +
+  geom_bracket(data = absolute_numbers,
+               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
+                   xmax = as.numeric(factor(knowledge)) + 0.4, 
+                   y.position = 1.05, label =  paste0("n = ", n)),
+               inherit.aes = FALSE,
+               step.increase = 0)+
+  facet_wrap(~study)
 ```
-
-Why did participants accept the scientific consensus?
-In studies 3 and 4, we had asked participants about cases where they agreed with the scientific consensus.
-A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (study 3: `r exp3_descriptives$acceptance_n`; study 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
-Of these, `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 4 said they accepted the scientific consensus because they trust scientists, while `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 4 said they independently verified the fact.
-We asked all participants who gave this answer to explain how they did it (answers can be found in the OSM).
-`r exp3_descriptives$acceptance_by_reason[["other"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in study 4 answered with other "other" and gave an open-ended explanation (see OSM).
-
-More detailed results addressing all our pre-registered research questions can be found in the OSM.
-
-# Discussion
-
-Survey data from the US suggests a decline of trust in science in recent years [@brianAmericansTrustScientists2023].
-Across the globe, many people only trust science ‘some of the time’ , and a sizable minority doesn’t trust it [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-This appears difficult to reconcile with the fact that, in many countries, most people nowadays receive at least a basic education [@ritchieGlobalEducation2023], and that science education has been found to be the most strongly associated variable with trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018].
-
-But how deep is distrust of science?
-In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
-Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
-That is true not only for basic science questions (studies 1-3), but also for more recent scientific findings that most people are unfamiliar with (study 4).
-In study 1 the acceptance rate after having given a wrong answer was considerably lower (`r exp1_descriptives$conditional_acceptance$false$No$share`) than in subsequent studies (study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`).
-While this could be just sampling variation, it might be that adding explanations and/or sources convinced participants more than merely stating the consensus.
-In general, participants with lower trust in science and who believe more in conspiracy theories tended to both know less about science and accept the scientific consensus less.
-
-Scholars have long interpreted the lack of basic science knowledge as the root cause of distrust in and more generally unfavorable attitudes towards science, an explanation known under the term "deficit-model" [@millerMeasurementCivicScientific1998a; @millerPublicUnderstandingAttitudes2004; @durantPublicUnderstandingScience1989].
-In line with proponents of the deficit-model, we find that people indeed err even on very basic science questions.
-However, we also find that almost everyone instantly accepted the scientific consensus once informed about it.
-Moreover, the same was true for non-basic science questions.
-This calls into question the deficit model: If participants trust science on basic (and non-basic) questions—as suggests immediate acceptance of the scientific consensus—then unfavorable attitudes towards science must have other explanations than a mere lack of knowledge.
-This argument is supported by past research.
-In a seminal meta-analysis, \@allumScienceKnowledgeAttitudes2008 found correlations between science knowledge and general attitudes towards science to be at best small and largely context dependent [\@allumScienceKnowledgeAttitudes2008].
-A more recent report by the National Academies of Sciences, Engineering, and Medicine states that science knowledge is "unlikely to substantially affect attitudes on scientific issues" [@nationalacademiesofsciencesCommunicatingScienceEffectively2017].
-
-How to make sense of the detachment of basic science knowledge from attitudes towards science?
-One explanation might be that most scientific findings are very abstract.
-For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second.
-It is therefor easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
-In other words, people might hold on to most scientific findings as reflective beliefs—beliefs that have little effect on our behaviors [@mercierHowGoodAre2021].
-It is probably when it becomes tangible that science shape's people's attitudes.
-This can happen, for example, when science appears to support conclusions in favor of certain policies or moral positions, as for example in the case of anthropogenic climate change or vaccines.
-
-It is tempting, especially for scientists, to explain mistrust in science with ignorance.
-After all, careful proof of objective truth is what makes science trustworthy—that is, for scientists.
-For most people, what shapes trust in science might not be so different from what shapes trust in other institutions, such as the government, banks or church.
-
-\FloatBarrier
-
-# References
-
-::: {#refs}
-:::
-
-\newpage
-
-# (APPENDIX) Appendix {.unnumbered}
-
-```{r child = "appendix_exp1.Rmd"}
-```
-
-\clearpage
-
-```{r child = "appendix_exp2.Rmd"}
-```
-
-\clearpage
-
-```{r child = "appendix_exp3.Rmd"}
-```
-
-\clearpage
-
-```{r child = "appendix_exp4.Rmd"}
-```

### 2024-09-12 paper.Rmd
first version within word limit of 4k

@@
   - "appendix_exp4.Rmd"
+  - "appendix_acceptance_knowledge.Rmd"
 
@@
 
-Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change (ref) or vaccination (ref) to following recommendations during COVID [@alganTrustScientistsTimes2021], which suggests that trust in science was the most important predictor of these behaviors).
+Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change [@colognaRoleTrustClimate2020] or vaccination [@sturgisTrustScienceSocial2021; @lindholtPublicAcceptanceCOVID192021] to following recommendations during COVID [@alganTrustScientistsTimes2021], which suggests that trust in science was the most important predictor of these behaviors).
 
@@
 
-This relative lack of trust in science could mean different things.
-It could mean that some people reject science wholesale, or at least to a significant extent.
-Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (e.g.
-creationism), or beliefs that are associated with disliked behaviors (e.g.
-vaccination, acting for climate change).
-
-We answer this question by looking at how much Americans accept basic scientific knowledge.
-We pay special attention to two groups of participants: people who say they don’t trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
-
-We briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
-
-## Trust in science
-
-Across the globe, most people say they trust science at least to some extent.
-In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018].
-In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants [@wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-Across all countries, in 2018, 32% of participants said they trust science "a lot" (41% in 2020), 45% trust "some" science (39% in 2020), and only 13% percent trust science "not much/ not at all" (13% in 2020), while 10% indicated "don't know" (7% in 2020)[^1].
-
-[^1]: <https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020#gid=6acd&pid=0>\>
-
-Trust in science also appears to be relatively stable in the US [@funkPublicConfidenceScientists2020].
-However, there are some potentially worrying trends: Recently, trust in US has dropped [@brianAmericansTrustScientists2023], and some evidence suggests increasing polarization of trust in science along political lines [@gauchatPoliticizationSciencePublic2012].
-For example, the partisan divide on the question of whether climate change is a major threat to society has been growing in the last decade[^2] [@kennedyWhatDataSays2023].
-A recent study suggests that by some measure, only about half of the US population believes in anthropogenic climate change, and that individuals do mostly not seem to change their mind on on the issue [@mottaChangingMindsChanging2021].
-
-[^2]: <https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/>
-
-Trust in science is typically measured in three ways: either asking explicitly about general trust in science or scientists, by asking about science attitudes (e.g.
-support for public funding of science), or asking questions about specific, typically contentious, science topics (e.g.
-vaccines, climate change, GMOs).
-It is unclear to which extent these measures capture trust in basic science.
-To assess this, we can look at the literature on science knowledge.
-
-[introduce SK as: that gives us a lower bound of how much people trust basic science]
-
-## Science knowledge
-
-Much of the research on public understanding of science measures to which extent people have basic science knowledge (e.g.
-"Does the Earth go around the Sun, or does the Sun go around the Earth?").
-For us, these surveys provide a lower bound to how much people trust basic science: people who give the right answer (and not by chance) supposedly trust science on that question.
-Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge [@millerPublicUnderstandingAttitudes2004; @millerMeasurementCivicScientific1998a].
-For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016].
-A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.
-
-[To do: Make a descriptive overview of science knowledge across time/countries.]
-
-## CT and TiS
-
-Recently, the relationship of one group to trust in science (TiS) has been particularly scrutinized: conspiracy theorists (CTists), either defined as people who endorse specific conspiracy theories, or as people who have a conspiracy mindset [@rutjensConspiracyBeliefsScience2022; @vranicDidMyOwn2022].
-CTists reject specific science knowledge (related to their CT), and they tend to trust science less on the whole.
-What remains unclear is to which extent CTists (dis)trust basic science.
-
 ## The present studies
@@
 In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts.
-For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted correct answer, and asked them whether they accepted it (measuring their trust in basic science).
+For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted answer, and asked them whether they accepted it (measuring their trust in basic science).
 We also measured participants’ trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF)).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF).
 The differences between the four studies are summarized presently, and the methods are detailed below.
@@
 In Study 4 we used more recent much less basic scientific discoveries (e.g.
-"How many more glial cells are there in the brain in comparison with neurons, as established in 2016?
-[The same amount; Twice as many; Tenth as many]"; correct answer: The same amount).
+"What is the electric charge of the Higgs Boson, as established in 2012?
+[1.602176634 × 10-19; 0; 3.2×10-19C]"; correct answer: 0, i.e.
+electrically neutral).
 
@@
 In Study 1, we simply told participants that they would be provided with the scientifically consensual answer.
-However, for participants to accept this answer, they must not only trust science, but also trust us (that we are presenting them with the actual scientifically consensual answer).
+However, for participants to accept this answer, they must not only trust science, but also trust that we are presenting them with the actual scientifically consensual answer.
 To remove this issue, in Studies 2 to 3, we presented participants with a short explanation of the correct answer, as well as links to three sources per answer (e.g.
 Wikipedia, National Geographic or NASA).
-In Study 4, as the topics were more complex, we did not provide an explanation, but we provided two sources per answer.
+In Study 4, as the topics were more complex, we did not provide an explanation, but still provided two sources per answer.
 
@@
 
-# Exploratory: Differences by reason for trust
+# Exploratory: Differences by reason
+
+# trust
 trust_by_reason_model <- run_regression(exp3_wide, dependent_var = "wgm_sciencegeneral", independent_var = "reason_agreement")
 
+# acceptance
 acceptance_by_reason_model <-run_regression(exp3_wide, dependent_var = "avg_acceptance", independent_var = "reason_agreement")
+
+# BCTI (conspiracy beliefs)
+BCTI_by_reason_model <- run_regression(exp3_wide, dependent_var = "BCTI_avg", independent_var = "reason_agreement")
+
+# survey time
+time_by_reason_model <- run_regression(exp3_wide %>% filter(duration_mins < 30), dependent_var = "duration_mins", independent_var = "reason_agreement")
 
@@
     group_by(reason_agreement) %>% 
-    summarize(across(c(wgm_sciencegeneral, avg_acceptance), 
-                     list(mean = mean, sd = sd), 
+    summarize(across(c(wgm_sciencegeneral, avg_acceptance, BCTI_avg), 
+                     list(mean = ~mean(.x, na.rm = TRUE), 
+                          sd = ~sd(.x, na.rm = TRUE)), 
                      .names = "{.col}_{.fn}")) %>%
@@
   reason_trust_model = trust_by_reason_model %>% split(.$term), 
-  reason_acceptance_model = acceptance_by_reason_model %>% split(.$term)
+  reason_acceptance_model = acceptance_by_reason_model %>% split(.$term),
+  reason_BCTI_model = BCTI_by_reason_model %>%  split(.$term),
+  reason_time_model = time_by_reason_model %>%  split(.$term), 
+  n_participants_over30mins = exp3_wide %>% 
+    filter(duration_mins >= 30) %>% 
+    nrow(.)
 )
@@
 
-# Exploratory: Differences by reason for trust
+# Exploratory: Differences by reason
+
+# trust
 trust_by_reason_model <- run_regression(exp4_wide, dependent_var = "wgm_sciencegeneral", independent_var = "reason_agreement")
 
-acceptance_by_reason_model <- run_regression(exp4_wide, dependent_var = "avg_acceptance", independent_var = "reason_agreement")
-
+# acceptance
+acceptance_by_reason_model <-run_regression(exp4_wide, dependent_var = "avg_acceptance", independent_var = "reason_agreement")
+
+# BCTI (conspiracy beliefs)
+BCTI_by_reason_model <- run_regression(exp4_wide, dependent_var = "BCTI_avg", independent_var = "reason_agreement")
+
+# survey time
+time_by_reason_model <- run_regression(exp4_wide %>% filter(duration_mins < 30), dependent_var = "duration_mins", independent_var = "reason_agreement")
+
+# clicks
 clicks_by_reason_model <- run_regression(exp4_wide, dependent_var = "sum_link_clicks", independent_var = "reason_agreement")
@@
     group_by(reason_agreement) %>% 
-    summarize(across(c(wgm_sciencegeneral, avg_acceptance, sum_link_clicks), 
+    summarize(across(c(wgm_sciencegeneral, avg_acceptance, sum_link_clicks, BCTI_avg), 
                      list(mean = mean, sd = sd), 
@@
   reason_acceptance_model = acceptance_by_reason_model %>% split(.$term), 
-  reason_clicks_model = clicks_by_reason_model %>% split(.$term)
+  reason_clicks_model = clicks_by_reason_model %>% split(.$term),
+  reason_BCTI_model = BCTI_by_reason_model %>%  split(.$term),
+  reason_time_model = time_by_reason_model %>%  split(.$term), 
+  n_participants_over30mins = exp4_wide %>% 
+    filter(duration_mins >= 30) %>% 
+    nrow(.)
 )
@@
                              mutate(study = "4")
-                           )
+                           ) %>% 
+  # make a unique id variable
+  mutate(id = paste0("study", study, "_", id))
 
@@
     mutate(share = paste0(share*100, " %")) %>% 
-    split(.$knowledge)
+    split(.$knowledge),
+  # flat earthers with no trust at all in science
+  # number of participants
+  flat_earth_no_trust_nparticipants = combined_data %>% 
+    filter(BCTI_flatearth == 9 & wgm_sciencegeneral == 1) %>%
+    summarize(n_participants = n_distinct(id)) %>% 
+    pull(),
+  # acceptance
+  flat_earth_no_trust_acceptance = combined_data %>% 
+    filter(BCTI_flatearth == 9 & wgm_sciencegeneral == 1) %>%
+    group_by(acceptance) %>% 
+    summarise(n = n()) %>% 
+    mutate(share = n/sum(n)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(share = paste0(share*100, " %")) %>% 
+    split(.$acceptance), 
+  # Lowest decile acceptance
+  lowest_decile_acceptance = combined_data %>% 
+    group_by(id) %>% 
+    summarise(acceptance_rate = sum(acceptance == "Yes")/n(), 
+              n = n()) %>% 
+    arrange(acceptance_rate) %>% 
+    slice(1:(n() * 0.1)) %>%  # Get the lowest 10% (decile)
+    summarize(mean_acceptance_rate = mean(acceptance_rate, na.rm = TRUE), 
+              n_smaller_or_equalto_0.5 = sum(acceptance_rate <=0.5)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(mean_acceptance_rate = paste0(mean_acceptance_rate*100, " %")), 
+  # Lowest decile knowledge
+  lowest_decile_knowledge = combined_data %>% 
+    group_by(id) %>% 
+    summarise(acceptance_rate = sum(acceptance == "Yes")/n(), 
+              knowledge_rate = mean(knowledge, na.rm=TRUE),
+              n = n()) %>% 
+    arrange(knowledge_rate) %>% 
+    slice(1:(n() * 0.1)) %>%  # Get the lowest 10% (decile)
+    summarize(mean_acceptance_rate = mean(acceptance_rate, na.rm = TRUE), 
+              mean_knowledge_rate = mean(knowledge_rate, na.rm = TRUE)) %>% 
+    mutate_if(is.numeric, round, digits = 3) %>% 
+    mutate(mean_acceptance_rate = paste0(mean_acceptance_rate*100, " %"), 
+           mean_knowledge_rate = paste0(mean_knowledge_rate*100, " %"))
 )
@@
 
-For study 2, we restricted our main hypotheses about acceptance to cases in which participants initially provided a wrong answer.
-The rationale was to see if trust in science or conspiracy belief were associated with a change of mind towards the consensus.
-We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM[^3].
-However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting simple, unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
-
-[^3]: Only in study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (study 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
+For Study 2, we restricted our main hypotheses about acceptance to cases in which participants initially provided a wrong answer.
+However, this meant the more participants had initially provided correct answers, the fewer opportunities they had for accepting correct answers.
+We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM[^1].
+However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
+
+[^1]: Only in Study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (Studies 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in Study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (Studies 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).
 
@@
 Participants then read the following instructions:"We will ask you 10 questions about science.
-After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions (in Study 1 they were randomly selected from a pool of 11 questions, as one ‘trick’ question had been added) in random order.
+After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions in random order.
 After each question, participants were presented with an answer reflecting the scientific consensus, and asked whether they accepted it.
@@
 While samples for Studies 1 and 2 were convenience samples, Studies 3 and 4 were conducted on a sample holding vaccine-skeptic beliefs.
-Prolific Academic allows to select participants based on their answers to a range of questions.
+Prolific Academic allows selecting participants based on their answers to a range of questions.
 We picked three of these questions and only recruited participants who met our criteria for each of them:
@@
 "Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: [For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say"].
-We selected particpants who answered "Against".
+We selected pariticpants who answered "Against".
 2.  "Have you received a coronavirus (COVID-19) vaccination? [Yes (at least one dose); No; Prefer not to answer]". We select only people who answered "No".
@@
 [1 (totally disagree); 2 (disagree); 3 (somewhat disagree); 4 (neither agree nor disagree); 5 (somewhat agree); 6 (agree); 7 (totally agree); rather not say]".
-We select only people who answered '1', '2', or '3'.
+We select only people who answered "1", "2", or "3".
 
@@
 Study 4 used 10 more recent scientific discoveries.
-Table \@ref(tab:knowledge) shows all questions and their answer options used in the four studies.
+Table \@ref(tab:knowledge) shows all questions and their answer options.
 
@@
 For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
-[1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true]".
+[1-9; labels: 1 - completely false, 5 - unsure, 9 - completely true]".
 
@@
 We used the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013] and the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016].
-Details and comparisons between the scales can be found in the OSM.
+Details and comparisons between the scales can be found in the ESM.
 
@@
   #                 show.legend = FALSE) +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.02)) +
   scale_color_viridis_d(option = "plasma", end = 0.8) +
@@
   scale_color_viridis_d(option = "plasma", end = 0.8) +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.01)) +
+  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.02)) +
   labs(x = "Conspiracy Thinking", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
@@
 Note that this includes both participants who had previously correctly answered the knowledge question, and participants who changed their mind when presented with the scientific consensus.
-In Studies 3 & 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: `r exp3_descriptives$means$acceptance_initial_mean`; in Study 4: `r exp4_descriptives$means$acceptance_initial_mean`).
-
-As shown in Figure \@ref(fig:summary-plot), these very high rates of acceptance hold for: participants who they do not trust science at all (`r combined_descriptives$participants_by_trust$'1'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_trust$'1'$Yes$share`), participants who rank in the top two deciles of the conspiracy thinking scale (`r combined_descriptives$participants_by_conspiracy_thinking$'80-100'$share` of participants, average acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_thinking$'80-100'$Yes$share`), participants who consider as "completely true" (the maximum of the 9-point scale) conspiracy theories stating that the earth is flat (`r combined_descriptives$participants_by_conspiracy_belief$flatearth$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share`), or that climate change due to fossil emissions is a hoax (`r combined_descriptives$participants_by_conspiracy_belief$climatechange$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$climatechange$'9'$share`).
-This doesn't merely reflect science knowledge: aggregated across all studies--participants only correctly answered `r combined_descriptives$mean_knowledge$true$share` of the questions (Study 1: `r exp1_descriptives$means$knowledge_mean`; Study 2: `r exp2_descriptives$means$knowledge_mean`; Study 3: `r exp3_descriptives$means$knowledge_mean`; Study 4: `r exp4_descriptives$means$knowledge_mean`) before they were provided with the scientifically consensual answer.
-Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
+In Studies 3 and 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: `r exp3_descriptives$means$acceptance_initial_mean`; in Study 4: `r exp4_descriptives$means$acceptance_initial_mean`).
+
+As shown in Figure \@ref(fig:summary-plot), these very high rates of acceptance hold for: participants who do not trust science at all (`r combined_descriptives$participants_by_trust$'1'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_trust$'1'$Yes$share`), participants who rank in the top two deciles of the conspiracy thinking scale (`r combined_descriptives$participants_by_conspiracy_thinking$'80-100'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_thinking$'80-100'$Yes$share`), participants who consider as "completely true" (the maximum of the 9-point scale) conspiracy theories stating that the earth is flat (`r combined_descriptives$participants_by_conspiracy_belief$flatearth$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share`), or that climate change due to fossil emissions is a hoax (`r combined_descriptives$participants_by_conspiracy_belief$climatechange$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$climatechange$'9'$share`).
+
+Participants in the lowest decile of acceptance still had an average acceptance rate of `r combined_descriptives$lowest_decile_acceptance$mean_acceptance_rate`.
+Even participants who considered as "completely true" that the earth is flat and who said they do "not trust science at all" (N = `r combined_descriptives$flat_earth_no_trust_nparticipants`) had an average acceptance rate of `r combined_descriptives$flat_earth_no_trust_acceptance$Yes$share`.
+
+These high acceptance rates do not merely reflect science knowledge: participants only correctly answered `r combined_descriptives$mean_knowledge$true$share` of the questions (Study 1: `r exp1_descriptives$means$knowledge_mean`; Study 2: `r exp2_descriptives$means$knowledge_mean`; Study 3: `r exp3_descriptives$means$knowledge_mean`; Study 4: `r exp4_descriptives$means$knowledge_mean`) before they were provided with the scientifically consensual answer.
+Even the lowest decile in science knowledge which, on average, answered correctly only on `r combined_descriptives$lowest_decile_knowledge$mean_knowledge_rate` of the questions, had an average acceptance rate of `r combined_descriptives$lowest_decile_knowledge$mean_acceptance_rate`.
+
+Did participants who had initially provided a wrong answer change their minds towards the scientific consensus?
 Yes.
 In most cases (study 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-In very few cases, participants changed their mind away from the scientific consensus.
-These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: `r exp1_descriptives$conditional_acceptance$true$No$share`; study 2: `r exp2_descriptives$conditional_acceptance$true$No$share`; study 3: `r exp3_descriptives$conditional_acceptance$true$acceptance$No$share`; study 4: `r exp4_descriptives$conditional_acceptance$true$acceptance$No$share`).
-
-Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`), but a less consistent relation between trust in science and science knowledge (Studies 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`).
-
-Regarding H2, the results are mixed both for the relation between conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) and acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`) as well as science knowledge (Studies 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`).
+
+Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`), but a less consistent relation between trust in science and science knowledge (Studies 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`).
+
+Regarding H2, the results are mixed for the relation of conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) with both acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`) and science knowledge (Studies 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`).
 
 Why did participants reject the scientific consensus?
-We got a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (study 2: `r exp2_descriptives$justifications_n`; study 3: `r exp3_descriptives$justifications_n`; study 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (study 2: `r exp2_descriptives$justifications_n_participants`; study 3: `r exp3_descriptives$justifications_n_participants`; study 4: `r exp4_descriptives$justifications_n_participants`) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+We collected a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (study 2: `r exp2_descriptives$justifications_n`; study 3: `r exp3_descriptives$justifications_n`; study 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (study 2: `r exp2_descriptives$justifications_n_participants`; study 3: `r exp3_descriptives$justifications_n_participants`; study 4: `r exp4_descriptives$justifications_n_participants`) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
 Based on the answers, we created five categories.
 Table \@ref(tab:justifications) summarizes these answers by five categories.
-All answers can be read in the OSM.
+All answers can be read in the ESM.
 
@@
   ) %>% 
+  # remove decimal places
+  mutate_if(is.numeric, ~as.integer(.x)) %>% 
   apa_table(caption = "Justifications by category for data from studies 2, 3 and 4 combined.")
@@
 Why did participants accept the scientific consensus?
-In Studies 3 and 4--the vaccine skeptic samples--we had asked participants about cases where they agreed with the scientific consensus.
+In Studies 3 and 4--the vaccine hesitant samples--we had asked participants about cases in which they agreed with the scientific consensus.
 A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (Study 3: `r exp3_descriptives$acceptance_n`; Study 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
-There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than there were participants saying it was because they trust scientists (Study 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`).
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than participants saying it was because they trust scientists (Study 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`)^[`r exp3_descriptives$acceptance_by_reason[["other"]]$share` in Study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in Study 4 answered with "other" and gave an open-ended explanation (see ESM).].
 We asked participants who said they verified independently to explain how they did it (answers can be found in the ESM).
-`r exp3_descriptives$acceptance_by_reason[["other"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in study 4 answered with other "other" and gave an open-ended explanation (see ESM).
-
-In an exploratory analysis, we ran linear regressions to test whether there are differences between these two groups.
-Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = `r exp3_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 4; Study 4: mean = `r exp4_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value`) than those who said they verified independently (Study 3 mean = `r exp3_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`; Study 4 mean = `r exp4_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`).
-We did not find a difference regarding acceptance (Study 3: $\hat{\beta}_{\text{Acceptance}}$ = `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value` on a scale from 0 to 1; Study 4: $\hat{\beta}_{\text{Acceptance}}$ = `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value`).
-In study 4, where we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
-On average, participants clicked only clicked on `r exp4_descriptives$reason_means$'trust in scientists'$sum_link_clicks_mean` links (out of 20 possible clicks) and there was no difference between the two groups ($\hat{\beta}_{\text{Clicks}}$ = `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$estimate`, `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$p.value`).
-
-More detailed results addressing all our pre-registered research questions can be found in the OSM.
+
+In an exploratory analysis, we ran linear regressions to test whether there are differences between participants who said they had trusted science and those who said they had verified the information independently.
+Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = `r exp3_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 4; Study 4: mean = `r exp4_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value`) than those who said they verified independently (Study 3 mean = `r exp3_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`; Study 4 mean = `r exp4_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`).
+We did not find a difference regarding acceptance (Study 3: $\hat{\beta}_{\text{Acceptance}}$ = `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value` on a scale from 0 to 1; Study 4: $\hat{\beta}_{\text{Acceptance}}$ = `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value`) or regarding beliefs in conspiracy theories (Study 3: $\hat{\beta}_{\text{BCTI}}$ = `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 9; Study 4: $\hat{\beta}_{\text{BCTI}}$ = `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value`).
+We also did not find a difference in time spent on the survey in Study 3 ($\hat{\beta}_{\text{Time}}$ = `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp3_wide$duration_mins)` mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate %>% round(digits = 0) %>% abs()` minutes less on the survey ($\hat{\beta}_{\text{Time}}$ = `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp4_wide$duration_mins)` mins).^[For these analyses, we excluded outliers who took over 30mins for the survey, which was estimated to take around 10 mins, and for which the median time was `r median(combined_data$duration_mins) %>% round(digits = 0)` minutes.
+As a result, we excluded `r exp3_descriptives$n_participants_over30mins` participant in Study 3 and `r exp4_descriptives$n_participants_over30mins` in Study 4.
+Significance levels remain the same when including these outliers.] In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+On average, participants clicked only on `r exp4_descriptives$reason_means$'trust in scientists'$sum_link_clicks_mean` links (out of 20 possible clicks) and there was no difference between the two groups ($\hat{\beta}_{\text{Clicks}}$ = `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$p.value`).
+
+More detailed results addressing all our pre-registered research questions can be found in the ESM.
 
@@
 
-Survey data from the US suggests a decline of trust in science in recent years [@brianAmericansTrustScientists2023].
-Across the globe, many people only trust science ‘some of the time’ , and a sizable minority doesn’t trust it [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-This appears difficult to reconcile with the fact that, in many countries, most people nowadays receive at least a basic education [@ritchieGlobalEducation2023], and that science education has been found to be the most strongly associated variable with trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018].
-
-But how deep is distrust of science?
-In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
-Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
-That is true not only for basic science questions (studies 1-3), but also for more recent scientific findings that most people are unfamiliar with (study 4).
-In study 1 the acceptance rate after having given a wrong answer was considerably lower (`r exp1_descriptives$conditional_acceptance$false$No$share`) than in subsequent studies (study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`).
-While this could be just sampling variation, it might be that adding explanations and/or sources convinced participants more than merely stating the consensus.
-In general, participants with lower trust in science and who believe more in conspiracy theories tended to both know less about science and accept the scientific consensus less.
-
-Scholars have long interpreted the lack of basic science knowledge as the root cause of distrust in and more generally unfavorable attitudes towards science, an explanation known under the term "deficit-model" [@millerMeasurementCivicScientific1998a; @millerPublicUnderstandingAttitudes2004; @durantPublicUnderstandingScience1989].
-In line with proponents of the deficit-model, we find that people indeed err even on very basic science questions.
-However, we also find that almost everyone accepted the scientific consensus once informed about it.
-Moreover, the same was true for non-basic science questions.
-This calls into question the deficit model: If participants trust science on basic (and non-basic) questions—as suggests immediate acceptance of the scientific consensus—then unfavorable attitudes towards science must have other explanations than a mere lack of knowledge.
-This argument is supported by past research.
-In a seminal meta-analysis, \@allumScienceKnowledgeAttitudes2008 found correlations between science knowledge and general attitudes towards science to be at best small and largely context dependent [\@allumScienceKnowledgeAttitudes2008].
-A more recent report by the National Academies of Sciences, Engineering, and Medicine states that science knowledge is "unlikely to substantially affect attitudes on scientific issues" [@nationalacademiesofsciencesCommunicatingScienceEffectively2017].
-
-How to make sense of the detachment of basic science knowledge from attitudes towards science?
-One explanation might be that most scientific findings are very abstract.
-For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second.
-It is therefore easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
-In other words, people might hold on to most scientific findings as reflective beliefs—beliefs that have little effect on our behaviors [@mercierHowGoodAre2021].
-It is probably when it becomes tangible that science shape's people's attitudes.
-This can happen, for example, when science appears to support conclusions in favor of certain policies or moral positions, as for example in the case of anthropogenic climate change or vaccines.
-
-It is tempting, especially for scientists, to explain mistrust in science with ignorance.
-After all, careful proof of objective truth is what makes science trustworthy—that is, for scientists.
-For most people, what shapes trust in science might not be so different from what shapes trust in other institutions, such as the government, banks or church.
+US participants were asked whether they accepted scientifically consensual answers on basic science questions.We found quasi-universal acceptance of basic science, with an overall rate of acceptance of `r combined_descriptives$mean_acceptance$Yes$share`, which remained very high for participants who declared not trusting science at all (`r combined_descriptives$acceptance_by_trust$'1'$Yes$share` of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (`r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share` of acceptance).
+
+Overall, the present findings support the motivated rejection of science model [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p.
+217).
+A number of participants in our studies rejected basic tenets of science (on evolution, the shape of the earth, etc.).
+However, they still accepted the vast majority of basic scientific knowledge presented to them.
+This suggests that these participants had reasons to reject specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
+These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+
+Our findings are also relevant for the knowledge–attitudes model of trust in science [see, e.g., @bauerWhatCanWe2007].
+The fact that even people with low science knowledge trust most of basic science might be taken as showing the power of even a modicum of science knowledge; on the other hand, it also means that attempts at increasing science knowledge in general might not have much effect on trust in science, by contrast with targeting the specific beliefs people are motivated to reject.
+
+Finally, some of the present results also speak to the alienation model [@gauchatCulturalAuthorityScience2011], and more specifically to the need for epistemic autonomy [@frickerTestimonyEpistemicAutonomy2006].
+Declaring not trusting science, or endorsing conspiracy theories [@harrisConspiracyTheoriesPopulism2023] might reflect a desire to maintain epistemic autonomy and not appear to ‘blindly’ accept epistemic authority.
+The acceptance of basic science facts would be reconciled with this need to appear epistemically autonomous by the suggestion that the acceptance stems from independent evaluation instead of trust (however implausible that might be: it’s not clear how participants could independently verify the ratio of glial cells to neurons, say, and few participants appeared to have engaged in even simple forms of verification).
+
+In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that could be tapped into in science communication.
+Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the ‘consensus gaps’ observed in these domains [on climate change, see, e.g.
+@ranneyClimateChangeConceptual2016].
+
+The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
+We hope that future studies will extend our findings to other populations.
 
@@
 
+```{r child = "appendix_acceptance_knowledge.Rmd"}
+```
+
+\clearpage
+
 ```{r child = "appendix_exp1.Rmd"}

### 2024-09-12 paper.tex
first version within word limit of 4k

@@
 
-Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change (ref) or vaccination (ref) to following recommendations during COVID (Algan, Cohen, Davoine, Foucault, \& Stantcheva, 2021), which suggests that trust in science was the most important predictor of these behaviors).
+Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change (Cologna \& Siegrist, 2020) or vaccination (Lindholt, Jørgensen, Bor, \& Petersen, 2021; Sturgis, Brunton-Smith, \& Jackson, 2021) to following recommendations during COVID (Algan, Cohen, Davoine, Foucault, \& Stantcheva, 2021), which suggests that trust in science was the most important predictor of these behaviors).
 
@@
 What does this apparent rejection of science actually entail?
-Do people who say they do not trust science much, or who believe in conspiracy theories at odds with well-established science, reject most of science?
+Do people who say they do not trust science, or who believe in conspiracy theories at odds with well-established science, reject most of science?
 Or, on the contrary, do they object to a few specific facets of science, while still accepting the overwhelming majority of basic science?
@@
 
-This relative lack of trust in science could mean different things.
-It could mean that some people reject science wholesale, or at least to a significant extent.
-Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (e.g.~creationism), or beliefs that are associated with disliked behaviors (e.g.~vaccination, acting for climate change).
-
-We answer this question by looking at how much Americans accept basic scientific knowledge.
-We pay special attention to two groups of participants: people who say they don't trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).
-
-We briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.
-
-\subsection{Trust in science}\label{trust-in-science}
-
-Across the globe, most people say they trust science at least to some extent.
-In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science (Wellcome Global Monitor, 2018).
-In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants (Wellcome Global Monitor, 2020).
-Across all countries, in 2018, 32\% of participants said they trust science ``a lot'' (41\% in 2020), 45\% trust ``some'' science (39\% in 2020), and only 13\% percent trust science ``not much/ not at all'' (13\% in 2020), while 10\% indicated ``don't know'' (7\% in 2020)\footnote{\url{https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020\#gid=6acd&pid=0}\textgreater{}}.
-
-Trust in science also appears to be relatively stable in the US (Funk \& Kennedy, 2020).
-However, there are some potentially worrying trends: Recently, trust in US has dropped (Brian \& Tyson, 2023), and some evidence suggests increasing polarization of trust in science along political lines (Gauchat, 2012).
-For example, the partisan divide on the question of whether climate change is a major threat to society has been growing in the last decade\footnote{\url{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}} (Kennedy, 2023).
-A recent study suggests that by some measure, only about half of the US population believes in anthropogenic climate change, and that individuals do mostly not seem to change their mind on on the issue (Motta, 2021).
-
-Trust in science is typically measured in three ways: either asking explicitly about general trust in science or scientists, by asking about science attitudes (e.g.~support for public funding of science), or asking questions about specific, typically contentious, science topics (e.g.~vaccines, climate change, GMOs).
-It is unclear to which extent these measures capture trust in basic science.
-To assess this, we can look at the literature on science knowledge.
-
-{[}introduce SK as: that gives us a lower bound of how much people trust basic science{]}
-
-\subsection{Science knowledge}\label{science-knowledge}
-
-Much of the research on public understanding of science measures to which extent people have basic science knowledge (e.g.~``Does the Earth go around the Sun, or does the Sun go around the Earth?'').
-For us, these surveys provide a lower bound to how much people trust basic science: people who give the right answer (and not by chance) supposedly trust science on that question.
-Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge (Miller, 1998, 2004).
+Studies of science knowledge provide relevant evidence, as people who answer science knowledge questions correctly presumably trust scientists in this respect.
+The average level of science knowledge is quite low, even in countries with extensive science education.
 For example, only 55\% in the US (in 2014) and only 46\% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses (Science Literacy et al., 2016).
-A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.
-
-{[}To do: Make a descriptive overview of science knowledge across time/countries.{]}
-
-\subsection{CT and TiS}\label{ct-and-tis}
-
-Recently, the relationship of one group to trust in science (TiS) has been particularly scrutinized: conspiracy theorists (CTists), either defined as people who endorse specific conspiracy theories, or as people who have a conspiracy mindset (Rutjens \& Većkalov, 2022; Vranic, Hromatko, \& Tonković, 2022).
-CTists reject specific science knowledge (related to their CT), and they tend to trust science less on the whole.
-What remains unclear is to which extent CTists (dis)trust basic science.
+However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.
 
@@
 In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts.
-For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted correct answer, and asked them whether they accepted it (measuring their trust in basic science).
+For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted answer, and asked them whether they accepted it (measuring their trust in basic science).
 We also measured participants' trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF)).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF).
 The differences between the four studies are summarized presently, and the methods are detailed below.
@@
 The scientific facts used in Studies 1 to 3 represent long-established and basic knowledge.
-In Study 4 we used more recent much less basic scientific discoveries (e.g.~``How many more glial cells are there in the brain in comparison with neurons, as established in 2016?
-{[}The same amount; Twice as many; Tenth as many{]}''; correct answer: The same amount).
+In Study 4 we used more recent much less basic scientific discoveries (e.g.~``What is the electric charge of the Higgs Boson, as established in 2012?
+{[}1.602176634 × 10-19; 0; 3.2×10-19C{]}''; correct answer: 0, i.e.~electrically neutral).
 
@@
 In Study 1, we simply told participants that they would be provided with the scientifically consensual answer.
-However, for participants to accept this answer, they must not only trust science, but also trust us (that we are presenting them with the actual scientifically consensual answer).
+However, for participants to accept this answer, they must not only trust science, but also trust that we are presenting them with the actual scientifically consensual answer.
 To remove this issue, in Studies 2 to 3, we presented participants with a short explanation of the correct answer, as well as links to three sources per answer (e.g.~Wikipedia, National Geographic or NASA).
-In Study 4, as the topics were more complex, we did not provide an explanation, but we provided two sources per answer.
+In Study 4, as the topics were more complex, we did not provide an explanation, but still provided two sources per answer.
 
@@
 
-For study 2, we restricted our main hypotheses about acceptance to cases in which participants initially provided a wrong answer.
-The rationale was to see if trust in science or conspiracy belief were associated with a change of mind towards the consensus.
-We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM\footnote{Only in study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = 0.06, p = 0.387; study 2: r = 0.16, p = 0.051; study 3: r = 0.04, p = 0.619; study 4: r = 0.15, p = 0.037) and only in study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (study 1: r = -0.14, p = 0.061; study 2: r = -0.22, p = 0.006; study 3: r = -0.04, p = 0.631; study 4: r = -0.05, p = 0.455).}.
-However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting simple, unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
+For Study 2, we restricted our main hypotheses about acceptance to cases in which participants initially provided a wrong answer.
+However, this meant the more participants had initially provided correct answers, the fewer opportunities they had for accepting correct answers.
+We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM\footnote{Only in Study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (Studies 1: r = 0.06, p = 0.387; 2: r = 0.16, p = 0.051; 3: r = 0.04, p = 0.619; 4: r = 0.15, p = 0.037) and only in Study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (Studies 1: r = -0.14, p = 0.061; 2: r = -0.22, p = 0.006; 3: r = -0.04, p = 0.631; 4: r = -0.05, p = 0.455).}.
+However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
 
@@
 Participants then read the following instructions:``We will ask you 10 questions about science.
-After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.'' Next, participants answered a set of 10 basic science questions (in Study 1 they were randomly selected from a pool of 11 questions, as one `trick' question had been added) in random order.
+After each question, we will provide you with the scientifically consensual answer and ask whether you accept it.'' Next, participants answered a set of 10 basic science questions in random order.
 After each question, participants were presented with an answer reflecting the scientific consensus, and asked whether they accepted it.
@@
   ``Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: {[}For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say''{]}.
-We selected particpants who answered''Against''.
+We selected pariticpants who answered''Against''.
 \item
@@
 {[}1 (totally disagree); 2 (disagree); 3 (somewhat disagree); 4 (neither agree nor disagree); 5 (somewhat agree); 6 (agree); 7 (totally agree); rather not say{]}''.
-We select only people who answered `1', `2', or `3'.
+We select only people who answered ``1'', ``2'', or ``3''.
 \end{enumerate}
@@
 Study 4 used 10 more recent scientific discoveries.
-Table \ref{tab:knowledge} shows all questions and their answer options used in the four studies.
+Table \ref{tab:knowledge} shows all questions and their answer options.
 
@@
 5 & Do lasers work by focusing sound waves?
-[Yes; No] & How many more glials cells are there in the brain in comparison with neurons, as established in 2016?
-[The same amount; Twice as many; Tenth as many]\\
+[Yes; No] & How many more glial cells are there in the brain in comparison with neurons, as established in 2016?
+[The same amount; Twice as many; Ten times as many]\\
 \addlinespace
@@
 \addlinespace
-11 & *Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air] & \\
+11 & **Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air] & \\
 \bottomrule
@@
 For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false.
-{[}1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true{]}''.
+{[}1-9; labels: 1 - completely false, 5 - unsure, 9 - completely true{]}''.
 
@@
 We used the four-item conspiracy mentality questionnaire (CMQ) (Bruder, Haffke, Neave, Nouripanah, \& Imhoff, 2013) and the single item conspiracy beliefs scale (SICBS) (Lantian, Muller, Nurra, \& Douglas, 2016).
-Details and comparisons between the scales can be found in the OSM.
+Details and comparisons between the scales can be found in the ESM.
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-1}
+\subsubsection{Trust in science}\label{trust-in-science}
 
@@
 Note that this includes both participants who had previously correctly answered the knowledge question, and participants who changed their mind when presented with the scientific consensus.
-In Studies 3 \& 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: 96 \%; in Study 4: 86 \%).
-
-As shown in Figure \ref{fig:summary-plot}, these very high rates of acceptance hold for: participants who they do not trust science at all (5.7 \% of participants, acceptance rate of 87.3 \%), participants who rank in the top two deciles of the conspiracy thinking scale (28.8 \% of participants, average acceptance rate of 93.2 \%), participants who consider as ``completely true'' (the maximum of the 9-point scale) conspiracy theories stating that the earth is flat (5.8 \% of participants, acceptance rate of 87.2 \%), or that climate change due to fossil emissions is a hoax (12 \% of participants, acceptance rate of 91.7 \%).
-This doesn't merely reflect science knowledge: aggregated across all studies--participants only correctly answered 65.8 \% of the questions (Study 1: 74 \%; Study 2: 79 \%; Study 3: 75 \%; Study 4: 36 \%) before they were provided with the scientifically consensual answer.
-Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
+In Studies 3 and 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: 96 \%; in Study 4: 86 \%).
+
+As shown in Figure \ref{fig:summary-plot}, these very high rates of acceptance hold for: participants who do not trust science at all (4.2 \% of participants, acceptance rate of 87.3 \%), participants who rank in the top two deciles of the conspiracy thinking scale (34.3 \% of participants, acceptance rate of 93.2 \%), participants who consider as ``completely true'' (the maximum of the 9-point scale) conspiracy theories stating that the earth is flat (3.7 \% of participants, acceptance rate of 87.2 \%), or that climate change due to fossil emissions is a hoax (11.4 \% of participants, acceptance rate of 91.7 \%).
+
+Participants in the lowest decile of acceptance still had an average acceptance rate of 67.4 \%.
+Even participants who considered as ``completely true'' that the earth is flat and who said they do ``not trust science at all'' (N = 3) had an average acceptance rate of 86.7 \%.
+
+These high acceptance rates do not merely reflect science knowledge: participants only correctly answered 65.8 \% of the questions (Study 1: 74 \%; Study 2: 79 \%; Study 3: 75 \%; Study 4: 36 \%) before they were provided with the scientifically consensual answer.
+Even the lowest decile in science knowledge which, on average, answered correctly only on 20 \% of the questions, had an average acceptance rate of 91.7 \%.
+
+Did participants who had initially provided a wrong answer change their minds towards the scientific consensus?
 Yes.
 In most cases (study 1: 76.3 \%; study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
-In very few cases, participants changed their mind away from the scientific consensus.
-These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: 1.6 \%; study 2: 0.5 \%; study 3: 0.7 \%; study 4: 5.7 \%).
-
-\begin{verbatim}
-## # A tibble: 1 x 2
-##   mean_acceptance mean_knowledge
-##             <dbl>          <dbl>
-## 1           0.805          0.416
-\end{verbatim}
-
-Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = 0.27, p \textless{} .001; 2: r = 0.30, p \textless{} .001; 3: r = 0.20, p = 0.019; 4: r = 0.16, p = 0.029), but a less consistent relation between trust in science and science knowledge (Studies 1: r = 0.29, p \textless{} .001; 2: r = 0.28, p \textless{} .001; 3: r = 0.14, p = 0.100; study 4: r = 0.10, p = 0.144).
-
-Regarding H2, the results are mixed both for the relation between conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) and acceptance of the scientific consensus (Studies 1: r = -0.33, p \textless{} .001; 2: r = -0.37, p \textless{} .001; 3: r = -0.02, p = 0.788; 4: r = -0.05, p = 0.498) as well as science knowledge (Studies 1: r = -0.38, p \textless{} .001; study 2: r = -0.40, p \textless{} .001; 3: r = -0.16, p = 0.055; 4: r = -0.02, p = 0.742).
+
+Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = 0.27, p \textless{} .001; 2: r = 0.30, p \textless{} .001; 3: r = 0.20, p = 0.019; 4: r = 0.16, p = 0.029), but a less consistent relation between trust in science and science knowledge (Studies 1: r = 0.29, p \textless{} .001; 2: r = 0.28, p \textless{} .001; 3: r = 0.14, p = 0.100; 4: r = 0.10, p = 0.144).
+
+Regarding H2, the results are mixed for the relation of conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) with both acceptance of the scientific consensus (Studies 1: r = -0.33, p \textless{} .001; 2: r = -0.37, p \textless{} .001; 3: r = -0.02, p = 0.788; 4: r = -0.05, p = 0.498) and science knowledge (Studies 1: r = -0.38, p \textless{} .001; 2: r = -0.40, p \textless{} .001; 3: r = -0.16, p = 0.055; 4: r = -0.02, p = 0.742).
 
 Why did participants reject the scientific consensus?
-We got a total of 364 answers (study 2: 35; study 3: 74; study 4: 255) from 167 (study 2: 25; study 3: 47; study 4: 95) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+We collected a total of 364 answers (study 2: 35; study 3: 74; study 4: 255) from 167 (study 2: 25; study 3: 47; study 4: 95) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
 Based on the answers, we created five categories.
 Table \ref{tab:justifications} summarizes these answers by five categories.
-All answers can be read in the OSM.
+All answers can be read in the ESM.
 
@@
 \midrule
-Not convinced & 158.00 & 43.4\% & 70.00\\
-No justification & 103.00 & 28.3\% & 51.00\\
-Personal convictions & 43.00 & 11.8\% & 31.00\\
-Mistake & 42.00 & 11.5\% & 34.00\\
-Religious Beliefs & 18.00 & 4.9\% & 13.00\\
+Not convinced & 158 & 43.4\% & 70\\
+No justification & 103 & 28.3\% & 51\\
+Personal convictions & 43 & 11.8\% & 31\\
+Mistake & 42 & 11.5\% & 34\\
+Religious Beliefs & 18 & 4.9\% & 13\\
 \bottomrule
@@
 Why did participants accept the scientific consensus?
-In Studies 3 and 4--the vaccine skeptic samples--we had asked participants about cases where they agreed with the scientific consensus.
+In Studies 3 and 4--the vaccine hesitant samples--we had asked participants about cases in which they agreed with the scientific consensus.
 A total of 320 (Study 3: 122; Study 4: 198) participants answered this question.
-There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: \texttt{exp3\_descriptives\$acceptance\_by\_reason{[}{[}"independent\ verification"{]}{]}\$share}; Study 4: 47\% ), than there were participants saying it was because they trust scientists (Study 3: 41.8\%; Study 4: 36.4\%).
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: 47.5\%; Study 4: 47\% ), than participants saying it was because they trust scientists (Study 3: 41.8\%; Study 4: 36.4\%)\footnote{10.7\% in Study 3 and 16.7\% in Study 4 answered with ``other'' and gave an open-ended explanation (see ESM).}.
 We asked participants who said they verified independently to explain how they did it (answers can be found in the ESM).
-10.7\% in study 3 and 16.7\% in study 4 answered with other ``other'' and gave an open-ended explanation (see ESM).
-
-In an exploratory analysis, we ran linear regressions to test whether there are differences between these two groups.
-Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = NA; \(\hat{\beta}_{\text{Trust}}\) = 0.19, = 0.330 on a scale from 1 to 4; Study 4: mean = 2.92; \(\hat{\beta}_{\text{Trust}}\) = 0.46, \textless{} .001) than those who said they verified independently (Study 3 mean = NA; Study 4 mean = 2.45).
-We did not find a difference regarding acceptance (Study 3: \(\hat{\beta}_{\text{Acceptance}}\) = 0.01, = 0.523 on a scale from 0 to 1; Study 4: \(\hat{\beta}_{\text{Acceptance}}\) = 0.04, = 0.116).
-In study 4, where we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
-On average, participants clicked only clicked on 1.36 links (out of 20 possible clicks) and there was no difference between the two groups (\(\hat{\beta}_{\text{Clicks}}\) = -1.09, = 0.062).
-
-More detailed results addressing all our pre-registered research questions can be found in the OSM.
+
+In an exploratory analysis, we ran linear regressions to test whether there are differences between participants who said they had trusted science and those who said they had verified the information independently.
+Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = 3; \(\hat{\beta}_{\text{Trust}}\) = 0.19, p = 0.330 on a scale from 1 to 4; Study 4: mean = 2.92; \(\hat{\beta}_{\text{Trust}}\) = 0.46, p \textless{} .001) than those who said they verified independently (Study 3 mean = 2.81; Study 4 mean = 2.45).
+We did not find a difference regarding acceptance (Study 3: \(\hat{\beta}_{\text{Acceptance}}\) = 0.01, p = 0.523 on a scale from 0 to 1; Study 4: \(\hat{\beta}_{\text{Acceptance}}\) = 0.04, p = 0.116) or regarding beliefs in conspiracy theories (Study 3: \(\hat{\beta}_{\text{BCTI}}\) = -0.16, p = 0.694 on a scale from 1 to 9; Study 4: \(\hat{\beta}_{\text{BCTI}}\) = -0.21, p = 0.322).
+We also did not find a difference in time spent on the survey in Study 3 (\(\hat{\beta}_{\text{Time}}\) = -0.02, p = 0.985; median = 7.64 mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average 2 minutes less on the survey (\(\hat{\beta}_{\text{Time}}\) = -2.16, p = 0.018; median = 9.38 mins).\footnote{For these analyses, we excluded outliers who took over 30mins for the survey, which was estimated to take around 10 mins, and for which the median time was 7 minutes.
+As a result, we excluded 1 participant in Study 3 and 4 in Study 4.
+Significance levels remain the same when including these outliers.} In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+On average, participants clicked only on 1.36 links (out of 20 possible clicks) and there was no difference between the two groups (\(\hat{\beta}_{\text{Clicks}}\) = -1.09, p = 0.062).
+
+More detailed results addressing all our pre-registered research questions can be found in the ESM.
 
@@
 
-Survey data from the US suggests a decline of trust in science in recent years (Brian \& Tyson, 2023).
-Across the globe, many people only trust science `some of the time' , and a sizable minority doesn't trust it (Wellcome Global Monitor, 2018, 2020).
-This appears difficult to reconcile with the fact that, in many countries, most people nowadays receive at least a basic education (Ritchie, Samborska, Ahuja, Ortiz-Ospina, \& Roser, 2023), and that science education has been found to be the most strongly associated variable with trust in science (Wellcome Global Monitor, 2018).
-
-But how deep is distrust of science?
-In four studies, we have shown that almost every participant accepts almost all of basic science questions we included.
-Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards.
-That is true not only for basic science questions (studies 1-3), but also for more recent scientific findings that most people are unfamiliar with (study 4).
-In study 1 the acceptance rate after having given a wrong answer was considerably lower (23.7 \%) than in subsequent studies (study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%).
-While this could be just sampling variation, it might be that adding explanations and/or sources convinced participants more than merely stating the consensus.
-In general, participants with lower trust in science and who believe more in conspiracy theories tended to both know less about science and accept the scientific consensus less.
-
-Scholars have long interpreted the lack of basic science knowledge as the root cause of distrust in and more generally unfavorable attitudes towards science, an explanation known under the term ``deficit-model'' (Durant et al., 1989; Miller, 1998, 2004).
-In line with proponents of the deficit-model, we find that people indeed err even on very basic science questions.
-However, we also find that almost everyone accepted the scientific consensus once informed about it.
-Moreover, the same was true for non-basic science questions.
-This calls into question the deficit model: If participants trust science on basic (and non-basic) questions---as suggests immediate acceptance of the scientific consensus---then unfavorable attitudes towards science must have other explanations than a mere lack of knowledge.
-This argument is supported by past research.
-In a seminal meta-analysis, @allumScienceKnowledgeAttitudes2008 found correlations between science knowledge and general attitudes towards science to be at best small and largely context dependent {[}@allumScienceKnowledgeAttitudes2008{]}.
-A more recent report by the National Academies of Sciences, Engineering, and Medicine states that science knowledge is ``unlikely to substantially affect attitudes on scientific issues'' (National Academies of Sciences, 2017).
-
-How to make sense of the detachment of basic science knowledge from attitudes towards science?
-One explanation might be that most scientific findings are very abstract.
-For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second.
-It is therefore easy to simply accept it, all the more if it appears to be consensual within the population of scientists.
-In other words, people might hold on to most scientific findings as reflective beliefs---beliefs that have little effect on our behaviors (Mercier, 2021).
-It is probably when it becomes tangible that science shape's people's attitudes.
-This can happen, for example, when science appears to support conclusions in favor of certain policies or moral positions, as for example in the case of anthropogenic climate change or vaccines.
-
-It is tempting, especially for scientists, to explain mistrust in science with ignorance.
-After all, careful proof of objective truth is what makes science trustworthy---that is, for scientists.
-For most people, what shapes trust in science might not be so different from what shapes trust in other institutions, such as the government, banks or church.
+US participants were asked whether they accepted scientifically consensual answers on basic science questions.We found quasi-universal acceptance of basic science, with an overall rate of acceptance of 95.1 \%, which remained very high for participants who declared not trusting science at all (87.3 \% of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (87.2 \% of acceptance).
+
+Overall, the present findings support the motivated rejection of science model (Lewandowsky \& Oberauer, 2016), in which ``people tend to reject findings that threaten their core beliefs or worldview'' (p.~217).
+A number of participants in our studies rejected basic tenets of science (on evolution, the shape of the earth, etc.).
+However, they still accepted the vast majority of basic scientific knowledge presented to them.
+This suggests that these participants had reasons to reject specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
+These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+
+Our findings are also relevant for the knowledge--attitudes model of trust in science (see, e.g., Bauer, Allum, \& Miller, 2007).
+The fact that even people with low science knowledge trust most of basic science might be taken as showing the power of even a modicum of science knowledge; on the other hand, it also means that attempts at increasing science knowledge in general might not have much effect on trust in science, by contrast with targeting the specific beliefs people are motivated to reject.
+
+Finally, some of the present results also speak to the alienation model (Gauchat, 2011), and more specifically to the need for epistemic autonomy (Fricker, 2006).
+Declaring not trusting science, or endorsing conspiracy theories (Harris, 2023) might reflect a desire to maintain epistemic autonomy and not appear to `blindly' accept epistemic authority.
+The acceptance of basic science facts would be reconciled with this need to appear epistemically autonomous by the suggestion that the acceptance stems from independent evaluation instead of trust (however implausible that might be: it's not clear how participants could independently verify the ratio of glial cells to neurons, say, and few participants appeared to have engaged in even simple forms of verification).
+
+In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that could be tapped into in science communication.
+Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the `consensus gaps' observed in these domains (on climate change, see, e.g.
+Ranney \& Clark, 2016).
+
+The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
+We hope that future studies will extend our findings to other populations.
 
@@
 
+\bibitem[\citeproctext]{ref-bauerWhatCanWe2007}
+Bauer, M.
+W., Allum, N., \& Miller, S.
+(2007).
+What can we learn from 25 years of PUS survey research?
+Liberating and expanding the agenda.
+\emph{Public Understanding of Science}, \emph{16}(1), 79--95.
+\url{https://doi.org/10.1177/0963662506071287}
+
 \bibitem[\citeproctext]{ref-brianAmericansTrustScientists2023}
@@
 
+\bibitem[\citeproctext]{ref-colognaRoleTrustClimate2020}
+Cologna, V., \& Siegrist, M.
+(2020).
+The role of trust for climate change mitigation and adaptation behaviour: A meta-analysis.
+\emph{Journal of Environmental Psychology}, \emph{69}, 101428.
+\url{https://doi.org/10.1016/j.jenvp.2020.101428}
+
 \bibitem[\citeproctext]{ref-durantPublicUnderstandingScience1989}
 Durant, J. R., Evans, G. A., \& Thomas, G. P. (1989). The public understanding of science. \emph{Nature}, \emph{340}(6228), 11--14. \url{https://doi.org/10.1038/340011a0}
+
+\bibitem[\citeproctext]{ref-frickerTestimonyEpistemicAutonomy2006}
+Fricker, E. (2006). \emph{Testimony and Epistemic Autonomy} (J. Lackey \& E. Sosa, Eds.). Oxford University PressOxford. \url{https://doi.org/10.1093/acprof:oso/9780199276011.003.0011}
 
@@
 \emph{5 key findings about public trust in scientists in the u.s.} Retrieved from \url{https://www.pewresearch.org/fact-tank/2019/08/05/5-key-findings-about-public-trust-in-scientists-in-the-u-s/}
-
-\bibitem[\citeproctext]{ref-funkPublicConfidenceScientists2020}
-Funk, C., \& Kennedy, B. (2020). \emph{Public confidence in scientists has remained stable for decades}.
 
@@
 
-\bibitem[\citeproctext]{ref-kennedyWhatDataSays2023}
-Kennedy, A.
-T.
-C.
-F.
-and B.
-(2023).
-\emph{What the data says about americans{'} views of climate change}.
-Retrieved from \url{https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/}
+\bibitem[\citeproctext]{ref-harrisConspiracyTheoriesPopulism2023}
+Harris, K. R. (2023). Conspiracy Theories, Populism, and Epistemic Autonomy. \emph{Journal of the American Philosophical Association}, \emph{9}(1), 21--36. \url{https://doi.org/10.1017/apa.2021.44}
 
@@
 
+\bibitem[\citeproctext]{ref-lewandowskyMotivatedRejectionScience2016}
+Lewandowsky, S., \& Oberauer, K. (2016). Motivated Rejection of Science. \emph{Current Directions in Psychological Science}, \emph{25}(4), 217--222. \url{https://doi.org/10.1177/0963721416654436}
+
 \bibitem[\citeproctext]{ref-liPolarizationPublicTrust2022}
@@
 
-\bibitem[\citeproctext]{ref-mercierHowGoodAre2021}
-Mercier, H. (2021). How Good Are We At Evaluating Communicated Information? \emph{Royal Institute of Philosophy Supplement}, \emph{89}, 257--272. \url{https://doi.org/10.1017/S1358246121000096}
+\bibitem[\citeproctext]{ref-lindholtPublicAcceptanceCOVID192021}
+Lindholt, M.
+F., Jørgensen, F., Bor, A., \& Petersen, M.
+B.
+(2021).
+Public acceptance of COVID-19 vaccines: cross-national evidence on levels and individual-level predictors using observational data.
+\emph{BMJ Open}, \emph{11}(6), e048172.
+\url{https://doi.org/10.1136/bmjopen-2020-048172}
 
@@
 Miller, J. D. (1998). The measurement of civic scientific literacy. \emph{Public Understanding of Science}, \emph{7}(3), 203--223. \url{https://doi.org/10.1088/0963-6625/7/3/001}
-
-\bibitem[\citeproctext]{ref-millerPublicUnderstandingAttitudes2004}
-Miller, J.
-D.
-(2004).
-Public Understanding of, and Attitudes toward, Scientific Research: What We Know and What We Need to Know.
-\emph{Public Understanding of Science}, \emph{13}(3), 273--294.
-\url{https://doi.org/10.1177/0963662504044908}
-
-\bibitem[\citeproctext]{ref-mottaChangingMindsChanging2021}
-Motta, M.
-(2021).
-Changing Minds or Changing Samples?
-Disentangling Microlevel Stability and Macrolevel Growth in Anthropogenic Climate Change Beliefs.
-\emph{International Journal of Public Opinion Research}, \emph{33}(2), 477--489.
-\url{https://doi.org/10.1093/ijpor/edaa020}
-
-\bibitem[\citeproctext]{ref-nationalacademiesofsciencesCommunicatingScienceEffectively2017}
-National Academies of Sciences, E. (2017). \emph{Communicating science effectively: A research agenda}. National Academies Press.
 
@@
 
-\bibitem[\citeproctext]{ref-ritchieGlobalEducation2023}
-Ritchie, H., Samborska, V., Ahuja, N., Ortiz-Ospina, E., \& Roser, M. (2023). Global education. \emph{Our World in Data}. Retrieved from \url{https://ourworldindata.org/global-education}
-
-\bibitem[\citeproctext]{ref-rutjensConspiracyBeliefsScience2022}
-Rutjens, B. T., \& Većkalov, B. (2022). Conspiracy beliefs and science rejection. \emph{Current Opinion in Psychology}, \emph{46}, 101392. \url{https://doi.org/10.1016/j.copsyc.2022.101392}
+\bibitem[\citeproctext]{ref-ranneyClimateChangeConceptual2016}
+Ranney, M.
+A., \& Clark, D.
+(2016).
+Climate Change Conceptual Change: Scientific Information Can Transform Attitudes.
+\emph{Topics in Cognitive Science}, \emph{8}(1), 49--75.
+\url{https://doi.org/10.1111/tops.12187}
 
@@
 
+\bibitem[\citeproctext]{ref-sturgisTrustScienceSocial2021}
+Sturgis, P., Brunton-Smith, I., \& Jackson, J.
+(2021).
+Trust in science, social consensus and vaccine confidence.
+\emph{Nature Human Behaviour}, \emph{5}(11), 1528--1534.
+\url{https://doi.org/10.1038/s41562-021-01115-7}
+
 \bibitem[\citeproctext]{ref-vanstekelenburgScientificConsensusCommunicationContested2022}
@@
 
-\bibitem[\citeproctext]{ref-vranicDidMyOwn2022}
-Vranic, A., Hromatko, I., \& Tonković, M.
-(2022).
-{``}I did my own research{''}: Overconfidence, (dis)trust in science, and endorsement of conspiracy theories.
-\emph{Frontiers in Psychology}, \emph{13}.
-\url{https://doi.org/10.3389/fpsyg.2022.931865}
-
 \bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2018}
@@
 \appendix
+
+\section{Acceptance conditional on knowledge}\label{acceptance-knowledge}
+
+Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
+Yes.
+In most cases (study 1: 76.3 \%; study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In very few cases, participants changed their mind away from the scientific consensus.
+These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: 1.6 \%; study 2: 0.5 \%; study 3: 0.7 \%; study 4: 5.7 \%).
+Figure \ref{fig:acceptance-knowledge} shows acceptance rates conditional on whether the initial answer to the questions was correct or not.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/acceptance-knowledge.pdf}
+\caption{\label{fig:acceptance-knowledge}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+\clearpage
 
@@
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:unnamed-chunk-25}Reasons for consensus rejection}\\
+\caption{\label{tab:unnamed-chunk-27}Reasons for consensus rejection}\\
 \toprule
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-9}
+\subsubsection{Trust in science}\label{trust-in-science-8}
 
@@
 
-\subsection{Trust in science}\label{trust-in-science-10}
+\subsection{Trust in science}\label{trust-in-science-9}
 

### 2024-09-19 appendix_exp1.Rmd
submission public understanding of science

@@
 
-Beside Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 which report in the main study, we also assessed two other consipiracy thinking measures:
+In addition to the Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 which report in the main study, we also assessed two other consipiracy thinking measures: The conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013 and the Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 (Table \@ref(tab:overview-conspiracy-measures)).
 
-1. The conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013:
-I think that . . .
+```{r overview-conspiracy-measures}
+# Create the data for the table
+conspiracy_measures <- data.frame(
+  Measure = c(
+    "Belief in Conspiracy Theory Inventory (BCTI)", 
+    "Conspiracy Mentality Questionnaire (CMQ)", 
+    "", "", "", "", "",
+    "Single Item Conspiracy Beliefs Scale (SICBS)"
+  ),
+  Description = c(
+    "By Pennycook et al., 2022, presented in the main article.", 
+    "By Bruder et al., 2013. Respondents are asked how much they agree with the following statements:", 
+    "- Many important things happen in the world, which the public is never informed about.",
+    "- Politicians usually do not tell us the true motives for their decisions.",
+    "- Government agencies closely monitor all citizens.",
+    "- Events which seem unrelated are often the result of secret activities.",
+    "- There are secret organizations that greatly influence political decisions.",
+    "By Lantian et al., 2016. The item is: 'I think that the official version of events given by the authorities very often hides the truth.'"
+  ),
+  Scale = c(
+    "1 - 9 (1 = Completely false, 5 = Unsure, 9 = Completely true)", 
+    "0% - 100% (0 = certainly not, 100 = certain)",
+    "", "", "", "", "",
+    "1 - 9 (1 = Completely false, 5 = Unsure, 9 = Completely true)"
+  )
+)
 
--   ... many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
--   ... government agencies closely monitor all citizens.
--   ... events which superficially seem to lack a connection are often the result of secret activities.
--   ... there are secret organizations that greatly influence political decisions.
+# # Create the table
+kbl(conspiracy_measures, booktabs = T, longtable = T,
+    caption = "Conspiracy measures", full_width = T) %>%
+  kable_styling(font_size = 8) %>% 
+  column_spec(1, width = "10em") %>%
+  column_spec(2, width = "20em") %>%
+  column_spec(3, width = "10em")
 
-[0% - 100%; 0 = certainly not, 100 = certain]
-
-2.  The Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 :
-
--   I think that the official version of the events given by the authorities very often hides the truth. [1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true]
-
-### Trust in science
-
-We rely on three items
-
-1. How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]
-
-2. In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]
-
-3. How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]
+# Use apa_table to format the table
+# apa_table(
+#   conspiracy_measures,
+#   caption = "Overview of Conspiracy Thinking Measures",
+#   align = c("m{4cm}", "m{10cm}", "m{4cm}")
+# )
+```
 
@@
 
-Fig.
-\@ref(fig:exp1-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
 In most cases (`r exp1_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having given the wrong answer to a question.
@@
 These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\@ref(fig:exp1-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+\@ref(fig:exp1-plot-overview), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
 
@@
 
-In Appendix \@ref(exp1), we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \@ref(exp1) also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+Below, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
+We also provide more descriptive statistics, such as knowledge and acceptance by science questions.
 
@@
 
-## By-question variation
-
-```{r}
-# calculate share
-plot_data <- exp1_long %>% 
-  group_by(subject) %>% 
-  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
-            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
-```
-
-### Knowledge
-
-(ref:exp1-questions-knowledge) Distribution of correct answers by question.
-
-```{r exp1-questions-knowledge, fig.cap="(ref:exp1-questions-knowledge)"}
-# plot knowledge
-ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of correct responses") +
-  coord_flip()
-```
-
-### Acceptance
-
-(ref:exp1-questions-acceptance) Distribution of consensus acceptance by question.
-
-```{r exp1-questions-acceptance, fig.cap="(ref:exp1-questions-acceptance)"}
-# plot acceptance
-ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of acceptance") +
-  coord_flip()
-```
-
-## Trust in science
-
-(ref:exp1-trust-scientists) Distribution of trust in scientists.
-
-```{r exp1-trust-scientists, fig.cap="(ref:exp1-trust-scientists)"}
-# calculate share
-plot_data <- exp1_wide %>% 
-  group_by(wgm_scientists) %>% 
-  summarize(n = n()) %>% 
-  ungroup() %>% 
-  mutate(share = n/sum(n)) %>% 
-  rounded_numbers()
-
-# plot knowledge
-ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Trust in science", y = "Share of participants") +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) 
-```
-
-## Conspiracy thinking
-
-### Distribution
-
-(ref:exp1-conspiracy-distribution) Distribution of conspiracy thinking.
-
-```{r exp1-conspiracy-distribution, fig.cap="(ref:exp1-conspiracy-distribution)"}
-# plot knowledge
-ggplot(exp1_wide, aes(x = BCTI_avg)) +
-  geom_histogram() +
-  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
-```
-
-### By-item variation
-
-(ref:exp1-conspiracy-items) Distribution of conspiracy thinking by item.
-
-```{r exp1-conspiracy-items, fig.cap="(ref:exp1-conspiracy-items)"}
-BCTI_data <- exp1_long %>% 
-  # bring to long format
-  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
-               names_to = "BCTI_item", values_to = "score") %>% 
-  select(id, c(starts_with("BCTI"), score)) %>% 
-  group_by(BCTI_item) %>% 
-  summarize(avg_score = mean(score)) %>% 
-  ungroup()
-
-ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
-    geom_pointrange(aes(ymin = 1, ymax = avg_score),
-                  # Make the lines a little thicker and the dots a little bigger
-                  fatten = 2, size = 1) +
-  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
-  labs(x = "Conspiracy item", y = "Average score") +
-  coord_cartesian(ylim = c(1, 9)) +
-  coord_flip()
-```
-
-## Open ended comments
-```{r}
-# levels(as.factor(exp1_wide$comments))
-table <- exp1_wide %>% 
-  drop_na(comments) %>% 
-  select(id, comments)
-
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "40em") 
-```
-
-## Additional plots
-
-(ref:exp1-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
-
-```{r exp1-conditional-acceptance, fig.cap="(ref:exp1-conditional-acceptance)"}
-plot_data <- exp1_long %>% 
-  group_by(knowledge, acceptance) %>% 
-  count() %>% 
-  group_by(knowledge) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-absolute_numbers <- plot_data %>% 
-  group_by(knowledge) %>% 
-  summarise(n = sum(n))
-
-ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  plot_theme +
-  geom_bracket(data = absolute_numbers,
-               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
-                   xmax = as.numeric(factor(knowledge)) + 0.4, 
-                   y.position = 1.05, label =  paste0("n = ", n)),
-               inherit.aes = FALSE,
-               step.increase = 0)
-```
-
-(ref:exp1-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus
+(ref:exp1-plot-overview) Summary plot for Study 1.
+**A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus
 
@@
        y = "Trust in science", 
-       fill = "N") +
-  theme(legend.position = "top")
+       fill = "N")
 
@@
 ```
+
+## Open ended comments
+
+```{r}
+# levels(as.factor(exp1_wide$comments))
+table <- exp1_wide %>% 
+  drop_na(comments) %>% 
+  select(id, comments)
+
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Comments of participants on Study 1") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2, width = "40em") 
+```

### 2024-09-19 appendix_exp2.Rmd
submission public understanding of science

@@
 To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
-We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \@ref(exp1)).
+We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong.
 
@@
 Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\@ref(fig:exp2-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-In Appendix \@ref(exp2) we show that these results hold for our alternative measures of trust and conspiracy thinking.
+These correlations are relatively weak, which might be partly due to ceiling effects: (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+Below, we show that these results hold for our alternative measures of trust and conspiracy thinking.
 We also include more descriptive statistics, such as knowledge and acceptance by science questions.
@@
 Regarding RQ1, participants answered on average `r exp2_descriptives$means$knowledge_mean` (sd = `r exp2_descriptives$means$knowledge_sd`) of the questions correctly, and accepted the scientific consensus on average for `r exp2_descriptives$means$acceptance_mean` (sd = `r exp2_descriptives$means$acceptance_sd`) of the questions.
-Fig.
-\@ref(fig:exp2-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
 In response to RQ2, in most cases (`r exp2_descriptives$conditional_acceptance$false$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
@@
 Table \@ref(tab:exp2-justifications) summarizes these answers by five categories.
-All answers can be read in Appendix \@ref(exp2).
+All answers can be read in below.
 
@@
 
-## By-question variation
+(ref:exp2-plot-overview) Summary plot for Study 2.
+**A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
 
-```{r}
-# calculate share
-plot_data <- exp2_long %>% 
-  group_by(subject) %>% 
-  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
-            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
-```
-
-### Knowledge
-
-(ref:exp2-questions-knowledge) Distribution of correct answers by question.
-
-```{r exp2-questions-knowledge, fig.cap="(ref:exp2-questions-knowledge)"}
-# plot knowledge
-ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of correct responses") +
-  coord_flip()
-```
-
-### Acceptance
-
-(ref:exp2-questions-acceptance) Distribution of consensus acceptance by question.
-
-```{r exp2-questions-acceptance, fig.cap="(ref:exp2-questions-acceptance)"}
-# plot acceptance
-ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of acceptance") +
-  coord_flip()
-```
-
-## Trust in science
-
-(ref:exp2-trust-scientists) Distribution of trust in scientists.
-
-```{r exp2-trust-scientists, fig.cap="(ref:exp2-trust-scientists)"}
-# calculate share
-plot_data <- exp2_wide %>% 
-  group_by(wgm_scientists) %>% 
-  summarize(n = n()) %>% 
-  ungroup() %>% 
-  mutate(share = n/sum(n)) %>% 
-  rounded_numbers()
-
-# plot knowledge
-ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Trust in science", y = "Share of participants") +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) 
-```
-
-## Conspiracy thinking
-
-### Distribution
-
-(ref:exp2-conspiracy-distribution) Distribution of conspiracy thinking.
-
-```{r exp2-conspiracy-distribution, fig.cap="(ref:exp2-conspiracy-distribution)"}
-# plot knowledge
-ggplot(exp2_wide, aes(x = BCTI_avg)) +
-  geom_histogram() +
-  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
-```
-
-### By-item variation
-
-(ref:exp2-conspiracy-items) Distribution of conspiracy thinking by item.
-
-```{r exp2-conspiracy-items, fig.cap="(ref:exp2-conspiracy-items)"}
-BCTI_data <- exp2_long %>% 
-  # bring to long format
-  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
-               names_to = "BCTI_item", values_to = "score") %>% 
-  select(id, c(starts_with("BCTI"), score)) %>% 
-  group_by(BCTI_item) %>% 
-  summarize(avg_score = mean(score, na.rm=TRUE)) %>% 
-  ungroup()
-
-ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
-    geom_pointrange(aes(ymin = 1, ymax = avg_score),
-                  # Make the lines a little thicker and the dots a little bigger
-                  fatten = 2, size = 1) +
-  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
-  labs(x = "Conspiracy item", y = "Average score") +
-  coord_cartesian(ylim = c(1, 9)) +
-  coord_flip()
-```
-
-## Reasons for consensus rejection
-
-### By category
-
-[To do]
-
-### All raw answers
-
-```{r exp2-reasons-rejection}
-# store all variables
-reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
-               "orbit", "diamonds", "speed", "salt", "water")
-
-table <- exp2_wide %>% 
-  select(id, all_of(reasons_rejection)) %>% 
-  pivot_longer(all_of(reasons_rejection), 
-               names_to = "question",
-               values_to = "answer") %>% 
-  drop_na(answer) %>% 
-  arrange(id)
-
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2) %>%
-  column_spec(3, width = "30em")
-
-```
-
-## Additional plots
-
-(ref:exp2-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
-
-```{r exp2-conditional-acceptance, fig.cap="(ref:exp2-conditional-acceptance)"}
-plot_data <- exp2_long %>% 
-  group_by(knowledge, acceptance) %>% 
-  count() %>% 
-  group_by(knowledge) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-absolute_numbers <- plot_data %>% 
-  group_by(knowledge) %>% 
-  summarise(n = sum(n))
-
-ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  plot_theme  +
-  geom_bracket(data = absolute_numbers,
-               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
-                   xmax = as.numeric(factor(knowledge)) + 0.4, 
-                   y.position = 1.05, label =  paste0("n = ", n)),
-               inherit.aes = FALSE,
-               step.increase = 0)
-```
-
-(ref:exp2-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
 ```{r exp2-plot-overview, fig.cap="(ref:exp2-plot-overview)", fig.height= 12, fig.width=12}
@@
 ```
+
+## Reasons for consensus rejection
+
+```{r exp2-reasons-rejection}
+# store all variables
+reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
+               "orbit", "diamonds", "speed", "salt", "water")
+
+table <- exp2_wide %>% 
+  select(id, all_of(reasons_rejection)) %>% 
+  pivot_longer(all_of(reasons_rejection), 
+               names_to = "question",
+               values_to = "answer") %>% 
+  drop_na(answer) %>% 
+  arrange(id)
+
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Reasons for consensus rejection (raw answers)") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2) %>%
+  column_spec(3, width = "30em")
+
+```

### 2024-09-19 appendix_exp3.Rmd
submission public understanding of science

@@
 
-Table \@ref(tab:knowledge-overview) shows all questions, their scientifically consensual answer, and their source.
+Table \@ref(tab:exp3-knowledge-items) shows all questions, their scientifically consensual answer, and their source.
 All but two questions were selected from existing science knowledge questionnaires.
@@
 
-```{r knowledge-overview}
+```{r exp3-knowledge-items}
 items <- read_csv("exp_3/materials/overview_questions.csv") %>% 
@@
 
-Fig.
-\@ref(fig:exp3-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
 In most cases (`r exp3_descriptives$conditional_acceptance$false$acceptance_initial$Yes$share`), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
@@
 We find no statistically significant correlation between science knowledge and trust in science (r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`).
-Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
-\@ref(fig:exp3-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+Again, these findings might be partly due to ceiling effects: (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
 We find no statistically significant correlation between conspiracy thinking and science knowledge (r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`), and between conspiracy thinking and acceptance of scientific consensus (r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`).
@@
 
-All answers can be read in the appendix.
+All answers can be read in the below.
 
@@
 
-## By-question variation
-
-```{r}
-# calculate share
-plot_data <- exp3_long %>% 
-  group_by(subject) %>% 
-  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
-            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
-```
-
-### Knowledge
-
-(ref:exp3-questions-knowledge) Distribution of correct answers by question.
-
-```{r exp3-questions-knowledge, fig.cap="(ref:exp3-questions-knowledge)"}
-# plot knowledge
-ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of correct responses") +
-  coord_flip()
-```
-
-(ref:exp3-questions-acceptance) Distribution of consensus acceptance by question.
-
-```{r exp3-questions-acceptance, fig.cap="(ref:exp3-questions-acceptance)"}
-# plot acceptance
-ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of acceptance") +
-  coord_flip()
-```
-
-## Trust in science
-
-(ref:exp3-trust-scientists) Distribution of trust in scientists.
-
-```{r exp3-trust-scientists, fig.cap="(ref:exp3-trust-scientists)"}
-# calculate share
-plot_data <- exp3_wide %>% 
-  group_by(wgm_scientists) %>% 
-  summarize(n = n()) %>% 
-  ungroup() %>% 
-  mutate(share = n/sum(n)) %>% 
-  rounded_numbers()
-
-# plot knowledge
-ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Trust in science", y = "Share of participants") +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) 
-```
-
-## Conspiracy thinking
-
-(ref:exp3-conspiracy-distribution) Distribution of conspiracy thinking.
-
-```{r exp3-conspiracy-distribution, fig.cap="(ref:exp3-conspiracy-distribution)"}
-# plot knowledge
-ggplot(exp3_wide, aes(x = BCTI_avg)) +
-  geom_histogram() +
-  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
-```
-
-### By-item variation
-
-(ref:exp3-conspiracy-items) Distribution of conspiracy thinking by item.
-
-```{r exp3-conspiracy-items, fig.cap="(ref:exp3-conspiracy-items)"}
-BCTI_data <- exp3_long %>% 
-  # bring to long format
-  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
-               names_to = "BCTI_item", values_to = "score") %>% 
-  select(id, c(starts_with("BCTI"), score)) %>% 
-  group_by(BCTI_item) %>% 
-  summarize(avg_score = mean(score, na.rm=TRUE)) %>% 
-  ungroup()
-
-ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
-    geom_pointrange(aes(ymin = 1, ymax = avg_score),
-                  # Make the lines a little thicker and the dots a little bigger
-                  fatten = 2, size = 1) +
-  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
-  labs(x = "Conspiracy item", y = "Average score") +
-  coord_cartesian(ylim = c(1, 9)) +
-  coord_flip()
-```
-
-## Reasons for consensus rejection
-
-### All raw answers
-
-```{r exp3-reasons-rejection}
-# store all variables
-reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
-               "orbit", "diamonds", "speed", "salt", "water")
-
-table <- exp3_wide %>% 
-  select(id, all_of(reasons_rejection)) %>% 
-  pivot_longer(all_of(reasons_rejection), 
-               names_to = "question",
-               values_to = "answer") %>% 
-  drop_na(answer) %>% 
-  arrange(id)
-
-# apa_table(table,
-#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
-#           longtable = TRUE
-#           )
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2) %>%
-  column_spec(3, width = "30em")
-
-```
-
-## Why do people accept the scientific consensus:
-
-### Participants who answered "other"
-
-```{r exp3-other-reasons-acceptance}
-# store all variables
-table <- exp3_wide %>% 
-  select(id, other_reason_agreement) %>% 
-  drop_na(other_reason_agreement) %>% 
-  arrange(id)
-
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "30em")
-```
-
-## How did participants independently verifiy?
-
-### Participants who answered "other"
-
-```{r exp3-independent-verification}
-# store all variables
-table <- exp3_wide %>% 
-  select(id, reason_followup) %>% 
-  drop_na(reason_followup) %>% 
-  arrange(id)
-
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "30em")
-```
-
-## Additional plots
-
-(ref:exp3-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
-
-```{r exp3-conditional-acceptance, fig.cap="(ref:exp3-conditional-acceptance)"}
-plot_data <- exp3_long %>% 
-  pivot_longer(c(acceptance, acceptance_initial), 
-               names_to = "measure", 
-               values_to = "acceptance") %>% 
-  # give nicer names
-  mutate(measure = case_when(measure == "acceptance" ~ "Acceptance including revisions", 
-                             measure == "acceptance_initial" ~ "Initial acceptance"), 
-         # make a factor
-         measure = factor(measure, levels = unique(measure)[c(2, 1)])
-         ) %>% 
-  group_by(measure, knowledge, acceptance) %>% 
-  count() %>% 
-  group_by(measure, knowledge) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-absolute_numbers <- plot_data %>% 
-  group_by(measure, knowledge) %>% 
-  summarise(n = sum(n))
-
-ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  plot_theme +
-  geom_bracket(data = absolute_numbers,
-               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
-                   xmax = as.numeric(factor(knowledge)) + 0.4, 
-                   y.position = 1.05, label =  paste0("n = ", n)),
-               inherit.aes = FALSE,
-               step.increase = 0) +
-  facet_wrap(~measure)
-```
-
-(ref:exp3-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
+(ref:exp3-plot-overview) Summary plot for Study 3.
+**A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
 
@@
 ```
+
+## Reasons for consensus rejection
+
+```{r exp3-reasons-rejection}
+# store all variables
+reasons_rejection <- c("electrons", "antibiotics", "continents", "sex", "lasers", 
+               "orbit", "diamonds", "speed", "salt", "water")
+
+table <- exp3_wide %>% 
+  select(id, all_of(reasons_rejection)) %>% 
+  pivot_longer(all_of(reasons_rejection), 
+               names_to = "question",
+               values_to = "answer") %>% 
+  drop_na(answer) %>% 
+  arrange(id)
+
+# apa_table(table,
+#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
+#           longtable = TRUE
+#           )
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Reasons for consensus rejection (raw answers)") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2) %>%
+  column_spec(3, width = "30em")
+
+```
+
+## Reasons for consensus acceptance
+
+```{r exp3-other-reasons-acceptance}
+# store all variables
+table <- exp3_wide %>% 
+  select(id, other_reason_agreement) %>% 
+  drop_na(other_reason_agreement) %>% 
+  arrange(id)
+
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Reasons for consensus acceptance (participants who answered 'other')") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2, width = "30em")
+```
+
+## How did participants independently verifiy?
+
+```{r exp3-independent-verification}
+# store all variables
+table <- exp3_wide %>% 
+  select(id, reason_followup) %>% 
+  drop_na(reason_followup) %>% 
+  arrange(id)
+
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Answers from participants saying they trusted the scientific consens because they verified independently on how they did so.") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2, width = "30em")
+```

### 2024-09-19 appendix_exp4.Rmd
submission public understanding of science

@@
 
-Fig.
-\@ref(fig:exp4-conditional-acceptance) illustrates the relationship between knowledge and acceptance.
 In most cases (`r exp4_descriptives$conditional_acceptance$false$acceptance_initial$Yes$share`), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
@@
 We find no statistically significant correlation between science knowledge and trust in science (r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`), but a small positive correlation between acceptance of scientific consensus and trust in science (r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`).
-Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
-\@ref(fig:exp4-plot), (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+Again, these findings might be partly due to ceiling effects: (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
 We do not find a statistically significant correlation between neither conspiracy thinking and science knowledge (r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`), nor between conspiracy thinking and acceptance of scientific consensus (r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`).
@@
 
-## By-question variation
-
-```{r}
-# calculate share
-plot_data <- exp4_long %>% 
-  group_by(subject) %>% 
-  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
-            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
-```
-
-### Knowledge
-
-(ref:exp4-questions-knowledge) Distribution of correct answers by question.
-
-```{r exp4-questions-knowledge, fig.cap="(ref:exp4-questions-knowledge)"}
-# plot knowledge
-ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of correct responses") +
-  coord_flip()
-```
-
-(ref:exp4-questions-acceptance) Distribution of consensus acceptance by question.
-
-```{r exp4-questions-acceptance, fig.cap="(ref:exp4-questions-acceptance)"}
-# plot acceptance
-ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Question", y = "Share of acceptance") +
-  coord_flip()
-```
-
-## Trust in science
-
-(ref:exp4-trust-scientists) Distribution of trust in scientists.
-
-```{r exp4-trust-scientists, fig.cap="(ref:exp4-trust-scientists)"}
-# calculate share
-plot_data <- exp4_wide %>% 
-  group_by(wgm_scientists) %>% 
-  summarize(n = n()) %>% 
-  ungroup() %>% 
-  mutate(share = n/sum(n)) %>% 
-  rounded_numbers()
-
-# plot knowledge
-ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
-  geom_col() +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  labs(x = "Trust in science", y = "Share of participants") +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) 
-```
-
-## Conspiracy thinking
-
-(ref:exp4-conspiracy-distribution) Distribution of conspiracy thinking.
-
-```{r exp4-conspiracy-distribution, fig.cap="(ref:exp4-conspiracy-distribution)"}
-# plot knowledge
-ggplot(exp4_wide, aes(x = BCTI_avg)) +
-  geom_histogram() +
-  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
-```
-
-### By-item variation
-
-(ref:exp4-conspiracy-items) Distribution of conspiracy thinking by item.
-
-```{r exp4-conspiracy-items, fig.cap="(ref:exp4-conspiracy-items)"}
-BCTI_data <- exp4_long %>% 
-  # bring to long format
-  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
-               names_to = "BCTI_item", values_to = "score") %>% 
-  select(id, c(starts_with("BCTI"), score)) %>% 
-  group_by(BCTI_item) %>% 
-  summarize(avg_score = mean(score, na.rm=TRUE)) %>% 
-  ungroup()
-
-ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
-    geom_pointrange(aes(ymin = 1, ymax = avg_score),
-                  # Make the lines a little thicker and the dots a little bigger
-                  fatten = 2, size = 1) +
-  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
-  labs(x = "Conspiracy item", y = "Average score") +
-  coord_cartesian(ylim = c(1, 9)) +
-  coord_flip()
-```
-
-## Reasons for consensus rejection
-
-### All raw answers
-
-```{r exp4-reasons-rejection}
-# store all variables
-reasons_rejection <- unique(exp4_long$subject)
-
-table <- exp4_wide %>% 
-  select(id, all_of(reasons_rejection)) %>% 
-  pivot_longer(all_of(reasons_rejection), 
-               names_to = "question",
-               values_to = "answer") %>% 
-  drop_na(answer) %>% 
-  arrange(id)
-
-# apa_table(table,
-#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
-#           longtable = TRUE
-#           )
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2) %>%
-  column_spec(3, width = "30em")
-
-```
-
-## Why do people accept the scientific consensus:
-
-### Participants who answered "other"
-
-```{r exp4-other-reasons-acceptance}
-# store all variables
-table <- exp4_wide %>% 
-  select(id, other_reason_agreement) %>% 
-  drop_na(other_reason_agreement) %>% 
-  arrange(id)
-
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "30em")
-```
-
-## How did participants independently verifiy?
-
-### Participants who answered "other"
-
-```{r exp4-independent-verification}
-# store all variables
-table <- exp4_wide %>% 
-  select(id, reason_followup) %>% 
-  drop_na(reason_followup) %>% 
-  arrange(id)
-
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "30em")
-```
-
-## Additional plots
-
-(ref:exp4-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.
-
-```{r exp4-conditional-acceptance, fig.cap="(ref:exp4-conditional-acceptance)"}
-plot_data <- exp4_long %>% 
-  pivot_longer(c(acceptance, acceptance_initial), 
-               names_to = "measure", 
-               values_to = "acceptance") %>% 
-  # give nicer names
-  mutate(measure = case_when(measure == "acceptance" ~ "Acceptance including revisions", 
-                             measure == "acceptance_initial" ~ "Initial acceptance"), 
-         # make a factor
-         measure = factor(measure, levels = unique(measure)[c(2, 1)])
-         ) %>% 
-  group_by(measure, knowledge, acceptance) %>% 
-  count() %>% 
-  group_by(measure, knowledge) %>% 
-  mutate (share = n/sum(n)) %>% 
-  mutate_if(is.numeric, round, digits = 3)
-
-absolute_numbers <- plot_data %>% 
-  group_by(measure, knowledge) %>% 
-  summarise(n = sum(n))
-
-ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
-  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
-            position = position_dodge(width = 0.9),
-            vjust = -0.5, size = 3) +
-  scale_fill_viridis_d(option = "cividis") +
-  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
-  plot_theme +
-  geom_bracket(data = absolute_numbers,
-               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
-                   xmax = as.numeric(factor(knowledge)) + 0.4, 
-                   y.position = 1.05, label =  paste0("n = ", n)),
-               inherit.aes = FALSE,
-               step.increase = 0) +
-  facet_wrap(~measure)
-```
-
-(ref:exp4-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
+(ref:exp4-plot-overview) Summary plot for Study 4.
+**A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer **B** Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus
 
@@
 ```
+
+## Reasons for consensus rejection
+
+```{r exp4-reasons-rejection}
+# store all variables
+reasons_rejection <- unique(exp4_long$subject)
+
+table <- exp4_wide %>% 
+  select(id, all_of(reasons_rejection)) %>% 
+  pivot_longer(all_of(reasons_rejection), 
+               names_to = "question",
+               values_to = "answer") %>% 
+  drop_na(answer) %>% 
+  arrange(id)
+
+# apa_table(table,
+#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
+#           longtable = TRUE
+#           )
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Reasons for consensus rejection (raw answers)") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2) %>%
+  column_spec(3, width = "30em")
+
+```
+
+## Reasons for consensus acceptance
+
+```{r exp4-other-reasons-acceptance}
+# store all variables
+table <- exp4_wide %>% 
+  select(id, other_reason_agreement) %>% 
+  drop_na(other_reason_agreement) %>% 
+  arrange(id)
+
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Reasons for consensus acceptance (participants who answered 'other')") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2, width = "30em")
+```
+
+## How did participants independently verifiy?
+
+```{r exp4-independent-verification}
+# store all variables
+table <- exp4_wide %>% 
+  select(id, reason_followup) %>% 
+  drop_na(reason_followup) %>% 
+  arrange(id)
+
+kbl(table, booktabs = T, longtable = TRUE,
+    caption = "Answers from participants saying they trusted the scientific consens because they verified independently on how they did so.") %>%
+  kable_paper(full_width = F) %>% 
+  column_spec(1) %>%
+  column_spec(2, width = "30em")
+```

### 2024-09-19 appendix_knowledge.Rmd
submission public understanding of science

# Science knowledge {#knowledge}

Figure \@ref(fig:knowledge-overview) shows the distributions of basic science knowledge across the four studies.

(ref:knowledge-overview) Distributions of science knowledge.

```{r knowledge-overview, fig.cap="(ref:knowledge-overview)", fig.width=8, fig.height=7}
plot_data <- combined_data %>%
  mutate(study = paste0("Study ", study)) %>% 
  group_by(study, id) %>% 
  summarize(n_correct = sum(knowledge)) %>% 
  group_by(study, n_correct) %>% 
  summarize(n = n_distinct(id)) %>% 
  group_by(study) %>% 
  mutate(share = n/sum(n)) %>% 
  rounded_numbers() %>% 
  ungroup()

ggplot(plot_data, aes(x = n_correct, y = share)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(0, 10, 1)) +
  labs(x = "Number of correct answers (out of 10 possible)", y = "Share of participants") +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  plot_theme +
  facet_wrap(~study)
```

### 2024-09-19 paper.Rmd
submission public understanding of science

@@
 abstract: |
-  XX
-
-keywords          :
+  Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
+This might indicate a wholesale rejection of science.
+In four studies, we asked 782 US participants questions about trust in science, conspiracy beliefs, and basic science (e.g.
+the relative size of electrons and atoms).
+Participants were provided with the scientifically consensual answer to the basic science questions, and asked whether they accept it.
+Acceptance of the scientific consensus was very high in the sample as a whole (95.1%), but also in every sub-sample (e.g., no trust in science: 87.3%; complete endorsement of flat Earth theory: 87.2%).
+This quasi-universal acceptance of basic science suggests that people are motivated to reject specific scientific beliefs, and not science as a whole.
+This could be leveraged in science communication.
+
+keywords          : Trust in science; conspiracy theories; conspiracy thinking; science knowledge.
 
@@
   - "appendix_acceptance_knowledge.Rmd"
+  - "appendix_conspiracy_thinking.Rmd"
+  - "appendix_trust.Rmd"
+  - "appendix_conspiracy_belief.Rmd"
+  - "appendix_knowledge.Rmd"
+  - "appendix_by-question.Rmd"
 
@@
 
-Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change [@colognaRoleTrustClimate2020] or vaccination [@sturgisTrustScienceSocial2021; @lindholtPublicAcceptanceCOVID192021] to following recommendations during COVID [@alganTrustScientistsTimes2021], which suggests that trust in science was the most important predictor of these behaviors).
+Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change [@colognaRoleTrustClimate2020] or vaccination [@sturgisTrustScienceSocial2021; @lindholtPublicAcceptanceCOVID192021] to following recommendations during COVID [@alganTrustScientistsTimes2021, which suggests that trust in science was the most important predictor of these behaviors].
 
@@
 The average level of science knowledge is quite low, even in countries with extensive science education.
-For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016].
+For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016].
 However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.
@@
 However, this meant the more participants had initially provided correct answers, the fewer opportunities they had for accepting correct answers.
-We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM[^1].
+We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM, section \@ref(exp1) to \@ref(exp4) [^1].
 However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
@@
 
-After removing failed attention checks, we were left with a total sample size of `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in Study 1, 6 failed attention checks; `r exp2_descriptives$n_subj` in Study 2, 11 failed attention checks; `r exp3_descriptives$n_subj` in Study 3, no failed attention checks; `r exp4_descriptives$n_subj` in Study 4, 2 failed attention checks) participants in the US via prolific.
+After removing failed attention checks, the total sample size was `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in Study 1, six failed attention checks; `r exp2_descriptives$n_subj` in Study 2, 11 failed attention checks; `r exp3_descriptives$n_subj` in Study 3, no failed attention checks; `r exp4_descriptives$n_subj` in Study 4, two failed attention checks) participants in the US, recruited through Prolific.
 Details and demographics can be found in the online supplemental material.
 While samples for Studies 1 and 2 were convenience samples, Studies 3 and 4 were conducted on a sample holding vaccine-skeptic beliefs.
-Prolific Academic allows selecting participants based on their answers to a range of questions.
+Prolific allows selecting participants based on their answers to a range of questions.
 We picked three of these questions and only recruited participants who met our criteria for each of them:
@@
 1.
-"Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: [For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say"].
-We selected pariticpants who answered "Against".
+"Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: [For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say]".
+We selected participants who answered "Against".
 2.  "Have you received a coronavirus (COVID-19) vaccination? [Yes (at least one dose); No; Prefer not to answer]". We select only people who answered "No".
@@
 
-We used the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013] and the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016].
-Details and comparisons between the scales can be found in the ESM.
+For all results presented here, we used the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013].
+We also assessed the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016].
+Details and comparisons between the scales can be found in the ESM, sections \@ref(exp1) to \@ref(exp4).
 
@@
 We chose this question as it seemed to be the most general one.
-In the appendix, we additionally report results for two alternative measures of trust included in our studies: Another from the WGM surveys ("How much do you trust scientists in this country?
+In the ESM (sections \@ref(exp1) to \@ref(exp4)), we additionally report results for two alternative measures of trust included in our studies: Another from the WGM surveys ("How much do you trust scientists in this country?
 Do you trust them a lot, some, not much, or not at all?
-[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), and one from the Pew Research Center ("How much confidence do you have in scientists to act in the best interests of the public?
-[1-5; 1 = No confidence at all, 5 = A great deal of confidence]"; see e.g.
-@funkKeyFindingsPublic2019 and used by e.g.
-@colognaTrustScientistsTheir2024).
+[1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), and one from the Pew Research Center [e.g.
+@funkKeyFindingsPublic2019] ("How much confidence do you have in scientists to act in the best interests of the public?
+[1-5; 1 = No confidence at all, 5 = A great deal of confidence]"), the latter having been used in a recent international study on trust in science [@colognaTrustScientistsTheir2024].
 We selected these items so that we could compare the answers in our sample to global survey results.
-We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM).
+We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM, sections \@ref(exp3) and \@ref(exp4)).
 
@@
 The main outcome of interest is acceptance of the scientifically consensual facts presented.
-Overall, acceptance was very high (aggregating across all studies: `r combined_descriptives$mean_acceptance$Yes$share`; Study 1: `r exp1_descriptives$means$acceptance_mean`; Study 2: `r exp2_descriptives$means$acceptance_mean`; Study 3: `r exp3_descriptives$means$acceptance_mean`; Study 4: `r exp4_descriptives$means$acceptance_mean`).
+Overall, acceptance was very high (aggregating across all studies: `r combined_descriptives$mean_acceptance$Yes$share`; Studies 1: `r exp1_descriptives$means$acceptance_mean`; 2: `r exp2_descriptives$means$acceptance_mean`; 3: `r exp3_descriptives$means$acceptance_mean`; 4: `r exp4_descriptives$means$acceptance_mean`).
 Note that this includes both participants who had previously correctly answered the knowledge question, and participants who changed their mind when presented with the scientific consensus.
-In Studies 3 and 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: `r exp3_descriptives$means$acceptance_initial_mean`; in Study 4: `r exp4_descriptives$means$acceptance_initial_mean`).
+In Studies 3 and 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Studies 3: `r exp3_descriptives$means$acceptance_initial_mean`; 4: `r exp4_descriptives$means$acceptance_initial_mean`).
 
@@
 
+```{r}
+#combined_descriptives$flat_earth_no_trust_nparticipants
+```
+
 Participants in the lowest decile of acceptance still had an average acceptance rate of `r combined_descriptives$lowest_decile_acceptance$mean_acceptance_rate`.
-Even participants who considered as "completely true" that the earth is flat and who said they do "not trust science at all" (N = `r combined_descriptives$flat_earth_no_trust_nparticipants`) had an average acceptance rate of `r combined_descriptives$flat_earth_no_trust_acceptance$Yes$share`.
-
-These high acceptance rates do not merely reflect science knowledge: participants only correctly answered `r combined_descriptives$mean_knowledge$true$share` of the questions (Study 1: `r exp1_descriptives$means$knowledge_mean`; Study 2: `r exp2_descriptives$means$knowledge_mean`; Study 3: `r exp3_descriptives$means$knowledge_mean`; Study 4: `r exp4_descriptives$means$knowledge_mean`) before they were provided with the scientifically consensual answer.
-Even the lowest decile in science knowledge which, on average, answered correctly only on `r combined_descriptives$lowest_decile_knowledge$mean_knowledge_rate` of the questions, had an average acceptance rate of `r combined_descriptives$lowest_decile_knowledge$mean_acceptance_rate`.
+Even the three participants who considered as "completely true" that the earth is flat and who said they do "not trust science at all" had an average acceptance rate of `r combined_descriptives$flat_earth_no_trust_acceptance$Yes$share`.
+
+These high acceptance rates do not merely reflect science knowledge: participants only correctly answered `r combined_descriptives$mean_knowledge$true$share` of the questions (Studies 1: `r exp1_descriptives$means$knowledge_mean`; 2: `r exp2_descriptives$means$knowledge_mean`; 3: `r exp3_descriptives$means$knowledge_mean`; 4: `r exp4_descriptives$means$knowledge_mean`) before they were provided with the scientifically consensual answer.
+Even the lowest decile in science knowledge, which on average answered correctly only on `r combined_descriptives$lowest_decile_knowledge$mean_knowledge_rate` of the questions, had an average acceptance rate of `r combined_descriptives$lowest_decile_knowledge$mean_acceptance_rate`.
 
@@
 Yes.
-In most cases (study 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In most cases (Studies 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
 
@@
 Why did participants reject the scientific consensus?
-We collected a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (study 2: `r exp2_descriptives$justifications_n`; study 3: `r exp3_descriptives$justifications_n`; study 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (study 2: `r exp2_descriptives$justifications_n_participants`; study 3: `r exp3_descriptives$justifications_n_participants`; study 4: `r exp4_descriptives$justifications_n_participants`) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Based on the answers, we created five categories.
-Table \@ref(tab:justifications) summarizes these answers by five categories.
-All answers can be read in the ESM.
+We collected a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (Studies 2: `r exp2_descriptives$justifications_n`; 3: `r exp3_descriptives$justifications_n`; 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (Studies 2: `r exp2_descriptives$justifications_n_participants`; 3: `r exp3_descriptives$justifications_n_participants`; 4: `r exp4_descriptives$justifications_n_participants`) participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Based on the answers, we created five categories (Table \@ref(tab:justifications)).
+All answers can be read in the ESM, sections \@ref(exp2) to \@ref(exp4).
 
@@
 
-justifications %>%
+table <- justifications %>%
   group_by(category)%>%
@@
     `Share (instances)` = Share, 
-    `N (unique participants)` = n_subjects
+    `N (participans)*` = n_subjects
   ) %>% 
   # remove decimal places
-  mutate_if(is.numeric, ~as.integer(.x)) %>% 
-  apa_table(caption = "Justifications by category for data from studies 2, 3 and 4 combined.")
-```
-
-Why did participants accept the scientific consensus?
+  mutate_if(is.numeric, ~as.integer(.x))
+
+table %>%
+  apa_table(caption = "Justifications for rejecting the scientific consensus by category, Studies 2, 3 and 4 combined.", 
+            note = "*Participans with at least one answer in that category")
+
+# table %>% kbl(
+#   booktabs = T, longtable = T,
+#     caption = "Justifications for rejecting the scientific consensus by category, Studies 2, 3 and 4 combined.", 
+#   full_width = T) %>%
+#   footnote(general = "*Participans with at least one answer in that category",
+#            escape = FALSE,
+#            footnote_as_chunk = TRUE,
+#            threeparttable = TRUE)
+```
+
+Why did participants say they accept the scientific consensus?
 In Studies 3 and 4--the vaccine hesitant samples--we had asked participants about cases in which they agreed with the scientific consensus.
-A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (Study 3: `r exp3_descriptives$acceptance_n`; Study 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
-There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than participants saying it was because they trust scientists (Study 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; Study 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`)^[`r exp3_descriptives$acceptance_by_reason[["other"]]$share` in Study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in Study 4 answered with "other" and gave an open-ended explanation (see ESM).].
-We asked participants who said they verified independently to explain how they did it (answers can be found in the ESM).
+A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (Studies 3: `r exp3_descriptives$acceptance_n`; 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Studies 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than participants saying it was because they trust scientists (Studies 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`)[^2].
+Answers to a question about how they had done so can be found in the ESM, sections \@ref(exp3) and \@ref(exp4).
+
+[^2]: `r exp3_descriptives$acceptance_by_reason[["other"]]$share` in Study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in Study 4 answered with "other" and gave an open-ended explanation (see ESM, sections \@ref(exp3) and \@ref(exp4)).
+
+```{r}
+#exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate %>% round(digits = 0) %>% abs()
+#median(combined_data$duration_mins) %>% round(digits = 0)
+#exp3_descriptives$n_participants_over30mins
+#exp4_descriptives$n_participants_over30mins
+```
 
 In an exploratory analysis, we ran linear regressions to test whether there are differences between participants who said they had trusted science and those who said they had verified the information independently.
-Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = `r exp3_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 4; Study 4: mean = `r exp4_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value`) than those who said they verified independently (Study 3 mean = `r exp3_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`; Study 4 mean = `r exp4_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`).
-We did not find a difference regarding acceptance (Study 3: $\hat{\beta}_{\text{Acceptance}}$ = `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value` on a scale from 0 to 1; Study 4: $\hat{\beta}_{\text{Acceptance}}$ = `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value`) or regarding beliefs in conspiracy theories (Study 3: $\hat{\beta}_{\text{BCTI}}$ = `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 9; Study 4: $\hat{\beta}_{\text{BCTI}}$ = `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value`).
-We also did not find a difference in time spent on the survey in Study 3 ($\hat{\beta}_{\text{Time}}$ = `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp3_wide$duration_mins)` mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate %>% round(digits = 0) %>% abs()` minutes less on the survey ($\hat{\beta}_{\text{Time}}$ = `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp4_wide$duration_mins)` mins).^[For these analyses, we excluded outliers who took over 30mins for the survey, which was estimated to take around 10 mins, and for which the median time was `r median(combined_data$duration_mins) %>% round(digits = 0)` minutes.
-As a result, we excluded `r exp3_descriptives$n_participants_over30mins` participant in Study 3 and `r exp4_descriptives$n_participants_over30mins` in Study 4.
-Significance levels remain the same when including these outliers.] In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+Participants who said they accepted the consensus because of trust in scientists reported trusting science more (Studies 3: mean = `r exp3_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 4; 4: mean = `r exp4_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value`) than those who said they verified independently (Studies 3 mean = `r exp3_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`; 4: mean = `r exp4_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`).
+We did not find a difference regarding acceptance (Studies 3: $\hat{\beta}_{\text{Acceptance}}$ = `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value` on a scale from 0 to 1; 4: $\hat{\beta}_{\text{Acceptance}}$ = `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value`) or regarding beliefs in conspiracy theories (Studies 3: $\hat{\beta}_{\text{BCTI}}$ = `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 9; 4: $\hat{\beta}_{\text{BCTI}}$ = `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value`).
+We also did not find a difference in time spent on the survey in Study 3 ($\hat{\beta}_{\text{Time}}$ = `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp3_wide$duration_mins)` mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average two minutes less on the survey ($\hat{\beta}_{\text{Time}}$ = `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp4_wide$duration_mins)` mins).[^3] In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
 On average, participants clicked only on `r exp4_descriptives$reason_means$'trust in scientists'$sum_link_clicks_mean` links (out of 20 possible clicks) and there was no difference between the two groups ($\hat{\beta}_{\text{Clicks}}$ = `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$p.value`).
 
-More detailed results addressing all our pre-registered research questions can be found in the ESM.
+[^3]: For these analyses, we excluded outliers who took over 30 mins for the survey, which was estimated to take around 10 mins, and for which the median time was seven minutes.
+As a result, we excluded one participant in Study 3 and four in Study 4.
+Significance levels are not affected by these exclusions.
+
+More detailed results addressing all our pre-registered research questions can be found in the ESM, sections \@ref(exp1) to \@ref(exp4).
 
@@
 These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+These results are in line with several recent studies in which conspiracy theorists are not less susceptible to social influence in general [@pummererConspiracyBeliefsMajority; @altayConspiracyBelieversClaim2023].
 
@@
 
-In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that could be tapped into in science communication.
+In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that can be tapped into in science communication.
 Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the ‘consensus gaps’ observed in these domains [on climate change, see, e.g.
@@
 
+```{r child = "appendix_trust.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_conspiracy_thinking.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_conspiracy_belief.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_knowledge.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_by-question.Rmd"}
+```
+
+\clearpage
+
 ```{r child = "appendix_exp1.Rmd"}

### 2024-09-19 paper.tex
submission public understanding of science

@@
 {}{\PatchFailed}
+\keywords{Trust in science; conspiracy theories; conspiracy thinking; science knowledge.}
 \usepackage{csquotes}
@@
   pdflang={en-EN},
+  pdfkeywords={Trust in science; conspiracy theories; conspiracy thinking; science knowledge.},
   hidelinks,
@@
 \abstract{%
-XX
+Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
+This might indicate a wholesale rejection of science.
+In four studies, we asked 782 US participants questions about trust in science, conspiracy beliefs, and basic science (e.g.~the relative size of electrons and atoms).
+Participants were provided with the scientifically consensual answer to the basic science questions, and asked whether they accept it.
+Acceptance of the scientific consensus was very high in the sample as a whole (95.1\%), but also in every sub-sample (e.g., no trust in science: 87.3\%; complete endorsement of flat Earth theory: 87.2\%).
+This quasi-universal acceptance of basic science suggests that people are motivated to reject specific scientific beliefs, and not science as a whole.
+This could be leveraged in science communication.
 }
@@
 
-Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change (Cologna \& Siegrist, 2020) or vaccination (Lindholt, Jørgensen, Bor, \& Petersen, 2021; Sturgis, Brunton-Smith, \& Jackson, 2021) to following recommendations during COVID (Algan, Cohen, Davoine, Foucault, \& Stantcheva, 2021), which suggests that trust in science was the most important predictor of these behaviors).
+Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change (Cologna \& Siegrist, 2020) or vaccination (Lindholt, Jørgensen, Bor, \& Petersen, 2021; Sturgis, Brunton-Smith, \& Jackson, 2021) to following recommendations during COVID (Algan, Cohen, Davoine, Foucault, \& Stantcheva, 2021, which suggests that trust in science was the most important predictor of these behaviors).
 
@@
 The average level of science knowledge is quite low, even in countries with extensive science education.
-For example, only 55\% in the US (in 2014) and only 46\% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses (Science Literacy et al., 2016).
+For example, only 55\% in the US (in 2014) and only 46\% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses (National Academies of Sciences, Engineering, and Medicine, 2016).
 However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.
@@
 However, this meant the more participants had initially provided correct answers, the fewer opportunities they had for accepting correct answers.
-We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM\footnote{Only in Study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (Studies 1: r = 0.06, p = 0.387; 2: r = 0.16, p = 0.051; 3: r = 0.04, p = 0.619; 4: r = 0.15, p = 0.037) and only in Study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (Studies 1: r = -0.14, p = 0.061; 2: r = -0.22, p = 0.006; 3: r = -0.04, p = 0.631; 4: r = -0.05, p = 0.455).}.
+We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM, section \ref{exp1} to \ref{exp4} \footnote{Only in Study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (Studies 1: r = 0.06, p = 0.387; 2: r = 0.16, p = 0.051; 3: r = 0.04, p = 0.619; 4: r = 0.15, p = 0.037) and only in Study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (Studies 1: r = -0.14, p = 0.061; 2: r = -0.22, p = 0.006; 3: r = -0.04, p = 0.631; 4: r = -0.05, p = 0.455).}.
 However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.
@@
 
-After removing failed attention checks, we were left with a total sample size of 782 (194 in Study 1, 6 failed attention checks; 190 in Study 2, 11 failed attention checks; 200 in Study 3, no failed attention checks; 198 in Study 4, 2 failed attention checks) participants in the US via prolific.
+After removing failed attention checks, the total sample size was 782 (194 in Study 1, six failed attention checks; 190 in Study 2, 11 failed attention checks; 200 in Study 3, no failed attention checks; 198 in Study 4, two failed attention checks) participants in the US, recruited through Prolific.
 Details and demographics can be found in the online supplemental material.
 While samples for Studies 1 and 2 were convenience samples, Studies 3 and 4 were conducted on a sample holding vaccine-skeptic beliefs.
-Prolific Academic allows selecting participants based on their answers to a range of questions.
+Prolific allows selecting participants based on their answers to a range of questions.
 We picked three of these questions and only recruited participants who met our criteria for each of them:
@@
 \item
-  ``Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: {[}For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say''{]}.
-We selected pariticpants who answered''Against''.
+  ``Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: {[}For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say{]}''.
+We selected participants who answered ``Against''.
 \item
@@
 \addlinespace
-11 & **Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air] & \\
+11 & *Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air] & \\
 \bottomrule
@@
 
-We used the four-item conspiracy mentality questionnaire (CMQ) (Bruder, Haffke, Neave, Nouripanah, \& Imhoff, 2013) and the single item conspiracy beliefs scale (SICBS) (Lantian, Muller, Nurra, \& Douglas, 2016).
-Details and comparisons between the scales can be found in the ESM.
+For all results presented here, we used the four-item conspiracy mentality questionnaire (CMQ) (Bruder, Haffke, Neave, Nouripanah, \& Imhoff, 2013).
+We also assessed the single item conspiracy beliefs scale (SICBS) (Lantian, Muller, Nurra, \& Douglas, 2016).
+Details and comparisons between the scales can be found in the ESM, sections \ref{exp1} to \ref{exp4}.
 
@@
 We chose this question as it seemed to be the most general one.
-In the appendix, we additionally report results for two alternative measures of trust included in our studies: Another from the WGM surveys (``How much do you trust scientists in this country?
+In the ESM (sections \ref{exp1} to \ref{exp4}), we additionally report results for two alternative measures of trust included in our studies: Another from the WGM surveys (``How much do you trust scientists in this country?
 Do you trust them a lot, some, not much, or not at all?
-{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''), and one from the Pew Research Center (``How much confidence do you have in scientists to act in the best interests of the public?
-{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}''; see e.g.
-Funk, Johnson, and Hefferon (2019) and used by e.g.
-Cologna et al.
-(2024)).
+{[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}''), and one from the Pew Research Center (e.g.
+Funk, Johnson, \& Hefferon, 2019) (``How much confidence do you have in scientists to act in the best interests of the public?
+{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}''), the latter having been used in a recent international study on trust in science (Cologna et al., 2024).
 We selected these items so that we could compare the answers in our sample to global survey results.
-We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM).
+We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM, sections \ref{exp3} and \ref{exp4}).
 
@@
 The main outcome of interest is acceptance of the scientifically consensual facts presented.
-Overall, acceptance was very high (aggregating across all studies: 95.1 \%; Study 1: 93 \%; Study 2: 98 \%; Study 3: 98 \%; Study 4: 91 \%).
+Overall, acceptance was very high (aggregating across all studies: 95.1 \%; Studies 1: 93 \%; 2: 98 \%; 3: 98 \%; 4: 91 \%).
 Note that this includes both participants who had previously correctly answered the knowledge question, and participants who changed their mind when presented with the scientific consensus.
-In Studies 3 and 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Study 3: 96 \%; in Study 4: 86 \%).
+In Studies 3 and 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Studies 3: 96 \%; 4: 86 \%).
 
@@
 Participants in the lowest decile of acceptance still had an average acceptance rate of 67.4 \%.
-Even participants who considered as ``completely true'' that the earth is flat and who said they do ``not trust science at all'' (N = 3) had an average acceptance rate of 86.7 \%.
-
-These high acceptance rates do not merely reflect science knowledge: participants only correctly answered 65.8 \% of the questions (Study 1: 74 \%; Study 2: 79 \%; Study 3: 75 \%; Study 4: 36 \%) before they were provided with the scientifically consensual answer.
-Even the lowest decile in science knowledge which, on average, answered correctly only on 20 \% of the questions, had an average acceptance rate of 91.7 \%.
+Even the three participants who considered as ``completely true'' that the earth is flat and who said they do ``not trust science at all'' had an average acceptance rate of 86.7 \%.
+
+These high acceptance rates do not merely reflect science knowledge: participants only correctly answered 65.8 \% of the questions (Studies 1: 74 \%; 2: 79 \%; 3: 75 \%; 4: 36 \%) before they were provided with the scientifically consensual answer.
+Even the lowest decile in science knowledge, which on average answered correctly only on 20 \% of the questions, had an average acceptance rate of 91.7 \%.
 
@@
 Yes.
-In most cases (study 1: 76.3 \%; study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+In most cases (Studies 1: 76.3 \%; 2: 92.9 \%; 3: 95.5 \%; 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
 
@@
 Why did participants reject the scientific consensus?
-We collected a total of 364 answers (study 2: 35; study 3: 74; study 4: 255) from 167 (study 2: 25; study 3: 47; study 4: 95) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
-Based on the answers, we created five categories.
-Table \ref{tab:justifications} summarizes these answers by five categories.
-All answers can be read in the ESM.
+We collected a total of 364 answers (Studies 2: 35; 3: 74; 4: 255) from 167 (Studies 2: 25; 3: 47; 4: 95) participants to the open-ended questions on why they had rejected the scientific consensus on a particular question.
+Based on the answers, we created five categories (Table \ref{tab:justifications}).
+All answers can be read in the ESM, sections \ref{exp2} to \ref{exp4}.
 
@@
 
-\caption{\label{tab:justifications}Justifications by category for data from studies 2, 3 and 4 combined.}
+\caption{\label{tab:justifications}Justifications for rejecting the scientific consensus by category, Studies 2, 3 and 4 combined.}
 
@@
 \toprule
-Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (unique participants)}\\
+Category & \multicolumn{1}{c}{N (instances)} & \multicolumn{1}{c}{Share (instances)} & \multicolumn{1}{c}{N (participans)*}\\
 \midrule
@@
 \bottomrule
+\addlinespace
 \end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} *Participans with at least one answer in that category}
+\end{tablenotes}
 
@@
 
-Why did participants accept the scientific consensus?
+Why did participants say they accept the scientific consensus?
 In Studies 3 and 4--the vaccine hesitant samples--we had asked participants about cases in which they agreed with the scientific consensus.
-A total of 320 (Study 3: 122; Study 4: 198) participants answered this question.
-There were more participants saying they accepted the scientific consensus because they independently verified the fact (Study 3: 47.5\%; Study 4: 47\% ), than participants saying it was because they trust scientists (Study 3: 41.8\%; Study 4: 36.4\%)\footnote{10.7\% in Study 3 and 16.7\% in Study 4 answered with ``other'' and gave an open-ended explanation (see ESM).}.
-We asked participants who said they verified independently to explain how they did it (answers can be found in the ESM).
+A total of 320 (Studies 3: 122; 4: 198) participants answered this question.
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Studies 3: 47.5\%; 4: 47\% ), than participants saying it was because they trust scientists (Studies 3: 41.8\%; 4: 36.4\%)\footnote{10.7\% in Study 3 and 16.7\% in Study 4 answered with ``other'' and gave an open-ended explanation (see ESM, sections \ref{exp3} and \ref{exp4}).}.
+Answers to a question about how they had done so can be found in the ESM, sections \ref{exp3} and \ref{exp4}.
 
 In an exploratory analysis, we ran linear regressions to test whether there are differences between participants who said they had trusted science and those who said they had verified the information independently.
-Perhaps unsurprisingly, those who said they accepted consensus because of trust in scientists reported to trust science more (Study 3: mean = 3; \(\hat{\beta}_{\text{Trust}}\) = 0.19, p = 0.330 on a scale from 1 to 4; Study 4: mean = 2.92; \(\hat{\beta}_{\text{Trust}}\) = 0.46, p \textless{} .001) than those who said they verified independently (Study 3 mean = 2.81; Study 4 mean = 2.45).
-We did not find a difference regarding acceptance (Study 3: \(\hat{\beta}_{\text{Acceptance}}\) = 0.01, p = 0.523 on a scale from 0 to 1; Study 4: \(\hat{\beta}_{\text{Acceptance}}\) = 0.04, p = 0.116) or regarding beliefs in conspiracy theories (Study 3: \(\hat{\beta}_{\text{BCTI}}\) = -0.16, p = 0.694 on a scale from 1 to 9; Study 4: \(\hat{\beta}_{\text{BCTI}}\) = -0.21, p = 0.322).
-We also did not find a difference in time spent on the survey in Study 3 (\(\hat{\beta}_{\text{Time}}\) = -0.02, p = 0.985; median = 7.64 mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average 2 minutes less on the survey (\(\hat{\beta}_{\text{Time}}\) = -2.16, p = 0.018; median = 9.38 mins).\footnote{For these analyses, we excluded outliers who took over 30mins for the survey, which was estimated to take around 10 mins, and for which the median time was 7 minutes.
-As a result, we excluded 1 participant in Study 3 and 4 in Study 4.
-Significance levels remain the same when including these outliers.} In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+Participants who said they accepted the consensus because of trust in scientists reported trusting science more (Studies 3: mean = 3; \(\hat{\beta}_{\text{Trust}}\) = 0.19, p = 0.330 on a scale from 1 to 4; 4: mean = 2.92; \(\hat{\beta}_{\text{Trust}}\) = 0.46, p \textless{} .001) than those who said they verified independently (Studies 3 mean = 2.81; 4: mean = 2.45).
+We did not find a difference regarding acceptance (Studies 3: \(\hat{\beta}_{\text{Acceptance}}\) = 0.01, p = 0.523 on a scale from 0 to 1; 4: \(\hat{\beta}_{\text{Acceptance}}\) = 0.04, p = 0.116) or regarding beliefs in conspiracy theories (Studies 3: \(\hat{\beta}_{\text{BCTI}}\) = -0.16, p = 0.694 on a scale from 1 to 9; 4: \(\hat{\beta}_{\text{BCTI}}\) = -0.21, p = 0.322).
+We also did not find a difference in time spent on the survey in Study 3 (\(\hat{\beta}_{\text{Time}}\) = -0.02, p = 0.985; median = 7.64 mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average two minutes less on the survey (\(\hat{\beta}_{\text{Time}}\) = -2.16, p = 0.018; median = 9.38 mins).\footnote{For these analyses, we excluded outliers who took over 30 mins for the survey, which was estimated to take around 10 mins, and for which the median time was seven minutes.
+As a result, we excluded one participant in Study 3 and four in Study 4.
+Significance levels are not affected by these exclusions.} In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
 On average, participants clicked only on 1.36 links (out of 20 possible clicks) and there was no difference between the two groups (\(\hat{\beta}_{\text{Clicks}}\) = -1.09, p = 0.062).
 
-More detailed results addressing all our pre-registered research questions can be found in the ESM.
+More detailed results addressing all our pre-registered research questions can be found in the ESM, sections \ref{exp1} to \ref{exp4}.
 
@@
 These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+These results are in line with several recent studies in which conspiracy theorists are not less susceptible to social influence in general (Altay, Nera, Ejaz, Schöpfer, \& Tomas, 2023; Pummerer, Fock, Winter, \& Sassenberg, n.d.).
 
@@
 
-In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that could be tapped into in science communication.
+In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that can be tapped into in science communication.
 Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the `consensus gaps' observed in these domains (on climate change, see, e.g.
@@
 \url{https://doi.org/10.1177/0963662506070159}
+
+\bibitem[\citeproctext]{ref-altayConspiracyBelieversClaim2023}
+Altay, S., Nera, K., Ejaz, W., Schöpfer, C., \& Tomas, F.
+(2023).
+Conspiracy believers claim to be free thinkers but (Under)Use advice like everyone else.
+\emph{British Journal of Social Psychology}, \emph{62}(4), 1782--1797.
+\url{https://doi.org/10.1111/bjso.12655}
 
@@
 
-\bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
-Pennycook, G., Binnendyk, J., \& Rand, D.
-(2022).
-\emph{Overconfidently conspiratorial: Conspiracy believers are dispositionally overconfident and massively overestimate how much others agree with them}.
-\url{https://doi.org/10.31234/osf.io/d5fz2}
-
-\bibitem[\citeproctext]{ref-ranneyClimateChangeConceptual2016}
-Ranney, M.
-A., \& Clark, D.
-(2016).
-Climate Change Conceptual Change: Scientific Information Can Transform Attitudes.
-\emph{Topics in Cognitive Science}, \emph{8}(1), 49--75.
-\url{https://doi.org/10.1111/tops.12187}
-
-\bibitem[\citeproctext]{ref-committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016}
-Science Literacy, C.
-on, Public Perception of Science, Board on Science Education, Behavioral, D.
-of, Sciences, S., Education, \& National Academies of Sciences, Engineering, and Medicine.
+\bibitem[\citeproctext]{ref-nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016}
+National Academies of Sciences, Engineering, and Medicine.
 (2016).
@@
 \url{https://doi.org/10.17226/23595}
+
+\bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
+Pennycook, G., Binnendyk, J., \& Rand, D.
+(2022).
+\emph{Overconfidently conspiratorial: Conspiracy believers are dispositionally overconfident and massively overestimate how much others agree with them}.
+\url{https://doi.org/10.31234/osf.io/d5fz2}
+
+\bibitem[\citeproctext]{ref-pummererConspiracyBeliefsMajority}
+Pummerer, L., Fock, L., Winter, K., \& Sassenberg, K.
+(n.d.).
+Conspiracy beliefs and majority influence.
+\emph{The Journal of Social Psychology}, \emph{0}(0), 1--16.
+\url{https://doi.org/10.1080/00224545.2024.2397491}
+
+\bibitem[\citeproctext]{ref-ranneyClimateChangeConceptual2016}
+Ranney, M.
+A., \& Clark, D.
+(2016).
+Climate Change Conceptual Change: Scientific Information Can Transform Attitudes.
+\emph{Topics in Cognitive Science}, \emph{8}(1), 49--75.
+\url{https://doi.org/10.1111/tops.12187}
 
@@
 
-Did participants who had initially gotten it wrong change their minds towards the scientific consensus?
-Yes.
-In most cases (study 1: 76.3 \%; study 2: 92.9 \%; study 3: 95.5 \%; study 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
+Did participants who gave a wrong answer change their minds towards the scientific consensus?
+In most cases (Studies 1: 76.3 \%; 2: 92.9 \%; 3: 95.5 \%; 4: 89.5 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
 In very few cases, participants changed their mind away from the scientific consensus.
-These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: 1.6 \%; study 2: 0.5 \%; study 3: 0.7 \%; study 4: 5.7 \%).
+These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (Studies 1: 1.6 \%; 2: 0.5 \%; 3: 0.7 \%; 4: 5.7 \%).
 Figure \ref{fig:acceptance-knowledge} shows acceptance rates conditional on whether the initial answer to the questions was correct or not.
@@
 \includegraphics{output/figures/acceptance-knowledge.pdf}
-\caption{\label{fig:acceptance-knowledge}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\caption{\label{fig:acceptance-knowledge}Acceptance/Rejection rate of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
+\end{figure}
+
+\clearpage
+
+\section{Trust in science}\label{trust}
+
+Figure \ref{fig:trust-overview} shows the distributions of our main trust in science measure across the four studies.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/trust-overview.pdf}
+\caption{\label{fig:trust-overview}Distributions of trust in science.}
+\end{figure}
+
+\clearpage
+
+\section{Conspiracy Thinking}\label{conspiracy-thinking}
+
+Figure \ref{fig:conspiracy-thinking-overview} shows the distributions of conspiracy thinking across the four studies.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/conspiracy-thinking-overview.pdf}
+\caption{\label{fig:conspiracy-thinking-overview}Distributions of conspiracy thinking.}
+\end{figure}
+
+\clearpage
+
+\section{Conspiracy Belief}\label{conspiracy-belief}
+
+Figure \ref{fig:conspiracy-belief-overview} shows the average belief in each of the conspiracy theories across the four studies.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/conspiracy-belief-overview.pdf}
+\caption{\label{fig:conspiracy-belief-overview}Average belief in conspiracy theories.}
+\end{figure}
+
+\clearpage
+
+\section{Science knowledge}\label{knowledge}
+
+Figure \ref{fig:knowledge-overview} shows the distributions of basic science knowledge across the four studies.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/knowledge-overview.pdf}
+\caption{\label{fig:knowledge-overview}Distributions of science knowledge.}
+\end{figure}
+
+\clearpage
+
+\section{Acceptance and knowledge by question}\label{by-question-variation}
+
+Figure \ref{fig:knowledge-overview} shows the share of correct answers, and Figure \ref{fig:knowledge-overview} the acceptance rate of the scientific consensus, for each knowledge question asked in the respective study.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/questions-knowledge.pdf}
+\caption{\label{fig:questions-knowledge}Share of correct answers by question.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/questions-acceptance.pdf}
+\caption{\label{fig:questions-acceptance}Acceptance rate of consensus by question.}
 \end{figure}
@@
 
-Beside Belief in Conspiracy Theory Inventory (BCTI) by Pennycook et al. (2022) which report in the main study, we also assessed two other consipiracy thinking measures:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  The conspiracy mentality questionnaire (CMQ) by Bruder et al. (2013):
-  I think that . . .
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  \ldots{} many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
-\item
-  \ldots{} government agencies closely monitor all citizens.
-\item
-  \ldots{} events which superficially seem to lack a connection are often the result of secret activities.
-\item
-  \ldots{} there are secret organizations that greatly influence political decisions.
-\end{itemize}
-
-{[}0\% - 100\%; 0 = certainly not, 100 = certain{]}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\tightlist
-\item
-  The Single Item Conspiracy Beliefs Scale (SICBS) by Lantian et al. (2016) :
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  I think that the official version of the events given by the authorities very often hides the truth. {[}1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true{]}
-\end{itemize}
-
-\subsubsection{Trust in science}\label{trust-in-science-1}
-
-We rely on three items
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
-\item
-  In general, would you say that you trust science a lot, some, not much, or not at all? {[}1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot{]}
-\item
-  How much confidence do you have in scientists to act in the best interests of the public? {[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}
-\end{enumerate}
+In addition to the Belief in Conspiracy Theory Inventory (BCTI) by Pennycook et al.
+(2022) which report in the main study, we also assessed two other consipiracy thinking measures: The conspiracy mentality questionnaire (CMQ) by Bruder et al.
+(2013) and the Single Item Conspiracy Beliefs Scale (SICBS) by Lantian et al.
+(2016) (Table \ref{tab:overview-conspiracy-measures}).
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{20em}>{\raggedright\arraybackslash}p{10em}}
+\caption{\label{tab:overview-conspiracy-measures}Conspiracy measures}\\
+\toprule
+Measure & Description & Scale\\
+\midrule
+Belief in Conspiracy Theory Inventory (BCTI) & By Pennycook et al., 2022, presented in the main article. & 1 - 9 (1 = Completely false, 5 = Unsure, 9 = Completely true)\\
+Conspiracy Mentality Questionnaire (CMQ) & By Bruder et al., 2013. Respondents are asked how much they agree with the following statements: & 0\% - 100\% (0 = certainly not, 100 = certain)\\
+ & - Many important things happen in the world, which the public is never informed about. & \\
+ & - Politicians usually do not tell us the true motives for their decisions. & \\
+ & - Government agencies closely monitor all citizens. & \\
+\addlinespace
+ & - Events which seem unrelated are often the result of secret activities. & \\
+ & - There are secret organizations that greatly influence political decisions. & \\
+Single Item Conspiracy Beliefs Scale (SICBS) & By Lantian et al., 2016.
+The item is: 'I think that the official version of events given by the authorities very often hides the truth.' & 1 - 9 (1 = Completely false, 5 = Unsure, 9 = Completely true)\\
+\bottomrule
+\end{longtable}
+\endgroup{}
 
@@
 
-Fig.
-\ref{fig:exp1-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
 In most cases (76.3 \%), participants readily accepted the scientific consensus after having given the wrong answer to a question.
@@
 These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp1-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+\ref{fig:exp1-plot-overview}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
 
@@
 
-In Appendix \ref{exp1}, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
-Appendix \ref{exp1} also includes more descriptive statistics, such as knowledge and acceptance by science questions.
+Below, we show that these associations are robust when using alternative measures of trust and conspiracy thinking.
+We also provide more descriptive statistics, such as knowledge and acceptance by science questions.
 
@@
 
-\subsection{By-question variation}\label{by-question-variation}
-
-\subsubsection{Knowledge}\label{knowledge}
-
 \begin{figure}
 \centering
-\includegraphics{output/figures/exp1-questions-knowledge.pdf}
-\caption{\label{fig:exp1-questions-knowledge}Distribution of correct answers by question.}
+\includegraphics{output/figures/exp1-plot-overview.pdf}
+\caption{\label{fig:exp1-plot-overview}Summary plot for Study 1.
+\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
 \end{figure}
 
-\subsubsection{Acceptance}\label{acceptance}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-questions-acceptance.pdf}
-\caption{\label{fig:exp1-questions-acceptance}Distribution of consensus acceptance by question.}
-\end{figure}
-
-\subsection{Trust in science}\label{trust-in-science-3}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-trust-scientists.pdf}
-\caption{\label{fig:exp1-trust-scientists}Distribution of trust in scientists.}
-\end{figure}
-
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-1}
-
-\subsubsection{Distribution}\label{distribution}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-conspiracy-distribution.pdf}
-\caption{\label{fig:exp1-conspiracy-distribution}Distribution of conspiracy thinking.}
-\end{figure}
-
-\subsubsection{By-item variation}\label{by-item-variation}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-conspiracy-items.pdf}
-\caption{\label{fig:exp1-conspiracy-items}Distribution of conspiracy thinking by item.}
-\end{figure}
-
 \subsection{Open ended comments}\label{open-ended-comments}
@@
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:unnamed-chunk-27}Reasons for consensus rejection}\\
+\caption{\label{tab:unnamed-chunk-52}Comments of participants on Study 1}\\
 \toprule
@@
 
-\subsection{Additional plots}\label{additional-plots}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-conditional-acceptance.pdf}
-\caption{\label{fig:exp1-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp1-plot-overview.pdf}
-\caption{\label{fig:exp1-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus}
-\end{figure}
-
 \clearpage
@@
 To better understand reasons for consensus rejection, after having answered all questions, we asked participants an open-ended question to explain why they rejected the consensus, for each question on which they did so.
-We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong (see Appendix \ref{exp1}).
+We also excluded the where do trees their materials from, as this question clearly seemed to be an outlier where most participants would get the answer wrong.
 
@@
 Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
-These correlations are relatively weak, which might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp2-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
-In Appendix \ref{exp2} we show that these results hold for our alternative measures of trust and conspiracy thinking.
+These correlations are relatively weak, which might be partly due to ceiling effects: (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+Below, we show that these results hold for our alternative measures of trust and conspiracy thinking.
 We also include more descriptive statistics, such as knowledge and acceptance by science questions.
@@
 Regarding RQ1, participants answered on average 79 \% (sd = 0.19) of the questions correctly, and accepted the scientific consensus on average for 98 \% (sd = 0.05) of the questions.
-Fig.
-\ref{fig:exp2-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
 In response to RQ2, in most cases (92.9 \%), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.
@@
 Table \ref{tab:exp2-justifications} summarizes these answers by five categories.
-All answers can be read in Appendix \ref{exp2}.
+All answers can be read in below.
 
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-4}
+\subsubsection{Trust in science}\label{trust-in-science-2}
 
@@
 
-\subsection{By-question variation}\label{by-question-variation-1}
-
-\subsubsection{Knowledge}\label{knowledge-1}
-
 \begin{figure}
 \centering
-\includegraphics{output/figures/exp2-questions-knowledge.pdf}
-\caption{\label{fig:exp2-questions-knowledge}Distribution of correct answers by question.}
+\includegraphics{output/figures/exp2-plot-overview.pdf}
+\caption{\label{fig:exp2-plot-overview}Summary plot for Study 2.
+\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
 \end{figure}
 
-\subsubsection{Acceptance}\label{acceptance-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-questions-acceptance.pdf}
-\caption{\label{fig:exp2-questions-acceptance}Distribution of consensus acceptance by question.}
-\end{figure}
-
-\subsection{Trust in science}\label{trust-in-science-5}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-trust-scientists.pdf}
-\caption{\label{fig:exp2-trust-scientists}Distribution of trust in scientists.}
-\end{figure}
-
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-2}
-
-\subsubsection{Distribution}\label{distribution-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-conspiracy-distribution.pdf}
-\caption{\label{fig:exp2-conspiracy-distribution}Distribution of conspiracy thinking.}
-\end{figure}
-
-\subsubsection{By-item variation}\label{by-item-variation-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-conspiracy-items.pdf}
-\caption{\label{fig:exp2-conspiracy-items}Distribution of conspiracy thinking by item.}
-\end{figure}
-
 \subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection}
 
-\subsubsection{By category}\label{by-category}
-
-{[}To do{]}
-
-\subsubsection{All raw answers}\label{all-raw-answers}
-
 \begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp2-reasons-rejection}Reasons for consensus rejection}\\
+\caption{\label{tab:exp2-reasons-rejection}Reasons for consensus rejection (raw answers)}\\
 \toprule
@@
 \end{longtable}
-
-\subsection{Additional plots}\label{additional-plots-1}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp2-conditional-acceptance.pdf}
-\caption{\label{fig:exp2-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
 
@@
 
-Table \ref{tab:knowledge-overview} shows all questions, their scientifically consensual answer, and their source.
+Table \ref{tab:exp3-knowledge-items} shows all questions, their scientifically consensual answer, and their source.
 All but two questions were selected from existing science knowledge questionnaires.
@@
 \begin{longtable}[t]{>{\raggedleft\arraybackslash}p{1em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{23em}}
-\caption{\label{tab:knowledge-overview}Science knowledge items}\\
+\caption{\label{tab:exp3-knowledge-items}Science knowledge items}\\
 \toprule
@@
 
-Fig.
-\ref{fig:exp3-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
 In most cases (87 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
@@
 We find no statistically significant correlation between science knowledge and trust in science (r = 0.14, p = 0.100), but a samll positive correlation between acceptance of scientific consensus and trust in science (r = 0.20, p = 0.019).
-Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp3-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+Again, these findings might be partly due to ceiling effects: (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
 We find no statistically significant correlation between conspiracy thinking and science knowledge (r = -0.16, p = 0.055), and between conspiracy thinking and acceptance of scientific consensus (r = -0.02, p = 0.788).
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-6}
+\subsubsection{Trust in science}\label{trust-in-science-3}
 
@@
 
-\subsection{By-question variation}\label{by-question-variation-2}
-
-\subsubsection{Knowledge}\label{knowledge-2}
-
 \begin{figure}
 \centering
-\includegraphics{output/figures/exp3-questions-knowledge.pdf}
-\caption{\label{fig:exp3-questions-knowledge}Distribution of correct answers by question.}
+\includegraphics{output/figures/exp3-plot-overview.pdf}
+\caption{\label{fig:exp3-plot-overview}Summary plot for Study 3.
+\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-questions-acceptance.pdf}
-\caption{\label{fig:exp3-questions-acceptance}Distribution of consensus acceptance by question.}
-\end{figure}
-
-\subsection{Trust in science}\label{trust-in-science-7}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-trust-scientists.pdf}
-\caption{\label{fig:exp3-trust-scientists}Distribution of trust in scientists.}
-\end{figure}
-
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-3}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-conspiracy-distribution.pdf}
-\caption{\label{fig:exp3-conspiracy-distribution}Distribution of conspiracy thinking.}
-\end{figure}
-
-\subsubsection{By-item variation}\label{by-item-variation-2}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-conspiracy-items.pdf}
-\caption{\label{fig:exp3-conspiracy-items}Distribution of conspiracy thinking by item.}
-\end{figure}
-
 \subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection-1}
 
-\subsubsection{All raw answers}\label{all-raw-answers-1}
-
 \begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp3-reasons-rejection}Reasons for consensus rejection}\\
+\caption{\label{tab:exp3-reasons-rejection}Reasons for consensus rejection (raw answers)}\\
 \toprule
@@
 
-\subsection{Why do people accept the scientific consensus:}\label{why-do-people-accept-the-scientific-consensus}
-
-\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other}
+\subsection{Reasons for consensus acceptance}\label{reasons-for-consensus-acceptance}
 
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp3-other-reasons-acceptance}Reasons for consensus rejection}\\
+\caption{\label{tab:exp3-other-reasons-acceptance}Reasons for consensus acceptance (participants who answered 'other')}\\
 \toprule
@@
 
-\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other-1}
-
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp3-independent-verification}Reasons for consensus rejection}\\
+\caption{\label{tab:exp3-independent-verification}Answers from participants saying they trusted the scientific consens because they verified independently on how they did so.}\\
 \toprule
@@
 \end{longtable}
-
-\subsection{Additional plots}\label{additional-plots-2}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-conditional-acceptance.pdf}
-\caption{\label{fig:exp3-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp3-plot-overview.pdf}
-\caption{\label{fig:exp3-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
-\end{figure}
 
@@
 
-Fig.
-\ref{fig:exp4-conditional-acceptance} illustrates the relationship between knowledge and acceptance.
 In most cases (83.7 \%), participants readily accepted the scientific consensus right after having given the wrong answer to a question.
@@
 We find no statistically significant correlation between science knowledge and trust in science (r = 0.10, p = 0.144), but a small positive correlation between acceptance of scientific consensus and trust in science (r = 0.16, p = 0.029).
-Again, these findings might be partly due to ceiling effects: As illustrated in Fig.
-\ref{fig:exp4-plot}, (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
+Again, these findings might be partly due to ceiling effects: (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
 We do not find a statistically significant correlation between neither conspiracy thinking and science knowledge (r = -0.02, p = 0.742), nor between conspiracy thinking and acceptance of scientific consensus (r = -0.05, p = 0.498).
@@
 
-\subsubsection{Trust in science}\label{trust-in-science-8}
+\subsubsection{Trust in science}\label{trust-in-science-4}
 
@@
 
-\subsection{By-question variation}\label{by-question-variation-3}
-
-\subsubsection{Knowledge}\label{knowledge-3}
-
 \begin{figure}
 \centering
-\includegraphics{output/figures/exp4-questions-knowledge.pdf}
-\caption{\label{fig:exp4-questions-knowledge}Distribution of correct answers by question.}
+\includegraphics{output/figures/exp4-plot-overview.pdf}
+\caption{\label{fig:exp4-plot-overview}Summary plot for Study 4.
+\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp4-questions-acceptance.pdf}
-\caption{\label{fig:exp4-questions-acceptance}Distribution of consensus acceptance by question.}
-\end{figure}
-
-\subsection{Trust in science}\label{trust-in-science-9}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp4-trust-scientists.pdf}
-\caption{\label{fig:exp4-trust-scientists}Distribution of trust in scientists.}
-\end{figure}
-
-\subsection{Conspiracy thinking}\label{conspiracy-thinking-4}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp4-conspiracy-distribution.pdf}
-\caption{\label{fig:exp4-conspiracy-distribution}Distribution of conspiracy thinking.}
-\end{figure}
-
-\subsubsection{By-item variation}\label{by-item-variation-3}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp4-conspiracy-items.pdf}
-\caption{\label{fig:exp4-conspiracy-items}Distribution of conspiracy thinking by item.}
-\end{figure}
-
 \subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection-2}
 
-\subsubsection{All raw answers}\label{all-raw-answers-2}
-
 \begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp4-reasons-rejection}Reasons for consensus rejection}\\
+\caption{\label{tab:exp4-reasons-rejection}Reasons for consensus rejection (raw answers)}\\
 \toprule
@@
 
-\subsection{Why do people accept the scientific consensus:}\label{why-do-people-accept-the-scientific-consensus-1}
-
-\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other-2}
+\subsection{Reasons for consensus acceptance}\label{reasons-for-consensus-acceptance-1}
 
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp4-other-reasons-acceptance}Reasons for consensus rejection}\\
+\caption{\label{tab:exp4-other-reasons-acceptance}Reasons for consensus acceptance (participants who answered 'other')}\\
 \toprule
@@
 
-\subsubsection{Participants who answered ``other''}\label{participants-who-answered-other-3}
-
 \begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp4-independent-verification}Reasons for consensus rejection}\\
+\caption{\label{tab:exp4-independent-verification}Answers from participants saying they trusted the scientific consens because they verified independently on how they did so.}\\
 \toprule
@@
 
-\subsection{Additional plots}\label{additional-plots-3}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp4-conditional-acceptance.pdf}
-\caption{\label{fig:exp4-conditional-acceptance}Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/exp4-plot-overview.pdf}
-\caption{\label{fig:exp4-plot-overview}\textbf{A} Shows the distribution of science knowledge (left) and acceptance of scientific consensus for participants who gave the wrong answer \textbf{B} Shows the relationship between trust in science and science knowledge/acceptance (if wrong at first, rounded to the first digit) of scientific consensus \textbf{C} Shows the relationship between conspiracy thinking and science knowledge/acceptance (if wrong at first) of scientific consensus}
-\end{figure}
-
 \end{document}

### 2024-09-19 paper.Rmd
add anonymized link before pus submission

@@
 We also measured participants’ trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (https://osf.io/8utsj/?view_only=364478f01fc548d0a42b9962918d871d).
 The differences between the four studies are summarized presently, and the methods are detailed below.

### 2024-09-19 paper.tex
add anonymized link before pus submission

@@
 We also measured participants' trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (\url{https://osf.io/8utsj/?view_only=364478f01fc548d0a42b9962918d871d}).
 The differences between the four studies are summarized presently, and the methods are detailed below.

### 2024-09-19 appendix_exp1.Rmd
make WORD compatible tables for PUS submission

@@
 
-# # Create the table
-kbl(conspiracy_measures, booktabs = T, longtable = T,
-    caption = "Conspiracy measures", full_width = T) %>%
-  kable_styling(font_size = 8) %>% 
-  column_spec(1, width = "10em") %>%
-  column_spec(2, width = "20em") %>%
-  column_spec(3, width = "10em")
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_styling
+  table_output <- conspiracy_measures %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Conspiracy measures",
+      full_width = TRUE
+    ) %>%
+    kableExtra::kable_styling(font_size = 8) %>%
+    kableExtra::column_spec(1, width = "10em") %>%
+    kableExtra::column_spec(2, width = "20em") %>%
+    kableExtra::column_spec(3, width = "10em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(conspiracy_measures) %>%
+    flextable::set_caption(caption = "Conspiracy measures") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 1, width = 2) %>%
+    flextable::width(j = 2, width = 4) %>%
+    flextable::width(j = 3, width = 2) # Adjust widths as needed
+}
 
-# Use apa_table to format the table
-# apa_table(
-#   conspiracy_measures,
-#   caption = "Overview of Conspiracy Thinking Measures",
-#   align = c("m{4cm}", "m{10cm}", "m{4cm}")
-# )
+# Display the table
+table_output
 ```
@@
 
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Comments of participants on Study 1") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "40em") 
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
+  table_output <- table %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Comments of participants on Study 1"
+    ) %>%
+    kableExtra::kable_paper(full_width = FALSE) %>%
+    kableExtra::column_spec(1) %>%
+    kableExtra::column_spec(2, width = "40em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(table) %>%
+    flextable::set_caption(caption = "Comments of participants on Study 1") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 2, width = 4.5) # Adjust column widths as needed
+}
+
+# Display the table
+table_output
 ```

### 2024-09-19 appendix_exp2.Rmd
make WORD compatible tables for PUS submission

@@
 
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection (raw answers)") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2) %>%
-  column_spec(3, width = "30em")
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
+  table_output <- table %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Reasons for consensus rejection (raw answers)"
+    ) %>%
+    kableExtra::kable_paper(full_width = FALSE) %>%
+    kableExtra::column_spec(1) %>%
+    kableExtra::column_spec(2) %>%
+    kableExtra::column_spec(3, width = "30em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(table) %>%
+    flextable::set_caption(caption = "Reasons for consensus rejection (raw answers)") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 3, width = 4.5) # Adjust column width for the third column
+}
 
+# Display the table
+table_output
 ```

### 2024-09-19 appendix_exp3.Rmd
make WORD compatible tables for PUS submission

@@
 
-# Output the table
-kbl(items, booktabs = T, longtable = TRUE,
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_styling
+  table_output <- items %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
       caption = "Science knowledge items", 
-    full_width = T) %>%
-  kable_styling(font_size = 8) %>% 
-  column_spec(1, width = "1em") %>%
-  column_spec(2, width = "10em") %>%
-  column_spec(3, width = "10em") %>%
-  column_spec(4, width = "23em") 
+      full_width = TRUE
+    ) %>%
+    kableExtra::kable_styling(font_size = 8) %>%
+    kableExtra::column_spec(1, width = "1em") %>%
+    kableExtra::column_spec(2, width = "10em") %>%
+    kableExtra::column_spec(3, width = "10em") %>%
+    kableExtra::column_spec(4, width = "23em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(items) %>%
+    flextable::set_caption(caption = "Science knowledge items") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 1, width = 0.8) %>%
+    flextable::width(j = 2, width = 3) %>%
+    flextable::width(j = 3, width = 3) %>%
+    flextable::width(j = 4, width = 6) # Adjust column widths as needed
+}
+
+# Display the table
+table_output
 ```
@@
 
-# apa_table(table,
-#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
-#           longtable = TRUE
-#           )
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection (raw answers)") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2) %>%
-  column_spec(3, width = "30em")
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
+  table_output <- table %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Reasons for consensus rejection (raw answers)"
+    ) %>%
+    kableExtra::kable_paper(full_width = FALSE) %>%
+    kableExtra::column_spec(1) %>%
+    kableExtra::column_spec(2) %>%
+    kableExtra::column_spec(3, width = "30em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(table) %>%
+    flextable::set_caption(caption = "Reasons for consensus rejection (raw answers)") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 3, width = 4.5) # Adjust column width for the third column
+}
+
+# Display the table
+table_output
 
@@
 
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus acceptance (participants who answered 'other')") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "30em")
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
+  table_output <- table %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Reasons for consensus acceptance (participants who answered 'other')"
+    ) %>%
+    kableExtra::kable_paper(full_width = FALSE) %>%
+    kableExtra::column_spec(1) %>%
+    kableExtra::column_spec(2, width = "30em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(table) %>%
+    flextable::set_caption(caption = "Reasons for consensus acceptance (participants who answered 'other')") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 2, width = 4.5) # Adjust column width for the second column
+}
+
+# Display the table
+table_output
 ```
@@
 
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Answers from participants saying they trusted the scientific consens because they verified independently on how they did so.") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2, width = "30em")
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
+  table_output <- table %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Answers from participants saying they trusted the scientific consensus because they verified independently on how they did so."
+    ) %>%
+    kableExtra::kable_paper(full_width = FALSE) %>%
+    kableExtra::column_spec(1) %>%
+    kableExtra::column_spec(2, width = "30em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(table) %>%
+    flextable::set_caption(caption = "Answers from participants saying they trusted the scientific consensus because they verified independently on how they did so.") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 2, width = 4.5) # Adjust column width for the second column
+}
+
+# Display the table
+table_output
 ```

### 2024-09-19 appendix_exp4.Rmd
make WORD compatible tables for PUS submission

@@
 
-# Output the table
-kbl(items_cleaned, booktabs = T, longtable = TRUE,
-    caption = "Science knowledge items", full_width = T) %>%
-  kable_styling(font_size = 8) %>%
-  column_spec(1, width = "1em") %>%
-  column_spec(2, width = "12em") %>%
-  column_spec(3, width = "10em") %>%
-  column_spec(4, width = "12em")
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_styling
+  table_output <- items_cleaned %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Science knowledge items", 
+      full_width = TRUE
+    ) %>%
+    kableExtra::kable_styling(font_size = 8) %>%
+    kableExtra::column_spec(1, width = "1em") %>%
+    kableExtra::column_spec(2, width = "12em") %>%
+    kableExtra::column_spec(3, width = "10em") %>%
+    kableExtra::column_spec(4, width = "12em")
+} else {
+ # For Word: Use apa_table for better Word compatibility
+  table_output <- papaja::apa_table(
+    items_cleaned, 
+    caption = "Science knowledge items"
+  )
+}
+
+# Display the table
+table_output
 ```
@@
 
-# apa_table(table,
-#           align = c("m{1cm}", "m{3cm}", "m{10cm}"), 
-#           longtable = TRUE
-#           )
-kbl(table, booktabs = T, longtable = TRUE,
-    caption = "Reasons for consensus rejection (raw answers)") %>%
-  kable_paper(full_width = F) %>% 
-  column_spec(1) %>%
-  column_spec(2) %>%
-  column_spec(3, width = "30em")
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_paper
+  table_output <- table %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      caption = "Reasons for consensus rejection (raw answers)"
+    ) %>%
+    kableExtra::kable_paper(full_width = FALSE) %>%
+    kableExtra::column_spec(1) %>%
+    kableExtra::column_spec(2) %>%
+    kableExtra::column_spec(3, width = "30em")
+} else {
+  # For Word: Use flextable for better Word compatibility
+  table_output <- flextable::flextable(table) %>%
+    flextable::set_caption(caption = "Reasons for consensus rejection (raw answers)") %>%
+    flextable::autofit() %>%
+    flextable::width(j = 3, width = 4.5) # Adjust column width for the third column
+}
 
+# Display the table
+table_output
 ```

### 2024-09-19 paper.Rmd
make WORD compatible tables for PUS submission

@@
 # Output the table
-knitr::kable(items_cleaned, booktabs = T, longtable = TRUE,
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kable, column_spec, and kable_styling
+  table_output <- items_cleaned %>%
+    knitr::kable(
+      booktabs = TRUE, 
+      longtable = TRUE,
       caption = "Science knowledge items",
-    full_width = T, 
-    col.names = c("", "Study 1-3", "Study 4")) %>%
-  kable_styling(font_size = 8) %>%
-  column_spec(1, width = "2em") %>%
-  column_spec(2, width = "22em")%>%
-  column_spec(3, width = "22em") %>% 
-  footnote(symbol = "Only used in Study 1")
+      full_width = TRUE,
+      col.names = c("", "Study 1-3", "Study 4")
+    ) %>%
+    kableExtra::kable_styling(font_size = 8) %>%
+    kableExtra::column_spec(1, width = "2em") %>%
+    kableExtra::column_spec(2, width = "22em") %>%
+    kableExtra::column_spec(3, width = "22em") %>%
+    kableExtra::footnote(
+      symbol = "Only used in Study 1"
+    )
+} else {
+  # For Word: Use apa_table
+  table_output <- papaja::apa_table(
+    items_cleaned, 
+    caption = "Science knowledge items",
+    note = "*Only used in Study 1",
+    col.names = c("", "Study 1-3", "Study 4")
+    #,align = c("l", "m{8cm}", "m{8cm}")
+  )
+}
+
+# Display the table
+table_output
+
 ```
@@
 
-# Output the table
-kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL,
-    caption = "Conspiracy items", full_width = T) %>%
-  kable_styling(font_size = 8) %>% 
-  column_spec(1, width = "3em") %>%
-  column_spec(2, width = "40em") %>% 
-  footnote(general = "Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.",
+# Create the basic table
+if (knitr::is_latex_output() || knitr::is_html_output()) {
+  # For LaTeX or HTML: Use kbl, column_spec, and kable_styling
+  table_output <- conspiracy_items %>%
+    kbl(
+      booktabs = TRUE, 
+      longtable = TRUE,
+      col.names = NULL,
+      caption = "Conspiracy items",
+      full_width = TRUE
+    ) %>%
+    kableExtra::kable_styling(font_size = 8) %>%
+    kableExtra::column_spec(1, width = "3em") %>%
+    kableExtra::column_spec(2, width = "40em") %>%
+    kableExtra::footnote(
+      general = "Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.",
       escape = FALSE,
       footnote_as_chunk = TRUE,
-           threeparttable = TRUE)
+      threeparttable = TRUE
+    )
+} else {
+  # For Word: Use apa_table
+  table_output <- papaja::apa_table(
+    conspiracy_items, 
+    caption = "Conspiracy items",
+    note = "Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.",
+    col.names = c("", "") # Replace with specific column names if desired
+  )
+}
+
+# Display the table
+table_output
 ```

### 2024-09-19 paper.tex
make WORD compatible tables for PUS submission

@@
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{2em}>{\raggedright\arraybackslash}p{22em}>{\raggedright\arraybackslash}p{22em}}
-\caption{\label{tab:knowledge}Science knowledge items}\\
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:knowledge}Science knowledge items}
+
+\begin{tabular}{m{0.5cm}m{8cm}m{8cm}}
 \toprule
@@
 [The mother's genes; the father's genes; both] & Using bomb-pusle dating with carbon 14, what is the age of the oldest known vertebrate, as established in 2016?
-[138 years; 205 years;  392 years]\\
+[138 years; 205 years;\ \ 392 years]\\
 5 & Do lasers work by focusing sound waves?
@@
 [The same amount; Twice as many; Ten times as many]\\
-\addlinespace
 6 & How long does it take for Earth to go around the sun: one day, one month, or one year?
 [One day; One month; One year] & As predicted by the general theory of relativity, how many times would the Earth keep orbiting if the Sun disappeared, as established in 2012?
-[47 seconds; 8 minutes;  2 hours]\\
+[47 seconds; 8 minutes;\ \ 2 hours]\\
 7 & Are diamonds made of carbon? [Yes; No] & What is the electric charge of the Higgs Boson, as established in 2012? [1.602176634 x 10-19; 0; 3.2x10-19C]\\
@@
 [15 microtesla; 30 microtesla; 45 microtesla]\\
-\addlinespace
-11 & *Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air] & \\
-\bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{*} Only used in Study 1}\\
-\end{longtable}
-\endgroup{}
+11 & *Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air] & NA\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} *Only used in Study 1}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}[para]
-\item \textit{Note: } 
-\item Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.
-\end{TableNotes}
-\begin{longtable}[t]{>{\raggedleft\arraybackslash}p{3em}>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:conspiracy}Conspiracy items}\\
-\toprule
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:conspiracy}Conspiracy items}
+
+\begin{tabular}{lp{10cm}}
+\toprule
+ & \\
+\midrule
 1 & The Apollo moon landings never happened and were staged in a Hollywood film studio.\\
@@
 5 & The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.\\
-\addlinespace
 6 & There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.\\
@@
 \bottomrule
-\insertTableNotes
-\end{longtable}
-\end{ThreePartTable}
-\endgroup{}
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 

### 2024-12-29 appendix_exp1.Rmd
add preprint and qsf files

@@
              stars = TRUE,
-             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values") 
+             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values", 
+             output = "kableExtra") 
 ```

### 2024-12-29 preprint.Rmd
add preprint and qsf files

--- 2024-09-19 paper.Rmd
@@
 shorttitle        : "Does anyone mistrust basic science?"
-
 header-includes:  | # to prevent floats from moving past certain points (for the appendix)
@@
 author: 
-  - name          : ""
-    affiliation   : ""
-
+  - name          : "Jan Pfänder"
+    affiliation   : "1"
+    #role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+  - name          : "Lou Kerzreho"
+    affiliation   : "1"
+    #role:         
+  - name          : "Hugo Mercier"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    email         : "hugo.mercier@gmail.com"
+    address: " "
 affiliation:
-  - id            : ""
-    institution   : ""
-
+
+  - id            : "1"
+    institution   : "Institut Jean Nicod, Département d’études cognitives, ENS, EHESS, PSL University, CNRS, France"
 abstract: |
+
   Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
@@
 This could be leveraged in science communication.
-
 keywords          : Trust in science; conspiracy theories; conspiracy thinking; science knowledge.
-
 floatsintext      : yes
@@
 mask              : no
-
 figurelist        : no
@@
 footnotelist      : no
-
 classoption       : "doc" # "doc" for nice look, "man" for manuscripty
-output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
-For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g.
-set always allow html as true in the yaml heading
-
+output: 
+  papaja::apa6_pdf: 
+    keep_tex: true
+  papaja::apa6_docx: default
 always_allow_html: true
-
 appendix:
-  - "appendix_exp1.Rmd"
-  - "appendix_exp2.Rmd"
-  - "appendix_exp3.Rmd"
-  - "appendix_exp4.Rmd"
-  - "appendix_acceptance_knowledge.Rmd"
-  - "appendix_conspiracy_thinking.Rmd"
-  - "appendix_trust.Rmd"
-  - "appendix_conspiracy_belief.Rmd"
-  - "appendix_knowledge.Rmd"
-  - "appendix_by-question.Rmd"
-
 bibliography: references.bib

### 2025-03-11 paper.Rmd
before final tweaks for re-submission

@@
     affiliation   : ""
-
 affiliation:
+
   - id            : ""
     institution   : ""
-
 abstract: |
+
   Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
@@
 footnotelist      : no
-
 classoption       : "doc" # "doc" for nice look, "man" for manuscripty
-output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
-For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g.
-set always allow html as true in the yaml heading
-
+output: 
+  papaja::apa6_docx: default
+  papaja::apa6_pdf: 
+    keep_tex: true
 always_allow_html: true
-
 appendix:
-  - "appendix_exp1.Rmd"
-  - "appendix_exp2.Rmd"
-  - "appendix_exp3.Rmd"
-  - "appendix_exp4.Rmd"
   - "appendix_acceptance_knowledge.Rmd"
@@
   - "appendix_by-question.Rmd"
-
+  - "appendix_conspiracy-scales.Rmd"
+  - "appendix_exp1.Rmd"
+  - "appendix_exp2.Rmd"
+  - "appendix_exp3.Rmd"
+  - "appendix_exp4.Rmd"
 bibliography: references.bib
@@
 
-Unfortunately, people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; although see @wellcomeglobalmonitorPublicTrustScientists2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
-Rejection of science can take the more extreme form of conspiracy theories, many of which question the scientific consensus on vaccination, climate change, and even the shape of the Earth.
-
-What does this apparent rejection of science actually entail?
+Although recent global evidence suggests that trust in science is moderately high [@colognaTrustScientistsTheir2024], it is far from being at ceiling.
+Large-scale polls have shown that people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
+Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; although see @wellcomeglobalmonitorPublicTrustScientists2021 and @funkPublicConfidenceScientists2020], including the US [@brianAmericansTrustScientists2023], where it is also increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
+
+Besides low answers on general trust in science questions, another indicator of distrust in science is the belief in conspiracy theories that question the scientific consensus on issues such as vaccination, climate change, and even the shape of the Earth.
+Conspiracy theories–in the realm of science or elsewhere–typically accuse a small group of powerful people to pursue nefarious goals in secrecy [@douglasUnderstandingConspiracyTheories2019; @medeSciencerelatedPopulismConceptualizing2020].
+Some of these conspiracy theories are widespread [@rutjensConspiracyBeliefsScience2022].
+In 2023, a survey in eight different countries found that up to 24% of respondents agreed that “climate change is a hoax and scientists touting its existence are lying” [@stockemerUnderstandingClimateChange2024].
+Similarly, in 2021, 40% of Americans believed that “the dangers of genetically-modified foods are being hidden from the public” [down from 45% in 2020, @uscinskiHaveBeliefsConspiracy2022].
+People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories [@lewandowskyRoleConspiracistIdeation2013; @hornseyPsychologicalRootsAntivaccination2018; @hornseyRelationshipsConspiratorialBeliefs2018] and tend to say that they distrust science more generally [@vranicDidMyOwn2022; @stockemerUnderstandingClimateChange2024].
+Despite these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
+Although conspiracy thinking has been identified as a "root cause" of anti-science attitudes [@hornseyAttitudeRootsJiu2017] in the past, these claims rest largely on observational data.
+
+What does this apparent lack of trust in science actually entail?
 Do people who say they do not trust science, or who believe in conspiracy theories at odds with well-established science, reject most of science?
 Or, on the contrary, do they object to a few specific facets of science, while still accepting the overwhelming majority of basic science?
-Answering this question has theoretical implications: if some people reject science wholesale, then it’s possible that their distrust of science causes their rejection of specific scientific theories, or facilitates their acceptance of conspiracy theories; by contrast, if everyone trusts most of science, then the mistrust of specific facets of science is more likely to be post-hoc, a rationalization of other beliefs or behaviors.
-The present question also has practical implications: many communication attempts leverage the scientific consensus [e.g.
+
+A common conception of trust is a willingness to be vulnerable to another party, whether an individual, a group, or an institution [@mayerIntegrativeModelOrganizational1995a; @rousseauIntroductionSpecialTopic1998].
+Accordingly, trust in science has been defined as “one’s willingness to rely on science and scientists (as representatives of the system) despite having a bounded understanding of science” [@wintterlinPredictingPublicTrust2022, p.
+2].
+Past research has disentangled this general concept of trust in science in various ways.
+Some research has identified different components of trust, the number of which varies, but which generally cover an epistemological and ethical dimension [@wilholtEpistemicTrustScience2013; @intemannScienceCommunicationPublic2023].
+For example, @hendriksMeasuringLaypeoplesTrust2015 suggest distinguishing between expertise/competence, integrity, and benevolence, while @besleyReassessingVariablesUsed2021 add openness.
+Other research has highlighted differences in trust between scientific disciplines [@altenmullerExplainingPolarizedTrust2024; @gligoricHowSocialEvaluations2024; @gauchatCulturalCognitiveMappingScientific2018].
+However, little research has assessed trust in specific scientific findings, besides contentious topics such as vaccines [e.g.
+@hornseyPsychologicalRootsAntivaccination2018], climate change [e.g.
+@stockemerUnderstandingClimateChange2024], evolution [e.g.
+@nadelsonTrustScienceScientists2015], genetically modified organisms (GMOs) [e.g.
+@fernbachExtremeOpponentsGenetically2019], or a combinations of such topics [see e.g @lewandowskyRoleConspiracistIdeation2013].
+To the best of our knowledge, no research has investigated the extent to which people trust basic science facts (e.g.
+electrons are smaller than atoms).
+An extensive literature on science literacy has assessed whether people know such facts [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016], but not whether they accept the facts once presented to them.
+
+Why does it matter whether trust in science in general is or is not related to trust in scientific findings?
+First, this question has theoretical implications.
+According to the most prominent explanation of (dis)trust in science–the deficit model–science knowledge is the main driver of attitudes towards science in general.
+A prediction of this model is that average trust in science on specific facts should be strongly associated with general trust in science.
+By contrast, a disconnect between the two would be in line with motivated reasoning accounts of trust in science.
+According to these accounts, science rejection serves to maintain coherence with other beliefs or behaviors [@lewandowskyMotivatedRejectionScience2016; @hornseyWhyFactsAre2020].
+For example, someone might say they don't trust science in general because they are skeptical towards vaccines, not because they actually distrust most of science.
+A prediction of the motivated reasoning account is that general trust in science should be strongly correlated with conspiracy beliefs.
+Second, the present question has practical implications: Many communication attempts leverage the scientific consensus [e.g.
 on vaccination, climate change, etc., for review, see @vanstekelenburgScientificConsensusCommunicationContested2022; see also @veckalov27countryTestCommunicating2024].
@@
 
-Studies of science knowledge provide relevant evidence, as people who answer science knowledge questions correctly presumably trust scientists in this respect.
-The average level of science knowledge is quite low, even in countries with extensive science education.
-For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016].
-However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.
-
 ## The present studies
 
-In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts.
+In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established, consensual scientific facts.
 For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted answer, and asked them whether they accepted it (measuring their trust in basic science).
 We also measured participants’ trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (https://osf.io/8utsj/?view_only=364478f01fc548d0a42b9962918d871d).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (<https://osf.io/8utsj/?view_only=364478f01fc548d0a42b9962918d871d>).
 The differences between the four studies are summarized presently, and the methods are detailed below.
@@
 
-In Study 1, we used questions drawn from questionnaires of scientific knowledge (e.g.
+In Study 1 (data collected on March 5, 2024), we used questions drawn from questionnaires of scientific knowledge (e.g.
 "Are electrons smaller, larger, or the same size as atoms?
@@
 [Earth; Water; Air]"; correct answer: Air).
-In Studies 2 and 3, this last question was removed.
+In Studies 2 (data collected on April 3, 2024) and 3 (data collected on April 22, 2024), this last question was removed.
 The scientific facts used in Studies 1 to 3 represent long-established and basic knowledge.
-In Study 4 we used more recent much less basic scientific discoveries (e.g.
+In Study 4 (data collected on August 13, 2024) we used more recent, much less basic scientific discoveries (e.g.
 "What is the electric charge of the Higgs Boson, as established in 2012?
@@
 
-In Study 1, we simply look at whether participants say that they accept the scientifically consensual answer.
-In the subsequent studies, we asked participants to explain why they disagreed with the scientific consensus.
+In Study 1, we simply looked at whether participants accept the scientifically consensual answer or not.
+In the subsequent studies, we asked participants to explain cases in which they disagreed with the scientific consensus.
 This revealed that a number of participants had made a mistake (misunderstanding, selecting the wrong answer).
@@
 The main goal of the present studies is descriptive: to find out whether participants who report not trusting science, or who believe in conspiracy theories, still accept most well-established scientific facts.
-However, we also tested two directional hypotheses (pre-registered as research questions in the Study 1):
+However, based on our literature review, we also tested two directional hypotheses (pre-registered as research questions in the Study 1):
 
@@
                               include.lowest = TRUE, right = FALSE))
+
+# Compute the correlation between conspiracy belief and general trust
+cor_trust_conspiracy_data <- combined_data |> 
+  group_by(id) |> 
+  summarise(BCTI_avg = mean(BCTI_avg, na.rm=TRUE), 
+            wgm_sciencegeneral = mean(wgm_sciencegeneral, na.rm=TRUE))
+
+cor_trust_conspiracy <- cor.test(cor_trust_conspiracy_data$wgm_sciencegeneral, cor_trust_conspiracy_data$BCTI_avg,
+  use = "complete.obs") %>% text_ready()
 
@@
     mutate(mean_acceptance_rate = paste0(mean_acceptance_rate*100, " %"), 
-           mean_knowledge_rate = paste0(mean_knowledge_rate*100, " %"))
+           mean_knowledge_rate = paste0(mean_knowledge_rate*100, " %")), 
+  # overall correlation between conspiracy belief and general trust
+  cor_trust_conspiracy = cor_trust_conspiracy 
 )
@@
 We selected these items so that we could compare the answers in our sample to global survey results.
-We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM, sections \@ref(exp3) and \@ref(exp4)).
+We find that all three items are highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures[^2] (see ESM).
+
+[^2]: With two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM, sections \@ref(exp3) and \@ref(exp4)
 
@@
 
+How do knowledge and acceptance relate to declared general trust in science and conspiracy belief (two strongly correlated variables, pooled r = `r combined_descriptives$cor_trust_conspiracy$estimate`, p `r combined_descriptives$cor_trust_conspiracy$p.value`)?
+
 Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`), but a less consistent relation between trust in science and science knowledge (Studies 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`).
@@
 Regarding H2, the results are mixed for the relation of conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) with both acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`) and science knowledge (Studies 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`).
+
+n an exploratory analysis, we find that
+
+While actual acceptance of scientific knowledge is only weakly correlated with trust in science,
 
@@
 Based on the answers, we created five categories (Table \@ref(tab:justifications)).
-All answers can be read in the ESM, sections \@ref(exp2) to \@ref(exp4).
+All individual answers can be accessed in cleaned data sheets via the OSF project page.
 
@@
 A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (Studies 3: `r exp3_descriptives$acceptance_n`; 4: `r exp4_descriptives$acceptance_n`) participants answered this question.
-There were more participants saying they accepted the scientific consensus because they independently verified the fact (Studies 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than participants saying it was because they trust scientists (Studies 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`)[^2].
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Studies 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than participants saying it was because they trust scientists (Studies 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`)[^3].
 Answers to a question about how they had done so can be found in the ESM, sections \@ref(exp3) and \@ref(exp4).
 
-[^2]: `r exp3_descriptives$acceptance_by_reason[["other"]]$share` in Study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in Study 4 answered with "other" and gave an open-ended explanation (see ESM, sections \@ref(exp3) and \@ref(exp4)).
+[^3]: `r exp3_descriptives$acceptance_by_reason[["other"]]$share` in Study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in Study 4 answered with "other" and gave an open-ended explanation.
 
@@
 We did not find a difference regarding acceptance (Studies 3: $\hat{\beta}_{\text{Acceptance}}$ = `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value` on a scale from 0 to 1; 4: $\hat{\beta}_{\text{Acceptance}}$ = `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value`) or regarding beliefs in conspiracy theories (Studies 3: $\hat{\beta}_{\text{BCTI}}$ = `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 9; 4: $\hat{\beta}_{\text{BCTI}}$ = `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value`).
-We also did not find a difference in time spent on the survey in Study 3 ($\hat{\beta}_{\text{Time}}$ = `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp3_wide$duration_mins)` mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average two minutes less on the survey ($\hat{\beta}_{\text{Time}}$ = `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp4_wide$duration_mins)` mins).[^3] In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+We also did not find a difference in time spent on the survey in Study 3 ($\hat{\beta}_{\text{Time}}$ = `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp3_wide$duration_mins)` mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average two minutes less on the survey ($\hat{\beta}_{\text{Time}}$ = `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp4_wide$duration_mins)` mins).[^4] In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links provided--a behavior that you would expect from people who report verifying facts independently.
 On average, participants clicked only on `r exp4_descriptives$reason_means$'trust in scientists'$sum_link_clicks_mean` links (out of 20 possible clicks) and there was no difference between the two groups ($\hat{\beta}_{\text{Clicks}}$ = `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$p.value`).
 
-[^3]: For these analyses, we excluded outliers who took over 30 mins for the survey, which was estimated to take around 10 mins, and for which the median time was seven minutes.
+[^4]: For these analyses, we excluded outliers who took over 30 mins for the survey, which was estimated to take around 10 mins, and for which the median time was seven minutes.
 As a result, we excluded one participant in Study 3 and four in Study 4.
@@
 
-US participants were asked whether they accepted scientifically consensual answers on basic science questions.We found quasi-universal acceptance of basic science, with an overall rate of acceptance of `r combined_descriptives$mean_acceptance$Yes$share`, which remained very high for participants who declared not trusting science at all (`r combined_descriptives$acceptance_by_trust$'1'$Yes$share` of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (`r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share` of acceptance).
-
-Overall, the present findings support the motivated rejection of science model [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p.
+In four studies, we asked US participants whether they accepted scientifically consensual answers on basic science questions.
+We found quasi-universal acceptance of basic science, with an overall rate of acceptance of `r combined_descriptives$mean_acceptance$Yes$share`, which remained very high for participants who declared not trusting science at all (`r combined_descriptives$acceptance_by_trust$'1'$Yes$share` of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (`r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share` of acceptance).
+
+This disconnect between declared general trust in science and average trust in (basic) scientific findings goes against predictions from the knowledge–attitudes model of trust in science [see, e.g., @bauerWhatCanWe2007], which posits that knowledge of scientific results is the main cause of trust in science.
+In line with past research [@allumScienceKnowledgeAttitudes2008], we find that both knowledge and acceptance of science are only weakly correlated with general trust in science.
+This supports contemporary definitions of science literacy, which suggest that for science literacy to be a meaningful concept for science attitudes, it needs to go beyond measuring knowledge of isolated science facts [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016].
+
+Instead, our findings align with a motivated reasoning account of trust in science [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p.
 217).
-A number of participants in our studies rejected basic tenets of science (on evolution, the shape of the earth, etc.).
-However, they still accepted the vast majority of basic scientific knowledge presented to them.
-This suggests that these participants had reasons to reject specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
-These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
-These results are in line with several recent studies in which conspiracy theorists are not less susceptible to social influence in general [@pummererConspiracyBeliefsMajority; @altayConspiracyBelieversClaim2023].
-
-Our findings are also relevant for the knowledge–attitudes model of trust in science [see, e.g., @bauerWhatCanWe2007].
-The fact that even people with low science knowledge trust most of basic science might be taken as showing the power of even a modicum of science knowledge; on the other hand, it also means that attempts at increasing science knowledge in general might not have much effect on trust in science, by contrast with targeting the specific beliefs people are motivated to reject.
-
-Finally, some of the present results also speak to the alienation model [@gauchatCulturalAuthorityScience2011], and more specifically to the need for epistemic autonomy [@frickerTestimonyEpistemicAutonomy2006].
+A number of participants in our studies endorsed specific conspiracy theories questioning basic tenets of science (on evolution, the shape of the earth, etc.), while still accepting the vast majority of basic scientific facts presented to them.
+This suggests that these participants had reasons to reject only specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
+Consistent with this explanation, we found a strong association between belief in the anti-science conspiracy theories and general trust in science.
+
+Some of the present results also speak to the alienation model [@gauchatCulturalAuthorityScience2011], and more specifically to the need for epistemic autonomy [@frickerTestimonyEpistemicAutonomy2006].
 Declaring not trusting science, or endorsing conspiracy theories [@harrisConspiracyTheoriesPopulism2023] might reflect a desire to maintain epistemic autonomy and not appear to ‘blindly’ accept epistemic authority.
-The acceptance of basic science facts would be reconciled with this need to appear epistemically autonomous by the suggestion that the acceptance stems from independent evaluation instead of trust (however implausible that might be: it’s not clear how participants could independently verify the ratio of glial cells to neurons, say, and few participants appeared to have engaged in even simple forms of verification).
-
-In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that can be tapped into in science communication.
-Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the ‘consensus gaps’ observed in these domains [on climate change, see, e.g.
+Such a need to appear epistemically autonomous would be reconciled with the acceptance of basic science facts if it stems from independent evaluation instead of trust.
+The majority of participants in Studies 3 and 4, two samples of vaccine skeptical participants who scored high across a range of different conspiracy beliefs (see ESM), claimed that this was their case, however implausible that might be: it’s not clear how participants could independently verify, say, the ratio of glial cells to neurons.
+These participants did also not engage any more in simple forms of verification, i.e.
+clicking on sources provided in the answers.
+This aligns with other results on conspiracy thinking: Several recent studies have shown that, although they claim not to be, conspiracy theorists are in fact just as susceptible to social influence as others [@pummererConspiracyBeliefsMajority2024; @altayConspiracyBelieversClaim2023].
+
+In applied terms, the present results have implications for science communication.
+Across various domains of science knowledge such as climate change [@veckalov27countryTestCommunicating2024] or vaccination [@salmonVaccineHesitancyCauses2015], researchers have observed a consensus gap: a gap between the scientific consensus and public opinion.
+For example, some of the most worrying consensus gaps relate to climate change, as substantial segments of the population disagree with scientists on what is happening and what to do about it [e.g., @eganClimateChangeUS2017].
+Yet, we found that for much of basic, non-contentious science, such consensus gaps are absent: The vast majority of our participants–including two vaccine-skeptic attitudes samples (Studies 3 and 4)--did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+Since even people who distrust science or reject contentious science facts appear to accept most of basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the consensus gaps observed in these domains [on climate change, see, e.g.
 @ranneyClimateChangeConceptual2016].
@@
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
-We hope that future studies will extend our findings to other populations.
+While our results lend support to motivated reasoning accounts of trust in science, they leave several important questions unaddressed, in particular: Why do people reject specific science beliefs but not others?
+Suggestions have already been made for a number of issues such as vaccination [@mitonCognitiveObstaclesProVaccination2015], GMOs [@blanckeFatalAttractionIntuitive2015], or nuclear energy [@hacquinDisgustSensitivityPublic2021].
+However, research is still needed to better understand what motivates these rejections [see e.g., @hornseyWhyFactsAre2020].
+Moreover, rejection of scientific facts can manifest in various ways, for example calling into question the integrity of scientists [@medeSciencerelatedPopulismConceptualization2023a; @medeWhoSupportsSciencerelated2022], or denying the very possibility of scientifically investigating certain issues [@munroScientificImpotenceExcuse2010].
+The classification of justifications we present here does not address these processes in detail.
+Finally, by design, most of the basic science knowledge presented in this study didn't directly relate to anything controversial for most people.
+Future studies could look at basic science that does relate to science topics which are the object of public and policy-relevant debates (e.g.
+GM food, climate science, vaccines), or scientific findings for which there is a more obvious potential conflict of interest (e.g.
+non-publicly funded medical research).
 
@@
 
+\newpage
+
 # (APPENDIX) Appendix {.unnumbered}
@@
 
+```{r child = "appendix_conspiracy_thinking.Rmd"}
+```
+
+\clearpage
+
 ```{r child = "appendix_trust.Rmd"}
@@
 
-```{r child = "appendix_conspiracy_thinking.Rmd"}
+```{r child = "appendix_conspiracy_belief.Rmd"}
 ```
@@
 
-```{r child = "appendix_conspiracy_belief.Rmd"}
+```{r child = "appendix_knowledge.Rmd"}
 ```
@@
 
-```{r child = "appendix_knowledge.Rmd"}
+```{r child = "appendix_by-question.Rmd"}
 ```
@@
 
-```{r child = "appendix_by-question.Rmd"}
+```{r child = "appendix_conspiracy-scales.Rmd"}
 ```

### 2025-03-11 paper.tex
before final tweaks for re-submission

--- 2024-09-19 paper.tex
@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 % definitions for citeproc citations
@@
   {\xapptocmd\section
-    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
     {}{\InnerPatchFailed}%
@@
 {}{\PatchFailed}
+\makeatother
 \keywords{Trust in science; conspiracy theories; conspiracy thinking; science knowledge.}
@@
 
-Unfortunately, people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia (Wellcome Global Monitor, 2018, 2020).
-Moreover, trust in science has recently been declining in some countries (Algan et al., 2021; although see Wellcome Global Monitor, 2021), including the US (Brian \& Tyson, 2023), where it is also increasingly polarizing (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
-Rejection of science can take the more extreme form of conspiracy theories, many of which question the scientific consensus on vaccination, climate change, and even the shape of the Earth.
-
-What does this apparent rejection of science actually entail?
+Although recent global evidence suggests that trust in science is moderately high (Cologna et al., 2024), it is far from being at ceiling.
+Large-scale polls have shown that people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia (Wellcome Global Monitor, 2018, 2020).
+Moreover, trust in science has recently been declining in some countries Funk \& Kennedy (2020), including the US (Brian \& Tyson, 2023), where it is also increasingly polarizing (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
+
+Besides low answers on general trust in science questions, another indicator of distrust in science is the belief in conspiracy theories that question the scientific consensus on issues such as vaccination, climate change, and even the shape of the Earth.
+Conspiracy theories--in the realm of science or elsewhere--typically accuse a small group of powerful people to pursue nefarious goals in secrecy (Douglas et al., 2019; Mede \& Schäfer, 2020).
+Some of these conspiracy theories are widespread (Rutjens \& Većkalov, 2022).
+In 2023, a survey in eight different countries found that up to 24\% of respondents agreed that ``climate change is a hoax and scientists touting its existence are lying'' (Stockemer \& Bordeleau, 2024).
+Similarly, in 2021, 40\% of Americans believed that ``the dangers of genetically-modified foods are being hidden from the public'' (down from 45\% in 2020, Uscinski et al., 2022).
+People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories (Hornsey, Harris, \& Fielding, 2018b, 2018a; Lewandowsky, Gignac, \& Oberauer, 2013) and tend to say that they distrust science more generally (Stockemer \& Bordeleau, 2024; Vranic, Hromatko, \& Tonković, 2022).
+Despite these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
+Although conspiracy thinking been identified as a ``root cause'' of anti-science attitudes (Hornsey \& Fielding, 2017) in the past, these claims rest largely on observational data.
+
+What does this apparent lack of trust in science actually entail?
 Do people who say they do not trust science, or who believe in conspiracy theories at odds with well-established science, reject most of science?
 Or, on the contrary, do they object to a few specific facets of science, while still accepting the overwhelming majority of basic science?
-Answering this question has theoretical implications: if some people reject science wholesale, then it's possible that their distrust of science causes their rejection of specific scientific theories, or facilitates their acceptance of conspiracy theories; by contrast, if everyone trusts most of science, then the mistrust of specific facets of science is more likely to be post-hoc, a rationalization of other beliefs or behaviors.
-The present question also has practical implications: many communication attempts leverage the scientific consensus (e.g.~on vaccination, climate change, etc., for review, see Van Stekelenburg, Schaap, Veling, Van 'T Riet, \& Buijzen, 2022; see also Većkalov et al., 2024).
+
+A common conception of trust is a willingness to be vulnerable to another party, whether an individual, a group, or an institution (Mayer, Davis, \& Schoorman, 1995; Rousseau, Sitkin, Burt, \& Camerer, 1998).
+Accordingly, trust in science has been defined as ``one's willingness to rely on science and scientists (as representatives of the system) despite having a bounded understanding of science'' (Wintterlin et al., 2022, p.
+2).
+Past research has disentangled this general concept of trust in science in various ways.
+Some research has identified different components of trust, the number of which varies, but which generally cover an epistemological and ethical dimension (Intemann, 2023; Wilholt, 2013).
+For example, Hendriks, Kienhues, and Bromme (2015) suggest distinguishing between expertise/competence, integrity, and benevolence, while Besley, Lee, and Pressgrove (2021) add openness.
+Other research has highlighted differences in trust between scientific disciplines (Altenmüller, Wingen, \& Schulte, 2024; Gauchat \& Andrews, 2018; Gligorić, Kleef, \& Rutjens, 2024).
+However, little research has assessed trust in specific scientific findings, besides contentious topics such as vaccines (e.g.
+Hornsey et al., 2018b), climate change (e.g.
+Stockemer \& Bordeleau, 2024), evolution (e.g.
+Nadelson \& Hardy, 2015), genetically modified organisms (GMOs) (e.g.
+Fernbach, Light, Scott, Inbar, \& Rozin, 2019), or a combinations of such topics (see e.g Lewandowsky et al., 2013).
+To the best of our knowledge, no research has investigated the extent to which people trust basic science facts (e.g.~electrons are smaller than atoms).
+An extensive literature on science literacy has assessed whether people know such facts (National Academies of Sciences, Engineering, and Medicine, 2016), but not whether they accept the facts once presented to them.
+
+Why does it matter whether trust in science in general is or is not related to trust in scientific findings?
+First, this question has theoretical implications.
+According to the most prominent explanation of (dis)trust in science--the deficit model--science knowledge is the main driver of attitudes towards science in general.
+A prediction of this model is that average trust in science on specific facts should be strongly associated with general trust in science.
+By contrast, a disconnect between the two would be in line with motivated reasoning accounts of trust in science.
+According to these accounts, science rejection serves to maintain coherence with other beliefs or behaviors (Hornsey, 2020; Lewandowsky \& Oberauer, 2016).
+For example, someone might say they don't trust science in general because they are skeptical towards vaccines, not because they actually distrust most of science.
+A prediction of the motivated reasoning account is that general trust in science should be strongly correlated with conspiracy beliefs.
+Second, the present question has practical implications: Many communication attempts leverage the scientific consensus (e.g.~on vaccination, climate change, etc., for review, see Van Stekelenburg, Schaap, Veling, Van 'T Riet, \& Buijzen, 2022; see also Većkalov et al., 2024).
 These attempts are more likely to be successful if everyone trusts basic science than if some people reject science wholesale.
 
-Studies of science knowledge provide relevant evidence, as people who answer science knowledge questions correctly presumably trust scientists in this respect.
-The average level of science knowledge is quite low, even in countries with extensive science education.
-For example, only 55\% in the US (in 2014) and only 46\% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses (National Academies of Sciences, Engineering, and Medicine, 2016).
-However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.
-
 \subsection{The present studies}\label{the-present-studies}
 
-In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts.
+In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established, consensual scientific facts.
 For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted answer, and asked them whether they accepted it (measuring their trust in basic science).
@@
 
-In Study 1, we used questions drawn from questionnaires of scientific knowledge (e.g.~``Are electrons smaller, larger, or the same size as atoms?
+In Study 1 (data collected on March 5, 2024), we used questions drawn from questionnaires of scientific knowledge (e.g.~``Are electrons smaller, larger, or the same size as atoms?
 {[}Smaller; Same size; Larger{]}''), supplemented by a `trick' question (``Where do trees mainly draw the materials with which they create their mass?
 {[}Earth; Water; Air{]}''; correct answer: Air).
-In Studies 2 and 3, this last question was removed.
+In Studies 2 (data collected on April 3, 2024) and 3 (data collected on April 22, 2024), this last question was removed.
 The scientific facts used in Studies 1 to 3 represent long-established and basic knowledge.
-In Study 4 we used more recent much less basic scientific discoveries (e.g.~``What is the electric charge of the Higgs Boson, as established in 2012?
+In Study 4 (data collected on August 13, 2024) we used more recent, much less basic scientific discoveries (e.g.~``What is the electric charge of the Higgs Boson, as established in 2012?
 {[}1.602176634 × 10-19; 0; 3.2×10-19C{]}''; correct answer: 0, i.e.~electrically neutral).
@@
 We selected these items so that we could compare the answers in our sample to global survey results.
-We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM, sections \ref{exp3} and \ref{exp4}).
+We find that all three items are highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures\footnote{With two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM, sections \ref{exp3} and \ref{exp4}} (see ESM).
 
@@
 
+How do knowledge and acceptance relate to declared general trust in science and conspiracy belief (two strongly correlated variables, pooled r = -0.64, p \textless{} .001)?
+
 Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = 0.27, p \textless{} .001; 2: r = 0.30, p \textless{} .001; 3: r = 0.20, p = 0.019; 4: r = 0.16, p = 0.029), but a less consistent relation between trust in science and science knowledge (Studies 1: r = 0.29, p \textless{} .001; 2: r = 0.28, p \textless{} .001; 3: r = 0.14, p = 0.100; 4: r = 0.10, p = 0.144).
@@
 Regarding H2, the results are mixed for the relation of conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) with both acceptance of the scientific consensus (Studies 1: r = -0.33, p \textless{} .001; 2: r = -0.37, p \textless{} .001; 3: r = -0.02, p = 0.788; 4: r = -0.05, p = 0.498) and science knowledge (Studies 1: r = -0.38, p \textless{} .001; 2: r = -0.40, p \textless{} .001; 3: r = -0.16, p = 0.055; 4: r = -0.02, p = 0.742).
+
+n an exploratory analysis, we find that
+
+While actual acceptance of scientific knowledge is only weakly correlated with trust in science,
 
@@
 Based on the answers, we created five categories (Table \ref{tab:justifications}).
-All answers can be read in the ESM, sections \ref{exp2} to \ref{exp4}.
+All individual answers can be accessed in cleaned data sheets via the OSF project page.
 
@@
 A total of 320 (Studies 3: 122; 4: 198) participants answered this question.
-There were more participants saying they accepted the scientific consensus because they independently verified the fact (Studies 3: 47.5\%; 4: 47\% ), than participants saying it was because they trust scientists (Studies 3: 41.8\%; 4: 36.4\%)\footnote{10.7\% in Study 3 and 16.7\% in Study 4 answered with ``other'' and gave an open-ended explanation (see ESM, sections \ref{exp3} and \ref{exp4}).}.
+There were more participants saying they accepted the scientific consensus because they independently verified the fact (Studies 3: 47.5\%; 4: 47\% ), than participants saying it was because they trust scientists (Studies 3: 41.8\%; 4: 36.4\%)\footnote{10.7\% in Study 3 and 16.7\% in Study 4 answered with ``other'' and gave an open-ended explanation.}.
 Answers to a question about how they had done so can be found in the ESM, sections \ref{exp3} and \ref{exp4}.
@@
 As a result, we excluded one participant in Study 3 and four in Study 4.
-Significance levels are not affected by these exclusions.} In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently.
+Significance levels are not affected by these exclusions.} In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links provided--a behavior that you would expect from people who report verifying facts independently.
 On average, participants clicked only on 1.36 links (out of 20 possible clicks) and there was no difference between the two groups (\(\hat{\beta}_{\text{Clicks}}\) = -1.09, p = 0.062).
@@
 
-US participants were asked whether they accepted scientifically consensual answers on basic science questions.We found quasi-universal acceptance of basic science, with an overall rate of acceptance of 95.1 \%, which remained very high for participants who declared not trusting science at all (87.3 \% of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (87.2 \% of acceptance).
-
-Overall, the present findings support the motivated rejection of science model (Lewandowsky \& Oberauer, 2016), in which ``people tend to reject findings that threaten their core beliefs or worldview'' (p.~217).
-A number of participants in our studies rejected basic tenets of science (on evolution, the shape of the earth, etc.).
-However, they still accepted the vast majority of basic scientific knowledge presented to them.
-This suggests that these participants had reasons to reject specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
-These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
-These results are in line with several recent studies in which conspiracy theorists are not less susceptible to social influence in general (Altay, Nera, Ejaz, Schöpfer, \& Tomas, 2023; Pummerer, Fock, Winter, \& Sassenberg, n.d.).
-
-Our findings are also relevant for the knowledge--attitudes model of trust in science (see, e.g., Bauer, Allum, \& Miller, 2007).
-The fact that even people with low science knowledge trust most of basic science might be taken as showing the power of even a modicum of science knowledge; on the other hand, it also means that attempts at increasing science knowledge in general might not have much effect on trust in science, by contrast with targeting the specific beliefs people are motivated to reject.
-
-Finally, some of the present results also speak to the alienation model (Gauchat, 2011), and more specifically to the need for epistemic autonomy (Fricker, 2006).
+In four studies, we asked US participants whether they accepted scientifically consensual answers on basic science questions.
+We found quasi-universal acceptance of basic science, with an overall rate of acceptance of 95.1 \%, which remained very high for participants who declared not trusting science at all (87.3 \% of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (87.2 \% of acceptance).
+
+This disconnect between declared general trust in science and average trust in (basic) scientific findings goes against predictions from the knowledge--attitudes model of trust in science (see, e.g., Bauer, Allum, \& Miller, 2007), which posits that knowledge of scientific results is the main cause of trust in science.
+In line with past research (Allum et al., 2008), we find that both knowledge and acceptance of science are only weakly correlated with general trust in science.
+This supports contemporary definitions of science literacy, which suggest that for science literacy to be a meaningful concept for science attitudes, it needs to go beyond measuring knowledge of isolated science facts (National Academies of Sciences, Engineering, and Medicine, 2016).
+
+Instead, our findings align with a motivated reasoning account of trust in science (Lewandowsky \& Oberauer, 2016), in which ``people tend to reject findings that threaten their core beliefs or worldview'' (p.~217).
+A number of participants in our studies endorsed specific conspiracy theories questioning basic tenets of science (on evolution, the shape of the earth, etc.), while still accepting the vast majority of basic scientific facts presented to them.
+This suggests that these participants had reasons to reject only specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
+Consistent with this explanation, we found a strong association between belief in the anti-science conspiracy theories and general trust in science.
+
+Some of the present results also speak to the alienation model (Gauchat, 2011), and more specifically to the need for epistemic autonomy (Fricker, 2006).
 Declaring not trusting science, or endorsing conspiracy theories (Harris, 2023) might reflect a desire to maintain epistemic autonomy and not appear to `blindly' accept epistemic authority.
-The acceptance of basic science facts would be reconciled with this need to appear epistemically autonomous by the suggestion that the acceptance stems from independent evaluation instead of trust (however implausible that might be: it's not clear how participants could independently verify the ratio of glial cells to neurons, say, and few participants appeared to have engaged in even simple forms of verification).
-
-In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that can be tapped into in science communication.
-Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the `consensus gaps' observed in these domains (on climate change, see, e.g.
+Such a need to appear epistemically autonomous would be reconciled with the acceptance of basic science facts if it stems from independent evaluation instead of trust.
+The majority of participants in Studies 3 and 4, two samples of vaccine skeptical participants who scored high across a range of different conspiracy beliefs (see ESM), claimed that this was their case, however implausible that might be: it's not clear how participants could independently verify, say, the ratio of glial cells to neurons.
+These participants did also not engage any more in simple forms of verification, i.e.~clicking on sources provided in the answers.
+This aligns with other results on conspiracy thinking: Several recent studies have shown that, although they claim not to be, conspiracy theorists are in fact just as susceptible to social influence as others (Altay, Nera, Ejaz, Schöpfer, \& Tomas, 2023; Pummerer, Fock, Winter, \& Sassenberg, 2024).
+
+In applied terms, the present results have implications for science communication.
+Across various domains of science knowledge such as climate change (Većkalov et al., 2024) or vaccination (Salmon, Dudley, Glanz, \& Omer, 2015), researchers have observed a consensus gap: a gap between the scientific consensus and public opinion.
+For example, some of the most worrying consensus gaps relate to climate change, as substantial segments of the population disagree with scientists on what is happening and what to do about it (e.g., Egan \& Mullin, 2017).
+Yet, we found that for much of basic, non-contentious science, such consensus gaps are absent: The vast majority of our participants--including two vaccine-skeptic attitudes samples (Studies 3 and 4)--did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+Since even people who distrust science or reject contentious science facts appear to accept most of basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the consensus gaps observed in these domains (on climate change, see, e.g.
 Ranney \& Clark, 2016).
@@
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
-We hope that future studies will extend our findings to other populations.
+While our results lend support to motivated reasoning accounts of trust in science, they leave several important questions unaddressed, in particular: Why do people reject specific science beliefs but not others?
+Suggestions have already been made for a number of issues such as vaccination (Miton \& Mercier, 2015), GMOs (Blancke, Van Breusegem, De Jaeger, Braeckman, \& Van Montagu, 2015), or nuclear energy (Hacquin, Altay, Aarøe, \& Mercier, 2021).
+However, research is still needed to better understand what motivates these rejections (see e.g., Hornsey, 2020).
+Moreover, rejection of scientific facts can manifest in various ways, for example calling into question the integrity of scientists (Mede, 2023; Mede, Schäfer, Metag, \& Klinger, 2022), or denying the very possibility of scientifically investigating certain issues (Munro, 2010).
+The classification of justifications we present here does not address these processes in detail.
+Finally, by design, most of the basic science knowledge presented in this study didn't directly relate to anything controversial for most people.
+Future studies could look at basic science that does relate to science topics which are the object of public and policy-relevant debates (e.g.~GM food, climate science, vaccines), or scientific findings for which there is a more obvious potential conflict of interest (e.g.~non-publicly funded medical research).
 
@@
 
+\bibitem[\citeproctext]{ref-altenmullerExplainingPolarizedTrust2024}
+Altenmüller, M.
+S., Wingen, T., \& Schulte, A.
+(2024).
+Explaining Polarized Trust in Scientists: A Political Stereotype-Approach.
+\emph{Science Communication}, \emph{46}(1), 92--115.
+\url{https://doi.org/10.1177/10755470231221770}
+
 \bibitem[\citeproctext]{ref-bauerWhatCanWe2007}
@@
 
+\bibitem[\citeproctext]{ref-besleyReassessingVariablesUsed2021}
+Besley, J.
+C., Lee, N.
+M., \& Pressgrove, G.
+(2021).
+Reassessing the Variables Used to Measure Public Perceptions of Scientists.
+\emph{Science Communication}, \emph{43}(1), 3--32.
+\url{https://doi.org/10.1177/1075547020949547}
+
+\bibitem[\citeproctext]{ref-blanckeFatalAttractionIntuitive2015}
+Blancke, S., Van Breusegem, F., De Jaeger, G., Braeckman, J., \& Van Montagu, M.
+(2015).
+Fatal attraction: the intuitive appeal of GMO opposition.
+\emph{Trends in Plant Science}, \emph{20}(7), 414--418.
+\url{https://doi.org/10.1016/j.tplants.2015.03.011}
+
 \bibitem[\citeproctext]{ref-brianAmericansTrustScientists2023}
@@
 
+\bibitem[\citeproctext]{ref-douglasUnderstandingConspiracyTheories2019}
+Douglas, K.
+M., Uscinski, J.
+E., Sutton, R.
+M., Cichocka, A., Nefes, T., Ang, C.
+S., \& Deravi, F.
+(2019).
+Understanding Conspiracy Theories.
+\emph{Political Psychology}, \emph{40}(S1), 3--35.
+\url{https://doi.org/10.1111/pops.12568}
+
 \bibitem[\citeproctext]{ref-durantPublicUnderstandingScience1989}
 Durant, J. R., Evans, G. A., \& Thomas, G. P. (1989). The public understanding of science. \emph{Nature}, \emph{340}(6228), 11--14. \url{https://doi.org/10.1038/340011a0}
+
+\bibitem[\citeproctext]{ref-eganClimateChangeUS2017}
+Egan, P. J., \& Mullin, M. (2017). Climate change: US public opinion. \emph{Annual Review of Political Science}, \emph{20}, 209--227.
+
+\bibitem[\citeproctext]{ref-fernbachExtremeOpponentsGenetically2019}
+Fernbach, P.
+M., Light, N., Scott, S.
+E., Inbar, Y., \& Rozin, P.
+(2019).
+Extreme opponents of genetically modified foods know the least but think they know the most.
+\emph{Nature Human Behaviour}, \emph{3}(3), 251--256.
+\url{https://doi.org/10.1038/s41562-018-0520-3}
 
@@
 \emph{5 key findings about public trust in scientists in the u.s.} Retrieved from \url{https://www.pewresearch.org/fact-tank/2019/08/05/5-key-findings-about-public-trust-in-scientists-in-the-u-s/}
+
+\bibitem[\citeproctext]{ref-funkPublicConfidenceScientists2020}
+Funk, C., \& Kennedy, B. (2020). \emph{Public confidence in scientists has remained stable for decades}.
 
@@
 
+\bibitem[\citeproctext]{ref-gauchatCulturalCognitiveMappingScientific2018}
+Gauchat, G., \& Andrews, K.
+T.
+(2018).
+The Cultural-Cognitive Mapping of Scientific Professions.
+\emph{American Sociological Review}, \emph{83}(3), 567--595.
+\url{https://doi.org/10.1177/0003122418773353}
+
+\bibitem[\citeproctext]{ref-gligoricHowSocialEvaluations2024}
+Gligorić, V., Kleef, G.
+A.
+van, \& Rutjens, B.
+T.
+(2024).
+How social evaluations shape trust in 45 types of scientists.
+\emph{PLOS ONE}, \emph{19}(4), e0299621.
+\url{https://doi.org/10.1371/journal.pone.0299621}
+
+\bibitem[\citeproctext]{ref-hacquinDisgustSensitivityPublic2021}
+Hacquin, A.-S., Altay, S., Aarøe, L., \& Mercier, H. (2021). Disgust sensitivity and public opinion on nuclear energy. \emph{Journal of Environmental Psychology}, 101749.
+
 \bibitem[\citeproctext]{ref-harrisConspiracyTheoriesPopulism2023}
 Harris, K. R. (2023). Conspiracy Theories, Populism, and Epistemic Autonomy. \emph{Journal of the American Philosophical Association}, \emph{9}(1), 21--36. \url{https://doi.org/10.1017/apa.2021.44}
+
+\bibitem[\citeproctext]{ref-hendriksMeasuringLaypeoplesTrust2015}
+Hendriks, F., Kienhues, D., \& Bromme, R.
+(2015).
+Measuring Laypeople{'}s Trust in Experts in a Digital Age: The Muenster Epistemic Trustworthiness Inventory (METI).
+\emph{PLOS ONE}, \emph{10}(10), e0139309.
+\url{https://doi.org/10.1371/journal.pone.0139309}
+
+\bibitem[\citeproctext]{ref-hornseyWhyFactsAre2020}
+Hornsey, M.
+J.
+(2020).
+Why Facts Are Not Enough: Understanding and Managing the Motivated Rejection of Science.
+\emph{Current Directions in Psychological Science}, \emph{29}(6), 583--591.
+\url{https://doi.org/10.1177/0963721420969364}
+
+\bibitem[\citeproctext]{ref-hornseyAttitudeRootsJiu2017}
+Hornsey, M.
+J., \& Fielding, K.
+S.
+(2017).
+Attitude roots and Jiu Jitsu persuasion: Understanding and overcoming the motivated rejection of science.
+\emph{American Psychologist}, \emph{72}(5), 459--473.
+\url{https://doi.org/10.1037/a0040437}
+
+\bibitem[\citeproctext]{ref-hornseyRelationshipsConspiratorialBeliefs2018}
+Hornsey, M.
+J., Harris, E.
+A., \& Fielding, K.
+S.
+(2018a).
+Relationships among conspiratorial beliefs, conservatism and climate scepticism across nations.
+\emph{Nature Climate Change}, \emph{8}(7), 614--620.
+\url{https://doi.org/10.1038/s41558-018-0157-2}
+
+\bibitem[\citeproctext]{ref-hornseyPsychologicalRootsAntivaccination2018}
+Hornsey, M.
+J., Harris, E.
+A., \& Fielding, K.
+S.
+(2018b).
+The psychological roots of anti-vaccination attitudes: A 24-nation investigation.
+\emph{Health Psychology}, \emph{37}(4), 307--315.
+\url{https://doi.org/10.1037/hea0000586}
+
+\bibitem[\citeproctext]{ref-intemannScienceCommunicationPublic2023}
+Intemann, K. (2023). Science communication and public trust in science. \emph{Interdisciplinary Science Reviews}, \emph{48}(2), 350--365. \url{https://doi.org/10.1080/03080188.2022.2152244}
 
@@
 
+\bibitem[\citeproctext]{ref-lewandowskyRoleConspiracistIdeation2013}
+Lewandowsky, S., Gignac, G.
+E., \& Oberauer, K.
+(2013).
+The Role of Conspiracist Ideation and Worldviews in Predicting Rejection of Science.
+\emph{PLOS ONE}, \emph{8}(10), e75637.
+\url{https://doi.org/10.1371/journal.pone.0075637}
+
 \bibitem[\citeproctext]{ref-lewandowskyMotivatedRejectionScience2016}
@@
 
+\bibitem[\citeproctext]{ref-mayerIntegrativeModelOrganizational1995a}
+Mayer, R. C., Davis, J. H., \& Schoorman, F. D. (1995). An Integrative Model of Organizational Trust. \emph{The Academy of Management Review}, \emph{20}(3), 709. \url{https://doi.org/10.2307/258792}
+
+\bibitem[\citeproctext]{ref-medeSciencerelatedPopulismConceptualization2023a}
+Mede, N.
+G.
+(2023).
+Science-related populism: Conceptualization, empirical investigation, and implications for science communication (Dissertation Summary).
+\emph{Studies in Communication Sciences}, \emph{23}(3).
+\url{https://doi.org/10.24434/j.scoms.2023.03.4403}
+
+\bibitem[\citeproctext]{ref-medeSciencerelatedPopulismConceptualizing2020}
+Mede, N.
+G., \& Schäfer, M.
+S.
+(2020).
+Science-related populism: Conceptualizing populist demands toward science.
+\emph{Public Understanding of Science}, \emph{29}(5), 473--491.
+\url{https://doi.org/10.1177/0963662520924259}
+
+\bibitem[\citeproctext]{ref-medeWhoSupportsSciencerelated2022}
+Mede, N.
+G., Schäfer, M.
+S., Metag, J., \& Klinger, K.
+(2022).
+Who supports science-related populism?
+A nationally representative survey on the prevalence and explanatory factors of populist attitudes toward science in Switzerland.
+\emph{PLOS ONE}, \emph{17}(8), e0271204.
+\url{https://doi.org/10.1371/journal.pone.0271204}
+
 \bibitem[\citeproctext]{ref-millerMeasurementCivicScientific1998a}
 Miller, J. D. (1998). The measurement of civic scientific literacy. \emph{Public Understanding of Science}, \emph{7}(3), 203--223. \url{https://doi.org/10.1088/0963-6625/7/3/001}
+
+\bibitem[\citeproctext]{ref-mitonCognitiveObstaclesProVaccination2015}
+Miton, H., \& Mercier, H. (2015). Cognitive Obstacles to Pro-Vaccination Beliefs. \emph{Trends in Cognitive Sciences}, \emph{19}(11), 633--636. \url{https://doi.org/10.1016/j.tics.2015.08.007}
+
+\bibitem[\citeproctext]{ref-munroScientificImpotenceExcuse2010}
+Munro, G.
+D.
+(2010).
+The scientific impotence excuse: Discounting belief-threatening scientific abstracts.
+\emph{Journal of Applied Social Psychology}, \emph{40}(3), 579--600.
+\url{https://doi.org/10.1111/j.1559-1816.2010.00588.x}
+
+\bibitem[\citeproctext]{ref-nadelsonTrustScienceScientists2015}
+Nadelson, L.
+S., \& Hardy, K.
+K.
+(2015).
+Trust in science and scientists and the acceptance of evolution.
+\emph{Evolution: Education and Outreach}, \emph{8}(1), 9.
+\url{https://doi.org/10.1186/s12052-015-0037-4}
 
@@
 
-\bibitem[\citeproctext]{ref-pummererConspiracyBeliefsMajority}
+\bibitem[\citeproctext]{ref-pummererConspiracyBeliefsMajority2024}
 Pummerer, L., Fock, L., Winter, K., \& Sassenberg, K.
-(n.d.).
+(2024).
 Conspiracy beliefs and majority influence.
-\emph{The Journal of Social Psychology}, \emph{0}(0), 1--16.
+\emph{The Journal of Social Psychology}, 1--16.
 \url{https://doi.org/10.1080/00224545.2024.2397491}
@@
 
+\bibitem[\citeproctext]{ref-rousseauIntroductionSpecialTopic1998}
+Rousseau, D.
+M., Sitkin, S.
+B., Burt, R.
+S., \& Camerer, C.
+(1998).
+Introduction to Special Topic Forum: Not so Different after All: A Cross-Discipline View of Trust.
+\emph{The Academy of Management Review}, \emph{23}(3), 393--404.
+Retrieved from \url{http://www.jstor.org/stable/259285}
+
+\bibitem[\citeproctext]{ref-rutjensConspiracyBeliefsScience2022}
+Rutjens, B. T., \& Većkalov, B. (2022). Conspiracy beliefs and science rejection. \emph{Current Opinion in Psychology}, \emph{46}, 101392. \url{https://doi.org/10.1016/j.copsyc.2022.101392}
+
+\bibitem[\citeproctext]{ref-salmonVaccineHesitancyCauses2015}
+Salmon, D. A., Dudley, M. Z., Glanz, J. M., \& Omer, S. B. (2015). Vaccine hesitancy: Causes, consequences, and a call to action. \emph{Vaccine}, \emph{33}, D66--D71.
+
+\bibitem[\citeproctext]{ref-stockemerUnderstandingClimateChange2024}
+Stockemer, D., \& Bordeleau, J.-N.
+(2024).
+Understanding climate change conspiracy beliefs: A comparative outlook.
+\emph{Harvard Kennedy School Misinformation Review}, \emph{5}(6).
+\url{https://doi.org/10.37016/mr-2020-162}
+
 \bibitem[\citeproctext]{ref-sturgisTrustScienceSocial2021}
@@
 \url{https://doi.org/10.1038/s41562-021-01115-7}
+
+\bibitem[\citeproctext]{ref-uscinskiHaveBeliefsConspiracy2022}
+Uscinski, J., Enders, A., Klofstad, C., Seelig, M., Drochon, H., Premaratne, K., \& Murthi, M.
+(2022).
+Have beliefs in conspiracy theories increased over time?
+\emph{PLoS ONE}, \emph{17}(7), e0270429.
+\url{https://doi.org/10.1371/journal.pone.0270429}
 
@@
 
+\bibitem[\citeproctext]{ref-vranicDidMyOwn2022}
+Vranic, A., Hromatko, I., \& Tonković, M.
+(2022).
+{``}I did my own research{''}: Overconfidence, (dis)trust in science, and endorsement of conspiracy theories.
+\emph{Frontiers in Psychology}, \emph{13}.
+\url{https://doi.org/10.3389/fpsyg.2022.931865}
+
 \bibitem[\citeproctext]{ref-wellcomeglobalmonitorWellcomeGlobalMonitor2018}
@@
 
+\bibitem[\citeproctext]{ref-wilholtEpistemicTrustScience2013}
+Wilholt, T. (2013). Epistemic trust in science. \emph{The British Journal for the Philosophy of Science}, \emph{64}(2), 233--253. \url{https://doi.org/10.1093/bjps/axs007}
+
+\bibitem[\citeproctext]{ref-wintterlinPredictingPublicTrust2022}
+Wintterlin, F., Hendriks, F., Mede, N.
+G., Bromme, R., Metag, J., \& Schäfer, M.
+S.
+(2022).
+Predicting public trust in science: The role of basic orientations toward science, perceived trustworthiness of scientists, and experiences with science.
+\emph{Frontiers in Communication}, \emph{6}.
+\url{https://doi.org/10.3389/fcomm.2021.822757}
+
 \end{CSLReferences}
+
+\newpage
 
@@
 
+\section{Conspiracy Thinking}\label{conspiracy-thinking}
+
+Figure \ref{fig:conspiracy-thinking-overview} shows the distributions of conspiracy thinking across the four studies.
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/conspiracy-thinking-overview.pdf}
+\caption{\label{fig:conspiracy-thinking-overview}Distributions of conspiracy thinking.}
+\end{figure}
+
+\clearpage
+
 \section{Trust in science}\label{trust}
@@
 
-\section{Conspiracy Thinking}\label{conspiracy-thinking}
-
-Figure \ref{fig:conspiracy-thinking-overview} shows the distributions of conspiracy thinking across the four studies.
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/conspiracy-thinking-overview.pdf}
-\caption{\label{fig:conspiracy-thinking-overview}Distributions of conspiracy thinking.}
-\end{figure}
-
-\clearpage
-
 \section{Conspiracy Belief}\label{conspiracy-belief}
@@
 
+Were participants better than chance in their answers?
+To obtain the expected value (EV) of someone answering at random, for each question, we calculate the probability of answering correctly (since each correct answer gives 1 point), and then sum these probabilities.
+In the case of Study 1, the EV is 4.666667 out of 11 (1/3 + 1/3 + 1/2 + 1/3 + 1/2 + 1/3 + 1/2 + 1/2 + 1/2 + 1/2 + 1/3) including the additional item on where trees draw most of their materials from.
+Without this item, the EV is, just as in Studies 2 and 3, 4.333333 out of 10 (1/3 + 1/3 + 1/2 + 1/3 + 1/2 + 1/3 + 1/2 + 1/2 + 1/2 + 1/2).
+In Study 4 the number of answer options was consistently three, so the EV is 3.333333.
+
+As visualized in Figure \ref{fig:knowledge-overview}, for Studies 1, 2, and 3, the vast majority of participants obtained a higher score than the EV, strongly suggesting answers not at random.
+For Study 4, as expected from the very specific, non-standard questions, the knowledge score is centered around the EV, suggesting that most people were at random, and slightly right-skewed, suggesting that a few individuals were either very knowledgeable or looked up some of the answers.
+
 \begin{figure}
@@
 
-\section{Experiment 1}\label{exp1}
-
-\subsection{Participants}\label{participants-1}
-
-We recruited 200 participants from the US via prolific.
-6 participants failed our attention check, resulting in a final sample of 194 participants (98 female, 96 male; \(age_\text{mean}\): 41.71, \(age_\text{sd}\): 13.07, \(age_\text{median}\): 39).
-Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.
-
-\subsection{Materials}\label{materials-2}
-
-\FloatBarrier
-
-\begin{figure}
-
-\includegraphics[width=1\linewidth]{./figures/study1_question_example} \hfill{}
-
-\caption{Example of a science question, the scientific consensus and the corresponding acceptance question.}\label{fig:stimulus-example}
-\end{figure}
-
-\subsubsection{Conspiracy scales}\label{conspiracy-scales}
+\section{Conspiracy Measures}\label{exp1}
 
@@
 
+\clearpage
+
+\section{Study 1}\label{exp1}
+
+\subsection{Participants}\label{participants-1}
+
+We recruited 200 participants from the US via prolific.
+6 participants failed our attention check, resulting in a final sample of 194 participants (98 female, 96 male; \(age_\text{mean}\): 41.71, \(age_\text{sd}\): 13.07, \(age_\text{median}\): 39).
+Since we did not have any prior assumptions on effect sizes, we did not do a power analysis.
+
+\subsection{Materials}\label{materials-2}
+
+\FloatBarrier
+
+\begin{figure}
+
+\includegraphics[width=1\linewidth]{./figures/study1_question_example} \hfill{}
+
+\caption{Example of a science question, the scientific consensus and the corresponding acceptance question.}\label{fig:stimulus-example}
+\end{figure}
+
 \subsection{Results}\label{results-1}
@@
 \midrule
-BCTI 
-(main conspiracy measure) & -0.38 (p< .001) & -0.33 (p< .001)\\
+BCTI (main conspiracy measure) & -0.38 (p< .001) & -0.33 (p< .001)\\
 CMQ & -0.11 (p= 0.121) & -0.07 (p= 0.317)\\
@@
 WGM trust scientists & 0.23 (p= 0.001) & 0.26 (p< .001)\\
-WGM trust general 
-(main trust measure) & 0.29 (p< .001) & 0.27 (p< .001)\\
+WGM trust general (main trust measure) & 0.29 (p< .001) & 0.27 (p< .001)\\
 PEW trust scientists & 0.16 (p= 0.027) & 0.21 (p= 0.003)\\
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
@@
 
-\subsection{Open ended comments}\label{open-ended-comments}
-
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{40em}}
-\caption{\label{tab:unnamed-chunk-52}Comments of participants on Study 1}\\
-\toprule
-id & comments\\
-\midrule
-1 & no\\
-20 & None\\
-22 & I answered thoughtfully, good luck!\\
-27 & interesting, thank you.\\
-30 & n/a\\
-\addlinespace
-32 & None\\
-39 & no\\
-41 & none\\
-42 & n/a\\
-67 & thanks\\
-\addlinespace
-69 & No comments\\
-74 & no\\
-76 & none\\
-78 & None - thank you!\\
-79 & Should've clicked agree about plants getting carbon from the air, but I wanted to look it up first.
-Whenever I encounter something I didn't know online, I like to cross-reference.
-In retrospect though, everything else you asked about was simple and on the level, and I'm sure you are correct.
-
-Thanks for your hard work.\\
-\addlinespace
-84 & No, thank you.\\
-86 & n/a\\
-89 & N/A\\
-95 & no\\
-98 & Great topics! Thank you for this study.\\
-\addlinespace
-109 & To provide context for my responses, I currently study physiology (STEM) and perform academic research as an undergrad at my university.\\
-110 & n/a\\
-114 & Everything in the survey went fine with no issues.
-We seem to have a lot of conspiracy theory nuts in the world today and I think it's a sign of the larger mental health crisis that's currently going on.\\
-115 & I like the subject content of the survey\\
-116 & No\\
-\addlinespace
-132 & None.\\
-133 & very interesting\\
-135 & Nope\\
-139 & none\\
-140 & Air AND water make up the bulk of trees mass and the water is drawn from the earth.
-Water has much more mass than air and I stand by my answer of Earth being that is where the water is drawn from.
-The question was very vague and as you see...
-open to interpretation\\
-\addlinespace
-141 & n/a\\
-148 & no\\
-150 & no\\
-152 & None.\\
-154 & none, tysm!\\
-\addlinespace
-155 & Not virus, just metals/DDT. And I think it would be impossible for dinosaurs to have sex with the way they are shaped.\\
-157 & none\\
-159 & I think scientists are acting in the best interests of science rather than the community directly. An unethical scientist could advance scientific knowledge while hurting humanity.\\
-161 & Mo Thanks\\
-164 & I appreciate science but understand that if scientists don't go along with what investors/govt want they won't receive grants for funding.
-I also very recently lived through several years of gaslighting about covid.
-While I realize I'm not the sharpest tool in the shed, I'm not a total idiot.
-Things that were censored at the beginning of the pandemic are now allowed to be discussed freely.
-Science should be publicly debated otherwise it isn't science.\\
-\addlinespace
-165 & it was interesting\\
-166 & Everything in the study was good I enjoyed taking it\\
-167 & none\\
-169 & good study\\
-170 & Thanks for the opportunity!\\
-\addlinespace
-173 & No\\
-180 & The study was great.\\
-183 & Very interesting study.\\
-185 & None\\
-186 & Thanks, have  good day\\
-\addlinespace
-189 & no\\
-\bottomrule
-\end{longtable}
-
 \clearpage
 
-\section{Experiment 2}\label{exp2}
-
-In experiment 1, we tested whether participants would accept the scientific consensus on basic science facts.
+\section{Study 2}\label{exp2}
+
+In study 1, we tested whether participants would accept the scientific consensus on basic science facts.
 In most instances they did, but not always.
-In experiment 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
+In study 2, we wanted to test whether this reluctance was because of participants not trusting us as a source of consensual science knowledge.
 To do so, we added an explanation and sources to each consensus statement, instead of only stating the consensus.
@@
 
-Based on experiment 1, we formulated the following hypotheses:
+Based on study 1, we formulated the following hypotheses:
 
@@
 
-The procedure was the same as in experiment 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
+The procedure was the same as in study 1, with the difference that, instead of just stating the scientific consensus, participants were presented with a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer (Fig.
 \ref{fig:exp2-stimulus-example}.
@@
 
-We relied on the same items as in experiment 1. The only difference was that we removed the question on trees.
+We relied on the same items as in study 1. The only difference was that we removed the question on trees.
 
@@
 
-As in experiment 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = 0.28, p \textless{} .001) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = -0.40, p \textless{} .001).
-By contrast to experiment 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
+As in study 1, we find a positive but small correlation between science knowledge and trust in science (H1a: r = 0.28, p \textless{} .001) and a small negative correlation between science knowledge and conspiracy thinking (H2a: r = -0.40, p \textless{} .001).
+By contrast to study 1, we conditioned on initially false answers when looking at the relationship of consensus acceptance with trust in science and conspiracy thinking, respectively.
 For trust in science, we find no statistically significant correlation (r = 0.16, p = 0.051).
@@
 
-Confirming results from experiment 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
+Confirming results from study 1, we find that the more people are knowledgeable about science and the more they tend to accept the scientific consensus even when they are not that knowledgeable in science, the more they tend to trust science.
 These correlations are relatively weak, which might be partly due to ceiling effects: (i) most people do trust science, and (ii) that is true even among people with low knowledge or acceptance rates.
@@
 
-Similar to experiment 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.~answered the knowledge question wrongly).
-The share of these latter is considerably larger in experiment 2 (92.9 \%) than in experiment 1 (76.3 \%).
+Similar to study 1, most people (i) do know and agree with the scientific consensus, and (ii) tend to accept the scientific consent even if they were not previously aware of it (i.e.~answered the knowledge question wrongly).
+The share of these latter is considerably larger in study 2 (92.9 \%) than in study 1 (76.3 \%).
 While this could be just sampling variation, it might be that adding explanations and sources convinced people more than merely stating the consensus.
@@
 \midrule
-BCTI 
-(main conspiracy measure) & -0.4 (p< .001) & -0.37 (p< .001)\\
+BCTI (main conspiracy measure) & -0.4 (p< .001) & -0.37 (p< .001)\\
 CMQ & -0.17 (p= 0.019) & -0.22 (p= 0.003)\\
@@
 WGM trust scientists & 0.23 (p= 0.002) & 0.25 (p< .001)\\
-WGM trust general 
-(main trust measure) & 0.28 (p< .001) & 0.3 (p< .001)\\
+WGM trust general (main trust measure) & 0.28 (p< .001) & 0.3 (p< .001)\\
 PEW trust scientists & 0.16 (p= 0.027) & 0.22 (p= 0.003)\\
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:exp2-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
@@
 
-\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection}
-
-\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp2-reasons-rejection}Reasons for consensus rejection (raw answers)}\\
-\toprule
-id & question & answer\\
-\midrule
-33 & sex & Because I didn't think the information provided here was accurate, more or less.\\
-45 & lasers & I read the article.\\
-45 & salt & Read the article provided.\\
-47 & lasers & The statement "lasers do not work by focusing sound waves" is correct according to scientific consensus.
-Lasers operate on the principles of stimulated emission of radiation, wherein photons are emitted in a coherent, focused beam through the process of stimulated emission.\\
-47 & diamonds & Are diamonds made of carbon? Scientific consensus: Diamonds are made of carbon.  Do you think you could explain why you answered differently?\\
-\addlinespace
-47 & water & it likely arose from a mistake or miscommunication.\\
-50 & orbit & I honestly couldn't remember.\\
-50 & speed & When I looked it up I got two different answers.\\
-71 & water & After I put "No", I realized that I was thinking about "H2o" incorrectly. So yes, I actually do agree.\\
-73 & electrons & Atoms are known to be the smallest unit of matter.\\
-\addlinespace
-77 & continents & I believe God make it they way it is and if He wanted it changed, He would have done it\\
-78 & sex & no\\
-78 & salt & no\\
-79 & orbit & It’s early and I’m tired. I misread it.\\
-105 & salt & I think I simply misunderstood the paragraph. I'm actually trying to read it thoroughly and answering truthfully, but I have misread it.\\
-\addlinespace
-109 & electrons & Maybe\\
-127 & continents & I'm a young earth Christian. I believe the Christian Bible (and outside sources) offers ample evidence that the earth is thousands of years old, not millions.\\
-127 & orbit & I do not believe in the heliocentric model of an earth spinning around a centralized sun.\\
-130 & continents & I don't think the continent move rather it is the atmosphere\\
-130 & sex & I think that the gene both parent release should determine the sex of the baby\\
-\addlinespace
-131 & antibiotics & There are some antibiotics that also kill viruses\\
-132 & salt & Because it says that table salt is not made of calcium carbonate, so I clicked no.\\
-133 & lasers & I feel like the excerpt that was provided did not definitively state the answer.\\
-140 & lasers & It was confusing to understand the explanation of why lasers don’t focus on sound\\
-152 & continents & I think that they were that way after the ice age.
-It split apart the whole surface and then water filled in and they have remained the same for ever.
-They are not moving around or have they for millions of years.\\
-\addlinespace
-152 & orbit & The  Earth circles the sun every day.  The dark side is when the moon comes out.  So everyday, you have daylight and dark (moon).  Dark comes from the dark side of the sun.\\
-170 & lasers & I may have misread, but I did not see confirmation of this.\\
-172 & salt & no\\
-178 & electrons & This answer is different because atoms is everywhere and you can't see it but electrons aren't out that as much at all but you could possibly know it's there\\
-178 & antibiotics & Because antibiotics can kill bacteria in which kills the virus as well in the process so you get healed with this product\\
-\addlinespace
-178 & sex & There is no 100\% evidence that this is the case and I still fully believe that it is from both genes to make this decision.\\
-179 & antibiotics & Antibiotics seem that if they are capable of killing tiny things such as bacteria then they are most likely to attach to other things in out bodies and cause it harm although unintentional.\\
-183 & salt & I thought the scientific consensus was right\\
-186 & sex & The scientific consensus is clear: both parents' genes determine the sex of a baby, specifically the sex chromosomes they contribute.\\
-189 & salt & I made an error in my selection.\\
-\bottomrule
-\end{longtable}
-
 \clearpage
 
-\section{Experiment 3}\label{exp3}
+\section{Study 3}\label{exp3}
 
@@
 
-The procedure was mostly the same as in experiment 2.
+The procedure was mostly the same as in study 2.
 In addition, after each open-ended question on cases where participants rejected the scientific consensus, participants were also asked if they want to change their answer and accept the scientific consensus.
@@
 \midrule
-BCTI 
-(main conspiracy measure) & -0.16 (p= 0.055) & -0.03 (p= 0.788)\\
+BCTI (main conspiracy measure) & -0.16 (p= 0.055) & -0.03 (p= 0.788)\\
 CMQ & 0.01 (p= 0.925) & -0.01 (p= 0.934)\\
@@
 WGM trust scientists & 0.11 (p= 0.193) & 0.2 (p= 0.007)\\
-WGM trust general 
-(main trust measure) & 0.14 (p= 0.100) & 0.17 (p= 0.019)\\
+WGM trust general (main trust measure) & 0.14 (p= 0.100) & 0.17 (p= 0.019)\\
 PEW trust scientists & -0.08 (p= 0.367) & 0.12 (p= 0.104)\\
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:exp3-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
@@
 
-\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection-1}
-
-\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp3-reasons-rejection}Reasons for consensus rejection (raw answers)}\\
-\toprule
-id & question & answer\\
-\midrule
-5 & lasers & It is what I believed I was taught\\
-12 & antibiotics & I do not know\\
-20 & water & I think I got it confused in my head. I was thinking 2 oxygen and 1 hydrogen.\\
-25 & electrons & I made a mistake.\\
-26 & antibiotics & I have always experienced receiving antibiotic for viral infections. So i do not agree that antibiotic does not kill viruses.\\
-\addlinespace
-32 & antibiotics & Because I realized that my thinking was not correct.\\
-32 & orbit & After thinking about it, I realized  that it does take a year for the sun to go around the earth.\\
-38 & continents & The earth is not millions of years old.\\
-38 & orbit & The earth is flat and stationary.\\
-43 & water & I got it confused. Its true. One oxygen and  2 hydrogen is what i learned in school.\\
-\addlinespace
-46 & sex & Yes.  Its normal to think that, however, not always.\\
-53 & continents & The Earth was created aged by God and has not been around for millions of years. I do not believe in evolution but rather creation based on the Word of God.\\
-61 & continents & I do not believe that the universe is millions of years old\\
-62 & continents & I do not believe the theory that the earth is that old.\\
-65 & electrons & Because I would like to see more research on this matter.\\
-\addlinespace
-65 & antibiotics & Because I would like to see more research on this matter.\\
-65 & continents & Because I would like to see more research on this matter.\\
-65 & diamonds & Because I would like to see more research on this matter.\\
-65 & water & Because I would like to see more research on this matter.\\
-72 & lasers & That’s the truth\\
-\addlinespace
-72 & salt & I wasn’t sure\\
-83 & continents & The Earth has only been around for 6-7 thousand years. As far as I know however, the Continents have been moving.\\
-83 & water & thought it was "2 parts oxide, one hydrogen." I was of course wrong. I also miss clicked the wrong answer for how long it takes the Earth the revolve around the sun... I think.\\
-87 & electrons & I believe atoms are the smallest particles.\\
-87 & continents & On second thought, I agree.\\
-\addlinespace
-87 & diamonds & On second thought, I believe I agree.\\
-87 & water & This does not fit according to my recollection.\\
-92 & salt & I meant to agree, I did not know the answer.\\
-96 & electrons & I don't think electrons are smaller, just because that is what we were taught in science\\
-96 & continents & This is something we are taught academically in science, it may not be true\\
-\addlinespace
-98 & antibiotics & I think there are two different kinds of antibiotics. One type kills bacteria infections and the other type kills viral infections\\
-99 & continents & I don't have a reason to believe Earth has been here for millions of years, so I don't agree with the consensus. Of course, I know I could be wrong.\\
-102 & continents & This information is based in evolution which is a theory.\\
-109 & continents & I partially agree that continents have moved due to plate tectonics.
-But I don't believe the Earth is as old as it mentions in the explanation, and is much younger.
-I also disagree with scientific consensus and there is really no such thing.
-Because with science, when you base things on evidence and then find new evidence, you have to change your theory.
-So there is really no such thing as consensus, there is data, facts and theories.\\
-112 & antibiotics & No\\
-\addlinespace
-112 & speed & No\\
-112 & salt & No\\
-117 & antibiotics & Not really. I thought they killed virus' but maybe I am wrong. heh.\\
-117 & sex & I thought both male and female decided the fate of the babies gender.\\
-122 & antibiotics & antibotics only kill bacteria\\
-\addlinespace
-122 & lasers & I thought they light was a lot faster than sound waves\\
-122 & salt & not sure\\
-129 & continents & YES, I believe in a young earth.  I do not accept the tax supported theory of evolution.\\
-130 & antibiotics & I was confused.\\
-132 & continents & Continental drift is an unproven theory. I have seen no actual evidence for it. Flooding and receding of waters are more likely the cause of coastline changes throughout history.\\
-\addlinespace
-132 & orbit & There is no scientific evidence that the earth moves at all except for earthquakes.
-The sun is clearly moving around the earth in a daily circular pattern.
-The sun is much smaller and closer than we have been taught.\\
-142 & salt & I thought it was made up of that.\\
-143 & lasers & Lasers operate based on the principles of stimulated emission of electromagnetic radiation, typically light.\\
-146 & antibiotics & Different knowledge and experience.\\
-146 & sex & A mothers choice not the fathers choice because of religious beliefs.\\
-\addlinespace
-146 & diamonds & Gained through knowledge and skills.\\
-150 & speed & To be honest, I did actually hear the concept that like travels fashion and sound so I’m not sure why I picked that answer.\\
-151 & lasers & I accidentally clicked the other answer I'm not going to lie. As soon as I clicked it on accident I was like "Sh*t". I meant to choose the other answer on this one, I'm sorry.\\
-151 & salt & I misunderstood this question. This one confused me the most even though it's right in front of me. Sorry.\\
-155 & salt & error\\
-\addlinespace
-158 & lasers & No I did not know this question.\\
-158 & salt & I knew this one and just blanked on the answer.\\
-159 & sex & I thought I read it said in most cases women transmit an X chromosome, but now I'm not sure.\\
-161 & antibiotics & I thought antibiotics kills every bit of poison.\\
-166 & electrons & I'm not sure, I dont' understand it\\
-\addlinespace
-166 & water & I'm not sure, I dont' understand it\\
-171 & diamonds & It just doesn't sit well with me.\\
-172 & sex & I was wrong. I agree with scientific consensus.\\
-173 & orbit & Yes the scientist now more than I do\\
-178 & electrons & I was incorrect on this.\\
-\addlinespace
-191 & antibiotics & I could be wrong, but I thought they killed viruses as well.\\
-191 & continents & I do not believe without a doubt the earth has been around for millions of years.\\
-196 & sex & Answered incorrectly in error.  I agree with the scientific consensus.\\
-197 & antibiotics & I’ve taken antibiotics numerous of times and although my sickness is usually cured, I always get a yeast infection as result.
-This always happens since the antibiotics kill the good and bad bacteria in the body.\\
-197 & sex & I believe that God decides whether parents will be blessed with a boy or girl. It had nothing to do with genetics.\\
-\addlinespace
-197 & diamonds & This just doesn’t seem correct even after viewing the article from the university. I can’t explain why but I can’t justify it.\\
-198 & diamonds & I just never heard of diamonds being made of carbon and it makes zero sense to me while thinking about it.\\
-200 & continents & The earth is not old but young. Not billions or millions but thanks of years.\\
-200 & orbit & One day because the sun is not taking a year to light up other parts of the earth.\\
-\bottomrule
-\end{longtable}
-
-\subsection{Reasons for consensus acceptance}\label{reasons-for-consensus-acceptance}
-
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp3-other-reasons-acceptance}Reasons for consensus acceptance (participants who answered 'other')}\\
-\toprule
-id & other\_reason\_agreement\\
-\midrule
-24 & some stuff I already knew, other stuff I don't care about like atoms and molecules\\
-37 & I mostly agreed based on my personal knowledge.\\
-62 & I have never seen compelling evidence to disprove the consensus.\\
-89 & Because it is true.\\
-101 & I can't say that I trust scientists on many things, but on most in this survey.
-However, while I believe the continents have moved around, I don't believe the earth is millions of years old, as I'm a Christian.
-Though God never notes exactly how old the earth is, so it could be millions of years old.\\
-\addlinespace
-102 & Agreed due to vague knowledge about the subject.\\
-166 & I'm not sure, you cant trust any party\\
-172 & Because i recall previously being taught that\\
-182 & I agree because the explaination makes sense\\
-188 & I don't trust scientists, but those are the current established 'facts'. Which, on the mundane side of things, are likely true.\\
-\addlinespace
-189 & I take nothing that anyone tells me at face value. Lots of studying and verifying accuracy of the things that I was taught my entire life.\\
-195 & I mostly agree with the consensus because I don't have the equipment to independently test it myself and I also know scientists have proven those who don't have the so called equipment to know what they know.
-We can only observe what we observe at the moment, but there's more to what the scientists are observing and theorizing.\\
-\bottomrule
-\end{longtable}
-
-\subsection{How did participants independently verifiy?}\label{how-did-participants-independently-verifiy}
-
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp3-independent-verification}Answers from participants saying they trusted the scientific consens because they verified independently on how they did so.}\\
-\toprule
-id & reason\_followup\\
-\midrule
-5 & By trying to remember\\
-10 & Curiosity and the internet\\
-12 & Guessed\\
-14 & I double-checked the information on Google by looking up various websites, I did go with my initial gut instincts in most cases however if it was wrong upon verifying it then I would change my answer\\
-18 & Prior knowledge from a formal education\\
-\addlinespace
-19 & I can check several different sources on the internet and see if they have the same verifiable information.\\
-21 & I use intuition.  I use common sense.\\
-22 & I have read it before, and did some experiments in chemistry class.
-I look for scientific consensus on everything.
-On these long established principles, I could have selected both answers.
-I believe sometimes scientists are bribed or exhibit bias depending on who is funding them and what answers the funding entity wants.
-COVID treatments were a big example of that.
-Hospitals were paid to execute certain protocols and disallow others.\\
-27 & Most of the questions, the consensus answer was the same as my answer.
-For the one or two questions where it was not, I looked at who their sources were for the answers, and decided whether I believed those sources and how unsure I was about my answer and made my decision from there.\\
-31 & Through exposure to scientific experiments and research that prove the claims valid and can be repeated with the same results occurring.\\
-\addlinespace
-36 & There is evidence and a good amount of understand about these issues. There isn't as much understanding when it comes to things like vaccines.\\
-38 & Through many sources, including my own common sense and a lifetime of research into these matters.\\
-39 & I used the information that I have learned in school and in the past.\\
-40 & Most things I knew already as some were common knowledge\\
-41 & My gut instinct.\\
-\addlinespace
-42 & Most was knowledge I already knew, however if I were to find out or look up info, I would simply consult the internet\\
-43 & I thought about what i was told in school.\\
-44 & I checked reliable sources, and I knew most of them from learning this information in science class.\\
-48 & some things are obvious and don't need verification\\
-49 & Research on my own as opposed to jist blindly following what the media is told to tell me.\\
-\addlinespace
-56 & extensive research\\
-58 & I verified it with what then knowledge I know\\
-67 & I remember some of the stuff from school.\\
-68 & I clicked on two out of three of the links for most statements that I was unsure about.\\
-71 & I have a lot of people who do deep research that verified the info.\\
-\addlinespace
-77 & Just various amounts of research online. There have been times where simple questions were simply ignored or answered in a perplexing manner. Those types of answers always raise a flag.\\
-81 & By  the knowledge I already had.\\
-82 & Some of these questions I remember from high school. And I learned this in science. Therefore I do remember alot of these questions.\\
-88 & I don't know\\
-97 & Either I already knew the information or after reading the information I can to my conclusion\\
-\addlinespace
-98 & My wife has taken two different kinds of antibiotics, one for viral infections and one for bacterial infections\\
-100 & I independently verified the information by doing a lot of research and comparing studies.\\
-103 & I do my own research.\\
-114 & I guess I haven't independently verified the information.\\
-116 & wikipedia articles and google searches\\
-\addlinespace
-120 & The questions I got wrong were facts that I had simply forgotten. Once made aware of the correct answer, I remembered that it was in fact the correct answer.\\
-122 & heard of it before\\
-123 & I used the links provided and then googled the question if I was unsure\\
-126 & It is an obvious pattern whenever these events happen that there is a coverup, you just have to research the track record\\
-129 & I'm unsure of which question, you asked several.
-it's a bit confusing.
-the vaccine question has been verified by personal treatment and independent studies that were banned on social platforms\\
-\addlinespace
-131 & I used the information and links given to verify it.\\
-132 & I have been studying this information for many years.
-Real science is not consensus alone.
-Peer review is rigged and fraudulent.
-We are not being given the truth about many things and most "scientists" don't even know that they are wrong and refuse to reinvestigate the topics.\\
-136 & Have an educational background in anthropology and took a lot of science courses. I trust scientists generally but not always the way the findings are relayed.\\
-137 & My own research, we were lied to in school\\
-144 & By looking through multiple sources on a given topic.\\
-\addlinespace
-152 & A lot of this I have already learned, but I can't personally verify it bc I can't run the tests.\\
-154 & I have done several science experiments to prove a lot of these ideas.\\
-156 & looking it up online from at least 2-3 sources\\
-159 & Instead of taking what the mainstream talking point is and immediately accepting it, I like to do further research myself.\\
-161 & I already knew of the information presented to me for years.\\
-\addlinespace
-163 & I googled it and read information from various sources.\\
-164 & Many of these tidbits of information are taught as introduction to different subjects. I have either studied this or have used Google  Academic to verify.\\
-178 & Some of the questions I've done my own research on, others I can see with my own eyes.\\
-185 & most of the questions i already knew the answer to because i researched it a long time ago\\
-186 & The information presented is basic knowledge we learn throughout ours lives.
-Light travels faster than sound.
-The fathers genes decide whether ba baby will be male or female.
-Diamonds are formed from carbon.
-Calcium carbonate is not sodium chloride.
-It takes the earth 365.2ish to orbit our sun.
-Yes, water is 2 hydrogen, 1 oxygen.
-And and electron is much smaller than an atom.
-All things we learn in school.
-Oh and lasers use light waves not sound waves.\\
-\addlinespace
-187 & For light traveling faster then sound it's obviously verifiable because you see lightning before you hear thunder, lasers are visible so obviously they are made of light not sound, the definition of water is H2O so if something isn't H2O then it isn't water.
-The vast majority of the questions were extremely common sense if you think about them for any amount of time.\\
-191 & I either realized I was wrong or it sounded logical in my mind.\\
-200 & The Bible.\\
-\bottomrule
-\end{longtable}
-
 \clearpage
 
-\section{Experiment 4}\label{exp4}
+\section{Study 4}\label{exp4}
 
@@
 
-The procedure was identical to the one in experiment 3.
+The procedure was identical to the one in study 3.
 
@@
 \midrule
-BCTI 
-(main conspiracy measure) & -0.02 (p= 0.742) & -0.05 (p= 0.498)\\
+BCTI (main conspiracy measure) & -0.02 (p= 0.742) & -0.05 (p= 0.498)\\
 CMQ & 0.11 (p= 0.112) & -0.1 (p= 0.156)\\
@@
 WGM trust scientists & 0.14 (p= 0.045) & 0.26 (p< .001)\\
-WGM trust general 
-(main trust measure) & 0.1 (p= 0.144) & 0.16 (p= 0.029)\\
+WGM trust general (main trust measure) & 0.1 (p= 0.144) & 0.16 (p= 0.029)\\
 PEW trust scientists & 0.14 (p= 0.043) & 0.21 (p= 0.003)\\
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:exp4-false-response-regression}Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values}
@@
 
-\subsection{Reasons for consensus rejection}\label{reasons-for-consensus-rejection-2}
-
-\begin{longtable}[t]{>{}r>{}l>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp4-reasons-rejection}Reasons for consensus rejection (raw answers)}\\
-\toprule
-id & question & answer\\
-\midrule
-5 & planet & I thought it was fake\\
-6 & magnetic & I dont believe the earth has been here 3.7 billion years\\
-10 & orbit & I honestly do not know anything about this stuff so i took a guess\\
-10 & cloning & I took a guess\\
-12 & material & seemed like too long of a time\\
-\addlinespace
-14 & cloning & I didn't think that many mice are actually cloned.\\
-16 & orbit & I don't believe the Earth orbits the Sun.\\
-16 & material & Nothing but their hypotheses without real evidence.\\
-16 & magnetic & I'm not convinced the Earth existed 3.7 billion years ago.\\
-23 & vertebrate & I don't believe in this kind of carbon dating. We can only slightly estimate the age, therefore I highly doubt the exact number of 392 years is accurate.\\
-\addlinespace
-23 & material & Earth is not billions of years old. At most, maybe 10,000 years. There is absolutely no way anything on Earth is billions of years old.\\
-23 & magnetic & There was no magnetic field 3.7 billion years ago because the Earth was formed only around 6,000-10,000 years ago.\\
-24 & cloning & I think it's less than that percentage\\
-26 & vertebrate & no, just a guess\\
-26 & brain & I have no idea.\\
-\addlinespace
-26 & material & No Idea\\
-28 & material & I really could not explain this\\
-28 & cloning & No, I could not explain this either\\
-29 & orbit & Not enough science to prove.\\
-29 & material & How can anyone put an  that age on something? Humans were not around the track.\\
-\addlinespace
-29 & cloning & Do not believe that anything can be cloned 100\%\\
-32 & drug & Science is not always accurate.\\
-32 & planet & N/A\\
-32 & vertebrate & That doesn't seem accurate.\\
-32 & brain & No.\\
-\addlinespace
-32 & boson & No.\\
-32 & orbit & No.\\
-32 & material & No.\\
-32 & cloning & No.\\
-32 & magnetic & No.\\
-\addlinespace
-35 & vertebrate & No, because I just guessed.\\
-35 & orbit & No, I just guessed.\\
-35 & material & No\\
-41 & planet & i guessed\\
-41 & vertebrate & i guessed\\
-\addlinespace
-41 & boson & i guessed\\
-41 & magnetic & i guessed\\
-42 & brain & There was a difference between what wiki said and the other article.\\
-43 & vertebrate & I did not see anything to support this answer.\\
-43 & brain & I did not feel like this was the correct answer.\\
-\addlinespace
-43 & material & Based on the article, I did not see this stated.\\
-43 & cloning & There was nothing stated to prove this.\\
-44 & vertebrate & I don’t see how it is possible to know the oldest vertebrate when there are so many species out there undiscovered\\
-44 & brain & I don’t think I can explain it differently because I’m not an expert in science\\
-44 & orbit & I am unsure if I can explain it differently, but how would we know if it never happened?\\
-\addlinespace
-44 & cloning & Again, this is a very detailed scientific question.
-I feel it so I will have to have extensive knowledge to answer this or I would have to have done an experiment on my own to to know this.\\
-45 & drug & I am sure I heard it was used for malaria.\\
-45 & cloning & I am not sure cloning is a definite technique.\\
-49 & material & Religious reasons.\\
-49 & magnetic & Religious reasons.\\
-\addlinespace
-54 & vertebrate & did not know the answer\\
-54 & brain & did not know the answer\\
-58 & cloning & The success rate just seems too high.\\
-59 & vertebrate & The websites I read, seemed to have disagreed, unless I may have misread what I was fact checking.\\
-60 & vertebrate & No.\\
-\addlinespace
-60 & orbit & No.\\
-63 & planet & I don't know, but in my opinion when you are talking distance from star that is unknowable.\\
-63 & vertebrate & I was guessing.  But knowing age via carbon dating is just guessing as well\\
-63 & brain & I was guessing.  Same answer as others.\\
-63 & boson & I was guessing.  But same answer.  They pretend to know things that are unknowable.\\
-\addlinespace
-63 & orbit & I was guessing.  Same answer as others.  Unknowable\\
-63 & material & I was guessing.  This again is something unknowable.\\
-63 & speed & I was guessing.  This is unknowable.\\
-63 & magnetic & I was guessing.  This is unknowable.\\
-65 & vertebrate & I couldn’t locate the proof\\
-\addlinespace
-65 & speed & The proof wasn’t clear to me\\
-65 & cloning & Didn’t see the info to back this up\\
-67 & material & Just a guess, I didn't think it made sense.\\
-67 & speed & Figured it was the speed of light, turns out I was close.\\
-68 & orbit & I'm not really able to give an explanation\\
-\addlinespace
-68 & speed & not really\\
-68 & cloning & Not really\\
-69 & orbit & It didn't make sense to me at the time.\\
-69 & cloning & I thought it was easier to clone mice\\
-70 & orbit & these were hypothesis on the ones i got wrong.\\
-\addlinespace
-70 & speed & it was a hypothesis, i had no idea\\
-70 & magnetic & it was a hypothesis i had no clue\\
-72 & drug & I do not believe in this consensus since it usually is fabricated\\
-72 & vertebrate & This is usually fabricated\\
-72 & brain & This is normally a fabricated fact\\
-\addlinespace
-72 & boson & This is probably fabricated\\
-72 & orbit & I do not agree with nearly scientific fact since it is made up\\
-72 & material & This is probably made up\\
-72 & speed & I do think this consensus is faked\\
-72 & magnetic & these are all made up terms\\
-\addlinespace
-74 & vertebrate & Because there was not evidence that I understood to support this theory.\\
-74 & brain & Because the research that was provided does not fully support this theory.\\
-74 & boson & Because the research that I read did not support our object this statement.\\
-74 & orbit & The research that was credited does not give a direct amount of time just a formula to get the answer.\\
-74 & material & I disagree with this statement because the research that was given says 4.6 GA years old not more than.\\
-\addlinespace
-74 & speed & I disagree because there are not enough facts to support this claim.\\
-75 & orbit & Looking at the website given, I could not find the exact time of how long the Earth would keep orbiting, so I didn't agree.\\
-75 & speed & I could not find the answer to that in the given websites\\
-75 & cloning & I could not find data that supported that percentage.\\
-78 & material & No, Just a thought.\\
-\addlinespace
-78 & cloning & Just a random thought\\
-79 & orbit & N/A\\
-80 & vertebrate & I actually meant to press agree after reading the wikipedia , it had mentioned how the oldest 'recorded' vertebrate was 392 years old\\
-81 & drug & Malaria sounds more like the answer to this question, since it was developed more recently.\\
-81 & brain & This question I have no idea. But I would be willing to answer differently if needed.\\
-\addlinespace
-81 & orbit & The Earth would still keep its orbit if the sun disappeared. I believe it would still take 365 days to do one full orbit on its axis.\\
-83 & material & The Bible says the Earth is 6000 years old\\
-84 & material & From a biblical perspective, this is not the case. Other than that, no.\\
-87 & vertebrate & No, I was just guessing.\\
-87 & speed & No. I was just guessing.\\
-\addlinespace
-88 & planet & God hasn't created stars that are millions of miles away from Earth or each other.\\
-88 & orbit & God created the Earth as flat and stationary!\\
-88 & material & God created the Earth 6,000 years ago.\\
-88 & magnetic & God created the Earth 6,000 years ago.\\
-92 & cloning & I feel that they've been cloning people with pretty great success for some time now.\\
-\addlinespace
-92 & magnetic & Because I don't agree that the earth is billions years old, since I'm a Christian.\\
-93 & material & There is no way for man to know if it was actually formed on earth or if it was formed somewhere else and then ended up on earth.\\
-94 & boson & I agree\\
-94 & cloning & The articles I read were quite vague but all stated the success of cloning mice has been incredibly low, which made me disagree with the consensus\\
-96 & orbit & I felt if the Sun disappeared than the Earth orbiting would be fast than 8 minutes.\\
-\addlinespace
-96 & speed & I would like to change my answer to agree.\\
-96 & cloning & I would like to change my answer to agree.\\
-97 & material & How do they know how old anything is?.  Everything we have been told by the powers that be for our whole lives has been a lie.  It is hard to believe anything anymore.\\
-98 & speed & I was not able to find that information in the reference material.\\
-98 & cloning & I was not able to find that answer in the reference material.
-And I guess I have a deep distrust of scientists, as well.
-Even if I read something that a scientist said on Wikipedia or Google I would not believe it.
-Remember it was the scientists who said safe and effective.
-And the Nazis have been cloning people since the 30s, so why believe anything they say about cloning mice.\\
-\addlinespace
-104 & vertebrate & It was just a guess.\\
-104 & brain & It was just a guess.\\
-104 & boson & It was just a guess.\\
-104 & speed & It was just a guess.\\
-105 & drug & I am unfamiliar.\\
-\addlinespace
-105 & orbit & Again I have no knowledge. It was a guess\\
-106 & planet & Not sure I took a guess\\
-106 & vertebrate & Not sure I took a guess\\
-106 & brain & Not sure I took a guess\\
-106 & material & Not sure I took a guess\\
-\addlinespace
-108 & material & I don’t trust the sources as anyone can make changes\\
-113 & vertebrate & Because I dont think they really know themselves\\
-113 & material & The earth has only been around a short time and it thousands not millions or billions of years\\
-113 & magnetic & No dont believe the bible is the true map of mans existance and its not billions of years old\\
-114 & planet & I tried to research it and didnt find the answer so i just chose anything\\
-\addlinespace
-114 & vertebrate & I just chose an answer because I didn't know\\
-114 & material & I just chose anything because I didn't know the answer\\
-114 & cloning & I just chose anything because I didn't know the answer\\
-115 & vertebrate & It said the oldest known vertebrate was 252 years old\\
-115 & speed & I didn’t see anywhere in the articles that backs this claim.\\
-\addlinespace
-118 & vertebrate & Aren't dinosaur bones older than 392 years old?\\
-119 & material & If I understood the resources correctly I thought the time period was less. I'm questioning my answer now though!\\
-121 & vertebrate & one of the articles i looked at said it was a smaller number I can't remember the number.\\
-121 & material & I don't believe the earth is that old.\\
-121 & cloning & I am not completely sure.  the number just seemed very large and I think that many of them die during the process.\\
-\addlinespace
-121 & magnetic & I'm not sure.  I still do not believe that the Earth is that old.\\
-122 & speed & When reading the material that were given, I could not find the answer to confirm it. Therefore, since I could not confirm it, I did not accept the scientific consensus.\\
-124 & planet & I didn't really understand the information on this one.\\
-124 & boson & While reading the information provided I didn't see exact confirmation on if it has an electric charge or not.\\
-124 & material & The answer was vague and not exact it just said more than 4 billion but not more than 4.6 billion.\\
-\addlinespace
-128 & drug & Because I thought it was the answer I had out.\\
-128 & planet & It’s that far away.\\
-128 & vertebrate & It shows that it was correct\\
-128 & brain & Because of what I read.\\
-128 & orbit & I had read it wrong.\\
-\addlinespace
-130 & brain & Neurons are everywhere.  I assumed that glials cells were not as prevalent as neurons.\\
-130 & cloning & It was hard to believe science had progressed that much regarding cloning.  Cloning brings to mind science in the far future.\\
-132 & orbit & I don't have anything scientifically to offer other than I just don't believe that's true.\\
-133 & brain & I remember briefly coming across information on this subject and at the time they were thing they out numbered them 10.1. however that could of changed the more research that was done.\\
-133 & cloning & i just assumed that cloning would not have a high success rate from reading about years ago, but I assume things have changed.\\
-\addlinespace
-134 & planet & I do not believe we can measure the distance of stars.\\
-134 & orbit & I took a guess because I do not believe we can predict such a thing.\\
-134 & material & I do not believe the earth is that old.\\
-136 & cloning & I’m not sure that I believe that they are that successful.\\
-137 & material & Not enough scientific proof\\
-\addlinespace
-137 & cloning & I don't believe in cloning techniques\\
-137 & magnetic & It's possible that the earth magnetic field 3.7 billion years ago was 15 micro Tesla\\
-139 & planet & I misread and thought they were talking about the distant from earth to Kepler.\\
-139 & brain & It depends on the area of the brain and there could be some differences depending on which brains you are looking at.\\
-139 & magnetic & I just wasn't sure.\\
-\addlinespace
-140 & vertebrate & No.\\
-140 & brain & No\\
-140 & material & No\\
-140 & cloning & No\\
-143 & planet & I don't trust that astronomers can accurately measure this.\\
-\addlinespace
-143 & vertebrate & I don't trust carbon 14 dating. I don't know what bomb-pusle is.\\
-143 & orbit & I don't trust that scientists can estimate this accurately.\\
-143 & material & I don't trust that scientists have a way to measure this accurately.\\
-143 & magnetic & I don't trust that scientists have a way to measure this accurately.\\
-145 & magnetic & i cant explain why i answered differently i probably forgot\\
-\addlinespace
-151 & boson & In some theoretical models or speculative physics, there might be hypotheses about different properties of the Higgs boson.\\
-151 & speed & The exact percentage of the speed of light can vary slightly depending on the specific conditions and the state of the accelerator at any given time.
-Precision in reporting such values is crucial.\\
-151 & cloning & The success rate might also depend on the specific strain of mice used or the conditions under which the cloning was performed.
-Variations in experimental conditions can affect the reported success rates.\\
-152 & boson & Neutrality is a toxic line of thinking and only establishes constraints\\
-152 & material & Universe is very old yes\\
-\addlinespace
-152 & speed & LHC is not what they say is it\\
-152 & cloning & "sucess" for one thing but also this is vague and doesn't really say much\\
-152 & magnetic & No way to really know this for sure and it also doesnt matter much\\
-153 & orbit & I think if the sun disappeared the gravitational field would disappear nearly immediately; and the Earth would continue on in a straight line.\\
-154 & orbit & I accept the scientific consensus but I do not accept Wikipedia as a confirming source of scientific data.\\
-\addlinespace
-154 & material & I don't have an explanation as to why I answered differently.\\
-154 & speed & I thought from memory of my previous reading that my answer was correct.\\
-155 & planet & I believe that there is not a way to accurately measure such a thing\\
-155 & vertebrate & I believe that a lot of scientists make up numbers and the numbers aren't even accurate\\
-155 & brain & This one might be accurate but I just wanted to disagree because I feel like it might not be accurate\\
-\addlinespace
-155 & orbit & I believe that a lot of those things about the Earth orbiting and stuff like that are misunderstood and not true\\
-155 & material & I believe that the Earth is much younger than we're told and that allow those numbers about the Earth being trillions and trillions of years old is just crazy talk\\
-155 & magnetic & I believe that the Earth does not have a magnetic field that's just a theory that is wrong in my opinion\\
-157 & orbit & Sure, it's a prediction and one that doesn't sound very plausible to me.\\
-157 & material & I don't believe the earth is that old.\\
-\addlinespace
-157 & magnetic & Once again, I don't believe the earth is that old.\\
-159 & planet & Couldn't havesais it better myself\\
-159 & material & No\\
-163 & cloning & Because one of the links said 8 out of 10 success rate on cloning mice.\\
-163 & magnetic & Because i didn't see anything showing that the researcher pulled samples from the northern most region to do the dating.
-There had been 2 other heating events that could have corrupted the data if they took it from the southern area.\\
-\addlinespace
-164 & orbit & I don't fully trust the sources and the Sun's never disappeared so how would they know.\\
-167 & material & They are just guessing. There is no way to verify their estimates.\\
-167 & magnetic & Earth age estimates are just guesses there is no way to verify them.\\
-168 & planet & Scientific "Consensus" has low credibility in my perception.\\
-168 & brain & Scientific "Consensus" has low credibility in my perception.\\
-\addlinespace
-168 & boson & Scientific "Consensus" has low credibility in my perception.\\
-168 & orbit & Scientific "Consensus" has low credibility in my perception.\\
-168 & material & Scientific "Consensus" has low credibility in my perception.\\
-168 & speed & Scientific "Consensus" has low credibility in my perception.\\
-168 & cloning & Scientific "Consensus" has low credibility in my perception.\\
-\addlinespace
-168 & magnetic & Scientific "Consensus" has low credibility in my perception.\\
-169 & material & I don't think has to be that old.\\
-171 & vertebrate & It did not sound right to me, and that science can be wrong too.\\
-171 & cloning & I will agree with science on this one because they  create weird things.\\
-172 & speed & I thought the article showed a different answer, but I looked back and read it again and confirmed the original statement is in fact true\\
-\addlinespace
-172 & cloning & I thought the article showed a different answer, but I looked back and read it again and confirmed the original statement is in fact true\\
-173 & vertebrate & I think an older one has been found since 2016.\\
-173 & brain & I think it is more likely there is a differing amount which may be very small but I really have no idea.
-I just know we humans are highly fallible and fallen and are wrong about things all the time.\\
-173 & material & The Earth is no where near this old, it is around 6000-7000 years old.
-I cannot wait until we come to realize how utterly foolish we are in believing in the millions-billions or years old myth.\\
-173 & magnetic & Same reason as above.\\
-\addlinespace
-174 & boson & The website was not convincing\\
-174 & orbit & I thought it would be longer\\
-175 & material & My final authority is the Bible.  It only goes back 6500 years to the creation.\\
-175 & cloning & I just dont believe cloning techniques are that good my researched showed its around 1-2\%.\\
-175 & magnetic & Earth was created around 6500 years ago.\\
-\addlinespace
-177 & material & I am a creationist and believe the earth is much younger than that and am not convinced the testing methods are correct\\
-177 & cloning & number just seemed too high\\
-181 & brain & There is a similar amount, but not the same amount.\\
-181 & orbit & It's impossible to state as fact.\\
-181 & material & It's impossible to state as fact.\\
-\addlinespace
-181 & magnetic & It's impossible to state as fact.\\
-183 & planet & i read the distance in kilometers instead of miles.\\
-183 & magnetic & i read that during this time the microteslas was 30-50\\
-184 & vertebrate & Knowing how long vertebrates live today it seems unlikely that they lived significantly longer in the past\\
-184 & material & There is so much variation when it comes to measuring the early days of earth that I would assume given enough time this estimate will change dramatically in the future\\
-\addlinespace
-188 & vertebrate & No I am not able to explain my answer.\\
-188 & orbit & No.\\
-192 & vertebrate & Based on what I read it’s not true.\\
-192 & material & Based on what I read it’s not true.\\
-192 & speed & I’m not sure if it’s true because I couldn’t find the proof for the claim in the material.\\
-\addlinespace
-192 & cloning & I don’t know if it is true because I couldn’t find the proof in the material.\\
-194 & planet & It how I was taught.\\
-194 & boson & I don't know\\
-195 & planet & im a flat earther and i believe all the stars are right in our own atmosphere\\
-195 & orbit & i dont believe that the earth orbits around the sun at all to be honest i think the sun and moon travel over the earth\\
-\addlinespace
-195 & material & i dont believe that the earth is that old\\
-195 & magnetic & dont believe the earth is that old\\
-197 & orbit & I could not find the reference to corroborate the information.\\
-197 & material & The oldest material I found was just over 4.4 Ga\\
-197 & speed & I could not find the reference to corroborate the information.\\
-\bottomrule
-\end{longtable}
-
-\subsection{Reasons for consensus acceptance}\label{reasons-for-consensus-acceptance-1}
-
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp4-other-reasons-acceptance}Reasons for consensus acceptance (participants who answered 'other')}\\
-\toprule
-id & other\_reason\_agreement\\
-\midrule
-8 & I didn’t know the answers at all so they could have been either one.\\
-12 & questions were out of my realm and I had no interest in researching it\\
-15 & I am not privy to information in he scientific field. I believe it to mostly be factual, but I'm unsure about discrepancies so I have to assume it may be true.\\
-20 & I agreed with certain things because it's just how I felt.\\
-23 & I didn't verify anything, nor do I really trust scientists, I guess I just believe that it's generally possible and seems like a reasonable calculation.\\
-\addlinespace
-25 & Not sure\\
-26 & I wasn't sure either way.\\
-35 & Neither, because I would have to do my own research beyond the untrustworthy sources you provided.\\
-39 & I mostly agree because I obtained information from opposing sources and formed an opinion.\\
-46 & I guessed\\
-\addlinespace
-63 & I haven't been able to independently verify it.  But I could if I wanted to and that information was something we could accurately figure out and prove it.\\
-85 & I mostly agree with the consensus because they weren't politically-charged topics, and also, I've been able to independently verify the statements.\\
-90 & There is no way for normal people to verify claims made by scientists with access to vast resources\\
-92 & Seemed to be accurate\\
-97 & I didn't really care what the answer was to those questions, so I just went with the flow.\\
-\addlinespace
-99 & I do not agree\\
-104 & It was just a guess.\\
-109 & i agreed with the consensus because i have not done my own research to refute it so what choice did i have.
-there’s not time enough on a survey for me to genuinely refute or agree so i mildly agree\\
-110 & I chose that I didn’t\\
-113 & I agree that things have been made and are scientifically real but not a billion or million years the bible does not confirm this and its smarter then science\\
-\addlinespace
-130 & I trust honest and ethical scientists.  The challenge is to find out who they are.  Science is an extraordinary discipline.  It should never be used to deceive and control us.\\
-132 & A combination of both as I didn't check all of them but the ones that I did in deed checked out.\\
-143 & I didn't have a reason to disagree with it.\\
-152 & ALmost everything is subjective and "science" is wrong again and again although some things are objectively true with empirical evidence like math and 2 + 2 = 4\\
-157 & I believe some of the questions are easier to verify than others.\\
-\addlinespace
-158 & I don't trust them that much because they are bought and paid for.\\
-162 & I agreed because of my trust in scientists.\\
-168 & I would agree if the subject involved could be independently verified, independent as in a non-aligned individual with any interests involving it.\\
-173 & A mix of both, it largely depends on the subject matter.
-I will say I do not have a whole lot of trust in scientists as they are by and large secular in their views which distorts them a tremendous amount.\\
-175 & The points I can agree with make sense and dont go against the Word of God.\\
-\addlinespace
-187 & I agreed because of my own research and knowledge coincides with the consensus.\\
-189 & Agreed because I didn’t want to verify the information.\\
-\bottomrule
-\end{longtable}
-
-\subsection{How did participants independently verifiy?}\label{how-did-participants-independently-verifiy-1}
-
-\begin{longtable}[t]{>{}r>{\raggedright\arraybackslash}p{30em}}
-\caption{\label{tab:exp4-independent-verification}Answers from participants saying they trusted the scientific consens because they verified independently on how they did so.}\\
-\toprule
-id & reason\_followup\\
-\midrule
-2 & Online ,Google, youtube\\
-3 & From the source and knowledge\\
-5 & I compared the articles and their sources for accuracy\\
-6 & Through research or facts I obtained previously\\
-11 & Through the links provided or a google search. Some of these things I've heard in passing on TV or the radio\\
-\addlinespace
-17 & its just information i have researched over the years.
-One example would be aliens.
-The US government denied aliens being real for years but a couple years ago they told the public they are in fact real.\\
-19 & I have not yet but will through factcheck.org.\\
-21 & Reading links and googling more articles\\
-24 & I have a friend and also Google.  I'm interested in scientific facts\\
-27 & After I was showed the answer, I copied and pasted in into search bar and looked at other sources.\\
-\addlinespace
-31 & I researched it on a fact checked website\\
-33 & I looked up each question. I have to just trust it at some point to answer the survey, but there are some things I'm not going to trust the internet for.\\
-34 & I dif a lot of research on some of the questions..\\
-36 & just an educated guess\\
-37 & Through painstakingly tiring, wikipedia combing and years of just generally being inquisitive\\
-\addlinespace
-38 & Do my own research times ten\\
-40 & I went to a couple of the provided links\\
-41 & i did not\\
-42 & By clicking on the links provided and reading the information.\\
-43 & I searched multiple sites.\\
-\addlinespace
-44 & I would have to conduct experiments sent independently, verify it or read trustworthy sources\\
-45 & Some of the information I have heard before.\\
-48 & By finding additional sources\\
-49 & My own research.\\
-50 & I go and read multiple articles or studies\\
-\addlinespace
-51 & Through study and resources.\\
-52 & I have done extensive research, read books from the library, watched DVDs and verified the information.
-After my mom died from a brain tumor I've researched the cures for cancer for 20 years now and know that doctors who come up with a cure are jailed or murdered to keep the status quo of the cancer industry.
-It's upsetting what our own government did in the 911 attacks on our own country as well.\\
-56 & There were 2 links shown to confirm the answers.  The references were from scientists as well.\\
-58 & There's lots of evidence.\\
-59 & I only quickly checked. I have a hard time following Science, as it is not something I find interesting. However, I do like to fact check things when possible, and when provoked to do so.\\
-\addlinespace
-60 & I verified the information by clicking the links provided and using critical reasoning.\\
-65 & Using the websites probided\\
-67 & I always like to fact check stuff because I don't necessarily believe things at face value.\\
-71 & I independently verified some of it a long time ago\\
-72 & By lots of reading, research, and logic common sense\\
-\addlinespace
-74 & I read the documents that were given myself\\
-75 & Read the provide websites to find agreeable information, also quick searches provided some verifications.\\
-76 & Wikepedia\\
-77 & I didn't\\
-79 & I used google. Which I would say is an elementary resource for information, but it was good for answering such questions\\
-\addlinespace
-82 & Non- mainstream media sources, government documents, intuition.\\
-83 & Research on my own\\
-86 & I researched on google.\\
-87 & By the websites you listed.\\
-95 & I knew some things and from personal opinions\\
-\addlinespace
-98 & I use the Brave browser. Less censorship.\\
-101 & I checked online with various sources to make sure they were true\\
-102 & Checking the websites provided\\
-107 & you need to add questions about where scientists get there funding from, and whether or not we would trust a statement like, all scientists agree\\
-108 & I just went with my gut\\
-\addlinespace
-111 & I like to read a lot from all sorts of different sources, then read arguments that disagree with them.
-Eventually I'll find a grain of truth in all of that and use pattern recognition to fill in gaps based on what I know.\\
-114 & Just did a google search\\
-115 & I read the articles.\\
-116 & Though previous knowledge and research\\
-117 & I clicked the links to read what was stated.\\
-\addlinespace
-118 & I did a quick look on Google and searched on my own\\
-119 & Google and the resources shown during the study\\
-120 & I do my own research.\\
-121 & i used the links in the study and looked briefly, at the articles.  Didn't understand most of it but made best decision I could.\\
-122 & I read the material that was given within each of the questions and found the information.\\
-\addlinespace
-123 & I checked various sources and informational circles\\
-124 & I searched up the information and researched myself.\\
-126 & I looked at the websites cited in the questions.\\
-127 & did web searches\\
-133 & Well not exactly verified , but Just things I have researched or came across while online. there is a lot of information out there and a lot of misinformation. You just have to take a side\\
-\addlinespace
-135 & doing research and reading articles and scientific studies.\\
-136 & By clicking the links provided\\
-137 & I somewhat trust scientist but overall, I independently do my own research to verify such existing's\\
-138 & Did much research and corroborated each finding or information and lined up against each other to see similar outcomes.\\
-139 & By looking up the links to reputable sources and seeing if they match up with what other reputable sources are saying.\\
-\addlinespace
-140 & Verified persons in specialized authority.\\
-141 & I just do my own research and look into things and how they could or couldn’t work\\
-145 & i was informed  by trusted sources\\
-147 & I checked the links given\\
-154 & Online research of more trusted sources\\
-\addlinespace
-155 & Basically it aligns with other things I believe to be true so I will agree with it I'm not saying that I went out of my way and did a bunch of different studies but the thing about a lot of scientific research and studies if the foundation of the study is unstable or unreliable then I can't trust that a lot of the results are accurate\\
-156 & I clicked on the articles and read for myself. Instead of trusting the information listed in the survey on its own I clicked the links to verify the information.\\
-163 & i either knew the information or i followed the links provided and believed the information.\\
-166 & I search multiple doctors and scientists and compare there work and theories\\
-167 & I read research papers and look for inconsistencies or bad methodology.\\
-\addlinespace
-170 & Google search it\\
-171 & I heard of these questions before, and I looked at the sources and my sources.\\
-172 & I read the attached files to see if the info could be backed up on there as well.
-The 2 I thought I read as false I ended up doing a google search to confirm that the statements were in fact true\\
-176 & I didn’t but I will\\
-179 & I research various sources that lean in either direction (left or right) as well as sources that appear, at least on the surface, to be impartial.
-After gathering all of this information, I make my informed decision.\\
-\addlinespace
-183 & asking ai such as chatgpt\\
-184 & Through my own research and comparison to other studies and others research.\\
-188 & I looked it up using the links provided to me.\\
-191 & By clicking and reading the link\\
-192 & I read the articles and found the actual answer to the question.\\
-\addlinespace
-194 & Online\\
-195 & google and research\\
-197 & With the references provided in the study.\\
-\bottomrule
-\end{longtable}
-
 \end{document}

### 2025-03-13 paper.Rmd
First Round Revisions

@@
 
-  Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
-This might indicate a wholesale rejection of science.
+  Many people report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
+This might indicate a wholesale rejection of science in substantial segments of the population.
 In four studies, we asked 782 US participants questions about trust in science, conspiracy beliefs, and basic science (e.g.
@@
 output: 
-  papaja::apa6_docx: default
   papaja::apa6_pdf: 
     keep_tex: true
+  papaja::apa6_docx: default
 always_allow_html: true
@@
 
-Although recent global evidence suggests that trust in science is moderately high [@colognaTrustScientistsTheir2024], it is far from being at ceiling.
+Although recent global evidence shows that trust in science is moderately high [@colognaTrustScientistsTheir2024], it is far from being at ceiling.
 Large-scale polls have shown that people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; although see @wellcomeglobalmonitorPublicTrustScientists2021 and @funkPublicConfidenceScientists2020], including the US [@brianAmericansTrustScientists2023], where it is also increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
+Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; @brianAmericansTrustScientists2023; although see @wellcomeglobalmonitorPublicTrustScientists2021; and @funkPublicConfidenceScientists2020] and in the US it is increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
 
@@
 Similarly, in 2021, 40% of Americans believed that “the dangers of genetically-modified foods are being hidden from the public” [down from 45% in 2020, @uscinskiHaveBeliefsConspiracy2022].
-People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories [@lewandowskyRoleConspiracistIdeation2013; @hornseyPsychologicalRootsAntivaccination2018; @hornseyRelationshipsConspiratorialBeliefs2018] and tend to say that they distrust science more generally [@vranicDidMyOwn2022; @stockemerUnderstandingClimateChange2024].
-Despite these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
+People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories [@lewandowskyRoleConspiracistIdeation2013; @hornseyPsychologicalRootsAntivaccination2018; @hornseyRelationshipsConspiratorialBeliefs2018], and tend to say that they distrust science more generally [@vranicDidMyOwn2022; @stockemerUnderstandingClimateChange2024].
+In spite of these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
 Although conspiracy thinking has been identified as a "root cause" of anti-science attitudes [@hornseyAttitudeRootsJiu2017] in the past, these claims rest largely on observational data.
@@
 @nadelsonTrustScienceScientists2015], genetically modified organisms (GMOs) [e.g.
-@fernbachExtremeOpponentsGenetically2019], or a combinations of such topics [see e.g @lewandowskyRoleConspiracistIdeation2013].
+@fernbachExtremeOpponentsGenetically2019], or a combination of such topics [see e.g.
+@lewandowskyRoleConspiracistIdeation2013].
 To the best of our knowledge, no research has investigated the extent to which people trust basic science facts (e.g.
@@
 
-Why does it matter whether trust in science in general is or is not related to trust in scientific findings?
+Why does it matter whether trust in science is or is not related to trust in specific scientific findings?
 First, this question has theoretical implications.
@@
 According to these accounts, science rejection serves to maintain coherence with other beliefs or behaviors [@lewandowskyMotivatedRejectionScience2016; @hornseyWhyFactsAre2020].
-For example, someone might say they don't trust science in general because they are skeptical towards vaccines, not because they actually distrust most of science.
+For example, someone might say they do not trust science in general because they are skeptical towards vaccines, not because they actually distrust most of science.
 A prediction of the motivated reasoning account is that general trust in science should be strongly correlated with conspiracy beliefs.
@@
 We also measured participants’ trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (<https://osf.io/8utsj/?view_only=364478f01fc548d0a42b9962918d871d>).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (<https://osf.io/8utsj/>).
 The differences between the four studies are summarized presently, and the methods are detailed below.
@@
   apa_table(caption = "Justifications for rejecting the scientific consensus by category, Studies 2, 3 and 4 combined.", 
-            note = "*Participans with at least one answer in that category")
+            note = "*Participants with at least one answer in that category")
 
@@
 In line with past research [@allumScienceKnowledgeAttitudes2008], we find that both knowledge and acceptance of science are only weakly correlated with general trust in science.
-This supports contemporary definitions of science literacy, which suggest that for science literacy to be a meaningful concept for science attitudes, it needs to go beyond measuring knowledge of isolated science facts [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016].
-
-Instead, our findings align with a motivated reasoning account of trust in science [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p.
+This supports recent definitions of science literacy, which suggest that for science literacy to be a meaningful concept for science attitudes, it needs to go beyond measuring knowledge of isolated science facts [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016].
+
+By contrast, our findings align with a motivated reasoning account of trust in science [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p.
 217).
@@
 This suggests that these participants had reasons to reject only specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
-Consistent with this explanation, we found a strong association between belief in the anti-science conspiracy theories and general trust in science.
+Consistent with this explanation, we found a strong association between belief in anti-science conspiracy theories and general trust in science.
 
@@
 Declaring not trusting science, or endorsing conspiracy theories [@harrisConspiracyTheoriesPopulism2023] might reflect a desire to maintain epistemic autonomy and not appear to ‘blindly’ accept epistemic authority.
-Such a need to appear epistemically autonomous would be reconciled with the acceptance of basic science facts if it stems from independent evaluation instead of trust.
-The majority of participants in Studies 3 and 4, two samples of vaccine skeptical participants who scored high across a range of different conspiracy beliefs (see ESM), claimed that this was their case, however implausible that might be: it’s not clear how participants could independently verify, say, the ratio of glial cells to neurons.
-These participants did also not engage any more in simple forms of verification, i.e.
+Such a need to appear epistemically autonomous could be reconciled with the acceptance of basic science facts if it was thought to stem from independent evaluation instead of trust.
+The majority of participants in Studies 3 and 4, two samples of vaccine-skeptical participants who scored high across a range of different conspiracy beliefs (see ESM), tended to claim that they had accepted the scientific consensus based on their own evaluation, however implausible that might be: it is not clear how participants could independently verify, say, the ratio of glial cells to neurons.
+These participants did also not engage (more than others) in simple forms of verification, i.e.
 clicking on sources provided in the answers.
-This aligns with other results on conspiracy thinking: Several recent studies have shown that, although they claim not to be, conspiracy theorists are in fact just as susceptible to social influence as others [@pummererConspiracyBeliefsMajority2024; @altayConspiracyBelieversClaim2023].
+This result aligns with other results on conspiracy thinking: Recent studies have shown that, although they claim not to be, conspiracy theorists are in fact just as susceptible to social influence as others [@pummererConspiracyBeliefsMajority2024; @altayConspiracyBelieversClaim2023].
 
@@
 Across various domains of science knowledge such as climate change [@veckalov27countryTestCommunicating2024] or vaccination [@salmonVaccineHesitancyCauses2015], researchers have observed a consensus gap: a gap between the scientific consensus and public opinion.
-For example, some of the most worrying consensus gaps relate to climate change, as substantial segments of the population disagree with scientists on what is happening and what to do about it [e.g., @eganClimateChangeUS2017].
-Yet, we found that for much of basic, non-contentious science, such consensus gaps are absent: The vast majority of our participants–including two vaccine-skeptic attitudes samples (Studies 3 and 4)--did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
-Since even people who distrust science or reject contentious science facts appear to accept most of basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the consensus gaps observed in these domains [on climate change, see, e.g.
+Some of the most worrying consensus gaps relate to climate change, as substantial segments of the population disagree with scientists on what is happening and what to do about it [e.g., @eganClimateChangeUS2017].
+Yet, we found that for much of basic, non-contentious science, such consensus gaps are absent: The vast majority of our participants--including two vaccine-skeptical samples (Studies 3 and 4)--did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+Since even people who say they distrust science, or who reject specific scientific facts appear to accept most of basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the consensus gaps observed in these domains [on climate change, see, e.g.
 @ranneyClimateChangeConceptual2016].
@@
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
-While our results lend support to motivated reasoning accounts of trust in science, they leave several important questions unaddressed, in particular: Why do people reject specific science beliefs but not others?
+While our results support motivated reasoning accounts of trust in science, they leave important questions unaddressed, in particular: Why do people reject specific scientific findings but not others?
 Suggestions have already been made for a number of issues such as vaccination [@mitonCognitiveObstaclesProVaccination2015], GMOs [@blanckeFatalAttractionIntuitive2015], or nuclear energy [@hacquinDisgustSensitivityPublic2021].
@@
 Finally, by design, most of the basic science knowledge presented in this study didn't directly relate to anything controversial for most people.
-Future studies could look at basic science that does relate to science topics which are the object of public and policy-relevant debates (e.g.
+Future studies could look at basic science that does relate to scientific topics which are the object of public and policy-relevant debates (e.g.
 GM food, climate science, vaccines), or scientific findings for which there is a more obvious potential conflict of interest (e.g.

### 2025-03-13 paper.tex
First Round Revisions

@@
 \abstract{%
-Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
-This might indicate a wholesale rejection of science.
+Many people report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
+This might indicate a wholesale rejection of science in substantial segments of the population.
 In four studies, we asked 782 US participants questions about trust in science, conspiracy beliefs, and basic science (e.g.~the relative size of electrons and atoms).
@@
 
-Although recent global evidence suggests that trust in science is moderately high (Cologna et al., 2024), it is far from being at ceiling.
+Although recent global evidence shows that trust in science is moderately high (Cologna et al., 2024), it is far from being at ceiling.
 Large-scale polls have shown that people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia (Wellcome Global Monitor, 2018, 2020).
-Moreover, trust in science has recently been declining in some countries Funk \& Kennedy (2020), including the US (Brian \& Tyson, 2023), where it is also increasingly polarizing (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
+Moreover, trust in science has recently been declining in some countries (Algan et al., 2021; Brian \& Tyson, 2023; although see Wellcome Global Monitor, 2021; and Funk \& Kennedy, 2020) and in the US it is increasingly polarizing (Gauchat, 2012; Krause, Brossard, Scheufele, Xenos, \& Franke, 2019; Li \& Qian, 2022).
 
@@
 Similarly, in 2021, 40\% of Americans believed that ``the dangers of genetically-modified foods are being hidden from the public'' (down from 45\% in 2020, Uscinski et al., 2022).
-People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories (Hornsey, Harris, \& Fielding, 2018b, 2018a; Lewandowsky, Gignac, \& Oberauer, 2013) and tend to say that they distrust science more generally (Stockemer \& Bordeleau, 2024; Vranic, Hromatko, \& Tonković, 2022).
-Despite these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
-Although conspiracy thinking been identified as a ``root cause'' of anti-science attitudes (Hornsey \& Fielding, 2017) in the past, these claims rest largely on observational data.
+People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories (Hornsey, Harris, \& Fielding, 2018b, 2018a; Lewandowsky, Gignac, \& Oberauer, 2013), and tend to say that they distrust science more generally (Stockemer \& Bordeleau, 2024; Vranic, Hromatko, \& Tonković, 2022).
+In spite of these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
+Although conspiracy thinking has been identified as a ``root cause'' of anti-science attitudes (Hornsey \& Fielding, 2017) in the past, these claims rest largely on observational data.
 
@@
 Nadelson \& Hardy, 2015), genetically modified organisms (GMOs) (e.g.
-Fernbach, Light, Scott, Inbar, \& Rozin, 2019), or a combinations of such topics (see e.g Lewandowsky et al., 2013).
+Fernbach, Light, Scott, Inbar, \& Rozin, 2019), or a combination of such topics (see e.g.
+Lewandowsky et al., 2013).
 To the best of our knowledge, no research has investigated the extent to which people trust basic science facts (e.g.~electrons are smaller than atoms).
@@
 According to these accounts, science rejection serves to maintain coherence with other beliefs or behaviors (Hornsey, 2020; Lewandowsky \& Oberauer, 2016).
-For example, someone might say they don't trust science in general because they are skeptical towards vaccines, not because they actually distrust most of science.
+For example, someone might say they do not trust science in general because they are skeptical towards vaccines, not because they actually distrust most of science.
 A prediction of the motivated reasoning account is that general trust in science should be strongly correlated with conspiracy beliefs.
@@
 We also measured participants' trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (\url{https://osf.io/8utsj/?view_only=364478f01fc548d0a42b9962918d871d}).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (\url{https://osf.io/8utsj/}).
 The differences between the four studies are summarized presently, and the methods are detailed below.
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} *Participans with at least one answer in that category}
+\normalsize{\textit{Note.} *Participants with at least one answer in that category}
 \end{tablenotes}
@@
 In line with past research (Allum et al., 2008), we find that both knowledge and acceptance of science are only weakly correlated with general trust in science.
-This supports contemporary definitions of science literacy, which suggest that for science literacy to be a meaningful concept for science attitudes, it needs to go beyond measuring knowledge of isolated science facts (National Academies of Sciences, Engineering, and Medicine, 2016).
-
-Instead, our findings align with a motivated reasoning account of trust in science (Lewandowsky \& Oberauer, 2016), in which ``people tend to reject findings that threaten their core beliefs or worldview'' (p.~217).
+This supports recent definitions of science literacy, which suggest that for science literacy to be a meaningful concept for science attitudes, it needs to go beyond measuring knowledge of isolated science facts (National Academies of Sciences, Engineering, and Medicine, 2016).
+
+By contrast, our findings align with a motivated reasoning account of trust in science (Lewandowsky \& Oberauer, 2016), in which ``people tend to reject findings that threaten their core beliefs or worldview'' (p.~217).
 A number of participants in our studies endorsed specific conspiracy theories questioning basic tenets of science (on evolution, the shape of the earth, etc.), while still accepting the vast majority of basic scientific facts presented to them.
 This suggests that these participants had reasons to reject only specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
-Consistent with this explanation, we found a strong association between belief in the anti-science conspiracy theories and general trust in science.
+Consistent with this explanation, we found a strong association between belief in anti-science conspiracy theories and general trust in science.
 
@@
 Declaring not trusting science, or endorsing conspiracy theories (Harris, 2023) might reflect a desire to maintain epistemic autonomy and not appear to `blindly' accept epistemic authority.
-Such a need to appear epistemically autonomous would be reconciled with the acceptance of basic science facts if it stems from independent evaluation instead of trust.
-The majority of participants in Studies 3 and 4, two samples of vaccine skeptical participants who scored high across a range of different conspiracy beliefs (see ESM), claimed that this was their case, however implausible that might be: it's not clear how participants could independently verify, say, the ratio of glial cells to neurons.
-These participants did also not engage any more in simple forms of verification, i.e.~clicking on sources provided in the answers.
-This aligns with other results on conspiracy thinking: Several recent studies have shown that, although they claim not to be, conspiracy theorists are in fact just as susceptible to social influence as others (Altay, Nera, Ejaz, Schöpfer, \& Tomas, 2023; Pummerer, Fock, Winter, \& Sassenberg, 2024).
+Such a need to appear epistemically autonomous could be reconciled with the acceptance of basic science facts if it was thought to stem from independent evaluation instead of trust.
+The majority of participants in Studies 3 and 4, two samples of vaccine-skeptical participants who scored high across a range of different conspiracy beliefs (see ESM), tended to claim that they had accepted the scientific consensus based on their own evaluation, however implausible that might be: it is not clear how participants could independently verify, say, the ratio of glial cells to neurons.
+These participants did also not engage (more than others) in simple forms of verification, i.e.~clicking on sources provided in the answers.
+This result aligns with other results on conspiracy thinking: Recent studies have shown that, although they claim not to be, conspiracy theorists are in fact just as susceptible to social influence as others (Altay, Nera, Ejaz, Schöpfer, \& Tomas, 2023; Pummerer, Fock, Winter, \& Sassenberg, 2024).
 
@@
 Across various domains of science knowledge such as climate change (Većkalov et al., 2024) or vaccination (Salmon, Dudley, Glanz, \& Omer, 2015), researchers have observed a consensus gap: a gap between the scientific consensus and public opinion.
-For example, some of the most worrying consensus gaps relate to climate change, as substantial segments of the population disagree with scientists on what is happening and what to do about it (e.g., Egan \& Mullin, 2017).
-Yet, we found that for much of basic, non-contentious science, such consensus gaps are absent: The vast majority of our participants--including two vaccine-skeptic attitudes samples (Studies 3 and 4)--did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
-Since even people who distrust science or reject contentious science facts appear to accept most of basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the consensus gaps observed in these domains (on climate change, see, e.g.
+Some of the most worrying consensus gaps relate to climate change, as substantial segments of the population disagree with scientists on what is happening and what to do about it (e.g., Egan \& Mullin, 2017).
+Yet, we found that for much of basic, non-contentious science, such consensus gaps are absent: The vast majority of our participants--including two vaccine-skeptical samples (Studies 3 and 4)--did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+Since even people who say they distrust science, or who reject specific scientific facts appear to accept most of basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the consensus gaps observed in these domains (on climate change, see, e.g.
 Ranney \& Clark, 2016).
@@
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
-While our results lend support to motivated reasoning accounts of trust in science, they leave several important questions unaddressed, in particular: Why do people reject specific science beliefs but not others?
+While our results support motivated reasoning accounts of trust in science, they leave important questions unaddressed, in particular: Why do people reject specific scientific findings but not others?
 Suggestions have already been made for a number of issues such as vaccination (Miton \& Mercier, 2015), GMOs (Blancke, Van Breusegem, De Jaeger, Braeckman, \& Van Montagu, 2015), or nuclear energy (Hacquin, Altay, Aarøe, \& Mercier, 2021).
@@
 Finally, by design, most of the basic science knowledge presented in this study didn't directly relate to anything controversial for most people.
-Future studies could look at basic science that does relate to science topics which are the object of public and policy-relevant debates (e.g.~GM food, climate science, vaccines), or scientific findings for which there is a more obvious potential conflict of interest (e.g.~non-publicly funded medical research).
+Future studies could look at basic science that does relate to scientific topics which are the object of public and policy-relevant debates (e.g.~GM food, climate science, vaccines), or scientific findings for which there is a more obvious potential conflict of interest (e.g.~non-publicly funded medical research).
 

### 2025-07-07 paper.Rmd
second round revisions

@@
 
-  Many people report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
-This might indicate a wholesale rejection of science in substantial segments of the population.
+  Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
+This might indicate a wholesale rejection of science.
 In four studies, we asked 782 US participants questions about trust in science, conspiracy beliefs, and basic science (e.g.
@@
 output: 
+  papaja::apa6_docx: default
   papaja::apa6_pdf: 
     keep_tex: true
-  papaja::apa6_docx: default
 always_allow_html: true
@@
 
-Although recent global evidence shows that trust in science is moderately high [@colognaTrustScientistsTheir2024], it is far from being at ceiling.
+Although recent global evidence shows that trust in science is moderately high [@colognaTrustScientistsTheir2025], it is far from being at ceiling.
 Large-scale polls have shown that people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
@@
 Some of these conspiracy theories are widespread [@rutjensConspiracyBeliefsScience2022].
+Data from the European Social Survey between 2020 and 2022 suggests that in several European countries, relatively large minorities of the population believed that COVID-19 was a result of deliberate and concealed efforts of some government or organisations: around 30% in Spain and Portugal, 22% in the UK, and 25% in Poland [@jabkowskiExploringCOVID19Conspiracy2023].
 In 2023, a survey in eight different countries found that up to 24% of respondents agreed that “climate change is a hoax and scientists touting its existence are lying” [@stockemerUnderstandingClimateChange2024].
 Similarly, in 2021, 40% of Americans believed that “the dangers of genetically-modified foods are being hidden from the public” [down from 45% in 2020, @uscinskiHaveBeliefsConspiracy2022].
-People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories [@lewandowskyRoleConspiracistIdeation2013; @hornseyPsychologicalRootsAntivaccination2018; @hornseyRelationshipsConspiratorialBeliefs2018], and tend to say that they distrust science more generally [@vranicDidMyOwn2022; @stockemerUnderstandingClimateChange2024].
+People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories [@lewandowskyRoleConspiracistIdeation2013; @hornseyPsychologicalRootsAntivaccination2018; @hornseyRelationshipsConspiratorialBeliefs2018; @littrellBeliefScienceRelatedConspiracy2025], and tend to say that they distrust science more generally [@vranicDidMyOwn2022; @stockemerUnderstandingClimateChange2024].
 In spite of these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
@@
 
-Studies 1 and 2 were conducted on the standard sample of US participants recruited on the platform Prolific Academic.
+The surveys were hosted via Qualtrics and participants were recruited via Prolific Academic, a crowdsourcing platform that allows researchers to pre-screen and invite eligible participants from their participant pool.
+Studies 1 and 2 were conducted on the standard sample of US participants.
 In order to increase the share of participants with low trust in science, and who endorse conspiracy theories, Studies 3 and 4 used the same platform, but only recruited participants who had declared previously being skeptical of vaccination.
+
+Because recruitment was conducted through Prolific, we do not have access to the number of individuals who viewed the study invitation but chose not to participate.
+As such, it is not possible for us to calculate response rates (the proportion of invited participants who completed the survey) for the studies.
+However, we employed attention checks and, for each study, we report exclusions based on these checks.
+Moreover, the study invitation was purposefully nondescript (“short psychology experiment”), so that it would be impossible for participants to self-select because of their attitudes towards science for instance.
 
@@
 @funkKeyFindingsPublic2019] ("How much confidence do you have in scientists to act in the best interests of the public?
-[1-5; 1 = No confidence at all, 5 = A great deal of confidence]"), the latter having been used in a recent international study on trust in science [@colognaTrustScientistsTheir2024].
+[1-5; 1 = No confidence at all, 5 = A great deal of confidence]"), the latter having been used in a recent international study on trust in science [@colognaTrustScientistsTheir2025].
 We selected these items so that we could compare the answers in our sample to global survey results.
@@
 
-n an exploratory analysis, we find that
-
-While actual acceptance of scientific knowledge is only weakly correlated with trust in science,
-
 Why did participants reject the scientific consensus?
@@
 @ranneyClimateChangeConceptual2016].
+Past research has shown that arguments building on specific scientific findings are more convincing even to populists who are generally skeptical towards expertise [@peresmanPopulistsListenExpertise2025].
+
+Why do most people (in the US) trust most of basic science?
+Based on the present studies, we can only speculate.
+But if this trust has a rational basis, then education–and science education in particular–is the most likely explanation.
+Large-scale surveys have identified science education as one of the strongest, or the strongest, correlates of trust in science [@noyScienceGoodEffects2019; @wellcomeglobalmonitorWellcomeGlobalMonitor2018] If these studies were to assess trust in basic scientific knowledge, rather than assessing trust in science more broadly, we might expect effects to be even stronger.
+Differentiating the impact of education on different facets of trust would be an interesting avenue for future research.
 
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
+The use of online samples recruited via Prolific does not allow us to calculate response rates and thus estimate non-response bias.
+However, participants could not have self-selected on the basis of the study topic and, by targeting vaccine-skeptic samples in some of the studies, we made an effort in recruiting those participants that can be assumed to be most reluctant to partake in a study confronting them with science facts.
 While our results support motivated reasoning accounts of trust in science, they leave important questions unaddressed, in particular: Why do people reject specific scientific findings but not others?

### 2025-07-07 paper.tex
second round revisions

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \def\languageshorthands#1{}
+\ifLuaTeX
+  \usepackage[english]{selnolig} % disable illegal ligatures
+\fi
 % Manuscript styling
@@
 
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage{bookmark}
@@
 \abstract{%
-Many people report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
-This might indicate a wholesale rejection of science in substantial segments of the population.
+Substantial segments of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
+This might indicate a wholesale rejection of science.
 In four studies, we asked 782 US participants questions about trust in science, conspiracy beliefs, and basic science (e.g.~the relative size of electrons and atoms).
@@
 
-Although recent global evidence shows that trust in science is moderately high (Cologna et al., 2024), it is far from being at ceiling.
+Although recent global evidence shows that trust in science is moderately high (Cologna et al., 2025), it is far from being at ceiling.
 Large-scale polls have shown that people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia (Wellcome Global Monitor, 2018, 2020).
@@
 Some of these conspiracy theories are widespread (Rutjens \& Većkalov, 2022).
+Data from the European Social Survey between 2020 and 2022 suggests that in several European countries, relatively large minorities of the population believed that COVID-19 was a result of deliberate and concealed efforts of some government or organisations: around 30\% in Spain and Portugal, 22\% in the UK, and 25\% in Poland (Jabkowski, Domaradzki, \& Baranowski, 2023).
 In 2023, a survey in eight different countries found that up to 24\% of respondents agreed that ``climate change is a hoax and scientists touting its existence are lying'' (Stockemer \& Bordeleau, 2024).
 Similarly, in 2021, 40\% of Americans believed that ``the dangers of genetically-modified foods are being hidden from the public'' (down from 45\% in 2020, Uscinski et al., 2022).
-People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories (Hornsey, Harris, \& Fielding, 2018b, 2018a; Lewandowsky, Gignac, \& Oberauer, 2013), and tend to say that they distrust science more generally (Stockemer \& Bordeleau, 2024; Vranic, Hromatko, \& Tonković, 2022).
+People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories (Hornsey, Harris, \& Fielding, 2018b, 2018a; Lewandowsky, Gignac, \& Oberauer, 2013; Littrell et al., 2025), and tend to say that they distrust science more generally (Stockemer \& Bordeleau, 2024; Vranic, Hromatko, \& Tonković, 2022).
 In spite of these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
@@
 
-Studies 1 and 2 were conducted on the standard sample of US participants recruited on the platform Prolific Academic.
+The surveys were hosted via Qualtrics and participants were recruited via Prolific Academic, a crowdsourcing platform that allows researchers to pre-screen and invite eligible participants from their participant pool.
+Studies 1 and 2 were conducted on the standard sample of US participants.
 In order to increase the share of participants with low trust in science, and who endorse conspiracy theories, Studies 3 and 4 used the same platform, but only recruited participants who had declared previously being skeptical of vaccination.
+
+Because recruitment was conducted through Prolific's system, we do not have access to the number of individuals who viewed the study invitation but chose not to participate.
+As such, it is not possible for us to calculate response rates (the proportion of invited participants who completed the survey) for the studies.
+However, we employed attention checks and, for each study, we report exclusions based on these checks.
 
@@
 Funk, Johnson, \& Hefferon, 2019) (``How much confidence do you have in scientists to act in the best interests of the public?
-{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}''), the latter having been used in a recent international study on trust in science (Cologna et al., 2024).
+{[}1-5; 1 = No confidence at all, 5 = A great deal of confidence{]}''), the latter having been used in a recent international study on trust in science (Cologna et al., 2025).
 We selected these items so that we could compare the answers in our sample to global survey results.
@@
 \centering
-\includegraphics{output/figures/summary-plot.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/summary-plot.pdf}}
 \caption{\label{fig:summary-plot}Points represent the average share of acceptance and numbers the absolute count of participants as a function of: \textbf{A} the level of trust in science (``In general, would you say that you trust science a lot, some, not much, or not at all?
@@
 Regarding H2, the results are mixed for the relation of conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) with both acceptance of the scientific consensus (Studies 1: r = -0.33, p \textless{} .001; 2: r = -0.37, p \textless{} .001; 3: r = -0.02, p = 0.788; 4: r = -0.05, p = 0.498) and science knowledge (Studies 1: r = -0.38, p \textless{} .001; 2: r = -0.40, p \textless{} .001; 3: r = -0.16, p = 0.055; 4: r = -0.02, p = 0.742).
-
-n an exploratory analysis, we find that
-
-While actual acceptance of scientific knowledge is only weakly correlated with trust in science,
 
@@
 Ranney \& Clark, 2016).
+Past research has shown that arguments building on specific scientific findings are more convincing even to populists who are generally skeptical towards expertise (Peresman, Larsen, Mazepus, \& Petersen, 2025).
+
+Why do most people (in the US) trust most of basic science?
+Based on the present studies, we can only speculate.
+But if this trust has a rational basis, then education is the most plausible candidate for explaining it.
+Large-scale surveys have identified education, and in particular science education, as one of the strongest correlates of trust in science (Noy \& O'Brien, 2019; Wellcome Global Monitor, 2018) but not consistently so (see e.g., Cologna et al., 2025 who only find a small positive relationship between tertiary education and trust in science).
+If these studies were to assess trust in scientific knowledge, rather than assessing trust more broadly, we might expect effects to be more strongly and more coherently positive.
+Differentiating the impact of education on different facets of trust would be an interesting avenue for future research.
 
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
+The use of online samples recruited via Prolific Academic does not allow us to calculate response rates and thus estimate non-response bias.
+However, by targeting vaccine-skeptic samples in some of the studies, we made an effort in recruiting those participants that can be assumed to be most reluctant to partake in a study confronting them with science facts.
 While our results support motivated reasoning accounts of trust in science, they leave important questions unaddressed, in particular: Why do people reject specific scientific findings but not others?
@@
 
-\bibitem[\citeproctext]{ref-colognaTrustScientistsTheir2024}
+\bibitem[\citeproctext]{ref-colognaTrustScientistsTheir2025}
 Cologna, V., Mede, N.
-G., Berger, S., Besley, J., Brick, C., Joubert, M., \ldots{} Linden, D.
-S.
-van der.
-(2024).
-\emph{Trust in scientists and their role in society across 67 countries}.
-\url{https://doi.org/10.31219/osf.io/6ay7s}
+G., Berger, S., Besley, J., Brick, C., Joubert, M., \ldots{} Zwaan, R.
+A.
+(2025).
+Trust in scientists and their role in society across 68 countries.
+\emph{Nature Human Behaviour}, 1--18.
+\url{https://doi.org/10.1038/s41562-024-02090-5}
 
@@
 
+\bibitem[\citeproctext]{ref-jabkowskiExploringCOVID19Conspiracy2023}
+Jabkowski, P., Domaradzki, J., \& Baranowski, M.
+(2023).
+Exploring COVID-19 conspiracy theories: education, religiosity, trust in scientists, and political orientation in 26 European countries.
+\emph{Scientific Reports}, \emph{13}(1), 18116.
+\url{https://doi.org/10.1038/s41598-023-44752-w}
+
 \bibitem[\citeproctext]{ref-krauseTrendsAmericansTrust2019a}
@@
 \url{https://doi.org/10.1136/bmjopen-2020-048172}
+
+\bibitem[\citeproctext]{ref-littrellBeliefScienceRelatedConspiracy2025}
+Littrell, S., Diekman, A., Seelig, M., Premaratne, K., Wuchty, S., Verdear, D., \ldots{} Uscinski, J.
+(2025).
+Belief in Science{-}Related Conspiracy Theories.
+\emph{Journal of Social Issues}, \emph{81}(1), e70004.
+\url{https://doi.org/10.1111/josi.70004}
 
@@
 
+\bibitem[\citeproctext]{ref-noyScienceGoodEffects2019}
+Noy, S., \& O'Brien, T.
+L.
+(2019).
+Science for good?
+The effects of education and national context on perceptions of science.
+\emph{Public Understanding of Science}, \emph{28}(8), 897--916.
+\url{https://doi.org/10.1177/0963662519863575}
+
 \bibitem[\citeproctext]{ref-pennycookOverconfidentlyConspiratorialConspiracy2022}
@@
 \url{https://doi.org/10.31234/osf.io/d5fz2}
+
+\bibitem[\citeproctext]{ref-peresmanPopulistsListenExpertise2025}
+Peresman, A., Larsen, L.
+T., Mazepus, H., \& Petersen, M.
+B.
+(2025).
+Do Populists Listen to Expertise?
+A Five-Country Study of Authority, Arguments, and Expert Sources.
+\emph{Political Studies}, 00323217251323402.
+\url{https://doi.org/10.1177/00323217251323402}
 
@@
 \centering
-\includegraphics{output/figures/acceptance-knowledge.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/acceptance-knowledge.pdf}}
 \caption{\label{fig:acceptance-knowledge}Acceptance/Rejection rate of scientific consensus, based on whether the initial response to the knowledge question was false or true.}
@@
 \centering
-\includegraphics{output/figures/conspiracy-thinking-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/conspiracy-thinking-overview.pdf}}
 \caption{\label{fig:conspiracy-thinking-overview}Distributions of conspiracy thinking.}
@@
 \centering
-\includegraphics{output/figures/trust-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/trust-overview.pdf}}
 \caption{\label{fig:trust-overview}Distributions of trust in science.}
@@
 \centering
-\includegraphics{output/figures/conspiracy-belief-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/conspiracy-belief-overview.pdf}}
 \caption{\label{fig:conspiracy-belief-overview}Average belief in conspiracy theories.}
@@
 \centering
-\includegraphics{output/figures/knowledge-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/knowledge-overview.pdf}}
 \caption{\label{fig:knowledge-overview}Distributions of science knowledge.}
@@
 \centering
-\includegraphics{output/figures/questions-knowledge.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/questions-knowledge.pdf}}
 \caption{\label{fig:questions-knowledge}Share of correct answers by question.}
@@
 \centering
-\includegraphics{output/figures/questions-acceptance.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/questions-acceptance.pdf}}
 \caption{\label{fig:questions-acceptance}Acceptance rate of consensus by question.}
@@
 \centering
-\includegraphics{output/figures/exp1-plot-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/exp1-plot-overview.pdf}}
 \caption{\label{fig:exp1-plot-overview}Summary plot for Study 1.
@@
 \centering
-\includegraphics{output/figures/exp2-plot-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/exp2-plot-overview.pdf}}
 \caption{\label{fig:exp2-plot-overview}Summary plot for Study 2.
@@
 \centering
-\includegraphics{output/figures/exp3-plot-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/exp3-plot-overview.pdf}}
 \caption{\label{fig:exp3-plot-overview}Summary plot for Study 3.
@@
 \centering
-\includegraphics{output/figures/exp4-plot-overview.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{output/figures/exp4-plot-overview.pdf}}
 \caption{\label{fig:exp4-plot-overview}Summary plot for Study 4.

### 2025-07-07 preprint.Rmd
second round revisions

@@
 appendix:
+  - "appendix_acceptance_knowledge.Rmd"
+  - "appendix_conspiracy_thinking.Rmd"
+  - "appendix_trust.Rmd"
+  - "appendix_conspiracy_belief.Rmd"
+  - "appendix_knowledge.Rmd"
+  - "appendix_by-question.Rmd"
+  - "appendix_conspiracy-scales.Rmd"
+  - "appendix_exp1.Rmd"
+  - "appendix_exp2.Rmd"
+  - "appendix_exp3.Rmd"
+  - "appendix_exp4.Rmd"
 bibliography: references.bib
@@
 
-Unfortunately, people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
-Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; although see @wellcomeglobalmonitorPublicTrustScientists2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
-Rejection of science can take the more extreme form of conspiracy theories, many of which question the scientific consensus on vaccination, climate change, and even the shape of the Earth.
-
-What does this apparent rejection of science actually entail?
+Although recent global evidence shows that trust in science is moderately high [@colognaTrustScientistsTheir2024], it is far from being at ceiling.
+Large-scale polls have shown that people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020].
+Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; @brianAmericansTrustScientists2023; although see @wellcomeglobalmonitorPublicTrustScientists2021; and @funkPublicConfidenceScientists2020] and in the US it is increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].
+
+Besides low answers on general trust in science questions, another indicator of distrust in science is the belief in conspiracy theories that question the scientific consensus on issues such as vaccination, climate change, and even the shape of the Earth.
+Conspiracy theories–in the realm of science or elsewhere–typically accuse a small group of powerful people to pursue nefarious goals in secrecy [@douglasUnderstandingConspiracyTheories2019; @medeSciencerelatedPopulismConceptualizing2020].
+Some of these conspiracy theories are widespread [@rutjensConspiracyBeliefsScience2022].
+In 2023, a survey in eight different countries found that up to 24% of respondents agreed that “climate change is a hoax and scientists touting its existence are lying” [@stockemerUnderstandingClimateChange2024].
+Similarly, in 2021, 40% of Americans believed that “the dangers of genetically-modified foods are being hidden from the public” [down from 45% in 2020, @uscinskiHaveBeliefsConspiracy2022].
+People who hold such views not only reject the relevant scientifically consensual facts, but also tend to believe in other conspiracy theories [@lewandowskyRoleConspiracistIdeation2013; @hornseyPsychologicalRootsAntivaccination2018; @hornseyRelationshipsConspiratorialBeliefs2018], and tend to say that they distrust science more generally [@vranicDidMyOwn2022; @stockemerUnderstandingClimateChange2024].
+In spite of these correlations, the causal relationship between declaring general distrust in science and believing in one or several anti-science conspiracy theories is not clear.
+Although conspiracy thinking has been identified as a "root cause" of anti-science attitudes [@hornseyAttitudeRootsJiu2017] in the past, these claims rest largely on observational data.
+
+What does this apparent lack of trust in science actually entail?
 Do people who say they do not trust science, or who believe in conspiracy theories at odds with well-established science, reject most of science?
 Or, on the contrary, do they object to a few specific facets of science, while still accepting the overwhelming majority of basic science?
-Answering this question has theoretical implications: if some people reject science wholesale, then it’s possible that their distrust of science causes their rejection of specific scientific theories, or facilitates their acceptance of conspiracy theories; by contrast, if everyone trusts most of science, then the mistrust of specific facets of science is more likely to be post-hoc, a rationalization of other beliefs or behaviors.
-The present question also has practical implications: many communication attempts leverage the scientific consensus [e.g.
+
+A common conception of trust is a willingness to be vulnerable to another party, whether an individual, a group, or an institution [@mayerIntegrativeModelOrganizational1995a; @rousseauIntroductionSpecialTopic1998].
+Accordingly, trust in science has been defined as “one’s willingness to rely on science and scientists (as representatives of the system) despite having a bounded understanding of science” [@wintterlinPredictingPublicTrust2022, p.
+2].
+Past research has disentangled this general concept of trust in science in various ways.
+Some research has identified different components of trust, the number of which varies, but which generally cover an epistemological and ethical dimension [@wilholtEpistemicTrustScience2013; @intemannScienceCommunicationPublic2023].
+For example, @hendriksMeasuringLaypeoplesTrust2015 suggest distinguishing between expertise/competence, integrity, and benevolence, while @besleyReassessingVariablesUsed2021 add openness.
+Other research has highlighted differences in trust between scientific disciplines [@altenmullerExplainingPolarizedTrust2024; @gligoricHowSocialEvaluations2024; @gauchatCulturalCognitiveMappingScientific2018].
+However, little research has assessed trust in specific scientific findings, besides contentious topics such as vaccines [e.g.
+@hornseyPsychologicalRootsAntivaccination2018], climate change [e.g.
+@stockemerUnderstandingClimateChange2024], evolution [e.g.
+@nadelsonTrustScienceScientists2015], genetically modified organisms (GMOs) [e.g.
+@fernbachExtremeOpponentsGenetically2019], or a combination of such topics [see e.g.
+@lewandowskyRoleConspiracistIdeation2013].
+To the best of our knowledge, no research has investigated the extent to which people trust basic science facts (e.g.
+electrons are smaller than atoms).
+An extensive literature on science literacy has assessed whether people know such facts [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016], but not whether they accept the facts once presented to them.
+
+Why does it matter whether trust in science is or is not related to trust in specific scientific findings?
+First, this question has theoretical implications.
+According to the most prominent explanation of (dis)trust in science–the deficit model–science knowledge is the main driver of attitudes towards science in general.
+A prediction of this model is that average trust in science on specific facts should be strongly associated with general trust in science.
+By contrast, a disconnect between the two would be in line with motivated reasoning accounts of trust in science.
+According to these accounts, science rejection serves to maintain coherence with other beliefs or behaviors [@lewandowskyMotivatedRejectionScience2016; @hornseyWhyFactsAre2020].
+For example, someone might say they do not trust science in general because they are skeptical towards vaccines, not because they actually distrust most of science.
+A prediction of the motivated reasoning account is that general trust in science should be strongly correlated with conspiracy beliefs.
+Second, the present question has practical implications: Many communication attempts leverage the scientific consensus [e.g.
 on vaccination, climate change, etc., for review, see @vanstekelenburgScientificConsensusCommunicationContested2022; see also @veckalov27countryTestCommunicating2024].
@@
 
-Studies of science knowledge provide relevant evidence, as people who answer science knowledge questions correctly presumably trust scientists in this respect.
-The average level of science knowledge is quite low, even in countries with extensive science education.
-For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016].
-However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.
-
 ## The present studies
 
-In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts.
+In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established, consensual scientific facts.
 For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted answer, and asked them whether they accepted it (measuring their trust in basic science).
 We also measured participants’ trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking.
-The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (https://osf.io/8utsj/?view_only=364478f01fc548d0a42b9962918d871d).
+The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (<https://osf.io/8utsj/>).
 The differences between the four studies are summarized presently, and the methods are detailed below.
@@
 
-US participants were asked whether they accepted scientifically consensual answers on basic science questions.We found quasi-universal acceptance of basic science, with an overall rate of acceptance of `r combined_descriptives$mean_acceptance$Yes$share`, which remained very high for participants who declared not trusting science at all (`r combined_descriptives$acceptance_by_trust$'1'$Yes$share` of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (`r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share` of acceptance).
-
-Overall, the present findings support the motivated rejection of science model [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p.
+In four studies, we asked US participants whether they accepted scientifically consensual answers on basic science questions.
+We found quasi-universal acceptance of basic science, with an overall rate of acceptance of `r combined_descriptives$mean_acceptance$Yes$share`, which remained very high for participants who declared not trusting science at all (`r combined_descriptives$acceptance_by_trust$'1'$Yes$share` of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (`r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share` of acceptance).
+
+This disconnect between declared general trust in science and average trust in (basic) scientific findings goes against predictions from the knowledge–attitudes model of trust in science [see, e.g., @bauerWhatCanWe2007], which posits that knowledge of scientific results is the main cause of trust in science.
+In line with past research [@allumScienceKnowledgeAttitudes2008], we find that both knowledge and acceptance of science are only weakly correlated with general trust in science.
+This supports recent definitions of science literacy, which suggest that for science literacy to be a meaningful concept for science attitudes, it needs to go beyond measuring knowledge of isolated science facts [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016].
+
+By contrast, our findings align with a motivated reasoning account of trust in science [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p.
 217).
-A number of participants in our studies rejected basic tenets of science (on evolution, the shape of the earth, etc.).
-However, they still accepted the vast majority of basic scientific knowledge presented to them.
-This suggests that these participants had reasons to reject specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
-These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
-These results are in line with several recent studies in which conspiracy theorists are not less susceptible to social influence in general [@pummererConspiracyBeliefsMajority; @altayConspiracyBelieversClaim2023].
-
-Our findings are also relevant for the knowledge–attitudes model of trust in science [see, e.g., @bauerWhatCanWe2007].
-The fact that even people with low science knowledge trust most of basic science might be taken as showing the power of even a modicum of science knowledge; on the other hand, it also means that attempts at increasing science knowledge in general might not have much effect on trust in science, by contrast with targeting the specific beliefs people are motivated to reject.
-
-Finally, some of the present results also speak to the alienation model [@gauchatCulturalAuthorityScience2011], and more specifically to the need for epistemic autonomy [@frickerTestimonyEpistemicAutonomy2006].
+A number of participants in our studies endorsed specific conspiracy theories questioning basic tenets of science (on evolution, the shape of the earth, etc.), while still accepting the vast majority of basic scientific facts presented to them.
+This suggests that these participants had reasons to reject only specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic.
+Consistent with this explanation, we found a strong association between belief in anti-science conspiracy theories and general trust in science.
+
+Some of the present results also speak to the alienation model [@gauchatCulturalAuthorityScience2011], and more specifically to the need for epistemic autonomy [@frickerTestimonyEpistemicAutonomy2006].
 Declaring not trusting science, or endorsing conspiracy theories [@harrisConspiracyTheoriesPopulism2023] might reflect a desire to maintain epistemic autonomy and not appear to ‘blindly’ accept epistemic authority.
-The acceptance of basic science facts would be reconciled with this need to appear epistemically autonomous by the suggestion that the acceptance stems from independent evaluation instead of trust (however implausible that might be: it’s not clear how participants could independently verify the ratio of glial cells to neurons, say, and few participants appeared to have engaged in even simple forms of verification).
-
-In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that can be tapped into in science communication.
-Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the ‘consensus gaps’ observed in these domains [on climate change, see, e.g.
+Such a need to appear epistemically autonomous could be reconciled with the acceptance of basic science facts if it was thought to stem from independent evaluation instead of trust.
+The majority of participants in Studies 3 and 4, two samples of vaccine-skeptical participants who scored high across a range of different conspiracy beliefs (see ESM), tended to claim that they had accepted the scientific consensus based on their own evaluation, however implausible that might be: it is not clear how participants could independently verify, say, the ratio of glial cells to neurons.
+These participants did also not engage (more than others) in simple forms of verification, i.e.
+clicking on sources provided in the answers.
+This result aligns with other results on conspiracy thinking: Recent studies have shown that, although they claim not to be, conspiracy theorists are in fact just as susceptible to social influence as others [@pummererConspiracyBeliefsMajority2024; @altayConspiracyBelieversClaim2023].
+
+In applied terms, the present results have implications for science communication.
+Across various domains of science knowledge such as climate change [@veckalov27countryTestCommunicating2024] or vaccination [@salmonVaccineHesitancyCauses2015], researchers have observed a consensus gap: a gap between the scientific consensus and public opinion.
+Some of the most worrying consensus gaps relate to climate change, as substantial segments of the population disagree with scientists on what is happening and what to do about it [e.g., @eganClimateChangeUS2017].
+Yet, we found that for much of basic, non-contentious science, such consensus gaps are absent: The vast majority of our participants--including two vaccine-skeptical samples (Studies 3 and 4)--did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them.
+Since even people who say they distrust science, or who reject specific scientific facts appear to accept most of basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the consensus gaps observed in these domains [on climate change, see, e.g.
 @ranneyClimateChangeConceptual2016].
@@
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
-We hope that future studies will extend our findings to other populations.
+While our results support motivated reasoning accounts of trust in science, they leave important questions unaddressed, in particular: Why do people reject specific scientific findings but not others?
+Suggestions have already been made for a number of issues such as vaccination [@mitonCognitiveObstaclesProVaccination2015], GMOs [@blanckeFatalAttractionIntuitive2015], or nuclear energy [@hacquinDisgustSensitivityPublic2021].
+However, research is still needed to better understand what motivates these rejections [see e.g., @hornseyWhyFactsAre2020].
+Moreover, rejection of scientific facts can manifest in various ways, for example calling into question the integrity of scientists [@medeSciencerelatedPopulismConceptualization2023a; @medeWhoSupportsSciencerelated2022], or denying the very possibility of scientifically investigating certain issues [@munroScientificImpotenceExcuse2010].
+The classification of justifications we present here does not address these processes in detail.
+Finally, by design, most of the basic science knowledge presented in this study didn't directly relate to anything controversial for most people.
+Future studies could look at basic science that does relate to scientific topics which are the object of public and policy-relevant debates (e.g.
+GM food, climate science, vaccines), or scientific findings for which there is a more obvious potential conflict of interest (e.g.
+non-publicly funded medical research).
+
+### Data availability
+
+Data for all experiments and the simulations is available on the OSF project page (https://osf.io/8utsj/).
+
+### Code availability
+
+The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (https://osf.io/8utsj/).
+
+### Competing interest
+
+The authors declare having no competing interests.
 

### 2025-07-07 paper.tex
second round  revisions update preprint

@@
 \abstract{%
-Substantial segments of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
+Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge.
 This might indicate a wholesale rejection of science.
@@
 
-Because recruitment was conducted through Prolific's system, we do not have access to the number of individuals who viewed the study invitation but chose not to participate.
+Because recruitment was conducted through Prolific, we do not have access to the number of individuals who viewed the study invitation but chose not to participate.
 As such, it is not possible for us to calculate response rates (the proportion of invited participants who completed the survey) for the studies.
 However, we employed attention checks and, for each study, we report exclusions based on these checks.
+Moreover, the study invitation was purposefully nondescript (``short psychology experiment''), so that it would be impossible for participants to self-select because of their attitudes towards science for instance.
 
@@
 Based on the present studies, we can only speculate.
-But if this trust has a rational basis, then education is the most plausible candidate for explaining it.
-Large-scale surveys have identified education, and in particular science education, as one of the strongest correlates of trust in science (Noy \& O'Brien, 2019; Wellcome Global Monitor, 2018) but not consistently so (see e.g., Cologna et al., 2025 who only find a small positive relationship between tertiary education and trust in science).
-If these studies were to assess trust in scientific knowledge, rather than assessing trust more broadly, we might expect effects to be more strongly and more coherently positive.
+But if this trust has a rational basis, then education--and science education in particular--is the most likely explanation.
+Large-scale surveys have identified science education as one of the strongest, or the strongest, correlates of trust in science (Noy \& O'Brien, 2019; Wellcome Global Monitor, 2018) If these studies were to assess trust in basic scientific knowledge, rather than assessing trust in science more broadly, we might expect effects to be even stronger.
 Differentiating the impact of education on different facets of trust would be an interesting avenue for future research.
@@
 The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country.
-The use of online samples recruited via Prolific Academic does not allow us to calculate response rates and thus estimate non-response bias.
-However, by targeting vaccine-skeptic samples in some of the studies, we made an effort in recruiting those participants that can be assumed to be most reluctant to partake in a study confronting them with science facts.
+The use of online samples recruited via Prolific does not allow us to calculate response rates and thus estimate non-response bias.
+However, participants could not have self-selected on the basis of the study topic and, by targeting vaccine-skeptic samples in some of the studies, we made an effort in recruiting those participants that can be assumed to be most reluctant to partake in a study confronting them with science facts.
 While our results support motivated reasoning accounts of trust in science, they leave important questions unaddressed, in particular: Why do people reject specific scientific findings but not others?

### 2025-07-17 preprint.Rmd
update preprint for second submission

@@
 title             : "Quasi-universal acceptance of basic science in the US"
-shorttitle        : "Does anyone mistrust basic science?"
+shorttitle        : "Quasi-universal acceptance of basic science in the US"
 header-includes:  | # to prevent floats from moving past certain points (for the appendix)
@@
     affiliation   : "1"
-    #role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
   - name          : "Lou Kerzreho"
     affiliation   : "1"
-    #role:         
   - name          : "Hugo Mercier"
@@
 This could be leveraged in science communication.
+
+authornote: |
+  Jan Pfänder is a PhD candidate in psychology, working on misinformation and trust in science. 
+  Lou Kerzreho is a Master’s student in cognitive science.
+  Hugo Mercier is a research director at the CNRS, whose work has focused on two main topics: the function and workings of reasoning, and how we evaluate communicated information.
+
 keywords          : Trust in science; conspiracy theories; conspiracy thinking; science knowledge.

### 2025-09-05 README.qmd
add README file and a file with producing additional figures

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("preprint.Rmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
```

# `r title`

---

## Abstract

`r abstract`

---

## How to download and replicate

To reproduce the findings and re-run the analysis, do the following:

1. Download and install this repository. You can use GitHub to clone or fork the repository (see the green "Clone or download" button at the top of the GitHub page).
2. Open `everyone_trusts_science_abit.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
3. Within the RStudio Project, open the `paper.Rmd` (alternatively, `preprint.Rmd` for a version with authors and affiliations) file and run it.

Please find the [session info below](#session-info).

*Estimated installation time* : If you have R and Rstudio installed, then adding additional packages required in our R scripts is quick (a couple of minutes)

*Estimated run time*: Rendering the `paper.Rmd` document takes less than 1min on our machine (see [session info](#session-info))

## Files glossary {#glossary}

- the two main files are the reproducible source scripts of the paper: `paper.Rmd` is an anonymized version for peer review, `preprint.Rmd` is the version for the public preprint on OSF. 
- all files with the prefix `paper_*` or `preprint_*` are output files that are generated by these source files. 
- `appendix*` files contain the appendices that are automatically added to `manuscript.Rmd`/`preprint.Rmd` when rendering.
- `experiment_*` folders contain all files related to the respective experiment. The `/data` folders contain the raw data (`qualtrics.csv`) and cleaned data (`cleaned.csv`). The cleaning can be reproduced by running the `cleaning.Rmd` files in the folder of each respective experiment. 
- `qualtrics` folders within each experiment folder contains copies of the surveys run via Qualtrics
- `functions/` folder contains functions that we rely upon in our analysis scripts. To keep the `paper.Rmd`/`preprint.Rmd` document neat and readable, we wrote all lengthy code bits as external functions.
- running the `paper.Rmd`/`preprint.Rmd` file will generate an `output` folder which lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.
- `references.bib` contains the references

## Session info {#session-info}

```{r, echo=FALSE}
sessionInfo()
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

### 2025-09-05 figures/additional_figures.qmd
add README file and a file with producing additional figures

---
title: "Additional figures"
format: html
---

```{r}
#| label: setup

library(tidyverse)
library(gt)
library(kableExtra)
library(readxl)
library(wesanderson) # colours
library(gghalves)
library(broom)
library(broom.mixed)
library(ggpubr)
```

```{r functions}
# load plot theme
source("../functions/plot_theme.R")

# load other functions
source("../functions/own_functions.R")
```

```{r}
# read data
exp1_long <- read_csv("../exp_1/data/cleaned_long.csv")
exp2_long <- read_csv("../exp_2/data/cleaned_long.csv")
exp3_long <- read_csv("../exp_3/data/cleaned_long.csv")
exp4_long <- read_csv("../exp_4/data/cleaned_long.csv")

## combine data of all three studies
combined_data <- bind_rows(exp1_long %>% 
                             mutate(study = "1"), 
                           exp2_long %>% 
                             mutate(study = "2"), 
                           exp3_long %>% 
                             mutate(study = "3"),
                           exp4_long %>% 
                             mutate(study = "4")
                           ) %>% 
  ## make a unique id variable
  mutate(id = paste0("study", study, "_", id))
```

```{r}
# make plot data for acceptance
plot_data_acceptance <- combined_data |> 
  drop_na(wgm_sciencegeneral, acceptance) |> 
  # make a labels version of trust in science
  mutate(wgm_sciencegeneral = case_when(
    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
    wgm_sciencegeneral == 2 ~ "2 (Not much)",
    wgm_sciencegeneral == 3 ~ "3 (Some)",
    wgm_sciencegeneral == 4 ~ "4 (A lot)",
    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
  )) |> 
  group_by(wgm_sciencegeneral, acceptance) |> 
  summarize(n = n()) |> 
  group_by(wgm_sciencegeneral) |> 
  mutate (share = n/sum(n)) |> 
  mutate_if(is.numeric, round, digits = 3) |> 
  filter(acceptance == "Yes") |> 
  select(-acceptance) |> 
  mutate(outcome = "Acceptance")

# make plot data for knowledge
plot_data_knowledge <- combined_data |> 
  drop_na(wgm_sciencegeneral, knowledge) |> 
  # make a labels version of trust in science
  mutate(wgm_sciencegeneral = case_when(
    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
    wgm_sciencegeneral == 2 ~ "2 (Not much)",
    wgm_sciencegeneral == 3 ~ "3 (Some)",
    wgm_sciencegeneral == 4 ~ "4 (A lot)",
    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
  )) |> 
  group_by(wgm_sciencegeneral, knowledge) |> 
  summarize(n = n()) |> 
  group_by(wgm_sciencegeneral) |> 
  mutate (share = n/sum(n)) |> 
  mutate_if(is.numeric, round, digits = 3) |> 
  filter(knowledge == TRUE) |> 
  select(-knowledge) |> 
  mutate(outcome = "Knowledge")

# combine all data
plot_data <- bind_rows(plot_data_acceptance, plot_data_knowledge) |> 
  mutate(outcome = factor(outcome, levels = c("Knowledge", "Acceptance")
                          )
         )

absolute_numbers <- combined_data %>% 
  group_by(wgm_sciencegeneral) %>% 
  summarise(n = n(), 
            n_participants = n_distinct(id))
```

```{r}
ggplot(plot_data |> 
         mutate(outcome = factor(ifelse(outcome == "Acceptance", "Acceptance", "Knowledge"),
                                 levels = c("Knowledge", "Acceptance"))), 
       aes(x = wgm_sciencegeneral, y = share, fill = outcome)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Trust in science", fill = "Share of", y = "") +
  plot_theme +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(wgm_sciencegeneral)) - 0.4, 
                   xmax = as.numeric(factor(wgm_sciencegeneral)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n_participants)),
               inherit.aes = FALSE,
               step.increase = 0) +
  theme(legend.position = "top")
```
